### 2022-10-05 manuscript/manuscript.Rmd
Adding manuscript to .gitignore

---
title: "Metabolite soup"
author: "William Kumler"
date: "10/27/2022"
output:
  word_document: default
  html_document: default
editor_options:
  chunk_output_type: inline
---

Hi Regina

```{r run analysis, eval=FALSE, echo=FALSE}
# Run untargeted analysis
knitr::purl("../untargeted/output_control.Rmd", output="../untargeted/output_script.R")
script_lines <- readLines("../untargeted/output_script.R")
polarity_set_line_idx <- grep(script_lines, pattern = "polarity <-")[1]
cruise_set_line_idx <- grep(script_lines, pattern = "cruise <-")[1]
to_write <- script_lines[-c(polarity_set_line_idx, cruise_set_line_idx)]
writeLines(to_write, "../untargeted/output_script.R")
current_dir <- getwd()
setwd("../untargeted")
for(cruise in c("FK", "MT", "MC")){
  for(polarity in c("pos", "neg")){
    print(paste("Working on", cruise, polarity))
    source("output_script.R", echo = FALSE)
  }
}

# Extract targeted results from targeted analysis

```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE, include = TRUE, fig.width = 6.5)

options(warn = -1)
options(dplyr.summarise.inform=F)
options(readr.num_columns = 0)
options(pillar.sigfig=7)
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(RaMS))
library(scales)
library(ggdendroplot)
library(ggrepel)
library(lmPerm)

depth_levels <- c("15m", "25m", "DCMless20m", "DCMless10m", "DCM", "DCMplus10m",
                  "DCMplus20m", "175m")
filled_file_metadata <- "../metadata/made_data/filled_file_metadata.csv" %>%
  read_csv(show_col_types = FALSE) %>%
  mutate(depth=factor(depth, levels = depth_levels))

factor_levels <- c(
    "no23_um", "po4_um", "si_um", "PC_um", "PN_um", "DON_um", "DOP_um", "oxy_um",
    "chl_ug", "phaeo_ug", "alk", "dic", "ph", "theta",
    paste0(c("pro", "syn", "euk", "het"), "_abund"), "beam_atten",
    "abs_depth", "time", "lat", "sla", "local_hour"
  )
factor_labels <- c(
    "Nitrate + nitrite", "Phosphate", "Silicate", "Particulate C", "Particulate N",
    "Dissolved organic N", "Dissolved organic P", "Oxygen", "Chlorophyll a",
    "Phaeophytin", "Alkalinity", "Dissolved inorganic C", "pH", "Density",
    paste(c("Pro.", "Syn.", "Picoeuk.", "Het. bact."), "abundance"),
    "Beam attenuation",
    "Depth", "Time", "Latitude", "Sea level anomaly", "Local hour"
  )
twoline_labels <- c(
    "Nitrate + nitrite", "Phosphate", "Silicate", "Particulate C", "Particulate N",
    "Dissolved\norganic N", "Dissolved\norganic P", "Oxygen", "Chlorophyll a",
    "Phaeophytin", "Alkalinity", "Dissolved\ninorganic C", "pH", "Density",
    paste(c("Prochlorococcus", "Synechococcus", "Picoeukaryote", 
            "Heterotrophic bact."), "\nabundance"),
    "Beam attenuation",
    "Depth", "Time", "Latitude", "Sea level anomaly", "Local hour"
  )

clean_stans <- "../metadata/made_data/clean_stans.csv" %>%
  read_csv(show_col_types = FALSE) %>%
  select(compound_type, compound_name, formula, polarity)

all_feats <- read_csv("../untargeted/all_feats.csv", show_col_types = FALSE)
all_peaks <- read_csv("../untargeted/all_peaks.csv", show_col_types = FALSE)

MT_feats <- all_feats %>%
  filter(cruise=="MT") %>%
  select(-classes, -CSI_name) %>%
  rename("calc_formula"=formula) %>%
  left_join(clean_stans, by=c("compound_name", "polarity"))
MT_peaks <- MT_feats %>%
  left_join(all_peaks, by="feature") %>%
  left_join(filled_file_metadata, by="filename") %>%
  filter(samp_type=="Smp") %>%
  select(feature, norm_area, filename)
```

## Abstract

TBD

## Introduction

A major goal of marine microbial metabolomics is a global ocean model of metabolites that resolves at the molecular level, answering the question: "At any given spot and any given time, what is the concentration of a given molecule 1) within a cell and 2) in the surrounding seawater?" With such a model, we could more easily discover marine biomarkers, predict the response of marine communities to perturbations both natural and anthropogenic, and even characterize the production of cloud-forming molecules such as DMS and improve our climate models [and similar stuff, don't love these examples but like the metabolite model justification].

This is a difficult problem because the metabolome is the result of a complex network of interactions between the ocean's biology and its biogeochemistry. Gradients of light and nutrient availability create a spectrum of fundamental niches throughout the water column that are sensitive to diel, seasonal, and interannual cycles. Additionally, vertical and horizontal mixing due to upwelling/downwelling, storms, and mesoscale features alter the distribution of water masses and the dissolved molecules contained within.

These complex interactions make it difficult to predict exactly which parameters are necessary to include in the final model. Some parameters simply don't affect the outcome enough to be worth including because their effect sizes are so small that they have no biological significance. Consider an absurd parameter like "distance to nearest butterfly" that somehow obtains enormous statistical significance. If, however, it only changes values in the model on the scale of attomoles it is simply not worth calculating in the model. [Need better example here]. We can constrain the number of possible parameters by sketching out a simple ocean model, the simplest of which simply predicts an average value for every molecule at every location and time. This is, approximately, where the field of marine microbial metabolomics currently rests. For many metabolites, we have an estimate of its concentration at a single location, forcing us to extrapolate that single value vertically and horizontally throughout the ocean in the absence of other information.

With some additional oceanographic knowledge, we can predict how metabolites respond to depth. The classic Martin curve and global DOM measurements allow us to predict that the surface ocean will have the highest concentrations of metabolites and that compound abundance will generally decrease with depth. This improves our model's accuracy significantly but fails to account for the variable lability of different compounds, with some highly recalcitrant molecules having essentially conservative profiles while other molecules are assimilated extremely rapidly. This spectrum of reactivity is expected based on [blah] but has [not yet been explored.] [Mention certain compound classes? Some amino acids more labile than others?]

Even more complex models have begun to emerge in recent years informed by additional data and our understanding of ocean physics, chemistry, and biology. We have just begun to characterize the response of metabolites to gradients in time and space, covering diel variation and transects across the ocean surface. We've also begun to integrate biological information from genomic and transcriptomic studies, allowing us to better resolve the theoretical metabolome as biology responds to shifts in light intensity and nutrient availability. The model shown below in black represents one potential graph of these interactions and the role of metabolites as mediators of the interaction between organisms.

```{r fig.height=3}
DiagrammeR::grViz(diagram = "digraph dot {
  graph [rankdir = LR]

  tab1 [label = '@@1', pos='0, 1!']
  tab2 [label = '@@2', pos='0, 0!']
  tab3 [label = '@@3', pos='2, 0.5!']
  tab4 [label = '@@4', pos='2,-0.5!']
  tab5 [label = '@@5', pos='4, 1!']
  tab6 [label = '@@6', pos='4,-1!']
  tab7 [label = '@@7', pos='4, 0!']
  tab8 [label = '@@8', pos='0,-1!', color='red', fontcolor='red', style='dashed']

  tab1 -> tab3;
  tab2 -> tab3;
  tab2 -> tab4;
  tab3 -> tab5;
  tab4 -> tab5;
  tab4 -> tab6;
  tab5 -> tab7;
  tab6 -> tab7;
  tab7 -> tab5;
  tab7 -> tab6;
  tab8 -> tab4 [color='red', style='dashed']
}

  [1]: 'Time of day'
  [2]: 'Depth'    
  [3]: 'Light'
  [4]: 'Nutrients'
  [5]: 'Phototrophs'
  [6]: 'Heterotrophs'
  [7]: 'Metabolites'
  [8]: 'Sea level anomaly'
  ", engine="neato")
```

However, the model as is predicts a uniform concentration of nutrients at a given depth. This is not observed in the real world due to mixing and the vertical movement of water masses in response to mesoscale eddies. These features can be identified by satellites as anomalies in sea surface height and reflect the fluctuations of density surfaces within the water column.

Cyclonic eddies, those circulating counterclockwise in the northern hemisphere and clockwise in the southern, bring the main pycnocline closer to the surface of the ocean and introduce extra nutrients to the sunlit layer of the ocean upon formation. Anticyclonic eddies have the opposite effect, depressing isopycnal surfaces and essentially piling up surface water in regions of positive sea level anomaly (SLA). These eddies can increase the concentrations of nutrients at the base of the photic zone by more than an order of magnitude and are known to affect phytoplankton species composition [and other species?].

The magnitude of this effect on the metabolome is currently unknown, potentially representing a major source of variation in the marine metabolite model. If metabolites are as sensitive to SLA as nutrients are, then the eddy state (shown in red in the figure) will be a critical parameter to include in the marine metabolomic model described above.

Certain metabolites have been studied previously with respect to SLA, most notably chlorophyll [and pheophytin] [for its role in estimating the state of the phytoplankton community]. Fluorometric measurements of chlorophyll are regularly taken during CTD casts and the deep chlorophyll maximum (DCM) represents an important water column feature for oceanographic sampling [and biological blah]. The DCM has been noted previously to fluctuate up to [XX] meters in response to fluctuations in sea surface height, indicating that at least one metabolite is sensitive to eddy state in a biologically significant way. However, the amplitude of this fluctuation is much less dramatic than that of the nutricline or isopycnal surfaces because it represents a compromise between the constant influx of light from above and the varying nutrient field, with light essentially serving as a stabilizing force for the DCM.

In this paper, we describe the metabolome's variability with respect to depth and sea level anomaly with the goal of constraining these parameters in a [future? hypothetical?] global ocean model of marine metabolites. The driving questions we answer are listed below:

1. Is the metabolome as a whole significantly affected by depth, time of day, and sea level anomaly? Which of these parameters causes the most variability in the metabolome and is therefore most important to include in a global ocean model?

2. How do individual metabolite depth profiles vary in the upper ocean? Can we distinguish labile compounds that are readily degraded from those that are more recalcitrant? Are there chemical or biological similarities between metabolites that respond similarly to changes in depth?

3. Which metabolites are sensitive to eddy state? Do they respond more like nutrients and isopycnal surfaces, or are they stabilized by light availability as the DCM is? Are there any metabolites that can be found more abundantly or even exclusively in cyclonic (or anticyclonic) eddies, perhaps serving as biomarkers for the biological community within?

These questions have not been answered previously because measuring compound abundances in the ocean is difficult. Marine metabolites typically have concentrations in the nanomolar-picomolar range and the small, polar molecules in particular are difficult to separate from the enormously more abundant salt ions. Fortunately, recent advances in mass spectrometry and liquid chromatography have produced methods that allow the measurement of these low-abundance molecules with sufficient accuracy and rapidity for this type of analysis. Here, we use a high-resolution accurate-mass instrument coupled to a hydrophilic interaction column to rapidly assay all the small molecules in a given sample.

Additionally, advances in data analysis and computational power have made untargeted mass-spectrometry increasingly viable as a technique for detecting and quantifying molecules that exist beyond the limited suite traditionally pursued in environmental metabolomics. Unlike targeted MS, which only analyzes a select few compounds known in advance, untargeted metabolomics is data-driven and can pick out signals produced by compounds that are unknown or unannotated. This is a significant advantage in environments such as the ocean where comprehensive databases have yet to be established and provides a way to make mass-spectrometry analysis perfectly reproducible via scripting.

## Methods

### Cruise information

To answer the questions established above, samples were collected from the North Pacific Subtropical Gyre near Station ALOHA during two research cruises that targeted strong mesoscale eddy features. The first cruise, KM1709, occurred during June and July 2017 and traversed an area between 28 °N, 157° W and 23 °N, 158.5 °W. For this cruise, an eddy dipole off the coast of Hawaii was detected using sea-level anomaly (SLA) satellite data and targeted for both a transect across the cyclonic and anticyclonic poles of the eddy dipole as well as two long-term Lagrangian stations at the center of each eddy. The cyclonic pole of the eddy had a maximum negative SLA anomaly of -20 centimeters and the anticyclonic center reached 24 centimeters. The transect samples were taken at various times of day as the ship transected the dipole, and the eddy centers were all sampled between 5 and 8 pm. (See Supplemental Table XX for cruise data) The second cruise occurred during March and April of 2018 and only sampled from the center of each eddy. This Falkor cruise (cruise number FK180310) also traveled between two eddies of different polarities, a cyclonic eddy at 22 °N 156 °W and an anticyclonic eddy at 24.5 °N and 161 °W. The cyclonic eddy was sampled first and had a maximum negative SLA difference of -12 cm while the anticyclonic eddy was sampled one day later and had a maximum SLA of 20 cm. Samples on this cruise were taken twice a day, once at 6AM and once at 6PM. (Supp table XX)

### Environmental data collection

For collaborators to fill in. Need details on every bit of environmental metadata:

  - Bottle 
    - Chlorophyll/phaeophytin
    - Nutrients
    - PC/PN (note outliers)
    - DON/DOP
    - Alkalinity/DIC/pH
    - Theta? This is corrected pressure, right? Not just pressure?
    - Oxygen
    - Pro/Syn/Picoeuk/Het
  - Beam attenuation
  - PAR?
  - SLA (pulled from CMAP)

Environmental parameters were interpolated to the locations at which metabolite samples were taken using a [bilinear algorithm].

### Metabolite sample collection

Samples were obtained using the onboard CTD rosette to collect water from 15 meters, the deep chlorophyll maximum, and 175 meters during KM1709 eddy transect, from the DCM ±10 and ±20 meters during the KM1709 eddy center sampling, and from 25 meters and the DCM during FK180310. The DCM was determined visually from fluorometer data (XX fluorometer from XX company) during the CTD downcast and Niskin bottles were tripped during the return trip to the surface. 10 liters of seawater were collected in each Niskin and each depth was sampled in triplicate by firing multiple bottles simultaneously. Samples were brought to the surface and decanted into prewashed (3x with DI, 3x with sampled seawater) polycarbonate bottles for filtration. Samples were filtered by peristaltic pump onto 142mm 0.2 µm Durapore [Omnipore?] filters [part number] held by polycarbonate filter holders [part number] on a Masterflex [part number] tubing line. Pressures were kept as low as possible while still producing a reasonable rate of flow through the filter, approximately 250 mL per minute. Samples were then removed from the filter holder using solvent-washed [ethanol?] tweezers and placed into pre-combusted aluminum foil packets that were then flash-frozen in liquid nitrogen before being stored at -80 °C until extraction. A methodological blank was also collected by running filtrate through a new filter and then treated identically to the samples.

### Sample extraction

Extraction followed a modified Bligh & Dyer approach as detailed in Boysen et al. (2018). Briefly, filters were added to PTFE centrifuge tubes with a 1:1 mix of 100 µm and 400 µm silica beads, approximately 2mL -20 °C Optima-grade DCM, and approximately 3mL -20 °C 1:1 methanol/water solution (both also Optima-grade). Extraction standards were added after this step, a table of which is available as SI Table XX. The samples were then bead-beaten three times, followed by triplicate washes with fresh methanol/water mixture. Samples were then dried down under clean N2 gas and kept warm using a Fisher-Scientific Reacti-Therm module. Dried aqueous fractions were re-dissolved in 380 µL of Optima-grade water and amended with 20 µL isotope-labeled injection standards, also documented in SI Table XX. The dry organic fraction was reconstituted into 380 µL 90:10 DCM:toluene [right?] rather than Boysen et al.’s 1:1 water:acetonitrile. Both fractions were also finally syringe-filtered [part number?] to remove any potential clogging material. The aqueous fraction was then aliquot into an HPLC vials one for injection on the HILIC column. The organic fraction was not run. [Were the samples themselves diluted? I think yes?] A pooled sample was created by combining 20 µL of each sample into the same HPLC vial, and a 1:1 dilution with water half-strength sample was aliquot from that to assess matrix effects and obscuring variation per Boysen et al. (2018). Also run alongside the environmental samples were two mixes of authentic standards in water and in an aliquot of the pooled sample at a variety of concentrations, available in SI Table XX, for quality control, annotation, and absolute concentration calculations. HPLC vials containing the samples were frozen at -80 °C until thawing shortly before injection.

### HPLC-MS methods

[Verbatim from CH_Ingalls_Lab_LC_Methods from Katherine’s MW submission] A SeQuant ZIC-pHILIC column (5 um particle size, 2.1 mm x 150 mm, from Millipore) was used with 10 mM ammonium carbonate in 85:15 acetonitrile to water (Solvent A) and 10 mM ammonium carbonate in 85:15 water to acetonitrile (Solvent B) at a flow rate of 0.15 mL/min. The column was held at 100% A for 2 minutes, ramped to 64% B over 18 minutes, ramped to 100% B over 1 minute, held at 100% B for 5 minutes, and equilibrated at 100% A for 25 minutes (50 minutes total). The column was maintained at 30 °C. The injection volume was 2 µL for samples and standard mixes. When starting a batch, the column was equilibrated at the starting conditions for at least 30 minutes. To improve the performance of the HILIC column, we maintained the same injection volume, kept the instrument running water blanks between samples as necessary, and injected standards in a representative matrix (the pooled sample) in addition to standards in water. After each batch, the column was flushed with 10 mM ammonium carbonate in 85:15 water to acetonitrile for 20 to 30 minutes.

Metabolomic data was collected on a Thermo Q Exactive HF hybrid Orbitrap (QE) mass spectrometer.

The capillary and auxiliary gas heater temperatures were maintained at 320°C and 100°C, respectively. The S-lens RF level was kept at 65 [units], but the H-ESI voltage was set to 3.3 kV instead and sheath gas, auxiliary gas, and sweep gas flow rates were decreased to 16, 3, and 1 [units], respectively. Polarity switching was again used but with a scan range of 60 to 900 m/z and a resolution of 60,000. Calibration was performed every 3-4 days using the [Thermo Calibration Mix (part number?)] at a target mass of 200 m/z. DDA data was collected by fragmenting the pooled samples with [details??? Same in HILIC and RP?]. All files were then converted to an open-source mzML format via Proteowizard’s msConvert tool [version number??] with flags and filters specified in Supplemental Table XX [centroiding, pos/neg separation].

### Targeted metabolomic methods

Targeted compounds were manually integrated using Skyline [version number] and MSDIAL [version number] with parameters specified in Supplemental Table XX. Skyline was used for the KM1709 and FK180310 eddy center samples, while MSDIAL was used for the KM1709 eddy transect data. Raw peak areas were then normalized to their best-matched internal standard per Boysen et al. (2018). The coefficient of variance (CV) threshold used to choose a suitable match was X.X, as determined by looking at the internal standards normalized to each other and selecting the minimum CV that did not introduce new variation, per the tools in the B-MIS normalization GitHub repository (https://github.com/IngallsLabUW/B-MIS-normalization). Normalized peaks were then calibrated via a single-point standard curve as determined uniquely for each compound from the authentic samples used to convert normalized peak area into moles per µL injected and scaled to the amount of seawater filtered to provide a final estimate for each compound of moles per liter of seawater filtered.

### Untargeted metabolomic methods

Untargeted compounds were initially analyzed using the xcms package [version] from the Bioconductor project [version] in the R programming language [version]. Parameters for import and peakpicking can be found in Supplemental Table XX, and the functions readMSdata and findPeaks were used to produce a list of raw peak areas. However, because the default xcms algorithm has several flaws in its signal-to-noise ratio (SNR) estimation, [cite Myers & Du 2017] custom code was written to remove many of these features likely to be noise. To estimate SNR more robustly, the raw m/z, retention time, and intensity data was extracted from the mzML files using the R package RaMS [version] and subset for each peak using the mzmin, mzmax, rtmin, and rtmax values produced by xcms. Each peak was matched against a Gaussian curve with varying degrees of skew [provide values!] to account for peak tailing effects, and the curve with the smallest residuals was chosen for further analysis. The raw data was then subtracted from this ideal curve to estimate the residual variation (noise) present in the peak, and the maximum height of the peak was divided by the standard deviation of these residuals to estimate SNR. The advantage of this method is that it can be used on peaks that do not have many data points outside of the peak itself, a feature common to HILIC data. This is unlike the existing algorithms in xcms and MzMine2 which require a large window of background points to correctly calculate SNR. This new SNR was used to calculate an overall “peak quality” index calculated as

$$ SNR * peak\ correlation^4*max\ intensity$$

where peak correlation is the Pearson’s correlation coefficient calculated between the Gaussian curve and the real data and max intensity is the height of the peak. Peaks with a quality score less than 20 were excluded from the remainder of the analysis. [Defend the choice of 20]. This list of peaks was then fed back into the default XCMSnExp object and followed by the default functions for alignment and correspondence, parameters for which can also be found in Supplemental Table XX. Finally, a custom version of xcms’ default fillPeaks algorithm was used due to bugs in the source code that prevented the appropriate parameters from being used.

After the curated list of peaks was produced, isotopes and adducts were removed. Isotopes and adducts were detected by again accessing the raw data via the R package RaMS [version number] and subsetting between rtmin and rtmax, but instead the new m/z limits were calculated by assuming the detected peak was the isotope or adduct of another and that assumed peak’s mass was used for the subsetting. The raw data from that m/z offset was then compared via Pearson correlation to the raw data of the original peak for a raw similarity comparison. Additionally, the Pearson correlation of the peak areas for each feature across multiple files was calculated as an orthogonal metric of similarity. If the raw similarity correlation was above 0.8 and the correlation between files was above 0.99, the feature was assumed to be an adduct or isotope and removed from downstream analysis as well. The list of adducts and isotopes checked for each feature can be found in Supplemental Table XX.

While untargeted data analysis makes it challenging to provide absolute quantification for given molecules, the authentic standards run for the targeted analysis made it possible to annotate a portion of the untargeted peaks as known compounds at a confidence level of 1 in the MSI reporting criteria (https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3772505/). This was accomplished by comparing each known metabolite to the data collected from the authentic standards in water data. The highest annotation certainty was assigned to those compounds with isotopically labeled internal standards – if a known metabolite had an internal standard, the m/z value for both the compound and its isotopologue were calculated. All peaks for each m/z value were collected and those with the closest retention time were assumed to be the isotope and its internal standard, removing them from further annotation consideration. The next highest certainty was calculated using the multiple authentic standard mixes we ran alongside the samples themselves. To discriminate between isomers of known compounds, these mixes were designed to alternate between isomers in retention time such that the distance between sequential peaks in a given mix was maximized. This allows us to calculate a between-mixes ratio that can then be compared to the known mix in which a given compound was detected. If the feature appears in the correct mix (defined as having an area in the correct mix one full standard deviation larger than that of the areas detected in the incorrect mix), it is assigned the metabolite name and removed from further annotation steps. The next highest certainty was calculated for known metabolites that had multiple features at the same m/z value, typically isomers. If the number of features matched the number of authentic standards expected at a given m/z, they were annotated by rank-order retention time matching (i.e. the first feature was assumed to be the authentic standard with the lowest retention time, followed by the second feature assumed to be the authentic standard with the second-lowest retention time, etc.). Finally, peaks were annotated using retention time similarity to the expected value in the standards list and ultimately annotated as the peak with the largest area if all the above failed. [Revisit the above to explain interactions and the whole if/else thing]. Compounds that were multiply assigned (either one feature assigned to multiple authentic standards or one standard assigned to multiple features) were manually deconvoluted.

Untargeted metabolites were also normalized to a best-matched internal standard (IS) as defined above in the targeted metabolomics section. However, the IS were manually integrated using Skyline because even small outliers in integration area significantly affected the downstream analysis after compounds were normalized to them. The cutoffs used for untargeted data analysis were [higher/lower] than those in the targeted suite, with an “already good” CV of 0.X and a minimum improvement of 0.X.

Chemical formulae were estimated from each metabolite's m/z ratio and isotope envelope information. We implemented this estimation using three separate approaches, using first a set of custom functions, then the R package Rdisop version X.XX.XXX, and finally SIRIUS version X.XX.XXX. To create isotope envelopes for detected features, we wrote code very similar to that used to remove the adducts and isotopes above, but rather than assuming the feature was an isotope or adduct we simply looked at the m/z values assuming that the feature was the [M+H]$^+$ adduct in positive mode and the [M-H]$^-$ adduct in negative mode. The adducts and isotopes tested can also be found in Supplemental Table XX. After calculating the m/z values for each feature, the raw data was again accessed using RaMS and the trapezoidal Riemann sum was calculated between the minimum retention time and the maximum retention time for the feature across each file and adduct or isotope. Raw data with a Pearson correlation above 0.75 and a correlation between files of 0.9 or greater was assumed to be an isotope of the base peak.

For the custom calculations, these values were used to estimate the ratio of heavy to light abundance. A Type I linear model was fit to each feature’s isotope ratios and the slope of the line was then used to estimate the number of atoms contained in each formula using the calculations provided in Supplemental Methods XX. This custom method provided a way to constrain the results produced by Rdisop and SIRIUS. For those functions, the file-specific isotope envelope had to be collapsed into a feature-based format. The average monoisotopic area was used for the base peak and multiplied by the slope of the line from the linear model to obtain an average isotope ratio across the multiple files. Next, the Rdisop’s “decomposeIsotopes” function was used with a ppm error of 5 if the m/z ratio was above 200 or a fixed tolerance of 0.001 Da otherwise. The potential formula candidates output from this function were then filtered by presence/absence in Pubchem while relaxing the formula to contain one more or fewer hydrogen atoms than initially estimated to account for [M+H]+ and [M-H]- adducts. Finally, SIRIUS command-line functions were used to provide a third orthogonal estimate of molecular formula. The “decomp” function was passed the m/z value for each peak and constrained with a ppm error of 5 and an element list of CHNOPS. Only formulae that appeared in both Rdisop and SIRIUS and also had an elemental breakdown matching the custom calculations were kept for further analysis.

The above untargeted process was applied separately to each combination of column type, polarity, and sample run. This resulted in six unique peak lists that required correspondence across analyses. The six HILIC runs were correspondedby selecting the feature with the largest area and features from the other runs near the largest feature in *m/z* and retention time were selected, with the features from each other run closest in area and retention time according to log$_2$ retention time and log$_{10}$ average area identified as most likely candidates and grouped. Those features were then removed from consideration and the correspondence was repeated for the next largest peak until no features remained ungrouped. Features that had no corresponding feature in other runs became their own group. Finally, grouped features that contained a targeted compound were used to identify precision and accuracy of the grouping. A manual retention time correction of 60 seconds was applied to the features detected in the FK180310 cruise after peak visualization showed a consistent offset of approximately that distance in the data.

### Statistics

Fourteen environmental variables were measured during the eddy transect alongside the metabolite samples ([Figure|Table XX]). We confirmed that these parameters were able to uniquely characterize differences due to depth in the eddy transect with both univariate and multivariate methods. First, we tested for differences due to sampling depth using Welch's ANOVA (@welch1951), chosen because variances could not be assumed equal for most measurements and corrected for multiple comparisons using Holm's method (@holm1979). Next, we distributed the samples into a two-dimensional space using non-metric multidimensional scaling (NMDS) as implemented in R's vegan package with the metaMDS function. Two dimensions were chosen because one dimension would not provide sufficient explanatory power and the stress value was below 0.1 when two dimensions were used, rendering a third was unnecessary. For this NMDS analysis, each environmental variable was z-score scaled using base R's scale function to have a mean of zero and unit standard deviation, thereby making each parameter unit-invariant and allowing us to disable autotransformation. Finally, the environmental vectors were fit onto this ordination using vegan's envfit function, allowing us to determine the direction in the ordination space in which each vector changes most quickly and estimate the significance of each fitted factor using permutation.

To assess the significance of depth, eddy polarity, and diel variability on the metabolome as a whole, we performed a Permutational Multivariate Analysis of Variance (PERMANOVA) using R's vegan package (version XX). For this analysis, the QC'd and BMIS'd peak areas were ranked using R's rank function with default ties.method argument of "average" to place all metabolites on the same scale. Discrete SLA bins were created by categorizing each station as cyclonic if the corrected SLA was less than -5cm, anticyclonic if greater than 5cm, or neither if between these endpoints (@barone). Discrete time categories were created by binning sample times into morning, midday, evening, and night if the sample was collected between 3:30am and 9:30am, 9:30am and 3:30pm, 3:30pm and 9:30pm, or 9:30pm and 3:30am, respectively. These bins were chosen [arbitrarily a priori, see if changes happen when using the MAHALO bins instead per Josh's recommendation]. Depth was already discretized by the depth at which the sample was taken: 15 meters, the DCM, and 175 meters. The adonis2 function was used with a model formula consisting of a ranked metabolite data matrix as described above as the LHS and a RHS of either no interactions, pairwise interactions, or all interactions. Significance was assessed marginally by setting the "by" argument to "margin" to avoid the potentially confounding effect of assessing terms sequentially, and pairwise distances were calculated using the Manhattan distance metric. Nonmetric multidimensional scaling (NMDS) was performed with the metaMDS function, also in the vegan package, again using the Manhattan distance metric. Three dimensions were chosen because two-dimensional solutions consistently exceeded a stress of 0.15. Autotransform was disabled because the data have already been normalized via ranking. The Manhattan distance metric was chosen for its more intuitive performance in higher dimensions, but the results did not change [very much? significantly?] when the Euclidean distance metric was used instead [double check this].

To characterize the response of metabolites to depth, we ran a Kruskal-Wallis on each compound/parameter followed by Dunn's post-hoc test for pairwise comparisons as implemented by the rstatix's package dunn_test function with no p-value adjustment if the Kruskal-Wallis was significant at an alpha level of 0.05 after adjusting p-values to control the false discovery rate using R's p.adjust function. The Kruskal-Wallis test was chosen instead of the traditional parametric ANOVA because compounds and parameters could not be assumed to have equal variance or normal error distributions, and Dunn's test was chosen as the post-hoc for the same reason. The Dunn test p-values were not corrected for multiple comparisons because [blah]. Compounds were then classified into one of twelve categories by the sign, significance, and magnitude of the post-hoc test results, as well as a null category (no detectable response with depth) and a fourteenth category for compounds with a significant Kruskal-Wallis but for which Dunn's test was unable to significantly discriminate between depth pairs. These categories [emerge from permuting boxplot thingys and there's twelve for REASONS and it's not arbitrary] [I need to describe how I assigned the categories in more detail but haven't figured out how to do that yet - maybe a table?].

To test the hypothesis that compounds most abundant at the surface would have the highest C:N ratios and compounds most abundant deeper in the ocean would have the lowest C:N ratios, we [blah blah blah, used the vertical axis of the clockplot as a proxy for "most abundant at surface" vs "most abundant at depth" and ran a regression of C:N ratio vs y-axis coordinate on 1) only known metabs, 2) all metabs with estimated molecular formulae].

To measure the effect of SLA at each sampled depth, we ran a simple linear regression (Model I) on each metabolite at each depth individually, using SLA as the predictor variable and rank-normalized metabolite abundance as the dependent. Separating the data set in this way allowed us to estimate the impact of SLA on metabolite abundance [without the confounding effect of variable responses due to depth]. Compounds were then categorized into a "positive trend with SLA" category if they had a significantly positive slope, a "negative trend with SLA" if they were significantly negative, or a null category of no significant trend if SLA did not have a detectable effect on metabolite abundance.

We assessed the relative importance of sea level anomaly (SLA) on the metabolome by using the separate set of samples collected from the center of the eddies on the KM1709 cruise. These depths were chosen to represent a large, but not enormous, amount of variability due to niche partitioning in the water column. We then compared metabolite abundances between the two eddies using a simple Student's t-test and grouped all five unique depths together.

We measured the similarity of the metabolome by calculating a distance matrix between each metabolite that appeared in all three data sets after normalizing each to the maximum value observed for that metabolite within each data set. We then calculated Pearson's correlation coefficient between these three distance matrices, allowing us to see how well each metabolome corresponded to each other. To compare the robustness of individual compounds, we ran permutational tests for differences due to SLA on all three data sets; a t-test for the Falkor and eddy center data and a regression for the eddy transect data. We then identified compounds that responded significantly to SLA at an alpha level of 0.05 across different data sets.

All statistics were performed in the R programming environment, [version].

## Results

### Environmental data

Measured values for all fourteen environmental variables are shown below in [Table|Figure] XX. Much of the variance observed in these values is explained by the collection of samples at three discrete depths: 15 meters below the surface, the deep chlorophyll maximum, and 175 meters. DCM depth was not a fixed distance from the surface but instead fluctuated with sea level anomaly in response to nutrient availability such that the DCM samples varied in their absolute depth from 100 to 133 meters.

```{r metadata boxplots & table}
distinct_MT_metadata <- filled_file_metadata %>%
  filter(cruise=="MT") %>%
  mutate(basename=str_remove(untripl, "180821_Smp_")) %>%
  select(basename, depth, sla:last_col()) %>%
  group_by(basename) %>%
  summarise(depth=unique(depth), across(-depth, mean)) %>%
  select(-PAR, -PAR_perc, -ph, -dic, -alk, -oxy_um, -theta) %>%
  filter(!is.na(basename))

# If we decide to use a figure (boxplots of values by depth)
distinct_MT_metadata %>%
  mutate(depth=factor(depth, levels = c("175m", "DCM", "15m"))) %>%
  select(basename, depth, sla, ends_with("abund"),
         ends_with("um"), ends_with("ug"), beam_atten) %>%
  pivot_longer(-c(basename, depth, sla)) %>%
  mutate(name=factor(name, levels = factor_levels, labels = twoline_labels)) %>%
  filter(!is.na(value)) %>%
  ggplot() +
  geom_boxplot(aes(x=value, y=depth)) +
  facet_wrap(~name, scales="free_x", nrow=3) +
  theme(axis.title = element_blank())

# If we decide to use a table
distinct_MT_metadata %>%
  select(basename, depth, ends_with("abund"),
         ends_with("um"), ends_with("ug"), beam_atten) %>%
  # Calculate summary statistics for each environmental parameter
  pivot_longer(-c(basename, depth)) %>%
  group_by(depth, name) %>%
  summarize(print_col=paste(
    signif(mean(value, na.rm=TRUE), digits = 3), "±", 
    signif(sd(value, na.rm=TRUE), digits = 1)
  ), .groups = "drop") %>%
  # Pivot into nice format
  pivot_wider(names_from="depth", values_from = print_col) %>%
  mutate(name=factor(name, levels = factor_levels, labels = factor_labels)) %>%
  arrange(name)
```

```{r welch ANOVA MT}
welch_aov <- distinct_MT_metadata %>%
  select(basename, depth, sla, ends_with("abund"),
         ends_with("um"), ends_with("ug"), beam_atten) %>%
  pivot_longer(-c(basename, depth, sla)) %>%
  # Perform the ANOVA
  nest(data=-name) %>%
  mutate(aovtest=map(data, ~broom::tidy(oneway.test(.x$value~.x$depth)))) %>%
  suppressMessages() %>%
  unnest(aovtest) %>%
  # Format p-values nicely
  mutate(p.value=p.adjust(p.value, method = "holm"))
```

All fourteen environmental parameters differed significantly with depth with beam attenuation, heterotrophic bacteria abundance, chlorophyll a concentration, and phaeophytin concentration differing most with depth and nitrate concentration differing least with respect to depth. [Note, however, that this assumes no interaction with sea level anomaly or time [because I couldn't figure out how to do this] [and it's possible that some parameters differ only with respect to a confounding variable].] 
N + N, phosphate, and DIC increased more or less linearly with depth while several environmental parameters (pH, density) and biological abundances (*Prochlorococcus*, *Synechococcus*, and heterotrophic bacteria) had the opposite trend, following DON and DOP as they linearly decreased. 
As expected, chlorophyll had a maximum at the DCM along with phaeophytin and O_2_. Both particulate carbon (PC) and particulate nitrogen (PN) had high values at the surface and through the DCM, dropping off at the 175m depth horizon, agreeing well with particle load measured via beam attenuation. Picoeukaryote abundance notably had its own profile somewhere between a DCM max and a decrease with depth, with highest abundances at the DCM but values at the surface also relatively high.

```{r}
# If we decide to include a heatmap showing correlations among the metadata
heatmap_cor <- filled_file_metadata %>%
  filter(cruise=="MT") %>%
  filter(samp_type=="Smp") %>%
  select(abs_depth, time, lat, sla, local_hour, 
         ends_with("abund"), ends_with("um"), ends_with("ug")) %>%
  data.matrix() %>%
  cor(use = "pair")

coph_coef <- cor(cophenetic(hclust(dist(heatmap_cor))), dist(heatmap_cor))
# pheatmap::pheatmap(
#   mat = heatmap_cor, color = viridisLite::viridis(100),border_color = "grey80"
# )

heatmap_labels <- c(factor_labels)
heatmap_levels <- c(factor_levels)
names(heatmap_labels) <- heatmap_levels

hc <- hclust(dist(t(heatmap_cor)))
hclust_labels <- heatmap_labels[hc$labels[hc$order]]

ggcor_df <- heatmap_cor %>%
  as.data.frame() %>%
  rownames_to_column("x") %>%
  pivot_longer(!x, names_to = "y") %>%
  # mutate(x=factor(x, levels = heatmap_levels, labels = heatmap_labels)) %>%
  # mutate(y=factor(y, levels = heatmap_levels, labels = heatmap_labels)) %>%
  mutate(x=factor(x, levels = names(hclust_labels), labels = hclust_labels)) %>%
  mutate(y=factor(y, levels = names(hclust_labels), labels = hclust_labels)) %>%
  arrange(x, y) %>%
  filter(as.numeric(x)<=as.numeric(y))
ggcor_labs <- ggcor_df %>% distinct(x) %>% mutate(y=row_number()-0.8)

ggplot() +
  geom_tile(aes(x=as.numeric(x), y=as.numeric(y), fill=value), data = ggcor_df) +
  geom_text(aes(x=as.numeric(x), y=as.numeric(y), label=x), data = ggcor_labs,
            angle=90, hjust=1, size=12*0.8/ggplot2::.pt) +
  geom_text(aes(x=0, y=as.numeric(y)+1, label=x), data = ggcor_labs,
            hjust=1, size=12*0.8/ggplot2::.pt) +
  geom_dendro(hc, ylim=c(19.5, 22), lwd=1) +
  annotate("text", x=Inf, y=Inf, label=paste("Coph coef:", round(coph_coef, 3)), 
           hjust=1, vjust=1) +
  scale_fill_gradient2(limits=c(-1, 1), name="Pearson's\ncorrelation\ncoefficient") +
  scale_x_discrete(position = "top") +
  theme_minimal() +
  theme(axis.title = element_blank(),
        legend.position = c(0.9, 0.3),
        axis.text = element_blank(),
        panel.grid = element_blank()) +
  coord_fixed() +
  scale_y_continuous(limits = c(-3, 22)) +
  scale_x_continuous(limits = c(-7, 21))
```

Two-dimensional non-metric multidimensional scaling (NMDS) of the environmental parameters distinctly separated the samples by depth, with samples taken from 15 meters, the DCM, and 175 meters falling into separate clusters.

```{r 2D NMDS plot}
nmds_metadata <- distinct_MT_metadata
nmds_output <- nmds_metadata %>%
  column_to_rownames("basename") %>%
  select(-depth) %>%
  data.matrix() %>%
  scale() %>%
  # rank() %>%
  vegan::metaMDS(distance = "manhattan", k = 2, trace = 0,
                 autotransform = FALSE, na.rm=TRUE)
nmds_df <- nmds_output$points %>%
  as.data.frame() %>%
  rownames_to_column("basename") %>%
  left_join(nmds_metadata, by="basename")
envfit_df <- nmds_metadata %>% 
  select(-depth) %>%
  column_to_rownames("basename") %>%
  vegan::envfit(ord=nmds_output, permutations=999, na.rm=TRUE) %>% 
  `$`("vectors") %>% `[`(1:4) %>%
  do.call(what = cbind) %>%
  as.data.frame() %>%
  set_names("MDS1", "MDS2", "r", "perm", "pval") %>%
  mutate(across(starts_with("MDS"), `*`, 20)) %>%
  select(-perm) %>%
  rownames_to_column("var") %>%
  mutate(var=factor(var, levels = factor_levels, labels = factor_labels)) %>%
  mutate(env_label=ifelse(pval<0.05, as.character(var), paste(var, "(ns)")))
ggplot(nmds_df, aes(x=MDS1, y=MDS2)) +
  geom_segment(aes(xend=0, yend=0), envfit_df, arrow = 
                 arrow(ends = "first", length = unit(0.02, "npc"), 
                       type = "closed")) + 
  geom_point(aes(color=depth)) +
  ggrepel::geom_label_repel(aes(label=env_label), envfit_df, 
                            min.segment.length = 0, nudge_y = 0.1) + 
  theme_minimal() +
  annotate("label", x = Inf, y=Inf, hjust=1, vjust=1,
           label=paste("2D NMDS Stress:", round(nmds_output$stress, 3)), 
           fill="black", color="white") +
  labs(color="Depth") +
  coord_fixed()
```

Fitting the environmental vectors onto this ordination illustrated that many parameters had significant collinearity [as expected from the correlation matrix and boxplots above but can't say that if I don't keep them all], with nitrate, phosphate, and silica all correlating with depth and DON and DOP doing so inversely. *Synechococcus* abundance correlated neatly with DON and DOP, while abundances of *Prochlorococcus* and heterotrophic bacteria remained high through the DCM alongside particulate nitrogen, particulate carbon, and beam attenuation. Chlorophyll and pheophytin concentrations were highest at the DCM, as expected, and corresponded to the highest level of picoeukaryotes as well. Neither sea level anomaly nor local hour were found to change significantly across the ordination (permutational p = 0.34 and 0.69, respectively).

### Overall metabolite response

```{r 3D NMDS plot}
nmds_output <- MT_peaks %>%
  left_join(MT_feats, by="feature") %>%
  filter(polarity=="pos") %>%
  group_by(feature) %>%
  mutate(value=rank(norm_area)) %>%
  # mutate(value=norm_area/max(norm_area)) %>%
  # mutate(value=scale(norm_area)[,1]) %>%
  ungroup() %>%
  select(feature, value, filename) %>%
  pivot_wider(names_from = feature) %>%
  column_to_rownames("filename") %>%
  vegan::metaMDS(distance = "manhattan", k = 3, trace = 0, autotransform = FALSE,
                 wascores = FALSE, noshare = FALSE)
nmds_df <- nmds_output$points %>%
  as.data.frame() %>%
  rownames_to_column("filename") %>%
  left_join(filled_file_metadata, by="filename") %>%
  select(filename, starts_with("MDS"), depth, sla, local_hour) %>%
  mutate(basename=str_extract(filename, "MS\\d+C\\d+.*(?=_[A-C])")) %>%
  mutate(disc_hour=case_when(
    between(local_hour, 4, 9)~"morning",
    between(local_hour, 10, 15)~"midday",
    between(local_hour, 16, 21)~"evening",
    TRUE~"night"
  ))
nmds_labels <- nmds_df %>%
  group_by(depth) %>%
  summarise(MDS1=mean(MDS1), MDS2=mean(MDS2), MDS3=mean(MDS3))
mds12 <- ggplot(nmds_df, aes(x=MDS1, y=MDS2)) +
  geom_polygon(aes(fill=depth, group=basename), alpha=0.5) +
  geom_polygon(aes(color=depth, group=basename), fill=NA) +
  geom_label_repel(data = nmds_labels, aes(color=depth, label=depth), 
                   size=6) +
  theme_minimal() +
  theme(legend.position="none") +
  coord_fixed() +
  ggtitle("A)")
mds13 <- ggplot(nmds_df, aes(x=MDS1, y=MDS3)) +
  geom_polygon(aes(fill=depth, group=basename), alpha=0.5) +
  geom_polygon(aes(color=depth, group=basename), fill=NA) +
  geom_label_repel(data = nmds_labels, aes(color=depth, label=depth), 
                   size=6) +
  theme_minimal() +
  theme(legend.position="none") +
  coord_fixed() +
  ggtitle("B)")
mds13_sla <- ggplot(nmds_df, aes(x=MDS1, y=MDS3)) +
  geom_polygon(aes(fill=sla, group=basename), alpha=0.5) +
  geom_polygon(aes(color=sla, group=basename), fill=NA) +
  theme_minimal() +
  coord_fixed() +
  scale_fill_gradient2(high = muted("red"), low = muted("blue"), midpoint = 0.05) +
  scale_color_gradient2(high = muted("red"), low = muted("blue"), midpoint = 0.05) +
  ggtitle("C)")
mds13_time <- ggplot(nmds_df, aes(x=MDS1, y=MDS3)) +
  geom_polygon(aes(fill=disc_hour, group=basename), alpha=0.5) +
  geom_polygon(aes(color=disc_hour, group=basename), fill=NA) +
  theme_minimal() +
  coord_fixed() +
  ggtitle("D)")
gridExtra::grid.arrange(mds12, mds13, mds13_sla, mds13_time,
                        layout_matrix=matrix(c(1,2,3,4), ncol=2))
```

*Figure XX: Non-metric multidimensional scaling plots of the metabolome. Axes 1 and 2 are shown in the subplot a) and axes 1 and 3 are shown for b) through d). Triangles connect biological replicates (same environmental context but collected in separate Niskin bottles) and are colored by sampling depth in a) and b), sea level anomaly (SLA) in c), and time of day in d).*

We also found similar clustering by depth in the metabolome, with each depth having a visually distinct combination of metabolites (Figure XX). In the NMDS plot shown, the overall stress of the system is approximately 0.114 and axes 1 and 3 neatly separate samples by depth while axis 2 represents a source of variation not clearly associated any environmental variable. Axis 1 distinguishes the 175 meter samples from the shallower depths, with a distinct cluster of deep samples on one side of the figure and significant overlap between the DCM and 15 meter samples on the other. Axis 3 distinguishes each depth but does so with some overlap between categories, placing the 15 meter samples all to one side, the 175 meter samples near zero, and the DCM samples opposite the 15 meter samples. When colored by sea level anomaly (Figure XX.C), the deep samples appear to separate within the cluster along axis 1, with samples taken from the anticyclonic eddy aligning more closely with the shallower samples than the samples from the cyclonic eddy.

However, the distribution of samples across NMDS axes was not especially robust under different normalization techniques. When each metabolite was scaled to a percentage of the maximum peak area for that metabolite or when scaled using z-score normalization it was axis 1 that contained the within-triplicate variation while axis 2 separated the 175m samples from the surface and DCM and axis 3 contained similar DCM-175m-15m sequences.

```{r}
rank_mat <- MT_peaks %>%
  left_join(MT_feats) %>%
  filter(polarity=="pos") %>%
  group_by(feature) %>%
  mutate(value=rank(norm_area)) %>%
  # mutate(value=value/max(value)) %>%
  # mutate(value=scale(value)[,1]) %>%
  ungroup() %>%
  select(filename, feature, value) %>%
  pivot_wider(names_from = feature) %>%
  column_to_rownames("filename")
permadata <- filled_file_metadata %>%
  filter(cruise=="MT") %>%
  filter(samp_type=="Smp") %>%
  mutate(disc_sla=case_when(
    sla < -0.05~"cyclone",
    sla > 0.05~"anticyclone",
    TRUE~"neither"
  )) %>%
  mutate(disc_hour=case_when(
    between(local_hour, 4, 9)~"morning",
    between(local_hour, 10, 15)~"midday",
    between(local_hour, 16, 21)~"evening",
    TRUE~"night"
  ))
permanout_all_int <- vegan::adonis2(
  rank_mat~depth*disc_sla*disc_hour, data = permadata, by = "margin", 
  permutations = 999, method = "manhattan"
)
permanout_depthsla_int <- vegan::adonis2(
  rank_mat~depth*disc_sla+disc_hour, data = permadata, by = "margin", 
  permutations = 999, method = "manhattan"
)
permanout_depthtime_int <- vegan::adonis2(
  rank_mat~depth*disc_hour+disc_sla, data = permadata, by = "margin", 
  permutations = 999, method = "manhattan"
)
permanout_timesla_int <- vegan::adonis2(
  rank_mat~depth+disc_sla*disc_hour, data = permadata, by = "margin", 
  permutations = 999, method = "manhattan"
)
permanout_no_int <- vegan::adonis2(
  rank_mat~depth+disc_sla+disc_hour, data = permadata, by = "margin", 
  permutations = 9999, method = "manhattan"
)
```

The less-subjective PERMANOVA confirmed that depth was a major factor in controlling the metabolome, explaining ~32% of the variation with a p-value <0.001. Time of day and SLA were both found to be significant factors as well but explained only 4-5% and 3-4% of the variance in the metabolome, respectively, with p-values in the 0.001-0.005 range. Testing for interactions revealed a significant three-way interaction with p-values in the 0.04-0.05 range explaining ~5% of the variance and a marginally significant two-way interaction between depth and SLA (p ~ 0.09), while time of day didn't interact significantly with either depth or SLA.

### Metabolite response to depth

```{r posthoc_classes and posthoc_table}
kw_diff_feats <- MT_peaks %>%
  left_join(filled_file_metadata, by="filename") %>% 
  select(feature, depth, value=norm_area) %>%
  nest(data=-feature) %>%
  mutate(tidyaov=map(data, ~ broom::tidy(kruskal.test(.x$value~.x$depth)))) %>%
  unnest(tidyaov) %>%
  mutate(pval=p.adjust(p.value, method = "fdr")) %>%
  filter(pval < 0.05) %>%
  select(feature, data) %>%
  unnest(data)
nondiff_feats <- MT_peaks %>%
  filter(!feature%in%kw_diff_feats$feature) %>%
  group_by(feature) %>%
  summarize(resp_class="All equal")

resp_class_levels <- c("All equal", 
    "15m > DCM = 175m", "15m > DCM > 175m", "175m < DCM = 15m",
    "DCM max 15m mid", "DCM > 15m = 175m", "DCM max 175m mid", 
    "15m < DCM = 175m", "15m < DCM < 175m", "175m > DCM = 15m", 
    "DCM min 15m mid", "DCM < 15m = 175m", "DCM min 175m mid", 
    "Hard to tell")

posthoc_classes <- kw_diff_feats %>%
  group_by(feature) %>%
  rstatix::dunn_test(value~depth, p.adjust.method = "none") %>%
  mutate(contrast=paste(group2, group1, sep = "_")) %>%
  select(feature, contrast, est=statistic, pval=p.adj) %>%
  pivot_wider(names_from = contrast, values_from = c(pval, est)) %>%
  mutate(resp_class=case_when(
    pval_DCM_15m<0.05 & pval_175m_15m<0.05 & pval_175m_DCM<0.05 & est_DCM_15m<0 & est_175m_15m<0 & est_175m_DCM<0 ~ "15m > DCM > 175m",
    pval_DCM_15m<0.05 & pval_175m_15m<0.05 & pval_175m_DCM<0.05 & est_DCM_15m>0 & est_175m_15m<0 & est_175m_DCM<0 ~ "DCM max 15m mid",
    pval_DCM_15m<0.05 & pval_175m_15m<0.05 & pval_175m_DCM<0.05 & est_DCM_15m<0 & est_175m_15m>0 & est_175m_DCM>0 ~ "DCM min 15m mid",
    pval_DCM_15m<0.05 & pval_175m_15m<0.05 & pval_175m_DCM<0.05 & est_DCM_15m>0 & est_175m_15m>0 & est_175m_DCM>0 ~ "15m < DCM < 175m",
    pval_DCM_15m<0.05 & pval_175m_15m<0.05 & pval_175m_DCM<0.05 & est_DCM_15m>0 & est_175m_15m>0 & est_175m_DCM<0 ~ "DCM max 175m mid",
    pval_DCM_15m<0.05 & pval_175m_15m<0.05 & pval_175m_DCM<0.05 & est_DCM_15m<0 & est_175m_15m<0 & est_175m_DCM>0 ~ "DCM min 175m mid",
    pval_DCM_15m<0.05 & pval_175m_15m<0.05 & pval_175m_DCM>0.05 & est_DCM_15m<0 & est_175m_15m<0 ~ "15m > DCM = 175m",
    pval_DCM_15m<0.05 & pval_175m_15m<0.05 & pval_175m_DCM>0.05 & est_DCM_15m>0 & est_175m_15m>0 ~ "15m < DCM = 175m",
    pval_DCM_15m>0.05 & pval_175m_15m<0.05 & pval_175m_DCM<0.05 & est_175m_15m<0 & est_175m_DCM<0 ~ "175m < DCM = 15m",
    pval_DCM_15m>0.05 & pval_175m_15m<0.05 & pval_175m_DCM<0.05 & est_175m_15m>0 & est_175m_DCM>0 ~ "175m > DCM = 15m",
    pval_DCM_15m<0.05 & pval_175m_15m>0.05 & pval_175m_DCM<0.05 & est_DCM_15m>0 & est_175m_DCM<0 ~ "DCM > 15m = 175m",
    pval_DCM_15m<0.05 & pval_175m_15m>0.05 & pval_175m_DCM<0.05 & est_DCM_15m<0 & est_175m_DCM>0 ~ "DCM < 15m = 175m",
    pval_DCM_15m>0.05 & pval_175m_15m>0.05 & pval_175m_DCM>0.05 ~ "All equal",
    TRUE ~ "Hard to tell"
  )) %>%
  bind_rows(nondiff_feats) %>%
  mutate(resp_class=factor(resp_class, levels = resp_class_levels)) %>%
  arrange(resp_class) %>%
  select(feature, resp_class)
posthoc_table <- table(posthoc_classes$resp_class)
posthoc_classes %>%
  left_join(MT_feats, by="feature") %>%
  mutate(feature=ifelse(is.na(compound_name), feature, compound_name)) %>%
  arrange(startsWith(feature, "FT"), desc(avgarea)) %>%
  group_by(resp_class) %>%
  summarize(n=n(), example=unique(feature)[1]) %>%
  setNames(c("Response type", "Number of metabolites", "Example metabolite"))
```

Approximately half (`r posthoc_table["All equal"]` / `r sum(posthoc_table)`) of all detected compounds showed no significant (p > 0.05 after FDR correction) response to depth. The metabolites in these different response classes is shown in the figure below and demonstrates a highly non-uniform distribution in this rank-ordered fold-change space. Most points are concentrated in the upper-right corner of the plot, indicating that many metabolites are most abundant in the surface or at the DCM. Compounds that decrease linearly with depth form a tight cluster at the very top of the plot, indicating very clear separation between the three depths, while transitioning more towards those compounds that have a maximum at the DCM causes separation to become less distinct. While more rare, compounds either not found or found at the lowest relative abundances at the surface are present in the data and shown in the bottom right corner of the plot. Notably, the left side of this plot is very sparse, with only three compounds falling into those categories that could be classified as having a minimum at the DCM.

```{r clockplot, fig.height=6, fig.width=6}
rank_peaks <- MT_peaks %>%
  group_by(feature) %>%
  mutate(norm_area=rank(norm_area)) %>%
  # mutate(value=value/max(value)) %>%
  # mutate(value=scale(value)[,1]) %>%
  ungroup()
clockplot_df <- rank_peaks %>%
  left_join(filled_file_metadata, by="filename") %>%
  select(feature, norm_area, depth, filename) %>%
  group_by(feature) %>%
  mutate(mean_area=mean(norm_area)) %>%
  group_by(feature, depth) %>%
  summarise(mean_fold_change=mean(norm_area/mean_area)) %>%
  pivot_wider(names_from=depth, values_from=mean_fold_change) %>%
  left_join(posthoc_classes, by="feature") %>%
  mutate(x=DCM-1) %>%       # cursed but works??
  mutate(y=`15m`+DCM/2-1.5) # super fucking cursed
clock_scale <- 0.66
axis_xends <- c(-1/2, 1.1, -1/2)*clock_scale
axis_yends <- c(sqrt(3)/2, 0, -sqrt(3)/2)*clock_scale
axis_y_dodge <- c(2, 2, -2)*clock_scale/25

color_scale <- c("grey50", hcl(h = seq(15, 375, length=13), l=65, c=100)[1:12], "grey80")
names(color_scale) <- levels(clockplot_df$resp_class)

clock_gp <- ggplot(clockplot_df) +
  annotate("segment", x=c(0,0,0), y=c(0,0,0), 
           xend=axis_xends, yend=axis_yends,
           arrow=arrow(length=unit(0.02, "npc"), type = "closed")) +
  annotate("segment", x=c(0,0,0), y=c(0,0,0),
           xend=-axis_xends, yend=-axis_yends,
           lty = 2) +
  annotate("label", x=axis_xends, y=axis_yends+axis_y_dodge,
           label=c("15m axis", "DCM axis", "175m axis"), 
           fill="#FFFFFFAA", label.size=0) +
  geom_point(aes(x=x, y=y, color=resp_class), alpha=0.8) +
  coord_fixed(xlim = c(-1.1, 1.3)*clock_scale, ylim = c(-1.1, 1.1)*clock_scale) +
  scale_color_manual(values = color_scale, breaks = names(color_scale)) +
  theme_void() +
  theme(legend.position = "right", legend.title = element_blank())

xampleMaker <- function(classes){
  which_cmpds <- posthoc_classes %>%
    filter(resp_class%in%classes) %>%
    left_join(MT_feats, by="feature") %>%
    select(feature, compound_name, resp_class, avgarea) %>%
    arrange(is.na(compound_name), desc(avgarea)) %>%
    group_by(resp_class) %>%
    slice(1)
  xamples_df <- which_cmpds %>%
    left_join(rank_peaks, by="feature") %>%
    group_by(resp_class) %>%
    mutate(value=norm_area-min(norm_area)+1) %>%
    ungroup() %>%
    mutate(feature=ifelse(is.na(compound_name), feature, compound_name)) %>%
    left_join(filled_file_metadata, by="filename") %>%
    select(resp_class, feature, filename, depth, value) %>%
    mutate(depth=factor(depth, levels = c("175m", "DCM", "15m")))
  xamples_label_df <- xamples_df %>%
    group_by(resp_class, feature) %>%
    summarize(midpt=(max(value)-min(value))/2+min(value))
  ggplot() +
    geom_boxplot(aes(x=value, y=depth), data = xamples_df, width=0.8) +
    geom_text(aes(x=midpt, label=resp_class, color=resp_class), y=4.8, data = xamples_label_df) +
    geom_text(aes(x=midpt, label=feature, color=resp_class), y=4.1, data = xamples_label_df) +
    annotate("segment", x=0, y=-Inf, xend=0, yend=3.5) +
    annotate("segment", x=0, y=3.5, xend=Inf, yend=3.5) +
    facet_wrap(~factor(resp_class, levels = classes), scales = "free_x") +
    scale_color_manual(values = color_scale, breaks = names(color_scale)) +
    theme_void() +
    coord_cartesian(ylim=c(0.5, 4.5)) +
    theme(strip.text = element_blank(),
          legend.position = "none",
          axis.text.y = element_text())
}
xample_gp_top <- xampleMaker(c("15m > DCM = 175m", "15m > DCM > 175m", "175m < DCM = 15m"))
xample_gp_bot <- xampleMaker(c("175m > DCM = 15m", "15m < DCM < 175m", "DCM > 15m = 175m"))

gridExtra::grid.arrange(xample_gp_top, clock_gp, xample_gp_bot,
                        layout_matrix=matrix(c(1,2,2,2,3), ncol = 1))
```

*Figure XX: Plot of metabolite response to depth shown across three axes corresponding to the significance, sign, and magnitude of the post-hoc test results. Points aligning with the 15m axis correspond to compounds with most of their abundance found in the surface ocea; points far to the right side correspond to compounds that are found only at the deep chlorophyll maximum; points found at the bottom of the plot are those compounds that increased more or less linearly with depth. Distance from plot center corresponds to increasing fold-change estimate between the depth categories, with most compounds having insignificant fold change with depth and falling in the center of the plot. A few categories are shown as boxplots with selected example compounds above and below the central figure.*

`r unname(posthoc_table["15m > DCM = 175m"])` compounds were found that had significantly higher values at 15 meters than the DCM or 175 meters. N6-Acetyl-L-lysine is the only known compound in this group and is a borderline inclusion, as its depth profile could also be characterized as linearly decreasing - however, the DCM and 175 meter samples were not statistically different.

```{r MAA search, eval=FALSE}
MAA_masses <- readxl::read_xlsx("../metadata/raw_data/MAA_masses.xlsx", sheet = 2) %>%
  select(1, 3, 5, 7, 9) %>%
  setNames(c("MAA", "formula", "mz", "rt", "MSMS")) %>%
  filter(MAA!="Common")

MAA_maybe <- MAA_masses %>%
  pmap_dfr(function(...){
    row_data <- data.frame(...)
    MT_feats %>%
      filter(mzmed%between%pmppm(row_data$mz, 10)) %>%
      mutate(MAA_name=row_data$MAA, MAA_rt=row_data$rt, MAA_formula=row_data$formula)
  }) %>%
  left_join(posthoc_classes, by="feature") %>%
  select(feature, mzmed, rtmed, avgarea, MAA_name, MAA_rt, resp_class)

msms_files <- list.files("../mzMLs/pos/MSMS", pattern = "180821.*DDA", full.names = TRUE)
msmsdata <- grabMSdata(msms_files)

MAA_msms <- MAA_maybe %>%
  mutate(rtmed=rtmed/60) %>%
  pmap_dfr(function(...){
    row_data <- data.frame(...)
    msmsdata$MS2[rt%between%(row_data$rtmed+c(-1, 1))][premz%between%pmppm(row_data$mzmed, 50)] %>%
      mutate(MAA=row_data$MAA_name)
  }) %>%
  group_by(MAA, filename, rt) %>%
  arrange(desc(int)) %>%
  # slice(5) %>%
  summarize(MSMS=paste(fragmz, int, sep = ", ", collapse = "; "))

MAA_msdata <- MAA_maybe %>%
  mutate(rtmed=rtmed/60) %>%
  pmap_dfr(function(...){
    row_data <- data.frame(...)
    v <- msmsdata$MS1[mz%between%pmppm(row_data$mzmed)]
    v$MAA <- row_data$MAA_name
    v
  })
# Plot all the peaks' MS1 and add lines at their MSMS scan RTs
ggplot() +
  geom_line(aes(x=rt, y=int, group=filename), data = MAA_msdata) +
  geom_vline(aes(xintercept=rt, group=filename), data = MAA_msms, color="red") +
  facet_wrap(~MAA, scales = "free_y") +
  xlim(6, 12)

aster_db_MSMS <- MAA_masses %>%
  filter(MAA=="Asterina-330" & !is.na(MSMS)) %>%
  separate_rows(MSMS, sep = "; ") %>%
  separate(MSMS, into = c("fragmz", "int"), sep = ", ", convert = TRUE)
MAA_msms %>%
  filter(MAA=="Asterina-330") %>%
  # mutate(which_peak=ifelse(rt>8, "tallskinnysecond", "shortfatfirst")) %>%
  separate_rows(MSMS, sep = "; ") %>%
  separate(MSMS, into = c("fragmz", "int"), sep = ", ", convert = TRUE) %>%
  group_by(filename) %>%
  mutate(int=int/max(int)) %>%
  filter(int>0.05) %>%
  ggplot(aes(x=fragmz, y=int)) +
  geom_vline(aes(xintercept=fragmz), data=aster_db_MSMS, color="red") +
  geom_point() +
  geom_segment(aes(yend=0, xend=fragmz)) +
  # facet_wrap(~which_peak) +
  theme_bw()
# plotly::ggplotly()
```

[Should I split off the MAA analysis into separate section to be discussed after walking through all the clockplot categories?]

However, several compounds in this category were putatively identified as mycosporine-like amino acids (MAAs) using *m/z* matching and MS^2^ fragmentation information. Asterina-330, shinorine, and palythine were simply identified by checking the fragment masses against those provided in @parailloux2020 because they were the only peak in the chromatogram and were assigned a confidence level of 2 according to MSI reporting standards (@sumner2007). Porphyra-334 had two possible peaks, only the earlier eluting of which was was fragmented in the DDA scan and which matched MSMS spectra in online databases, but without fragmentation for the second peak we cannot be confident of this identification. Both peaks have the same type of response to depth (essentially zero abundance outside of the 15 meter samples) but have minimal correlation with each other in the surface samples, confirming that this is unlikely to be the same compound somehow split or poorly-integrated.

```{r porphyra dupe check, eval=FALSE}
porph_data <- MAA_maybe %>%
  filter(MAA_name=="Porphyra-334") %>%
  left_join(MT_peaks, by="feature") %>%
  select(feature, norm_area, filename) %>%
  pivot_wider(values_from = norm_area, names_from = feature) %>%
  left_join(filled_file_metadata, by="filename")

 porph_data %>%
  ggplot() +
  geom_point(aes(x=FT1409, y=FT1410)) +
  # geom_point(aes(x=log10(FT1409), y=log10(FT1410))) +
  facet_wrap(~depth, scales = "free")

porph_data %>%
 filter(depth=="15m") %>%
  with(cor(FT1409, FT1410))
```

Mycosporine-glycine and mycosporine-2-glycine were more difficult to annotate with a specific confidence level. The peak tentatively identified as mycosporine-2-glycine shared no fragments with the list of common MAA fragments that they put forward and probably remains at confidence level 3. Mycosporine-glycine had two features in the chromatogram at an *m/z* ratio of 246.0972, one short and wide feature eluting earlier around 9.4 minutes and a second taller and sharper feature at 10.2 minutes. DDA data was collected for both features and both shared the common MAA fragment of *m/z* 122.0602 but no others. The earlier eluting short feature had a unique low-intensity fragment replicated across multiple collision energies at an *m/z* of 228.0861, likely representing a water loss. It therefore seems likely that one of these features is mycosporine-glycine but this information alone cannot be used to identify which of the two it is.

It was also difficult to assign a specific confidence level to the isomer couple palythene/usujirene. Three features were found in the *m/z* window around 285.1444 +/- 10 ppm, one of these was unlikely because its retention time was high (12.4 minutes), the feature itself was 10-20x smaller in area than the other two, and no MSMS scans were obtained. We first considered whether our chromatographic conditions could have separated these isomers to produce the two remaining peaks, but closer investigation of the MSMS spectra revealed that several distinguishing fragments were produced upon fragmentation at multiple voltages. The earlier eluting feature with a retention time of 6.9 minutes had two fragments not found in the later eluting feature (*m/z*s of 241.0817 and 226.1310) and the later feature eluting around 8.5 minutes had fragments not seen in the earlier (*m/z*s of 252.1101, 239.1025, and 225.1234). Given that palythene and usujirene are isomers differing only in the cis/trans positioning of a double bond, these distinguishing fragments make it unlikely that one feature is palythene and the other is usujirene and more likely that one of the features is an unknown metabolite. Both features shared several fragments but none of them were MAA-diagnostic as named in @parailloux2020 (*m/z*s of 270.1208, 197.0921, and a curiously small fragment at 60.0451).

Nearly all of the putative MAA peaks described above fell into this first category of compounds that were found abundantly at the surface but with very low abundances at the DCM indistinguishable from the abundances found in the 175 meter samples. For shinorine and the later-eluting mycosporine-glycine, the samples collected at the DCM contained significantly more abundant quantities of the compound than those at 175m, which is our second category of interest in Figure XX above.

Compounds that decreased linearly with depth composed the second-largest category of compounds that responded to depth, with `r unname(posthoc_table["15m > DCM > 175m"])` compounds. These compounds had highest concentrations at the surface, middling concentrations at the DCM, and lowest concentrations at 175 meters. Several known compounds fall into this category, notably dimethylsulfoniopropionate (DMSP) and gonyol, the amino acid leucine, and several betaine molecules - proline betaine and carnitine. Three unknown molecules have also been tentatively identified as betaines in this category - beta-Alanine betaine, for which our lab obtained a standard after these samples were run and can match via retention time ordering, *m/z*, and MSMS for a confidence level very close to 1, and homoserine betaine and threonine betaine which have been identified by MS1 *m/z* matching and MSMS betaine fragments (58.0659, 59.0737, 60.0815) for a confidence level of 2. With retention time ordering, we were also able to annotate each peak uniquely because we have homoserine and threonine in our standards and know that threonine elutes before homoserine.

```{r MSMS for betaines, eval=FALSE}
msms_files <- list.files("../mzMLs/pos/MSMS", pattern = "180821.*DDA", full.names = TRUE)
msmsdata <- grabMSdata(msms_files)

mz_i <- 162.113019 #homoserine betaine, threonine betaine, carnitine
# mz_i <- 132.1018 # Leucine, isoleucine, beta-alanine betaine, alanine betaine(?)

msmsdata$MS1[mz%between%pmppm(mz_i)] %>%
  ggplot() +
  geom_line(aes(x=rt, y=int, group=filename)) +
  geom_vline(aes(xintercept=rt), 
             data = msmsdata$MS2[premz%between%pmppm(mz_i)],
             color="red")

obs_MSMS <- msmsdata$MS2[premz%between%pmppm(mz_i)] %>%
  filter(rt%between%(8.2+c(-1, 1))) %>%
  filter(!fragmz%between%pmppm(mz_i)) %>%
  group_by(filename) %>%
  mutate(int=int/max(int))
theo_MSMS <- "https://raw.githubusercontent.com/IngallsLabUW/Ingalls_Standards/master/MSMS/data_processed/Ingalls_Lab_Standards_MSMS.csv" %>%
  read_csv(show_col_types = FALSE) %>%
  filter(compound_name=="Glycine betaine") %>% 
  separate_rows(MS2, sep = "; ") %>%
  separate(MS2, into = c("fragmz", "int"), sep = ", ", convert = TRUE) %>%
  filter(!fragmz%between%pmppm(132.1018)) %>%
  group_by(voltage) %>%
  mutate(int=int/max(int))

ggplot() +
  geom_point(aes(x=fragmz, y=int), data = obs_MSMS) +
  geom_segment(aes(x=fragmz, y=int, xend=fragmz, yend=0), data = obs_MSMS) +
  geom_point(aes(x=fragmz, y=-int), data = theo_MSMS, col="blue") +
  geom_segment(aes(x=fragmz, y=-int, xend=fragmz, yend=0), data = theo_MSMS, col="blue") +
  geom_hline(yintercept = 0) +
  facet_wrap(~voltage)
```

The group of compounds with equally high values at the surface and the DCM and significantly lower values at 175 meters contained the most compounds (`r unname(posthoc_table["175m < DCM = 15m"])`, or `r paste0(signif(unname(posthoc_table["175m < DCM = 15m"])/sum(posthoc_table), 3)*100, "%")`) other than the category of those that didn't respond to depth at all. This category was dominated in raw peak area [change to concentration w targeted data] by two known compounds, trigonelline and dimethylsulfonioacetate (DMS-Ac), together composing nearly 30% of the signal detected in this size class. Other known compounds in this class include the nucleobase cytosine, the nucleoside adenosine and its deoxy form, creatine, choline, and taurine as well as the amino acids alanine, glutamate, and glutamine. Betonicine and turicine fell into this category but could not be separated by HILIC chromatography as stereoisomers. Acetylcholine was also identified in this category at slightly lower confidence (FT0432, authentic standard purchased and run after this data set and tentative annotation done by *m/z* and retention time matching) as well as hypoxanthine (FT0357) in the same way. An additional large unknown was identified with the same mass as betonicine but eluting a minute earlier perhaps corresponding to 3-Dehydrocarnitine or N-Acetyl-valine based on *m/z* matching.

Compounds abundant at the DCM were also common, with a total of `r sum(unname(posthoc_table[c("DCM max 15m mid", "DCM > 15m = 175m", "DCM max 175m mid")]))` features across the three depth categories that had maxima at the DCM. Many traditional osmolytes were found in this category, including glycine betaine, trimethylamine N-oxide, homarine, and ectoine. Several amino acids and their derivatives fell here - glycine, proline, arginine, tyrosine, asparagine, threonine (combined with homoserine as the two did not appear to separate in retention time [revisit this - I claim that we can tell the betaine versions apart because these DO separate]) and beta-alanine. The nucleobase guanine also had a maximum at the DCM. This collection of compounds was notable to us as containing several nitrogen-rich compounds, which inspired the followup analysis of C:N ratio in section XX below. Several unknown compounds with very large average peak areas were also categorized as having DCM maxima, most notably MT_POS_FT0483 (*m/z* of 151.1441, retention time about 7 minutes) and MT_POS_FT0510 ([need mz and rt info here]) were in the top 20 largest peak areas but neither belonged to our targeted list.

Compounds most abundant at depth are unexpected given general trends of particulate matter decreasing with depth, but prior research has noted several compounds that accumulate with depth (@healarseno?). Here, we find significantly more compounds with this trend than we expected with a total of `r sum(unname(posthoc_table[c("175m > DCM = 15m", "15m < DCM < 175m")]))` compounds that had highest peak areas in samples taken from 175 meters. Arsenobetaine was the only identified compound in these two depth response classes and had a very linear response (Fig above). It's unknown whether these other compounds have a similar role as toxic element shunts or are simply some kind of degradation product, but some of them nonetheless have quite large peak areas. Features MT_POS_FT0617 and MT_POS_FT0618 are especially interesting because they share an *m/z* ratio (170.11757) and differ in retention time by less than a minute (8.3 vs 9 minutes), but both belong to this category of compounds that are most abundant at 175 meters.

Compounds that have a minimum at the DCM are even rarer than those that accumulate with depth, requiring either rapid drawdown at the DCM (as might be the case for a chlorophyll or other DCM-max precursor) or increased synthesis above and below it. Nonetheless, a total of `r sum(unname(posthoc_table[c("DCM min 15m mid", "DCM < 15m = 175m", "DCM min 175m mid")]))` compounds were identified with this trend - none of which belonged to our list of targeted compounds and therefore none of which could be identified at high confidence. Despite the use of stringent false-discovery correction methods outlined above, it is entirely possible these compounds are simply statistical anomalies and it should be noted that the fold-change differences shown in the figure above are relatively weak.

### Predicting metabolite response classes from molecular formulae

[Need to update intro with ideas for this section, justify why we did it]

[I don't know how to explain why I'm using the y-axis coordinates here but it's basically  proxy for whether a metabolite's more abundant at the surface or at depth. High values correspond to three groups where 15m values are super high, and low (negative) values correspond to groups where a metabolite is mostly found at 175 meters] When we regressed the vertical axis of Figure XX above against the derived C:N ratio of each metabolite, we expected to find a positive trend with those metabolites most abundant at the surface having the highest C:N ratio based on the increased abundance of light energy and the decreased availability of nitrate at the surface relative to depth. However, we found no significant trend between C:N and the y-coordinate of the metabolites. This absence of a trend was robust when using only known metabolites or when using all metabolites for which a molecular formula could be calculated (see Methods).

```{r fig.height=3, fig.width=5}
formula2elements <- function(formula_vec){
  split_formulas <- formula_vec %>%
    gregexpr(pattern = "[A-Z][a-z]*[0-9]*") %>%
    regmatches(x = formula_vec)
  elements_in <- lapply(split_formulas, gsub, pattern = "[0-9]", replacement = "")
  element_counts <- lapply(split_formulas, gsub, pattern = "[A-z]", replacement = "")
  element_counts <- lapply(element_counts, function(x){
    x[!nchar(x)]<-1
    return(as.numeric(x))
  })
  mapply(`names<-`, element_counts, elements_in, SIMPLIFY = FALSE)
}

# Targeted compounds
targ_cn_ratios <- clockplot_df %>%
  select(feature, y, resp_class) %>%
  left_join(MT_feats, by="feature") %>%
  filter(compound_type!="Internal Standard") %>%
  mutate(elem_list=formula2elements(formula)) %>%
  unnest_wider(elem_list) %>%
  mutate(CN_ratio=C/N) %>%
  arrange(CN_ratio) %>%
  select(feature, y, CN_ratio, resp_class) %>%
  mutate(method_type="Targeted")

# Untargeted compounds
untarg_cn_ratios <- clockplot_df %>%
  select(feature, y, resp_class) %>%
  left_join(MT_feats, by="feature") %>%
  mutate(elem_list=formula2elements(calc_formula)) %>%
  unnest_wider(elem_list) %>%
  mutate(CN_ratio=C/N) %>%
  arrange(CN_ratio) %>%
  select(feature, y, CN_ratio, resp_class) %>%
  mutate(method_type="Untargeted")

all_cn_ratios <- targ_cn_ratios %>%
  bind_rows(untarg_cn_ratios) %>%
  filter(!is.na(CN_ratio))

lm_output <- all_cn_ratios %>%
  nest(data=-method_type) %>%
  mutate(lm_pval=map_dbl(data, ~broom::tidy(lm(.x$CN_ratio~.x$y))$p.value[2])) %>%
  mutate(lm_label=paste("Regression\np-value:", round(lm_pval, 3)))

all_cn_ratios %>%
  ggplot(aes(x=y, y=CN_ratio)) +
  geom_point(aes(color=resp_class)) +
  scale_color_manual(values = color_scale, breaks = names(color_scale), limits=force) +
  geom_smooth(method = "lm", formula = "y~x") +
  geom_label(aes(label=lm_label), data = lm_output, 
             x=Inf, y=Inf, hjust=1, vjust=1, label.size = 0) +
  facet_wrap(~method_type, nrow=1) +
  xlim(-0.67, 0.67) +
  ylab("C:N ratio") +
  xlab("y-axis coordinate from Figure XX above") +
  theme_bw() +
  theme(legend.position = "top", legend.title = element_blank()) +
  guides(color = guide_legend(nrow = 3))
```

However, while performing preliminary data exploration we did find that compounds that decrease monotonically with depth did have significantly higher C:N ratios than those that remain at comparable abundances between the surface and the DCM.

```{r, fig.width=5}
t_labels <- all_cn_ratios %>%
  filter(resp_class%in%c("15m > DCM > 175m", "175m < DCM = 15m")) %>%
  nest(data=-method_type) %>%
  mutate(t_pval=map_dbl(data, ~broom::tidy(t.test(.x$CN_ratio~.x$resp_class))$p.value)) %>%
  mutate(t_label=paste("t-test p-value:", round(t_pval, 5)))

all_cn_ratios %>%
  filter(resp_class%in%c("15m > DCM > 175m", "175m < DCM = 15m")) %>%
  ggplot() +
  geom_violin(aes(x=resp_class, y=CN_ratio, color=resp_class), lwd=0.8) +
  geom_boxplot(aes(x=resp_class, y=CN_ratio, color=resp_class), lwd=0.8, 
               width=0.2, outlier.shape = NA, fill="#FFFFFF33") +
  geom_jitter(aes(x=resp_class, y=CN_ratio), color="grey40", width = 0.1) +
  scale_color_manual(values = color_scale, breaks = names(color_scale)) +
  geom_label(aes(label=t_label), data = t_labels, 
           x=Inf, y=Inf, hjust=1, vjust=1, label.size = 0) +
  facet_wrap(~method_type, nrow=1) +
  ylab("C:N ratio") +
  xlab(NULL) +
  theme_bw() +
  theme(legend.position = "none")
```

[Optional additional discussion could occur based on the compounds with a DCM max continuing this trend, implying that if anything it's the compounds that are most abundant at the DCM that have the lowest C:N ratios, but then goes back up again at 175m unexpectedly, see temporary fig below]

```{r}
all_cn_ratios %>%
  ggplot() +
  geom_boxplot(aes(x=resp_class, y=CN_ratio, color=resp_class), lwd=1.2, varwidth = TRUE) +
  scale_color_manual(values = color_scale, breaks = names(color_scale)) +
  facet_wrap(~method_type, nrow=1) +
  ylab("C:N ratio") +
  xlab(NULL) +
  theme_bw() +
  theme(legend.position = "none") +
  coord_flip()
```

### Metabolite response to SLA

```{r metabolite sla response}
# For metabolites
trend_w_sla <- MT_peaks %>%
  left_join(filled_file_metadata %>% select(filename, depth, sla), by="filename") %>%
  group_by(feature) %>%
  mutate(value=rank(norm_area)) %>%
  ungroup() %>%
  nest(data=-c(feature, depth)) %>%
  mutate(lmtest=map(data, ~lm(.x$value~.x$sla))) %>%
  mutate(lmtidy=map(lmtest, broom::tidy)) %>%
  suppressWarnings() %>%
  unnest(lmtidy) %>%
  mutate(term=ifelse(term=="(Intercept)", "intercept", "slope")) %>%
  pivot_wider(names_from = term, values_from = estimate) %>%
  fill(intercept) %>%
  filter(!is.na(slope)) %>%
  select(feature, depth, data, slope, intercept, p.value) %>%
  mutate(p.value=p.adjust(p.value, method = "fdr")) %>%
  mutate(p_type=case_when(
    p.value<0.05 & slope>0 ~ "Positive trend",
    p.value<0.05 & slope<0 ~ "Negative trend",
    TRUE ~ "No trend"
  )) %>%
  select(feature, depth, p_type, slope, intercept) %>%
  mutate(p_type=factor(p_type, levels = c(
    "No trend", "Negative trend", "Positive trend")
  ))
alldepth_sla_response <- trend_w_sla %>%
  group_by(feature) %>%
  summarize(n_depths=sum(p_type!="No trend")) %>%
  count(n_depths) %>%
  pull(n, n_depths)
sla_trend_table <- trend_w_sla %>%
  count(depth, p_type)
sla_resp_types <- trend_w_sla %>%
  select(-slope, -intercept) %>%
  pivot_wider(names_from = depth, values_from = p_type) %>%
  mutate(sla_resp_type=paste(`15m`, `DCM`, `175m`)) %>%
  group_by(sla_resp_type) %>%
  # filter(`DCM`=="Negative trend" & `175m`=="Positive trend") %>% left_join(MT_feats)
  summarize(`15m`=unique(`15m`), `DCM`=unique(`DCM`), `175m`=unique(`175m`), n=n()) %>%
  select(-sla_resp_type) %>%
  arrange(desc(n)) %>%
  filter(n>1) %>%
  mutate(across(-n, ~case_when(
    .x=="No trend" ~ "---",
    .x=="Positive trend" ~ "Pos",
    .x=="Negative trend" ~ "Neg"
  )))

# Metabolite plot
metab_gp <- MT_peaks %>%
  group_by(feature) %>%
  mutate(value=rank(norm_area)) %>%
  ungroup() %>%
  left_join(filled_file_metadata %>% select(filename, depth, sla), by="filename") %>%
  right_join(trend_w_sla, by = c("feature", "depth")) %>%
  mutate(plotcol=ifelse(p_type!="No trend", feature, NA)) %>%
  ungroup() %>%
  ggplot(aes(x=sla, y=value, group=feature, color=plotcol, label=feature)) +
  geom_point() +
  geom_abline(aes(slope=slope, intercept=intercept, color=plotcol), lwd=0.8) +
  facet_grid(depth~p_type) +
  theme_bw() +
  theme(legend.position = "none") +
  ggtitle("A) Metabolites")
```

As expected from the multivariate PERMANOVA results detailed above, we find that relatively few compounds are sensitive to sea level anomalies. `r unname(alldepth_sla_response["0"])` metabolites were independent of SLA at all three depths, `r unname(alldepth_sla_response["1"])` had a significant trend at exactly one depth, `r unname(alldepth_sla_response["2"])` responded to SLA at two depths, and only `r unname(alldepth_sla_response["3"])` metabolite (MT_POS_FT0709, an unknown with *m/z* of 180.1019 and retention time ~2 minutes) had a trend with SLA at all depths, always negative. Notably, compounds at the DCM are much more likely to have a negative trend with SLA, while compounds at 175 meters are likely to have an opposite trend. The distribution of metabolites across these categories is shown below in Figure XX. This effect was rarely due to a given compound switching from a positive to negative trend (just `r sla_resp_types$n[4]` of the `r nrow(MT_feats)` metabolites switched), but was instead largely driven by compounds with no trend at the DCM switching to a positive trend at 175 meters (`r sla_resp_types$n[2]` / `r nrow(MT_feats)`) and compounds with no trend at other depths trending negatively with SLA at the DCM (`r sla_resp_types$n[3]` / `r nrow(MT_feats)`).

```{r metadata sla response}
# Metadata calcs and plot
metadata_sla_df <- filled_file_metadata %>%
  filter(cruise=="MT") %>%
  filter(samp_type=="Smp") %>%
  select(untripl, depth, sla, no23_um:last_col()) %>%
  pivot_longer(cols = no23_um:last_col()) %>%
  group_by(name) %>%
  mutate(value=rank(value)) %>%
  ungroup() %>%
  distinct()
trend_w_sla_meta <- metadata_sla_df %>%
  nest(data=-c(name, depth)) %>%
  mutate(lmtest=map(data, ~lm(.x$value~.x$sla))) %>%
  mutate(lmtidy=map(lmtest, broom::tidy)) %>%
  suppressWarnings() %>%
  unnest(lmtidy) %>%
  mutate(term=ifelse(term=="(Intercept)", "intercept", "slope")) %>%
  pivot_wider(names_from = term, values_from = estimate) %>%
  fill(intercept) %>%
  filter(!is.na(slope)) %>%
  select(name, depth, data, slope, intercept, p.value) %>%
  mutate(p.value=p.adjust(p.value, method = "fdr")) %>%
  mutate(p_type=case_when(
    p.value<0.05 & slope>0 ~ "Positive trend",
    p.value<0.05 & slope<0 ~ "Negative trend",
    TRUE ~ "No trend"
  )) %>%
  select(name, depth, p_type, slope, intercept) %>%
  mutate(p_type=factor(p_type, levels = c(
    "No trend", "Negative trend", "Positive trend")
  ))
meta_gp <- metadata_sla_df %>%
  right_join(trend_w_sla_meta, by = c("name", "depth")) %>%
  mutate(plotcol=ifelse(p_type!="No trend", name, NA)) %>%
  ungroup() %>%
  mutate(name=factor(name, levels = factor_levels, labels = factor_labels)) %>%
  ggplot(aes(x=sla, y=value, group=name, color=name, label=name)) +
  geom_vline(xintercept=0, color="grey75") +
  geom_point() +
  geom_abline(aes(slope=slope, intercept=intercept, color=name), lwd=0.8) +
  facet_grid(depth~p_type) +
  theme_bw() +
  theme(axis.title.y = element_blank(), axis.text.y = element_blank()) +
  ggtitle("B) Environmental parameters")

gridExtra::grid.arrange(metab_gp, meta_gp, layout_matrix=matrix(c(1,1,2,2,2), nrow = 1))
```

The response to SLA was weakest at the surface when measured by the absolute number of compounds in each category. At a depth of 15 meters, only `r sla_trend_table$n[2]` compounds had a significant negative trend with SLA (more abundant in cyclone than anticyclone) while `r sla_trend_table$n[3]` had a positive trend, together representing `r paste0(round(sum(sla_trend_table$n[2:3])/(nrow(trend_w_sla)/3), 3)*100, "%")` of the total compounds. No known compounds had negative SLA trends despite particulate N and picoeukaryote abundance falling into this category, but known compounds with positive SLA trends were 5-Oxoproline and N6-Acetyl-L-lysine as well as the environmental parameters DOP/DON, nitrate + nitrite, and phosphate.

At the DCM, the proportion of compounds that responded to SLA increased to `r paste0(round(sum(sla_trend_table$n[5:6])/(nrow(trend_w_sla)/3), 3)*100, "%")`, with `r sla_trend_table$n[5]` metabolites trending negatively and `r sla_trend_table$n[6]` in the opposite direction. Metabolites in this first category included the abundant osmolyte homarine and its [similar molecules] nicotinic acid and trigonelline. Carnitine and O-acetylcarnitine are also included as well as adenosine and its degradation product inosine, plus the other nucleoside cytidine. Other known molecules were the osmolyte DMS-Ac, the amino acids alanine and proline, the glycine derivative sarcosine, and the compound tentatively identified as threonine betaine (see above). The environmental parameters that were most strongly negative were particulate carbon and particulate nitrogen. Only one unidentified metabolite had the opposite trend of increased abundance in the anticyclone, as well as DON and DOP.

Finally, at 175 meters, nearly 10% of all measured compounds (`r paste0(round(sum(sla_trend_table$n[8:9])/(nrow(trend_w_sla)/3), 3)*100, "%")`) changed with SLA. `r sla_trend_table$n[8]` were less abundant in the cyclone than the anticyclone, and `r sla_trend_table$n[9]` increased in abundance across the eddy dipole. Compounds with a negative SLA trend consisted of two unknown molecules as well as N + N, and phosphate. Compounds with a positive trend were much more numerous and included several betaines (glycine betaine, proline betaine, betonicine/turicine, and the pseudo-betaine arsenobetaine) plus guanine, gonyol, dexpanthenol, and the amino acids leucine and tyrosine. Alanine, O-acetylcarnitine, and the putatively identified threonine betaine also fell into this class, making them among the few compounds that actually switched the sign of their trend between the DCM and 175 meters.

### Investigating the relative importance of sea level anomaly on a higher-resolution depth scale

```{r MC diff stats}
MC_feats <- all_feats %>%
  filter(cruise=="MC") %>%
  # filter(polarity=="pos") %>%
  select(-classes, -CSI_name) %>%
  rename("calc_formula"=formula) %>%
  left_join(clean_stans, by=c("compound_name", "polarity"))
MC_peaks <- MC_feats %>%
  left_join(all_peaks, by="feature") %>%
  left_join(filled_file_metadata, by="filename") %>%
  filter(samp_type=="Smp") %>%
  select(feature, norm_area, filename)

most_stat_diff <- MC_peaks %>%
  left_join(filled_file_metadata, by="filename") %>%
  select(feature, filename, station, depth, norm_area) %>% 
  nest(data=-feature) %>%
  ### Classic t-test
  # mutate(test=map(data, ~broom::tidy(t.test(.x$norm_area~.x$station)))) %>%
  # unnest(test) %>%
  ### Non-parametric rank-sum test
  # mutate(test=map(data, ~broom::tidy(wilcox.test(.x$norm_area~.x$station)))) %>%
  # unnest(test) %>%
  # rename(estimate=statistic) %>%
  ### Permutational t-test
  mutate(test=map(data, ~broom::tidy(aovp(.x$norm_area~.x$station, seqs = FALSE,
                                          settings=FALSE)))) %>%
  unnest(test) %>%
  rename(p.value=`Pr(Prob)`) %>%
  mutate(estimate=1) %>%
  filter(term==".x$station1") %>%
  ### No matter what test, run the below
  mutate(pval=p.adjust(p.value, method = "fdr")) %>%
  arrange(pval) %>%
  select(feature, data, estimate, pval) %>%
  mutate(pval_cats = case_when(
    pval < 0.001 ~ "Highly signif",
    pval < 0.05 ~ "Signif",
    pval >=0.05 ~ "Nonsig"
  )) %>%
  mutate(fold_change=map_dbl(data, ~mean(.x$norm_area[.x$station=="L1"])/
                             mean(.x$norm_area[.x$station=="L2"])))

pval_cat_counts <- most_stat_diff %>%
  pull(pval_cats) %>% table()

# most_stat_diff %>%
#   left_join(all_feats) %>%
#   filter(!is.na(compound_name)) %>%
#   select(pval_cats, compound_name, fold_change) %>%
#   arrange(desc(fold_change)) %>%
#   group_by(pval_cats) %>%
#   group_split()
```

We found `r sum(pval_cat_counts)` features in this new set of data using the untargeted methods described above. Of those, `r unname(pval_cat_counts["Highly signif"])` were significantly different between the two eddies with p-values less than 0.001 and `r unname(pval_cat_counts["Signif"])` were significantly different with p-values between 0.001 and 0.05 after controlling the false discovery rate. All known compounds found to be significantly different were elevated in the cyclonic samples, but several unknown molecules had the opposite trend and were enriched in the anticyclone.

```{r MC targ concentrations, eval=FALSE}
highly_sig_feats <- most_stat_diff %>%
  left_join(all_feats, by="feature") %>%
  filter(!is.na(compound_name)) %>%
  filter(pval_cats=="Highly signif") %>%
  pull(compound_name)
MC_targ <- "../targeted/made_data/longdata.csv" %>%
  read_csv(show_col_types = FALSE) %>%
  filter(compound_name%in%highly_sig_feats) %>%
  mutate(filename=paste0(filename, ".mzML")) %>%
  left_join(filled_file_metadata, by="filename") %>%
  filter(cruise=="MC") %>%
  filter(samp_type=="Smp")
MC_targ %>%
  group_by(compound_name, station) %>%
  summarize(mean_val=mean(nM)) %>%
  pivot_wider(names_from = station, values_from = mean_val) %>%
  mutate(fold_change=L1/L2) %>%
  arrange(desc(fold_change))
# MC_targ %>%
#   filter(compound_name=="Isethionic acid") %>%
#   group_by(station) %>%
#   summarise(max_nM=max(nM), min_nM=min(nM))
```

Eleven known compounds were highly sensitive (permutational t-test p-value < 0.001 after false discovery rate correction) to sampling location. Isethionic acid increased from 0.03 nM in the anticyclone to 0.12 nM in the cyclone, a 4x fold change and the largest difference among the known compounds. This molecule separated completely between the eddies, with values less than 0.08 nM found exclusively in the anticyclonic feature and all larger values exclusively in the cyclone.

Compounds that were significantly enrighed at the DCM in the cyclone relative to the anticyclone at a less-dramatic p-value cutoff of 0.05 included several amino acids (glycine as noted above, alanine, asparagine, and serine) as well as amino acid derivatives (sarcosine, choline, homoserine, theanine, and 4-aminobutyric acid). Homarine and its demethylated form nicotinic acid were also distinctive between the eddies, as well as a surprising number of sulfur-containing compounds (the aforementioned isethionic acid, plus DMSP, DHPS, and N-methyltaurine.) Other compounds included the betaines proline betaine, carnitine, and the putative threonine betaine, plus AMP, cytidine, and 2-O-alpha-D-Glucosylglycerol.

```{r MC volcano plot}
most_stat_diff %>%
  left_join(MC_feats, by = "feature") %>%
  filter(!is.na(compound_name)) %>%
  slice(1:20) %>%
  unnest(data) %>%
  mutate(eddy_type=ifelse(station=="L1", "Cyclonic", "Anticyclonic")) %>%
  ggplot() +
  geom_boxplot(aes(x=norm_area, y=depth, color=eddy_type)) +
  facet_wrap(~compound_name, scales = "free_x") +
  scale_color_manual(breaks = c("Cyclonic", "Anticyclonic"), values = c(muted("blue"), muted("red")))

MT_type <- trend_w_sla %>%
  filter(depth=="DCM") %>%
  left_join(all_feats, by="feature") %>%
  select(feature, p_type, group_feat) %>%
  mutate(p_type=as.character(p_type))

most_stat_diff %>%
  mutate(pval=ifelse(pval==0, 0.001, pval)) %>%
  left_join(MC_feats, by = "feature") %>%
  mutate(which_stat_bigger=ifelse(estimate>0, "L1", "L2")) %>%
  mutate(fold_change=log2(fold_change)) %>%
  mutate(cmpd_known=!is.na(compound_name)) %>%
  mutate(compound_name=ifelse(is.na(compound_name), "", compound_name)) %>%
  left_join(MT_type, by="group_feat") %>%
  mutate(p_type=ifelse(is.na(p_type), "Not found in MT", p_type)) %>%
  ggplot(aes(x=fold_change, y=-log10(pval), label=compound_name)) +
  geom_vline(xintercept = 0) +
  geom_hline(yintercept = 0) +
  geom_hline(yintercept = -log10(0.05), lty=2) +
  geom_point(aes(color=p_type)) +
  geom_text_repel(max.overlaps = 10, min.segment.length = 0) +
  annotate("label", x=Inf, y=Inf, label="Enriched in cyclone", 
           hjust=1, vjust=1, label.r=unit(0, "in")) +
  annotate("label", x=-Inf, y=Inf, label="Enriched in anticyclone", 
           hjust=0, vjust=1, label.r=unit(0, "in")) +
  theme_bw() +
  xlim(-4.5, 4.5) +
  scale_color_manual(breaks=c("Negative trend", "No trend", "Not found in MT", "Positive trend"),
                     values = c(muted("red"), "grey50", "grey80", muted("blue")), name="Trend in eddy\ntransect")
```

*Figure XX: Volcano plot of metabolites, with the log2 fold-change difference between the cyclone and anticyclone shown along the x-axis and the log10-scaled p-value (Student's t-test followed by FDR correction) on the y-axis. Compounds that have the largest biological differences are found on the far left and far right sides of the graph while compounds with the largest statistical differences are found at the top of the plot. The dotted line denotes an alpha level of 0.05.*

```{r check different unknowns, eval=FALSE}
feature_msdata <- most_stat_diff %>%
  left_join(MC_feats, by="feature") %>%
  filter(is.na(compound_name)) %>%
  select(feature, mzmed, rtmed, calc_formula) %>%
  # filter(feature=="MC_POS_FT410")
  slice(1)

# msms_files <- list.files("../mzMLs/pos/MSMS", pattern = "180205.*dda", full.names = TRUE)
# msmsdata <- grabMSdata(msms_files)

mz_i <- feature_msdata$mzmed
ggplot() +
  geom_line(aes(x=rt, y=int, group=filename),
            data = msmsdata$MS1[mz%between%pmppm(mz_i)]) +
  geom_vline(aes(xintercept=rt), color="steelblue",
             data = msmsdata$MS2[premz%between%pmppm(mz_i)] %>% distinct(rt)) +
  geom_vline(xintercept = feature_msdata$rtmed/60+c(-0.5, 0.5), color="red")
msmsdata$MS2[premz%between%pmppm(mz_i)] %>%
  filter(!fragmz%between%pmppm(mz_i, 10)) %>%
  filter(rt%between%(feature_msdata$rtmed/60+c(-0.5, 0.5))) %>%
  ggplot(aes(x=fragmz, y=int)) +
  geom_segment(aes(xend=fragmz, yend=0)) +
  geom_point()

mgf_maker <- function(feature_msdata, ms1, ms2, output_file){
  ms1_to_write <- paste(
    "BEGIN IONS", paste0("PEPMASS=", feature_msdata$mzmed),
    "MSLEVEL=1", "CHARGE=1+", 
    paste0(apply(ms1, 1, paste, collapse=" "), collapse = "\n"),
    "END IONS", "", 
    sep = "\n")
  ms2_to_write <- lapply(unique(ms2$filename), function(volt_i){
    paste0(c("BEGIN IONS", paste0("PEPMASS=", feature_msdata$mzmed),
           "MSLEVEL=2", "CHARGE=1+",
           paste0(apply(ms2[ms2$filename==volt_i, c("fragmz", "int")], 
                 1, paste, collapse=" "), collapse="\n"),
           "END IONS",
           ""), collapse="\n")
  })
  outtext <- paste0(ms1_to_write, "\n", paste0(unlist(ms2_to_write), collapse = "\n"))
  writeLines(outtext, con = output_file)
}

ms1 <- matrix(c(feature_msdata$mzmed, feature_msdata$avg_area), ncol = 2)
ms2 <- msmsdata$MS2[premz%between%pmppm(mz_i)] %>%
  filter(rt%between%c(7, 8))
mgf_maker(feature_msdata, ms1, ms2, output_file = "~/../Desktop/temp.mgf")
```

We also did some preliminary searches for the identity of the low p-value compounds that were enriched in the anticyclone using CSI:FingerID software. MC_POS_FT410 (*m/z* 179.0047, retention time 7.3 minutes) had a fold-change enrichment of 4x in the anticyclonic eddy and we annotated it as arsenobetaine after purchasing the standard which was run at a later time and matched in MS^1^ and MS^2^ *m/z* values at a confidence level very close to 1. MC_POS_FT328 (*m/z* 160.0968, retention time 7.8 minutes, putative formula $C_7H_{14}NO_3$), another compound that separated completely between the two eddies and had a p-value of approximately 10^-7^ shares an exact mass with betonicine/turicine but is not likely to be this compound, as we later purchased this standard and its elution time corresponds to a separate peak in the same chromatogram. Instead, CSI:FingerID and CANOPUS provided top hits of acetylvaline and 5-acetamidopentanoate with a 90% posterior probability of containing an N-acyl amino acid group fora confidence of approximately 3. MC_POS_FT400 (*m/z* 176.1281, retention time 8.1 minutes, putative formula $C_8H_{17}NO_3$, confidence level 3) appears to be some unidentified betaine, with a 97+% chance of belonging to the quaternary ammonium salts class, and may in fact represent a methylated version of the MC_POS_FT328 described just above.

### How robust are these results?

We tested the robustness of our results by returning to the same approximate location one year later and again sampled two eddies of opposite polarities. Each eddy was sampled at 25 meters and the deep chlorophyll maximum (~105 meters in the cyclonic eddy, ~122 meters in the anticyclone) and once at 6 AM local time and again at 6 PM local time. This time, only nutrient data (nitrate + nitrite and soluble reactive phosphorus), particulate data (PC and PN), and flow cytometry data was collected.

```{r metadata boxplots}
filled_FK_metadata <- filled_file_metadata %>%
  filter(cruise=="FK") %>%
  filter(samp_type=="Smp")

distinct_FK_metadata <- filled_FK_metadata %>%
  mutate(basename=str_remove(untripl, "190715_Smp_")) %>%
  mutate(cyc_type=ifelse(station%in%c("62", "64"), "cyc", "anti")) %>%
  select(basename, cyc_type, depth, sla, ends_with("abund"), ends_with("um")) %>%
  select(-oxy_um, -si_um, -DON_um, -DOP_um) %>%
  distinct()

# If we decide to use a figure (boxplots of values by depth)
distinct_FK_metadata %>%
  mutate(depth=factor(depth, levels = c("DCM", "25m"))) %>%
  select(basename, cyc_type, depth, sla, ends_with("abund"), ends_with("um")) %>%
  pivot_longer(-c(basename, cyc_type, depth, sla)) %>%
  mutate(name=factor(name, levels = factor_levels, labels = twoline_labels)) %>%
  filter(!is.na(value)) %>%
  ggplot() +
  # geom_boxplot(aes(x=value, y=depth)) +
  geom_jitter(aes(x=value, y=depth, color=cyc_type), height = 0.1, width = 0) +
  facet_wrap(~name, scales="free_x", nrow=2) +
  theme(axis.title = element_blank(), legend.position = "none")
```

```{r metadata table}
# If we decide to use a table
distinct_FK_metadata %>%
  select(basename, cyc_type, depth, sla, ends_with("abund"), ends_with("um")) %>%
  pivot_longer(-c(basename, cyc_type, depth, sla)) %>%
  group_by(depth, name) %>%
  summarize(print_col=paste(
    signif(mean(value, na.rm=TRUE), digits = 3), "±", 
    signif(sd(value, na.rm=TRUE), digits = 1)
  ), .groups = "drop") %>%
  # Pivot into nice format
  pivot_wider(names_from="depth", values_from = print_col) %>%
  mutate(name=factor(name, levels = factor_levels, labels = factor_labels)) %>%
  arrange(name)
```

```{r welch ANOVA FK}
welch_aov <- distinct_FK_metadata %>%
  select(basename, depth, sla, ends_with("abund"),
         ends_with("um"), ends_with("ug")) %>%
  pivot_longer(-c(basename, depth, sla)) %>%
  # Perform the ANOVA
  nest(data=-name) %>%
  mutate(aovtest=map(data, ~broom::tidy(oneway.test(.x$value~.x$depth)))) %>%
  suppressMessages() %>%
  unnest(aovtest) %>%
  # Format p-values nicely
  mutate(p.value=p.adjust(p.value, method = "holm"))
```

The environmental data show many similar trends to the previous year and are well within the ranges documented elsewhere. Nitrate + nitrite and phosphate are both lower in the anticyclonic eddy than the cyclonic one, and PC and PN are both generally lower as well with a much larger difference at the DCM. Notably, heterotrophic bacteria and *Synechococcus* were consistently more abundant in the anticyclonic eddy.

```{r FK v MT v MC distance correlation}
nested_peaks <- all_feats %>%
  group_by(group_feat) %>%
  filter(n()==3) %>%
  ungroup(group_feat) %>%
  left_join(all_peaks, by="feature") %>%
  left_join(filled_file_metadata, by = c("cruise", "filename")) %>%
  filter(samp_type=="Smp") %>%
  group_by(cruise, feature) %>%
  mutate(norm_area=norm_area/max(norm_area)) %>%
  ungroup()

fk_dist <- nested_peaks %>%
  filter(cruise=="FK") %>%
  select(feature, filename, norm_area) %>%
  pivot_wider(names_from = filename, values_from = norm_area) %>%
  column_to_rownames("feature") %>%
  data.matrix() %>%
  dist()
mt_dist <- nested_peaks %>%
  filter(cruise=="MT") %>%
  select(feature, filename, norm_area) %>%
  pivot_wider(names_from = filename, values_from = norm_area) %>%
  column_to_rownames("feature") %>%
  data.matrix() %>%
  dist()
mc_dist <- nested_peaks %>%
  filter(cruise=="MC") %>%
  select(feature, filename, norm_area) %>%
  pivot_wider(names_from = filename, values_from = norm_area) %>%
  column_to_rownames("feature") %>%
  data.matrix() %>%
  dist()

dist_mat <- cbind(as.numeric(fk_dist), as.numeric(mc_dist), as.numeric(mt_dist))
colnames(dist_mat) <- c("FK", "MC", "MT")
cor_mat <- round(cor(dist_mat), 3)
```

When we calculated the similarity of the three separate metabolomes, we obtained correlation coefficients of `r cor_mat["MC", "MT"]` for the eddy center and eddy transect data, `r cor_mat["MC", "FK"]` for the eddy center and Falkor data, and `r cor_mat["MT", "FK"]` for the transect and Falkor data.

```{r SLA p-value crosscruise comparison, eval=FALSE}
FK_sla_diff <- all_peaks %>%
  left_join(all_feats, by = "feature") %>%
  left_join(filled_file_metadata, by = c("filename", "cruise")) %>%
  mutate(eddy=ifelse(station%in%c("62", "64"), "cyc", "anti")) %>%
  filter(samp_type=="Smp") %>%
  filter(cruise=="FK") %>%
  nest(data=-c(group_feat, feature)) %>%
  # mutate(pval=map_dbl(data, ~broom::tidy(t.test(.x$norm_area~.x$eddy))$p.value)) %>%
  mutate(pval=map_dbl(data, ~broom::tidy(lmp(.x$norm_area~.x$eddy, seqs=FALSE, settings=FALSE))$statistic[2])) %>%
  select(group_feat, feature, pval) %>%
  mutate(pval=p.adjust(pval, method = "fdr"))

MT_sla_diff <- all_peaks %>%
  left_join(all_feats, by = "feature") %>%
  left_join(filled_file_metadata, by = c("filename", "cruise")) %>%
  filter(samp_type=="Smp") %>%
  filter(cruise=="MT") %>%
  nest(data=-c(group_feat, feature)) %>%
  mutate(pval=map_dbl(data, ~broom::tidy(lmp(.x$norm_area~.x$sla, seqs=FALSE, settings=FALSE))$statistic[2])) %>%
  select(group_feat, feature, pval) %>%
  mutate(pval=p.adjust(pval, method = "fdr")) %>%
  filter(pval<0.05)

MC_sla_diff <- all_peaks %>%
  left_join(all_feats, by = "feature") %>%
  left_join(filled_file_metadata, by = c("filename", "cruise")) %>%
  filter(samp_type=="Smp") %>%
  filter(cruise=="MC") %>%
  nest(data=-c(group_feat, feature)) %>%
  mutate(pval=map_dbl(data, ~broom::tidy(lmp(.x$norm_area~.x$station, seqs=FALSE, settings=FALSE))$statistic[2])) %>%
  select(group_feat, feature, pval) %>%
  mutate(pval=p.adjust(pval, method = "fdr")) %>%
  filter(pval<0.05)

all_feats %>%
  filter(group_feat%in%MT_sla_diff$group_feat) %>%
  filter(group_feat%in%MC_sla_diff$group_feat) %>%
  filter(group_feat%in%(FK_sla_diff %>% filter(pval<0.05) %>% pull(group_feat))) %>%
  select(group_feat, feature, mzmed, rtmed, compound_name, formula) %>%
  arrange(mzmed)

all_feats %>%
  filter(group_feat%in%MT_sla_diff$group_feat) %>%
  filter(group_feat%in%MC_sla_diff$group_feat) %>%
  left_join(FK_sla_diff %>% select(group_feat, pval), by="group_feat") %>%
  select(group_feat, feature, mzmed, rtmed, compound_name, formula, pval) %>%
  arrange(mzmed)
```

Although we did not find any compounds that differed significantly across all three data sets, we found several shared compounds between the eddy transect and eddy center samples. Sarcosine, inosine, nicotinic acid, and isethionic acid were all more abundant in the cyclone than the anticyclone, as well as seven unknown compounds. Of these, 4 were unfound in the Falkor data set, while the other 7 were simply not different (p-values ranging from 0.13 - 0.62).

## Discussion

### Environmental data

The results obtained in this study from the environmental data are generally in good agreement with those previously described (e.g. @Barone2022, @HenderikxFreitas2020, and @harke2021) despite the different statistical methodology and the use here of interpolated values instead of raw data.

### Overall metabolite response

It is relatively unsurprising that we're able to separate the samples by depth given the results from the environmental data. Plotting the metabolome along NMDS axes 1 and 3 creates clear visual clusters corresponding to each depth (Fig. XX). NMDS axis 2, however, represents an unknown source of variation. Many of the triplicate triangles separate along this axis and often span zero. Due to the variance between triplicates here, it is unlikely that this variation is biological or environmental and instead may be accounting for differences in sample processing, extraction, or injection despite most compounds being normalized to a best-matched internal standard.

The SLA-driven separation within the 175 meter samples is also quite informative. The deep samples from stations with a larger (anticyclonic) SLA all fall closer to the shallow samples along NMDS axis 1 than those from stations with a negative (cyclonic) SLA. This result makes sense given the context of anticyclones pushing surface water downward upon formation, deepening pycnoclines and causing the metabolome at 175m to look more like a metabolome sampled at a shallower depth. This trend is not visually present in either the DCM or 15m samples, however, indicating that the effect of SLA may actually be strongest at 175 meters overall. There was no clear separation by time of day in the NMDS plots that we could detect visually which was a little surprising given prior research showing the complex diel cycling of metabolites over the course of the day.

The PERMANOVA results validate this subjective assessment with a very large R^2^ value (~30% variance explained) for depth, while time of day and eddy polarity had higher p-values (still significant at an alpha-level of 0.05) and lower R^2^ values, in total accounting for approximately 40% of the variance in the metabolome. A significant three-way interaction between depth, time, and SLA complicates this interpretation slightly, with each parameter potentially depending on the interaction of the other two, but with a low R^2^ and marginal p-value the interaction can probably be ignored for an exploratory analysis like this. The remaining ~60% of the variance in the data could be explained by true biological variation in the original seawater, differential treatment during the sample processing steps, an unexplored confounding variable, or most likely some combination of the above.

### Metabolite response to depth

The distinctly nonhomogenous distribution of the metabolome in Figure XX emphasizes the diversity of metabolite responses to depth. Rather than displaying a single, universal trend, compounds vary wildly in their abundance throughout the water column. Many appear to track biomass, correlating well with PC, PN and particle abundance with maximum concentrations somewhere between the surface and the DCM. Others have more complicated responses, whether found exclusively at the surface or *increasing* in concentration with depth, that illustrate the nuance missed by bulk particle measurements. Our conservative use of rank-order statistics here intentionally sacrifices some experimental power but guarantees that the signal observed raw peak-area space or z-scoring space would be at least as strong as the separations observed here. Compounds found mostly in the surface layer are synthesized in excess of biomass in the surface ocean [mixed layer], where light is highly abundant but nutrients are scarce. The presence of MAAs in this category provides one possible reason for excess synthesis in the surface ocean, as these compounds are known to be synthesized under UV light and act as "sunscreen" compounds in the environment [ref]. At the deep chlorophyll maximum, we found several compounds again synthesized in excess of biomass - most obviously chlorophyll, but also several nitrogen-rich compounds like amino acids, osmolytes, and the nucleobase guanine. Despite this apparent trend, our followup analysis revealed no strong correlation between depth category and C:N content. [Anitra, can you talk about or add citations about a lot of these compounds seeming to come from euks? Would be a nice connection back to the metadata where picoeuks are found almost exclusively at the DCM]. The absence of compounds that have a minimum at the DCM is relatively unsurprising. To have a DCM minimum requires either extensive drawdown at the DCM (which might occur for precursor molecules of DCM maximum compounds) or synthesis at the surface and depth which does not occur at the DCM itself (perhaps some general stress compound). Finally, compounds (both known and unknown) that did not respond to depth may do so for several reasons. First, it is possible but unlikely that a compound is being synthesized at a rate exactly inversely proportional to the decrease in biomass with depth - this would cause the observed concentration to be fixed across depths. Similarly, highly recalcitrant compounds could degrade at a rate too slow to be detected over the ~150 meters that we sampled across. Alternatively, these points could represent compounds that are highly sensitive to the sampling, extraction, or measurement process - because this classification scheme relies on the variance within each depth relative to the variance between depths, some highly sensitive compounds may have much higher noise levels and the depth signal is simply not visible. Our decision to use the nonparametric statistics here even when some data may be normally distributed exacerbates this problem, protecting strongly against false positives in the case of non-normality but undoubtedly causing some truly depth-sensitive compounds to be observed as uniform in abundance with depth. Finally, some of these compounds could be contaminants introduced during the sampling and measurement process but not all, because we know the identities of several molecules that belong to this category and do not expect them to be contaminants introduced by tubing, filters, or solvents. Most likely is that this null hypothesis category consists of a combination of the above, but we were not able to distinguish those potential explanations here.

### Metabolite response to SLA, both alone and compared to depth

Many of the trends with SLA observed here can be explained as a function of biomass. At the DCM, biomass is high in the cyclone as nutrient-rich water is upwelled into the euphotic zone and lower in the anticyclone. Conversely, in the 175 meter samples biomass is depleted from the cyclone and instead enriched at the DCM (? ask benedetto to explain this again). As expected, the response to SLA was weakest within the mixed layer at the surface where processes other than eddy upwelling/downwelling dominate. Despite this, we find slightly higher nitrate and nitrite concentrations in the anticyclone, as well as DON and DOP, so it's not completely unsurprising that some metabolites follow this trend. The increasing effect of SLA on metabolites and environmental variables is equally unsurprising, as density surfaces are displaced more at depth, though it's unclear how deep this effect is likely to propagate and future sampling at depths below 175 meters could improve our understanding of the depth at which eddies cease to affect the metabolome because that cannot be concluded from this study. Given that all of these samples were from the particulate fraction, it's also unsurprising that most metabolites track with PC/PN at the DCM and 175 meters. Conversely, those few metabolites that are either independent of PC/PN or negatively correlated with these variables deserve future study to determine whether they are real biological signal and, if so, what the biological functions they correspond to.

We used a very conservative statistical method to compare the variance introduced by ~40 centimeters of SLA to the variance introduced by sampling across 50 meters of the water column and found a surprising number of compounds for which the difference between the two eddies was larger than the difference with depth. For the `r sum(pval_cat_counts["Highly signif"], pval_cat_counts["Signif"])` compounds that were statistically different despite not stratifying by depth in any way, it is clear that SLA is at least as important as sampling depth around the DCM and the care that is currently taken to sample at exact depths using the CTD should be the same care taken to account for differences in SLA. We expected from the earlier analysis on the eddy transect to find that most compounds were enriched in the cyclone, but when we used the two endpoints established at the center of each eddy we found significantly more compounds enriched in the anticyclone. This may be due to the increased power of the t-tests (compared to the ranked regressions run above) or due to the increased difference expected from sampling the endpoints alone.

### Replicating/robustness of results

We expected to find a strong correlation between the metabolome measured during the eddy transect and the eddy centers, as these two data sets sampled the same water mass, albeit separated by a few days. We expected a lower correlation coefficient for comparisons to the Falkor data sets, as that data was collected a full year after the original eddy center and eddy transect. Surprisingly, we learned that the two data sets that sampled the centers of an eddy (the eddy center and Falkor data sets). The eddy transect metabolome was actually the odd one out, with correlation coefficients around 0.45 compared to the (quite a bit higher) 0.55 obtained when comparing the two sets of eddy center samples. This could imply that the areas outside of an eddy contain a large number of relatively unique compounds, or could be due to differences like time of day that was controlled for the eddy center and Falkor samples but not for the eddy transect.

The lack of any specific metabolite differing with SLA across all three data sets is more troubling. We would expect to find several such metabolites given the low p-values presented individually in each data set, but even at higher alpha values there was very little correlation between the p-value observed and its likelihood of differing with respect to SLA in the Falkor data. This may be due to the smaller sample size of the Falkor data set - with only 24 samples, the analysis is undoubtedly underpowered to detect the small differences observed in the eddy transect data with a sample size of 99. More powerful statistics such as multilevel modeling may also be able to better manage within-sample variation but also often require strong priors for the mechanisms by which compounds interact with each other and the environment or priors that are simply unavailable for these understudied molecules. That some compounds reproduced between the transect and eddy center data is heartening, but we are at this point unable to claim that we have discovered a particular biomarker for eddy state.

## Conclusions

TBD

### 2022-10-05 metadata/metadata_control.Rmd
Now tracking metadata

---
title: "Metadata control"
author: "William Kumler"
date: "5/5/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

library(tidyverse)
library(cmap4r)
cmap_api <- "6acd63a0-4c63-11ea-aa09-71d9d763e28d"
set_authorization(cmap_key = cmap_api)
library(fuzzyjoin)
library(ggmap)
register_google("AIzaSyBVNJV0Myoy3M50IBBDIjXZhUyU8Mnagvs")
library(ggrepel)
library(gridExtra)  # For grid.arrange
library(data.table) # For fast indexing of seaflow data
library(RaMS)       # For MS file timestamps
library(scales)     # For "muted" function
library(readxl)     # For "readxl"

scope_uname <- "scope"
scope_pw <- "SCOPE2014"

if(!dir.exists("raw_data")){
  dir.create("raw_data")
}
if(!dir.exists("made_data")){
  dir.create("made_data")
}
if(!dir.exists("figures")){
  dir.create("figures")
}

options(pillar.sigfig = 7)
```

```{r functions}
toDecDeg <- function(deg, min, dir){
  num <- deg+min/60
  if(dir%in%c("S", "W")){
    num*-1
  } else {
    num
  }
}
```

## CTD data (cruise, station, time, lat, lon, cast, SLA, DCM depth)

Cruise is manually entered, station/time/lat/lon/cast come from the SCOPE website cruise
summaries, SLA is pulled down from CMAP, and DCM depth was manually found by looking
at the cruise logbook.

Also now includes data from .gof files for exact depth

```{r station_metadata}
# DCMs are entered manually ----
if(!file.exists("raw_data/all_DCMs.csv")){
  falkor_DCMs <- data.frame(
    station=c(62, 64, 77, 80),
    cast=1,
    DCM_depth=c(98, 115, 125, 120)
  )
  transect_DCMs <- data.frame(
    station = c(4:13, 13, 14),
    cast=c(1,1,3,1,1,2,2,2,1,2,3,2),
    DCM_depth=c(100, 122, 119, 133, 122, 110, 110, 110, 103, 112, 112, 105)
  )
  center_DCMs <- data.frame(
    station = paste0("L", c(1,1,2,2)),
    cast=c(32, 33, 1, 2),
    DCM_depth=c(108, 104, 123, 123)
  )
  all_DCMs <- rbind(falkor_DCMs, transect_DCMs, center_DCMs)

  write.csv(all_DCMs, "raw_data/all_DCMs.csv", row.names = FALSE)
}

# SCOPE CTD data consists of station, cast, time, max_depth, bottles, lat, and lon ----
getScopeCTD <- function(url){
  col_names <- c("station", "cast", "month", "day", "year", "time", 
               "lat_deg", "lat_min", "lat_dir", 
                "lon_deg", "lon_min", "lon_dir", "max_depth", "bottles")
  read_table(url, skip = 3, col_names = col_names, col_types = cols(
    station = col_character(),
    cast = col_double(),
    month = col_character(),
    day = col_character(),
    year = col_double(),
    time = col_time(format = ""),
    lat_deg = col_double(),
    lat_min = col_double(),
    lat_dir = col_character(),
    lon_deg = col_double(),
    lon_min = col_double(),
    lon_dir = col_character(),
    max_depth = col_double(),
    bottles = col_double())) %>%
    mutate(time=strptime(paste(month, day, year, time), format = "%b %d %Y %X")) %>%
    select(-month, -day, -year) %>%
    mutate(lat=mapply(toDecDeg, lat_deg, lat_min, lat_dir)) %>%
    mutate(lon=mapply(toDecDeg, lon_deg, lon_min, lon_dir)) %>%
    select(-matches("^lon_|^lat_"))
}

if(!file.exists("raw_data/falkor_SCOPEweb_ctd_summary.csv")){
  # Data source: scope.soest.hawaii.edu/collaborators/datainventory/Data/SCOPEcore/FK_CTDsummary_current.txt
  falkor_SCOPEweb_ctd_summary_url <- paste0("http://", scope_uname, ":", scope_pw,
                "@scope.soest.hawaii.edu/collaborators/datainventory/Data/",
                "SCOPEcore/FK_CTDsummary_current.txt")
  falkor_SCOPEweb_ctd_summary <- getScopeCTD(falkor_SCOPEweb_ctd_summary_url)
  falkor_SCOPEweb_ctd_summary <- filter(falkor_SCOPEweb_ctd_summary, time > "2018-03-23 04:01:51")
  write.csv(falkor_SCOPEweb_ctd_summary, 
            file = "raw_data/falkor_SCOPEweb_ctd_summary.csv", 
            row.names = FALSE)
}
if(!file.exists("raw_data/meso_SCOPEweb_ctd_summary.csv")){
  # Data url source: scope.soest.hawaii.edu/collaborators/datainventory/Data/SCOPEcore/MS_CTDsummary_current.txt
  meso_SCOPEweb_ctd_summary_url <- paste0("http://", scope_uname, ":", scope_pw,
                                 "@scope.soest.hawaii.edu/collaborators/datainventory/Data",
                                 "/SCOPEcore/MS_CTDsummary_current.txt")
  meso_SCOPEweb_ctd_summary <- getScopeCTD(meso_SCOPEweb_ctd_summary_url)
  write.csv(meso_SCOPEweb_ctd_summary, 
            file = "raw_data/meso_SCOPEweb_ctd_summary.csv", 
            row.names = FALSE)
}

# SLA data is pulled down from CMAP for Falkor, FTP for meso ----
# SLA units are meters
if(!file.exists("raw_data/falkor_SLA.csv")){
  falkor_cruisename = "SCOPE_Falkor2"
  falkor_cruise_data <- get_cruise_by_name(falkor_cruisename)
  falkor_CMAP_SLA <- get_spacetime(tableName = 'tblAltimetry_REP',
                     varName = 'sla',
                     dt1=as.character(falkor_cruise_data$Start_Time),
                     dt2=as.character(falkor_cruise_data$End_Time),
                     lat1=falkor_cruise_data$Lat_Min,
                     lat2=falkor_cruise_data$Lat_Max,
                     lon1=falkor_cruise_data$Lon_Min,
                     lon2=falkor_cruise_data$Lon_Max)

  falkor_ctddata <- read.csv("raw_data/falkor_SCOPEweb_ctd_summary.csv")
  falkor_SLA <- falkor_ctddata %>%
    mutate(time=as.POSIXct(time)) %>%
    group_by(station) %>%
    summarize(lat=median(lat), lon=median(lon), time=median(time)) %>%
    difference_left_join(falkor_CMAP_SLA, max_dist = 0.125, by=c("lat", "lon")) %>%
    filter(as.Date(time.x)==as.Date(time.y)) %>%
    select(station, sla)

  write.csv(falkor_SLA, file="raw_data/falkor_SLA.csv", row.names = FALSE)
}
if(!file.exists("raw_data/meso_SLA.csv")){
  meso_SLA <- "http://scope.soest.hawaii.edu/data/mesoscope/km1709_CtdSla.txt" %>%
    read.table(skip = 1, col.names = c("station", "cast", "sla")) %>%
    mutate(sla=sla/100)
  write.csv(meso_SLA, file = "raw_data/meso_SLA.csv", row.names = FALSE)
}

# Combine and write out ----
falkor_SLA <- read.csv("raw_data/falkor_SLA.csv") %>%
  mutate(station=as.character(station)) %>%
  mutate(cast=1)
meso_SLA <- read.csv("raw_data/meso_SLA.csv")
dcm_depths <- read.csv("raw_data/all_DCMs.csv")
falkor_ctddata <- read.csv("raw_data/falkor_SCOPEweb_ctd_summary.csv")
meso_ctddata <- read.csv("raw_data/meso_SCOPEweb_ctd_summary.csv")

station_metadata <- meso_ctddata %>%
  mutate(cruise=ifelse(startsWith(station, "L")|station==15, "MC", "MT")) %>%
  rbind(falkor_ctddata %>% mutate(cruise="FK")) %>%
  left_join(dcm_depths, by=c("station", "cast")) %>%
  left_join(rbind(meso_SLA, falkor_SLA), by=c("station", "cast")) %>%
  filter(!(cruise=="FK" & station < 28)) %>%
  select(cruise, station, cast, time, lat, lon, DCM_depth, sla) %>%
  mutate(time=as.POSIXct(time, tz = "UTC")) %>% 
  mutate(local_hour=as.numeric(format(time, tz="HST", format = "%H")))

write.csv(station_metadata, file = "made_data/station_metadata.csv", row.names = FALSE)
```

```{r all_core and bottle_depth_data}
# Core Water Column Data from .gof file on the internet
if(!file.exists("raw_data/falkor_core.txt")){
  fk_meta <- "http://scope.soest.hawaii.edu/FTP/scope/water/fk180310.gof"
  download.file(fk_meta, destfile = "raw_data/falkor_core.txt")
}
if(!file.exists("raw_data/mesoscope_core.txt")){
  meso_meta <- "http://scope.soest.hawaii.edu/FTP/scope/water/mesoscope.gof"
  download.file(meso_meta, destfile = "raw_data/mesoscope_core.txt")
}
meso_core <- read_table("raw_data/mesoscope_core.txt", skip = 4, 
                        col_names = FALSE) %>%
  set_names(c("station", "cast", "rosette", "lat", "lon", "pressure", "temp",
              "sal", "ctd_oxy", "ctd_chl", "theta", "sigma", "oxy_um", "dic", "alk", "ph",
              "po4_um", "no23_um", "sio4_um", "LLN", "LLP", "PC", "PN", "PP",
              "PSi", "chl_ug", "phaeo_ug", "het_abund", 
              "pro_abund", "syn_abund", "euk_abund",
              "q1", "q2", "q3", "q4")) %>%
  mutate(across(everything(), function(x)ifelse(x==-9, NA, x))) %>%
  mutate(cruise=case_when(
    station <= 14 ~ "MT",
    station %in% 15:17 ~ "MC",
    TRUE ~ "Unknown"
  )) %>%
  mutate(station=case_when(
    station <= 15 ~ as.character(station),
    station == 16 ~ "L1",
    station == 17 ~ "L2"
  )) %>%
  select(cruise, station, cast, rosette, pressure, 
         temp, sal, theta, dic, alk, ph, oxy_um, ends_with("abund"))
fk_core <- read_table("raw_data/falkor_core.txt", skip = 4, 
                        col_names = FALSE) %>%
  set_names(c("station", "cast", "rosette", "lat", "lon", "pressure", "temp",
              "sal", "ctd_oxy", "ctd_chl", "theta", "sigma", "oxy_um", "dic", "alk", "ph",
              "po4_um", "no23_um", "sio4_um", "LLN", "LLP", "PC", "PN", "PP",
              "PSi", "chl_ug", "phaeo_ug", "het_abund", 
              "pro_abund", "syn_abund", "euk_abund",
              "q1", "q2", "q3", "q4")) %>%
  mutate(across(everything(), function(x)ifelse(x==-9, NA, x))) %>%
  mutate(cruise="FK") %>%
  filter(station>28) %>%
  mutate(station=as.character(station)) %>%
  select(cruise, station, cast, rosette, pressure, 
         temp, sal, theta, dic, alk, ph, oxy_um, ends_with("abund"))
all_core <- bind_rows(fk_core, meso_core)
write.csv(all_core, "made_data/all_core.csv", row.names = FALSE)

bottle_depth_data <- all_core %>%
  distinct(cruise, station, cast, rosette, pressure)
write.csv(bottle_depth_data, "made_data/bottle_depth_data.csv", row.names = FALSE)
```

```{r plot station_metadata}
station_metadata <- read_csv("made_data/station_metadata.csv", show_col_types = FALSE)
stations_summarized <- station_metadata %>%
  group_by(cruise, station, cast) %>%
  summarize(lat=median(lat), lon=median(lon), time=median(time), sla=mean(sla)) %>%
  arrange(time)

station_labels <- stations_summarized %>%
  filter(station%in%c(seq(1, 15, 2), "L1", "L2", 62, 64, 77, 80)) %>%
  filter(cast==1)

gm <- get_map(location = c(-158, 24), zoom = 6)
map_gp <- ggmap(ggmap = gm) +
  geom_point(aes(x=lon, y=lat), size=4, color="#FFFFFFAA", data = station_metadata) +
  geom_path(aes(x=lon, y=lat), data = station_metadata) +
  geom_point(aes(x=lon, y=lat, color=sla), data = station_metadata) +
  geom_label_repel(aes(x=lon, y=lat, label=station), color="steelblue", 
                   data = station_labels, min.segment.length = 0) +
  scale_color_gradient2(low = muted("blue"), high = muted("red"), mid = "grey80") +
  facet_wrap(~factor(cruise, levels = c("MT", "MC", "FK"), 
                     labels = c("Eddy transect", "Eddy center", "Falkor")), 
             scales = "free_x", nrow=1) +
  theme(legend.position = "none") +
  coord_cartesian(xlim = c(-162, -155), ylim=c(21, 28))

sla_ts_gp <- station_metadata %>%
  ggplot() +
  geom_point(aes(x=time, y=sla, color=sla)) +
  scale_color_gradient2(low = muted("blue"), high = muted("red"), mid = "grey80") +
  facet_wrap(~factor(cruise, levels = c("MT", "MC", "FK"), 
                     labels = c("Eddy transect", "Eddy center", "Falkor")), 
             scales = "free_x", nrow=1) +
  theme_bw() +
  theme(legend.position = c(0.5, 0.55), axis.title.x = element_blank())
station_metadata_map <- grid.arrange(map_gp, sla_ts_gp)
ggsave(plot = station_metadata_map, filename = "station_metadata_map.png", 
       path = "figures", device = "png", width = 8, height = 6, dpi = 144, 
       type = "cairo")
```

## Other bottle data (cruise, station, cast, depth, PC, PN, PO4, N+N, DON, DOP, ATP, chl)

This data is mostly available on the SCOPE website via Excel files

```{r raw data download}
if(!file.exists("raw_data/Karl_FK_PCPN_WC_current.xlsx")){
  # Data source: http://scope.soest.hawaii.edu/collaborators/datainventory/Data/Karl/Karl_FK_Nutrients_Final.xlsx
  falkor_CN_url <- paste0("http://", scope_uname, ":", scope_pw,
                "@scope.soest.hawaii.edu/collaborators/datainventory/",
                "Data/Karl/Karl_FK_PCPN_WC_current.xlsx")
  download.file(falkor_CN_url, destfile = "raw_data/Karl_FK_PCPN_WC_current.xlsx", mode = "wb")
}
if(!file.exists("raw_data/Karl_FK_Nutrients_Final.xlsx")){
  # Data source: http://scope.soest.hawaii.edu/collaborators/datainventory/Data/Karl/Karl_FK_Nutrients_Final.xlsx
  falkor_nutrient_url <- paste0("http://", scope_uname, ":", scope_pw,
                "@scope.soest.hawaii.edu/collaborators/datainventory/Data/",
                "Karl/Karl_FK_Nutrients_Final.xlsx")
  download.file(falkor_nutrient_url, destfile = "raw_data/Karl_FK_Nutrients_Final.xlsx", mode = "wb")
}
if(!file.exists("raw_data/2017MS_CoreData_current.xls")){
  # Data source: http://scope.soest.hawaii.edu/collaborators/datainventory/Data/SCOPEcore/2017MS_CoreData_current.xls
  transect_pcpn_url <- paste0("http://", scope_uname, ":", scope_pw,
                "@scope.soest.hawaii.edu/collaborators/datainventory/Data/",
                "SCOPEcore/2017MS_CoreData_current.xls")
  download.file(transect_pcpn_url, destfile = "raw_data/2017MS_CoreData_current.xls", mode = "wb")
}
if(!file.exists("raw_data/Karl_MS_Nutrient_Summary_KB.xlsx")){
  # Data source: http://scope.soest.hawaii.edu/collaborators/datainventory/Data/Karl/Karl_MS_Nutrient_Summary_KB.xlsx
  transect_nutrient_url <- paste0("http://", scope_uname, ":", scope_pw,
                "@scope.soest.hawaii.edu/collaborators/datainventory/Data/Karl/",
                "Karl_MS_Nutrient_Summary_KB.xlsx")
  download.file(transect_nutrient_url, destfile = "raw_data/Karl_MS_Nutrient_Summary_KB.xlsx", mode = "wb")
}
if(!file.exists("raw_data/Karl_MS_pATP.xlsx")){
  # Data source: http://scope.soest.hawaii.edu/collaborators/datainventory/Data/Karl/Karl_MS_pATP.xlsx
  atp_transect_url <- paste0("http://", scope_uname, ":", scope_pw,
         "@scope.soest.hawaii.edu/collaborators/datainventory/",
         "Data/Karl/Karl_MS_pATP.xlsx")
  download.file(atp_transect_url, destfile = "raw_data/Karl_MS_pATP.xlsx", mode = "wb")
}
```

```{r all_pcn}
fk_pcn <- "raw_data/Karl_FK_PCPN_WC_current.xlsx" %>%
  read_excel(skip=1) %>%
  slice(-1:-2) %>% 
  filter(is.na(Comments)) %>%
  mutate(station=as.numeric(str_extract(Name, "(?<=FK-)\\d+"))) %>%
  filter(station>=28) %>%
  mutate(rosette=as.numeric(str_extract(Name, "\\d+$"))) %>%
  mutate(cruise="FK", cast=1) %>%
  select(cruise, station, cast, rosette, abs_depth=`Depth (m)`, 
         PC_um=`C value (µmol/L)`, PN_um=`N value (µmol/L)`)
meso_pcn <- "raw_data/2017MS_CoreData_current.xls" %>%
  read_excel(skip = 1, sheet = "PCPN") %>%
  slice(-1) %>%
  select(abs_depth=depth_sample, station, PC_um=particulate_carbon, 
         PN_um=particulate_nitrogen, cast=cast_num, rosette=rosette_position) %>%
  mutate(across(everything(), as.numeric)) %>%
  mutate(cruise=case_when(
    station <= 14 ~ "MT",
    station %in% 15:17 ~ "MC",
    TRUE ~ "Unknown"
  )) %>%
  mutate(station=case_when(
    station <= 15 ~ as.character(station),
    station %in% 16 ~ "L1",
    station == 17 ~ "L2"
  ))
all_pcn <- rbind(meso_pcn, fk_pcn)
write.csv(all_pcn, "made_data/all_pcn.csv", row.names = FALSE)
```

```{r plot all_pcn}
pc_gp <- all_pcn %>%
  mutate(cruise=factor(cruise, levels = c("MT", "MC", "FK"), 
                   labels = c("Eddy transect", "Eddy center", "Falkor"))) %>%
  mutate(station=factor(station, levels=unique(station))) %>%
  ggplot(aes(x=PC_um, y=abs_depth, color=station, group=interaction(station, cast))) +
  geom_path() +
  geom_point() +
  facet_wrap(~cruise, scales="free_x") +
  ggtitle("Particulate carbon") +
  scale_y_reverse(name="Depth")  +
  xlim(0, NA) +
  theme(axis.title.x = element_blank(), legend.position = "bottom") +
  guides(color=guide_legend(nrow=2,byrow=TRUE, title = "Station"))
pn_gp <- all_pcn %>%
  mutate(cruise=factor(cruise, levels = c("MT", "MC", "FK"), 
                   labels = c("Eddy transect", "Eddy center", "Falkor"))) %>%
  mutate(station=factor(station, levels=unique(station))) %>%
  ggplot(aes(x=PN_um, y=abs_depth, color=station, group=interaction(station, cast))) +
  geom_path() +
  geom_point() +
  facet_wrap(~cruise, scales="free_x") +
  ggtitle("Particulate nitrogen") +
  scale_y_reverse(name="Depth")  +
  xlim(0, NA) +
  theme(axis.title.x = element_blank(), legend.position = "none")
all_pcn_plot <- grid.arrange(
  pc_gp, pn_gp, layout_matrix=matrix(c(1,1,1,1,2,2,2), ncol = 1)
)
ggsave(plot = all_pcn_plot, filename = "all_pcn_plot.png", 
       path = "figures", device = "png", width = 8, height = 6, dpi = 144, 
       type = "cairo")
```

```{r all_nuts}
fk_nuts <- "raw_data/Karl_FK_Nutrients_Final.xlsx" %>%
  read_excel(skip = 3) %>%
  select(abs_depth=`Depth`, sample=`Sample ID`, po4_um=`µM PO4`, no23_um=`µM N+N`) %>%
  filter(!is.na(abs_depth)) %>%
  filter(!abs_depth=="Depth") %>%
  mutate(rosette=as.numeric(str_extract(sample, "\\d+$"))) %>% 
  mutate(sample=ifelse(startsWith(sample, "FK"), sample, NA)) %>%
  fill(sample) %>%
  mutate(station=str_extract(sample, "(?<=FK 180310[- ])\\d+")) %>%
  mutate(across(c(abs_depth, po4_um, no23_um, station), as.numeric)) %>%
  filter(station>=28) %>%
  mutate(station=as.character(station)) %>%
  select(-sample) %>%
  mutate(cast=1, cruise="FK")
transect_profile_sheet <- "raw_data/Karl_MS_Nutrient_Summary_KB.xlsx" %>%
  read_excel(range = "KM1709 Transect, L1,L2 profiles!A4:N180") %>%
  slice(-1:-2) %>%
  mutate(sd=as.numeric(sd))
highres_sheet <- "raw_data/Karl_MS_Nutrient_Summary_KB.xlsx" %>%
  read_excel(range = "KM1709 High res nuts!A4:P53") %>%
  slice(-1:-2)
diel_sheet <- "raw_data/Karl_MS_Nutrient_Summary_KB.xlsx" %>%
  read_excel(range = "KM1709 Diel nuts!A4:O76") %>%
  slice(-1:-3)
meso_nuts <- bind_rows(transect_profile_sheet, highres_sheet, diel_sheet) %>% 
  select(abs_depth=Depth, station=Station, cast=Cast, rosette=Niskin,
         no23_um=`N+N`, LLN, DON_um=DON, TDN, 
         po4_um=`SRP`, LLP, DOP_um=DOP, TDP,
         si_um=`Si`) %>%
  mutate(no23_um=ifelse(no23_um=="BLD"|no23_um=="NA"|is.na(no23_um), LLN, no23_um)) %>%
  mutate(po4_um=ifelse(po4_um=="BLD"|po4_um=="NA"|is.na(po4_um), LLP, po4_um)) %>%
  mutate(across(-station, as.numeric)) %>%
  arrange(station, abs_depth, cast) %>%
  filter(!is.na(abs_depth)) %>%
  mutate(cruise=case_when(
    station %in% as.character(1:14) ~ "MT",
    station %in% c(15, "L1", "L2") ~ "MC",
    TRUE ~ "Unknown"
  )) %>%
  mutate(no23_um=ifelse(is.na(no23_um) & cruise=="MC", 0, no23_um)) %>%
  # Fixing a few rosette numbers per Benedetto conversation
  mutate(rosette=case_when(
    station=="13"&rosette==200~3,
    station=="13"&rosette==250~2,
    station=="13"&rosette==500~1,
    station=="14"&rosette==200~4,
    station=="14"&rosette==250~3,
    station=="14"&rosette==500~2,
    TRUE~rosette
  ))
all_nuts <- bind_rows(meso_nuts, fk_nuts) %>%
  left_join(station_metadata) %>%
  arrange(time) %>%
  select(cruise, station, cast, rosette, abs_depth, no23_um, po4_um, 
         DON_um, DOP_um, si_um) %>%
  mutate(station=factor(station, levels=unique(station)))
write.csv(all_nuts, "made_data/all_nuts.csv", row.names = FALSE)
```

```{r plot all_nuts}
refactored_nuts <- all_nuts %>%
  mutate(cruise=factor(cruise, levels = c("MT", "MC", "FK"), 
         labels = c("Eddy transect", "Eddy center", "Falkor")))

refactored_nuts %>%
  filter(is.na(no23_um))

no23_gp <- refactored_nuts %>%
  ggplot(aes(x=no23_um, y=abs_depth, color=station, group=interaction(station, cast))) +
  geom_path() +
  geom_point() +
  facet_wrap(~cruise, scales="free_x") +
  ggtitle("Nitrate + nitrite") +
  scale_y_reverse(name="Depth")  +
  xlim(0, NA) +
  theme(axis.title.x = element_blank(), legend.position = "none")
po4_gp <- refactored_nuts %>%
  ggplot(aes(x=po4_um, y=abs_depth, color=station, group=interaction(station, cast))) +
  geom_path() +
  geom_point() +
  facet_wrap(~cruise, scales="free_x") +
  ggtitle("Phosphate") +
  scale_y_reverse(name="Depth")  +
  xlim(0, NA) +
  theme(axis.title.x = element_blank(), legend.position = "none")
DON_gp <- refactored_nuts %>%
  ggplot(aes(x=DON_um, y=abs_depth, color=station, group=interaction(station, cast))) +
  geom_path() +
  geom_point() +
  facet_wrap(~cruise, scales="free_x") +
  ggtitle("Dissolved organic nitrogen") +
  scale_y_reverse(name="Depth")  +
  xlim(0, NA) +
  theme(axis.title.x = element_blank(), legend.position = "none")
DOP_gp <- refactored_nuts %>%
  ggplot(aes(x=DOP_um, y=abs_depth, color=station, group=interaction(station, cast))) +
  geom_path() +
  geom_point() +
  facet_wrap(~cruise, scales="free_x") +
  ggtitle("Dissolved organic phosphorus") +
  scale_y_reverse(name="Depth")  +
  xlim(0, NA) +
  theme(axis.title.x = element_blank(), legend.position = "none") +
  guides(color=guide_legend(nrow=2,byrow=TRUE, title = "Station"))
si_gp <- refactored_nuts %>%
  ggplot(aes(x=si_um, y=abs_depth, color=station, group=interaction(station, cast))) +
  geom_path() +
  geom_point() +
  facet_wrap(~cruise, scales="free_x") +
  ggtitle("Silica") +
  scale_y_reverse(name="Depth")  +
  xlim(0, NA) +
  theme(axis.title.x = element_blank(), legend.position = "bottom") +
  guides(color=guide_legend(nrow=2,byrow=TRUE, title = "Station"))
all_nuts_plot <- grid.arrange(
  no23_gp, po4_gp, DON_gp, DOP_gp, si_gp,
  layout_matrix=matrix(rep(1:5, c(3,3,3,3,5)), ncol = 1)
)
ggsave(plot = all_nuts_plot, filename = "all_nuts_plot.png", 
       path = "figures", device = "png", width = 6, height = 8, dpi = 144, 
       type = "cairo")
```

```{r meso_chl}
meso_chl <- "raw_data/2017MS_CoreData_current.xls" %>%
  read_excel(skip = 1, sheet = "Chl_Pheo") %>%
  slice(-1) %>%
  mutate(filter_min=as.numeric(filter_min)) %>%
  filter(filter_min>0.5) %>%
  select(abs_depth=depth_sample, station, chl_ug=fluorometric_chlorophyll_a, 
         phaeo_ug=total_phaeopigment, cast=cast_num, rosette=rosette_position) %>%
  mutate(across(everything(), as.numeric)) %>%
  mutate(station=case_when(
    station <= 15 ~ as.character(station),
    station == 16 ~ "L1",
    station == 17 ~ "L2"
  )) %>%
  mutate(cruise=case_when(
    station %in% as.character(1:14) ~ "MT",
    station %in% c(15, "L1", "L2") ~ "MC",
    TRUE ~ "Unknown"
  )) %>%
  # Removing one data point because it's negative
  filter(chl_ug>0) %>%
  # Removing one data point because it doesn't have an associated surface value
  filter(!(station=="L1" & cast==19)) %>%
  # Fixing one nonsense rosette number using all_core to find the rosette fired at 15m for S11C2
  mutate(rosette=ifelse(rosette==33, 23, rosette)) %>%
  select(-abs_depth)

# Deciding to replace the now-missing value with average of nearby stations 
# instead bc interpolation sux
s14_chl_surf_row <- meso_chl %>% 
  filter(station%in%c("12", "14")) %>%
  left_join(bottle_depth_data) %>%
  filter(pressure<30) %>%
  summarise(station="13", cast=2, rosette=23, cruise="MT", 
            chl_ug=round(mean(chl_ug), 5), phaeo_ug=round(mean(phaeo_ug), 5))
meso_chl <- bind_rows(meso_chl, s14_chl_surf_row)

write.csv(meso_chl, "made_data/meso_chl.csv", row.names = FALSE)
```

```{r plot meso_chl}
station_metadata %>%
  right_join(meso_chl) %>% 
  left_join(bottle_depth_data) %>%
  arrange(time) %>%
  mutate(station=factor(station, levels = unique(station))) %>%
  ggplot(aes(x=chl_ug, y=pressure, color=station, group=interaction(station, cast))) + 
  geom_point() + 
  geom_path() + 
  scale_y_reverse() +
  theme_bw() +
  theme(legend.position = "left") +
  facet_wrap(~cruise,nrow = 1)
ggsave(filename = "meso_chl_plot.png", path = "figures", device = "png", 
       width = 8, height = 6, dpi = 144, type = "cairo")
```

```{r beam_core from Benedetto}
if(!file.exists("raw_data/MESO-SCOPE_BeamAttenuation_forWill.csv")){
  stop("Go download the beam data from email, search 'beam attenuation'")
}
beam_dat <- read.csv("raw_data/MESO-SCOPE_BeamAttenuation_forWill.csv", 
                    skip = 3, header = FALSE)
beam_station_cast <- paste0(beam_dat[1,], " ", beam_dat[2,])
beam_dat <- beam_dat[3:nrow(beam_dat),]
colnames(beam_dat) <- beam_station_cast
beam_dat[] <- sapply(beam_dat, as.numeric)
beam_long <- beam_dat %>%
  pivot_longer(cols = 2:ncol(beam_dat), values_to = "beam_atten") %>%
  filter(!is.na(beam_atten)) %>%
  separate(name, into = c("station", "cast"), sep = " ") %>%
  rename(pressure=`NaN NaN`) %>%
  mutate(cast=as.numeric(cast))

beam_core <- all_core %>%
  group_by(station, cast) %>%
  group_split() %>% #`[[`(1) -> core_subset
  map_dfr(function(core_subset){
    beam_subset <- beam_long %>%
      filter(station==unique(core_subset$station), cast==unique(core_subset$cast))
    if(nrow(beam_subset)==0)return(NULL)
    approxout <- with(beam_subset, approx(pressure, beam_atten, core_subset$pressure))
    core_subset %>%
      select(cruise, station, cast, rosette, pressure) %>%
      left_join(set_names(as.data.frame(approxout), c("pressure", "beam_atten")), 
                by="pressure")
  })
write.csv(beam_core, "made_data/beam_core.csv", row.names = FALSE)
```

```{r plot beam_core and beam_long}
beam_long %>%
  filter(station%in%c("L1", "L2")) %>%
  filter(beam_atten < 0.1) %>%
  ggplot(aes(x=beam_atten, y=pressure, group=cast, color=beam_atten)) +
  geom_path() +
  facet_wrap(~station) +
  scale_color_viridis_c() +
  scale_y_reverse()
last_plot() +
  geom_point(data=beam_core %>% filter(station%in%c("L1", "L2")), size=1.5)
```

```{r light_core from Benedetto}
if(!file.exists("raw_data/MESO-SCOPE_DailyPAR_forWill.csv")){
  stop("Go download the light data from email, search 'beam attenuation'")
}
if(!file.exists("raw_data/MESO-SCOPE_PercentPAR_forWill.csv")){
  stop("Go download the light data from email, search 'beam attenuation'")
}
PAR_dat <- read.csv("raw_data/MESO-SCOPE_DailyPAR_forWill.csv", 
                    skip = 5, header = FALSE)
PAR_month_day <- paste0(PAR_dat[1,], " ", PAR_dat[2,])
PAR_month_day_pseudocast <- paste(PAR_month_day, c(1, rep(1:2, 6)))
PAR_dat <- PAR_dat[3:nrow(PAR_dat),]
colnames(PAR_dat) <- PAR_month_day_pseudocast
PAR_dat[] <- sapply(PAR_dat, as.numeric)
PAR_long <- PAR_dat %>%
  pivot_longer(cols = 2:ncol(PAR_dat), values_to = "PAR") %>%
  filter(!is.na(PAR)) %>%
  separate(name, into = c("month", "day", "pseudocast"), sep = " ") %>%
  rename(pressure=`NaN NaN 1`) %>%
  mutate(station=ifelse(as.numeric(day)>8, "L1", "L2")) %>%
  group_by(station, pressure) %>%
  summarise(PAR=mean(PAR))

PAR_perc_dat <- read.csv("raw_data/MESO-SCOPE_PercentPAR_forWill.csv", 
                    skip = 5, header = FALSE)
PAR_perc_month_day <- paste0(PAR_perc_dat[1,], " ", PAR_perc_dat[2,])
PAR_perc_month_day_pseudocast <- paste(PAR_perc_month_day, c(1, rep(1:2, 6)))
PAR_perc_dat <- PAR_perc_dat[3:nrow(PAR_perc_dat),]
colnames(PAR_perc_dat) <- PAR_perc_month_day_pseudocast
PAR_perc_dat[] <- sapply(PAR_perc_dat, as.numeric)
PAR_perc_long <- PAR_perc_dat %>%
  pivot_longer(cols = 2:ncol(PAR_perc_dat), values_to = "PAR_perc") %>%
  filter(!is.na(PAR_perc)) %>%
  separate(name, into = c("month", "day", "pseudocast"), sep = " ") %>%
  rename(pressure=`NaN NaN 1`) %>%
  mutate(station=ifelse(as.numeric(day)>8, "L1", "L2")) %>%
  group_by(station, pressure) %>%
  summarise(PAR_perc=mean(PAR_perc))

PAR_long <- PAR_long %>%
  full_join(PAR_perc_long, by=c("station", "pressure"))

light_core <- all_core %>%
  filter(station%in%c("L1", "L2")) %>%
  group_by(station) %>%
  group_split() %>% #`[[`(1) -> core_subset
  map_dfr(function(core_subset){
    PAR_subset <- PAR_long %>%
      filter(station==unique(core_subset$station))
    PARout <- with(PAR_subset, approx(pressure, PAR, core_subset$pressure))
    percout <- with(PAR_subset, approx(pressure, PAR_perc, core_subset$pressure))
    core_subset %>%
      select(cruise, station, cast, rosette, pressure) %>%
      left_join(data.frame(pressure=PARout$x, PAR=PARout$y, PAR_perc=percout$y), 
                by="pressure")
  })
light_core %>%
  ggplot() +
  geom_point(aes(x=PAR_perc, y=pressure, color=station)) +
  facet_wrap(~station) +
  scale_y_reverse()
write.csv(light_core, "made_data/light_core.csv", row.names = FALSE)
```

```{r pcn outlier removal (requires beam_core)}
# Remove two outlier samples because Eric thinks the values are suspicious
# filter(!(station%in%c("10", "11") & abs_depth==15))
# Advance prediction is that they will be far off the beam_atten:PC line
# Plan: replace existing PC values with those from a regression
# on the non-weird PC values for that given SLA

pcn_init <- all_pcn %>%
  left_join(bottle_depth_data) %>%
  filter(cruise=="MT") %>% #filter(station=="4") -> station_subset
  group_by(station) %>%
  group_split() %>%
  map_dfr(function(station_subset){
    beam_subset <- beam_core %>%
      filter(station==unique(station_subset$station)) %>%
      distinct(pressure, beam_atten)
    approx(beam_subset$pressure, beam_subset$beam_atten, station_subset$pressure) %>%
      as.data.frame() %>%
      setNames(c("pressure", "beam_atten")) %>%
      left_join(station_subset, by="pressure")
  }) %>%
  group_by(temp=station%in%c("10", "11") & abs_depth==15) %>%
  group_split()

beam_pc_model <- lm(PC_um~beam_atten, data = pcn_init[[1]])
pcn_init[[2]]$PC_um <- predict(beam_pc_model, pcn_init[[2]])

# Plot it to double-check outputs
plot(PC_um~beam_atten, data = pcn_init[[1]])
abline(beam_pc_model)
legend("topleft", legend = paste("R^2 =", round(summary(beam_pc_model)$r.squared, 2)))
points(pcn_init[[2]]$beam_atten, predict(beam_pc_model, pcn_init[[2]]), col="red")

all_pcn <- bind_rows(pcn_init) %>%
  select(-temp, -beam_atten, -pressure) %>%
  bind_rows(all_pcn %>% filter(cruise!="MT")) %>%
  select(-abs_depth)

write.csv(all_pcn, "made_data/all_pcn.csv", row.names = FALSE)
```

## Combine all the above into env_metadata

```{r load made_data and make env_metadata}
station_metadata <- read_csv("made_data/station_metadata.csv")
all_pcn <- read_csv("made_data/all_pcn.csv")
all_nuts <- read_csv("made_data/all_nuts.csv")
meso_chl <- read_csv("made_data/meso_chl.csv")
all_core <- read_csv("made_data/all_core.csv")
beam_core <- read_csv("made_data/beam_core.csv")
light_core <- read_csv("made_data/light_core.csv")

env_metadata <- all_core %>%
  left_join(station_metadata) %>%
  left_join(all_nuts) %>%
  left_join(all_pcn) %>% 
  left_join(meso_chl) %>%
  left_join(beam_core) %>%
  left_join(light_core) %>%
  arrange(cruise, station, cast, rosette) %>%
  # Organize columns, not dropping anything
  select(cruise, station, cast, time, lat, lon, DCM_depth, sla, local_hour,
         rosette, pressure, abs_depth, temp, sal, theta, dic, alk, ph, oxy_um,
         PAR, PAR_perc, no23_um, po4_um, DON_um, DOP_um, si_um,
         PC_um, PN_um, chl_ug, phaeo_ug, beam_atten,
         everything()) %>%
  mutate(abs_depth=ifelse(is.na(abs_depth), pressure, abs_depth))

write.csv(env_metadata, "made_data/env_metadata.csv", row.names = FALSE)
```

```{r check via plots, eval=FALSE}
env_metadata <- read_csv("made_data/env_metadata.csv")

# Check on cruise, time, lat, lon, sla
env_metadata %>%
  arrange(time) %>%
  distinct(cruise, lon, lat, sla) %>%
  ggplot(aes(x=lon, y=lat, group=cruise)) +
  geom_path(aes(color=cruise), lwd=0.7) +
  geom_point(size=3, aes(color=cruise)) +
  scale_color_manual(values = hcl.colors(5)[2:4]) +
  ggnewscale::new_scale_color() +
  geom_point(aes(color=sla), size=2.2) +
  coord_fixed() +
  scale_color_gradient2(low = "steelblue4", high = "firebrick") +
  theme_minimal()

# PCPN and nutrient data
env_metadata %>%
  pivot_longer(c(ends_with("um"))) %>%
  filter(!is.na(value)) %>%
  filter(pressure<260) %>%
  select(cruise, station, cast, pressure, name, value) %>%
  ggplot(aes(x=value, y=pressure)) +
  geom_path(aes(group=interaction(station, cast))) +
  geom_point(aes(color=station)) +
  facet_grid(cruise~name, scales="free_x") +
  scale_y_reverse()

env_metadata %>%
  select(cruise, time, pressure:last_col()) %>%
  pivot_longer(cols = temp:last_col()) %>%
  filter(pressure<260) %>%
  group_by(name) %>%
  mutate(value=(value-min(value, na.rm = TRUE))/
           (max(value, na.rm = TRUE)-min(value, na.rm = TRUE))) %>%
  ungroup() %>%
  mutate(name=factor(name, levels=unique(name))) %>%
  mutate(cruise=factor(cruise, levels=c("MT", "MC", "FK"))) %>%
  filter(!is.na(value)) %>%
  ggplot() +
  geom_point(aes(x=time, y=pressure, color=value)) +
  facet_grid(name~cruise, scales = "free_x") +
  scale_y_reverse() +
  scale_color_viridis_c() +
  theme(legend.position = "none")
ggsave("figures/env_metadata_all.pdf", device = "pdf", width = 8, height = 20,
       units = "in")
```

## Metabolite sample data creation (file_metadata)

```{r samp_log from sample log}
if(!file.exists("raw_data/MESO_SCOPE_Cruise_KM1709_Sample_Log.xlsx")){
  stop("Go download the data from email, search 'MESO_SCOPE_Cruise_KM1709_Sample_Log'")
  # Needs to be converted to .xlsx and downloaded
}
if(!file.exists("raw_data/MESO_SCOPE_Cruise_KM1709_Sample_Log.xlsx")){
  stop("Go download the data from email, search 'FALKOR- Sample Log_INGALLS'")
  # Needs to be converted to .xlsx and downloaded
}
bottle_depth_data <- read_csv("made_data/bottle_depth_data.csv")
MT_samps <- "raw_data/MESO_SCOPE_Cruise_KM1709_Sample_Log.xlsx" %>%
  read_excel(sheet = "Metabolites_and_Proteins", range = "A2:J114") %>%
  mutate(cruise="MT") %>%
  select(cruise, station=Station, cast=Cast, depth=`Depth (m)`, 
         rosette=`Niskin #`, id=`Sample ID`) %>%
  filter(complete.cases(.)) %>%
  mutate(station=str_remove(station, "^S")) %>%
  mutate(cast=as.numeric(str_remove(cast, "^C"))) %>%
  # We can grab DCM values here or from the mini DCM spreadsheet I made later
  # mutate(abs_depth=as.numeric(str_remove(depth, "DCM="))) %>%
  mutate(depth=str_replace(depth, ".0", "m")) %>%
  mutate(depth=str_replace(depth, "=.*", "")) %>%
  # Guessing at rosette number here - doesn't really matter bc we're matching
  # based on station more than exact rosette mostly and all bottles
  # were fired at similar depths and times
  mutate(rosette=ifelse(rosette=="?", 1, rosette)) %>% 
  mutate(rosette=as.numeric(rosette)) %>%
  # Fixing likely transcription error - sample looks like 15m samples and
  # all other triplicates were 19, 20, and 21
  mutate(rosette=ifelse(rosette==4 & cast==3 & station=="6", 21, rosette)) %>%
  left_join(bottle_depth_data)
MC_samps <- "raw_data/MESO_SCOPE_Cruise_KM1709_Sample_Log.xlsx" %>%
  read_excel(sheet = "Metabolites_High_Resolution", range = "A2:J35") %>%
  mutate(cruise="MC") %>%
  select(cruise, station=Station, cast=Cast, depth=`Depth (m)`, 
         rosette=`Niskin #`, id=`Sample ID`) %>%
  filter(!is.na(station)) %>%
  mutate(station=str_remove(station, "^S")) %>%
  mutate(cast=as.numeric(str_remove(cast, "^C"))) %>%
  mutate(rosette=as.numeric(rosette)) %>%
  mutate(depth=str_replace(depth, "\\-", "less")) %>%
  mutate(depth=str_replace(depth, "\\+", "plus")) %>%
  mutate(id=str_remove(id, "^MHR")) %>%
  left_join(bottle_depth_data)
samp_log <- bind_rows(MT_samps, MC_samps)
```

```{r file_metadata from mzML list}
samp_metadata <- list.files("../mzMLs/pos", pattern = "Smp.*mzML") %>%
  data.frame(filename=.) %>%
  mutate(samp_type="Smp") %>%
  mutate(cruise=case_when(
    str_detect(filename, "180205") ~ "MC",
    str_detect(filename, "180821") ~ "MT",
    str_detect(filename, "190715") ~ "FK",
    TRUE ~ "Unknown"
  )) %>%
  mutate(station=str_extract(filename, "((?<=S)\\d+)|L\\d")) %>%
  mutate(cast=case_when(
    cruise=="MC" ~ str_extract(filename, "(?<=C)\\d+"),
    cruise=="MT" ~ str_extract(filename, "(?<=C)\\d"),
    cruise=="FK" ~ "1",
  )) %>%
  mutate(cast=as.numeric(cast)) %>%
  mutate(depth=str_extract(filename, "15m|25m|DCM(plus|less)?(10m|20m)?|175m")) %>%
  mutate(tripl=str_extract(filename, "(?<=_)[A-C](?=.mzML$)")) %>%
  mutate(untripl=str_replace(filename, "_[A-C].mzML$", "")) %>%
  mutate(date_run=str_extract(filename, "^\\d+")) %>%
  mutate(date_run=strptime(date_run, format = "%y%m%d")) %>%
  mutate(id=str_extract(filename, "(?<=_)(MS|L).*_[A-C](?=.mzML$)")) %>%
  mutate(id=str_replace(id, "less", "-")) %>%
  mutate(id=str_replace(id, "plus", "+"))
absolute_depths <- samp_metadata %>%
  distinct(station, depth) %>%
  mutate(depth_diff=sapply(depth, function(depth){
    depth_num <- str_extract(depth, "(?<=DCM(plus|less))\\d*")
    depth_num <- ifelse(is.na(depth_num), 0, depth_num)
    depth_sign <- sapply(str_extract(depth, "less|plus"), 
                         switch, "plus"=1, "less"=-1, 0, USE.NAMES = FALSE)
    as.numeric(depth_num)*depth_sign
  })) %>%
  left_join(read.csv("raw_data/all_DCMs.csv")) %>%
  mutate(abs_depth=DCM_depth+depth_diff) %>%
  mutate(abs_depth=ifelse(str_detect(depth, "DCM"), abs_depth, 
                          as.numeric(str_extract(depth, "^\\d+")))) %>%
  select(station, cast, depth, abs_depth)
samp_metadata <- samp_metadata %>%
  left_join(absolute_depths) %>%
  left_join(samp_log, by=c("cruise", "station", "cast", "depth", "id")) %>%
  # If we have an exact depth for a given bottle, use it, otherwise trust
  # the depth extracted from sample sheets or file names
  mutate(abs_depth = ifelse(is.na(pressure), abs_depth, pressure)) %>%
  select(-pressure)

extra_metadata <- list.files("../mzMLs/pos", pattern = "Blk|Std|Poo.*mzML") %>%
  data.frame(filename=.) %>%
  mutate(samp_type=str_extract(filename, "Std|Blk|Poo")) %>%
  mutate(cruise=case_when(
    str_detect(filename, "170706") ~ "MC",
    str_detect(filename, "180205") ~ "MC",
    str_detect(filename, "180821") ~ "MT",
    str_detect(filename, "190715") ~ "FK",
    TRUE ~ "Unknown"
  )) %>%
  mutate(date_run=str_extract(filename, "^\\d+")) %>%
  mutate(date_run=strptime(date_run, format = "%y%m%d"))

if(!file.exists("made_data/file_timestamps.csv")){
  list.files("../mzMLs/pos", pattern = "mzML", full.names = TRUE) %>%
    grabMSdata(grab_what = "metadata") %>%
    `$`("metadata") %>%
    select(filename, timestamp) %>%
    write.csv("made_data/file_timestamps.csv", row.names = FALSE)
}
file_timestamps <- read_csv("made_data/file_timestamps.csv") %>%
  fill(timestamp, .direction = "up")

file_metadata <- bind_rows(samp_metadata, extra_metadata) %>%
  left_join(file_timestamps) %>%
  arrange(timestamp) %>%
  # Just reordering, nothing getting dropped
  select(c("filename", "samp_type", "cruise", "station", "cast", "depth", 
           "abs_depth", "rosette", "tripl", "id", "untripl", "date_run", 
           "timestamp"))
write.csv(file_metadata, file = "made_data/file_metadata.csv", row.names = FALSE)
```

## Interpolate metadata for metabolite samples

```{r summary similarity distance dot plot}
env_metadata <- read_csv("made_data/env_metadata.csv")
env_metadata_long <- env_metadata %>%
  pivot_longer(cols = temp:het_abund) %>%
  filter(!is.na(value)) %>%
  select(cruise, station, cast, lat, abs_depth, name, value)

samp_metadata_w_time <- read_csv("made_data/file_metadata.csv") %>%
  filter(samp_type=="Smp") %>%
  left_join(env_metadata %>% distinct(cruise, station, cast, lat, local_hour))

cruise_bounds <- samp_metadata_w_time %>%
  group_by(cruise) %>%
  summarise(min_bound=min(lat)-0.25, max_bound=max(lat)+0.25)

metadata_gp <- env_metadata_long %>%
  left_join(cruise_bounds) %>%
  # filter(time > min_bound & time < max_bound) %>%
  filter(lat > min_bound & lat < max_bound) %>%
  filter(abs_depth < 300) %>%
  ggplot(aes(x=lat, y=abs_depth)) +
  geom_jitter(aes(color=name), width = 0, height = 0) +
  geom_point(color="black", data = samp_metadata_w_time) +
  scale_y_reverse() +
  facet_wrap(~factor(cruise, levels = c("MT", "MC", "FK"), 
                     labels = c("Eddy transect", "Eddy center", "Falkor")), 
             scales = "free_x", nrow=1)
cruisetrack_gp <- env_metadata %>%
  arrange(time) %>%
  group_by(cruise, station) %>%
  mutate(lat=mean(lat), sla=mean(sla)) %>%
  left_join(cruise_bounds) %>%
  filter(lat > min_bound & lat < max_bound) %>%
  ggplot(aes(x=lat, y=sla, label=station)) +
  geom_path() +
  geom_point(aes(color=sla)) +
  facet_wrap(~cruise, scales = "free_x") +
  scale_color_gradient2(low = muted("blue"), high = muted("red"), mid = "grey80") +
  facet_wrap(~factor(cruise, levels = c("MT", "MC", "FK"), 
                     labels = c("Eddy transect", "Eddy center", "Falkor")), 
             scales = "free_x", nrow=1)

top_left_indent <- 1
top_right_indent <- 40
layout_mat <- rbind(
  c(rep(NA, top_left_indent), rep(1, 1000-top_left_indent-top_right_indent),
    rep(NA, top_right_indent)),
  2, 2, 2
)
comb_gp <- grid.arrange(cruisetrack_gp, metadata_gp, layout_matrix=layout_mat)
ggsave(comb_gp, filename = "env_metadata.png", device = "png", type="cairo",
       width = 10, height = 8, path = "figures")
# plotly::ggplotly(metadata_gp)
```

```{r interpolate env_metadata to metab sample values}
station_metadata <- read_csv(file = "made_data/station_metadata.csv")
file_metadata <- read_csv("made_data/file_metadata.csv") %>%
  left_join(station_metadata)
env_metadata <- read_csv("made_data/env_metadata.csv")

temp_env_meta <- env_metadata %>%
  select(cruise, station, value_depth=abs_depth, temp:euk_abund) %>%
  mutate(pseudostation=case_when(
    station%in%as.character(74:86)~"anticyc",
    station%in%as.character(27:59)~"cyc",
    TRUE~station
  ))
temp_file_meta <- file_metadata %>%
  filter(samp_type=="Smp") %>%
  select(station, cruise, filename, samp_depth=abs_depth) %>%
  mutate(pseudostation=case_when(
    station%in%as.character(74:100)~"anticyc",
    station%in%as.character(25:70)~"cyc",
    TRUE~station
  ))

interp_out <- temp_env_meta %>%
  bind_rows(temp_file_meta) %>%
  group_by(pseudostation) %>%
  group_split() %>%
  map_dfr(function(single_station){
    single_station %>%
      mutate(samp_depth=ifelse(is.na(samp_depth), value_depth, samp_depth)) %>%
      mutate(across(temp:euk_abund, function(col){
        if(sum(!is.na(col))<2){return(NA)}
        approx(x = value_depth, y=col, xout = samp_depth, rule = 2)$y
      })) %>%
      suppressWarnings() %>%
      filter(!is.na(filename))
  }) %>%
  select(-value_depth, -pseudostation, -samp_depth)

filled_file_metadata <- file_metadata %>% left_join(interp_out)

write.csv(filled_file_metadata, "made_data/filled_file_metadata.csv", row.names = FALSE)
```

```{r load data for plotting}
filled_file_metadata <- read_csv("made_data/filled_file_metadata.csv")
filled_MT_metadata <- filled_file_metadata %>% filter(cruise=="MT")
filled_MC_metadata <- filled_file_metadata %>% filter(cruise=="MC")
env_metadata <- read_csv("made_data/env_metadata.csv")
```

```{r plot MT interpolated data as transects}
meta_nms <- c("no23_um", "po4_um", "chl_ug", "phaeo_ug", 
             "PC_um", "PN_um", "DON_um", "DOP_um",
             "dic", "alk", "ph", "oxy_um", 
              paste0(c("pro", "syn", "het", "euk"), "_abund"))

meta_MT_gps <- lapply(meta_nms, function(chosen_var){
  env_metadata %>%
    filter(cruise=="MT") %>%
    filter(!is.na(get(chosen_var))) %>%
    ggplot(aes(x=time, y=abs_depth, color=get(chosen_var))) + 
    geom_point(aes(y=abs_depth-2), size=4.5, color="grey50") +
    geom_point(aes(y=abs_depth-2), size=4) +
    geom_point(aes(y=abs_depth), filled_MT_metadata, size=4.5, color="red") +
    geom_point(aes(y=abs_depth), filled_MT_metadata, size=4) +
    ylim(250, 0) +
    scale_color_viridis_c() +
    ggtitle(chosen_var) +
    theme_bw() +
    theme(legend.position = "none", axis.title = element_blank())
})

comb_gp <- do.call(grid.arrange, c(meta_MT_gps, ncol=4))
ggsave(comb_gp, filename = "figures/interp_MT_data.png", device = "png", 
       width = 10, height = 10, units = "in", type="cairo")
```

```{r plot MT interped data as depth profiles}
meta_MT_depthgps <- lapply(meta_nms, function(chosen_var){
  env_metadata %>%
    filter(cruise=="MT") %>%
    filter(!is.na(get(chosen_var))) %>%
    ggplot(aes(x=get(chosen_var), y=abs_depth, color=get(chosen_var))) + 
    geom_path(aes(group=station), color="black", lwd=1) +
    geom_point(size=4) +
    geom_point(data = filled_MT_metadata, color="red") +
    ylim(250, 0) +
    scale_color_viridis_c() +
    ggtitle(chosen_var) +
    theme_bw() +
    theme(legend.position = "none", axis.title = element_blank())
})
comb_gp <- do.call(grid.arrange, c(meta_MT_depthgps, ncol=4))
ggsave(comb_gp, filename = "figures/interp_MT_depthprofs.png", device = "png", 
       width = 10, height = 10, units = "in", type="cairo")
```

```{r plot MC interpolated data as transects}
meta_nms <- c("no23_um", "po4_um", "chl_ug", "phaeo_ug", 
             "PC_um", "PN_um", "DOP_um", 
             paste0(c("pro", "syn", "het", "euk"), "_abund"),
             "beam_atten", "PAR")

meta_MC_gps <- lapply(meta_nms, function(chosen_var){
  env_metadata %>%
    filter(cruise=="MC") %>%
    filter(!is.na(get(chosen_var))) %>%
    ggplot(aes(x=time, y=abs_depth, color=get(chosen_var))) + 
    geom_point(aes(y=abs_depth-2), size=4.5, color="grey50") +
    geom_point(aes(y=abs_depth-2), size=4) +
    geom_point(aes(y=abs_depth), filled_MC_metadata, size=4.5, color="red") +
    geom_point(aes(y=abs_depth), filled_MC_metadata, size=4) +
    ylim(250, 0) +
    scale_color_viridis_c() +
    ggtitle(chosen_var) +
    theme_bw() +
    theme(legend.position = "none", axis.title = element_blank())
})

comb_gp <- do.call(grid.arrange, c(meta_MC_gps, ncol=4))
ggsave(comb_gp, filename = "figures/interp_MC_data.png", device = "png", 
       width = 10, height = 10, units = "in", type="cairo")
```

```{r plot MC interped data as depth profiles}
meta_MC_depthgps <- lapply(meta_nms, function(chosen_var){
  env_metadata %>%
    filter(cruise=="MC") %>%
    filter(!is.na(get(chosen_var))) %>%
    ggplot(aes(x=get(chosen_var), y=abs_depth, color=get(chosen_var))) + 
    geom_path(aes(group=station), color="black", lwd=1) +
    geom_point(size=4) +
    geom_point(data = filled_MC_metadata, color="red") +
    ylim(250, 0) +
    scale_color_viridis_c() +
    ggtitle(chosen_var) +
    theme_bw() +
    theme(legend.position = "none", axis.title = element_blank())
})
comb_gp <- do.call(grid.arrange, c(meta_MC_depthgps, ncol=4))
ggsave(comb_gp, filename = "figures/interp_MC_depthprofs.png", device = "png", 
       width = 10, height = 10, units = "in", type="cairo")
```

Metadata pairs plot and heatmaply things

```{r heatmaply metadata}
filled_MT_metadata <- filled_file_metadata %>%
  filter(cruise=="FK") %>%
  filter(samp_type=="Smp") %>%
  select(where(~!all(is.na(.x))))

heatmap_cor <- filled_MT_metadata %>%
  select(abs_depth, time, lat, sla, local_hour, #alk, dic, ph
         ends_with("abund"), #beam_atten,
         ends_with("um"), ends_with("ug")) %>%
  data.matrix() %>%
  cor(use = "pair") %>%
  abs()

cor(cophenetic(hclust(dist(heatmap_cor))), dist(heatmap_cor))
pheatmap::pheatmap(
  mat = heatmap_cor, color = viridisLite::viridis(100),
  filename = "figures/meta_cor_clust.pdf", width = 7, height = 7
)
heatmaply::heatmaply(heatmap_cor)
```

```{r ggally pairsplot}
filled_MT_metadata <- filled_file_metadata %>%
  filter(cruise=="MT") %>%
  mutate(depth=factor(depth, levels = c("15m", "DCM", "175m")))

library(GGally)
filled_MT_metadata %>%
  select(station, depth, abs_depth, time, lat, sla, local_hour, alk, dic, ph, 
         ends_with("um"), ends_with("ug"), ends_with("abund")) %>%
  ggpairs(aes(color=depth)) %>%
  ggsave(filename = "MT_ggpairsplot.pdf", device = "pdf", path = "figures", 
         width = 30, height = 30, units = "in", dpi = 300)
```

## PCA groupings

```{r pca, eval=FALSE}
filled_MT_metadata <- read_csv("made_data/filled_file_metadata.csv") %>%
  filter(cruise=="MT") %>%
  filter(samp_type=="Smp") %>%
  mutate(depth=factor(depth, levels = c("15m", "DCM", "175m")))

pca_output <- filled_MT_metadata %>%
  select(filename, abs_depth, time, lat, sla, local_hour, #alk, dic, ph, 
         ends_with("abund"), ends_with("um"), ends_with("ug"),
         beam_atten, -oxy_um) %>%
  column_to_rownames("filename") %>%
  data.matrix() %>%
  scale() %>%
  prcomp()
# plot(pca_output)
biplot(pca_output)
biplot(pca_output, choices = c(3, 4))

ggpca_data <- pca_output$x %>%
  as.data.frame() %>%
  rownames_to_column("filename") %>%
  select(filename, PC1:PC4) %>%
  left_join(filled_MT_metadata)
ggpca_arrows <- pca_output$rotation %>% 
  `*`(10) %>%
  as.data.frame() %>%
  rownames_to_column("env_param")
ggpca_labels <- ggpca_arrows %>% 
  mutate(PC2=PC2+0.2) %>%
  mutate(PC2=case_when(
    env_param=="sla"~PC2-0.4,
    env_param=="PC_um"~PC2+0.2,
    env_param=="phaeo_ug"~PC2+0.2,
    TRUE~PC2)) %>%
  mutate(PC4=PC4+0.3)
ggplot(mapping = aes(x=PC1, y=PC2)) + 
  geom_point(data = ggpca_data, aes(color=depth)) +
  geom_segment(data = ggpca_arrows, aes(xend=0, yend=0), 
               arrow = arrow(length = unit(0.02, "npc"), ends = "first")) +
  geom_label(data = ggpca_labels, aes(label=env_param)) +
  theme_bw() +
  theme(legend.position = c(0.8, 0.8))
ggsave(filename = "MT_PC1_PC2.png", device = "png", path = "figures", 
       width = 5, height = 5, dpi = 144, units = "in", type="cairo")
ggplot(mapping=aes(x=PC3, y=PC4)) + 
  geom_point(data=ggpca_data, aes(color=sla, size=lat)) +
  geom_segment(data = ggpca_arrows, aes(xend=0, yend=0), 
               arrow = arrow(length = unit(0.02, "npc"), ends = "first")) +
  geom_label(data = ggpca_labels, aes(label=env_param)) +
  theme_bw()
ggsave(filename = "MT_PC3_PC4.png", device = "png", path = "figures", 
       width = 5.5, height = 5, dpi = 144, units = "in", type="cairo")
```

## Get standards and clean up
It's also nice to pull down a list of standards that will be found in the data.

```{r grabstans}
if(!file.exists("raw_data/raw_stans.csv")){
  raw_stans <- read.csv(paste0(
    "https://raw.githubusercontent.com/IngallsLabUW/Ingalls_Standards",
    "/master/Ingalls_Lab_Standards.csv"
  ))
  write.csv(raw_stans, file = "raw_data/raw_stans.csv", row.names = FALSE)
}
raw_stans <- read.csv("raw_data/raw_stans.csv")
clean_stans <- raw_stans %>%
  filter(Column=="HILIC") %>%
  mutate(polarity=ifelse(str_detect(z, "-"), "neg", "pos")) %>%
  mutate(mz=as.numeric(mz)) %>%
  mutate(peak_quality=str_extract(Peak_Quality, "Yes|No|Problematic")) %>%
  select(compound_type=Compound_Type, compound_name=Compound_Name,
         compound_name_old=Compound_Name_Original,
         formula=Empirical_Formula, rt=RT_minute, mz=mz, 
         ionization_form=Ionization_Form,
         charge=z, kegg_id=KEGG_Code, polarity, conc_um=Concentration_uM,
         date_added=Date_Added, mix=HILIC_Mix, priority=Priority,
         peak_quality) %>%
  add_row(compound_type="Custom", compound_name="Bisulfate",
          compound_name_old="Bisulfate", formula="HSO4", rt=13, 
          mz=95.95172965+1.007276, ionization_form="[M]", 
          charge=-1, date_added=150101, polarity="neg") %>%
  add_row(compound_type="Custom", compound_name="Bromine81",
          compound_name_old="Bromine upper", formula="Br(81)", rt=7, 
          mz=80.91563, ionization_form="[M]", charge=-1, 
          date_added=150101, polarity="neg") %>%
  add_row(compound_type="Custom", compound_name="Bromine79",
          compound_name_old="Bromine lower", formula="Br(79)", rt=7, 
          mz=78.91768, ionization_form="[M]", charge=-1, 
          date_added=150101, polarity="neg") %>%
  add_row(compound_type="Custom", compound_name="Pyro from glutamine?",
          compound_name_old=compound_name, formula="C5H7NO3", rt=11, 
          mz=130.050419, ionization_form="[M-NH3+H]", charge=-1, 
          date_added=150101, polarity="pos") %>%
  add_row(compound_type="Custom", compound_name="Pyro from glutamate?",
          compound_name_old=compound_name, formula="C5H7NO3", rt=12, 
          mz=130.050419, ionization_form="[M-H2O+H]", charge=-1, 
          date_added=150101, polarity="pos") %>%
  # add_row(compound_type="Custom", compound_name="Deoxyguanosine?",
  #         compound_name_old=compound_name, formula="C10H13N5O4", rt = 4.5,
  #         mz=268.10403, ionization_form="[M+H]", charge=1, 
  #         date_added=150101, polarity="pos") %>%
  add_row(compound_type="Custom", compound_name="Deoxycytidine?",
          compound_name_old=compound_name, formula="C10H13N5O4", rt = 6.6,
          mz=228.09788, ionization_form="[M+H]", charge=1, 
          date_added=150101, polarity="pos") %>%
  add_row(compound_type="Custom", compound_name="Homoserine betaine?",
          compound_name_old=compound_name, formula="C7H15NO3", rt = 6.7,
          mz=162.1125, ionization_form="[M+H]", charge=1, 
          date_added=150101, polarity="pos") %>%
  add_row(compound_type="Custom", compound_name="Threonine betaine?",
          compound_name_old=compound_name, formula="C7H15NO3", rt = 5,
          mz=162.1125, ionization_form="[M+H]", charge=1, 
          date_added=150101, polarity="pos") %>%
  add_row(compound_type="Custom", compound_name="Theanine?",
          compound_name_old=compound_name, formula="C7H14N2O3", rt = 6.8,
          mz=175.10771, ionization_form="[M+H]", charge=1, 
          date_added=150101, polarity="pos") %>%
  mutate(date_added=strptime(x = date_added, format = "%y%m%d"))
write.csv(clean_stans, file = "made_data/clean_stans.csv", row.names = FALSE)
```

### 2022-10-05 targeted/output_control.Rmd
Now tracking targeted data

---
title: "Peak area control doc"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options("readr.num_columns" = 0)

library(tidyverse)

already_good <- 0.1
min_improvement <- 0.4

cv <- function(x)sd(x)/mean(x)

clean_stans <- read.csv("../metadata/made_data/clean_stans.csv")
name_conversion <- clean_stans %>%
  select(new=compound_name, old=compound_name_old) %>%
  distinct()
```

# Convert raw data ----

First step is getting the data out of Skyline/MSDIAL. This is doable by script
if the original .sky and .mdf files are available, as they should be in this
case in the raw_data folder broken down by cruise.

We also need the console interfaces to both Skyline and MSDIAL.
MSDIAL's comes with the app and is pointed to with a raw file path to wherever
the MSDIAL files were extracted, and Skyline requires a separate download of
their command line version.

Mesotransect was integrated with MSDIAL, Falkor and Mesocenter used Skyline.

```{r grabRAnotes}
if(!file.exists("raw_data/Raafay_HILIC_notes.csv")){
  library(googlesheets4)
  if(!gs4_has_token())stop("Refresh Google token")
  # googlesheets4::gs4_auth()

  manual_pos <- read_sheet(ss = "1lU7g5rfem81s_0fqMP93yzegogCN29JJEdsIESfAX7c", 
                           skip = 1, sheet = "HILIC_POS") %>%
    select(compound_name=`Compound name`,
           integrated=`Integrated? (Y/N)`,
           in_samps=`In samples? (Y/N)`) %>%
    mutate(polarity="pos")
  manual_neg <- read_sheet(ss = "1lU7g5rfem81s_0fqMP93yzegogCN29JJEdsIESfAX7c",
                           sheet = "HILIC_NEG") %>%
    select(compound_name=`Compound name`,
           integrated=`Integrated? (Y/N)`,
           in_samps=`In samples? (Y/N)`) %>%
    mutate(polarity="neg")
  manual_quality <- rbind(manual_pos, manual_neg) %>%
    mutate(compound_name=str_replace(compound_name, "_ ", ", ")) %>%
    filter(!is.na(compound_name)) %>%
    filter(compound_name!="Internal Standards") %>%
  left_join(name_conversion, by=c(compound_name="old")) %>%
  mutate(compound_name=new) %>%
  select(-new)
  write.csv(manual_quality, file = "raw_data/Raafay_HILIC_notes.csv", 
            row.names = FALSE)
}
manual_quality <- read.csv("raw_data/Raafay_HILIC_notes.csv") 
```

SkylineRunner.exe output relating to exporting files:
Exporting reports
ÚÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÂÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄ¿
³--report-name=<name>          ³The name of a report to export as it appears ³
³                              ³in the Skyline Export Report form.           ³
ÃÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÅÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄ´
³--report-file=path/to/file.csv³The path to export the report to. Required if³
³                              ³--report-name is specified.                  ³
ÃÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÅÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄ´
³--report-add=path/to/file.skyr³Add reports from a saved report template file³
³                              ³to the running program.                      ³
ÃÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÅÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄ´
³--report-conflict-resolution= ³Specifies how to resolve report name         ³
³<overwrite | skip>            ³conflicts, by either overwriting or skipping ³
³                              ³them, when using --report-add (default is to ³
³                              ³output an error message for conflicts)       ³
ÃÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÅÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄ´
³--report-format=<csv | tsv>   ³CSV for comma-separated reports (or semicolon³
³                              ³separated, depending on your localization) or³
³                              ³TSV for tab separated reports [default CSV]  ³

```{r Falkor Skyline data extraction}
FK_POS_cmd <- paste(
  "SkylineRunner.exe",
  "--in=raw_data/Falkor/HILIC_QE_POSITIVE_Falkor_1.sky", 
  '--report-name="Ingalls_Lab_QE_Transition Results"',
  '--report-file="raw_data/Falkor/Ingalls_Lab_QE_Transition Results_POS.csv"',
  '--report-format="csv"'
)
system(FK_POS_cmd)

FK_neg_cmd <- paste(
  "SkylineRunner.exe",
  "--in=raw_data/Falkor/HILIC_QE_NEGATIVE_Falkor.sky", 
  '--report-name="Ingalls_Lab_QE_Transition Results"',
  '--report-file="raw_data/Falkor/Ingalls_Lab_QE_Transition Results_NEG.csv"',
  '--report-format="csv"'
)
system(FK_neg_cmd)
```

```{r Mesocenter Skyline data extraction}
MC_POS_cmd <- paste(
  "SkylineRunner.exe",
  "--in=raw_data/Mesocenters/HILIC_QE_POSITIVE_Wei_HRM_DepthProfile.sky", 
  '--report-name="Ingalls_Lab_QE_Transition Results"',
  '--report-file="raw_data/Mesocenters/Ingalls_Lab_QE_Transition Results_POS.csv"',
  '--report-format="csv"'
)
system(MC_POS_cmd)

FK_NEG_cmd <- paste(
  "SkylineRunner.exe",
  "--in=raw_data/Mesocenters/HILIC_QE_NEGATIVE_Wei_HRM_DepthProfile.sky", 
  '--report-name="Ingalls_Lab_QE_Transition Results"',
  '--report-file="raw_data/Mesocenters/Ingalls_Lab_QE_Transition Results_NEG.csv"',
  '--report-format="csv"'
)
system(FK_NEG_cmd)
```

```{r MSDIAL autoexport things not yet}
# Not implemented yet, check on thread below for updates
# http://www.metabolomics-forum.com/index.php?topic=1820.0
# Instead, open an existing .mtd2 file, click the "Export" tab, select "Alignment result", specify the output directory as "Area_HILICPos_EddyTransect.csv", check a couple of the boxes (Raw Data Area is the only critical one), change the output format to msp(?), and hit the "Export" button at the bottom. You then need to open this file in Excel and save as CSV.

# Haven't been able to find the raw data for negative mode so I'm currently just copying this off the Google Drive (G:\Shared drives\Ingalls Lab\Collaborative Projects\MESO-SCOPE\EddyTransect\HILIC_Neg_MSDial)
```

# Falkor things

Start with the Falkor data, since it's nice and neat. Here we're pulling from
the raw_data/Falkor folder, which has the Skyline files, their output, and
will soon have Raafay's notes on peak quality.

These CSV files were obtained from Skyline by exporting the transition results,
with a "_POS" or "_NEG" added to distinguish them.

```{r falkorSkylineImport}
# Read in the data and clean it up a little bit
manual_pos <- read.csv("raw_data/Falkor/Ingalls_Lab_QE_Transition Results_POS.csv") %>%
  mutate(polarity="pos")
manual_neg <- read.csv("raw_data/Falkor/Ingalls_Lab_QE_Transition Results_NEG.csv") %>%
  mutate(polarity="neg")
manual_ints <- rbind(manual_pos, manual_neg) %>%
  select(filename=Replicate.Name, compound_name=Precursor.Ion.Name, 
         area=Area, polarity) %>% 
  mutate(filename = ifelse(str_detect(filename, "DDA"),
                           str_replace(filename, "Poo", "DDA"),
                           filename)) %>%
  mutate(filename = ifelse(str_detect(filename, "QC"),
                           str_replace(filename, "Poo", "QC"),
                           filename)) %>% 
  mutate(compound_name=str_replace(
    compound_name, "Monophosphate, 15N5", "monophosphate, 15N5")) %>%
  left_join(name_conversion, by=c("compound_name"="old")) %>% 
  mutate(compound_name=new) %>% 
  select(-new) %>% 
  left_join(manual_quality, by=c("compound_name", "polarity"))

# Filter for only those peaks that Raafay actually integrated
clean_ints_falkor <- manual_ints %>%
  filter(integrated=="Y"&in_samps%in%c("Y", "Maybe?", "Kinda?", "Barely", "Kinda")) %>%
  select(-integrated, -in_samps) %>% 
  mutate(area=as.numeric(area))
```

```{r falkorBMIS}
pooled <- clean_ints_falkor %>%
  filter(str_detect(filename, "_Poo_"))

pooled_IS <- pooled %>%
  left_join(clean_stans %>% select(compound_name, compound_type) %>% distinct(), by=c("compound_name")) %>% 
  filter(compound_type=="Internal Standard") %>%
  select(-compound_type, -polarity) %>%
  rename_with(~paste0("IS_", .), -filename) %>%
  add_row(filename=unique(.$filename), IS_compound_name="None", IS_area=1)

initial_CVs <- pooled %>%
  mutate(area=ifelse(str_detect(filename, "Half"), area*2, area)) %>%
  group_by(compound_name) %>%
  summarize(init_CV=cv(area))

standardized_CVs <- full_join(pooled, pooled_IS, by=c("filename")) %>%
  group_by(compound_name, IS_compound_name) %>%
  summarize(norm_CV=cv(area/IS_area), .groups="drop")

chosen_BMIS_falkor <- standardized_CVs %>%
  left_join(initial_CVs) %>%
  group_by(compound_name) %>%
  mutate(MIS_improvement=(init_CV - norm_CV)/init_CV) %>% 
  filter(MIS_improvement>min_improvement|IS_compound_name=="None") %>%
  arrange(desc(MIS_improvement)) %>%
  group_by(compound_name) %>%
  slice(1) %>%
  select(compound_name, BMIS=IS_compound_name)

BMISed_falkor <- clean_ints_falkor %>%
  left_join(chosen_BMIS_falkor, by="compound_name") %>%
  left_join(clean_ints_falkor %>% rename(IS_area=area) %>% select(-polarity), 
            by=c(BMIS="compound_name", "filename")) %>%
  group_by(compound_name) %>%
  mutate(BMISed_area=area/IS_area*mean(IS_area)) %>%
  mutate(BMISed_area = ifelse(BMIS=="None", area, BMISed_area)) %>%
  mutate(BMISed_area = ifelse(str_detect(filename, "Half")&BMIS=="None", 
                              BMISed_area*2, BMISed_area)) %>%
  select(-IS_area)
```

```{r falkorRFs}
falkor_rundate <- strptime("190715", "%y%m%d")
falkor_stans <- clean_stans %>%
  filter(date_added<=falkor_rundate)

rfs_falkor <- falkor_stans %>%
  left_join(BMISed_falkor, by=c("compound_name", "polarity")) %>%
  filter(str_detect(filename, "_Std_")) %>%
  filter(compound_type!="Internal Standard") %>%
  select(-compound_type) %>%
  mutate(mix=paste0("H2OinMatrix|", paste0(mix, "InMatrix"), "|", paste0(mix, "InH2O"))) %>%
  filter(str_detect(filename, mix)) %>%
  mutate(type=str_replace(str_extract(filename, mix), "Mix[12]", "Mix")) %>%
  select(-mix) %>%
  group_by(compound_name, type) %>%
  summarize(mean_area=mean(area, na.rm = TRUE), conc_um=unique(conc_um)) %>%
  pivot_wider(names_from = type, values_from=mean_area) %>%
  mutate(rf_h2o=MixInH2O/conc_um) %>%
  mutate(rf_matrix=(MixInMatrix - H2OinMatrix)/conc_um) %>%
  mutate(rf_ratio=rf_matrix/rf_h2o)

# 400uL vial, 10L sample, diluted x2
falkor_data <- rfs_falkor %>% 
  left_join(BMISed_falkor, by="compound_name") %>%
  left_join(clean_stans) %>%
  mutate(uM = BMISed_area/rf_h2o * 0.0004/10 * 1/rf_ratio * 2) %>% 
  mutate(nM=uM*1000) %>%
  select(compound_name, nM, filename, polarity)
```

# Mesotransect things

Continue with the MESOSCOPE transect data. Here again, we use Raafay's notes
on peak quality to remove the gross peaks.

```{r transectSkylineImport}
manual_pos <- read_csv("raw_data/Mesotransect/Area_HILICPos_EddyTransect.csv", skip = 4) %>%
  filter(`Metabolite name`!="Unknown") %>%
  select(compound_name=`Metabolite name`, starts_with("180821")) %>%
  pivot_longer(cols=starts_with("180821"), names_to = "filename", values_to="area") %>%
  mutate(polarity="pos")
manual_neg <- read_csv("raw_data/Mesotransect/Area_HILICNeg_EddyTransect.csv", skip = 4) %>%
  filter(`Metabolite name`!="Unknown") %>%
  select(compound_name=`Metabolite name`, starts_with("180821")) %>%
  pivot_longer(cols=starts_with("180821"), names_to = "filename", values_to="area") %>%
  mutate(polarity="neg")
manual_ints <- rbind(manual_pos, manual_neg) %>%
  mutate(compound_name=str_replace_all(compound_name, "Ingalls_|w/o MS2:|Unsettled: ", "")) %>%
  mutate(filename = ifelse(str_detect(filename, "DDA"),
                           str_replace(filename, "Poo", "DDA"),
                           filename)) %>%
  mutate(filename = ifelse(str_detect(filename, "QC"),
                           str_replace(filename, "Poo", "QC"),
                           filename))

# Once again, filter only those that Raafay integrated
clean_ints_transect <- manual_ints %>%
  filter(!str_detect(filename, "DDA")) %>%
  filter(compound_name!="Methyl 3-methylthiopropionate") %>%
  mutate(compound_name=str_replace(compound_name, "Monophosphate, 15N5", "monophosphate, 15N5")) %>%
  left_join(name_conversion, by=c("compound_name"="old")) %>% 
  mutate(compound_name=new) %>% 
  select(-new) %>% 
  left_join(manual_quality, by=c("compound_name", "polarity")) %>%
  filter(integrated=="Y" &in_samps%in%c("Y", "Maybe?", "Kinda?", "Barely", "Kinda")) %>%
  select(-integrated, -in_samps) %>%
  mutate(area=as.numeric(area)) %>%
  anti_join(distinct(., compound_name, polarity) %>% filter(duplicated(compound_name)))

```

```{r transectBMIS}
pooled <- clean_ints_transect %>%
  filter(str_detect(filename, "_Poo_"))

pooled_IS <- pooled %>%
  left_join(clean_stans %>% select(compound_name, compound_type) %>% distinct(), by=c("compound_name")) %>% 
  filter(compound_type=="Internal Standard") %>%
  select(-compound_type, -polarity) %>%
  rename_with(~paste0("IS_", .), -filename) %>%
  add_row(filename=unique(.$filename), IS_compound_name="None", IS_area=1)

initial_CVs <- pooled %>%
  mutate(area=ifelse(str_detect(filename, "Half"), area*2, area)) %>%
  group_by(compound_name) %>%
  summarize(init_CV=cv(area))

standardized_CVs <- full_join(pooled, pooled_IS, by=c("filename")) %>%
  group_by(compound_name, IS_compound_name) %>%
  summarize(norm_CV=cv(area/IS_area), .groups="drop")

chosen_BMIS_transect <- standardized_CVs %>%
  left_join(initial_CVs) %>%
  group_by(compound_name) %>%
  mutate(MIS_improvement=(init_CV - norm_CV)/init_CV) %>% 
  filter(MIS_improvement>min_improvement|IS_compound_name=="None") %>%
  arrange(desc(MIS_improvement)) %>% 
  group_by(compound_name) %>% 
  slice(1) %>%
  select(compound_name, BMIS=IS_compound_name)

BMISed_transect <- clean_ints_transect %>% 
  left_join(chosen_BMIS_transect, by=c("compound_name")) %>%
  left_join(clean_ints_transect %>% rename(IS_area=area) %>% select(-polarity), 
            by=c(BMIS="compound_name", "filename")) %>%
  group_by(compound_name) %>%
  mutate(BMISed_area=area/IS_area*mean(IS_area)) %>%
  mutate(BMISed_area = ifelse(BMIS=="None", area, BMISed_area)) %>%
  mutate(BMISed_area = ifelse(str_detect(filename, "Half")&BMIS=="None", 
                              BMISed_area*2, BMISed_area)) %>%
  select(-IS_area)
```

```{r transectRFs}
transect_rundate <- strptime("180821", "%y%m%d")
transect_stans <- clean_stans %>%
  filter(date_added<=transect_rundate)

rfs_transect <- transect_stans %>%
  left_join(clean_ints_transect, by="compound_name") %>%
  filter(str_detect(filename, "_Std_")) %>% 
  # filter(compound_name=="N6-Acetyl-L-lysine") %>%
  # filter(compound_name=="Dimethylsulfonioacetate") %>%
  # ggplot() + geom_col(aes(x=area,y=filename))
  filter(compound_type!="Internal Standard") %>%
  select(-compound_type) %>%
  mutate(mix=paste0("H2OInMatrix|", paste0(mix, "InMatrix"), "|", paste0(mix, "InH2O"))) %>%
  filter(str_detect(filename, mix)) %>%
  mutate(type=str_replace(str_extract(filename, mix), "Mix[12]", "Mix")) %>%
  select(-mix) %>% 
  group_by(compound_name, type) %>%
  summarise(mean_area=mean(area, na.rm = TRUE), conc_um=unique(conc_um)) %>%
  pivot_wider(names_from = type, values_from=mean_area) %>% 
  mutate(rf_h2o=MixInH2O/conc_um) %>%
  mutate(rf_matrix=(MixInMatrix - H2OInMatrix)/conc_um) %>%
  mutate(rf_ratio=rf_matrix/rf_h2o)

mesotransect_data <- rfs_transect %>% 
  left_join(BMISed_transect, by="compound_name") %>%
  left_join(clean_stans) %>%
  mutate(uM = BMISed_area/rf_h2o * 0.0004/10 * 1/rf_ratio * 2) %>% # 400uL vial, 10L sample, diluted 2x
  mutate(nM=uM*1000) %>%
  select(compound_name, nM, filename, polarity)
```

# Mesocenter things

```{r centerSkylineImport}
manual_pos <- read.csv("raw_data/Mesocenters/Ingalls_Lab_QE_Transition Results_POS.csv") %>%
  mutate(polarity="pos")
manual_neg <- read.csv("raw_data/Mesocenters/Ingalls_Lab_QE_Transition Results_NEG.csv") %>%
  mutate(polarity="neg")
manual_ints <- rbind(manual_pos, manual_neg) %>%
  select(filename=Replicate.Name, compound_name=Precursor.Ion.Name, 
         area=Area, polarity) %>%
  mutate(filename = ifelse(str_detect(filename, "DDA"),
                           str_replace(filename, "Poo", "DDA"),
                           filename)) %>%
  mutate(filename = ifelse(str_detect(filename, "QC"),
                           str_replace(filename, "Poo", "QC"),
                           filename)) %>%
  mutate(compound_name=str_replace(
    compound_name, "Monophosphate, 15N5", "monophosphate, 15N5")) %>%
  left_join(name_conversion, by=c("compound_name"="old")) %>% 
  mutate(compound_name=new) %>% 
  select(-new) %>% 
  left_join(manual_quality, by=c("compound_name", "polarity"))

clean_ints_center <- manual_ints %>%
  filter(integrated=="Y"&in_samps%in%c("Y", "Maybe?", "Kinda?", "Barely", "Kinda")) %>%
  select(-integrated, -in_samps) %>% 
  mutate(area=as.numeric(area))
```

```{r centerBMIS}
pooled <- clean_ints_center %>%
  filter(str_detect(filename, "_Poo_"))

pooled_IS <- pooled %>%
  left_join(clean_stans %>% select(compound_name, compound_type) %>% 
              distinct(), by=c("compound_name")) %>% 
  filter(compound_type=="Internal Standard") %>%
  select(-compound_type, -polarity) %>%
  rename_with(~paste0("IS_", .), -filename) %>%
  add_row(filename=unique(.$filename), IS_compound_name="None", IS_area=1)

initial_CVs <- pooled %>%
  mutate(area=ifelse(str_detect(filename, "Half"), area*2, area)) %>%
  group_by(compound_name) %>%
  summarize(init_CV=cv(area))

standardized_CVs <- full_join(pooled, pooled_IS, by=c("filename")) %>%
  group_by(compound_name, IS_compound_name) %>%
  summarize(norm_CV=cv(area/IS_area), .groups="drop")

chosen_BMIS_center <- standardized_CVs %>%
  left_join(initial_CVs) %>%
  group_by(compound_name) %>%
  mutate(MIS_improvement=(init_CV - norm_CV)/init_CV) %>% 
  filter(MIS_improvement>min_improvement|IS_compound_name=="None") %>%
  arrange(desc(MIS_improvement)) %>%
  group_by(compound_name) %>%
  slice(1) %>%
  select(compound_name, BMIS=IS_compound_name)

BMISed_center <- clean_ints_center %>% 
  left_join(chosen_BMIS_center, by="compound_name") %>%
  left_join(clean_ints_center %>% rename(IS_area=area) %>% select(-polarity), 
            by=c(BMIS="compound_name", "filename")) %>%
  group_by(compound_name) %>%
  mutate(BMISed_area=area/IS_area*mean(IS_area)) %>%
  mutate(BMISed_area = ifelse(BMIS=="None", area, BMISed_area)) %>%
  mutate(BMISed_area = ifelse(str_detect(filename, "Half")&BMIS=="None", 
                              BMISed_area*2, BMISed_area)) %>%
  select(-IS_area)
```

The RFs for the Mesocenter data are a little weird because they weren't run
with H2OinMatrix samples. The good news is that in theory, the RF ratio for
every compound is the same so we can use the RF ratios from Falkor to normalize
these samples instead.

```{r centerRFs}
# Slightly weird because Mesocenter data doesn't have H2OinMatrix samples
# Instead, use rf_ratio from Falkor to swap in
center_rundate <- strptime("180205", "%y%m%d")
center_stans <- clean_stans %>%
  filter(date_added<=center_rundate)

rfs_center <- center_stans %>% 
  left_join(clean_ints_center, by="compound_name") %>% 
  filter(str_detect(filename, "_Std_")) %>%
  filter(compound_type!="Internal Standard") %>%
  select(-compound_type) %>% 
  mutate(mix="InMatrix|InH2O") %>% 
  filter(str_detect(filename, mix)) %>%
  mutate(type=str_extract(filename, mix)) %>%
  select(-mix) %>% 
  group_by(compound_name, type) %>%
  summarize(mean_area=mean(area, na.rm = TRUE), conc_um=unique(conc_um)) %>%
  pivot_wider(names_from = type, values_from=mean_area) %>%
  mutate(rf_h2o=InH2O/conc_um) %>%
  left_join(rfs_falkor %>% select("rf_ratio", "compound_name"), by="compound_name")

mesocenter_data <- rfs_center %>% 
  left_join(BMISed_center, by="compound_name") %>% 
  left_join(clean_stans) %>% 
  mutate(uM = BMISed_area/rf_h2o * 0.0004/10 * 1/rf_ratio * 2) %>% 
  mutate(nM=uM*1000) %>%
  select(compound_name, nM, filename, polarity)
```

# Combine and write out

I'm interested in exporting not only final concentrations, but also the raw
areas and BMIS assignments for debugging purposes. These all go into the
made_data folder.

```{r collect}
## Write out just the raw areas
raw_area_data <- rbind(
  cbind(clean_ints_falkor, cruise="Falkor"),
  cbind(clean_ints_center, cruise="Mesocenter"),
  cbind(clean_ints_transect, cruise="Mesotransect")
) %>%
  left_join(clean_stans, by="compound_name")
write.csv(raw_area_data, file = "made_data/raw_area_data.csv", row.names = FALSE)

## Write out BMISed areas
BMISed_areas <- rbind(
  cbind(BMISed_falkor, cruise="Falkor"),
  cbind(BMISed_center, cruise="Mesocenter"),
  cbind(BMISed_transect, cruise="Mesotransect")
)
write.csv(BMISed_areas, file = "made_data/BMIS_areas.csv", row.names = FALSE)

## Write out RF ratios
rfrs <- rbind(cbind(rfs_falkor, cruise="Falkor"),
              cbind(rfs_transect, cruise="Mesotransect"))
write.csv(rfrs, "made_data/rf_ratios.csv", row.names=FALSE)

## Write out quant data
quant_data <- bind_rows(falkor_data, mesotransect_data, mesocenter_data) %>%
  filter(compound_name!="Choline")
write.csv(quant_data, file = "made_data/longdata.csv", row.names = FALSE)
```

# Perform some quality checks

```{r}
# Fine to have separate H2OinMatrix cols, different filenames between each
rfrs <- read_csv("made_data/rf_ratios.csv") %>%
  mutate(H2OinMatrix=ifelse(is.na(H2OinMatrix), H2OInMatrix, H2OinMatrix)) %>%
  select(-H2OInMatrix) %>%
  select(compound_name, rf_ratio, cruise) %>%
  pivot_wider(names_from = cruise, values_from = rf_ratio)

with(rfrs, cor(Falkor, Mesotransect, use="pair"))

ggplot(rfrs, aes(x=Falkor, y=Mesotransect, label=compound_name)) +
  geom_point()
plotly::ggplotly()
rfrs %>%
  filter(compound_name!="Choline") %>%
  ggplot(aes(x=Falkor, y=Mesotransect, label=compound_name)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0) +
  lims(x=c(0, NA), y=c(0, NA)) +
  coord_fixed()
```

### 2022-10-05 untargeted/output_control.Rmd
Now tracking untargeted things

---
title: "Control Script for SCOPE data"
output: html_document
---

```{r setup, include=FALSE, purl=TRUE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, error = TRUE, purl = FALSE)
options(dplyr.summarise.inform=F)
options(pillar.sigfig=10)
suppressPackageStartupMessages(library(tidyverse))
library(data.table)
library(pbapply)
library(RaMS)

polarity <- "neg"
cruise <- "MC"
output_folder <- paste0(c(cruise, polarity, "output/"), collapse = "_")
mzml_path <- paste0("../mzMLs/", polarity, "/")

source("scripts/functions.R")

cruise_metadata <- read.csv("../metadata/made_data/file_metadata.csv") %>%
  filter(cruise==!!cruise)

run_date <- cruise_metadata %>%
  pull(date_run) %>%
  unique()

ms_files <- cruise_metadata$filename

BiocParallel::register(BPPARAM = BiocParallel::SnowParam(tasks = length(ms_files), progressbar = TRUE))
```

## Peakpicking

**Obviously** the most important step, you can't do anything else without first finding things to look at and talk about. The script below, when sourced, runs XCMS's centWave peakpicking. It also calculates an improved signal-to-noise and metric of Gaussian-ness that seems to sift through the noise more accurately than the default signal-to-noise ratio, which has known bugs (https://doi.org/10.1021/acs.analchem.7b01069). Finally, it performs retention time correction and peak correspondence (grouping peaks across files).

It takes in the metadataframe created above as well as the list of paths to the .mzML files and returns a (rather large) data frame full of peaks.

```{r peakpicking, error=TRUE, purl=FALSE}
library(xcms)

# Create folder if it doesn't already exist
if(!dir.exists(output_folder)){dir.create(output_folder)}

# Define the peakpicking parameters
cwp <- CentWaveParam(ppm = 2.5, peakwidth = c(15, 45), 
                     snthresh = 1, prefilter = c(5, 10000), 
                     integrate = 2, mzCenterFun = "wMean", 
                     mzdiff = 0.001, fitgauss = FALSE, 
                     noise = 5000, firstBaselineCheck = FALSE, 
                     extendLengthMSW = TRUE)

# Set the new quality threshold
qscore_threshold <- 30

# Define the retention time correction parameters
obp <- ObiwarpParam(binSize = 0.1, centerSample = length(ms_files)/2, 
                    response = 1, distFun = "cor_opt")

# Define the correspondence parameters
pdp <- PeakDensityParam(sampleGroups = cruise_metadata$samp_type, 
                        bw = 12, minFraction = 0.1, 
                        binSize = 0.001, minSamples = 2)

# Make sure that filled peaks have at least a 2.5ppm window from mzmin to mzmax around mz
fpp <- FillChromPeaksParam(ppm = 2.5)

# Perform peakpicking
source("scripts/peakpicking.R")
unique(warnings())

# Fill in missing data - filename/feature combinations with no entry
# First, choose largest peak from those with multiples per file
# Second, remove all peaks that show up in less than 50% of the *samples*
# Fill in mz with the average for each feature
# Fill in retention times using samples with nearest timestamp
n_samps <- length(str_subset(unique(cruise_metadata$filename), "Smp"))
good_feats <- raw_peaks %>%
  left_join(cruise_metadata) %>%
  filter(samp_type=="Smp") %>%
  count(feature) %>%
  filter(n>floor(n_samps*0.5)) %>%
  distinct(feature)
filled_peaks <- good_feats %>%
  left_join(raw_peaks) %>%
  arrange(desc(into)) %>%
  group_by(feature, filename) %>%
  slice(1) %>%
  ungroup() %>%
  complete(feature, filename) %>%
  left_join(cruise_metadata) %>%
  arrange(timestamp) %>%
  group_by(feature) %>%
  mutate(mz=ifelse(is.na(mz), mean(mz, na.rm=TRUE), mz)) %>%
  fill(starts_with("rt"), .direction = "downup") %>%
  arrange(feature) %>%
  select(feature, filename, mz, rt, rtmin, rtmax)

# Save intermediate results
saveRDS(xdata, file = paste0(output_folder, "xdata.rds"))
saveRDS(files_qscores, file = paste0(output_folder, "files_qscores.rds"))
saveRDS(xdata_cleanpeak, file = paste0(output_folder, "xdata_cleanpeak.rds"))
saveRDS(xdata_rt, file = paste0(output_folder, "xdata_rt.rds"))
saveRDS(xdata_cor, file = paste0(output_folder, "xdata_cor.rds"))
saveRDS(xdata_filled, file = paste0(output_folder, "xdata_filled.rds"))
write.csv(raw_peaks, file = paste0(output_folder, "raw_peaks.csv"), row.names = FALSE)
write.csv(filled_peaks, file = paste0(output_folder, "filled_peaks.csv"), row.names = FALSE)
unique(warnings())
```

## De-isotoping and de-adducting

After peaks have been identified, many of them will be isotopes and adducts of other peaks. Removing these is important to minimize pseudoreplication issues and data artifacts. I didn't like any of the packages I tried to do this, so I wrote my own code to process this more robustly.

This script does two things. First, it identifies peaks that are likely isotopes or adducts of other peaks in the data set. It does this by assuming that every single peak is an isotope/adduct of another peak. If a peak is an adduct/isotope, there will be another peak that looks very similar but is separated by a very specific mass difference. For example, if we find a peak at 140.06875, we "guess" that it's an adduct or isotope and check all the places where the M+H would be found, depending on the isotope or adduct. If we suspect it of being a 13C isotope of another peak, we'd look for a peak at 140.06875-1.003355. If we suspect it of being a sodium isotope, we'd look for a peak at 140.06875-22.99787+1.007276. In this case, we would indeed find a signal at the M+H peak if we assume sodium, but no peak at the M+H if we assume a 13C isotope - allowing us to conclude tentatively that this is actually an adduct.

However, simply finding data at the expected mass isn't specific enough because we'll often stumble upon noise or a different peak in the general area. Thus, we check for peak similarity before assuming that there's real signal there. The general theory is that isotopes and adducts will be similar to the base peak in two ways. First, the individual x/y (rt/intensity) data points should match up almost exactly, as adducts and isotopes elute at the same time as the base peak. We can check this with a Pearson correlation - simply, how nicely do the two peaks correlate with each other? This method is quite sensitive to even small differences in retention time. Second, we can compare peak area ratios across files. Here, we run a similar correlation but check x/y as base peak area/isotope area. This works because a compound will always create similarly stable adducts (some will love M+H, others M+Na) and will always have a fixed isotope abundance. Across multiple files, the correlation between these should be very strong (and is, in fact, usually stronger than the first method).

Deciding on the thresholds to use as cutoffs for "yes, this is an adduct/isotope" is complicated. If set too high, not all adducts and isotopes will be flagged and removed - spuriously increasing the impact of a compound that has an addiso that snuck through. If set too low, compounds that aren't adducts or isotopes will be erroneously removed. Ideally, there's a clear separation between the spurious peaks and the ones that are good - spurious all hovering around zero and good all close to 1. However, it seems like there's a lot more overlap and so there's two things we can do. First, plotting histograms of peak area and shape correlations can provide some clue as to where the valley in between the two is and that's a good starting spot. For a more intensive analysis we can use our standards to calibrate the thresholds because we can predict the isotope envelope of each standard and know what isotopes to expect and which ones shouldn't show up at all.

```{r deisoadduct, error=TRUE, purl=FALSE}
library(xcms)
xdata_filled <- readRDS(file = paste0(output_folder, "xdata_filled.rds"))
filled_peaks <- read.csv(file = paste0(output_folder, "filled_peaks.csv"))

# How similar in shape do the original and addiso peaks need to be?
shape_remove_threshold <- 0.8
# How good does the M area ~ adduct area correlation across files need to be to assume adduct?
area_remove_threshold <- 0.95

# Which adducts and isotopes should we look for?
addiso_masses_all <- tribble(
  ~adduct_or_iso, ~polarity, ~addiso, ~nat_abund, ~mz_diff,
  "iso",          "both",    "C13",   0.0110,     1.003355, 
  "iso",          "both",    "X2C13", 0.0110,     2*1.003355, 
  "iso",          "both",    "S34",   0.0421,     1.995796, 
  "iso",          "both",    "S33",   0.0075,     0.999387,
  "iso",          "both",    "N15",   0.0037,     0.997035, 
  "iso",          "both",    "O18",   0.0020,     2.004244,
  "iso",          "both",    "Cl37",  0.2423,     1.99705,
  "iso",          "both",    "Br81",  0.4931,     1.997954,
  "adduct",       "both",    "M-H2O", NA,         -18.0106, 
  "adduct",       "pos",     "M+Na",  NA,          22.98922-1.007276 # Goes from [M+H] to [M+Na]
  # "adduct",       "pos",     "M+NH4", NA,          18.0338-1.007276,
  # "adduct",       "pos",     "M+K",   NA,          38.963708-1.007276,
  # "adduct",       "neg",     "M+Cl",  NA,          34.969402+1.007276,
  # "adduct",       "neg",     "M+Ac",  NA,          59.013851+1.007276,
  # "adduct",       "neg",     "M+Br",  NA,          78.918885+1.007276
)
addiso_masses <- addiso_masses_all %>% filter(polarity%in%c(!!polarity, "both"))

# Make sure certain known compounds aren't removed spuriously
not_addisos <- tribble(
  ~name,           ~mz,        ~addiso,
  "Glutamine",     147.076968, "M+NH4",
  "Citrulline",    176.103517, "M+NH4",
  "Guanine",       152.0567,   "M+NH4",
  "Glutamic acid", 148.061,    "M+NH4",
  "4-Aminobutyric acid", 104.071, "M+NH4",
  "Adenine",       136.0618,   "M+NH4",
) %>%
  pmap_dfr(function(...){
    row_data <- data.frame(...)
    filled_peaks %>%
      filter(mz%between%pmppm(row_data$mz, 5)) %>%
      mutate(name=row_data$name, addiso=row_data$addiso)
  }) %>% distinct(feature, name, addiso)

# Run the script
source("scripts/deisoadduct.R")
unique(warnings())

# Some summary stats
nrow(has_addiso)
nrow(is_addiso)
sort(table(has_addiso$addiso), decreasing = TRUE)

# Save intermediate files
saveRDS(isoadduct_df, file = paste0(output_folder, "isoadduct_df.rds"))

# Write out addiso information
write.csv(is_addiso, file = paste0(output_folder, "is_addiso.csv"), row.names = FALSE)
write.csv(has_addiso, file = paste0(output_folder, "has_addiso.csv"), row.names = FALSE)
write.csv(M_areas, file = paste0(output_folder, "M_areas.csv"), row.names = FALSE)
# Contains feature x sample x addiso data frame
write.csv(feature_envelopes, file = paste0(output_folder, "feature_envelopes.csv"), row.names = FALSE)
# Update filled_peaks with reintegrated areas
M_area_peaks <- filled_peaks %>% left_join(M_areas)
write.csv(M_area_peaks, paste0(output_folder, "M_area_peaks.csv"), row.names = FALSE)

unique(warnings())
```

```{r addiso check}
given_file <- str_subset(unique(filled_peaks$filename), "Poo_Full")
file_data_table <- as.data.table(filterFile(xdata_filled, given_file))[,c("rt", "mz", "i")]
file_peaks <- filled_peaks %>% filter(filename%in%given_file)

# test_peak <- file_peaks %>% filter(feature=="FT0207")
test_peak <- file_peaks %>% filter(feature=="FT0702")
feature_i <- test_peak$feature
mz_i <- test_peak$mz
rtmin_i <- test_peak$rtmin
rtmax_i <- test_peak$rtmax
split_table_i <- file_data_table

addiso_masses %>%
  {rbind(cbind(., type="self", dir=-1), cbind(., type="seek", dir=1))} %>%
  mutate(mz_diff=mz_diff*dir) %>%
  select(-dir) %>%
  add_row(polarity="both", addiso="M+/-H", mz_diff=0, type="self") %>%
  group_by(addiso, type) %>%
  group_split() %>% 
  lapply(function(row_data){
    v <- split_table_i[
      mz%between%pmppm(row_data$mz_diff+mean(mz_i), 5)][
        rt%between%c(mean(rtmin_i), mean(rtmax_i))]
    v$addiso <- row_data$addiso
    v$type <- row_data$type
    v
  }) %>%
  rbindlist() %>%
  filter(!is.na(i)) %>%
  group_by(addiso, type) %>%
  filter(n()>=5) %>%
  ggplot(aes(x=rt, y=i)) +
  geom_path() +
  geom_point() +
  facet_grid(addiso~type, scales = "free_y")
```

## Noise audit peakpicking?

```{r peakpicking noise audit, eval=FALSE}
filled_peaks <- read.csv(paste0(output_folder, "filled_peaks.csv"))

filled_peaks %>%
  group_by(feature) %>%
  summarise(mean_mz=mean(mz), mean_rt=mean(rt)/60, med_sn=median(sn,na.rm = TRUE)) %>%
  filter(!mean_rt%between%c(14.5, 14.7)) %>%
  filter(mean_rt>1) %>%
  arrange(desc(med_sn))

mini_mzml_names <- paste(output_folder, ms_files)
set.seed(123)
feats_to_check <- slice_sample(picked_feats, n = 100)
```

```{r check peakpicking}
filled_peaks <- read.csv(paste0(output_folder, "filled_peaks.csv"))

picked_feats <- filled_peaks %>%
  filter(str_detect(filename, "Smp")) %>%
  group_by(feature) %>%
  summarise(mean_mz=mean(mz), mean_rt=mean(rt)/60)

picked_feats %>%
  ggplot() + 
  geom_point(aes(x=mean_rt, y=mean_mz)) +
  xlim(0, NA)

# msdata <- grabMSdata(list.files("../tmzMLs/pos", pattern = "180821", full.names = TRUE))
# test_data <- slice_sample(picked_feats, n = 1)
# test_data <- picked_feats %>% filter(feature=="FT0179")
# bet_data <- msdata$MS1[mz%between%pmppm(test_data$mean_mz, 5)]
# bet_data %>%
#   filter(rt%between%c(8, 13)) %>%
#   mutate(filename=str_replace(filename, "tmzML", "mzML")) %>%
#   left_join(cruise_metadata) %>%
#   ggplot() +
#   geom_line(aes(x=rt, y=int, group=filename, color=samp_type)) +
#   geom_vline(xintercept = test_data$mean_rt) +
#   theme_bw() +
#   ggtitle(do.call(what = paste, test_data))
```

## Annotate standards

This block runs a custom function that tries to resolve discrepancies between various metrics using basically a bunch of if/else statements.
  - Highest priority: isotope validated (isotope_choice)
  - Next highest: if two peaks are found and two standards should be found, use RT ordering to match them up (match_choice)
  - Next highest: if peaks are dramatically larger in the correct mix and smaller in the other mix, assume it's the larger one (mix_choice) sometimes produces multiple, thus later area and RT matching
  - Next highest: if one peak is larger than the others, assume it's the
              standard (area_choice)
  - Lowest: RT matching based on the number in the standards list

For each standard, grab all the features that it could be based on 5ppm m/z window. If there aren't any features, return NA for everything. If there are features, calculate 5 metrics for each one:
Isotope choice:
  If the compound has an isotope, use the peak closest to the isotope peak in RT
    (i.e. another compound exists with the same name + C13, N15, or O18)
  If the compound IS an isotope (i.e. has C13, N15, or O18 in the name)
    find the two peaks closest in RT and assume those are the isotopologues
  Otherwise, NA
Match choice:
  If there are two standards expected in a mix and two features found,
  assume that the two found peaks are the standards and use rank-order RT matching
  If more or fewer peaks found, NA
Mix choice:
  Generally, which peaks are bigger in the mix they've been added to?
  Calculate average z-statistic between the mixes after controlling for H2O vs Matrix
  Return all feature numbers with a z-score above 10 (arbitrarily chosen)
Area choice:
  Which peak has the largest area?
RT choice:
  Which peak is closest in RT to the value in the standards list?

```{r targeted, purl=TRUE}
M_area_peaks <- read.csv(paste0(output_folder, "M_area_peaks.csv"))
given_stans <- read.csv("../metadata/made_data/clean_stans.csv") %>% 
  filter(polarity==!!polarity) %>%
  filter(run_date[1]>=date_added)

feature_data <- M_area_peaks %>%
  group_by(feature) %>%
  summarise(mzmed=median(mz), rtmed=round(median(rt)), avgarea=round(mean(M_area)))

source("scripts/standard_assignments.R")

write.csv(stan_annotations, paste0(output_folder, "stan_annotations.csv"), 
          row.names = FALSE)
```

```{r useful dedup code}
library(xcms)
xdata_filled <- readRDS(file = paste0(output_folder, "xdata_filled.rds"))
pooled_msdata <- str_subset(ms_files, "Std") %>%
  lapply(function(filename){
    filterFile(xdata_filled, filename) %>%
      as.data.frame() %>%
      mutate(filename=filename)
  }) %>%
  rbindlist()

mz_i <- given_stans %>%
  filter(compound_name=="Trehalose") %>%
  pull(mz)
# mz_i <- 341.1084

# Plot chromatogram
peak_bounds <- M_area_peaks %>% 
  filter(mz%between%pmppm(mz_i, 10)) %>%
  filter(str_detect(filename, "Std"))
pooled_msdata[mz%between%pmppm(mz_i, 20)] %>%
  ggplot() + geom_line(aes(x=rt, y =i, group=filename)) +
  # facet_wrap(~filename) +
  facet_wrap(~filename, ncol=2) +
  theme_bw() +
  geom_vline(aes(xintercept=rtmin), color="green", data = peak_bounds) +
  geom_vline(aes(xintercept=rtmax), color="red", data = peak_bounds)
M_area_peaks %>% 
  filter(mz%between%pmppm(mz_i, 10)) %>% 
  group_by(feature) %>%
  summarize(mzmed=median(mz), rtmed=median(rt), 
            areamed=median(M_area, na.rm=TRUE),
            minrt=median(rtmin), maxrt=median(rtmax)) %>%
  arrange(rtmed)

# Plotting peak areas by file type
M_area_peaks %>%
  filter(mz%between%pmppm(mz_i)) %>%
  mutate(fillcol=str_extract(filename, "Mix1|Mix2|Poo|Smp|Blk|H2O|Stds")) %>%
  mutate(fillcol=factor(fillcol, levels=unique(fillcol))) %>%
  mutate(feature=factor(feature, levels=unique(peak_bounds$feature))) %>%
  ggplot() +
  geom_col(aes(x=filename, y=M_area, fill=fillcol)) +
  theme(axis.text.x = element_text(angle=90)) +
  facet_wrap(~feature, ncol=1, scales = "free_y")

# Grab possible stans that should have been annotated
given_stans %>%
  filter(mz%between%pmppm(mz_i, 20)) %>%
  arrange(rt)
```

```{r how many annod}
cat(paste0("Annotated: ", sum(!is.na(stan_annotations$feature)),
           "\nUnannotated: ", sum(is.na(stan_annotations$feature))))
stan_annotations %>% filter(is.na(stan_annotations$feature))
```

## Finding B-MISs

```{r B-MIS, purl=TRUE}
M_area_peaks <- read.csv(paste0(output_folder, "M_area_peaks.csv"))
stan_annotations <- read.csv(paste0(output_folder, "stan_annotations.csv"))
given_stans <- read.csv("../metadata/made_data/clean_stans.csv")
is_peaks <- read.csv(paste0("IS_integrations/all_is.csv")) %>%
  filter(cruise==!!cruise) %>%
  mutate(filename=paste0(Replicate.Name, ".mzML")) %>%
  left_join(given_stans, by=c("Precursor.Ion.Name"="compound_name_old")) %>%
  select(filename, compound_name, M_area=Area) %>%
  mutate(M_area=as.numeric(M_area)) %>%
  mutate(M_area=ifelse(is.na(M_area), 0, M_area)) %>%
  group_by(compound_name) %>%
  mutate(M_area=ifelse(M_area==0, mean(M_area), M_area)) %>%
  ungroup()

min_improvement <- 0.2
already_good <- 0.1 #Not currently used

source("scripts/bmisscript.R")

table(chosen_BMIS$BMIS)
table(chosen_BMIS$BMIS!="None")

write.csv(is_peaks, file = paste0(output_folder, "is_peaks.csv"), 
          row.names = FALSE)
write.csv(chosen_BMIS, file = paste0(output_folder, "chosen_BMIS.csv"), 
          row.names = FALSE)
```

```{r bmischeck}
is_peaks <- read.csv(paste0(output_folder, "is_peaks.csv"))

# Create plot of absolute IS areas
facet_labels <- paste(unique(is_peaks$compound_name), sep=": ")
names(facet_labels) <- unique(is_peaks$compound_name)
IS_areas_gp <- is_peaks %>%
  mutate(type=str_extract(filename, "Blk|Poo|Smp|Std")) %>%
  mutate(xax=str_extract(filename, "(?<=_.{3}_).*(?=\\.mzML)")) %>%
  # mutate(xax=paste0(str_extract(filename, "^\\d+"), "_", xax)) %>%
  mutate(xax=factor(xax, levels = unique(xax))) %>%
  ggplot() +
  geom_bar(aes(x=xax, y=M_area, fill=type), stat = "identity") +
  theme(axis.text.x = element_text(angle = 90, hjust=1, vjust=0.5)) +
  facet_wrap(~compound_name, ncol = 1, scales = "free_y",
             labeller = as_labeller(facet_labels))

IS_areas_gp

ggsave(plot = IS_areas_gp, filename = paste0(output_folder, "IS_areas.pdf"), 
       device = "pdf", width = 10, height = 10, dpi = 144)

# Create BMIS plot for IS with pooled samps vs all samps
IS_pooled_cvs <- is_peaks %>% 
  filter(str_detect(filename, "_Poo_")) %>%
  select(compound_name, M_area, filename) %>%
  {full_join(., ., by=c("filename"))} %>%
  group_by(compound_name.x, compound_name.y) %>%
  mutate(M_area_normed=(M_area.x/M_area.y)*mean(M_area.y)) %>%
  summarise(calc_cv=cv(M_area_normed)) %>%
  mutate(type="pooled_only")
IS_all_cvs <- is_peaks %>% 
  filter(!str_detect(filename, "Blk")) %>%
  select(compound_name, M_area, filename) %>%
  filter(M_area>0) %>%
  {full_join(., ., by=c("filename"))} %>%
  group_by(compound_name.x, compound_name.y) %>%
  mutate(M_area_normed=(M_area.x/M_area.y)*mean(M_area.y)) %>%
  summarise(calc_cv=cv(M_area_normed)) %>%
  mutate(type="all_files")
BMIS_on_IS_gp <- rbind(IS_pooled_cvs, IS_all_cvs) %>%
  pivot_wider(values_from = calc_cv, names_from = "type") %>%
  ggplot() +
  geom_point(aes(x=pooled_only, y=all_files, label=compound_name.y)) +
  facet_wrap(~compound_name.x, scales = "free") +
  geom_hline(yintercept = 1) +
  geom_vline(xintercept = 1) +
  geom_abline(slope = 1, intercept = 0)
BMIS_on_IS_gp
ggsave(plot=BMIS_on_IS_gp, filename = paste0(output_folder, "BMIS_on_IS.pdf"), 
       device = "pdf", width = 8.5, height = 8.5)

# plotly::ggplotly(IS_areas_gp)
# plotly::ggplotly(BMIS_on_IS_gp)
```

## Annotate formulae

```{r assignformulae, purl=FALSE}
M_area_peaks <- read.csv(paste0(output_folder, "M_area_peaks.csv"))
feature_envelopes <- read.csv(paste0(output_folder, "feature_envelopes.csv"))
feature_data <- M_area_peaks %>%
  group_by(feature) %>%
  summarise(mzmed=median(mz), rtmed=median(rt), avg_area=mean(M_area))
stan_annotations <- read.csv(paste0(output_folder, "stan_annotations.csv"))
# database_formulae <- readRDS("unique_formulae.rds") %>%
#   gsub(pattern = "\\+.?", replacement = "")
database_formulae <- unique(readRDS("hmdb_formulae.rds"))

source("scripts/formula_assignments.R")

saveRDS(sirius_formulas, file = paste0(output_folder, "sirius_formulas.rds"))
saveRDS(rdisop_formulas, file = paste0(output_folder, "rdisop_formulas.rds"))
saveRDS(manual_check, file = paste0(output_folder, "iso_formulas.rds"))
write.csv(feature_formulas, file = paste0(output_folder, "feature_formulas.csv"), row.names = FALSE)
```

```{r assessformulas}
feature_formulas <- read.csv(paste0(output_folder, "feature_formulas.csv"))
stan_annotations <- read.csv(paste0(output_folder, "stan_annotations.csv")) %>%
  filter(!is.na(feature)) %>%
  select(compound_name, feature)
given_stans <- read.csv("../metadata/made_data/clean_stans.csv") %>% 
  filter(polarity==!!polarity) %>%
  filter(compound_type!="Internal Standard") %>%
  select(compound_name, compound_type, official_formula=formula)

anno_quality <- stan_annotations %>%
  left_join(feature_formulas, by="feature") %>%
  left_join(given_stans,by="compound_name") %>%
  filter(!is.na(official_formula)) %>%
  select(compound_name, feature, official_formula, formula) %>%
  rowwise() %>%
  mutate(form_diffH=ifelse(polarity=="pos", minH(formula), addH(formula))) %>%
  mutate(success=case_when(
    is.na(formula) ~ "Unannotated",
    formula==official_formula | form_diffH==official_formula ~ "Success",
    TRUE ~ "Failed"
  )) %>%
  print(n=Inf)

anno_quality %>% count(success)
anno_quality %>% filter(success=="Failed")
```

## Annotate classes and structures

```{r assignids}
filled_peaks <- read.csv(paste0(output_folder, "filled_peaks.csv"))
feature_envelopes <- read.csv(paste0(output_folder, "feature_envelopes.csv"))
feature_data <- filled_peaks %>%
  group_by(feature) %>%
  summarise(mzmed=median(mz), rtmed=median(rt), avg_area=mean(M_area)) %>%
  left_join(feature_envelopes, by = "feature")
stan_annotations <- read.csv(paste0(output_folder, "stan_annotations.csv"))

MSMS_files <- paste0(mzml_path, "/MSMS") %>%
  list.files(pattern = unique(str_extract(ms_files, "^\\d+"))[1], 
             full.names = TRUE)
sirius_project_dir <- paste0(output_folder, "/sirius_project")

source("scripts/run_sirius.R")

write.csv(classy_confs, file = paste0(output_folder, "classy_confs.csv"), row.names = FALSE)
write.csv(structure_ids, file = paste0(output_folder, "structure_ids.csv"), row.names = FALSE)
```

```{r assessclasses}
stan_classes <- read.csv(paste0("https://raw.githubusercontent.com/",
                             "IngallsLabUW/Ingalls_Standards/",
                             "b098927ea0089b6e7a31e1758e7c7eaad5408535/",
                             "Ingalls_Lab_Standards_NEW.csv")) %>%
  filter(Column=="HILIC") %>%
  mutate(polarity=tolower(gsub(pattern = "HILIC", "", .$Fraction1))) %>%
  mutate(m.z=as.numeric(m.z)) %>%
  select(compound_name=Compound.Name, classyfire=Classyfire)
stan_annotations <- read.csv(paste0(output_folder, "stan_annotations.csv")) %>%
  select(compound_name, feature) %>%
  separate_rows(feature, sep = "; ")
classy_confs <- read.csv(file = paste0(output_folder, "classy_confs.csv"))

stan_classes %>%
  left_join(stan_annotations, by="compound_name") %>%
  filter(!is.na(feature)) %>%
  separate_rows(classyfire, sep = "; ") %>%
  mutate(classyfire=str_replace(classyfire, "^.*: ", "")) %>%
  mutate(classyfire=str_replace_all(classyfire, "_", " ")) %>%
  distinct() %>%
  left_join(classy_confs, by=c(classyfire="classes", "feature")) %>%
  print(n=100)
```

```{r assessstructures}
structure_ids <- read.csv(paste0(output_folder, "structure_ids.csv"))
stan_annotations <- read.csv(paste0(output_folder, "stan_annotations.csv")) %>%
  select(compound_name, feature) %>%
  separate_rows(feature, sep = "; ") %>%
  left_join(structure_ids, by="feature") %>%
  # filter(compound_name=="Glycerophosphocholine") %>%
  print(n=200)

```

## Collect into the clean database

After running all the code above, we've got a bunch of separate data frames that
are a pain to work with when all we really need is two - one to hold the values
corresponding to each *peak* and one to hold the values corresponding to each
*feature*.

Peak values: mz, rt, area, norm_area, filename

Feature values: mzmed, rtmed, avgarea, formula, classes, structure/id

These data frames, combined with metadata_complete, can then be left joined
together and filtered/selected based on what information is actually needed.
This makes the stats scripts a lot cleaner and avoids needing to recalculate
medmz, medrt, etc. every single time.

We begin with peakvals, which starts out as the xcms output raw_peaks.
Step 1: remove adducts/isotopes named in addiso_features
Step 2: calculate norm_area using the chosen_BMIS and IS_peaks

```{r peakvals, purl=TRUE}
# Step 0: load relevant data
M_area_peaks <- read.csv(paste0(output_folder, "M_area_peaks.csv")) %>%
  select(feature, mz, rt, M_area, filename)
chosen_BMIS <- read.csv(paste0(output_folder, "chosen_BMIS.csv"))
is_peaks <- read.csv(paste0(output_folder, "is_peaks.csv")) %>%
  select(compound_name, filename, is_area=M_area)
is_addiso <- read.csv(paste0(output_folder, "is_addiso.csv"))

# Step 1
M_area_peaks <- M_area_peaks %>%
  filter(!feature%in%is_addiso$feature)

# Step 2
norm_peaks <- M_area_peaks %>%
  left_join(chosen_BMIS, by="feature") %>%
  left_join(is_peaks, by = c(BMIS="compound_name", "filename")) %>%
  group_by(feature) %>%
  mutate(norm_area=(M_area/is_area)*mean(is_area, na.rm=TRUE)) %>%
  select(feature, mz, rt, norm_area, filename)

write.csv(norm_peaks, paste0(output_folder, "norm_peaks.csv"), row.names = FALSE)
```

Featurevals then builds on this new norm_peaks data frame.
Step 1: Group by feature and produce summary statistics
 - mzmed, rtmed, avgarea
Step 2: Annotate known standards with stan_annotations
Step 3: Annotate formulae with feature_formulas
Step 4: Annotate classes with classy_confs (currently removed because annoying)
Step 5: Annotate hypothetical structures with structure_ids

```{r featurevals, purl=TRUE}
# Step 0: Load everything
stan_annotations <- read.csv(paste0(output_folder, "stan_annotations.csv")) %>%
  select(compound_name, feature)
feature_formulas <- read.csv(paste0(output_folder, "feature_formulas.csv"))

# Step 1: group and summarize
norm_feats <- norm_peaks %>%
  filter(str_detect(filename, "_Smp_")) %>%
  group_by(feature) %>%
  summarize(mzmed=mean(mz, na.rm=TRUE), rtmed=mean(rt, na.rm=TRUE), avgarea=mean(norm_area, na.rm=TRUE))

# Step 2: authentic standards
norm_feats <- norm_feats %>%
  left_join(stan_annotations, by="feature")

# Step 3: formulae
norm_feats <- norm_feats %>%
  left_join(feature_formulas, by="feature")

# Step 4: classes
# 0.99 cutoff is arbitrary
if(file.exists(paste0(output_folder, "classy_confs.csv"))){
  classy_confs <- read.csv(paste0(output_folder, "classy_confs.csv"))
  clean_classes <- classy_confs %>%
    filter(conf>0.99) %>%
    group_by(feature) %>%
    summarize(classes=paste(classes, collapse = "; "))
# # Currently commented out because it's messy and super long and printing sucks
# norm_feats <- norm_feats %>%
#   left_join(clean_classes, by="feature")
  # Instead they're all being set to NA
  norm_feats$classes <- NA
} else {
  norm_feats$classes <- NA
}

# Step 5: structures
# -20 cutoff is due to Raafay's manual checks, -25 ~= 80% correct
if(file.exists(paste0(output_folder, "structure_ids.csv"))){
  classy_confs <- read.csv(paste0(output_folder, "structure_ids.csv"))
  clean_CSI <- structure_ids %>%
    filter(confidence>-25) %>%
    select(feature, CSI_name=name)
  norm_feats <- left_join(norm_feats, clean_CSI, by="feature")
} else {
  norm_feats$CSI_name <- NA
}

write.csv(norm_feats, paste0(output_folder, "norm_feats.csv"), row.names = FALSE)
```

## Cruise correspondence

Finally, it's useful to combine all of the output into a single data frame
rather than working with several little ones. The real challenge here is
asserting that a feature is the same across the different runs given differences
in retention time, m/z, and polarity. But mostly retention time.

This can only be done on output folders that have been produced by the script
above, so if you haven't run those yet then please make sure you do.

```{r correspond, purl=TRUE}
all_cruises <- c("FK", "MT", "MC")
polarities <- rep(c("pos", "neg"), each=length(all_cruises))
output_all <- paste(paste(all_cruises, polarities, sep="_"), "output/", sep = "_")

peaks_all_files <- paste0(output_all, "norm_peaks.csv")
feats_all_files <- paste0(output_all, "norm_feats.csv")

if(!all(file.exists(feats_all_files))){
  message(paste("Not all feature lists found:", paste(feats_all_files[!file.exists(feats_all_files)], collapse = ", "), "missing"))
  next # Allows running file within for loop after purling
  # Code below only runs if all files exist
}
if(!all(file.exists(peaks_all_files))){
  message(paste("Not all feature lists found:", paste(feats_all_files[!file.exists(feats_all_files)], collapse = ", "), "missing"))
  stop("Peak lists should be made simultaneously with feature lists")
}

feats_all <- lapply(feats_all_files, function(x){
  given_cruise <- str_extract(x, "FK|MT|MC")
  v <- read.csv(x)
  v$cruise <- given_cruise
  v$feature <- paste(given_cruise, toupper(str_extract(x, "pos|neg")), v$feature, sep = "_")
  v$polarity <- str_extract(x, "pos|neg")
  v
}) %>% bind_rows() %>%
  arrange(desc(avgarea))

temp_feats <- feats_all %>%
  mutate(rtmed=ifelse(cruise=="FK", rtmed+60, rtmed))
output_groups <- list()
pb <- txtProgressBar(min = 0, max = nrow(temp_feats), style = 3)
while(nrow(temp_feats>1)){
  row_data <- temp_feats[1,]
  # if(!is.na(row_data$compound_name)){
  #   if(str_detect(row_data$compound_name, "13C12")){
  #     dput(row_data)
  #     stop()
  #   }
  # }
  out_group <- temp_feats %>%
    filter(polarity==row_data$polarity) %>%
    filter(cruise!=row_data$cruise) %>%
    filter(mzmed%between%pmppm(row_data$mzmed, 5)) %>%
    mutate(dist_from_main=map2_dbl(rtmed, avgarea, function(rtmed, avgarea){
      rt_dist <- abs(log2(row_data$rtmed)-log2(rtmed))
      area_dist <- abs(log10(row_data$avgarea)-log10(avgarea))
      sum(rt_dist, area_dist)
    })) %>%
    arrange(dist_from_main)
  if(nrow(out_group>0)){
    out_group <- out_group %>%
      filter(rtmed%between%(row_data$rtmed+c(-60, 60))) %>%
      group_by(cruise) %>%
      slice(1) %>%
      select(-dist_from_main) %>%
      ungroup() %>%
      add_row(row_data) %>%
      mutate(rtmed=ifelse(cruise=="FK", rtmed-60, rtmed))
  } else {
    out_group <- row_data
  }
  temp_feats <- temp_feats[!temp_feats$feature%in%out_group$feature,]
  output_groups[[length(output_groups)+1]] <- out_group
  setTxtProgressBar(pb, nrow(feats_all)-nrow(temp_feats))
}
close(pb)

feat_groups_init <- output_groups %>%
  map2_dfr(seq_along(.), function(group, id){
    cbind(group, group_feat=sprintf("FT%04d", id))
  })
```

```{r fix correspondence, purl=TRUE}
(same_feat_diff_names <- feat_groups_init %>%
  group_by(group_feat) %>%
  summarise(n=length(unique(compound_name[!is.na(compound_name)])),
            nms=paste0(unique(compound_name), collapse = ";")) %>%
  filter(n>1))

(same_name_diff_feats <- feat_groups_init %>%
  group_by(compound_name, polarity) %>%
  summarise(n=length(unique(group_feat)),
            nms=paste0(unique(group_feat), collapse = ";")) %>%
  filter(n>1 & !is.na(compound_name)))

# Fixes will go here if necessary
feat_groups <- feat_groups_init %>%
  select(feature, group_feat)
```

```{r final peak list compilation, purl=TRUE}
all_cruises <- c("FK", "MT", "MC")
polarities <- rep(c("pos", "neg"), each=length(all_cruises))

all_peak_list <- list()
all_feat_list <- list()
for(cruise in all_cruises){
  for(polarity in polarities){
    peaklist_path <- paste(cruise, polarity, "output/norm_peaks.csv", sep = "_")
    all_peak_list[[cruise]][[polarity]] <- peaklist_path %>%
      read_csv(show_col_types = FALSE) %>%
      mutate(polarity=polarity) %>%
      mutate(cruise=cruise) %>%
      mutate(feature=paste(cruise, toupper(polarity), feature, sep = "_"))

    featlist_path <- paste(cruise, polarity, "output/norm_feats.csv", sep = "_")
    all_feat_list[[cruise]][[polarity]] <- featlist_path %>%
      read_csv(show_col_types = FALSE) %>%
      mutate(polarity=polarity) %>%
      mutate(cruise=cruise) %>%
      mutate(feature=paste(cruise, toupper(polarity), feature, sep = "_"))
  }

  neg_dupes <- all_feat_list[[cruise]]$pos %>%
    pmap_dfr(function(...){
      row_data <- data.frame(...)
      all_feat_list[[cruise]]$neg %>%
        filter(mzmed%between%pmppm(row_data$mzmed-1.007276*2, 10)) %>%
        filter(rtmed%between%(row_data$rtmed+c(-5, 5))) %>%
        distinct(feature)
    })
  all_feat_list[[cruise]]$neg <- all_feat_list[[cruise]]$neg %>%
    anti_join(neg_dupes, by = "feature")
  all_peak_list[[cruise]]$neg <- all_peak_list[[cruise]]$neg %>%
    anti_join(neg_dupes, by = "feature")
}

all_peaks <- bind_rows(lapply(all_peak_list, bind_rows)) %>%
  select(feature, filename, norm_area)
all_feats <- bind_rows(lapply(all_feat_list, bind_rows)) %>%
  left_join(feat_groups)

write.csv(all_peaks, file = "all_peaks.csv", row.names = FALSE)
write.csv(all_feats, file = "all_feats.csv", row.names = FALSE)
```

### 2022-10-05 untargeted/scripts/bmisscript.R

@@ -1,0 +1,37 @@
+# Script to find the B-MIS for each standard in the untargeted data set
+# Called by Control.Rmd
+
+# all_peaks, given_stans, polarity, pmppm all defined in Control.Rmd
+
+
+# Perform BMIS
+pooled <- M_area_peaks %>%
+  filter(str_detect(filename, "_Poo_"))
+
+pooled_IS <- is_peaks %>%
+  rename_with(~paste0("IS_", .), -filename) %>%
+  add_row(filename=unique(.$filename), IS_compound_name="None", IS_M_area=1)
+
+initial_CVs <- pooled %>%
+  mutate(M_area=ifelse(str_detect(filename, "Half"), M_area*2, M_area)) %>%
+  group_by(feature) %>%
+  summarize(init_CV=cv(M_area))
+
+standardized_CVs <- full_join(pooled, pooled_IS, by=c("filename")) %>%
+  group_by(feature, IS_compound_name) %>%
+  summarize(norm_CV=cv(M_area/IS_M_area), .groups="drop")
+
+BMIS_data <- standardized_CVs %>%
+  left_join(initial_CVs) %>%
+  group_by(feature) %>%
+  mutate(MIS_improvement=(init_CV - norm_CV)/init_CV) %>% 
+  filter(MIS_improvement>min_improvement|IS_compound_name=="None") %>%
+  arrange(desc(MIS_improvement)) %>%
+  group_by(feature) %>%
+  slice(1) %>%
+  select(feature, BMIS=IS_compound_name, init_CV, norm_CV, MIS_improvement)
+chosen_BMIS <- BMIS_data %>% select(feature, BMIS)
+
+is_peaks <- is_peaks %>%
+  add_row(compound_name="None", filename=unique(.$filename), 
+          M_area=ifelse(str_detect(filename, "Half"), 0.5, 1))

### 2022-10-05 untargeted/scripts/compound_assignments.R

@@ -1,0 +1,52 @@
+# Collects structural best-guesses from CSI:FingerID and compares them with
+# metfRag and other software?
+
+library(tidyverse)
+library(rvest)
+
+getSynonymsName <- function(cmpd_name, how_many=20){
+  prolog <- "https://pubchem.ncbi.nlm.nih.gov/rest/pug/compound/name/"
+  tail <- "/cids/TXT"
+  url <- paste0(prolog, cmpd_name, tail)
+  CID <- readLines(url)
+  getSynonymsCID(CID, how_many)
+}
+
+getSynonymsCID <- function(CID, how_many=20){
+  prolog <- "https://pubchem.ncbi.nlm.nih.gov/rest/pug/compound/cid/"
+  tail <- "/synonyms/TXT"
+  url <- paste0(prolog, CID, tail)
+  content <- head(readLines(url), how_many)
+  paste0(content, collapse = "; ")
+}
+
+cmpd_file <- "XCMS/data_intermediate/sirius_project/output_dir/compound_identifications_adducts.tsv"
+cmpd_raw <- read_tsv(cmpd_file, col_types = 
+         cols(
+           rank = col_double(),
+           `#adducts` = col_double(),
+           `#predictedFPs` = col_double(),
+           Confidence_Score = col_character(),
+           `CSI:FingerID_Score` = col_double(),
+           Zodiac_Score = col_double(),
+           TreeIsotope_Score = col_double(),
+           molecularFormula = col_character(),
+           adduct = col_character(),
+           InChIkey2D = col_character(),
+           InChI = col_character(),
+           name = col_character(),
+           smiles = col_character(),
+           xlogp = col_logical(),
+           pubchemids = col_character(),
+           links = col_character(),
+           dbflags = col_double(),
+           id = col_character()
+         ))
+cmpd_data <- cmpd_raw %>%
+  mutate(feature=sub(pattern = "[[:digit:]]*_", replacement = "", x = .$id)) %>%
+  mutate(feature=sub(pattern = "_FEATURE1", replacement = "", x = .$feature)) %>%
+  #mutate(top_cid=sub(pattern = ";.*", replacement = "", x=.$pubchemids)) %>%
+  head() %>%
+  #mutate(synonyms=sapply(X = .$top_cid, getSynonymsCID))
+  mutate(synonyms=sapply(X = .$name, getSynonymsName))
+cmpd_data %>% select(name, synonyms)

### 2022-10-05 untargeted/scripts/deisoadduct.R

@@ -1,0 +1,116 @@
+# Script to de-isotope and de-adduct XCMS-picked peaks
+# Called by Control.Rmd
+
+# Include isIsoAdduct code here so we can safely source() the functions script
+# rather than providing each of the functions because bplapply doesn't copy over
+# the working environment
+# given_file <- filterFile(xdata_filled, "180821_Smp_MS12C115m_A.mzML")
+isIsoAdduct <- function(file_peaks, xdata_filled, addiso_masses){
+  source("scripts/functions.R")
+  library(data.table)
+  setDTthreads(1)
+  library(dplyr, warn.conflicts = FALSE)
+  suppressPackageStartupMessages(library(xcms))
+  
+  file_addisos <- addiso_masses %>%
+    {rbind(cbind(., type="self", dir=-1), cbind(., type="seek", dir=1))} %>%
+    mutate(mz_diff=mz_diff*dir) %>%
+    select(-dir) %>%
+    group_by(addiso, type) %>%
+    group_split()
+  addiso_range <- range(sapply(file_addisos, function(x)x$mz_diff))
+  
+  given_file <- filterFile(xdata_filled, unique(file_peaks$filename))
+  file_data_table <- as.data.table(given_file)[,c("rt", "mz", "i")]
+  
+  # Split up the large data.table into smaller ones to help speed up binary search
+  split_vec <- round(1:nrow(file_peaks)/(nrow(file_peaks)/100))
+  split_file_peaks <- split(file_peaks[order(file_peaks$rt),], split_vec)
+  split_data_table <- lapply(split_file_peaks, function(file_peaks_i){
+    file_data_table[rt%between%c(min(file_peaks_i$rtmin), max(file_peaks_i$rtmax))]
+  })
+  
+  # test_peak <- file_peaks %>% slice_sample(n=1) # Debug
+  # test_peak <- file_peaks %>% filter(mz%between%pmppm(118.0865, 5))
+  # test_peak <- file_peaks %>% filter(feature=="FT0487")
+  # feature_i <- test_peak$feature
+  # mz_i <- test_peak$mz
+  # rtmin_i <- test_peak$rtmin
+  # rtmax_i <- test_peak$rtmax
+  # split_table_i <- file_data_table
+  
+  addiso_all <- mapply(function(split_peaks_i, split_table_i){
[…]

### 2022-10-05 untargeted/scripts/formula_assignments.R

@@ -1,0 +1,145 @@
+
+# Load addisos - this should be done earlier in the script but that's not guaranteed
+if(!exists(addiso_masses)){
+  warning("Using manually written addisos, may need updating")
+  addiso_masses_all <- tribble(
+    ~adduct_or_iso, ~polarity, ~addiso, ~nat_abund, ~mz_diff,
+    "iso",          "both",    "C13",   0.0110,     1.003355, 
+    "iso",          "both",    "X2C13", 0.0110,     2*1.003355, 
+    "iso",          "both",    "S34",   0.0421,     1.995796, 
+    "iso",          "both",    "S33",   0.0075,     0.999387,
+    "iso",          "both",    "N15",   0.0037,     0.997035, 
+    "iso",          "both",    "O18",   0.0020,     2.004244,
+    "iso",          "both",    "Cl37",  0.2423,     1.99705,
+    "iso",          "both",    "Br81",  0.4931,     1.997954,
+    "adduct",       "both",    "M-H2O", NA,         -18.0106, 
+    "adduct",       "pos",     "M+Na",  NA,          22.98922-1.007276 # Goes from [M+H] to [M+Na]
+    # "adduct",       "pos",     "M+NH4", NA,          18.0338-1.007276,
+    # "adduct",       "pos",     "M+K",   NA,          38.963708-1.007276,
+    # "adduct",       "neg",     "M+Cl",  NA,          34.969402+1.007276,
+    # "adduct",       "neg",     "M+Ac",  NA,          59.013851+1.007276,
+    # "adduct",       "neg",     "M+Br",  NA,          78.918885+1.007276
+  )
+  addiso_masses <- addiso_masses_all %>% filter(polarity%in%c(!!polarity, "both"))
+}
+
+
+# Run SIRIUS ----
+message("Running SIRIUS formulas...")
+mzs <- feature_data$mzmed
+sirius_cmd <- paste0('sirius --noCite decomp',
+                     ' --mass=', paste(mzs, collapse = ","),
+                     ' --ppm=5',
+                     ' --elements=CHNOPS')
+if(nchar(sirius_cmd)>8096){
+  # Windows doesn't support commands longer than 8,096 characters
+  n_chunks <- ceiling(nchar(sirius_cmd)/8096)
[…]

### 2022-10-05 untargeted/scripts/funcs_old.R

@@ -1,0 +1,505 @@
+# Functions called by Control.Rmd or scripts within
+# Should be sourced every time!
+
+cv <- function(x, na.rm=FALSE, inf.rm=FALSE){
+  if(inf.rm){
+    x[is.infinite(x)] <- NA
+  }
+  sd(x = x, na.rm = na.rm)/mean(x = x, na.rm = na.rm)
+}
+
+pmppm <- RaMS::pmppm
+
+grabSingleFileData <- function(filename){
+  msdata <- mzR:::openMSfile(filename)
+  fullhd <- mzR::header(msdata)
+  spectra_list <- lapply(seq_len(nrow(fullhd)), function(x){
+    given_peaks <- mzR::peaks(msdata, x)
+    rtime <- fullhd[x, "retentionTime"]
+    return(cbind(rtime, given_peaks))
+  })
+  all_data <- `names<-`(as.data.frame(do.call(rbind, spectra_list)), 
+                        c("rt", "mz", "int"))
+  return(all_data)
+}
+speedyQscoreCalculator <- function(file_peaks, grabSingleFileData, qscoreCalculator){
+  library(data.table)
+  file_data <- grabSingleFileData(unique(file_peaks$filename))
+  file_data_table <- as.data.table(file_data)
+  peak_splits <- split(file_peaks, ceiling(seq_len(nrow(file_peaks))/100))
+  file_qscores <- lapply(peak_splits, function(i){
+    eic_many <- file_data_table[mz%between%c(min(i$mzmin), max(i$mzmax))]
+    individual_peaks <- split(i, seq_len(nrow(i)))
+    output <- sapply(individual_peaks, function(peak_row){
+      eic <- eic_many[rt%between%c(peak_row$rtmin, peak_row$rtmax) & 
+                        mz%between%c(peak_row$mzmin, peak_row$mzmax)]
+      qscoreCalculator(eic)
+    }, USE.NAMES = FALSE)
+  })
+  file_peaks$sn <- unlist(file_qscores)
+  return(file_peaks)
+}
+qscoreCalculator <- function(eic){
+  #Check for bogus EICs
+  if(nrow(eic)<5){
+    return(0)
+  }
+  #Calculate where each rt would fall on a beta dist (accounts for missed scans)
+  scaled_rts <- (eic$rt-min(eic$rt))/(max(eic$rt)-min(eic$rt))
+  
+  # Create a couple different skews and test fit
[…]

### 2022-10-05 untargeted/scripts/functions.R

@@ -1,0 +1,345 @@
+# Functions called by Control.Rmd or scripts within
+# Should be sourced every time!
+
+cv <- function(x, na.rm=FALSE, inf.rm=FALSE){
+  if(inf.rm){
+    x[is.infinite(x)] <- NA
+  }
+  sd(x = x, na.rm = na.rm)/mean(x = x, na.rm = na.rm)
+}
+speedyQscoreCalculator <- function(file_peaks, grabSingleFileData, qscoreCalculator){
+  library(data.table)
+  file_data <- grabSingleFileData(unique(file_peaks$filename))
+  file_data_table <- as.data.table(file_data)
+  peak_splits <- split(file_peaks, ceiling(seq_len(nrow(file_peaks))/100))
+  file_qscores <- lapply(peak_splits, function(i){
+    eic_many <- file_data_table[mz%between%c(min(i$mzmin), max(i$mzmax))]
+    individual_peaks <- split(i, seq_len(nrow(i)))
+    output <- sapply(individual_peaks, function(peak_row){
+      eic <- eic_many[rt%between%c(peak_row$rtmin, peak_row$rtmax) & 
+                        mz%between%c(peak_row$mzmin, peak_row$mzmax)]
+      qscoreCalculator(eic)
+    }, USE.NAMES = FALSE)
+  })
+  file_peaks$sn <- unlist(file_qscores)
+  return(file_peaks)
+}
+qscoreCalculator <- function(eic){
+  #Check for bogus EICs
+  if(nrow(eic)<5){
+    return(0)
+  }
+  #Calculate where each rt would fall on a beta dist (accounts for missed scans)
+  scaled_rts <- (eic$rt-min(eic$rt))/(max(eic$rt)-min(eic$rt))
+  
+  # Create a couple different skews and test fit
+  maybe_skews <- c(2.5,3,4,5) #Add 7 to catch more multipeaks and more noise
+  #Add 2 to catch very slopey peaks and more noise
+  best_skew <- maybe_skews[which.max(sapply(maybe_skews, function(x){
+    cor(dbeta(scaled_rts, shape1 = x, shape2 = 5), eic$int)
+  }))]
+  perf_peak <- dbeta(scaled_rts, shape1 = best_skew, shape2 = 5)
+  peak_cor <- cor(perf_peak, eic$int)
+  
+  
+  #Calculate the normalized residuals
+  residuals <- eic$int/max(eic$int)-perf_peak/max(perf_peak)
[…]

### 2022-10-05 untargeted/scripts/peakpicking.R

@@ -1,0 +1,82 @@
+# Script to run XCMS peakpicking, retention time correction, and grouping
+# Called by output_control.Rmd
+# Requires dev version of XCMS for improved peakpicking given settings
+
+# ms_files comes from output_control.Rmd
+# qscore_threshold comes from output_control.Rmd
+
+
+# Read in the raw data ----
+message("Reading files...")
+start_time <- Sys.time()
+raw_data <- readMSData(files = normalizePath(paste0(mzml_path, ms_files)), 
+                       msLevel. = 1, centroided. = TRUE, mode = "onDisk",
+                       pdata = as(cruise_metadata, "AnnotatedDataFrame"))
+time_total <- round(difftime(Sys.time(), start_time), digits = 2)
+cat("Time to read in files: ", time_total, units(time_total), "\n")
+
+# Perform peakpicking ----
+message("Picking peaks...")
+start_time <- Sys.time()
+xdata <- suppressMessages(findChromPeaks(raw_data, param = cwp))
+time_total <- round(difftime(Sys.time(), start_time), digits = 2)
+cat("Time to perform peakpicking: ", time_total, units(time_total), "\n")
+# 13 minutes
+
+
+# Assign new quality scores ----
+# speedyQscoreCalculator is custom function which should be in functions script
+# sourced by output_control.Rmd
+message("Assigning quality scores...")
+start_time <- Sys.time()
+xcms_peakdf <- chromPeaks(xdata) %>%
+  as.data.frame(stringsAsFactors=FALSE) %>%
+  mutate(filename=fileNames(xdata)[.[["sample"]]]) %>%
+  arrange(mz)
+split_xcms_filedata <- split(xcms_peakdf, xcms_peakdf$filename)
+files_qscores <- bplapply(X = split_xcms_filedata, FUN = speedyQscoreCalculator,
+                          grabSingleFileData, qscoreCalculator)
+peakdf_qscored <- files_qscores %>%
+  do.call(what = rbind) %>%
+  as.data.frame(stringsAsFactors=FALSE) %>% 
+  filter(sn>qscore_threshold) %>%
+  select(-filename) %>%
+  arrange(sample, rtmin, rtmax) %>%
+  as.matrix()
+xdata_cleanpeak <- `chromPeaks<-`(xdata, peakdf_qscored)
+
[…]

### 2022-10-05 untargeted/scripts/run_sirius.R

@@ -1,0 +1,133 @@
+# Grab MSMS data ----
+print("Reading in MS2 data...")
+msdata <- grabMSdata(MSMS_files, grab_what = c("MS1", "MS2"))
+msdata$MS1$rt <- msdata$MS1$rt*60
+msdata$MS2$rt <- msdata$MS2$rt*60
+
+print("Determining which have MS2 frags...")
+has_msms <- feature_data %>%
+  select(feature, mzmed, rtmed) %>%
+  split(.$feature) %>%
+  pbsapply(function(x){
+    msdata$MS2[premz%between%pmppm(x$mzmed, ppm = 5)] %>%
+      filter(rt%between%(x$rtmed+c(-30, 30))) %>%
+      nrow() %>%
+      as.logical()
+  })
+data.frame(feature=names(has_msms), has_msms) %>%
+  filter(has_msms) %>%
+  left_join(stan_annotations, by=c("feature")) %>%
+  # filter(!is.na(compound_name)) %>%
+  select(feature, compound_name)
+feature_data <- feature_data %>%
+  filter(feature%in%names(has_msms)[has_msms])
+
+
+
+# Create .mgf files for SIRIUS to read ----
+if(dir.exists(sirius_project_dir)){
+  unlink(sirius_project_dir, recursive = TRUE)
+}
+dir.create(sirius_project_dir)
+dir.create(paste0(sirius_project_dir, "//raw_files"))
+message("Writing .mgf files...")
+pbsapply(feature_data$feature, function(feature_num){
+  output_file <- paste0(sirius_project_dir, "\\raw_files\\", feature_num, ".mgf")
+  feature_msdata <- feature_data[feature_data$feature==feature_num, ]
+  ms1 <- rbind(c(feature_msdata$mzmed, feature_msdata$avg_area),
+               c(feature_msdata$mzmed+1.003355, feature_msdata$C13),
+               c(feature_msdata$mzmed+1.003355*2, feature_msdata$X2C13),
+               c(feature_msdata$mzmed+1.995796, feature_msdata$S34),
+               c(feature_msdata$mzmed+0.997035, feature_msdata$N15),
+               c(feature_msdata$mzmed+2.004244, feature_msdata$O18))
+  ms1 <- ms1[ms1[,2]!=0, , drop=FALSE]
+  ms2 <- msdata$MS2[premz%between%pmppm(feature_msdata$mzmed)&
[…]

### 2022-10-05 untargeted/scripts/standard_assignments.R

@@ -1,0 +1,393 @@
+
+# Functions ----
+stan_guesser <- function(isotope_choice, mix_choice, match_choice, area_choice, rt_choice){
+  if(all(is.na(c(isotope_choice, mix_choice, match_choice, area_choice, rt_choice)))){
+    return("No peaks found")
+  }
+  if(!is.na(isotope_choice)){
+    return(isotope_choice)
+  }
+  if(!is.na(match_choice)) {
+    return(match_choice)
+  }
+  if(is.na(mix_choice)){
+    mix_options <- NA
+  } else if(!nchar(mix_choice)){
+    mix_options <- NA
+  } else {
+    mix_options <- unlist(strsplit(mix_choice, split = "; "))
+  }
+  if(length(mix_options)==1&!all(is.na(mix_options))){
+    return(mix_options)
+  }
+  
+  if(length(mix_options)>1&!all(is.na(mix_options))){
+    if(rt_choice==area_choice){
+      return(rt_choice)
+    }
+    if(rt_choice%in%mix_options&!area_choice%in%mix_options){
+      return(rt_choice)
+    }
+    if(area_choice%in%mix_options&!rt_choice%in%mix_options){
+      return(area_choice)
+    }
+    if(area_choice%in%mix_options&rt_choice%in%mix_options){
+      return(paste(c(area_choice, rt_choice), collapse = "; "))
+    }
+  }
+  return(rt_choice)
+}
+reAnno <- function(anno_df, cmpd_name, which_new=NULL, new_feature_num=NULL){
+  if(is.null(new_feature_num) & is.null(which_new)){
+    stop(paste0("For ", cmpd_name, ", both `new_feature_name` and ",
+                "`which_new` must not both be NULL"))
+  }
+  if(!is.null(which_new) && !which_new%in%c("largest", "first", "last", "second")){
+    stop(paste0("For ", cmpd_name, ", ", which_new, " is not a known keyword"))
+  }
+  if(!is.null(new_feature_num)){
+    new_df <- anno_df %>% rows_update(
+      data.frame(compound_name=cmpd_name, feature=new_feature_num, isotope_validated="Manual"), 
+      by = "compound_name"
+    )
+    return(new_df)
+  }
+  if(!is.null(which_new)){
+    chosen_feature <- anno_df %>% filter(compound_name==cmpd_name) %>% pull(feature)
+    if(length(chosen_feature)==0){
+      stop(paste0(cmpd_name, " not found in supplied anno_df"))
+    }
[…]

### 2022-10-05 untargeted/scripts/treatment_diff_search.R

@@ -1,0 +1,74 @@
+# Code that looks for differences in peak area across treatments
+# Needs a peak list (typically produced by peakpicking.R)
+# Calculates p-values between treatments
+#  p-values are amended via the Holm-Bonferroni method to avoid multiple comparisons
+#  Treatments are deduced from the file_name column of the peak list
+# Data that only shows up in one treatment is given a p-value of 0
+
+library(tidyverse)
+final_peaks <- read.csv(file = "XCMS/data_pretty/final_peaks.csv")
+final_features <- read.csv(file = "XCMS/data_pretty/final_features.csv")
+feature_formulas <- read.csv(file = "XCMS/data_pretty/feature_formulas.csv")
+feature_putatives <- "XCMS/data_intermediate/sirius_project/output_dir/" %>%
+  paste0("compound_identifications.tsv") %>%
+  readLines() %>%
+  strsplit(split = "\t")
+header <- feature_putatives[[1]]
+feature_putatives <- feature_putatives[-1] %>%
+  do.call(what = rbind) %>%
+  as.data.frame() %>%
+  `names<-`(header) %>%
+  select(id, Confidence_Score, name) %>%
+  mutate(feature=gsub(pattern = "^[[:digit:]]+_", "", .$id)) %>%
+  mutate(feature=gsub(pattern = "_FEATURE1$", "", .$feature)) %>%
+  mutate(Confidence_Score=round(as.numeric(Confidence_Score), digits = 3)) %>%
+  select(-id)
+
+diffreport <- function(peaks, pattern_a, pattern_b){
+  split(peaks, peaks$feature) %>%
+    lapply(function(x){
+      DCM_areas <- x$BMISed_area[grep(pattern = pattern_a, x$file_name)]
+      m25_areas <- x$BMISed_area[grep(pattern = pattern_b, x$file_name)]
+      if(length(DCM_areas)<3|length(m25_areas)<3)return(c(0, 1))
+      c(pval=t.test(DCM_areas, m25_areas)$p.value, 
+        diff=mean(DCM_areas)/mean(m25_areas))
+    }) %>% do.call(what = rbind) %>% as.data.frame() %>%
+    mutate(feature=rownames(.)) %>%
+    arrange(pval) %>%
+    mutate(p_adj=p.adjust(pval, method = "BH")) %>%
[…]

### 2022-10-05 untargeted/output_control.Rmd
Moving hmdb formulae to metadata

@@
 #   gsub(pattern = "\\+.?", replacement = "")
-database_formulae <- unique(readRDS("hmdb_formulae.rds"))
+database_formulae <- unique(readRDS("../metadata/raw_data/hmdb_formulae.rds"))
 

### 2022-10-13 metadata/metadata_control.Rmd
Adding MAAs to standards list

@@
           date_added=150101, polarity="pos") %>%
+  add_row(compound_type="Custom", compound_name="Asterina-330",
+          compound_name_old=compound_name, formula="C12H20N2O6", rt = 10.5,
+          mz=289.1393, ionization_form="[M+H]", charge=1) %>%
+  add_row(compound_type="Custom", compound_name="Shinorine",
+        compound_name_old=compound_name, formula="C13H20N2O8", rt = 11.6,
+        mz=333.1292, ionization_form="[M+H]", charge=1) %>%
+  add_row(compound_type="Custom", compound_name="Palythine",
+        compound_name_old=compound_name, formula="C10H16N2O5", rt = 11.0,
+        mz=245.1132, ionization_form="[M+H]", charge=1) %>%
+  add_row(compound_type="Custom", compound_name="Porphyra-334?",
+        compound_name_old=compound_name, formula="C14H22N2O8", rt = 8.1,
+        mz=347.1449, ionization_form="[M+H]", charge=1) %>%
+  add_row(compound_type="Custom", compound_name="Mycosporine-2-glycine?",
+        compound_name_old=compound_name, formula="C12H18N2O7", rt = 11.3,
+        mz=303.1186, ionization_form="[M+H]", charge=1) %>%
+  add_row(compound_type="Custom", compound_name="Palythene/Usujirene?",
+        compound_name_old=compound_name, formula="C13H20N2O5", rt = 8.3,
+        mz=285.1445, ionization_form="[M+H]", charge=1) %>%
+  add_row(compound_type="Custom", compound_name="Also Palythene/Usujirene?",
+        compound_name_old=compound_name, formula="C13H20N2O5", rt = 6.9,
+        mz=285.1445, ionization_form="[M+H]", charge=1) %>%
+  add_row(compound_type="Custom", compound_name="Mycosporine-glycine?",
+        compound_name_old=compound_name, formula="C10H15NO6", rt = 10,
+        mz=246.0972, ionization_form="[M+H]", charge=1) %>%
+  add_row(compound_type="Custom", compound_name="Also mycosporine-glycine?",
+        compound_name_old=compound_name, formula="C10H15NO6", rt = 9.5,
+        mz=246.0972, ionization_form="[M+H]", charge=1) %>%
   mutate(date_added=strptime(x = date_added, format = "%y%m%d"))

### 2022-10-13 metadata/metadata_control.Rmd
Adding culture data to metadata

@@
 ```
+
+## Culture data ----
+
+```{r}
+clean_stans <- read.csv(file = "made_data/clean_stans.csv")
+
+culture_desc <- readxl::read_xlsx("raw_data/Heal_mSystems uncorrected Metabs_SuppTabs_20210223.xlsx",
+                  sheet = "S7_Culturedescriptions_full")
+culture_data <- readxl::read_xlsx("raw_data/Heal_mSystems uncorrected Metabs_SuppTabs_20210223.xlsx",
+                                  sheet = "S5_FullEnviroCulture_PeakAreas")
+taxa_metabs <- culture_data %>%
+  filter(!str_detect(MassFeature_Column, "RP$")) %>%
+  mutate(polarity=tolower(str_extract(MassFeature_Column, "Neg$|Pos$"))) %>%
+  select(Identification, polarity, `7803_B`:Nmar_3) %>%
+  pivot_longer(cols = `7803_B`:Nmar_3) %>%
+  left_join(clean_stans %>% select(compound_name, compound_name_old, polarity), 
+            by=c(Identification="compound_name_old", "polarity")) %>%
+  filter(!is.na(compound_name)) %>%
+  left_join(culture_desc %>% select(name=`Sample ID`, taxa=`Broad taxon`, species=Species),
+            by="name") %>%
+  filter(!is.na(value)) %>%
+  select(compound_name, polarity, broad_taxa=taxa, species, id=name, value)
+
+write.csv(taxa_metabs, file = "made_data/taxa_metabs.csv")
+```

### 2022-10-13 manuscript/manuscript.Rmd
Reformatting clockplot and adding metadata

@@
 ---
-
-Hi Regina
 
@@
 ```{r posthoc_classes and posthoc_table}
+long_MT_meta <- filled_file_metadata %>%
+  filter(samp_type=="Smp") %>%
+  filter(cruise=="MT") %>%
+  select(filename, no23_um:euk_abund) %>%
+  pivot_longer(cols = no23_um:euk_abund, names_to = "feature", values_to = "norm_area")
 kw_diff_feats <- MT_peaks %>%
+  bind_rows(long_MT_meta) %>%
   left_join(filled_file_metadata, by="filename") %>% 
@@
 rank_peaks <- MT_peaks %>%
+  bind_rows(long_MT_meta) %>%
   group_by(feature) %>%
@@
   pivot_wider(names_from=depth, values_from=mean_fold_change) %>%
+  mutate(env_cmpd=ifelse(startsWith(feature, "MT"), FALSE, TRUE)) %>%
   left_join(posthoc_classes, by="feature") %>%
@@
 axis_y_dodge <- c(2, 2, -2)*clock_scale/25
+
+env_label_df <- clockplot_df %>%
+  filter(!startsWith(feature, "MT")) %>%
+  mutate(xend=x*1.3, yend=y*1.3) %>%
+  mutate(xend=case_when(
+    feature=="DON_um"~-0.5,
+    feature=="syn_abund"~-0.3,
+    feature=="DOP_um"~-0.1,
+    feature=="pro_abund"~0.1,
+    feature=="PC_um"~0.3,
+    feature=="het_abund"~0.5,
+    feature=="PN_um"~0.5,
+    feature=="beam_atten"~0.7,
+    TRUE~xend
+  )) %>%
+  mutate(yend=case_when(
+    feature=="DON_um"~0.7,
+    feature=="syn_abund"~0.8,
+    feature=="DOP_um"~0.7,
+    feature=="pro_abund"~0.8,
+    feature=="PC_um"~0.7,
+    feature=="het_abund"~0.8,
+    feature=="PN_um"~0.6,
+    feature=="beam_atten"~0.5,
+    TRUE~yend
+  ))
 
@@
   geom_point(aes(x=x, y=y, color=resp_class), alpha=0.8) +
-  coord_fixed(xlim = c(-1.1, 1.3)*clock_scale, ylim = c(-1.1, 1.1)*clock_scale) +
+  geom_segment(aes(x=x, y=y, xend=xend, yend=yend), data = env_label_df) +
+  geom_label(aes(x=xend, y=yend, color=resp_class, label=feature), 
+                   data = env_label_df, min.segment.length = 0) +
+  # geom_label_repel(aes(x=x, y=y, color=resp_class, label=feature), 
+  #                  data = env_label_df, min.segment.length = 0) +
+  coord_fixed(xlim = c(-1.1, 1.3)*clock_scale, ylim = c(-1.2, 1.2)*clock_scale) +
   scale_color_manual(values = color_scale, breaks = names(color_scale)) +
@@
   # slice(5) %>%
-  summarize(MSMS=paste(fragmz, int, sep = ", ", collapse = "; "))
+  summarize(premz=mean(premz), MSMS=paste(fragmz, int, sep = ", ", collapse = "; "))
 

### 2022-10-14 metadata/metadata_control.Rmd
Adding MAAs to known compounds and untagging acetyllysine

@@
           compound_name_old=compound_name, formula="C12H20N2O6", rt = 10.5,
-          mz=289.1393, ionization_form="[M+H]", charge=1) %>%
+          mz=289.1393, ionization_form="[M+H]", charge=1, polarity="pos") %>%
   add_row(compound_type="Custom", compound_name="Shinorine",
         compound_name_old=compound_name, formula="C13H20N2O8", rt = 11.6,
-        mz=333.1292, ionization_form="[M+H]", charge=1) %>%
+        mz=333.1292, ionization_form="[M+H]", charge=1, polarity="pos") %>%
   add_row(compound_type="Custom", compound_name="Palythine",
         compound_name_old=compound_name, formula="C10H16N2O5", rt = 11.0,
-        mz=245.1132, ionization_form="[M+H]", charge=1) %>%
+        mz=245.1132, ionization_form="[M+H]", charge=1, polarity="pos") %>%
   add_row(compound_type="Custom", compound_name="Porphyra-334?",
         compound_name_old=compound_name, formula="C14H22N2O8", rt = 8.1,
-        mz=347.1449, ionization_form="[M+H]", charge=1) %>%
+        mz=347.1449, ionization_form="[M+H]", charge=1, polarity="pos") %>%
   add_row(compound_type="Custom", compound_name="Mycosporine-2-glycine?",
         compound_name_old=compound_name, formula="C12H18N2O7", rt = 11.3,
-        mz=303.1186, ionization_form="[M+H]", charge=1) %>%
+        mz=303.1186, ionization_form="[M+H]", charge=1, polarity="pos") %>%
   add_row(compound_type="Custom", compound_name="Palythene/Usujirene?",
         compound_name_old=compound_name, formula="C13H20N2O5", rt = 8.3,
-        mz=285.1445, ionization_form="[M+H]", charge=1) %>%
+        mz=285.1445, ionization_form="[M+H]", charge=1, polarity="pos") %>%
   add_row(compound_type="Custom", compound_name="Also Palythene/Usujirene?",
         compound_name_old=compound_name, formula="C13H20N2O5", rt = 6.9,
-        mz=285.1445, ionization_form="[M+H]", charge=1) %>%
+        mz=285.1445, ionization_form="[M+H]", charge=1, polarity="pos") %>%
   add_row(compound_type="Custom", compound_name="Mycosporine-glycine?",
         compound_name_old=compound_name, formula="C10H15NO6", rt = 10,
-        mz=246.0972, ionization_form="[M+H]", charge=1) %>%
+        mz=246.0972, ionization_form="[M+H]", charge=1, polarity="pos") %>%
   add_row(compound_type="Custom", compound_name="Also mycosporine-glycine?",
         compound_name_old=compound_name, formula="C10H15NO6", rt = 9.5,
-        mz=246.0972, ionization_form="[M+H]", charge=1) %>%
+        mz=246.0972, ionization_form="[M+H]", charge=1, polarity="pos") %>%
   mutate(date_added=strptime(x = date_added, format = "%y%m%d"))
@@
 
-## Culture data ----
+## Culture data
 

### 2022-10-14 untargeted/output_control.Rmd
Adding MAAs to known compounds and untagging acetyllysine

@@
 
-polarity <- "neg"
-cruise <- "MC"
+polarity <- "pos"
+cruise <- "MT"
 output_folder <- paste0(c(cruise, polarity, "output/"), collapse = "_")
@@
   filter(polarity==!!polarity) %>% 
-  filter(run_date[1]>=date_added)
+  filter(run_date[1]>=date_added | is.na(date_added))
 

### 2022-10-14 untargeted/scripts/standard_assignments.R

@@ -323,7 +323,7 @@ if(cruise=="MT" & polarity=="pos") {
     rmAnno("4-Aminobutyric acid") %>%
     reAnno("Allopurinol", new_feature_num = "FT0359") %>%
     reAnno("N-Methyltaurine", new_feature_num = "FT0385") %>%
-    reAnno(cmpd_name = "N6-Acetyl-L-lysine", new_feature_num = "FT0759") %>%
+    reAnno("N6-Acetyl-L-lysine", new_feature_num = "FT0760") %>%
     reAnno("Nicotinic acid", new_feature_num = "FT0252") %>%
     reAnno("S-Adenosylmethionine", new_feature_num = "FT1537") %>%
     rmAnno("O-Acetyl-L-serine") %>%

### 2023-05-04 targeted/output_control.Rmd
Adding targeted nM outputs

@@
   distinct()
+file_metadata <- read.csv("../metadata/made_data/file_metadata.csv")
 ```
@@
 
+```{r write MESOSCOPE to Excel}
+library(openxlsx)
+wb <- createWorkbook(creator = "William Kumler, wkumler@uw.edu")
+addWorksheet(wb, sheetName = "Sample data")
+bind_rows(mesotransect_data, mesocenter_data) %>%
+  pivot_wider(names_from = filename, values_from = nM) %>%
+  writeData(wb = wb, sheet = "Sample data")
+freezePane(wb, sheet = "Sample data", firstRow = TRUE, firstCol = TRUE)
+addWorksheet(wb, "File metadata")
+file_metadata %>%
+  filter(cruise%in%c("MT", "MC")) %>%
+  mutate(filename=str_remove(filename, ".mzML")) %>%
+  distinct(filename, samp_type, station, cast, depth, abs_depth, rosette,date_run, timestamp) %>%
+  rename(sample_name=filename, file_timestamp=timestamp) %>%
+  writeData(wb = wb, sheet = "File metadata")
+addWorksheet(wb = wb, sheetName = "Compound metadata")
+clean_stans %>%
+  select(compound_name, compound_type, formula, rt, mz, kegg_id, polarity, peak_quality) %>%
+  writeData(wb = wb, sheet = "Compound metadata")
+
+addWorksheet(wb=wb, sheetName = "README")
+data.frame(`README for Ingalls MESOSCOPE data`=c(
+  "This file was produced by William Kumler in the Ingalls lab at the University of Washington",
+  "This file contains concentrations of compounds in seawater measured via HPLC-MS",
+  "in a targeted (selected subset of known compounds) method.",
+  "",
+  "The first sheet (Sample data) contains the concentrations of various molecules in each sample",
+  "taken during the MESOSCOPE cruise in 2017. These values are in nanomoles per",
+  "liter of seawater (nM) from the particulate phase (particles collected onto",
+  "0.2um Durapore filters).",
+  "Missing values are shown as empty cells and infinite values (usually in the standards) are shown as #NUM!",
+  "",
+  "The second sheet (File metadata) contains information about the samples themselves, including",
+  "the depth, time, and station from which they were collected. ",
+  "sample_name = Name of the sample as shown in the Sample data Excel sheet",
+  "samp_type = Type of sample, whether biological sample (Smp), blank (Blk), pooled (Poo), or standard (Std)",
+  "station = Station from which the sample was collected, 1-15 plus L1 and L2",
+  "cast = Which cast the sample was taken from",
+  "depth = Depth descriptor of where in the water column the sample was collected",
+  "abs_depth = Absolute depth. Accounts for differences from DCM and is determined from the CTD itself",
+  "rosette = Which Niskin was this sample taken from? Largely incomplete",
+  "date_run = The date on which the HPLC-MS run was started",
+  "file_timestamp = The time at which the sample was actually run",
+  "",
+  "The third sheet (Compound metadata) contains information about the compounds measured",
+  "during this analysis. Not all compounds were detected and passed quality control",
+  "so there are more compounds here than found in the Sample data sheet.",
+  "compound_name = Name of the compound, as found in the Sample data sheet",
+  "compound_type = Broad compound class as determined rather arbitrarily by myself",
+  "formula = Molecular formula of the compound. Isotopes are labeled in parentheses",
+  "rt = Estimated retention time of the compound, in minutes",
+  "mz = Mass-to-charge ratio of the molecule at which it's detected in the mass spec",
+  "kegg_id = KEGG (Kyoto Encyclopedia of Genes and Genomes) ID: https://www.genome.jp/kegg/",
+  "polarity = The polarity at which the molecule was detected in the mass spec, either 'pos' for positive mode or 'neg' for negative mode",
+  "peak_quality = The typical quality of the chromatographic peak. Yes is good, No is bad, and see the Ingalls Standards repo for more details on Problematic compounds",
+  "",
+  paste("File created on", Sys.Date())
+)) %>%
+  writeData(wb = wb, sheet = "README")
+saveWorkbook(wb, file = "MESOSCOPE_Targeted_nM.xlsx", overwrite = TRUE)
+```
+
+```{r write Falkor to Excel}
+library(openxlsx)
+wb <- createWorkbook(creator = "William Kumler, wkumler@uw.edu")
+addWorksheet(wb, sheetName = "Sample data")
+falkor_data %>%
+  pivot_wider(names_from = filename, values_from = nM) %>%
+  writeData(wb = wb, sheet = "Sample data")
+freezePane(wb, sheet = "Sample data", firstRow = TRUE, firstCol = TRUE)
+addWorksheet(wb, "File metadata")
+file_metadata %>%
+  filter(cruise=="FK") %>%
+  mutate(filename=str_remove(filename, ".mzML")) %>%
+  distinct(filename, samp_type, station, cast, depth, abs_depth, rosette, date_run, timestamp) %>%
+  rename(sample_name=filename, file_timestamp=timestamp) %>%
+  writeData(wb = wb, sheet = "File metadata")
+addWorksheet(wb = wb, sheetName = "Compound metadata")
+clean_stans %>%
+  select(compound_name, compound_type, formula, rt, mz, kegg_id, polarity, peak_quality) %>%
+  writeData(wb = wb, sheet = "Compound metadata")
+
+addWorksheet(wb=wb, sheetName = "README")
+data.frame(`README for Ingalls Falkor data`=c(
+  "This file was produced by William Kumler in the Ingalls lab at the University of Washington",
+  "This file contains concentrations of compounds in seawater measured via HPLC-MS",
+  "in a targeted (selected subset of known compounds, manually integrated) method.",
+  "",
+  "The first sheet (Sample data) contains the concentrations of various molecules in each sample",
+  "taken during the Falkor cruise in 2018. These values are in nanomoles per",
+  "liter of seawater (nM) from the particulate phase (particles collected onto",
+  "0.2um Durapore filters).",
+  "Missing values are shown as empty cells and infinite values (usually in the standards) are shown as #NUM!",
+  "",
+  "The second sheet (File metadata) contains information about the samples themselves, including",
+  "the depth, time, and station from which they were collected. ",
+  "sample_name = Name of the sample as shown in the Sample data Excel sheet",
+  "samp_type = Type of sample, whether biological sample (Smp), blank (Blk), pooled (Poo), or standard (Std)",
+  "station = Station from which the sample was collected",
+  "cast = Which cast the sample was taken from",
+  "depth = Depth descriptor of where in the water column the sample was collected",
+  "abs_depth = Absolute depth. Accounts for differences from DCM and is determined from the CTD itself",
+  "rosette = Which Niskin was this sample taken from? Largely incomplete",
+  "date_run = The date on which the HPLC-MS run was started",
+  "file_timestamp = The time at which the sample was actually run",
+  "",
+  "The third sheet (Compound metadata) contains information about the compounds measured",
+  "during this analysis. Not all compounds were detected and passed quality control",
+  "so there are more compounds here than found in the Sample data sheet.",
+  "compound_name = Name of the compound, as found in the Sample data sheet",
+  "compound_type = Broad compound class as determined rather arbitrarily by myself",
+  "formula = Molecular formula of the compound. Isotopes are labeled in parentheses",
+  "rt = Estimated retention time of the compound, in minutes",
+  "mz = Mass-to-charge ratio of the molecule at which it's detected in the mass spec",
+  "kegg_id = KEGG (Kyoto Encyclopedia of Genes and Genomes) ID: https://www.genome.jp/kegg/",
+  "polarity = The polarity at which the molecule was detected in the mass spec, either 'pos' for positive mode or 'neg' for negative mode",
+  "peak_quality = The typical quality of the chromatographic peak. Yes is good, No is bad, and see the Ingalls Standards repo for more details on Problematic compounds",
+  "",
+  paste("File created on", Sys.Date())
+)) %>%
+  writeData(wb = wb, sheet = "README")
+saveWorkbook(wb, file = "Falkor_Targeted_nM.xlsx", overwrite = TRUE)
+```
+
 # Perform some quality checks

### 2023-05-04 metadata/metadata_control.Rmd
Removing rownames from taxa metabs
Refining metadata, moving custom stans to own spreadsheet

--- 2022-10-13 metadata/metadata_control.Rmd
@@
 }
-if(!file.exists("raw_data/MESO_SCOPE_Cruise_KM1709_Sample_Log.xlsx")){
-  stop("Go download the data from email, search 'FALKOR- Sample Log_INGALLS'")
-  # Needs to be converted to .xlsx and downloaded
-}
 bottle_depth_data <- read_csv("made_data/bottle_depth_data.csv")
@@
 ```{r file_metadata from mzML list}
-samp_metadata <- list.files("../mzMLs/pos", pattern = "Smp.*mzML") %>%
+samp_metadata <- list.files("../mzMLs", pattern = "Smp.*mzML", recursive = TRUE) %>%
   data.frame(filename=.) %>%
+  mutate(polarity=dirname(filename)) %>%
+  mutate(filename=basename(filename)) %>%
   mutate(samp_type="Smp") %>%
@@
 
-extra_metadata <- list.files("../mzMLs/pos", pattern = "Blk|Std|Poo.*mzML") %>%
+extra_metadata <- list.files("../mzMLs", pattern = "Blk|Std|Poo.*mzML", recursive = TRUE) %>%
   data.frame(filename=.) %>%
+  mutate(polarity=dirname(filename)) %>%
+  mutate(filename=basename(filename)) %>%
   mutate(samp_type=str_extract(filename, "Std|Blk|Poo")) %>%
@@
 if(!file.exists("made_data/file_timestamps.csv")){
-  list.files("../mzMLs/pos", pattern = "mzML", full.names = TRUE) %>%
+  bind_rows(samp_metadata, extra_metadata) %>%
+    mutate(filepath=paste("../mzMLs/", polarity, filename, sep = "/")) %>%
+    pull(filepath) %>%
     grabMSdata(grab_what = "metadata") %>%
@@
   # Just reordering, nothing getting dropped
-  select(c("filename", "samp_type", "cruise", "station", "cast", "depth", 
-           "abs_depth", "rosette", "tripl", "id", "untripl", "date_run", 
-           "timestamp"))
+  select(c("filename", "polarity", "samp_type", "cruise", "station", "cast",
+           "depth", "abs_depth", "rosette", "tripl", "id", "untripl",
+           "date_run", "timestamp"))
 write.csv(file_metadata, file = "made_data/file_metadata.csv", row.names = FALSE)
@@
 
-```{r heatmaply metadata}
+```{r heatmaply metadata, purl=FALSE}
 filled_MT_metadata <- filled_file_metadata %>%
@@
 
-```{r ggally pairsplot}
-filled_MT_metadata <- filled_file_metadata %>%
-  filter(cruise=="MT") %>%
-  mutate(depth=factor(depth, levels = c("15m", "DCM", "175m")))
-
-library(GGally)
-filled_MT_metadata %>%
-  select(station, depth, abs_depth, time, lat, sla, local_hour, alk, dic, ph, 
-         ends_with("um"), ends_with("ug"), ends_with("abund")) %>%
-  ggpairs(aes(color=depth)) %>%
-  ggsave(filename = "MT_ggpairsplot.pdf", device = "pdf", path = "figures", 
-         width = 30, height = 30, units = "in", dpi = 300)
-```
-
-## PCA groupings
-
-```{r pca, eval=FALSE}
-filled_MT_metadata <- read_csv("made_data/filled_file_metadata.csv") %>%
-  filter(cruise=="MT") %>%
-  filter(samp_type=="Smp") %>%
-  mutate(depth=factor(depth, levels = c("15m", "DCM", "175m")))
-
-pca_output <- filled_MT_metadata %>%
-  select(filename, abs_depth, time, lat, sla, local_hour, #alk, dic, ph, 
-         ends_with("abund"), ends_with("um"), ends_with("ug"),
-         beam_atten, -oxy_um) %>%
-  column_to_rownames("filename") %>%
-  data.matrix() %>%
-  scale() %>%
-  prcomp()
-# plot(pca_output)
-biplot(pca_output)
-biplot(pca_output, choices = c(3, 4))
-
-ggpca_data <- pca_output$x %>%
-  as.data.frame() %>%
-  rownames_to_column("filename") %>%
-  select(filename, PC1:PC4) %>%
-  left_join(filled_MT_metadata)
-ggpca_arrows <- pca_output$rotation %>% 
-  `*`(10) %>%
-  as.data.frame() %>%
-  rownames_to_column("env_param")
-ggpca_labels <- ggpca_arrows %>% 
-  mutate(PC2=PC2+0.2) %>%
-  mutate(PC2=case_when(
-    env_param=="sla"~PC2-0.4,
-    env_param=="PC_um"~PC2+0.2,
-    env_param=="phaeo_ug"~PC2+0.2,
-    TRUE~PC2)) %>%
-  mutate(PC4=PC4+0.3)
-ggplot(mapping = aes(x=PC1, y=PC2)) + 
-  geom_point(data = ggpca_data, aes(color=depth)) +
-  geom_segment(data = ggpca_arrows, aes(xend=0, yend=0), 
-               arrow = arrow(length = unit(0.02, "npc"), ends = "first")) +
-  geom_label(data = ggpca_labels, aes(label=env_param)) +
-  theme_bw() +
-  theme(legend.position = c(0.8, 0.8))
-ggsave(filename = "MT_PC1_PC2.png", device = "png", path = "figures", 
-       width = 5, height = 5, dpi = 144, units = "in", type="cairo")
-ggplot(mapping=aes(x=PC3, y=PC4)) + 
-  geom_point(data=ggpca_data, aes(color=sla, size=lat)) +
-  geom_segment(data = ggpca_arrows, aes(xend=0, yend=0), 
-               arrow = arrow(length = unit(0.02, "npc"), ends = "first")) +
-  geom_label(data = ggpca_labels, aes(label=env_param)) +
-  theme_bw()
-ggsave(filename = "MT_PC3_PC4.png", device = "png", path = "figures", 
-       width = 5.5, height = 5, dpi = 144, units = "in", type="cairo")
-```
-
 ## Get standards and clean up
@@
 }
+if(!file.exists("raw_data/custom_stans.csv")){
+  custom_stans <- read.csv(paste0(
+    "https://raw.githubusercontent.com/IngallsLabUW/Ingalls_Standards",
+    "/master/data_extra/custom_stans.csv"
+  ))
+  write.csv(custom_stans, file = "raw_data/custom_stans.csv", row.names = FALSE)
+}
 raw_stans <- read.csv("raw_data/raw_stans.csv")
+custom_stans <- read.csv("raw_data/custom_stans.csv")
 clean_stans <- raw_stans %>%
-  filter(Column=="HILIC") %>%
+  bind_rows(custom_stans) %>%
+  filter(Column=="HILIC"|Compound_Type=="Custom") %>%
   mutate(polarity=ifelse(str_detect(z, "-"), "neg", "pos")) %>%
@@
          peak_quality) %>%
-  add_row(compound_type="Custom", compound_name="Bisulfate",
-          compound_name_old="Bisulfate", formula="HSO4", rt=13, 
-          mz=95.95172965+1.007276, ionization_form="[M]", 
-          charge=-1, date_added=150101, polarity="neg") %>%
-  add_row(compound_type="Custom", compound_name="Bromine81",
-          compound_name_old="Bromine upper", formula="Br(81)", rt=7, 
-          mz=80.91563, ionization_form="[M]", charge=-1, 
-          date_added=150101, polarity="neg") %>%
-  add_row(compound_type="Custom", compound_name="Bromine79",
-          compound_name_old="Bromine lower", formula="Br(79)", rt=7, 
-          mz=78.91768, ionization_form="[M]", charge=-1, 
-          date_added=150101, polarity="neg") %>%
-  add_row(compound_type="Custom", compound_name="Pyro from glutamine?",
-          compound_name_old=compound_name, formula="C5H7NO3", rt=11, 
-          mz=130.050419, ionization_form="[M-NH3+H]", charge=-1, 
-          date_added=150101, polarity="pos") %>%
-  add_row(compound_type="Custom", compound_name="Pyro from glutamate?",
-          compound_name_old=compound_name, formula="C5H7NO3", rt=12, 
-          mz=130.050419, ionization_form="[M-H2O+H]", charge=-1, 
-          date_added=150101, polarity="pos") %>%
-  # add_row(compound_type="Custom", compound_name="Deoxyguanosine?",
-  #         compound_name_old=compound_name, formula="C10H13N5O4", rt = 4.5,
-  #         mz=268.10403, ionization_form="[M+H]", charge=1, 
-  #         date_added=150101, polarity="pos") %>%
-  add_row(compound_type="Custom", compound_name="Deoxycytidine?",
-          compound_name_old=compound_name, formula="C10H13N5O4", rt = 6.6,
-          mz=228.09788, ionization_form="[M+H]", charge=1, 
-          date_added=150101, polarity="pos") %>%
-  add_row(compound_type="Custom", compound_name="Homoserine betaine?",
-          compound_name_old=compound_name, formula="C7H15NO3", rt = 6.7,
-          mz=162.1125, ionization_form="[M+H]", charge=1, 
-          date_added=150101, polarity="pos") %>%
-  add_row(compound_type="Custom", compound_name="Threonine betaine?",
-          compound_name_old=compound_name, formula="C7H15NO3", rt = 5,
-          mz=162.1125, ionization_form="[M+H]", charge=1, 
-          date_added=150101, polarity="pos") %>%
-  add_row(compound_type="Custom", compound_name="Theanine?",
-          compound_name_old=compound_name, formula="C7H14N2O3", rt = 6.8,
-          mz=175.10771, ionization_form="[M+H]", charge=1, 
-          date_added=150101, polarity="pos") %>%
-  add_row(compound_type="Custom", compound_name="Asterina-330",
-          compound_name_old=compound_name, formula="C12H20N2O6", rt = 10.5,
-          mz=289.1393, ionization_form="[M+H]", charge=1) %>%
-  add_row(compound_type="Custom", compound_name="Shinorine",
-        compound_name_old=compound_name, formula="C13H20N2O8", rt = 11.6,
-        mz=333.1292, ionization_form="[M+H]", charge=1) %>%
-  add_row(compound_type="Custom", compound_name="Palythine",
-        compound_name_old=compound_name, formula="C10H16N2O5", rt = 11.0,
-        mz=245.1132, ionization_form="[M+H]", charge=1) %>%
-  add_row(compound_type="Custom", compound_name="Porphyra-334?",
-        compound_name_old=compound_name, formula="C14H22N2O8", rt = 8.1,
-        mz=347.1449, ionization_form="[M+H]", charge=1) %>%
-  add_row(compound_type="Custom", compound_name="Mycosporine-2-glycine?",
-        compound_name_old=compound_name, formula="C12H18N2O7", rt = 11.3,
-        mz=303.1186, ionization_form="[M+H]", charge=1) %>%
-  add_row(compound_type="Custom", compound_name="Palythene/Usujirene?",
-        compound_name_old=compound_name, formula="C13H20N2O5", rt = 8.3,
-        mz=285.1445, ionization_form="[M+H]", charge=1) %>%
-  add_row(compound_type="Custom", compound_name="Also Palythene/Usujirene?",
-        compound_name_old=compound_name, formula="C13H20N2O5", rt = 6.9,
-        mz=285.1445, ionization_form="[M+H]", charge=1) %>%
-  add_row(compound_type="Custom", compound_name="Mycosporine-glycine?",
-        compound_name_old=compound_name, formula="C10H15NO6", rt = 10,
-        mz=246.0972, ionization_form="[M+H]", charge=1) %>%
-  add_row(compound_type="Custom", compound_name="Also mycosporine-glycine?",
-        compound_name_old=compound_name, formula="C10H15NO6", rt = 9.5,
-        mz=246.0972, ionization_form="[M+H]", charge=1) %>%
   mutate(date_added=strptime(x = date_added, format = "%y%m%d"))
@@
 
-write.csv(taxa_metabs, file = "made_data/taxa_metabs.csv")
-```
+write.csv(taxa_metabs, file = "made_data/taxa_metabs.csv", row.names = FALSE)
+```

### 2023-05-04 untargeted/output_control.Rmd
Converting purl chunks to TRUE

@@
 ```{r setup, include=FALSE, purl=TRUE}
-knitr::opts_chunk$set(echo = TRUE, warning = FALSE, error = TRUE, purl = FALSE)
+knitr::opts_chunk$set(echo = TRUE, warning = FALSE, error = TRUE, purl=TRUE)
 options(dplyr.summarise.inform=F)
@@
 cruise_metadata <- read.csv("../metadata/made_data/file_metadata.csv") %>%
-  filter(cruise==!!cruise)
+  filter(cruise==!!cruise) %>%
+  filter(polarity==!!polarity)
 
@@
 
-```{r peakpicking, error=TRUE, purl=FALSE}
+```{r peakpicking, error=TRUE, purl=TRUE}
 library(xcms)
@@
 
-```{r deisoadduct, error=TRUE, purl=FALSE}
+```{r deisoadduct, error=TRUE, purl=TRUE}
 library(xcms)
@@
 
-```{r assignformulae, purl=FALSE}
+```{r assignformulae, purl=TRUE}
 M_area_peaks <- read.csv(paste0(output_folder, "M_area_peaks.csv"))
@@
 
-```{r assignids}
-filled_peaks <- read.csv(paste0(output_folder, "filled_peaks.csv"))
+```{r assignids, purl=FALSE}
+filled_peaks <- read.csv(paste0(output_folder, "M_area_peaks.csv"))
 feature_envelopes <- read.csv(paste0(output_folder, "feature_envelopes.csv"))
@@
   group_by(feature) %>%
-  summarise(mzmed=median(mz), rtmed=median(rt), avg_area=mean(M_area)) %>%
-  left_join(feature_envelopes, by = "feature")
+  summarise(mzmed=median(mz), rtmed=median(rt), avg_area=mean(M_area))
 stan_annotations <- read.csv(paste0(output_folder, "stan_annotations.csv"))
+
+want_stans <- c("MT_POS_FT0489", "MT_POS_FT0519", "MT_POS_FT0525", "MT_POS_FT0567", 
+"MT_POS_FT0584", "MT_POS_FT0720", "MT_POS_FT0862", "MT_POS_FT0973", 
+"MT_POS_FT0433", "MT_POS_FT0618", "MT_POS_FT0664", "MT_POS_FT0691", 
+"MT_POS_FT0744", "MT_POS_FT0607", "MT_POS_FT0617", "MT_POS_FT1113", 
+"MT_POS_FT1178")
+feature_data <- feature_data %>% filter(feature%in%str_remove(want_stans, "MT_POS_"))
 

### 2023-05-04 metadata/metadata_control.Rmd
Cleaning up metadata pre-rerun

@@
 ---
-title: "Metadata control"
+title: "Metadata control document"
 author: "William Kumler"
-date: "5/5/2022"
+date: "5/5/2023"
 output: html_document
@@
 options(pillar.sigfig = 7)
-```
-
-```{r functions}
+
 toDecDeg <- function(deg, min, dir){
@@
 ```{r station_metadata}
-# DCMs are entered manually ----
+# DCM depths are entered manually ----
 if(!file.exists("raw_data/all_DCMs.csv")){
@@
 ```{r}
+download.file("https://datadryad.org/stash/downloads/file_stream/2192675", 
+              destfile = "raw_data/Heal_mSystems_Supplement.xlsx", 
+              mode = "wb")
 clean_stans <- read.csv(file = "made_data/clean_stans.csv")
 
-culture_desc <- readxl::read_xlsx("raw_data/Heal_mSystems uncorrected Metabs_SuppTabs_20210223.xlsx",
+culture_desc <- readxl::read_xlsx("raw_data/Heal_mSystems_Supplement.xlsx",
                   sheet = "S7_Culturedescriptions_full")
-culture_data <- readxl::read_xlsx("raw_data/Heal_mSystems uncorrected Metabs_SuppTabs_20210223.xlsx",
+culture_data <- readxl::read_xlsx("raw_data/Heal_mSystems_Supplement.xlsx",
                                   sheet = "S5_FullEnviroCulture_PeakAreas")

### 2023-05-15 metadata/metadata_control_v2.Rmd
Implementing new version of metadata collection and cleaning

---
title: "Metadata control document"
author: "William Kumler"
date: "5/5/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

library(tidyverse)
library(readxl)     # For "readxl"

scope_uname <- "scope"
scope_pw <- "SCOPE2014"

options(pillar.sigfig = 7)
```

## Download/collect raw data

Full list of files we need!
From SCOPE website:
  - FK_CTDsummary_current.txt
  - MS_CTDsummary_current.txt
  - Karl_FK_PCPN_WC_current.xlsx
  - Karl_FK_Nutrients_Final.xlsx
  - 2017MS_CoreData_current.xls
  - Karl_MS_Nutrient_Summary_KB.xlsx
  - Karl_MS_pATP.xlsx
  - Karl_FK_chlorophyll.xlsx
From email
  - FALKOR- Sample Log_INGALLS (From Tim)
  - MESO_SCOPE_Cruise_KM1709_Sample_Log (From Laura)
  - MESO-SCOPE_BeamAttenuation_forWill (From Benedetto, as are remainder)
  - MESO-SCOPE_DailyPAR_forWill
  - MESO-SCOPE_PercentPAR_forWill
  - kok1607_ctd_sla
  - km1709_ctd_sla
  - fk180310_ctd_sla
Manually created
  - hmdb_formulae.rds
  - MAA_masses.xlsx

```{r From SCOPE data inventory, eval = !file.exists("raw_data/FK_CTDsummary_current.txt")}
inventory_url <- paste0("http://", scope_uname, ":", scope_pw,
                        "@scope.soest.hawaii.edu/collaborators/datainventory/Data/")
from_scope_inventory <- c(
  "SCOPEcore/FK_CTDsummary_current.txt",
  "SCOPEcore/MS_CTDsummary_current.txt",
  "Karl/Karl_FK_PCPN_WC_current.xlsx",
  "Karl/Karl_FK_Nutrients_Final.xlsx",
  "SCOPEcore/2017MS_CoreData_current.xls",
  "Karl/Karl_MS_Nutrient_Summary_KB.xlsx",
  "Karl/Karl_MS_pATP.xlsx",
  "Karl/Karl_FK_Chlorophyll.xlsx"
)
sapply(from_scope_inventory, function(end_url){
  download.file(paste0(inventory_url, end_url), destfile = paste0("raw_data/", basename(end_url)), mode = "wb")
})
```

```{r From SCOPE FTP, eval=!file.exists("raw_data/mesoscope.gof")}
ftp_url <- "http://scope.soest.hawaii.edu/FTP/scope/water/"
from_scope_ftp <- c("mesoscope.gof","fk180310.gof")
sapply(from_scope_ftp, function(end_url){
  download.file(paste0(ftp_url, end_url), destfile = paste0("raw_data/", basename(end_url)))
})
```

```{r From email}
# Go search for the files in email, we can't script this part
```

```{r Manually created}
# Go search for the files in Slack, we can't script this part
```

## SCOPE data cleaning and combining

```{r Sample log and DCM extraction}
col_types <- c("date", "text", "numeric", "text",
               "text", "date", "date", "text")
col_names <- c("date", "station", "cast", "sample_id", "depth",
               "start_time", "end_time", "volume")
falkor_log <- c("Anti-Cyclonic Eddy", "Cyclonic Eddy") %>%
  lapply(function(sheet_i){
    read_xlsx("raw_data/FALKOR- Sample Log_INGALLS.xlsx",
             sheet=sheet_i, skip = 5, 
             col_types = col_types, col_names = col_names) 
}) %>%
  bind_rows() %>%
  drop_na() %>%
  filter(sample_id!="1") %>%
  mutate(date=as.Date(date)) %>%
  mutate(abs_depth=as.numeric(str_extract(depth, "\\d+"))) %>%
  mutate(str_depth=ifelse(startsWith(depth, "DCM"), "DCM", paste0(depth, "m"))) %>%
  mutate(depth=ifelse(depth=="25", "Surface", "DCM")) %>%
  mutate(volume=as.numeric(str_extract(volume, "\\d+"))) %>%
  mutate(sample_id=paste0("S", station, "C", cast, "-", str_depth, "_", sample_id)) %>%
  select(-str_depth)

col_types <- c("text", "text", "text", "text",
               "text", "numeric", "date", "date",
               "text", "text", "skip", "skip", "skip")
col_names <- c("date", "station", "cast", "depth", "sample_id",
               "volume", "start_time", "end_time", "rosette", 
               "dissolved")
MT_log <- read_xlsx("raw_data/MESO_SCOPE_Cruise_KM1709_Sample_Log.xlsx",
          sheet = "Metabolites_and_Proteins", skip = 2,
          col_types = col_types, col_names=col_names) %>%
  drop_na() %>%
  mutate(date=ifelse(startsWith(date, "6/29"), "42916.0", date)) %>%
  mutate(date=as.Date(as.numeric(date), origin="1899-12-31")) %>%
  mutate(station=str_extract(station, "\\d+")) %>%
  mutate(cast=as.numeric(str_extract(cast, "\\d+"))) %>%
  mutate(abs_depth=as.numeric(str_extract(depth, "\\d+"))) %>%
  mutate(depth=case_when(
    depth=="15.0"~"Surface", depth=="175.0"~"Deep", TRUE~"DCM")
  ) %>%
  mutate(rosette=case_when(rosette=="?"~NA, rosette=="43026.0"~"10", rosette=="16- 24"~"16")) %>%
  mutate(rosette=as.numeric(rosette)) %>%
  # Fixing likely transcription error - sample looks like 15m samples and
  # all other triplicates were 19, 20, and 21
  mutate(rosette=ifelse(rosette==4 & cast==3 & station=="6", 21, rosette))

col_names <- c("station", "cast", "depth", "sample_id",
               "volume", "start_time", "end_time", "rosette", 
               "dissolved")
col_types <- c("text", "text", "text", "text",
               "numeric", "date", "date", "numeric", 
               "text", "skip")

MC_log <- read_xlsx("raw_data/MESO_SCOPE_Cruise_KM1709_Sample_Log.xlsx",
          sheet = "Metabolites_High_Resolution", skip = 2,
          col_types = col_types, col_names=col_names) %>%
  filter(!is.na(station)) %>%
  mutate(station=str_remove(station, "^S")) %>%
  mutate(cast=as.numeric(str_remove(cast, "^C"))) %>%
  mutate(sample_id=str_replace(sample_id, "\\-", "less")) %>%
  mutate(sample_id=str_replace(sample_id, "\\+", "plus")) %>%
  mutate(depth_shift=str_extract(depth, "(\\-|\\+)\\d+")) %>%
  mutate(depth_shift=ifelse(is.na(depth_shift), 0, as.numeric(depth_shift))) %>%
  mutate(abs_depth=case_when(
    station=="L1"~108+depth_shift,
    station=="L2"~123+depth_shift
  )) %>%
  select(-depth_shift) %>%
  mutate(depth=str_replace(depth, "\\-", "less")) %>%
  mutate(depth=str_replace(depth, "\\+", "plus")) %>%
  mutate(rosette=as.numeric(rosette)) %>%
  mutate(sample_id=str_remove(sample_id, "^MHR")) %>%
  mutate(date=as.Date(case_when(
    station=="L1"~"2018-07-08",
    station=="L2"~"2018-07-09"
  )))

samp_log <- bind_rows(falkor_log, MC_log, MT_log)
dcm_depths <- samp_log %>%
  distinct(station, depth, abs_depth) %>%
  filter(depth=="DCM") %>%
  select(station, dcm_depth=abs_depth)
```

```{r station locations and CTD summaries}
col_names <- c("station", "cast", "month", "day", "year", "time", 
               "lat_deg", "lat_min", "lat_dir", 
                "lon_deg", "lon_min", "lon_dir", "max_depth", "bottles")
col_types <- cols(
  station = col_character(),
  cast = col_double(),
  month = col_character(),
  day = col_character(),
  year = col_double(),
  time = col_time(format = ""),
  lat_deg = col_double(),
  lat_min = col_double(),
  lat_dir = col_character(),
  lon_deg = col_double(),
  lon_min = col_double(),
  lon_dir = col_character(),
  max_depth = col_double(),
  bottles = col_double())
toDecDeg <- function(deg, min, dir){
  (deg+min/60)*ifelse(dir%in%c("S", "W"), -1, 1)
}
parseCTD <- function(file_path){
  file_path %>%
    read_table(skip = 3, col_names = col_names, col_types=col_types) %>%
    mutate(time=strptime(paste(month, day, year, time), format = "%b %d %Y %X")) %>%
    select(-month, -day, -year) %>%
    mutate(lat=mapply(toDecDeg, lat_deg, lat_min, lat_dir)) %>%
    mutate(lon=mapply(toDecDeg, lon_deg, lon_min, lon_dir)) %>%
    select(-matches("^lon_|^lat_"))
}
meso_ctddata <- parseCTD("raw_data/MS_CTDsummary_current.txt")
falkor_ctddata <- parseCTD("raw_data/FK_CTDsummary_current.txt") %>%
  filter(time > "2018-03-23 04:01:51")
complete_ctddata <- bind_rows(meso_ctddata, falkor_ctddata)
```

```{r SLA parsing}
col_types <- cols(
  stn = col_character(),
  cast = col_double(),
  sla_corr = col_double()
)
falkor_sla <- read_csv("raw_data/fk180310_ctd_sla.txt", col_types = col_types) %>%
  mutate(cruise="FK")
meso_sla <- read_csv("raw_data/km1709_ctd_sla.txt", col_types = col_types) %>%
  mutate(cruise=ifelse(stn%in%c("15", "L1", "L2"), "MC", "MT"))
complete_sla <- bind_rows(falkor_sla, meso_sla) %>%
  rename(station="stn")
```

```{r Core water column parsing (.gof files)}
col_names <- c("station", "cast", "rosette", "lat", "lon", "pressure", "temp",
              "sal", "ctd_oxy", "ctd_chl", "theta", "sigma", "oxy_um", "dic", "alk", "ph",
              "po4_um", "no23_um", "sio4_um", "LLN", "LLP", "PC", "PN", "PP",
              "PSi", "chl_ug", "phaeo_ug", "het_abund", 
              "pro_abund", "syn_abund", "euk_abund",
              "q1", "q2", "q3", "q4")
qual_names <- setdiff(col_names, c(head(col_names, 6), tail(col_names, 4)))
gofReader <- function(filename){
  # From ftp://ftp.soest.hawaii.edu/hot/water/Readme.water.jgofs
  # Quality Indicators:
  # Flag  Meaning
  # 1    not quality controled
  # 2    good data
  # 3    suspect (i.e.  questionable) data
  # 4    bad data
  # 5    missing data
  # 9    variable not measured during this cast
  raw_gof <- read_table(filename, skip = 4, col_names = col_names,
                        col_types = cols(.default = col_double()))
  qual_df <- select(raw_gof, matches("^q\\d+$")) %>%
    unite("value", sep = "") %>%
    separate(sep = "", col = "value", into = c(NA, qual_names)) %>%
    mutate(across(everything(), as.numeric))
  raw_gof[,qual_names][qual_df!=2] <- NA
  raw_gof %>% select(-matches("^q\\d+$"))
}
meso_core <- gofReader("raw_data/mesoscope.gof") %>%
  mutate(cruise=case_when(
    station <= 14 ~ "MT",
    station %in% 15:17 ~ "MC",
    TRUE ~ "Unknown"
  )) %>%
  mutate(station=case_when(
    station <= 15 ~ as.character(station),
    station == 16 ~ "L1",
    station == 17 ~ "L2"
  ))
falkor_core <- gofReader("raw_data/fk180310.gof") %>%
  filter(station>28) %>%
  mutate(station=as.character(station)) %>%
  mutate(cruise="FK")
all_core <- bind_rows(meso_core, falkor_core)
```

```{r Parsing beam attenuation data}
beam_long <- read.csv("raw_data/MESO-SCOPE_BeamAttenuation_forWill.csv", 
                    skip = 3, header = FALSE) %>%
  `colnames<-`(paste(.[1,], .[2,], sep = " ")) %>%
  tail(-2) %>%
  mutate(across(everything(), as.numeric)) %>%
  pivot_longer(cols = -1, values_to = "beam_atten") %>%
  filter(!is.na(beam_atten)) %>%
  separate(name, into = c("station", "cast"), sep = " ") %>%
  rename(abs_depth=`NaN NaN`) %>%
  mutate(cast=as.numeric(cast))

# Convert to bottle locations
beam_core <- all_core %>%
  group_by(station, cast) %>%
  # filter(station==4, cast==2) %>%
  group_split() %>% #`[[`(1) -> core_subset
  map(function(core_subset){
    beam_subset <- beam_long %>%
      filter(station==unique(core_subset$station), cast==unique(core_subset$cast))
    if(nrow(beam_subset)==0)return(NULL)
    approxout <- with(beam_subset, approx(abs_depth, beam_atten, core_subset$pressure, ties = mean))
    core_subset %>%
      select(cruise, station, cast, rosette, pressure) %>%
      left_join(set_names(as.data.frame(approxout), c("pressure", "beam_atten")), 
                by="pressure", multiple="all")
  }, .progress = TRUE) %>%
  bind_rows() %>%
  distinct() %>%
  filter(beam_atten<0.2)
```

```{r PCPN parsing}
fk_pcn <- "raw_data/Karl_FK_PCPN_WC_current.xlsx" %>%
  read_excel(skip=1) %>%
  slice(-1:-2) %>% 
  filter(is.na(Comments)) %>%
  mutate(station=str_extract(Name, "(?<=FK-)\\d+")) %>%
  mutate(rosette=as.numeric(str_extract(Name, "\\d+$"))) %>%
  mutate(cruise="FK", cast=1) %>%
  select(cruise, station, cast, rosette, abs_depth=`Depth (m)`, 
         PC_um=`C value (µmol/L)`, PN_um=`N value (µmol/L)`)
meso_pcn <- "raw_data/2017MS_CoreData_current.xls" %>%
  read_excel(skip = 1, sheet = "PCPN") %>%
  slice(-1) %>%
  select(abs_depth=depth_sample, station, PC_um=particulate_carbon, 
         PN_um=particulate_nitrogen, cast=cast_num, rosette=rosette_position) %>%
  mutate(across(everything(), as.numeric)) %>%
  mutate(cruise=case_when(
    station <= 14 ~ "MT",
    station %in% 15:17 ~ "MC",
    TRUE ~ "Unknown"
  )) %>%
  mutate(station=case_when(
    station <= 15 ~ as.character(station),
    station %in% 16 ~ "L1",
    station == 17 ~ "L2"
  )) %>%
  mutate(rosette=case_when(
    station=="L1"&cast==13&rosette==10~20, 
    station=="L1"&cast==6&rosette==15~16,
    station=="L1"&cast==6&rosette==14~15,
    TRUE~rosette))
all_pcn <- bind_rows(meso_pcn, fk_pcn)
```

```{r Nutrient parsing}
# Not really necessary because .gof file has nutrients in it already
# These files have more sigfigs but less depth accuracy (5, 25 instead of 4.7, 25.2)
# Actually, very necessary because the Falkor .gof file has N+N and PO4
# values switched??? Compare to Karl_FK_Nutrients_Final.xlsx
all_core %>%
  filter(cruise=="FK") %>%
  filter(!is.na(no23_um)) %>%
  filter(pressure>160) %>%
  select(station, cast, rosette, pressure, no23_um, po4_um)

falkor_nuts <- "raw_data/Karl_FK_Nutrients_Final.xlsx" %>%
  read_excel(skip = 3) %>%
  select(abs_depth=`Depth`, sample=`Sample ID`, po4_um=`µM PO4`, no23_um=`µM N+N`) %>%
  filter(!is.na(abs_depth)) %>%
  filter(!abs_depth=="Depth") %>%
  mutate(rosette=as.numeric(str_extract(sample, "\\d+$"))) %>% 
  mutate(sample=ifelse(startsWith(sample, "FK"), sample, NA)) %>%
  fill(sample) %>%
  mutate(station=str_extract(sample, "(?<=FK 180310[- ])\\d+")) %>%
  mutate(across(c(abs_depth, po4_um, no23_um, station), as.numeric)) %>%
  filter(station>=28) %>%
  mutate(station=as.character(station)) %>%
  select(-sample) %>%
  mutate(cast=1, cruise="FK") %>%
  arrange(station, abs_depth)

transect_profile_sheet <- "raw_data/Karl_MS_Nutrient_Summary_KB.xlsx" %>%
  read_excel(range = "KM1709 Transect, L1,L2 profiles!A4:N180") %>%
  slice(-1:-2) %>%
  mutate(sd=as.numeric(sd)) %>%
  mutate(Si=case_when(
    Station=="8" & Cast==2 & Niskin==4 ~ NA,
    Station=="10" & Cast==1 & Niskin==3 ~ NA,
    Station=="14" & Cast==1 & Niskin==8 ~ NA,
    TRUE~Si
  )) %>%
  mutate(LLP=case_when(
    Station=="L2" & Cast==26 & Niskin==1 ~ NA,
    TRUE~LLP
  ))
highres_sheet <- "raw_data/Karl_MS_Nutrient_Summary_KB.xlsx" %>%
  read_excel(range = "KM1709 High res nuts!A4:P53") %>%
  slice(-1:-2) %>%
  filter(!is.na(Station))
diel_sheet <- "raw_data/Karl_MS_Nutrient_Summary_KB.xlsx" %>%
  read_excel(range = "KM1709 Diel nuts!A4:O76") %>%
  slice(-1:-3) %>%
  filter(!is.na(Station)) %>%
  mutate(LLP=case_when(
    Station=="L1"&Cast==11&Niskin...3==20~NA,
    Station=="L1"&Cast==19&Niskin...3==9~NA,
    TRUE~LLP
  )) %>%
  mutate(Si=case_when(
    Station=="L1" & Cast==10 & Niskin...3==10 ~ NA,
    Station=="L1" & Cast==15 & Niskin...3==10 ~ NA,
    Station=="L1" & Cast==17 & Niskin...3==11 ~ NA,
    Station=="L2" & Cast==6 & Niskin...3==12 ~ NA,
    Station=="L2" & Cast==12 & Niskin...3==12 ~ NA,
    Station=="L2" & Cast==17 & Niskin...3==12 ~ NA,
    Station=="L2" & Cast==20 & Niskin...3==18 ~ NA,
    Station=="L2" & Cast==25 & Niskin...3==12 ~ NA,
    Station=="L2" & Cast==11 & Niskin...3==3 ~ NA,
    TRUE~Si
  ))

meso_nuts <- bind_rows(transect_profile_sheet, highres_sheet, diel_sheet) %>% 
  select(abs_depth=Depth, station=Station, cast=Cast, rosette=Niskin,
         no23_um=`N+N`, LLN, DON_um=DON, TDN, 
         po4_um=`SRP`, LLP, DOP_um=DOP, TDP,
         si_um=`Si`) %>%
  mutate(no23_um=ifelse(no23_um=="BLD"|no23_um=="NA"|is.na(no23_um), LLN, no23_um)) %>%
  mutate(po4_um=ifelse(po4_um=="BLD"|po4_um=="NA"|is.na(po4_um), LLP, po4_um)) %>%
  select(-LLN, -LLP) %>%
  mutate(across(-station, as.numeric)) %>%
  arrange(station, abs_depth, cast) %>%
  filter(!is.na(abs_depth)) %>%
  mutate(cruise=case_when(
    station %in% as.character(1:14) ~ "MT",
    station %in% c(15, "L1", "L2") ~ "MC",
    TRUE ~ "Unknown"
  )) %>%
  mutate(no23_um=ifelse(is.na(no23_um) & cruise=="MC", 0, no23_um)) %>%
  # Fixing a few rosette numbers
  mutate(rosette=case_when(
    station=="13"&rosette==200~3,
    station=="13"&rosette==250~2,
    station=="13"&rosette==500~1,
    station=="14"&rosette==200~4,
    station=="14"&rosette==250~3,
    station=="14"&rosette==500~2,
    TRUE~rosette
  ))
# For determining the above
# all_core %>% 
#   filter(station%in%c("13", "14")) %>% 
#   select(station, rosette, pressure) %>% 
#   filter(pressure>180)
all_nuts <- bind_rows(falkor_nuts, meso_nuts)
```

```{r Chlorophyll parsing}
meso_chl <- "raw_data/2017MS_CoreData_current.xls" %>%
  read_excel(skip = 1, sheet = "Chl_Pheo") %>%
  slice(-1) %>%
  mutate(filter_min=as.numeric(filter_min)) %>%
  filter(filter_min>0.5) %>%
  select(abs_depth=depth_sample, station, chl_ug=fluorometric_chlorophyll_a, 
         phaeo_ug=total_phaeopigment, cast=cast_num, rosette=rosette_position) %>%
  mutate(across(everything(), as.numeric)) %>%
  mutate(station=case_when(
    station <= 15 ~ as.character(station),
    station == 16 ~ "L1",
    station == 17 ~ "L2"
  )) %>%
  mutate(cruise=case_when(
    station %in% as.character(1:14) ~ "MT",
    station %in% c(15, "L1", "L2") ~ "MC",
    TRUE ~ "Unknown"
  )) %>%
  # Removing one data point because it's negative
  filter(chl_ug>0) %>%
  # Removing one data point because it doesn't have an associated surface value
  filter(!(station=="L1" & cast==19)) %>%
  # Fixing one nonsense rosette number using all_core to find the rosette fired at 15m for S11C2
  mutate(rosette=ifelse(rosette==33, 23, rosette)) %>%
  select(-abs_depth)

falkor_chl <- "raw_data/Karl_FK_Chlorophyll.xlsx" %>%
  read_excel(skip = 19, sheet = "10au") %>%
  slice(-1) %>%
  select(station=Station, cast=Cast, rosette=Bottle, chl_ug=`chla (ug/)l`, phaeo_ug=`phaeo (ug/l)`, abs_depth=Depth) %>%
  mutate(cruise="FK") %>%
  select(-abs_depth) %>%
  mutate(station=as.character(station))

all_chl <- bind_rows(meso_chl, falkor_chl)
```

```{r PARsing}
light_long <- read.csv("raw_data/MESO-SCOPE_PercentPAR_forWill.csv", 
                    skip = 7, header = FALSE) %>%
  `colnames<-`(c("abs_depth", paste0("L1_", 1:6), paste0("L2_", 1:6))) %>%
  mutate(across(everything(), as.numeric)) %>%
  pivot_longer(cols = -1, values_to = "beam_atten") %>%
  filter(!is.na(beam_atten)) %>%
  separate(name, into = c("station", "rep"), sep = "_") %>%
  mutate(rep=as.numeric(rep))
```

```{r Create complete_env_data by doing hella joins}
complete_env_data <- all_core %>%
  select(cruise, station, cast, rosette, lat, lon, pressure,
         temp, sal, ctd_oxy, ctd_chl, theta, sigma, oxy_um,
         dic, alk, ph, ends_with("abund")) %>%
  left_join(dcm_depths) %>%
  left_join(complete_ctddata %>% select(station, cast, time)) %>%
  # filter(station=="1", cast==1) %>%
  left_join(complete_sla) %>%
  left_join(all_pcn) %>% 
  left_join(all_nuts %>% select(-abs_depth)) %>%
  left_join(all_chl) %>%
  left_join(beam_core) %>%
  mutate(abs_depth=ifelse(is.na(abs_depth), pressure, abs_depth))
```

## MS file metadata creation and import

```{r Create file data for samples}
samp_filedata <- list.files("../mzMLs", pattern = "Smp.*mzML", recursive = TRUE) %>%
  data.frame(filename=.) %>%
  mutate(polarity=dirname(filename)) %>%
  mutate(filename=basename(filename)) %>%
  mutate(samp_type="Smp") %>%
  mutate(cruise=case_when(
    str_detect(filename, "180205") ~ "MC",
    str_detect(filename, "180821") ~ "MT",
    str_detect(filename, "190715") ~ "FK",
    TRUE ~ "Unknown"
  )) %>%
  mutate(station=str_extract(filename, "((?<=S)\\d+)|L\\d")) %>%
  mutate(cast=case_when(
    cruise=="MC" ~ str_extract(filename, "(?<=C)\\d+"),
    cruise=="MT" ~ str_extract(filename, "(?<=C)\\d"),
    cruise=="FK" ~ "1",
  )) %>%
  mutate(cast=as.numeric(cast)) %>%
  mutate(depth=str_extract(filename, "15m|25m|DCM(plus|less)?(10m|20m)?|175m")) %>%
  mutate(depth=ifelse(depth=="15m"|depth=="25m", "Surface", depth)) %>%
  mutate(depth=ifelse(depth=="175m", "Deep", depth)) %>%
  mutate(tripl=str_extract(filename, "(?<=_)[A-C](?=.mzML$)")) %>%
  mutate(untripl=str_replace(filename, "_[A-C].mzML$", "")) %>%
  mutate(date_run=str_extract(filename, "^\\d+")) %>%
  mutate(date_run=strptime(date_run, format = "%y%m%d")) %>%
  mutate(sample_id=str_extract(filename, "(S|MS|L)\\d.*_[A-C](?=.mzML$)"))

extra_filedata <- list.files("../mzMLs", pattern = "Blk|Std|Poo.*mzML", recursive = TRUE) %>%
  data.frame(filename=.) %>%
  mutate(polarity=dirname(filename)) %>%
  mutate(filename=basename(filename)) %>%
  mutate(samp_type=str_extract(filename, "Std|Blk|Poo")) %>%
  mutate(cruise=case_when(
    str_detect(filename, "170706") ~ "MC",
    str_detect(filename, "180205") ~ "MC",
    str_detect(filename, "180821") ~ "MT",
    str_detect(filename, "190715") ~ "FK",
    TRUE ~ "Unknown"
  )) %>%
  mutate(date_run=str_extract(filename, "^\\d+")) %>%
  mutate(date_run=strptime(date_run, format = "%y%m%d"))

complete_file_data <- bind_rows(samp_filedata, extra_filedata) %>%
  left_join(samp_log)
```

## Interpolation of SCOPE data to MS sample locations

```{r Interpolate to sample sites}
interp_df <- lst(complete_env_data, complete_file_data) %>%
  bind_rows(.id = "data_type") %>%
  mutate(station_grouping=case_when(
    station%in%as.character(74:86)~"falkor_anticyc",
    station%in%as.character(28:70)~"falkor_cyc",
    TRUE~station
  )) %>%
  split(.$station_grouping) %>%
  map(function(single_station){
    single_station %>%
      mutate(across(c(lat:beam_atten), function(col){
        if(all(is.na(col)))return(NA)
        approx(abs_depth, col, xout = abs_depth, rule = 2, ties = mean)$y
        # pmax(0, predict(loess(col~abs_depth, data = ., span = 1), newdata = .))
      }))
  }) %>%
  bind_rows() %>%
  filter(data_type=="complete_file_data") %>%
  select(-data_type)
# Correct wonky PC values (those greater than 4) with a linear model built on beam_atten
interp_df <- interp_df %>%
  mutate(pred_PC=interp_df %>% 
           filter(PC_um<4) %>%
           lm(formula=PC_um~beam_atten) %>%
           predict(newdata=interp_df)) %>%
  mutate(PC_um=ifelse(PC_um>4, pred_PC, PC_um))
write_csv(interp_df, "filled_file_metadata.csv")
```

## Visualization

```{r Single-column metadata plot}
col_name <- "PC_um"
complete_env_data %>%
  mutate(station=factor(station, levels=c(
    1:80, "L1", "L2"
  ))) %>%
  filter(!is.na(get(col_name))) %>%
  filter(abs_depth<190) %>%
  mutate(abs_depth=abs_depth+3) %>%
  mutate(station_coord=as.numeric(station)-0.2) %>%
  ggplot(mapping = aes(x=station, y=abs_depth, color=get(col_name))) +
  geom_point(size=3, position = position_nudge(x=-0.1)) +
  geom_point(data = interp_df, color="red", size=4, position = position_nudge(x=0.1)) +
  geom_point(data = interp_df, size=3, position = position_nudge(x=0.1)) +
  facet_grid(~cruise, scales = "free_x", space = "free_x") +
  scale_color_viridis_c() +
  coord_cartesian(ylim = c(180, 0)) +
  theme_bw() +
  theme(legend.position = "none") +
  ggtitle(str_remove(col_name, "_u.*"))
```

```{r PDF of all the metadata}
pdf("allmeso_metadata.pdf", onefile = TRUE, width = 6.5, height = 3)
setdiff(names(complete_env_data), c("abs_depth", "time")) %>%
  walk(function(col_name){
    gp_i <-complete_env_data %>%
      mutate(station=factor(station, levels=c(
        1:80, "L1", "L2"
      ))) %>%
      filter(!is.na(get(col_name))) %>%
      filter(abs_depth<190) %>%
      mutate(abs_depth=abs_depth+3) %>%
      mutate(station_coord=as.numeric(station)-0.2) %>%
      ggplot(mapping = aes(x=station, y=abs_depth, color=get(col_name))) +
      geom_point(size=3, position = position_nudge(x=-0.1)) +
      geom_point(data = interp_df, color="red", size=4, position = position_nudge(x=0.1)) +
      geom_point(data = interp_df, size=3, position = position_nudge(x=0.1)) +
      facet_grid(~cruise, scales = "free_x", space = "free_x") +
      coord_cartesian(ylim = c(180, 0)) +
      theme_bw() +
      theme(legend.position = "none") +
      ggtitle(str_remove(col_name, "_u.*"))
    print(gp_i)
  }, .progress=TRUE)
dev.off()
```

```{r Interpolation comparison loess v linear for particular stations}
dat_vals <- data.frame(station_id="Station 9", value_id="N+N (µm)")
map(c("L1", "9"), function(station_i){
  map(c("PC_um", "no23_um"), function(col_i){
    lst(complete_env_data, complete_file_data) %>%
      bind_rows(.id = "data_type") %>%
      filter(station==station_i) %>%
      mutate(col=get(col_i)) %>%
      mutate(lin_approx=approx(abs_depth, col, xout = abs_depth, rule = 2, ties = mean)$y) %>%
      mutate(loess_approx_05=predict(loess(col~abs_depth, data = ., span = 0.5), newdata = .)) %>%
      mutate(loess_approx_1=predict(loess(col~abs_depth, data = ., span = 1), newdata = .)) %>%
      filter(abs_depth<190) %>%
      arrange(abs_depth) %>%
      mutate(value_id=col_i)
  }) %>%
    bind_rows() %>%
    mutate(station_id=station_i)
}) %>%
  bind_rows() %>%
  mutate(station_id=factor(station_id, levels=c("9", "L1"), labels = c("Station 9", "Station L1"))) %>%
  mutate(value_id=factor(value_id, levels=c("no23_um", "PC_um"), 
                         labels = c("N+N (µm)", "PC (µm)"))) %>%
  ggplot(aes(y=abs_depth)) +
  geom_vline(xintercept = 0) +
  geom_point(aes(x=col)) +
  geom_path(aes(x=lin_approx), color="#00588B") +
  geom_path(aes(x=loess_approx_05), color="#009B95") +
  geom_path(aes(x=loess_approx_1), color="#53CC67") +
  coord_cartesian(ylim=c(200, 0)) +
  geom_label(x=Inf, y=50, label="Linear", color="#00588B", label.r=unit(0, "in"), hjust=1, data=dat_vals) + 
  geom_label(x=Inf, y=75, label="Loess (span=0.5)", color="#009B95", label.r=unit(0, "in"), hjust=1, data=dat_vals) + 
  geom_label(x=Inf, y=100, label="Loess (span=1)", color="#53CC67", label.r=unit(0, "in"), hjust=1, data=dat_vals) + 
  theme_bw() +
  facet_grid(station_id~value_id, scales="free_x") +
  labs(x=NULL, y="Depth")
```

### 2023-05-23 untargeted/scripts/bmisscript.R
Moving scripts to old folder

@@ -1,37 +1,0 @@
-# Script to find the B-MIS for each standard in the untargeted data set
-# Called by Control.Rmd
-
-# all_peaks, given_stans, polarity, pmppm all defined in Control.Rmd
-
-
-# Perform BMIS
-pooled <- M_area_peaks %>%
-  filter(str_detect(filename, "_Poo_"))
-
-pooled_IS <- is_peaks %>%
-  rename_with(~paste0("IS_", .), -filename) %>%
-  add_row(filename=unique(.$filename), IS_compound_name="None", IS_M_area=1)
-
-initial_CVs <- pooled %>%
-  mutate(M_area=ifelse(str_detect(filename, "Half"), M_area*2, M_area)) %>%
-  group_by(feature) %>%
-  summarize(init_CV=cv(M_area))
-
-standardized_CVs <- full_join(pooled, pooled_IS, by=c("filename")) %>%
-  group_by(feature, IS_compound_name) %>%
-  summarize(norm_CV=cv(M_area/IS_M_area), .groups="drop")
-
-BMIS_data <- standardized_CVs %>%
-  left_join(initial_CVs) %>%
-  group_by(feature) %>%
-  mutate(MIS_improvement=(init_CV - norm_CV)/init_CV) %>% 
-  filter(MIS_improvement>min_improvement|IS_compound_name=="None") %>%
-  arrange(desc(MIS_improvement)) %>%
-  group_by(feature) %>%
-  slice(1) %>%
-  select(feature, BMIS=IS_compound_name, init_CV, norm_CV, MIS_improvement)
-chosen_BMIS <- BMIS_data %>% select(feature, BMIS)
-
-is_peaks <- is_peaks %>%
-  add_row(compound_name="None", filename=unique(.$filename), 
-          M_area=ifelse(str_detect(filename, "Half"), 0.5, 1))

### 2023-05-23 untargeted/scripts/compound_assignments.R

@@ -1,52 +1,0 @@
-# Collects structural best-guesses from CSI:FingerID and compares them with
-# metfRag and other software?
-
-library(tidyverse)
-library(rvest)
-
-getSynonymsName <- function(cmpd_name, how_many=20){
-  prolog <- "https://pubchem.ncbi.nlm.nih.gov/rest/pug/compound/name/"
-  tail <- "/cids/TXT"
-  url <- paste0(prolog, cmpd_name, tail)
-  CID <- readLines(url)
-  getSynonymsCID(CID, how_many)
-}
-
-getSynonymsCID <- function(CID, how_many=20){
-  prolog <- "https://pubchem.ncbi.nlm.nih.gov/rest/pug/compound/cid/"
-  tail <- "/synonyms/TXT"
-  url <- paste0(prolog, CID, tail)
-  content <- head(readLines(url), how_many)
-  paste0(content, collapse = "; ")
-}
-
-cmpd_file <- "XCMS/data_intermediate/sirius_project/output_dir/compound_identifications_adducts.tsv"
-cmpd_raw <- read_tsv(cmpd_file, col_types = 
-         cols(
-           rank = col_double(),
-           `#adducts` = col_double(),
-           `#predictedFPs` = col_double(),
-           Confidence_Score = col_character(),
-           `CSI:FingerID_Score` = col_double(),
-           Zodiac_Score = col_double(),
-           TreeIsotope_Score = col_double(),
-           molecularFormula = col_character(),
-           adduct = col_character(),
-           InChIkey2D = col_character(),
-           InChI = col_character(),
-           name = col_character(),
-           smiles = col_character(),
-           xlogp = col_logical(),
-           pubchemids = col_character(),
-           links = col_character(),
-           dbflags = col_double(),
-           id = col_character()
-         ))
-cmpd_data <- cmpd_raw %>%
-  mutate(feature=sub(pattern = "[[:digit:]]*_", replacement = "", x = .$id)) %>%
-  mutate(feature=sub(pattern = "_FEATURE1", replacement = "", x = .$feature)) %>%
-  #mutate(top_cid=sub(pattern = ";.*", replacement = "", x=.$pubchemids)) %>%
-  head() %>%
-  #mutate(synonyms=sapply(X = .$top_cid, getSynonymsCID))
-  mutate(synonyms=sapply(X = .$name, getSynonymsName))
-cmpd_data %>% select(name, synonyms)

### 2023-05-23 untargeted/scripts/deisoadduct.R

@@ -1,116 +1,0 @@
-# Script to de-isotope and de-adduct XCMS-picked peaks
-# Called by Control.Rmd
-
-# Include isIsoAdduct code here so we can safely source() the functions script
-# rather than providing each of the functions because bplapply doesn't copy over
-# the working environment
-# given_file <- filterFile(xdata_filled, "180821_Smp_MS12C115m_A.mzML")
-isIsoAdduct <- function(file_peaks, xdata_filled, addiso_masses){
-  source("scripts/functions.R")
-  library(data.table)
-  setDTthreads(1)
-  library(dplyr, warn.conflicts = FALSE)
-  suppressPackageStartupMessages(library(xcms))
-  
-  file_addisos <- addiso_masses %>%
-    {rbind(cbind(., type="self", dir=-1), cbind(., type="seek", dir=1))} %>%
-    mutate(mz_diff=mz_diff*dir) %>%
-    select(-dir) %>%
-    group_by(addiso, type) %>%
-    group_split()
-  addiso_range <- range(sapply(file_addisos, function(x)x$mz_diff))
-  
-  given_file <- filterFile(xdata_filled, unique(file_peaks$filename))
-  file_data_table <- as.data.table(given_file)[,c("rt", "mz", "i")]
-  
-  # Split up the large data.table into smaller ones to help speed up binary search
-  split_vec <- round(1:nrow(file_peaks)/(nrow(file_peaks)/100))
-  split_file_peaks <- split(file_peaks[order(file_peaks$rt),], split_vec)
-  split_data_table <- lapply(split_file_peaks, function(file_peaks_i){
-    file_data_table[rt%between%c(min(file_peaks_i$rtmin), max(file_peaks_i$rtmax))]
-  })
-  
-  # test_peak <- file_peaks %>% slice_sample(n=1) # Debug
-  # test_peak <- file_peaks %>% filter(mz%between%pmppm(118.0865, 5))
-  # test_peak <- file_peaks %>% filter(feature=="FT0487")
-  # feature_i <- test_peak$feature
-  # mz_i <- test_peak$mz
-  # rtmin_i <- test_peak$rtmin
-  # rtmax_i <- test_peak$rtmax
-  # split_table_i <- file_data_table
-  
-  addiso_all <- mapply(function(split_peaks_i, split_table_i){
[…]

### 2023-05-23 untargeted/scripts/formula_assignments.R

@@ -1,145 +1,0 @@
-
-# Load addisos - this should be done earlier in the script but that's not guaranteed
-if(!exists(addiso_masses)){
-  warning("Using manually written addisos, may need updating")
-  addiso_masses_all <- tribble(
-    ~adduct_or_iso, ~polarity, ~addiso, ~nat_abund, ~mz_diff,
-    "iso",          "both",    "C13",   0.0110,     1.003355, 
-    "iso",          "both",    "X2C13", 0.0110,     2*1.003355, 
-    "iso",          "both",    "S34",   0.0421,     1.995796, 
-    "iso",          "both",    "S33",   0.0075,     0.999387,
-    "iso",          "both",    "N15",   0.0037,     0.997035, 
-    "iso",          "both",    "O18",   0.0020,     2.004244,
-    "iso",          "both",    "Cl37",  0.2423,     1.99705,
-    "iso",          "both",    "Br81",  0.4931,     1.997954,
-    "adduct",       "both",    "M-H2O", NA,         -18.0106, 
-    "adduct",       "pos",     "M+Na",  NA,          22.98922-1.007276 # Goes from [M+H] to [M+Na]
-    # "adduct",       "pos",     "M+NH4", NA,          18.0338-1.007276,
-    # "adduct",       "pos",     "M+K",   NA,          38.963708-1.007276,
-    # "adduct",       "neg",     "M+Cl",  NA,          34.969402+1.007276,
-    # "adduct",       "neg",     "M+Ac",  NA,          59.013851+1.007276,
-    # "adduct",       "neg",     "M+Br",  NA,          78.918885+1.007276
-  )
-  addiso_masses <- addiso_masses_all %>% filter(polarity%in%c(!!polarity, "both"))
-}
-
-
-# Run SIRIUS ----
-message("Running SIRIUS formulas...")
-mzs <- feature_data$mzmed
-sirius_cmd <- paste0('sirius --noCite decomp',
-                     ' --mass=', paste(mzs, collapse = ","),
-                     ' --ppm=5',
-                     ' --elements=CHNOPS')
-if(nchar(sirius_cmd)>8096){
-  # Windows doesn't support commands longer than 8,096 characters
-  n_chunks <- ceiling(nchar(sirius_cmd)/8096)
[…]

### 2023-05-23 untargeted/scripts/funcs_old.R

@@ -1,505 +1,0 @@
-# Functions called by Control.Rmd or scripts within
-# Should be sourced every time!
-
-cv <- function(x, na.rm=FALSE, inf.rm=FALSE){
-  if(inf.rm){
-    x[is.infinite(x)] <- NA
-  }
-  sd(x = x, na.rm = na.rm)/mean(x = x, na.rm = na.rm)
-}
-
-pmppm <- RaMS::pmppm
-
-grabSingleFileData <- function(filename){
-  msdata <- mzR:::openMSfile(filename)
-  fullhd <- mzR::header(msdata)
-  spectra_list <- lapply(seq_len(nrow(fullhd)), function(x){
-    given_peaks <- mzR::peaks(msdata, x)
-    rtime <- fullhd[x, "retentionTime"]
-    return(cbind(rtime, given_peaks))
-  })
-  all_data <- `names<-`(as.data.frame(do.call(rbind, spectra_list)), 
-                        c("rt", "mz", "int"))
-  return(all_data)
-}
-speedyQscoreCalculator <- function(file_peaks, grabSingleFileData, qscoreCalculator){
-  library(data.table)
-  file_data <- grabSingleFileData(unique(file_peaks$filename))
-  file_data_table <- as.data.table(file_data)
-  peak_splits <- split(file_peaks, ceiling(seq_len(nrow(file_peaks))/100))
-  file_qscores <- lapply(peak_splits, function(i){
-    eic_many <- file_data_table[mz%between%c(min(i$mzmin), max(i$mzmax))]
-    individual_peaks <- split(i, seq_len(nrow(i)))
-    output <- sapply(individual_peaks, function(peak_row){
-      eic <- eic_many[rt%between%c(peak_row$rtmin, peak_row$rtmax) & 
-                        mz%between%c(peak_row$mzmin, peak_row$mzmax)]
-      qscoreCalculator(eic)
-    }, USE.NAMES = FALSE)
-  })
-  file_peaks$sn <- unlist(file_qscores)
-  return(file_peaks)
-}
-qscoreCalculator <- function(eic){
-  #Check for bogus EICs
-  if(nrow(eic)<5){
-    return(0)
-  }
-  #Calculate where each rt would fall on a beta dist (accounts for missed scans)
-  scaled_rts <- (eic$rt-min(eic$rt))/(max(eic$rt)-min(eic$rt))
-  
-  # Create a couple different skews and test fit
[…]

### 2023-05-23 untargeted/scripts/functions.R

@@ -1,345 +1,0 @@
-# Functions called by Control.Rmd or scripts within
-# Should be sourced every time!
-
-cv <- function(x, na.rm=FALSE, inf.rm=FALSE){
-  if(inf.rm){
-    x[is.infinite(x)] <- NA
-  }
-  sd(x = x, na.rm = na.rm)/mean(x = x, na.rm = na.rm)
-}
-speedyQscoreCalculator <- function(file_peaks, grabSingleFileData, qscoreCalculator){
-  library(data.table)
-  file_data <- grabSingleFileData(unique(file_peaks$filename))
-  file_data_table <- as.data.table(file_data)
-  peak_splits <- split(file_peaks, ceiling(seq_len(nrow(file_peaks))/100))
-  file_qscores <- lapply(peak_splits, function(i){
-    eic_many <- file_data_table[mz%between%c(min(i$mzmin), max(i$mzmax))]
-    individual_peaks <- split(i, seq_len(nrow(i)))
-    output <- sapply(individual_peaks, function(peak_row){
-      eic <- eic_many[rt%between%c(peak_row$rtmin, peak_row$rtmax) & 
-                        mz%between%c(peak_row$mzmin, peak_row$mzmax)]
-      qscoreCalculator(eic)
-    }, USE.NAMES = FALSE)
-  })
-  file_peaks$sn <- unlist(file_qscores)
-  return(file_peaks)
-}
-qscoreCalculator <- function(eic){
-  #Check for bogus EICs
-  if(nrow(eic)<5){
-    return(0)
-  }
-  #Calculate where each rt would fall on a beta dist (accounts for missed scans)
-  scaled_rts <- (eic$rt-min(eic$rt))/(max(eic$rt)-min(eic$rt))
-  
-  # Create a couple different skews and test fit
-  maybe_skews <- c(2.5,3,4,5) #Add 7 to catch more multipeaks and more noise
-  #Add 2 to catch very slopey peaks and more noise
-  best_skew <- maybe_skews[which.max(sapply(maybe_skews, function(x){
-    cor(dbeta(scaled_rts, shape1 = x, shape2 = 5), eic$int)
-  }))]
-  perf_peak <- dbeta(scaled_rts, shape1 = best_skew, shape2 = 5)
-  peak_cor <- cor(perf_peak, eic$int)
-  
-  
-  #Calculate the normalized residuals
-  residuals <- eic$int/max(eic$int)-perf_peak/max(perf_peak)
[…]

### 2023-05-23 untargeted/scripts/peakpicking.R

@@ -1,82 +1,0 @@
-# Script to run XCMS peakpicking, retention time correction, and grouping
-# Called by output_control.Rmd
-# Requires dev version of XCMS for improved peakpicking given settings
-
-# ms_files comes from output_control.Rmd
-# qscore_threshold comes from output_control.Rmd
-
-
-# Read in the raw data ----
-message("Reading files...")
-start_time <- Sys.time()
-raw_data <- readMSData(files = normalizePath(paste0(mzml_path, ms_files)), 
-                       msLevel. = 1, centroided. = TRUE, mode = "onDisk",
-                       pdata = as(cruise_metadata, "AnnotatedDataFrame"))
-time_total <- round(difftime(Sys.time(), start_time), digits = 2)
-cat("Time to read in files: ", time_total, units(time_total), "\n")
-
-# Perform peakpicking ----
-message("Picking peaks...")
-start_time <- Sys.time()
-xdata <- suppressMessages(findChromPeaks(raw_data, param = cwp))
-time_total <- round(difftime(Sys.time(), start_time), digits = 2)
-cat("Time to perform peakpicking: ", time_total, units(time_total), "\n")
-# 13 minutes
-
-
-# Assign new quality scores ----
-# speedyQscoreCalculator is custom function which should be in functions script
-# sourced by output_control.Rmd
-message("Assigning quality scores...")
-start_time <- Sys.time()
-xcms_peakdf <- chromPeaks(xdata) %>%
-  as.data.frame(stringsAsFactors=FALSE) %>%
-  mutate(filename=fileNames(xdata)[.[["sample"]]]) %>%
-  arrange(mz)
-split_xcms_filedata <- split(xcms_peakdf, xcms_peakdf$filename)
-files_qscores <- bplapply(X = split_xcms_filedata, FUN = speedyQscoreCalculator,
-                          grabSingleFileData, qscoreCalculator)
-peakdf_qscored <- files_qscores %>%
-  do.call(what = rbind) %>%
-  as.data.frame(stringsAsFactors=FALSE) %>% 
-  filter(sn>qscore_threshold) %>%
-  select(-filename) %>%
-  arrange(sample, rtmin, rtmax) %>%
-  as.matrix()
-xdata_cleanpeak <- `chromPeaks<-`(xdata, peakdf_qscored)
-
[…]

### 2023-05-23 untargeted/scripts/run_sirius.R

@@ -1,133 +1,0 @@
-# Grab MSMS data ----
-print("Reading in MS2 data...")
-msdata <- grabMSdata(MSMS_files, grab_what = c("MS1", "MS2"))
-msdata$MS1$rt <- msdata$MS1$rt*60
-msdata$MS2$rt <- msdata$MS2$rt*60
-
-print("Determining which have MS2 frags...")
-has_msms <- feature_data %>%
-  select(feature, mzmed, rtmed) %>%
-  split(.$feature) %>%
-  pbsapply(function(x){
-    msdata$MS2[premz%between%pmppm(x$mzmed, ppm = 5)] %>%
-      filter(rt%between%(x$rtmed+c(-30, 30))) %>%
-      nrow() %>%
-      as.logical()
-  })
-data.frame(feature=names(has_msms), has_msms) %>%
-  filter(has_msms) %>%
-  left_join(stan_annotations, by=c("feature")) %>%
-  # filter(!is.na(compound_name)) %>%
-  select(feature, compound_name)
-feature_data <- feature_data %>%
-  filter(feature%in%names(has_msms)[has_msms])
-
-
-
-# Create .mgf files for SIRIUS to read ----
-if(dir.exists(sirius_project_dir)){
-  unlink(sirius_project_dir, recursive = TRUE)
-}
-dir.create(sirius_project_dir)
-dir.create(paste0(sirius_project_dir, "//raw_files"))
-message("Writing .mgf files...")
-pbsapply(feature_data$feature, function(feature_num){
-  output_file <- paste0(sirius_project_dir, "\\raw_files\\", feature_num, ".mgf")
-  feature_msdata <- feature_data[feature_data$feature==feature_num, ]
-  ms1 <- rbind(c(feature_msdata$mzmed, feature_msdata$avg_area),
-               c(feature_msdata$mzmed+1.003355, feature_msdata$C13),
-               c(feature_msdata$mzmed+1.003355*2, feature_msdata$X2C13),
-               c(feature_msdata$mzmed+1.995796, feature_msdata$S34),
-               c(feature_msdata$mzmed+0.997035, feature_msdata$N15),
-               c(feature_msdata$mzmed+2.004244, feature_msdata$O18))
-  ms1 <- ms1[ms1[,2]!=0, , drop=FALSE]
-  ms2 <- msdata$MS2[premz%between%pmppm(feature_msdata$mzmed)&
[…]

### 2023-05-23 untargeted/scripts/standard_assignments.R

@@ -1,393 +1,0 @@
-
-# Functions ----
-stan_guesser <- function(isotope_choice, mix_choice, match_choice, area_choice, rt_choice){
-  if(all(is.na(c(isotope_choice, mix_choice, match_choice, area_choice, rt_choice)))){
-    return("No peaks found")
-  }
-  if(!is.na(isotope_choice)){
-    return(isotope_choice)
-  }
-  if(!is.na(match_choice)) {
-    return(match_choice)
-  }
-  if(is.na(mix_choice)){
-    mix_options <- NA
-  } else if(!nchar(mix_choice)){
-    mix_options <- NA
-  } else {
-    mix_options <- unlist(strsplit(mix_choice, split = "; "))
-  }
-  if(length(mix_options)==1&!all(is.na(mix_options))){
-    return(mix_options)
-  }
-  
-  if(length(mix_options)>1&!all(is.na(mix_options))){
-    if(rt_choice==area_choice){
-      return(rt_choice)
-    }
-    if(rt_choice%in%mix_options&!area_choice%in%mix_options){
-      return(rt_choice)
-    }
-    if(area_choice%in%mix_options&!rt_choice%in%mix_options){
-      return(area_choice)
-    }
-    if(area_choice%in%mix_options&rt_choice%in%mix_options){
-      return(paste(c(area_choice, rt_choice), collapse = "; "))
-    }
-  }
-  return(rt_choice)
-}
-reAnno <- function(anno_df, cmpd_name, which_new=NULL, new_feature_num=NULL){
-  if(is.null(new_feature_num) & is.null(which_new)){
-    stop(paste0("For ", cmpd_name, ", both `new_feature_name` and ",
-                "`which_new` must not both be NULL"))
-  }
-  if(!is.null(which_new) && !which_new%in%c("largest", "first", "last", "second")){
-    stop(paste0("For ", cmpd_name, ", ", which_new, " is not a known keyword"))
-  }
-  if(!is.null(new_feature_num)){
-    new_df <- anno_df %>% rows_update(
-      data.frame(compound_name=cmpd_name, feature=new_feature_num, isotope_validated="Manual"), 
-      by = "compound_name"
-    )
-    return(new_df)
-  }
-  if(!is.null(which_new)){
-    chosen_feature <- anno_df %>% filter(compound_name==cmpd_name) %>% pull(feature)
-    if(length(chosen_feature)==0){
-      stop(paste0(cmpd_name, " not found in supplied anno_df"))
-    }
[…]

### 2023-05-23 untargeted/scripts/treatment_diff_search.R

@@ -1,74 +1,0 @@
-# Code that looks for differences in peak area across treatments
-# Needs a peak list (typically produced by peakpicking.R)
-# Calculates p-values between treatments
-#  p-values are amended via the Holm-Bonferroni method to avoid multiple comparisons
-#  Treatments are deduced from the file_name column of the peak list
-# Data that only shows up in one treatment is given a p-value of 0
-
-library(tidyverse)
-final_peaks <- read.csv(file = "XCMS/data_pretty/final_peaks.csv")
-final_features <- read.csv(file = "XCMS/data_pretty/final_features.csv")
-feature_formulas <- read.csv(file = "XCMS/data_pretty/feature_formulas.csv")
-feature_putatives <- "XCMS/data_intermediate/sirius_project/output_dir/" %>%
-  paste0("compound_identifications.tsv") %>%
-  readLines() %>%
-  strsplit(split = "\t")
-header <- feature_putatives[[1]]
-feature_putatives <- feature_putatives[-1] %>%
-  do.call(what = rbind) %>%
-  as.data.frame() %>%
-  `names<-`(header) %>%
-  select(id, Confidence_Score, name) %>%
-  mutate(feature=gsub(pattern = "^[[:digit:]]+_", "", .$id)) %>%
-  mutate(feature=gsub(pattern = "_FEATURE1$", "", .$feature)) %>%
-  mutate(Confidence_Score=round(as.numeric(Confidence_Score), digits = 3)) %>%
-  select(-id)
-
-diffreport <- function(peaks, pattern_a, pattern_b){
-  split(peaks, peaks$feature) %>%
-    lapply(function(x){
-      DCM_areas <- x$BMISed_area[grep(pattern = pattern_a, x$file_name)]
-      m25_areas <- x$BMISed_area[grep(pattern = pattern_b, x$file_name)]
-      if(length(DCM_areas)<3|length(m25_areas)<3)return(c(0, 1))
-      c(pval=t.test(DCM_areas, m25_areas)$p.value, 
-        diff=mean(DCM_areas)/mean(m25_areas))
-    }) %>% do.call(what = rbind) %>% as.data.frame() %>%
-    mutate(feature=rownames(.)) %>%
-    arrange(pval) %>%
-    mutate(p_adj=p.adjust(pval, method = "BH")) %>%
[…]

### 2023-05-23 metadata/metadata_control_v2.Rmd
Fixed date bug, added discrete time and sla

@@
   mutate(date=as.Date(case_when(
-    station=="L1"~"2018-07-08",
-    station=="L2"~"2018-07-09"
+    station=="L1"~"2017-07-08",
+    station=="L2"~"2017-07-09"
   )))
@@
            predict(newdata=interp_df)) %>%
-  mutate(PC_um=ifelse(PC_um>4, pred_PC, PC_um))
+  mutate(PC_um=ifelse(PC_um>4, pred_PC, PC_um)) %>%
+  select(-pred_PC) %>%
+  mutate(sla_class=case_when(
+    sla_corr < -5 ~ "Cyclone",
+    sla_corr > 5 ~ "Anticyclone",
+    TRUE ~ "Neither"
+  )) %>%
+  mutate(time=as.POSIXct(time, origin="1970-01-01")) %>%
+  mutate(time_class=c("Night", "Morning", "Midday", "Evening")[
+    floor((as.numeric(format(time, "%H"))+3)/6)%%4+1
+  ])
 write_csv(interp_df, "filled_file_metadata.csv")

### 2023-05-23 metadata/metadata_control_v2.Rmd
Fixed bug in time interpolation

@@
   filter(data_type=="complete_file_data") %>%
-  select(-data_type)
+  select(-data_type, -time) %>%
+  left_join(complete_env_data %>% distinct(station, cast, time), by=c("station", "cast"))
 # Correct wonky PC values (those greater than 4) with a linear model built on beam_atten
@@
   select(-pred_PC) %>%
+# Discretize SLA and time
   mutate(sla_class=case_when(

### 2023-05-25 metadata/metadata_control.Rmd
Fully switched to metadata v2

@@
 library(tidyverse)
-library(cmap4r)
-cmap_api <- "6acd63a0-4c63-11ea-aa09-71d9d763e28d"
-set_authorization(cmap_key = cmap_api)
-library(fuzzyjoin)
-library(ggmap)
-register_google("AIzaSyBVNJV0Myoy3M50IBBDIjXZhUyU8Mnagvs")
-library(ggrepel)
-library(gridExtra)  # For grid.arrange
-library(data.table) # For fast indexing of seaflow data
-library(RaMS)       # For MS file timestamps
-library(scales)     # For "muted" function
 library(readxl)     # For "readxl"
@@
 
-if(!dir.exists("raw_data")){
-  dir.create("raw_data")
-}
-if(!dir.exists("made_data")){
-  dir.create("made_data")
-}
-if(!dir.exists("figures")){
-  dir.create("figures")
-}
-
 options(pillar.sigfig = 7)
-
-toDecDeg <- function(deg, min, dir){
-  num <- deg+min/60
-  if(dir%in%c("S", "W")){
-    num*-1
-  } else {
-    num
-  }
-}
 ```
 
-## CTD data (cruise, station, time, lat, lon, cast, SLA, DCM depth)
+## Download/collect raw data
 
-Cruise is manually entered, station/time/lat/lon/cast come from the SCOPE website cruise
-summaries, SLA is pulled down from CMAP, and DCM depth was manually found by looking
-at the cruise logbook.
+Full list of files we need!
+From SCOPE website:
+  - FK_CTDsummary_current.txt
+  - MS_CTDsummary_current.txt
+  - Karl_FK_PCPN_WC_current.xlsx
+  - Karl_FK_Nutrients_Final.xlsx
+  - 2017MS_CoreData_current.xls
+  - Karl_MS_Nutrient_Summary_KB.xlsx
+  - Karl_MS_pATP.xlsx
+  - Karl_FK_chlorophyll.xlsx
+From email
+  - FALKOR- Sample Log_INGALLS (From Tim)
+  - MESO_SCOPE_Cruise_KM1709_Sample_Log (From Laura)
+  - MESO-SCOPE_BeamAttenuation_forWill (From Benedetto, as are remainder)
+  - MESO-SCOPE_DailyPAR_forWill
+  - MESO-SCOPE_PercentPAR_forWill
+  - kok1607_ctd_sla
+  - km1709_ctd_sla
+  - fk180310_ctd_sla
+Manually created
+  - hmdb_formulae.rds
+  - MAA_masses.xlsx
 
-Also now includes data from .gof files for exact depth
+```{r From SCOPE data inventory, eval = !file.exists("raw_data/FK_CTDsummary_current.txt")}
+inventory_url <- paste0("http://", scope_uname, ":", scope_pw,
+                        "@scope.soest.hawaii.edu/collaborators/datainventory/Data/")
+from_scope_inventory <- c(
+  "SCOPEcore/FK_CTDsummary_current.txt",
+  "SCOPEcore/MS_CTDsummary_current.txt",
+  "Karl/Karl_FK_PCPN_WC_current.xlsx",
+  "Karl/Karl_FK_Nutrients_Final.xlsx",
+  "SCOPEcore/2017MS_CoreData_current.xls",
+  "Karl/Karl_MS_Nutrient_Summary_KB.xlsx",
+  "Karl/Karl_MS_pATP.xlsx",
+  "Karl/Karl_FK_Chlorophyll.xlsx"
+)
+sapply(from_scope_inventory, function(end_url){
+  download.file(paste0(inventory_url, end_url), destfile = paste0("raw_data/", basename(end_url)), mode = "wb")
+})
+```
 
-```{r station_metadata}
-# DCM depths are entered manually ----
-if(!file.exists("raw_data/all_DCMs.csv")){
-  falkor_DCMs <- data.frame(
-    station=c(62, 64, 77, 80),
-    cast=1,
-    DCM_depth=c(98, 115, 125, 120)
-  )
-  transect_DCMs <- data.frame(
-    station = c(4:13, 13, 14),
-    cast=c(1,1,3,1,1,2,2,2,1,2,3,2),
-    DCM_depth=c(100, 122, 119, 133, 122, 110, 110, 110, 103, 112, 112, 105)
-  )
-  center_DCMs <- data.frame(
-    station = paste0("L", c(1,1,2,2)),
-    cast=c(32, 33, 1, 2),
-    DCM_depth=c(108, 104, 123, 123)
-  )
-  all_DCMs <- rbind(falkor_DCMs, transect_DCMs, center_DCMs)
+```{r From SCOPE FTP, eval=!file.exists("raw_data/mesoscope.gof")}
+ftp_url <- "http://scope.soest.hawaii.edu/FTP/scope/water/"
+from_scope_ftp <- c("mesoscope.gof","fk180310.gof")
+sapply(from_scope_ftp, function(end_url){
+  download.file(paste0(ftp_url, end_url), destfile = paste0("raw_data/", basename(end_url)))
+})
+```
 
-  write.csv(all_DCMs, "raw_data/all_DCMs.csv", row.names = FALSE)
-}
+```{r From email}
+# Go search for the files in email, we can't script this part
+```
 
-# SCOPE CTD data consists of station, cast, time, max_depth, bottles, lat, and lon ----
-getScopeCTD <- function(url){
+```{r Manually created}
+# Go search for the files in Slack, we can't script this part
+```
+
+## SCOPE data cleaning and combining
+
+```{r Sample log and DCM extraction}
+col_types <- c("date", "text", "numeric", "text",
+               "text", "date", "date", "text")
+col_names <- c("date", "station", "cast", "sample_id", "depth",
+               "start_time", "end_time", "volume")
+falkor_log <- c("Anti-Cyclonic Eddy", "Cyclonic Eddy") %>%
+  lapply(function(sheet_i){
+    read_xlsx("raw_data/FALKOR- Sample Log_INGALLS.xlsx",
+             sheet=sheet_i, skip = 5, 
+             col_types = col_types, col_names = col_names) 
+}) %>%
+  bind_rows() %>%
+  drop_na() %>%
+  filter(sample_id!="1") %>%
+  mutate(date=as.Date(date)) %>%
+  mutate(abs_depth=as.numeric(str_extract(depth, "\\d+"))) %>%
+  mutate(str_depth=ifelse(startsWith(depth, "DCM"), "DCM", paste0(depth, "m"))) %>%
+  mutate(depth=ifelse(depth=="25", "Surface", "DCM")) %>%
+  mutate(volume=as.numeric(str_extract(volume, "\\d+"))) %>%
+  mutate(sample_id=paste0("S", station, "C", cast, "-", str_depth, "_", sample_id)) %>%
+  select(-str_depth)
+
+col_types <- c("text", "text", "text", "text",
+               "text", "numeric", "date", "date",
+               "text", "text", "skip", "skip", "skip")
+col_names <- c("date", "station", "cast", "depth", "sample_id",
+               "volume", "start_time", "end_time", "rosette", 
+               "dissolved")
+MT_log <- read_xlsx("raw_data/MESO_SCOPE_Cruise_KM1709_Sample_Log.xlsx",
+          sheet = "Metabolites_and_Proteins", skip = 2,
+          col_types = col_types, col_names=col_names) %>%
+  drop_na() %>%
+  mutate(date=ifelse(startsWith(date, "6/29"), "42916.0", date)) %>%
+  mutate(date=as.Date(as.numeric(date), origin="1899-12-31")) %>%
+  mutate(station=str_extract(station, "\\d+")) %>%
+  mutate(cast=as.numeric(str_extract(cast, "\\d+"))) %>%
+  mutate(abs_depth=as.numeric(str_extract(depth, "\\d+"))) %>%
+  mutate(depth=case_when(
+    depth=="15.0"~"Surface", depth=="175.0"~"Deep", TRUE~"DCM")
+  ) %>%
+  mutate(rosette=case_when(rosette=="?"~NA, rosette=="43026.0"~"10", rosette=="16- 24"~"16")) %>%
+  mutate(rosette=as.numeric(rosette)) %>%
+  # Fixing likely transcription error - sample looks like 15m samples and
+  # all other triplicates were 19, 20, and 21
+  mutate(rosette=ifelse(rosette==4 & cast==3 & station=="6", 21, rosette))
+
+col_names <- c("station", "cast", "depth", "sample_id",
+               "volume", "start_time", "end_time", "rosette", 
+               "dissolved")
+col_types <- c("text", "text", "text", "text",
+               "numeric", "date", "date", "numeric", 
+               "text", "skip")
+
+MC_log <- read_xlsx("raw_data/MESO_SCOPE_Cruise_KM1709_Sample_Log.xlsx",
+          sheet = "Metabolites_High_Resolution", skip = 2,
+          col_types = col_types, col_names=col_names) %>%
+  filter(!is.na(station)) %>%
+  mutate(station=str_remove(station, "^S")) %>%
+  mutate(cast=as.numeric(str_remove(cast, "^C"))) %>%
+  mutate(sample_id=str_replace(sample_id, "\\-", "less")) %>%
+  mutate(sample_id=str_replace(sample_id, "\\+", "plus")) %>%
+  mutate(depth_shift=str_extract(depth, "(\\-|\\+)\\d+")) %>%
+  mutate(depth_shift=ifelse(is.na(depth_shift), 0, as.numeric(depth_shift))) %>%
+  mutate(abs_depth=case_when(
+    station=="L1"~108+depth_shift,
+    station=="L2"~123+depth_shift
+  )) %>%
+  select(-depth_shift) %>%
+  mutate(depth=str_replace(depth, "\\-", "less")) %>%
+  mutate(depth=str_replace(depth, "\\+", "plus")) %>%
+  mutate(rosette=as.numeric(rosette)) %>%
+  mutate(sample_id=str_remove(sample_id, "^MHR")) %>%
+  mutate(date=as.Date(case_when(
+    station=="L1"~"2017-07-08",
+    station=="L2"~"2017-07-09"
+  )))
+
+samp_log <- bind_rows(falkor_log, MC_log, MT_log)
+dcm_depths <- samp_log %>%
+  distinct(station, depth, abs_depth) %>%
+  filter(depth=="DCM") %>%
+  select(station, dcm_depth=abs_depth)
+```
+
+```{r station locations and CTD summaries}
 col_names <- c("station", "cast", "month", "day", "year", "time", 
@@
                 "lon_deg", "lon_min", "lon_dir", "max_depth", "bottles")
-  read_table(url, skip = 3, col_names = col_names, col_types = cols(
+col_types <- cols(
   station = col_character(),
@@
   max_depth = col_double(),
-    bottles = col_double())) %>%
+  bottles = col_double())
+toDecDeg <- function(deg, min, dir){
+  (deg+min/60)*ifelse(dir%in%c("S", "W"), -1, 1)
+}
+parseCTD <- function(file_path){
+  file_path %>%
+    read_table(skip = 3, col_names = col_names, col_types=col_types) %>%
     mutate(time=strptime(paste(month, day, year, time), format = "%b %d %Y %X")) %>%
@@
 }
-
-if(!file.exists("raw_data/falkor_SCOPEweb_ctd_summary.csv")){
-  # Data source: scope.soest.hawaii.edu/collaborators/datainventory/Data/SCOPEcore/FK_CTDsummary_current.txt
-  falkor_SCOPEweb_ctd_summary_url <- paste0("http://", scope_uname, ":", scope_pw,
-                "@scope.soest.hawaii.edu/collaborators/datainventory/Data/",
-                "SCOPEcore/FK_CTDsummary_current.txt")
-  falkor_SCOPEweb_ctd_summary <- getScopeCTD(falkor_SCOPEweb_ctd_summary_url)
-  falkor_SCOPEweb_ctd_summary <- filter(falkor_SCOPEweb_ctd_summary, time > "2018-03-23 04:01:51")
-  write.csv(falkor_SCOPEweb_ctd_summary, 
-            file = "raw_data/falkor_SCOPEweb_ctd_summary.csv", 
-            row.names = FALSE)
-}
-if(!file.exists("raw_data/meso_SCOPEweb_ctd_summary.csv")){
-  # Data url source: scope.soest.hawaii.edu/collaborators/datainventory/Data/SCOPEcore/MS_CTDsummary_current.txt
-  meso_SCOPEweb_ctd_summary_url <- paste0("http://", scope_uname, ":", scope_pw,
-                                 "@scope.soest.hawaii.edu/collaborators/datainventory/Data",
-                                 "/SCOPEcore/MS_CTDsummary_current.txt")
-  meso_SCOPEweb_ctd_summary <- getScopeCTD(meso_SCOPEweb_ctd_summary_url)
-  write.csv(meso_SCOPEweb_ctd_summary, 
-            file = "raw_data/meso_SCOPEweb_ctd_summary.csv", 
-            row.names = FALSE)
-}
-
-# SLA data is pulled down from CMAP for Falkor, FTP for meso ----
-# SLA units are meters
-if(!file.exists("raw_data/falkor_SLA.csv")){
-  falkor_cruisename = "SCOPE_Falkor2"
-  falkor_cruise_data <- get_cruise_by_name(falkor_cruisename)
-  falkor_CMAP_SLA <- get_spacetime(tableName = 'tblAltimetry_REP',
-                     varName = 'sla',
-                     dt1=as.character(falkor_cruise_data$Start_Time),
-                     dt2=as.character(falkor_cruise_data$End_Time),
-                     lat1=falkor_cruise_data$Lat_Min,
-                     lat2=falkor_cruise_data$Lat_Max,
-                     lon1=falkor_cruise_data$Lon_Min,
-                     lon2=falkor_cruise_data$Lon_Max)
-
-  falkor_ctddata <- read.csv("raw_data/falkor_SCOPEweb_ctd_summary.csv")
-  falkor_SLA <- falkor_ctddata %>%
-    mutate(time=as.POSIXct(time)) %>%
-    group_by(station) %>%
-    summarize(lat=median(lat), lon=median(lon), time=median(time)) %>%
-    difference_left_join(falkor_CMAP_SLA, max_dist = 0.125, by=c("lat", "lon")) %>%
-    filter(as.Date(time.x)==as.Date(time.y)) %>%
-    select(station, sla)
-
-  write.csv(falkor_SLA, file="raw_data/falkor_SLA.csv", row.names = FALSE)
-}
-if(!file.exists("raw_data/meso_SLA.csv")){
-  meso_SLA <- "http://scope.soest.hawaii.edu/data/mesoscope/km1709_CtdSla.txt" %>%
-    read.table(skip = 1, col.names = c("station", "cast", "sla")) %>%
-    mutate(sla=sla/100)
-  write.csv(meso_SLA, file = "raw_data/meso_SLA.csv", row.names = FALSE)
-}
-
-# Combine and write out ----
-falkor_SLA <- read.csv("raw_data/falkor_SLA.csv") %>%
-  mutate(station=as.character(station)) %>%
-  mutate(cast=1)
-meso_SLA <- read.csv("raw_data/meso_SLA.csv")
-dcm_depths <- read.csv("raw_data/all_DCMs.csv")
-falkor_ctddata <- read.csv("raw_data/falkor_SCOPEweb_ctd_summary.csv")
-meso_ctddata <- read.csv("raw_data/meso_SCOPEweb_ctd_summary.csv")
-
-station_metadata <- meso_ctddata %>%
-  mutate(cruise=ifelse(startsWith(station, "L")|station==15, "MC", "MT")) %>%
-  rbind(falkor_ctddata %>% mutate(cruise="FK")) %>%
-  left_join(dcm_depths, by=c("station", "cast")) %>%
-  left_join(rbind(meso_SLA, falkor_SLA), by=c("station", "cast")) %>%
-  filter(!(cruise=="FK" & station < 28)) %>%
-  select(cruise, station, cast, time, lat, lon, DCM_depth, sla) %>%
-  mutate(time=as.POSIXct(time, tz = "UTC")) %>% 
-  mutate(local_hour=as.numeric(format(time, tz="HST", format = "%H")))
-
-write.csv(station_metadata, file = "made_data/station_metadata.csv", row.names = FALSE)
+meso_ctddata <- parseCTD("raw_data/MS_CTDsummary_current.txt")
+falkor_ctddata <- parseCTD("raw_data/FK_CTDsummary_current.txt") %>%
+  filter(time > "2018-03-23 04:01:51")
+complete_ctddata <- bind_rows(meso_ctddata, falkor_ctddata)
 ```
 
-```{r all_core and bottle_depth_data}
-# Core Water Column Data from .gof file on the internet
-if(!file.exists("raw_data/falkor_core.txt")){
-  fk_meta <- "http://scope.soest.hawaii.edu/FTP/scope/water/fk180310.gof"
-  download.file(fk_meta, destfile = "raw_data/falkor_core.txt")
-}
-if(!file.exists("raw_data/mesoscope_core.txt")){
-  meso_meta <- "http://scope.soest.hawaii.edu/FTP/scope/water/mesoscope.gof"
-  download.file(meso_meta, destfile = "raw_data/mesoscope_core.txt")
-}
-meso_core <- read_table("raw_data/mesoscope_core.txt", skip = 4, 
-                        col_names = FALSE) %>%
-  set_names(c("station", "cast", "rosette", "lat", "lon", "pressure", "temp",
+```{r SLA parsing}
+col_types <- cols(
+  stn = col_character(),
+  cast = col_double(),
+  sla_corr = col_double()
+)
+falkor_sla <- read_csv("raw_data/fk180310_ctd_sla.txt", col_types = col_types) %>%
+  mutate(cruise="FK")
+meso_sla <- read_csv("raw_data/km1709_ctd_sla.txt", col_types = col_types) %>%
+  mutate(cruise=ifelse(stn%in%c("15", "L1", "L2"), "MC", "MT"))
+complete_sla <- bind_rows(falkor_sla, meso_sla) %>%
+  rename(station="stn")
+```
+
+```{r Core water column parsing (.gof files)}
+col_names <- c("station", "cast", "rosette", "lat", "lon", "pressure", "temp",
               "sal", "ctd_oxy", "ctd_chl", "theta", "sigma", "oxy_um", "dic", "alk", "ph",
@@
               "pro_abund", "syn_abund", "euk_abund",
-              "q1", "q2", "q3", "q4")) %>%
-  mutate(across(everything(), function(x)ifelse(x==-9, NA, x))) %>%
+              "q1", "q2", "q3", "q4")
+qual_names <- setdiff(col_names, c(head(col_names, 6), tail(col_names, 4)))
+gofReader <- function(filename){
+  # From ftp://ftp.soest.hawaii.edu/hot/water/Readme.water.jgofs
+  # Quality Indicators:
+  # Flag  Meaning
+  # 1    not quality controled
+  # 2    good data
+  # 3    suspect (i.e.  questionable) data
+  # 4    bad data
+  # 5    missing data
+  # 9    variable not measured during this cast
+  raw_gof <- read_table(filename, skip = 4, col_names = col_names,
+                        col_types = cols(.default = col_double()))
+  qual_df <- select(raw_gof, matches("^q\\d+$")) %>%
+    unite("value", sep = "") %>%
+    separate(sep = "", col = "value", into = c(NA, qual_names)) %>%
+    mutate(across(everything(), as.numeric))
+  raw_gof[,qual_names][qual_df!=2] <- NA
+  raw_gof %>% select(-matches("^q\\d+$"))
+}
+meso_core <- gofReader("raw_data/mesoscope.gof") %>%
   mutate(cruise=case_when(
@@
     station == 17 ~ "L2"
-  )) %>%
-  select(cruise, station, cast, rosette, pressure, 
-         temp, sal, theta, dic, alk, ph, oxy_um, ends_with("abund"))
-fk_core <- read_table("raw_data/falkor_core.txt", skip = 4, 
-                        col_names = FALSE) %>%
-  set_names(c("station", "cast", "rosette", "lat", "lon", "pressure", "temp",
-              "sal", "ctd_oxy", "ctd_chl", "theta", "sigma", "oxy_um", "dic", "alk", "ph",
-              "po4_um", "no23_um", "sio4_um", "LLN", "LLP", "PC", "PN", "PP",
-              "PSi", "chl_ug", "phaeo_ug", "het_abund", 
-              "pro_abund", "syn_abund", "euk_abund",
-              "q1", "q2", "q3", "q4")) %>%
-  mutate(across(everything(), function(x)ifelse(x==-9, NA, x))) %>%
-  mutate(cruise="FK") %>%
+  ))
+falkor_core <- gofReader("raw_data/fk180310.gof") %>%
   filter(station>28) %>%
   mutate(station=as.character(station)) %>%
-  select(cruise, station, cast, rosette, pressure, 
-         temp, sal, theta, dic, alk, ph, oxy_um, ends_with("abund"))
-all_core <- bind_rows(fk_core, meso_core)
-write.csv(all_core, "made_data/all_core.csv", row.names = FALSE)
-
-bottle_depth_data <- all_core %>%
-  distinct(cruise, station, cast, rosette, pressure)
-write.csv(bottle_depth_data, "made_data/bottle_depth_data.csv", row.names = FALSE)
+  mutate(cruise="FK")
+all_core <- bind_rows(meso_core, falkor_core)
 ```
 
-```{r plot station_metadata}
-station_metadata <- read_csv("made_data/station_metadata.csv", show_col_types = FALSE)
-stations_summarized <- station_metadata %>%
-  group_by(cruise, station, cast) %>%
-  summarize(lat=median(lat), lon=median(lon), time=median(time), sla=mean(sla)) %>%
-  arrange(time)
+```{r Parsing beam attenuation data}
+beam_long <- read.csv("raw_data/MESO-SCOPE_BeamAttenuation_forWill.csv", 
+                    skip = 3, header = FALSE) %>%
+  `colnames<-`(paste(.[1,], .[2,], sep = " ")) %>%
+  tail(-2) %>%
+  mutate(across(everything(), as.numeric)) %>%
+  pivot_longer(cols = -1, values_to = "beam_atten") %>%
+  filter(!is.na(beam_atten)) %>%
+  separate(name, into = c("station", "cast"), sep = " ") %>%
+  rename(abs_depth=`NaN NaN`) %>%
+  mutate(cast=as.numeric(cast))
 
-station_labels <- stations_summarized %>%
-  filter(station%in%c(seq(1, 15, 2), "L1", "L2", 62, 64, 77, 80)) %>%
-  filter(cast==1)
-
-gm <- get_map(location = c(-158, 24), zoom = 6)
-map_gp <- ggmap(ggmap = gm) +
-  geom_point(aes(x=lon, y=lat), size=4, color="#FFFFFFAA", data = station_metadata) +
-  geom_path(aes(x=lon, y=lat), data = station_metadata) +
-  geom_point(aes(x=lon, y=lat, color=sla), data = station_metadata) +
-  geom_label_repel(aes(x=lon, y=lat, label=station), color="steelblue", 
-                   data = station_labels, min.segment.length = 0) +
-  scale_color_gradient2(low = muted("blue"), high = muted("red"), mid = "grey80") +
-  facet_wrap(~factor(cruise, levels = c("MT", "MC", "FK"), 
-                     labels = c("Eddy transect", "Eddy center", "Falkor")), 
-             scales = "free_x", nrow=1) +
-  theme(legend.position = "none") +
-  coord_cartesian(xlim = c(-162, -155), ylim=c(21, 28))
-
-sla_ts_gp <- station_metadata %>%
-  ggplot() +
-  geom_point(aes(x=time, y=sla, color=sla)) +
-  scale_color_gradient2(low = muted("blue"), high = muted("red"), mid = "grey80") +
-  facet_wrap(~factor(cruise, levels = c("MT", "MC", "FK"), 
-                     labels = c("Eddy transect", "Eddy center", "Falkor")), 
-             scales = "free_x", nrow=1) +
-  theme_bw() +
-  theme(legend.position = c(0.5, 0.55), axis.title.x = element_blank())
-station_metadata_map <- grid.arrange(map_gp, sla_ts_gp)
-ggsave(plot = station_metadata_map, filename = "station_metadata_map.png", 
-       path = "figures", device = "png", width = 8, height = 6, dpi = 144, 
-       type = "cairo")
+# Convert to bottle locations
+beam_core <- all_core %>%
+  group_by(station, cast) %>%
+  # filter(station==4, cast==2) %>%
+  group_split() %>% #`[[`(1) -> core_subset
+  map(function(core_subset){
+    beam_subset <- beam_long %>%
+      filter(station==unique(core_subset$station), cast==unique(core_subset$cast))
+    if(nrow(beam_subset)==0)return(NULL)
+    approxout <- with(beam_subset, approx(abs_depth, beam_atten, core_subset$pressure, ties = mean))
+    core_subset %>%
+      select(cruise, station, cast, rosette, pressure) %>%
+      left_join(set_names(as.data.frame(approxout), c("pressure", "beam_atten")), 
+                by="pressure", multiple="all")
+  }, .progress = TRUE) %>%
+  bind_rows() %>%
+  distinct() %>%
+  filter(beam_atten<0.2)
 ```
 
-## Other bottle data (cruise, station, cast, depth, PC, PN, PO4, N+N, DON, DOP, ATP, chl)
-
-This data is mostly available on the SCOPE website via Excel files
-
-```{r raw data download}
-if(!file.exists("raw_data/Karl_FK_PCPN_WC_current.xlsx")){
-  # Data source: http://scope.soest.hawaii.edu/collaborators/datainventory/Data/Karl/Karl_FK_Nutrients_Final.xlsx
-  falkor_CN_url <- paste0("http://", scope_uname, ":", scope_pw,
-                "@scope.soest.hawaii.edu/collaborators/datainventory/",
-                "Data/Karl/Karl_FK_PCPN_WC_current.xlsx")
-  download.file(falkor_CN_url, destfile = "raw_data/Karl_FK_PCPN_WC_current.xlsx", mode = "wb")
-}
-if(!file.exists("raw_data/Karl_FK_Nutrients_Final.xlsx")){
-  # Data source: http://scope.soest.hawaii.edu/collaborators/datainventory/Data/Karl/Karl_FK_Nutrients_Final.xlsx
-  falkor_nutrient_url <- paste0("http://", scope_uname, ":", scope_pw,
-                "@scope.soest.hawaii.edu/collaborators/datainventory/Data/",
-                "Karl/Karl_FK_Nutrients_Final.xlsx")
-  download.file(falkor_nutrient_url, destfile = "raw_data/Karl_FK_Nutrients_Final.xlsx", mode = "wb")
-}
-if(!file.exists("raw_data/2017MS_CoreData_current.xls")){
-  # Data source: http://scope.soest.hawaii.edu/collaborators/datainventory/Data/SCOPEcore/2017MS_CoreData_current.xls
-  transect_pcpn_url <- paste0("http://", scope_uname, ":", scope_pw,
-                "@scope.soest.hawaii.edu/collaborators/datainventory/Data/",
-                "SCOPEcore/2017MS_CoreData_current.xls")
-  download.file(transect_pcpn_url, destfile = "raw_data/2017MS_CoreData_current.xls", mode = "wb")
-}
-if(!file.exists("raw_data/Karl_MS_Nutrient_Summary_KB.xlsx")){
-  # Data source: http://scope.soest.hawaii.edu/collaborators/datainventory/Data/Karl/Karl_MS_Nutrient_Summary_KB.xlsx
-  transect_nutrient_url <- paste0("http://", scope_uname, ":", scope_pw,
-                "@scope.soest.hawaii.edu/collaborators/datainventory/Data/Karl/",
-                "Karl_MS_Nutrient_Summary_KB.xlsx")
-  download.file(transect_nutrient_url, destfile = "raw_data/Karl_MS_Nutrient_Summary_KB.xlsx", mode = "wb")
-}
-if(!file.exists("raw_data/Karl_MS_pATP.xlsx")){
-  # Data source: http://scope.soest.hawaii.edu/collaborators/datainventory/Data/Karl/Karl_MS_pATP.xlsx
-  atp_transect_url <- paste0("http://", scope_uname, ":", scope_pw,
-         "@scope.soest.hawaii.edu/collaborators/datainventory/",
-         "Data/Karl/Karl_MS_pATP.xlsx")
-  download.file(atp_transect_url, destfile = "raw_data/Karl_MS_pATP.xlsx", mode = "wb")
-}
-```
-
-```{r all_pcn}
+```{r PCPN parsing}
 fk_pcn <- "raw_data/Karl_FK_PCPN_WC_current.xlsx" %>%
@@
   filter(is.na(Comments)) %>%
-  mutate(station=as.numeric(str_extract(Name, "(?<=FK-)\\d+"))) %>%
-  filter(station>=28) %>%
+  mutate(station=str_extract(Name, "(?<=FK-)\\d+")) %>%
   mutate(rosette=as.numeric(str_extract(Name, "\\d+$"))) %>%
@@
     station == 17 ~ "L2"
-  ))
-all_pcn <- rbind(meso_pcn, fk_pcn)
-write.csv(all_pcn, "made_data/all_pcn.csv", row.names = FALSE)
+  )) %>%
+  mutate(rosette=case_when(
+    station=="L1"&cast==13&rosette==10~20, 
+    station=="L1"&cast==6&rosette==15~16,
+    station=="L1"&cast==6&rosette==14~15,
+    TRUE~rosette))
+all_pcn <- bind_rows(meso_pcn, fk_pcn)
 ```
 
-```{r plot all_pcn}
-pc_gp <- all_pcn %>%
-  mutate(cruise=factor(cruise, levels = c("MT", "MC", "FK"), 
-                   labels = c("Eddy transect", "Eddy center", "Falkor"))) %>%
-  mutate(station=factor(station, levels=unique(station))) %>%
-  ggplot(aes(x=PC_um, y=abs_depth, color=station, group=interaction(station, cast))) +
-  geom_path() +
-  geom_point() +
-  facet_wrap(~cruise, scales="free_x") +
-  ggtitle("Particulate carbon") +
-  scale_y_reverse(name="Depth")  +
-  xlim(0, NA) +
-  theme(axis.title.x = element_blank(), legend.position = "bottom") +
-  guides(color=guide_legend(nrow=2,byrow=TRUE, title = "Station"))
-pn_gp <- all_pcn %>%
-  mutate(cruise=factor(cruise, levels = c("MT", "MC", "FK"), 
-                   labels = c("Eddy transect", "Eddy center", "Falkor"))) %>%
-  mutate(station=factor(station, levels=unique(station))) %>%
-  ggplot(aes(x=PN_um, y=abs_depth, color=station, group=interaction(station, cast))) +
-  geom_path() +
-  geom_point() +
-  facet_wrap(~cruise, scales="free_x") +
-  ggtitle("Particulate nitrogen") +
-  scale_y_reverse(name="Depth")  +
-  xlim(0, NA) +
-  theme(axis.title.x = element_blank(), legend.position = "none")
-all_pcn_plot <- grid.arrange(
-  pc_gp, pn_gp, layout_matrix=matrix(c(1,1,1,1,2,2,2), ncol = 1)
-)
-ggsave(plot = all_pcn_plot, filename = "all_pcn_plot.png", 
-       path = "figures", device = "png", width = 8, height = 6, dpi = 144, 
-       type = "cairo")
-```
+```{r Nutrient parsing}
+# Not really necessary because .gof file has nutrients in it already
+# These files have more sigfigs but less depth accuracy (5, 25 instead of 4.7, 25.2)
+# Actually, very necessary because the Falkor .gof file has N+N and PO4
+# values switched??? Compare to Karl_FK_Nutrients_Final.xlsx
+all_core %>%
+  filter(cruise=="FK") %>%
+  filter(!is.na(no23_um)) %>%
+  filter(pressure>160) %>%
+  select(station, cast, rosette, pressure, no23_um, po4_um)
 
-```{r all_nuts}
-fk_nuts <- "raw_data/Karl_FK_Nutrients_Final.xlsx" %>%
+falkor_nuts <- "raw_data/Karl_FK_Nutrients_Final.xlsx" %>%
   read_excel(skip = 3) %>%
@@
   select(-sample) %>%
-  mutate(cast=1, cruise="FK")
+  mutate(cast=1, cruise="FK") %>%
+  arrange(station, abs_depth)
+
 transect_profile_sheet <- "raw_data/Karl_MS_Nutrient_Summary_KB.xlsx" %>%
@@
   slice(-1:-2) %>%
-  mutate(sd=as.numeric(sd))
+  mutate(sd=as.numeric(sd)) %>%
+  mutate(Si=case_when(
+    Station=="8" & Cast==2 & Niskin==4 ~ NA,
+    Station=="10" & Cast==1 & Niskin==3 ~ NA,
+    Station=="14" & Cast==1 & Niskin==8 ~ NA,
+    TRUE~Si
+  )) %>%
+  mutate(LLP=case_when(
+    Station=="L2" & Cast==26 & Niskin==1 ~ NA,
+    TRUE~LLP
+  ))
 highres_sheet <- "raw_data/Karl_MS_Nutrient_Summary_KB.xlsx" %>%
   read_excel(range = "KM1709 High res nuts!A4:P53") %>%
-  slice(-1:-2)
+  slice(-1:-2) %>%
+  filter(!is.na(Station))
 diel_sheet <- "raw_data/Karl_MS_Nutrient_Summary_KB.xlsx" %>%
   read_excel(range = "KM1709 Diel nuts!A4:O76") %>%
-  slice(-1:-3)
+  slice(-1:-3) %>%
+  filter(!is.na(Station)) %>%
+  mutate(LLP=case_when(
+    Station=="L1"&Cast==11&Niskin...3==20~NA,
+    Station=="L1"&Cast==19&Niskin...3==9~NA,
+    TRUE~LLP
+  )) %>%
+  mutate(Si=case_when(
+    Station=="L1" & Cast==10 & Niskin...3==10 ~ NA,
+    Station=="L1" & Cast==15 & Niskin...3==10 ~ NA,
+    Station=="L1" & Cast==17 & Niskin...3==11 ~ NA,
+    Station=="L2" & Cast==6 & Niskin...3==12 ~ NA,
+    Station=="L2" & Cast==12 & Niskin...3==12 ~ NA,
+    Station=="L2" & Cast==17 & Niskin...3==12 ~ NA,
+    Station=="L2" & Cast==20 & Niskin...3==18 ~ NA,
+    Station=="L2" & Cast==25 & Niskin...3==12 ~ NA,
+    Station=="L2" & Cast==11 & Niskin...3==3 ~ NA,
+    TRUE~Si
+  ))
+
 meso_nuts <- bind_rows(transect_profile_sheet, highres_sheet, diel_sheet) %>% 
@@
   mutate(po4_um=ifelse(po4_um=="BLD"|po4_um=="NA"|is.na(po4_um), LLP, po4_um)) %>%
+  select(-LLN, -LLP) %>%
   mutate(across(-station, as.numeric)) %>%
@@
   mutate(no23_um=ifelse(is.na(no23_um) & cruise=="MC", 0, no23_um)) %>%
-  # Fixing a few rosette numbers per Benedetto conversation
+  # Fixing a few rosette numbers
   mutate(rosette=case_when(
@@
   ))
-all_nuts <- bind_rows(meso_nuts, fk_nuts) %>%
-  left_join(station_metadata) %>%
-  arrange(time) %>%
-  select(cruise, station, cast, rosette, abs_depth, no23_um, po4_um, 
-         DON_um, DOP_um, si_um) %>%
-  mutate(station=factor(station, levels=unique(station)))
-write.csv(all_nuts, "made_data/all_nuts.csv", row.names = FALSE)
+# For determining the above
+# all_core %>% 
+#   filter(station%in%c("13", "14")) %>% 
+#   select(station, rosette, pressure) %>% 
+#   filter(pressure>180)
+all_nuts <- bind_rows(falkor_nuts, meso_nuts)
 ```
 
-```{r plot all_nuts}
-refactored_nuts <- all_nuts %>%
-  mutate(cruise=factor(cruise, levels = c("MT", "MC", "FK"), 
-         labels = c("Eddy transect", "Eddy center", "Falkor")))
-
-refactored_nuts %>%
-  filter(is.na(no23_um))
-
-no23_gp <- refactored_nuts %>%
-  ggplot(aes(x=no23_um, y=abs_depth, color=station, group=interaction(station, cast))) +
-  geom_path() +
-  geom_point() +
-  facet_wrap(~cruise, scales="free_x") +
-  ggtitle("Nitrate + nitrite") +
-  scale_y_reverse(name="Depth")  +
-  xlim(0, NA) +
-  theme(axis.title.x = element_blank(), legend.position = "none")
-po4_gp <- refactored_nuts %>%
-  ggplot(aes(x=po4_um, y=abs_depth, color=station, group=interaction(station, cast))) +
-  geom_path() +
-  geom_point() +
-  facet_wrap(~cruise, scales="free_x") +
-  ggtitle("Phosphate") +
-  scale_y_reverse(name="Depth")  +
-  xlim(0, NA) +
-  theme(axis.title.x = element_blank(), legend.position = "none")
-DON_gp <- refactored_nuts %>%
-  ggplot(aes(x=DON_um, y=abs_depth, color=station, group=interaction(station, cast))) +
-  geom_path() +
-  geom_point() +
-  facet_wrap(~cruise, scales="free_x") +
-  ggtitle("Dissolved organic nitrogen") +
-  scale_y_reverse(name="Depth")  +
-  xlim(0, NA) +
-  theme(axis.title.x = element_blank(), legend.position = "none")
-DOP_gp <- refactored_nuts %>%
-  ggplot(aes(x=DOP_um, y=abs_depth, color=station, group=interaction(station, cast))) +
-  geom_path() +
-  geom_point() +
-  facet_wrap(~cruise, scales="free_x") +
-  ggtitle("Dissolved organic phosphorus") +
-  scale_y_reverse(name="Depth")  +
-  xlim(0, NA) +
-  theme(axis.title.x = element_blank(), legend.position = "none") +
-  guides(color=guide_legend(nrow=2,byrow=TRUE, title = "Station"))
-si_gp <- refactored_nuts %>%
-  ggplot(aes(x=si_um, y=abs_depth, color=station, group=interaction(station, cast))) +
-  geom_path() +
-  geom_point() +
-  facet_wrap(~cruise, scales="free_x") +
-  ggtitle("Silica") +
-  scale_y_reverse(name="Depth")  +
-  xlim(0, NA) +
-  theme(axis.title.x = element_blank(), legend.position = "bottom") +
-  guides(color=guide_legend(nrow=2,byrow=TRUE, title = "Station"))
-all_nuts_plot <- grid.arrange(
-  no23_gp, po4_gp, DON_gp, DOP_gp, si_gp,
-  layout_matrix=matrix(rep(1:5, c(3,3,3,3,5)), ncol = 1)
-)
-ggsave(plot = all_nuts_plot, filename = "all_nuts_plot.png", 
-       path = "figures", device = "png", width = 6, height = 8, dpi = 144, 
-       type = "cairo")
-```
-
-```{r meso_chl}
+```{r Chlorophyll parsing}
 meso_chl <- "raw_data/2017MS_CoreData_current.xls" %>%
@@
 
-# Deciding to replace the now-missing value with average of nearby stations 
-# instead bc interpolation sux
-s14_chl_surf_row <- meso_chl %>% 
-  filter(station%in%c("12", "14")) %>%
-  left_join(bottle_depth_data) %>%
-  filter(pressure<30) %>%
-  summarise(station="13", cast=2, rosette=23, cruise="MT", 
-            chl_ug=round(mean(chl_ug), 5), phaeo_ug=round(mean(phaeo_ug), 5))
-meso_chl <- bind_rows(meso_chl, s14_chl_surf_row)
+falkor_chl <- "raw_data/Karl_FK_Chlorophyll.xlsx" %>%
+  read_excel(skip = 19, sheet = "10au") %>%
+  slice(-1) %>%
+  select(station=Station, cast=Cast, rosette=Bottle, chl_ug=`chla (ug/)l`, phaeo_ug=`phaeo (ug/l)`, abs_depth=Depth) %>%
+  mutate(cruise="FK") %>%
+  select(-abs_depth) %>%
+  mutate(station=as.character(station))
 
-write.csv(meso_chl, "made_data/meso_chl.csv", row.names = FALSE)
+all_chl <- bind_rows(meso_chl, falkor_chl)
 ```
 
-```{r plot meso_chl}
-station_metadata %>%
-  right_join(meso_chl) %>% 
-  left_join(bottle_depth_data) %>%
-  arrange(time) %>%
-  mutate(station=factor(station, levels = unique(station))) %>%
-  ggplot(aes(x=chl_ug, y=pressure, color=station, group=interaction(station, cast))) + 
-  geom_point() + 
-  geom_path() + 
-  scale_y_reverse() +
-  theme_bw() +
-  theme(legend.position = "left") +
-  facet_wrap(~cruise,nrow = 1)
-ggsave(filename = "meso_chl_plot.png", path = "figures", device = "png", 
-       width = 8, height = 6, dpi = 144, type = "cairo")
+```{r PARsing}
+light_long <- read.csv("raw_data/MESO-SCOPE_PercentPAR_forWill.csv", 
+                    skip = 7, header = FALSE) %>%
+  `colnames<-`(c("abs_depth", paste0("L1_", 1:6), paste0("L2_", 1:6))) %>%
+  mutate(across(everything(), as.numeric)) %>%
+  pivot_longer(cols = -1, values_to = "beam_atten") %>%
+  filter(!is.na(beam_atten)) %>%
+  separate(name, into = c("station", "rep"), sep = "_") %>%
+  mutate(rep=as.numeric(rep))
 ```
 
-```{r beam_core from Benedetto}
-if(!file.exists("raw_data/MESO-SCOPE_BeamAttenuation_forWill.csv")){
-  stop("Go download the beam data from email, search 'beam attenuation'")
-}
-beam_dat <- read.csv("raw_data/MESO-SCOPE_BeamAttenuation_forWill.csv", 
-                    skip = 3, header = FALSE)
-beam_station_cast <- paste0(beam_dat[1,], " ", beam_dat[2,])
-beam_dat <- beam_dat[3:nrow(beam_dat),]
-colnames(beam_dat) <- beam_station_cast
-beam_dat[] <- sapply(beam_dat, as.numeric)
-beam_long <- beam_dat %>%
-  pivot_longer(cols = 2:ncol(beam_dat), values_to = "beam_atten") %>%
-  filter(!is.na(beam_atten)) %>%
-  separate(name, into = c("station", "cast"), sep = " ") %>%
-  rename(pressure=`NaN NaN`) %>%
-  mutate(cast=as.numeric(cast))
-
-beam_core <- all_core %>%
-  group_by(station, cast) %>%
-  group_split() %>% #`[[`(1) -> core_subset
-  map_dfr(function(core_subset){
-    beam_subset <- beam_long %>%
-      filter(station==unique(core_subset$station), cast==unique(core_subset$cast))
-    if(nrow(beam_subset)==0)return(NULL)
-    approxout <- with(beam_subset, approx(pressure, beam_atten, core_subset$pressure))
-    core_subset %>%
-      select(cruise, station, cast, rosette, pressure) %>%
-      left_join(set_names(as.data.frame(approxout), c("pressure", "beam_atten")), 
-                by="pressure")
-  })
-write.csv(beam_core, "made_data/beam_core.csv", row.names = FALSE)
+```{r Create complete_env_data by doing hella joins}
+complete_env_data <- all_core %>%
+  select(cruise, station, cast, rosette, lat, lon, pressure,
+         temp, sal, ctd_oxy, ctd_chl, theta, sigma, oxy_um,
+         dic, alk, ph, ends_with("abund")) %>%
+  left_join(dcm_depths) %>%
+  left_join(complete_ctddata %>% select(station, cast, time)) %>%
+  # filter(station=="1", cast==1) %>%
+  left_join(complete_sla) %>%
+  left_join(all_pcn) %>% 
+  left_join(all_nuts %>% select(-abs_depth)) %>%
+  left_join(all_chl) %>%
+  left_join(beam_core) %>%
+  mutate(abs_depth=ifelse(is.na(abs_depth), pressure, abs_depth))
 ```
 
-```{r plot beam_core and beam_long}
-beam_long %>%
-  filter(station%in%c("L1", "L2")) %>%
-  filter(beam_atten < 0.1) %>%
-  ggplot(aes(x=beam_atten, y=pressure, group=cast, color=beam_atten)) +
-  geom_path() +
-  facet_wrap(~station) +
-  scale_color_viridis_c() +
-  scale_y_reverse()
-last_plot() +
-  geom_point(data=beam_core %>% filter(station%in%c("L1", "L2")), size=1.5)
-```
+## MS file metadata creation and import
 
-```{r light_core from Benedetto}
-if(!file.exists("raw_data/MESO-SCOPE_DailyPAR_forWill.csv")){
-  stop("Go download the light data from email, search 'beam attenuation'")
-}
-if(!file.exists("raw_data/MESO-SCOPE_PercentPAR_forWill.csv")){
-  stop("Go download the light data from email, search 'beam attenuation'")
-}
-PAR_dat <- read.csv("raw_data/MESO-SCOPE_DailyPAR_forWill.csv", 
-                    skip = 5, header = FALSE)
-PAR_month_day <- paste0(PAR_dat[1,], " ", PAR_dat[2,])
-PAR_month_day_pseudocast <- paste(PAR_month_day, c(1, rep(1:2, 6)))
-PAR_dat <- PAR_dat[3:nrow(PAR_dat),]
-colnames(PAR_dat) <- PAR_month_day_pseudocast
-PAR_dat[] <- sapply(PAR_dat, as.numeric)
-PAR_long <- PAR_dat %>%
-  pivot_longer(cols = 2:ncol(PAR_dat), values_to = "PAR") %>%
-  filter(!is.na(PAR)) %>%
-  separate(name, into = c("month", "day", "pseudocast"), sep = " ") %>%
-  rename(pressure=`NaN NaN 1`) %>%
-  mutate(station=ifelse(as.numeric(day)>8, "L1", "L2")) %>%
-  group_by(station, pressure) %>%
-  summarise(PAR=mean(PAR))
-
-PAR_perc_dat <- read.csv("raw_data/MESO-SCOPE_PercentPAR_forWill.csv", 
-                    skip = 5, header = FALSE)
-PAR_perc_month_day <- paste0(PAR_perc_dat[1,], " ", PAR_perc_dat[2,])
-PAR_perc_month_day_pseudocast <- paste(PAR_perc_month_day, c(1, rep(1:2, 6)))
-PAR_perc_dat <- PAR_perc_dat[3:nrow(PAR_perc_dat),]
-colnames(PAR_perc_dat) <- PAR_perc_month_day_pseudocast
-PAR_perc_dat[] <- sapply(PAR_perc_dat, as.numeric)
-PAR_perc_long <- PAR_perc_dat %>%
-  pivot_longer(cols = 2:ncol(PAR_perc_dat), values_to = "PAR_perc") %>%
-  filter(!is.na(PAR_perc)) %>%
-  separate(name, into = c("month", "day", "pseudocast"), sep = " ") %>%
-  rename(pressure=`NaN NaN 1`) %>%
-  mutate(station=ifelse(as.numeric(day)>8, "L1", "L2")) %>%
-  group_by(station, pressure) %>%
-  summarise(PAR_perc=mean(PAR_perc))
-
-PAR_long <- PAR_long %>%
-  full_join(PAR_perc_long, by=c("station", "pressure"))
-
-light_core <- all_core %>%
-  filter(station%in%c("L1", "L2")) %>%
-  group_by(station) %>%
-  group_split() %>% #`[[`(1) -> core_subset
-  map_dfr(function(core_subset){
-    PAR_subset <- PAR_long %>%
-      filter(station==unique(core_subset$station))
-    PARout <- with(PAR_subset, approx(pressure, PAR, core_subset$pressure))
-    percout <- with(PAR_subset, approx(pressure, PAR_perc, core_subset$pressure))
-    core_subset %>%
-      select(cruise, station, cast, rosette, pressure) %>%
-      left_join(data.frame(pressure=PARout$x, PAR=PARout$y, PAR_perc=percout$y), 
-                by="pressure")
-  })
-light_core %>%
-  ggplot() +
-  geom_point(aes(x=PAR_perc, y=pressure, color=station)) +
-  facet_wrap(~station) +
-  scale_y_reverse()
-write.csv(light_core, "made_data/light_core.csv", row.names = FALSE)
-```
-
-```{r pcn outlier removal (requires beam_core)}
-# Remove two outlier samples because Eric thinks the values are suspicious
-# filter(!(station%in%c("10", "11") & abs_depth==15))
-# Advance prediction is that they will be far off the beam_atten:PC line
-# Plan: replace existing PC values with those from a regression
-# on the non-weird PC values for that given SLA
-
-pcn_init <- all_pcn %>%
-  left_join(bottle_depth_data) %>%
-  filter(cruise=="MT") %>% #filter(station=="4") -> station_subset
-  group_by(station) %>%
-  group_split() %>%
-  map_dfr(function(station_subset){
-    beam_subset <- beam_core %>%
-      filter(station==unique(station_subset$station)) %>%
-      distinct(pressure, beam_atten)
-    approx(beam_subset$pressure, beam_subset$beam_atten, station_subset$pressure) %>%
-      as.data.frame() %>%
-      setNames(c("pressure", "beam_atten")) %>%
-      left_join(station_subset, by="pressure")
-  }) %>%
-  group_by(temp=station%in%c("10", "11") & abs_depth==15) %>%
-  group_split()
-
-beam_pc_model <- lm(PC_um~beam_atten, data = pcn_init[[1]])
-pcn_init[[2]]$PC_um <- predict(beam_pc_model, pcn_init[[2]])
-
-# Plot it to double-check outputs
-plot(PC_um~beam_atten, data = pcn_init[[1]])
-abline(beam_pc_model)
-legend("topleft", legend = paste("R^2 =", round(summary(beam_pc_model)$r.squared, 2)))
-points(pcn_init[[2]]$beam_atten, predict(beam_pc_model, pcn_init[[2]]), col="red")
-
-all_pcn <- bind_rows(pcn_init) %>%
-  select(-temp, -beam_atten, -pressure) %>%
-  bind_rows(all_pcn %>% filter(cruise!="MT")) %>%
-  select(-abs_depth)
-
-write.csv(all_pcn, "made_data/all_pcn.csv", row.names = FALSE)
-```
-
-## Combine all the above into env_metadata
-
-```{r load made_data and make env_metadata}
-station_metadata <- read_csv("made_data/station_metadata.csv")
-all_pcn <- read_csv("made_data/all_pcn.csv")
-all_nuts <- read_csv("made_data/all_nuts.csv")
-meso_chl <- read_csv("made_data/meso_chl.csv")
-all_core <- read_csv("made_data/all_core.csv")
-beam_core <- read_csv("made_data/beam_core.csv")
-light_core <- read_csv("made_data/light_core.csv")
-
-env_metadata <- all_core %>%
-  left_join(station_metadata) %>%
-  left_join(all_nuts) %>%
-  left_join(all_pcn) %>% 
-  left_join(meso_chl) %>%
-  left_join(beam_core) %>%
-  left_join(light_core) %>%
-  arrange(cruise, station, cast, rosette) %>%
-  # Organize columns, not dropping anything
-  select(cruise, station, cast, time, lat, lon, DCM_depth, sla, local_hour,
-         rosette, pressure, abs_depth, temp, sal, theta, dic, alk, ph, oxy_um,
-         PAR, PAR_perc, no23_um, po4_um, DON_um, DOP_um, si_um,
-         PC_um, PN_um, chl_ug, phaeo_ug, beam_atten,
-         everything()) %>%
-  mutate(abs_depth=ifelse(is.na(abs_depth), pressure, abs_depth))
-
-write.csv(env_metadata, "made_data/env_metadata.csv", row.names = FALSE)
-```
-
-```{r check via plots, eval=FALSE}
-env_metadata <- read_csv("made_data/env_metadata.csv")
-
-# Check on cruise, time, lat, lon, sla
-env_metadata %>%
-  arrange(time) %>%
-  distinct(cruise, lon, lat, sla) %>%
-  ggplot(aes(x=lon, y=lat, group=cruise)) +
-  geom_path(aes(color=cruise), lwd=0.7) +
-  geom_point(size=3, aes(color=cruise)) +
-  scale_color_manual(values = hcl.colors(5)[2:4]) +
-  ggnewscale::new_scale_color() +
-  geom_point(aes(color=sla), size=2.2) +
-  coord_fixed() +
-  scale_color_gradient2(low = "steelblue4", high = "firebrick") +
-  theme_minimal()
-
-# PCPN and nutrient data
-env_metadata %>%
-  pivot_longer(c(ends_with("um"))) %>%
-  filter(!is.na(value)) %>%
-  filter(pressure<260) %>%
-  select(cruise, station, cast, pressure, name, value) %>%
-  ggplot(aes(x=value, y=pressure)) +
-  geom_path(aes(group=interaction(station, cast))) +
-  geom_point(aes(color=station)) +
-  facet_grid(cruise~name, scales="free_x") +
-  scale_y_reverse()
-
-env_metadata %>%
-  select(cruise, time, pressure:last_col()) %>%
-  pivot_longer(cols = temp:last_col()) %>%
-  filter(pressure<260) %>%
-  group_by(name) %>%
-  mutate(value=(value-min(value, na.rm = TRUE))/
-           (max(value, na.rm = TRUE)-min(value, na.rm = TRUE))) %>%
-  ungroup() %>%
-  mutate(name=factor(name, levels=unique(name))) %>%
-  mutate(cruise=factor(cruise, levels=c("MT", "MC", "FK"))) %>%
-  filter(!is.na(value)) %>%
-  ggplot() +
-  geom_point(aes(x=time, y=pressure, color=value)) +
-  facet_grid(name~cruise, scales = "free_x") +
-  scale_y_reverse() +
-  scale_color_viridis_c() +
-  theme(legend.position = "none")
-ggsave("figures/env_metadata_all.pdf", device = "pdf", width = 8, height = 20,
-       units = "in")
-```
-
-## Metabolite sample data creation (file_metadata)
-
-```{r samp_log from sample log}
-if(!file.exists("raw_data/MESO_SCOPE_Cruise_KM1709_Sample_Log.xlsx")){
-  stop("Go download the data from email, search 'MESO_SCOPE_Cruise_KM1709_Sample_Log'")
-  # Needs to be converted to .xlsx and downloaded
-}
-bottle_depth_data <- read_csv("made_data/bottle_depth_data.csv")
-MT_samps <- "raw_data/MESO_SCOPE_Cruise_KM1709_Sample_Log.xlsx" %>%
-  read_excel(sheet = "Metabolites_and_Proteins", range = "A2:J114") %>%
-  mutate(cruise="MT") %>%
-  select(cruise, station=Station, cast=Cast, depth=`Depth (m)`, 
-         rosette=`Niskin #`, id=`Sample ID`) %>%
-  filter(complete.cases(.)) %>%
-  mutate(station=str_remove(station, "^S")) %>%
-  mutate(cast=as.numeric(str_remove(cast, "^C"))) %>%
-  # We can grab DCM values here or from the mini DCM spreadsheet I made later
-  # mutate(abs_depth=as.numeric(str_remove(depth, "DCM="))) %>%
-  mutate(depth=str_replace(depth, ".0", "m")) %>%
-  mutate(depth=str_replace(depth, "=.*", "")) %>%
-  # Guessing at rosette number here - doesn't really matter bc we're matching
-  # based on station more than exact rosette mostly and all bottles
-  # were fired at similar depths and times
-  mutate(rosette=ifelse(rosette=="?", 1, rosette)) %>% 
-  mutate(rosette=as.numeric(rosette)) %>%
-  # Fixing likely transcription error - sample looks like 15m samples and
-  # all other triplicates were 19, 20, and 21
-  mutate(rosette=ifelse(rosette==4 & cast==3 & station=="6", 21, rosette)) %>%
-  left_join(bottle_depth_data)
-MC_samps <- "raw_data/MESO_SCOPE_Cruise_KM1709_Sample_Log.xlsx" %>%
-  read_excel(sheet = "Metabolites_High_Resolution", range = "A2:J35") %>%
-  mutate(cruise="MC") %>%
-  select(cruise, station=Station, cast=Cast, depth=`Depth (m)`, 
-         rosette=`Niskin #`, id=`Sample ID`) %>%
-  filter(!is.na(station)) %>%
-  mutate(station=str_remove(station, "^S")) %>%
-  mutate(cast=as.numeric(str_remove(cast, "^C"))) %>%
-  mutate(rosette=as.numeric(rosette)) %>%
-  mutate(depth=str_replace(depth, "\\-", "less")) %>%
-  mutate(depth=str_replace(depth, "\\+", "plus")) %>%
-  mutate(id=str_remove(id, "^MHR")) %>%
-  left_join(bottle_depth_data)
-samp_log <- bind_rows(MT_samps, MC_samps)
-```
-
-```{r file_metadata from mzML list}
-samp_metadata <- list.files("../mzMLs", pattern = "Smp.*mzML", recursive = TRUE) %>%
+```{r Create file data for samples}
+samp_filedata <- list.files("../mzMLs", pattern = "Smp.*mzML", recursive = TRUE) %>%
   data.frame(filename=.) %>%
@@
   mutate(depth=str_extract(filename, "15m|25m|DCM(plus|less)?(10m|20m)?|175m")) %>%
+  mutate(depth=ifelse(depth=="15m"|depth=="25m", "Surface", depth)) %>%
+  mutate(depth=ifelse(depth=="175m", "Deep", depth)) %>%
   mutate(tripl=str_extract(filename, "(?<=_)[A-C](?=.mzML$)")) %>%
@@
   mutate(date_run=strptime(date_run, format = "%y%m%d")) %>%
-  mutate(id=str_extract(filename, "(?<=_)(MS|L).*_[A-C](?=.mzML$)")) %>%
-  mutate(id=str_replace(id, "less", "-")) %>%
-  mutate(id=str_replace(id, "plus", "+"))
-absolute_depths <- samp_metadata %>%
-  distinct(station, depth) %>%
-  mutate(depth_diff=sapply(depth, function(depth){
-    depth_num <- str_extract(depth, "(?<=DCM(plus|less))\\d*")
-    depth_num <- ifelse(is.na(depth_num), 0, depth_num)
-    depth_sign <- sapply(str_extract(depth, "less|plus"), 
-                         switch, "plus"=1, "less"=-1, 0, USE.NAMES = FALSE)
-    as.numeric(depth_num)*depth_sign
-  })) %>%
-  left_join(read.csv("raw_data/all_DCMs.csv")) %>%
-  mutate(abs_depth=DCM_depth+depth_diff) %>%
-  mutate(abs_depth=ifelse(str_detect(depth, "DCM"), abs_depth, 
-                          as.numeric(str_extract(depth, "^\\d+")))) %>%
-  select(station, cast, depth, abs_depth)
-samp_metadata <- samp_metadata %>%
-  left_join(absolute_depths) %>%
-  left_join(samp_log, by=c("cruise", "station", "cast", "depth", "id")) %>%
-  # If we have an exact depth for a given bottle, use it, otherwise trust
-  # the depth extracted from sample sheets or file names
-  mutate(abs_depth = ifelse(is.na(pressure), abs_depth, pressure)) %>%
-  select(-pressure)
+  mutate(sample_id=str_extract(filename, "(S|MS|L)\\d.*_[A-C](?=.mzML$)"))
 
-extra_metadata <- list.files("../mzMLs", pattern = "Blk|Std|Poo.*mzML", recursive = TRUE) %>%
+extra_filedata <- list.files("../mzMLs", pattern = "Blk|Std|Poo.*mzML", recursive = TRUE) %>%
   data.frame(filename=.) %>%
@@
 
-if(!file.exists("made_data/file_timestamps.csv")){
-  bind_rows(samp_metadata, extra_metadata) %>%
-    mutate(filepath=paste("../mzMLs/", polarity, filename, sep = "/")) %>%
-    pull(filepath) %>%
-    grabMSdata(grab_what = "metadata") %>%
-    `$`("metadata") %>%
-    select(filename, timestamp) %>%
-    write.csv("made_data/file_timestamps.csv", row.names = FALSE)
-}
-file_timestamps <- read_csv("made_data/file_timestamps.csv") %>%
-  fill(timestamp, .direction = "up")
-
-file_metadata <- bind_rows(samp_metadata, extra_metadata) %>%
-  left_join(file_timestamps) %>%
-  arrange(timestamp) %>%
-  # Just reordering, nothing getting dropped
-  select(c("filename", "polarity", "samp_type", "cruise", "station", "cast",
-           "depth", "abs_depth", "rosette", "tripl", "id", "untripl",
-           "date_run", "timestamp"))
-write.csv(file_metadata, file = "made_data/file_metadata.csv", row.names = FALSE)
+complete_file_data <- bind_rows(samp_filedata, extra_filedata) %>%
+  left_join(samp_log)
 ```
 
-## Interpolate metadata for metabolite samples
+## Interpolation of SCOPE data to MS sample locations
 
-```{r summary similarity distance dot plot}
-env_metadata <- read_csv("made_data/env_metadata.csv")
-env_metadata_long <- env_metadata %>%
-  pivot_longer(cols = temp:het_abund) %>%
-  filter(!is.na(value)) %>%
-  select(cruise, station, cast, lat, abs_depth, name, value)
-
-samp_metadata_w_time <- read_csv("made_data/file_metadata.csv") %>%
-  filter(samp_type=="Smp") %>%
-  left_join(env_metadata %>% distinct(cruise, station, cast, lat, local_hour))
-
-cruise_bounds <- samp_metadata_w_time %>%
-  group_by(cruise) %>%
-  summarise(min_bound=min(lat)-0.25, max_bound=max(lat)+0.25)
-
-metadata_gp <- env_metadata_long %>%
-  left_join(cruise_bounds) %>%
-  # filter(time > min_bound & time < max_bound) %>%
-  filter(lat > min_bound & lat < max_bound) %>%
-  filter(abs_depth < 300) %>%
-  ggplot(aes(x=lat, y=abs_depth)) +
-  geom_jitter(aes(color=name), width = 0, height = 0) +
-  geom_point(color="black", data = samp_metadata_w_time) +
-  scale_y_reverse() +
-  facet_wrap(~factor(cruise, levels = c("MT", "MC", "FK"), 
-                     labels = c("Eddy transect", "Eddy center", "Falkor")), 
-             scales = "free_x", nrow=1)
-cruisetrack_gp <- env_metadata %>%
-  arrange(time) %>%
-  group_by(cruise, station) %>%
-  mutate(lat=mean(lat), sla=mean(sla)) %>%
-  left_join(cruise_bounds) %>%
-  filter(lat > min_bound & lat < max_bound) %>%
-  ggplot(aes(x=lat, y=sla, label=station)) +
-  geom_path() +
-  geom_point(aes(color=sla)) +
-  facet_wrap(~cruise, scales = "free_x") +
-  scale_color_gradient2(low = muted("blue"), high = muted("red"), mid = "grey80") +
-  facet_wrap(~factor(cruise, levels = c("MT", "MC", "FK"), 
-                     labels = c("Eddy transect", "Eddy center", "Falkor")), 
-             scales = "free_x", nrow=1)
-
-top_left_indent <- 1
-top_right_indent <- 40
-layout_mat <- rbind(
-  c(rep(NA, top_left_indent), rep(1, 1000-top_left_indent-top_right_indent),
-    rep(NA, top_right_indent)),
-  2, 2, 2
-)
-comb_gp <- grid.arrange(cruisetrack_gp, metadata_gp, layout_matrix=layout_mat)
-ggsave(comb_gp, filename = "env_metadata.png", device = "png", type="cairo",
-       width = 10, height = 8, path = "figures")
-# plotly::ggplotly(metadata_gp)
+```{r Interpolate to sample sites}
+interp_df <- lst(complete_file_data, complete_env_data) %>%
+  bind_rows(.id = "data_type") %>%
+  mutate(station_grouping=case_when(
+    station%in%as.character(74:86)~"falkor_anticyc",
+    station%in%as.character(28:70)~"falkor_cyc",
+    TRUE~station
+  )) %>%
+  split(.$station_grouping) %>%
+  map(function(single_station){
+    single_station %>%
+      mutate(across(c(lat:beam_atten), function(col){
+        if(all(is.na(col)))return(NA)
+        approx(abs_depth, col, xout = abs_depth, rule = 2, ties = mean)$y
+        # pmax(0, predict(loess(col~abs_depth, data = ., span = 1), newdata = .))
+      }))
+  }) %>%
+  bind_rows() %>%
+  filter(data_type=="complete_file_data") %>%
+  select(-data_type, -time) %>%
+  left_join(complete_env_data %>% distinct(station, cast, time), by=c("station", "cast"))
+# Correct wonky PC values (those greater than 4) with a linear model built on beam_atten
+interp_df <- interp_df %>%
+  mutate(pred_PC=interp_df %>% 
+           filter(PC_um<4) %>%
+           lm(formula=PC_um~beam_atten) %>%
+           predict(newdata=interp_df)) %>%
+  mutate(PC_um=ifelse(PC_um>4, pred_PC, PC_um)) %>%
+  select(-pred_PC) %>%
+# Discretize SLA and time
+  mutate(sla_class=case_when(
+    sla_corr < -5 ~ "Cyclone",
+    sla_corr > 5 ~ "Anticyclone",
+    TRUE ~ "Neither"
+  )) %>%
+  mutate(time=as.POSIXct(time, origin="1970-01-01")) %>%
+  mutate(time_class=c("Night", "Morning", "Midday", "Evening")[
+    floor((as.numeric(format(time, "%H"))+3)/6)%%4+1
+  ]) %>%
+  bind_rows(extra_filedata)
+write_csv(interp_df, "filled_file_metadata.csv")
 ```
 
-```{r interpolate env_metadata to metab sample values}
-station_metadata <- read_csv(file = "made_data/station_metadata.csv")
-file_metadata <- read_csv("made_data/file_metadata.csv") %>%
-  left_join(station_metadata)
-env_metadata <- read_csv("made_data/env_metadata.csv")
+## Visualization
 
-temp_env_meta <- env_metadata %>%
-  select(cruise, station, value_depth=abs_depth, temp:euk_abund) %>%
-  mutate(pseudostation=case_when(
-    station%in%as.character(74:86)~"anticyc",
-    station%in%as.character(27:59)~"cyc",
-    TRUE~station
-  ))
-temp_file_meta <- file_metadata %>%
-  filter(samp_type=="Smp") %>%
-  select(station, cruise, filename, samp_depth=abs_depth) %>%
-  mutate(pseudostation=case_when(
-    station%in%as.character(74:100)~"anticyc",
-    station%in%as.character(25:70)~"cyc",
-    TRUE~station
-  ))
-
-interp_out <- temp_env_meta %>%
-  bind_rows(temp_file_meta) %>%
-  group_by(pseudostation) %>%
-  group_split() %>%
-  map_dfr(function(single_station){
-    single_station %>%
-      mutate(samp_depth=ifelse(is.na(samp_depth), value_depth, samp_depth)) %>%
-      mutate(across(temp:euk_abund, function(col){
-        if(sum(!is.na(col))<2){return(NA)}
-        approx(x = value_depth, y=col, xout = samp_depth, rule = 2)$y
-      })) %>%
-      suppressWarnings() %>%
-      filter(!is.na(filename))
-  }) %>%
-  select(-value_depth, -pseudostation, -samp_depth)
-
-filled_file_metadata <- file_metadata %>% left_join(interp_out)
-
-write.csv(filled_file_metadata, "made_data/filled_file_metadata.csv", row.names = FALSE)
+```{r Single-column metadata plot}
+col_name <- "PC_um"
+complete_env_data %>%
+  mutate(station=factor(station, levels=c(
+    1:80, "L1", "L2"
+  ))) %>%
+  filter(!is.na(get(col_name))) %>%
+  filter(abs_depth<190) %>%
+  mutate(abs_depth=abs_depth+3) %>%
+  mutate(station_coord=as.numeric(station)-0.2) %>%
+  ggplot(mapping = aes(x=station, y=abs_depth, color=get(col_name))) +
+  geom_point(size=3, position = position_nudge(x=-0.1)) +
+  geom_point(data = interp_df, color="red", size=4, position = position_nudge(x=0.1)) +
+  geom_point(data = interp_df, size=3, position = position_nudge(x=0.1)) +
+  facet_grid(~cruise, scales = "free_x", space = "free_x") +
+  scale_color_viridis_c() +
+  coord_cartesian(ylim = c(180, 0)) +
+  theme_bw() +
+  theme(legend.position = "none") +
+  ggtitle(str_remove(col_name, "_u.*"))
 ```
 
-```{r load data for plotting}
-filled_file_metadata <- read_csv("made_data/filled_file_metadata.csv")
-filled_MT_metadata <- filled_file_metadata %>% filter(cruise=="MT")
-filled_MC_metadata <- filled_file_metadata %>% filter(cruise=="MC")
-env_metadata <- read_csv("made_data/env_metadata.csv")
+```{r PDF of all the metadata}
+pdf("allmeso_metadata.pdf", onefile = TRUE, width = 6.5, height = 3)
+setdiff(names(complete_env_data), c("abs_depth", "time")) %>%
+  walk(function(col_name){
+    gp_i <-complete_env_data %>%
+      mutate(station=factor(station, levels=c(
+        1:80, "L1", "L2"
+      ))) %>%
+      filter(!is.na(get(col_name))) %>%
+      filter(abs_depth<190) %>%
+      mutate(abs_depth=abs_depth+3) %>%
+      mutate(station_coord=as.numeric(station)-0.2) %>%
+      ggplot(mapping = aes(x=station, y=abs_depth, color=get(col_name))) +
+      geom_point(size=3, position = position_nudge(x=-0.1)) +
+      geom_point(data = interp_df, color="red", size=4, position = position_nudge(x=0.1)) +
+      geom_point(data = interp_df, size=3, position = position_nudge(x=0.1)) +
+      facet_grid(~cruise, scales = "free_x", space = "free_x") +
+      coord_cartesian(ylim = c(180, 0)) +
+      theme_bw() +
+      theme(legend.position = "none") +
+      ggtitle(str_remove(col_name, "_u.*"))
+    print(gp_i)
+  }, .progress=TRUE)
+dev.off()
 ```
 
-```{r plot MT interpolated data as transects}
-meta_nms <- c("no23_um", "po4_um", "chl_ug", "phaeo_ug", 
-             "PC_um", "PN_um", "DON_um", "DOP_um",
-             "dic", "alk", "ph", "oxy_um", 
-              paste0(c("pro", "syn", "het", "euk"), "_abund"))
-
-meta_MT_gps <- lapply(meta_nms, function(chosen_var){
-  env_metadata %>%
-    filter(cruise=="MT") %>%
-    filter(!is.na(get(chosen_var))) %>%
-    ggplot(aes(x=time, y=abs_depth, color=get(chosen_var))) + 
-    geom_point(aes(y=abs_depth-2), size=4.5, color="grey50") +
-    geom_point(aes(y=abs_depth-2), size=4) +
-    geom_point(aes(y=abs_depth), filled_MT_metadata, size=4.5, color="red") +
-    geom_point(aes(y=abs_depth), filled_MT_metadata, size=4) +
-    ylim(250, 0) +
-    scale_color_viridis_c() +
-    ggtitle(chosen_var) +
+```{r Interpolation comparison loess v linear for particular stations}
+dat_vals <- data.frame(station_id="Station 9", value_id="N+N (µm)")
+map(c("L1", "9"), function(station_i){
+  map(c("PC_um", "no23_um"), function(col_i){
+    lst(complete_env_data, complete_file_data) %>%
+      bind_rows(.id = "data_type") %>%
+      filter(station==station_i) %>%
+      mutate(col=get(col_i)) %>%
+      mutate(lin_approx=approx(abs_depth, col, xout = abs_depth, rule = 2, ties = mean)$y) %>%
+      mutate(loess_approx_05=predict(loess(col~abs_depth, data = ., span = 0.5), newdata = .)) %>%
+      mutate(loess_approx_1=predict(loess(col~abs_depth, data = ., span = 1), newdata = .)) %>%
+      filter(abs_depth<190) %>%
+      arrange(abs_depth) %>%
+      mutate(value_id=col_i)
+  }) %>%
+    bind_rows() %>%
+    mutate(station_id=station_i)
+}) %>%
+  bind_rows() %>%
+  mutate(station_id=factor(station_id, levels=c("9", "L1"), labels = c("Station 9", "Station L1"))) %>%
+  mutate(value_id=factor(value_id, levels=c("no23_um", "PC_um"), 
+                         labels = c("N+N (µm)", "PC (µm)"))) %>%
+  ggplot(aes(y=abs_depth)) +
+  geom_vline(xintercept = 0) +
+  geom_point(aes(x=col)) +
+  geom_path(aes(x=lin_approx), color="#00588B") +
+  geom_path(aes(x=loess_approx_05), color="#009B95") +
+  geom_path(aes(x=loess_approx_1), color="#53CC67") +
+  coord_cartesian(ylim=c(200, 0)) +
+  geom_label(x=Inf, y=50, label="Linear", color="#00588B", label.r=unit(0, "in"), hjust=1, data=dat_vals) + 
+  geom_label(x=Inf, y=75, label="Loess (span=0.5)", color="#009B95", label.r=unit(0, "in"), hjust=1, data=dat_vals) + 
+  geom_label(x=Inf, y=100, label="Loess (span=1)", color="#53CC67", label.r=unit(0, "in"), hjust=1, data=dat_vals) + 
   theme_bw() +
-    theme(legend.position = "none", axis.title = element_blank())
-})
-
-comb_gp <- do.call(grid.arrange, c(meta_MT_gps, ncol=4))
-ggsave(comb_gp, filename = "figures/interp_MT_data.png", device = "png", 
-       width = 10, height = 10, units = "in", type="cairo")
+  facet_grid(station_id~value_id, scales="free_x") +
+  labs(x=NULL, y="Depth")
 ```
 
-```{r plot MT interped data as depth profiles}
-meta_MT_depthgps <- lapply(meta_nms, function(chosen_var){
-  env_metadata %>%
-    filter(cruise=="MT") %>%
-    filter(!is.na(get(chosen_var))) %>%
-    ggplot(aes(x=get(chosen_var), y=abs_depth, color=get(chosen_var))) + 
-    geom_path(aes(group=station), color="black", lwd=1) +
-    geom_point(size=4) +
-    geom_point(data = filled_MT_metadata, color="red") +
-    ylim(250, 0) +
-    scale_color_viridis_c() +
-    ggtitle(chosen_var) +
-    theme_bw() +
-    theme(legend.position = "none", axis.title = element_blank())
-})
-comb_gp <- do.call(grid.arrange, c(meta_MT_depthgps, ncol=4))
-ggsave(comb_gp, filename = "figures/interp_MT_depthprofs.png", device = "png", 
-       width = 10, height = 10, units = "in", type="cairo")
-```
-
-```{r plot MC interpolated data as transects}
-meta_nms <- c("no23_um", "po4_um", "chl_ug", "phaeo_ug", 
-             "PC_um", "PN_um", "DOP_um", 
-             paste0(c("pro", "syn", "het", "euk"), "_abund"),
-             "beam_atten", "PAR")
-
-meta_MC_gps <- lapply(meta_nms, function(chosen_var){
-  env_metadata %>%
-    filter(cruise=="MC") %>%
-    filter(!is.na(get(chosen_var))) %>%
-    ggplot(aes(x=time, y=abs_depth, color=get(chosen_var))) + 
-    geom_point(aes(y=abs_depth-2), size=4.5, color="grey50") +
-    geom_point(aes(y=abs_depth-2), size=4) +
-    geom_point(aes(y=abs_depth), filled_MC_metadata, size=4.5, color="red") +
-    geom_point(aes(y=abs_depth), filled_MC_metadata, size=4) +
-    ylim(250, 0) +
-    scale_color_viridis_c() +
-    ggtitle(chosen_var) +
-    theme_bw() +
-    theme(legend.position = "none", axis.title = element_blank())
-})
-
-comb_gp <- do.call(grid.arrange, c(meta_MC_gps, ncol=4))
-ggsave(comb_gp, filename = "figures/interp_MC_data.png", device = "png", 
-       width = 10, height = 10, units = "in", type="cairo")
-```
-
-```{r plot MC interped data as depth profiles}
-meta_MC_depthgps <- lapply(meta_nms, function(chosen_var){
-  env_metadata %>%
-    filter(cruise=="MC") %>%
-    filter(!is.na(get(chosen_var))) %>%
-    ggplot(aes(x=get(chosen_var), y=abs_depth, color=get(chosen_var))) + 
-    geom_path(aes(group=station), color="black", lwd=1) +
-    geom_point(size=4) +
-    geom_point(data = filled_MC_metadata, color="red") +
-    ylim(250, 0) +
-    scale_color_viridis_c() +
-    ggtitle(chosen_var) +
-    theme_bw() +
-    theme(legend.position = "none", axis.title = element_blank())
-})
-comb_gp <- do.call(grid.arrange, c(meta_MC_depthgps, ncol=4))
-ggsave(comb_gp, filename = "figures/interp_MC_depthprofs.png", device = "png", 
-       width = 10, height = 10, units = "in", type="cairo")
-```
-
-Metadata pairs plot and heatmaply things
-
-```{r heatmaply metadata, purl=FALSE}
-filled_MT_metadata <- filled_file_metadata %>%
-  filter(cruise=="FK") %>%
-  filter(samp_type=="Smp") %>%
-  select(where(~!all(is.na(.x))))
-
-heatmap_cor <- filled_MT_metadata %>%
-  select(abs_depth, time, lat, sla, local_hour, #alk, dic, ph
-         ends_with("abund"), #beam_atten,
-         ends_with("um"), ends_with("ug")) %>%
-  data.matrix() %>%
-  cor(use = "pair") %>%
-  abs()
-
-cor(cophenetic(hclust(dist(heatmap_cor))), dist(heatmap_cor))
-pheatmap::pheatmap(
-  mat = heatmap_cor, color = viridisLite::viridis(100),
-  filename = "figures/meta_cor_clust.pdf", width = 7, height = 7
-)
-heatmaply::heatmaply(heatmap_cor)
-```
-
-## Get standards and clean up
-It's also nice to pull down a list of standards that will be found in the data.
-
-```{r grabstans}
-if(!file.exists("raw_data/raw_stans.csv")){
-  raw_stans <- read.csv(paste0(
-    "https://raw.githubusercontent.com/IngallsLabUW/Ingalls_Standards",
-    "/master/Ingalls_Lab_Standards.csv"
-  ))
-  write.csv(raw_stans, file = "raw_data/raw_stans.csv", row.names = FALSE)
-}
-if(!file.exists("raw_data/custom_stans.csv")){
-  custom_stans <- read.csv(paste0(
-    "https://raw.githubusercontent.com/IngallsLabUW/Ingalls_Standards",
-    "/master/data_extra/custom_stans.csv"
-  ))
-  write.csv(custom_stans, file = "raw_data/custom_stans.csv", row.names = FALSE)
-}
-raw_stans <- read.csv("raw_data/raw_stans.csv")
-custom_stans <- read.csv("raw_data/custom_stans.csv")
-clean_stans <- raw_stans %>%
-  bind_rows(custom_stans) %>%
-  filter(Column=="HILIC"|Compound_Type=="Custom") %>%
-  mutate(polarity=ifelse(str_detect(z, "-"), "neg", "pos")) %>%
-  mutate(mz=as.numeric(mz)) %>%
-  mutate(peak_quality=str_extract(Peak_Quality, "Yes|No|Problematic")) %>%
-  select(compound_type=Compound_Type, compound_name=Compound_Name,
-         compound_name_old=Compound_Name_Original,
-         formula=Empirical_Formula, rt=RT_minute, mz=mz, 
-         ionization_form=Ionization_Form,
-         charge=z, kegg_id=KEGG_Code, polarity, conc_um=Concentration_uM,
-         date_added=Date_Added, mix=HILIC_Mix, priority=Priority,
-         peak_quality) %>%
-  mutate(date_added=strptime(x = date_added, format = "%y%m%d"))
-write.csv(clean_stans, file = "made_data/clean_stans.csv", row.names = FALSE)
-```
-
-## Culture data
+## Create nice list of authentic standards
 
 ```{r}
-download.file("https://datadryad.org/stash/downloads/file_stream/2192675", 
-              destfile = "raw_data/Heal_mSystems_Supplement.xlsx", 
-              mode = "wb")
-clean_stans <- read.csv(file = "made_data/clean_stans.csv")
+clean_stans <- "https://raw.githubusercontent.com/IngallsLabUW/" %>%
+  paste0("Ingalls_Standards/master/Ingalls_Lab_Standards.csv") %>%
+  read_csv() %>%
+  filter(Column=="HILIC") %>%
+  mutate(polarity=case_when(
+    z==1 ~ "pos",
+    z==-1 ~ "neg",
+    TRUE~"unknown"
+  )) %>%
+  select(compound=Compound_Name, compound_type=Compound_Type, 
+         formula=Empirical_Formula, stan_mz=mz, stan_rt=RT_minute, 
+         mix=HILIC_Mix, conc=Concentration_uM, Date_Added, 
+         name_old=Compound_Name_Original, polarity) %>%
+  mutate(compound=case_when(
+    compound=="Ethyl aminobenzoate"~"Ethyl aminobenzoate?",
+    TRUE~compound
+  ))
+extra_stans <- "https://raw.githubusercontent.com/IngallsLabUW/" %>%
+  paste0("Ingalls_Standards/master/data_extra/custom_stans.csv") %>%
+  read_csv()
 
-culture_desc <- readxl::read_xlsx("raw_data/Heal_mSystems_Supplement.xlsx",
-                  sheet = "S7_Culturedescriptions_full")
-culture_data <- readxl::read_xlsx("raw_data/Heal_mSystems_Supplement.xlsx",
-                                  sheet = "S5_FullEnviroCulture_PeakAreas")
-taxa_metabs <- culture_data %>%
-  filter(!str_detect(MassFeature_Column, "RP$")) %>%
-  mutate(polarity=tolower(str_extract(MassFeature_Column, "Neg$|Pos$"))) %>%
-  select(Identification, polarity, `7803_B`:Nmar_3) %>%
-  pivot_longer(cols = `7803_B`:Nmar_3) %>%
-  left_join(clean_stans %>% select(compound_name, compound_name_old, polarity), 
-            by=c(Identification="compound_name_old", "polarity")) %>%
-  filter(!is.na(compound_name)) %>%
-  left_join(culture_desc %>% select(name=`Sample ID`, taxa=`Broad taxon`, species=Species),
-            by="name") %>%
-  filter(!is.na(value)) %>%
-  select(compound_name, polarity, broad_taxa=taxa, species, id=name, value)
-
-write.csv(taxa_metabs, file = "made_data/taxa_metabs.csv", row.names = FALSE)
+write_csv(clean_stans, "clean_stans.csv")
 ```

### 2023-05-25 targeted/targeted_control.Rmd
Updating targeted to modern syntax

--- 2023-05-04 targeted/output_control.Rmd
@@
 
-clean_stans <- read.csv("../metadata/made_data/clean_stans.csv")
+clean_stans <- read.csv("../metadata/clean_stans.csv")
 name_conversion <- clean_stans %>%
-  select(new=compound_name, old=compound_name_old) %>%
+  select(new=compound_name, old=name_old) %>%
   distinct()
-file_metadata <- read.csv("../metadata/made_data/file_metadata.csv")
+file_metadata <- read.csv("../metadata/filled_file_metadata.csv")
 ```
@@
 
-```{r MSDIAL autoexport things not yet}
+```{r Mesotransect MSDIAL autoexport things not yet}
 # Not implemented yet, check on thread below for updates
 # http://www.metabolomics-forum.com/index.php?topic=1820.0
-# Instead, open an existing .mtd2 file, click the "Export" tab, select "Alignment result", specify the output directory as "Area_HILICPos_EddyTransect.csv", check a couple of the boxes (Raw Data Area is the only critical one), change the output format to msp(?), and hit the "Export" button at the bottom.
-You then need to open this file in Excel and save as CSV.
-
-# Haven't been able to find the raw data for negative mode so I'm currently just copying this off the Google Drive (G:\Shared drives\Ingalls Lab\Collaborative Projects\MESO-SCOPE\EddyTransect\HILIC_Neg_MSDial)
+# Instead, open an existing .mtd2 file, click the "Export" tab, select "Alignment result", specify the output directory as the current working directory, check the Raw Data matrix (Area) box , keep the output format as msp, and hit the "Export" button at the bottom.
 ```
@@
 
-```{r transectSkylineImport}
-manual_pos <- read_csv("raw_data/Mesotransect/Area_HILICPos_EddyTransect.csv", skip = 4) %>%
+```{r transectMsdialImport}
+
+manual_neg <- read_delim("raw_data/Mesotransect/MSDial_neg_data/Area_0_20235251247.txt", skip = 4, delim = "\t") %>%
+  filter(`Metabolite name`!="Unknown") %>%
+  select(compound_name=`Metabolite name`, starts_with("180821")) %>%
+  pivot_longer(cols=starts_with("180821"), names_to = "filename", values_to="area") %>%
+  mutate(polarity="neg")
+manual_pos <- read_delim("raw_data/Mesotransect/MSDial_pos_data/Area_2_2023525133.txt", skip = 4, delim = "\t") %>%
   filter(`Metabolite name`!="Unknown") %>%
@@
   mutate(polarity="pos")
-manual_neg <- read_csv("raw_data/Mesotransect/Area_HILICNeg_EddyTransect.csv", skip = 4) %>%
-  filter(`Metabolite name`!="Unknown") %>%
-  select(compound_name=`Metabolite name`, starts_with("180821")) %>%
-  pivot_longer(cols=starts_with("180821"), names_to = "filename", values_to="area") %>%
-  mutate(polarity="neg")
 manual_ints <- rbind(manual_pos, manual_neg) %>%
@@
   left_join(name_conversion, by=c("compound_name"="old")) %>% 
-  mutate(compound_name=new) %>% 
+  mutate(compound_name=ifelse(is.na(new), compound_name, new)) %>% 
   select(-new) %>% 
@@
   anti_join(distinct(., compound_name, polarity) %>% filter(duplicated(compound_name)))
-
 ```
@@
 mesotransect_data <- rfs_transect %>% 
-  left_join(BMISed_transect, by="compound_name") %>%
+  left_join(BMISed_transect, by="compound_name", multiple="all") %>%
   left_join(clean_stans) %>%
@@
 write.csv(quant_data, file = "made_data/longdata.csv", row.names = FALSE)
+write_csv(quant_data, file = "made_data/all_peaks.csv")
 ```
@@
   mutate(filename=str_remove(filename, ".mzML")) %>%
-  distinct(filename, samp_type, station, cast, depth, abs_depth, rosette,date_run, timestamp) %>%
-  rename(sample_name=filename, file_timestamp=timestamp) %>%
+  distinct(filename, samp_type, station, cast, depth, abs_depth, rosette,date_run) %>%
+  rename(sample_name=filename) %>%
   writeData(wb = wb, sheet = "File metadata")
@@
 clean_stans %>%
-  select(compound_name, compound_type, formula, rt, mz, kegg_id, polarity, peak_quality) %>%
+  select(compound_name, compound_type, formula, stan_rt, stan_mz, 
+         kegg_id, polarity, peak_quality) %>%
   writeData(wb = wb, sheet = "Compound metadata")
@@
   "date_run = The date on which the HPLC-MS run was started",
-  "file_timestamp = The time at which the sample was actually run",
   "",
@@
   "formula = Molecular formula of the compound. Isotopes are labeled in parentheses",
-  "rt = Estimated retention time of the compound, in minutes",
-  "mz = Mass-to-charge ratio of the molecule at which it's detected in the mass spec",
+  "stan_rt = Estimated retention time of the compound, in minutes",
+  "stan_mz = Mass-to-charge ratio of the molecule at which it's detected in the mass spec",
   "kegg_id = KEGG (Kyoto Encyclopedia of Genes and Genomes) ID: https://www.genome.jp/kegg/",
@@
   mutate(filename=str_remove(filename, ".mzML")) %>%
-  distinct(filename, samp_type, station, cast, depth, abs_depth, rosette, date_run, timestamp) %>%
-  rename(sample_name=filename, file_timestamp=timestamp) %>%
+  distinct(filename, samp_type, station, cast, depth, abs_depth, rosette, date_run) %>%
+  rename(sample_name=filename) %>%
   writeData(wb = wb, sheet = "File metadata")

### 2023-05-25 metadata/metadata_control.Rmd
Switching back to compound_name and adding other columns

@@
   )) %>%
-  select(compound=Compound_Name, compound_type=Compound_Type, 
+  select(compound_name=Compound_Name, compound_type=Compound_Type, 
          formula=Empirical_Formula, stan_mz=mz, stan_rt=RT_minute, 
-         mix=HILIC_Mix, conc=Concentration_uM, Date_Added, 
-         name_old=Compound_Name_Original, polarity) %>%
-  mutate(compound=case_when(
-    compound=="Ethyl aminobenzoate"~"Ethyl aminobenzoate?",
-    TRUE~compound
+         mix=HILIC_Mix, conc_um=Concentration_uM, date_added=Date_Added, 
+         name_old=Compound_Name_Original, kegg_id=KEGG_Code, polarity,
+         peak_quality=Peak_Quality) %>%
+  mutate(compound_name=case_when(
+    compound_name=="Ethyl aminobenzoate"~"Ethyl aminobenzoate?",
+    TRUE~compound_name
   ))

### 2023-05-25 metadata/metadata_control.Rmd
Switching over to MS cruise instead of MC and MT separately

@@
   ]) %>%
-  bind_rows(extra_filedata)
+  bind_rows(extra_filedata) %>%
+  mutate(cruise=ifelse(cruise%in%c("MC", "MT"), "MS", cruise))
 write_csv(interp_df, "filled_file_metadata.csv")

### 2023-05-25 targeted/targeted_control.Rmd
Clearing out made data in prep for rerun

@@
 write.csv(quant_data, file = "made_data/longdata.csv", row.names = FALSE)
-write_csv(quant_data, file = "made_data/all_peaks.csv")
+write_csv(quant_data, file = "all_peaks.csv")
 ```

### 2023-05-26 manuscript/manuscript.Rmd
Manuscript "Run all" now functional
Complete rerun overnight success

--- 2022-10-05 manuscript/manuscript.Rmd
@@
 ---
-title: "Metabolite soup"
+title: "AllMeso manuscript"
 author: "William Kumler"
-date: "10/27/2022"
-output:
-  word_document: default
-  html_document: default
-editor_options:
-  chunk_output_type: inline
+date: "`r Sys.Date()`"
+output: html_document
 ---
 
-Hi Regina
-
-```{r run analysis, eval=FALSE, echo=FALSE}
-# Run untargeted analysis
-knitr::purl("../untargeted/output_control.Rmd", output="../untargeted/output_script.R")
-script_lines <- readLines("../untargeted/output_script.R")
-polarity_set_line_idx <- grep(script_lines, pattern = "polarity <-")[1]
-cruise_set_line_idx <- grep(script_lines, pattern = "cruise <-")[1]
-to_write <- script_lines[-c(polarity_set_line_idx, cruise_set_line_idx)]
-writeLines(to_write, "../untargeted/output_script.R")
-current_dir <- getwd()
+```{r Run analysis}
+if(!file.exists("../metadata/filled_file_metadata.csv")){
+  setwd("../metadata")
+  knitr::purl("metadata_control.Rmd")
+  source("metadata_control.R")
+  file.remove("metadata_control.R")
+  setwd("../manuscript")
+}
+if(!file.exists("../untargeted/all_peaks.csv")){
   setwd("../untargeted")
-for(cruise in c("FK", "MT", "MC")){
+  data_iters <- c("FK_pos", "FK_neg", "MS_pos", "MS_neg")
+  knitr::purl("untargeted_control.Rmd")
+  for(cruise in c("FK", "MS")){
     for(polarity in c("pos", "neg")){
-    print(paste("Working on", cruise, polarity))
-    source("output_script.R", echo = FALSE)
+      source("untargeted_control.R")
     }
   }
-
-# Extract targeted results from targeted analysis
-
+  file.remove("untargeted_control.R")
+  map(data_iters, function(output_dir){
+    comb_peaks <- read_csv(paste0(output_dir, "_output/clean_peak_data.csv")) %>%
+      mutate(temp=output_dir) %>%
+      separate(temp, into = c("cruise", "polarity"), sep = "_") %>%
+      mutate(feature=str_replace(feature, "FT", paste0(toupper(output_dir), "_")))
+  }) %>%
+    bind_rows() %>%
+    write_csv("all_peaks.csv")
+  map(data_iters, function(output_dir){
+    comb_peaks <- read_csv(paste0(output_dir, "_output/clean_feat_data.csv")) %>%
+      mutate(temp=output_dir) %>%
+      separate(temp, into = c("cruise", "polarity"), sep = "_") %>%
+      mutate(feature=str_replace(feature, "FT", paste0(toupper(output_dir), "_")))
+  }) %>%
+    bind_rows() %>%
+    write_csv("all_feats.csv")
+  setwd("../manuscript")
+}
+if(!file.exists("../targeted/all_peaks.csv")){
+  setwd("../targeted")
+  knitr::purl("../targeted/targeted_control.Rmd")
+  source("targeted_control.R")
+  file.remove("targeted_control.R")
+  setwd("../manuscript")
+}
 ```
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE, include = TRUE, fig.width = 6.5)
+knitr::opts_chunk$set(echo = TRUE)
+library(tidyverse)
 
-options(warn = -1)
-options(dplyr.summarise.inform=F)
-options(readr.num_columns = 0)
-options(pillar.sigfig=7)
-suppressPackageStartupMessages(library(tidyverse))
-suppressPackageStartupMessages(library(RaMS))
-library(scales)
-library(ggdendroplot)
-library(ggrepel)
-library(lmPerm)
-
-depth_levels <- c("15m", "25m", "DCMless20m", "DCMless10m", "DCM", "DCMplus10m",
-                  "DCMplus20m", "175m")
-filled_file_metadata <- "../metadata/made_data/filled_file_metadata.csv" %>%
-  read_csv(show_col_types = FALSE) %>%
-  mutate(depth=factor(depth, levels = depth_levels))
-
-factor_levels <- c(
-    "no23_um", "po4_um", "si_um", "PC_um", "PN_um", "DON_um", "DOP_um", "oxy_um",
-    "chl_ug", "phaeo_ug", "alk", "dic", "ph", "theta",
-    paste0(c("pro", "syn", "euk", "het"), "_abund"), "beam_atten",
-    "abs_depth", "time", "lat", "sla", "local_hour"
-  )
-factor_labels <- c(
-    "Nitrate + nitrite", "Phosphate", "Silicate", "Particulate C", "Particulate N",
-    "Dissolved organic N", "Dissolved organic P", "Oxygen", "Chlorophyll a",
-    "Phaeophytin", "Alkalinity", "Dissolved inorganic C", "pH", "Density",
-    paste(c("Pro.", "Syn.", "Picoeuk.", "Het. bact."), "abundance"),
-    "Beam attenuation",
-    "Depth", "Time", "Latitude", "Sea level anomaly", "Local hour"
-  )
-twoline_labels <- c(
-    "Nitrate + nitrite", "Phosphate", "Silicate", "Particulate C", "Particulate N",
-    "Dissolved\norganic N", "Dissolved\norganic P", "Oxygen", "Chlorophyll a",
-    "Phaeophytin", "Alkalinity", "Dissolved\ninorganic C", "pH", "Density",
-    paste(c("Prochlorococcus", "Synechococcus", "Picoeukaryote", 
-            "Heterotrophic bact."), "\nabundance"),
-    "Beam attenuation",
-    "Depth", "Time", "Latitude", "Sea level anomaly", "Local hour"
-  )
-
-clean_stans <- "../metadata/made_data/clean_stans.csv" %>%
-  read_csv(show_col_types = FALSE) %>%
-  select(compound_type, compound_name, formula, polarity)
-
-all_feats <- read_csv("../untargeted/all_feats.csv", show_col_types = FALSE)
-all_peaks <- read_csv("../untargeted/all_peaks.csv", show_col_types = FALSE)
-
-MT_feats <- all_feats %>%
-  filter(cruise=="MT") %>%
-  select(-classes, -CSI_name) %>%
-  rename("calc_formula"=formula) %>%
-  left_join(clean_stans, by=c("compound_name", "polarity"))
-MT_peaks <- MT_feats %>%
-  left_join(all_peaks, by="feature") %>%
-  left_join(filled_file_metadata, by="filename") %>%
-  filter(samp_type=="Smp") %>%
-  select(feature, norm_area, filename)
+filled_file_metadata <- read_csv("../metadata/filled_file_metadata.csv")
+clean_stans <- read_csv("../metadata/clean_stans.csv")
+untarg_feats <- read_csv("../untargeted/all_feats.csv")
+untarg_peaks <- read_csv("../untargeted/all_peaks.csv")
+targ_peaks <- read_csv("../targeted/all_peaks.csv")
 ```
-
-## Abstract
-
-TBD
-
-## Introduction
-
-A major goal of marine microbial metabolomics is a global ocean model of metabolites that resolves at the molecular level, answering the question: "At any given spot and any given time, what is the concentration of a given molecule 1) within a cell and 2) in the surrounding seawater?" With such a model, we could more easily discover marine biomarkers, predict the response of marine communities to perturbations both natural and anthropogenic, and even characterize the production of cloud-forming molecules such as DMS and improve our climate models [and similar stuff, don't love these examples but like the metabolite model justification].
-
-This is a difficult problem because the metabolome is the result of a complex network of interactions between the ocean's biology and its biogeochemistry.
-Gradients of light and nutrient availability create a spectrum of fundamental niches throughout the water column that are sensitive to diel, seasonal, and interannual cycles.
-Additionally, vertical and horizontal mixing due to upwelling/downwelling, storms, and mesoscale features alter the distribution of water masses and the dissolved molecules contained within.
-
-These complex interactions make it difficult to predict exactly which parameters are necessary to include in the final model.
-Some parameters simply don't affect the outcome enough to be worth including because their effect sizes are so small that they have no biological significance.
-Consider an absurd parameter like "distance to nearest butterfly" that somehow obtains enormous statistical significance.
-If, however, it only changes values in the model on the scale of attomoles it is simply not worth calculating in the model.
-[Need better example here].
-We can constrain the number of possible parameters by sketching out a simple ocean model, the simplest of which simply predicts an average value for every molecule at every location and time.
-This is, approximately, where the field of marine microbial metabolomics currently rests.
-For many metabolites, we have an estimate of its concentration at a single location, forcing us to extrapolate that single value vertically and horizontally throughout the ocean in the absence of other information.
-
-With some additional oceanographic knowledge, we can predict how metabolites respond to depth.
-The classic Martin curve and global DOM measurements allow us to predict that the surface ocean will have the highest concentrations of metabolites and that compound abundance will generally decrease with depth.
-This improves our model's accuracy significantly but fails to account for the variable lability of different compounds, with some highly recalcitrant molecules having essentially conservative profiles while other molecules are assimilated extremely rapidly.
-This spectrum of reactivity is expected based on [blah] but has [not yet been explored.] [Mention certain compound classes?
-Some amino acids more labile than others?]
-
-Even more complex models have begun to emerge in recent years informed by additional data and our understanding of ocean physics, chemistry, and biology.
-We have just begun to characterize the response of metabolites to gradients in time and space, covering diel variation and transects across the ocean surface.
-We've also begun to integrate biological information from genomic and transcriptomic studies, allowing us to better resolve the theoretical metabolome as biology responds to shifts in light intensity and nutrient availability.
-The model shown below in black represents one potential graph of these interactions and the role of metabolites as mediators of the interaction between organisms.
-
-```{r fig.height=3}
-DiagrammeR::grViz(diagram = "digraph dot {
-  graph [rankdir = LR]
-
-  tab1 [label = '@@1', pos='0, 1!']
-  tab2 [label = '@@2', pos='0, 0!']
-  tab3 [label = '@@3', pos='2, 0.5!']
-  tab4 [label = '@@4', pos='2,-0.5!']
-  tab5 [label = '@@5', pos='4, 1!']
-  tab6 [label = '@@6', pos='4,-1!']
-  tab7 [label = '@@7', pos='4, 0!']
-  tab8 [label = '@@8', pos='0,-1!', color='red', fontcolor='red', style='dashed']
-
-  tab1 -> tab3;
-  tab2 -> tab3;
-  tab2 -> tab4;
-  tab3 -> tab5;
-  tab4 -> tab5;
-  tab4 -> tab6;
-  tab5 -> tab7;
-  tab6 -> tab7;
-  tab7 -> tab5;
-  tab7 -> tab6;
-  tab8 -> tab4 [color='red', style='dashed']
-}
-
-  [1]: 'Time of day'
-  [2]: 'Depth'    
-  [3]: 'Light'
-  [4]: 'Nutrients'
-  [5]: 'Phototrophs'
-  [6]: 'Heterotrophs'
-  [7]: 'Metabolites'
-  [8]: 'Sea level anomaly'
-  ", engine="neato")
-```
-
-However, the model as is predicts a uniform concentration of nutrients at a given depth.
-This is not observed in the real world due to mixing and the vertical movement of water masses in response to mesoscale eddies.
-These features can be identified by satellites as anomalies in sea surface height and reflect the fluctuations of density surfaces within the water column.
-
-Cyclonic eddies, those circulating counterclockwise in the northern hemisphere and clockwise in the southern, bring the main pycnocline closer to the surface of the ocean and introduce extra nutrients to the sunlit layer of the ocean upon formation.
-Anticyclonic eddies have the opposite effect, depressing isopycnal surfaces and essentially piling up surface water in regions of positive sea level anomaly (SLA).
-These eddies can increase the concentrations of nutrients at the base of the photic zone by more than an order of magnitude and are known to affect phytoplankton species composition [and other species?].
-
-The magnitude of this effect on the metabolome is currently unknown, potentially representing a major source of variation in the marine metabolite model.
-If metabolites are as sensitive to SLA as nutrients are, then the eddy state (shown in red in the figure) will be a critical parameter to include in the marine metabolomic model described above.
-
-Certain metabolites have been studied previously with respect to SLA, most notably chlorophyll [and pheophytin] [for its role in estimating the state of the phytoplankton community].
-Fluorometric measurements of chlorophyll are regularly taken during CTD casts and the deep chlorophyll maximum (DCM) represents an important water column feature for oceanographic sampling [and biological blah].
-The DCM has been noted previously to fluctuate up to [XX] meters in response to fluctuations in sea surface height, indicating that at least one metabolite is sensitive to eddy state in a biologically significant way.
-However, the amplitude of this fluctuation is much less dramatic than that of the nutricline or isopycnal surfaces because it represents a compromise between the constant influx of light from above and the varying nutrient field, with light essentially serving as a stabilizing force for the DCM.
-
-In this paper, we describe the metabolome's variability with respect to depth and sea level anomaly with the goal of constraining these parameters in a [future?
-hypothetical?] global ocean model of marine metabolites.
-The driving questions we answer are listed below:
-
-1.
-Is the metabolome as a whole significantly affected by depth, time of day, and sea level anomaly?
-Which of these parameters causes the most variability in the metabolome and is therefore most important to include in a global ocean model?
-
-2.
-How do individual metabolite depth profiles vary in the upper ocean?
-Can we distinguish labile compounds that are readily degraded from those that are more recalcitrant?
-Are there chemical or biological similarities between metabolites that respond similarly to changes in depth?
-
-3.
-Which metabolites are sensitive to eddy state?
-Do they respond more like nutrients and isopycnal surfaces, or are they stabilized by light availability as the DCM is?
-Are there any metabolites that can be found more abundantly or even exclusively in cyclonic (or anticyclonic) eddies, perhaps serving as biomarkers for the biological community within?
-
-These questions have not been answered previously because measuring compound abundances in the ocean is difficult.
-Marine metabolites typically have concentrations in the nanomolar-picomolar range and the small, polar molecules in particular are difficult to separate from the enormously more abundant salt ions.
-Fortunately, recent advances in mass spectrometry and liquid chromatography have produced methods that allow the measurement of these low-abundance molecules with sufficient accuracy and rapidity for this type of analysis.
-Here, we use a high-resolution accurate-mass instrument coupled to a hydrophilic interaction column to rapidly assay all the small molecules in a given sample.
-
-Additionally, advances in data analysis and computational power have made untargeted mass-spectrometry increasingly viable as a technique for detecting and quantifying molecules that exist beyond the limited suite traditionally pursued in environmental metabolomics.
-Unlike targeted MS, which only analyzes a select few compounds known in advance, untargeted metabolomics is data-driven and can pick out signals produced by compounds that are unknown or unannotated.
-This is a significant advantage in environments such as the ocean where comprehensive databases have yet to be established and provides a way to make mass-spectrometry analysis perfectly reproducible via scripting.
-
-## Methods
-
-### Cruise information
-
-To answer the questions established above, samples were collected from the North Pacific Subtropical Gyre near Station ALOHA during two research cruises that targeted strong mesoscale eddy features.
-The first cruise, KM1709, occurred during June and July 2017 and traversed an area between 28 °N, 157° W and 23 °N, 158.5 °W.
-For this cruise, an eddy dipole off the coast of Hawaii was detected using sea-level anomaly (SLA) satellite data and targeted for both a transect across the cyclonic and anticyclonic poles of the eddy dipole as well as two long-term Lagrangian stations at the center of each eddy.
-The cyclonic pole of the eddy had a maximum negative SLA anomaly of -20 centimeters and the anticyclonic center reached 24 centimeters.
-The transect samples were taken at various times of day as the ship transected the dipole, and the eddy centers were all sampled between 5 and 8 pm.
-(See Supplemental Table XX for cruise data) The second cruise occurred during March and April of 2018 and only sampled from the center of each eddy.
-This Falkor cruise (cruise number FK180310) also traveled between two eddies of different polarities, a cyclonic eddy at 22 °N 156 °W and an anticyclonic eddy at 24.5 °N and 161 °W.
-The cyclonic eddy was sampled first and had a maximum negative SLA difference of -12 cm while the anticyclonic eddy was sampled one day later and had a maximum SLA of 20 cm.
-Samples on this cruise were taken twice a day, once at 6AM and once at 6PM.
-(Supp table XX)
-
-### Environmental data collection
-
-For collaborators to fill in. Need details on every bit of environmental metadata:
-
-  - Bottle 
-    - Chlorophyll/phaeophytin
-    - Nutrients
-    - PC/PN (note outliers)
-    - DON/DOP
-    - Alkalinity/DIC/pH
-    - Theta? This is corrected pressure, right? Not just pressure?
-    - Oxygen
-    - Pro/Syn/Picoeuk/Het
-  - Beam attenuation
-  - PAR?
-  - SLA (pulled from CMAP)
-
-Environmental parameters were interpolated to the locations at which metabolite samples were taken using a [bilinear algorithm].
-
-### Metabolite sample collection
-
-Samples were obtained using the onboard CTD rosette to collect water from 15 meters, the deep chlorophyll maximum, and 175 meters during KM1709 eddy transect, from the DCM ±10 and ±20 meters during the KM1709 eddy center sampling, and from 25 meters and the DCM during FK180310.
-The DCM was determined visually from fluorometer data (XX fluorometer from XX company) during the CTD downcast and Niskin bottles were tripped during the return trip to the surface.
-10 liters of seawater were collected in each Niskin and each depth was sampled in triplicate by firing multiple bottles simultaneously.
-Samples were brought to the surface and decanted into prewashed (3x with DI, 3x with sampled seawater) polycarbonate bottles for filtration.
-Samples were filtered by peristaltic pump onto 142mm 0.2 µm Durapore [Omnipore?] filters [part number] held by polycarbonate filter holders [part number] on a Masterflex [part number] tubing line.
-Pressures were kept as low as possible while still producing a reasonable rate of flow through the filter, approximately 250 mL per minute.
-Samples were then removed from the filter holder using solvent-washed [ethanol?] tweezers and placed into pre-combusted aluminum foil packets that were then flash-frozen in liquid nitrogen before being stored at -80 °C until extraction.
-A methodological blank was also collected by running filtrate through a new filter and then treated identically to the samples.
-
-### Sample extraction
-
-Extraction followed a modified Bligh & Dyer approach as detailed in Boysen et al.
-(2018).
-Briefly, filters were added to PTFE centrifuge tubes with a 1:1 mix of 100 µm and 400 µm silica beads, approximately 2mL -20 °C Optima-grade DCM, and approximately 3mL -20 °C 1:1 methanol/water solution (both also Optima-grade).
-Extraction standards were added after this step, a table of which is available as SI Table XX.
-The samples were then bead-beaten three times, followed by triplicate washes with fresh methanol/water mixture.
-Samples were then dried down under clean N2 gas and kept warm using a Fisher-Scientific Reacti-Therm module.
-Dried aqueous fractions were re-dissolved in 380 µL of Optima-grade water and amended with 20 µL isotope-labeled injection standards, also documented in SI Table XX.
-The dry organic fraction was reconstituted into 380 µL 90:10 DCM:toluene [right?] rather than Boysen et al.’s 1:1 water:acetonitrile.
-Both fractions were also finally syringe-filtered [part number?] to remove any potential clogging material.
-The aqueous fraction was then aliquot into an HPLC vials one for injection on the HILIC column.
-The organic fraction was not run.
-[Were the samples themselves diluted?
-I think yes?] A pooled sample was created by combining 20 µL of each sample into the same HPLC vial, and a 1:1 dilution with water half-strength sample was aliquot from that to assess matrix effects and obscuring variation per Boysen et al.
-(2018).
-Also run alongside the environmental samples were two mixes of authentic standards in water and in an aliquot of the pooled sample at a variety of concentrations, available in SI Table XX, for quality control, annotation, and absolute concentration calculations.
-HPLC vials containing the samples were frozen at -80 °C until thawing shortly before injection.
-
-### HPLC-MS methods
-
-[Verbatim from CH_Ingalls_Lab_LC_Methods from Katherine’s MW submission] A SeQuant ZIC-pHILIC column (5 um particle size, 2.1 mm x 150 mm, from Millipore) was used with 10 mM ammonium carbonate in 85:15 acetonitrile to water (Solvent A) and 10 mM ammonium carbonate in 85:15 water to acetonitrile (Solvent B) at a flow rate of 0.15 mL/min.
-The column was held at 100% A for 2 minutes, ramped to 64% B over 18 minutes, ramped to 100% B over 1 minute, held at 100% B for 5 minutes, and equilibrated at 100% A for 25 minutes (50 minutes total).
-The column was maintained at 30 °C.
-The injection volume was 2 µL for samples and standard mixes.
-When starting a batch, the column was equilibrated at the starting conditions for at least 30 minutes.
-To improve the performance of the HILIC column, we maintained the same injection volume, kept the instrument running water blanks between samples as necessary, and injected standards in a representative matrix (the pooled sample) in addition to standards in water.
-After each batch, the column was flushed with 10 mM ammonium carbonate in 85:15 water to acetonitrile for 20 to 30 minutes.
-
-Metabolomic data was collected on a Thermo Q Exactive HF hybrid Orbitrap (QE) mass spectrometer.
-
-The capillary and auxiliary gas heater temperatures were maintained at 320°C and 100°C, respectively.
-The S-lens RF level was kept at 65 [units], but the H-ESI voltage was set to 3.3 kV instead and sheath gas, auxiliary gas, and sweep gas flow rates were decreased to 16, 3, and 1 [units], respectively.
-Polarity switching was again used but with a scan range of 60 to 900 m/z and a resolution of 60,000.
-Calibration was performed every 3-4 days using the [Thermo Calibration Mix (part number?)] at a target mass of 200 m/z.
-DDA data was collected by fragmenting the pooled samples with [details???
-Same in HILIC and RP?].
-All files were then converted to an open-source mzML format via Proteowizard’s msConvert tool [version number??] with flags and filters specified in Supplemental Table XX [centroiding, pos/neg separation].
-
-### Targeted metabolomic methods
-
-Targeted compounds were manually integrated using Skyline [version number] and MSDIAL [version number] with parameters specified in Supplemental Table XX.
-Skyline was used for the KM1709 and FK180310 eddy center samples, while MSDIAL was used for the KM1709 eddy transect data.
-Raw peak areas were then normalized to their best-matched internal standard per Boysen et al.
-(2018).
-The coefficient of variance (CV) threshold used to choose a suitable match was X.X, as determined by looking at the internal standards normalized to each other and selecting the minimum CV that did not introduce new variation, per the tools in the B-MIS normalization GitHub repository (https://github.com/IngallsLabUW/B-MIS-normalization).
-Normalized peaks were then calibrated via a single-point standard curve as determined uniquely for each compound from the authentic samples used to convert normalized peak area into moles per µL injected and scaled to the amount of seawater filtered to provide a final estimate for each compound of moles per liter of seawater filtered.
-
-### Untargeted metabolomic methods
-
-Untargeted compounds were initially analyzed using the xcms package [version] from the Bioconductor project [version] in the R programming language [version].
-Parameters for import and peakpicking can be found in Supplemental Table XX, and the functions readMSdata and findPeaks were used to produce a list of raw peak areas.
-However, because the default xcms algorithm has several flaws in its signal-to-noise ratio (SNR) estimation, [cite Myers & Du 2017] custom code was written to remove many of these features likely to be noise.
-To estimate SNR more robustly, the raw m/z, retention time, and intensity data was extracted from the mzML files using the R package RaMS [version] and subset for each peak using the mzmin, mzmax, rtmin, and rtmax values produced by xcms.
-Each peak was matched against a Gaussian curve with varying degrees of skew [provide values!] to account for peak tailing effects, and the curve with the smallest residuals was chosen for further analysis.
-The raw data was then subtracted from this ideal curve to estimate the residual variation (noise) present in the peak, and the maximum height of the peak was divided by the standard deviation of these residuals to estimate SNR.
-The advantage of this method is that it can be used on peaks that do not have many data points outside of the peak itself, a feature common to HILIC data.
-This is unlike the existing algorithms in xcms and MzMine2 which require a large window of background points to correctly calculate SNR.
-This new SNR was used to calculate an overall “peak quality” index calculated as
-
-$$ SNR * peak\ correlation^4*max\ intensity$$
-
-where peak correlation is the Pearson’s correlation coefficient calculated between the Gaussian curve and the real data and max intensity is the height of the peak.
-Peaks with a quality score less than 20 were excluded from the remainder of the analysis.
-[Defend the choice of 20].
-This list of peaks was then fed back into the default XCMSnExp object and followed by the default functions for alignment and correspondence, parameters for which can also be found in Supplemental Table XX.
-Finally, a custom version of xcms’ default fillPeaks algorithm was used due to bugs in the source code that prevented the appropriate parameters from being used.
-
-After the curated list of peaks was produced, isotopes and adducts were removed.
-Isotopes and adducts were detected by again accessing the raw data via the R package RaMS [version number] and subsetting between rtmin and rtmax, but instead the new m/z limits were calculated by assuming the detected peak was the isotope or adduct of another and that assumed peak’s mass was used for the subsetting.
-The raw data from that m/z offset was then compared via Pearson correlation to the raw data of the original peak for a raw similarity comparison.
-Additionally, the Pearson correlation of the peak areas for each feature across multiple files was calculated as an orthogonal metric of similarity.
-If the raw similarity correlation was above 0.8 and the correlation between files was above 0.99, the feature was assumed to be an adduct or isotope and removed from downstream analysis as well.
-The list of adducts and isotopes checked for each feature can be found in Supplemental Table XX.
-
-While untargeted data analysis makes it challenging to provide absolute quantification for given molecules, the authentic standards run for the targeted analysis made it possible to annotate a portion of the untargeted peaks as known compounds at a confidence level of 1 in the MSI reporting criteria (https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3772505/).
-This was accomplished by comparing each known metabolite to the data collected from the authentic standards in water data.
-The highest annotation certainty was assigned to those compounds with isotopically labeled internal standards – if a known metabolite had an internal standard, the m/z value for both the compound and its isotopologue were calculated.
-All peaks for each m/z value were collected and those with the closest retention time were assumed to be the isotope and its internal standard, removing them from further annotation consideration.
-The next highest certainty was calculated using the multiple authentic standard mixes we ran alongside the samples themselves.
-To discriminate between isomers of known compounds, these mixes were designed to alternate between isomers in retention time such that the distance between sequential peaks in a given mix was maximized.
-This allows us to calculate a between-mixes ratio that can then be compared to the known mix in which a given compound was detected.
-If the feature appears in the correct mix (defined as having an area in the correct mix one full standard deviation larger than that of the areas detected in the incorrect mix), it is assigned the metabolite name and removed from further annotation steps.
-The next highest certainty was calculated for known metabolites that had multiple features at the same m/z value, typically isomers.
-If the number of features matched the number of authentic standards expected at a given m/z, they were annotated by rank-order retention time matching (i.e.
-the first feature was assumed to be the authentic standard with the lowest retention time, followed by the second feature assumed to be the authentic standard with the second-lowest retention time, etc.).
-Finally, peaks were annotated using retention time similarity to the expected value in the standards list and ultimately annotated as the peak with the largest area if all the above failed.
-[Revisit the above to explain interactions and the whole if/else thing].
-Compounds that were multiply assigned (either one feature assigned to multiple authentic standards or one standard assigned to multiple features) were manually deconvoluted.
-
-Untargeted metabolites were also normalized to a best-matched internal standard (IS) as defined above in the targeted metabolomics section.
-However, the IS were manually integrated using Skyline because even small outliers in integration area significantly affected the downstream analysis after compounds were normalized to them.
-The cutoffs used for untargeted data analysis were [higher/lower] than those in the targeted suite, with an “already good” CV of 0.X and a minimum improvement of 0.X.
-
-Chemical formulae were estimated from each metabolite's m/z ratio and isotope envelope information.
-We implemented this estimation using three separate approaches, using first a set of custom functions, then the R package Rdisop version X.XX.XXX, and finally SIRIUS version X.XX.XXX.
-To create isotope envelopes for detected features, we wrote code very similar to that used to remove the adducts and isotopes above, but rather than assuming the feature was an isotope or adduct we simply looked at the m/z values assuming that the feature was the [M+H]$^+$ adduct in positive mode and the [M-H]$^-$ adduct in negative mode.
-The adducts and isotopes tested can also be found in Supplemental Table XX.
-After calculating the m/z values for each feature, the raw data was again accessed using RaMS and the trapezoidal Riemann sum was calculated between the minimum retention time and the maximum retention time for the feature across each file and adduct or isotope.
-Raw data with a Pearson correlation above 0.75 and a correlation between files of 0.9 or greater was assumed to be an isotope of the base peak.
-
-For the custom calculations, these values were used to estimate the ratio of heavy to light abundance.
-A Type I linear model was fit to each feature’s isotope ratios and the slope of the line was then used to estimate the number of atoms contained in each formula using the calculations provided in Supplemental Methods XX.
-This custom method provided a way to constrain the results produced by Rdisop and SIRIUS.
-For those functions, the file-specific isotope envelope had to be collapsed into a feature-based format.
-The average monoisotopic area was used for the base peak and multiplied by the slope of the line from the linear model to obtain an average isotope ratio across the multiple files.
-Next, the Rdisop’s “decomposeIsotopes” function was used with a ppm error of 5 if the m/z ratio was above 200 or a fixed tolerance of 0.001 Da otherwise.
-The potential formula candidates output from this function were then filtered by presence/absence in Pubchem while relaxing the formula to contain one more or fewer hydrogen atoms than initially estimated to account for [M+H]+ and [M-H]- adducts.
-Finally, SIRIUS command-line functions were used to provide a third orthogonal estimate of molecular formula.
-The “decomp” function was passed the m/z value for each peak and constrained with a ppm error of 5 and an element list of CHNOPS.
-Only formulae that appeared in both Rdisop and SIRIUS and also had an elemental breakdown matching the custom calculations were kept for further analysis.
-
-The above untargeted process was applied separately to each combination of column type, polarity, and sample run.
-This resulted in six unique peak lists that required correspondence across analyses.
-The six HILIC runs were correspondedby selecting the feature with the largest area and features from the other runs near the largest feature in *m/z* and retention time were selected, with the features from each other run closest in area and retention time according to log$_2$ retention time and log$_{10}$ average area identified as most likely candidates and grouped.
-Those features were then removed from consideration and the correspondence was repeated for the next largest peak until no features remained ungrouped.
-Features that had no corresponding feature in other runs became their own group.
-Finally, grouped features that contained a targeted compound were used to identify precision and accuracy of the grouping.
-A manual retention time correction of 60 seconds was applied to the features detected in the FK180310 cruise after peak visualization showed a consistent offset of approximately that distance in the data.
-
-### Statistics
-
-Fourteen environmental variables were measured during the eddy transect alongside the metabolite samples ([Figure|Table XX]).
-We confirmed that these parameters were able to uniquely characterize differences due to depth in the eddy transect with both univariate and multivariate methods.
-First, we tested for differences due to sampling depth using Welch's ANOVA (@welch1951), chosen because variances could not be assumed equal for most measurements and corrected for multiple comparisons using Holm's method (@holm1979).
-Next, we distributed the samples into a two-dimensional space using non-metric multidimensional scaling (NMDS) as implemented in R's vegan package with the metaMDS function.
-Two dimensions were chosen because one dimension would not provide sufficient explanatory power and the stress value was below 0.1 when two dimensions were used, rendering a third was unnecessary.
-For this NMDS analysis, each environmental variable was z-score scaled using base R's scale function to have a mean of zero and unit standard deviation, thereby making each parameter unit-invariant and allowing us to disable autotransformation.
-Finally, the environmental vectors were fit onto this ordination using vegan's envfit function, allowing us to determine the direction in the ordination space in which each vector changes most quickly and estimate the significance of each fitted factor using permutation.
-
-To assess the significance of depth, eddy polarity, and diel variability on the metabolome as a whole, we performed a Permutational Multivariate Analysis of Variance (PERMANOVA) using R's vegan package (version XX).
-For this analysis, the QC'd and BMIS'd peak areas were ranked using R's rank function with default ties.method argument of "average" to place all metabolites on the same scale.
-Discrete SLA bins were created by categorizing each station as cyclonic if the corrected SLA was less than -5cm, anticyclonic if greater than 5cm, or neither if between these endpoints (@barone).
-Discrete time categories were created by binning sample times into morning, midday, evening, and night if the sample was collected between 3:30am and 9:30am, 9:30am and 3:30pm, 3:30pm and 9:30pm, or 9:30pm and 3:30am, respectively.
-These bins were chosen [arbitrarily a priori, see if changes happen when using the MAHALO bins instead per Josh's recommendation].
-Depth was already discretized by the depth at which the sample was taken: 15 meters, the DCM, and 175 meters.
-The adonis2 function was used with a model formula consisting of a ranked metabolite data matrix as described above as the LHS and a RHS of either no interactions, pairwise interactions, or all interactions.
-Significance was assessed marginally by setting the "by" argument to "margin" to avoid the potentially confounding effect of assessing terms sequentially, and pairwise distances were calculated using the Manhattan distance metric.
-Nonmetric multidimensional scaling (NMDS) was performed with the metaMDS function, also in the vegan package, again using the Manhattan distance metric.
-Three dimensions were chosen because two-dimensional solutions consistently exceeded a stress of 0.15.
-Autotransform was disabled because the data have already been normalized via ranking.
-The Manhattan distance metric was chosen for its more intuitive performance in higher dimensions, but the results did not change [very much?
-significantly?] when the Euclidean distance metric was used instead [double check this].
-
-To characterize the response of metabolites to depth, we ran a Kruskal-Wallis on each compound/parameter followed by Dunn's post-hoc test for pairwise comparisons as implemented by the rstatix's package dunn_test function with no p-value adjustment if the Kruskal-Wallis was significant at an alpha level of 0.05 after adjusting p-values to control the false discovery rate using R's p.adjust function.
-The Kruskal-Wallis test was chosen instead of the traditional parametric ANOVA because compounds and parameters could not be assumed to have equal variance or normal error distributions, and Dunn's test was chosen as the post-hoc for the same reason.
-The Dunn test p-values were not corrected for multiple comparisons because [blah].
-Compounds were then classified into one of twelve categories by the sign, significance, and magnitude of the post-hoc test results, as well as a null category (no detectable response with depth) and a fourteenth category for compounds with a significant Kruskal-Wallis but for which Dunn's test was unable to significantly discriminate between depth pairs.
-These categories [emerge from permuting boxplot thingys and there's twelve for REASONS and it's not arbitrary] [I need to describe how I assigned the categories in more detail but haven't figured out how to do that yet - maybe a table?].
-
-To test the hypothesis that compounds most abundant at the surface would have the highest C:N ratios and compounds most abundant deeper in the ocean would have the lowest C:N ratios, we [blah blah blah, used the vertical axis of the clockplot as a proxy for "most abundant at surface" vs "most abundant at depth" and ran a regression of C:N ratio vs y-axis coordinate on 1) only known metabs, 2) all metabs with estimated molecular formulae].
-
-To measure the effect of SLA at each sampled depth, we ran a simple linear regression (Model I) on each metabolite at each depth individually, using SLA as the predictor variable and rank-normalized metabolite abundance as the dependent.
-Separating the data set in this way allowed us to estimate the impact of SLA on metabolite abundance [without the confounding effect of variable responses due to depth].
-Compounds were then categorized into a "positive trend with SLA" category if they had a significantly positive slope, a "negative trend with SLA" if they were significantly negative, or a null category of no significant trend if SLA did not have a detectable effect on metabolite abundance.
-
-We assessed the relative importance of sea level anomaly (SLA) on the metabolome by using the separate set of samples collected from the center of the eddies on the KM1709 cruise.
-These depths were chosen to represent a large, but not enormous, amount of variability due to niche partitioning in the water column.
-We then compared metabolite abundances between the two eddies using a simple Student's t-test and grouped all five unique depths together.
-
-We measured the similarity of the metabolome by calculating a distance matrix between each metabolite that appeared in all three data sets after normalizing each to the maximum value observed for that metabolite within each data set.
-We then calculated Pearson's correlation coefficient between these three distance matrices, allowing us to see how well each metabolome corresponded to each other.
-To compare the robustness of individual compounds, we ran permutational tests for differences due to SLA on all three data sets; a t-test for the Falkor and eddy center data and a regression for the eddy transect data.
-We then identified compounds that responded significantly to SLA at an alpha level of 0.05 across different data sets.
-
-All statistics were performed in the R programming environment, [version].
-
-## Results
-
-### Environmental data
-
-Measured values for all fourteen environmental variables are shown below in [Table|Figure] XX.
-Much of the variance observed in these values is explained by the collection of samples at three discrete depths: 15 meters below the surface, the deep chlorophyll maximum, and 175 meters.
-DCM depth was not a fixed distance from the surface but instead fluctuated with sea level anomaly in response to nutrient availability such that the DCM samples varied in their absolute depth from 100 to 133 meters.
-
-```{r metadata boxplots & table}
-distinct_MT_metadata <- filled_file_metadata %>%
-  filter(cruise=="MT") %>%
-  mutate(basename=str_remove(untripl, "180821_Smp_")) %>%
-  select(basename, depth, sla:last_col()) %>%
-  group_by(basename) %>%
-  summarise(depth=unique(depth), across(-depth, mean)) %>%
-  select(-PAR, -PAR_perc, -ph, -dic, -alk, -oxy_um, -theta) %>%
-  filter(!is.na(basename))
-
-# If we decide to use a figure (boxplots of values by depth)
-distinct_MT_metadata %>%
-  mutate(depth=factor(depth, levels = c("175m", "DCM", "15m"))) %>%
-  select(basename, depth, sla, ends_with("abund"),
-         ends_with("um"), ends_with("ug"), beam_atten) %>%
-  pivot_longer(-c(basename, depth, sla)) %>%
-  mutate(name=factor(name, levels = factor_levels, labels = twoline_labels)) %>%
-  filter(!is.na(value)) %>%
-  ggplot() +
-  geom_boxplot(aes(x=value, y=depth)) +
-  facet_wrap(~name, scales="free_x", nrow=3) +
-  theme(axis.title = element_blank())
-
-# If we decide to use a table
-distinct_MT_metadata %>%
-  select(basename, depth, ends_with("abund"),
-         ends_with("um"), ends_with("ug"), beam_atten) %>%
-  # Calculate summary statistics for each environmental parameter
-  pivot_longer(-c(basename, depth)) %>%
-  group_by(depth, name) %>%
-  summarize(print_col=paste(
-    signif(mean(value, na.rm=TRUE), digits = 3), "±", 
-    signif(sd(value, na.rm=TRUE), digits = 1)
-  ), .groups = "drop") %>%
-  # Pivot into nice format
-  pivot_wider(names_from="depth", values_from = print_col) %>%
-  mutate(name=factor(name, levels = factor_levels, labels = factor_labels)) %>%
-  arrange(name)
-```
-
-```{r welch ANOVA MT}
-welch_aov <- distinct_MT_metadata %>%
-  select(basename, depth, sla, ends_with("abund"),
-         ends_with("um"), ends_with("ug"), beam_atten) %>%
-  pivot_longer(-c(basename, depth, sla)) %>%
-  # Perform the ANOVA
-  nest(data=-name) %>%
-  mutate(aovtest=map(data, ~broom::tidy(oneway.test(.x$value~.x$depth)))) %>%
-  suppressMessages() %>%
-  unnest(aovtest) %>%
-  # Format p-values nicely
-  mutate(p.value=p.adjust(p.value, method = "holm"))
-```
-
-All fourteen environmental parameters differed significantly with depth with beam attenuation, heterotrophic bacteria abundance, chlorophyll a concentration, and phaeophytin concentration differing most with depth and nitrate concentration differing least with respect to depth.
-[Note, however, that this assumes no interaction with sea level anomaly or time [because I couldn't figure out how to do this] [and it's possible that some parameters differ only with respect to a confounding variable].]
-N + N, phosphate, and DIC increased more or less linearly with depth while several environmental parameters (pH, density) and biological abundances (*Prochlorococcus*, *Synechococcus*, and heterotrophic bacteria) had the opposite trend, following DON and DOP as they linearly decreased.
-As expected, chlorophyll had a maximum at the DCM along with phaeophytin and O_2_.
-Both particulate carbon (PC) and particulate nitrogen (PN) had high values at the surface and through the DCM, dropping off at the 175m depth horizon, agreeing well with particle load measured via beam attenuation.
-Picoeukaryote abundance notably had its own profile somewhere between a DCM max and a decrease with depth, with highest abundances at the DCM but values at the surface also relatively high.
-
-```{r}
-# If we decide to include a heatmap showing correlations among the metadata
-heatmap_cor <- filled_file_metadata %>%
-  filter(cruise=="MT") %>%
-  filter(samp_type=="Smp") %>%
-  select(abs_depth, time, lat, sla, local_hour, 
-         ends_with("abund"), ends_with("um"), ends_with("ug")) %>%
-  data.matrix() %>%
-  cor(use = "pair")
-
-coph_coef <- cor(cophenetic(hclust(dist(heatmap_cor))), dist(heatmap_cor))
-# pheatmap::pheatmap(
-#   mat = heatmap_cor, color = viridisLite::viridis(100),border_color = "grey80"
-# )
-
-heatmap_labels <- c(factor_labels)
-heatmap_levels <- c(factor_levels)
-names(heatmap_labels) <- heatmap_levels
-
-hc <- hclust(dist(t(heatmap_cor)))
-hclust_labels <- heatmap_labels[hc$labels[hc$order]]
-
-ggcor_df <- heatmap_cor %>%
-  as.data.frame() %>%
-  rownames_to_column("x") %>%
-  pivot_longer(!x, names_to = "y") %>%
-  # mutate(x=factor(x, levels = heatmap_levels, labels = heatmap_labels)) %>%
-  # mutate(y=factor(y, levels = heatmap_levels, labels = heatmap_labels)) %>%
-  mutate(x=factor(x, levels = names(hclust_labels), labels = hclust_labels)) %>%
-  mutate(y=factor(y, levels = names(hclust_labels), labels = hclust_labels)) %>%
-  arrange(x, y) %>%
-  filter(as.numeric(x)<=as.numeric(y))
-ggcor_labs <- ggcor_df %>% distinct(x) %>% mutate(y=row_number()-0.8)
-
-ggplot() +
-  geom_tile(aes(x=as.numeric(x), y=as.numeric(y), fill=value), data = ggcor_df) +
-  geom_text(aes(x=as.numeric(x), y=as.numeric(y), label=x), data = ggcor_labs,
-            angle=90, hjust=1, size=12*0.8/ggplot2::.pt) +
-  geom_text(aes(x=0, y=as.numeric(y)+1, label=x), data = ggcor_labs,
-            hjust=1, size=12*0.8/ggplot2::.pt) +
-  geom_dendro(hc, ylim=c(19.5, 22), lwd=1) +
-  annotate("text", x=Inf, y=Inf, label=paste("Coph coef:", round(coph_coef, 3)), 
-           hjust=1, vjust=1) +
-  scale_fill_gradient2(limits=c(-1, 1), name="Pearson's\ncorrelation\ncoefficient") +
-  scale_x_discrete(position = "top") +
-  theme_minimal() +
-  theme(axis.title = element_blank(),
-        legend.position = c(0.9, 0.3),
-        axis.text = element_blank(),
-        panel.grid = element_blank()) +
-  coord_fixed() +
-  scale_y_continuous(limits = c(-3, 22)) +
-  scale_x_continuous(limits = c(-7, 21))
-```
-
-Two-dimensional non-metric multidimensional scaling (NMDS) of the environmental parameters distinctly separated the samples by depth, with samples taken from 15 meters, the DCM, and 175 meters falling into separate clusters.
-
-```{r 2D NMDS plot}
-nmds_metadata <- distinct_MT_metadata
-nmds_output <- nmds_metadata %>%
-  column_to_rownames("basename") %>%
-  select(-depth) %>%
-  data.matrix() %>%
-  scale() %>%
-  # rank() %>%
-  vegan::metaMDS(distance = "manhattan", k = 2, trace = 0,
-                 autotransform = FALSE, na.rm=TRUE)
-nmds_df <- nmds_output$points %>%
-  as.data.frame() %>%
-  rownames_to_column("basename") %>%
-  left_join(nmds_metadata, by="basename")
-envfit_df <- nmds_metadata %>% 
-  select(-depth) %>%
-  column_to_rownames("basename") %>%
-  vegan::envfit(ord=nmds_output, permutations=999, na.rm=TRUE) %>% 
-  `$`("vectors") %>% `[`(1:4) %>%
-  do.call(what = cbind) %>%
-  as.data.frame() %>%
-  set_names("MDS1", "MDS2", "r", "perm", "pval") %>%
-  mutate(across(starts_with("MDS"), `*`, 20)) %>%
-  select(-perm) %>%
-  rownames_to_column("var") %>%
-  mutate(var=factor(var, levels = factor_levels, labels = factor_labels)) %>%
-  mutate(env_label=ifelse(pval<0.05, as.character(var), paste(var, "(ns)")))
-ggplot(nmds_df, aes(x=MDS1, y=MDS2)) +
-  geom_segment(aes(xend=0, yend=0), envfit_df, arrow = 
-                 arrow(ends = "first", length = unit(0.02, "npc"), 
-                       type = "closed")) + 
-  geom_point(aes(color=depth)) +
-  ggrepel::geom_label_repel(aes(label=env_label), envfit_df, 
-                            min.segment.length = 0, nudge_y = 0.1) + 
-  theme_minimal() +
-  annotate("label", x = Inf, y=Inf, hjust=1, vjust=1,
-           label=paste("2D NMDS Stress:", round(nmds_output$stress, 3)), 
-           fill="black", color="white") +
-  labs(color="Depth") +
-  coord_fixed()
-```
-
-Fitting the environmental vectors onto this ordination illustrated that many parameters had significant collinearity [as expected from the correlation matrix and boxplots above but can't say that if I don't keep them all], with nitrate, phosphate, and silica all correlating with depth and DON and DOP doing so inversely.
-*Synechococcus* abundance correlated neatly with DON and DOP, while abundances of *Prochlorococcus* and heterotrophic bacteria remained high through the DCM alongside particulate nitrogen, particulate carbon, and beam attenuation.
-Chlorophyll and pheophytin concentrations were highest at the DCM, as expected, and corresponded to the highest level of picoeukaryotes as well.
-Neither sea level anomaly nor local hour were found to change significantly across the ordination (permutational p = 0.34 and 0.69, respectively).
-
-### Overall metabolite response
-
-```{r 3D NMDS plot}
-nmds_output <- MT_peaks %>%
-  left_join(MT_feats, by="feature") %>%
-  filter(polarity=="pos") %>%
-  group_by(feature) %>%
-  mutate(value=rank(norm_area)) %>%
-  # mutate(value=norm_area/max(norm_area)) %>%
-  # mutate(value=scale(norm_area)[,1]) %>%
-  ungroup() %>%
-  select(feature, value, filename) %>%
-  pivot_wider(names_from = feature) %>%
-  column_to_rownames("filename") %>%
-  vegan::metaMDS(distance = "manhattan", k = 3, trace = 0, autotransform = FALSE,
-                 wascores = FALSE, noshare = FALSE)
-nmds_df <- nmds_output$points %>%
-  as.data.frame() %>%
-  rownames_to_column("filename") %>%
-  left_join(filled_file_metadata, by="filename") %>%
-  select(filename, starts_with("MDS"), depth, sla, local_hour) %>%
-  mutate(basename=str_extract(filename, "MS\\d+C\\d+.*(?=_[A-C])")) %>%
-  mutate(disc_hour=case_when(
-    between(local_hour, 4, 9)~"morning",
-    between(local_hour, 10, 15)~"midday",
-    between(local_hour, 16, 21)~"evening",
-    TRUE~"night"
-  ))
-nmds_labels <- nmds_df %>%
-  group_by(depth) %>%
-  summarise(MDS1=mean(MDS1), MDS2=mean(MDS2), MDS3=mean(MDS3))
-mds12 <- ggplot(nmds_df, aes(x=MDS1, y=MDS2)) +
-  geom_polygon(aes(fill=depth, group=basename), alpha=0.5) +
-  geom_polygon(aes(color=depth, group=basename), fill=NA) +
-  geom_label_repel(data = nmds_labels, aes(color=depth, label=depth), 
-                   size=6) +
-  theme_minimal() +
-  theme(legend.position="none") +
-  coord_fixed() +
-  ggtitle("A)")
-mds13 <- ggplot(nmds_df, aes(x=MDS1, y=MDS3)) +
-  geom_polygon(aes(fill=depth, group=basename), alpha=0.5) +
-  geom_polygon(aes(color=depth, group=basename), fill=NA) +
-  geom_label_repel(data = nmds_labels, aes(color=depth, label=depth), 
-                   size=6) +
-  theme_minimal() +
-  theme(legend.position="none") +
-  coord_fixed() +
-  ggtitle("B)")
-mds13_sla <- ggplot(nmds_df, aes(x=MDS1, y=MDS3)) +
-  geom_polygon(aes(fill=sla, group=basename), alpha=0.5) +
-  geom_polygon(aes(color=sla, group=basename), fill=NA) +
-  theme_minimal() +
-  coord_fixed() +
-  scale_fill_gradient2(high = muted("red"), low = muted("blue"), midpoint = 0.05) +
-  scale_color_gradient2(high = muted("red"), low = muted("blue"), midpoint = 0.05) +
-  ggtitle("C)")
-mds13_time <- ggplot(nmds_df, aes(x=MDS1, y=MDS3)) +
-  geom_polygon(aes(fill=disc_hour, group=basename), alpha=0.5) +
-  geom_polygon(aes(color=disc_hour, group=basename), fill=NA) +
-  theme_minimal() +
-  coord_fixed() +
-  ggtitle("D)")
-gridExtra::grid.arrange(mds12, mds13, mds13_sla, mds13_time,
-                        layout_matrix=matrix(c(1,2,3,4), ncol=2))
-```
-
-*Figure XX: Non-metric multidimensional scaling plots of the metabolome.
-Axes 1 and 2 are shown in the subplot a) and axes 1 and 3 are shown for b) through d).
-Triangles connect biological replicates (same environmental context but collected in separate Niskin bottles) and are colored by sampling depth in a) and b), sea level anomaly (SLA) in c), and time of day in d).*
-
-We also found similar clustering by depth in the metabolome, with each depth having a visually distinct combination of metabolites (Figure XX).
-In the NMDS plot shown, the overall stress of the system is approximately 0.114 and axes 1 and 3 neatly separate samples by depth while axis 2 represents a source of variation not clearly associated any environmental variable.
-Axis 1 distinguishes the 175 meter samples from the shallower depths, with a distinct cluster of deep samples on one side of the figure and significant overlap between the DCM and 15 meter samples on the other.
-Axis 3 distinguishes each depth but does so with some overlap between categories, placing the 15 meter samples all to one side, the 175 meter samples near zero, and the DCM samples opposite the 15 meter samples.
-When colored by sea level anomaly (Figure XX.C), the deep samples appear to separate within the cluster along axis 1, with samples taken from the anticyclonic eddy aligning more closely with the shallower samples than the samples from the cyclonic eddy.
-
-However, the distribution of samples across NMDS axes was not especially robust under different normalization techniques.
-When each metabolite was scaled to a percentage of the maximum peak area for that metabolite or when scaled using z-score normalization it was axis 1 that contained the within-triplicate variation while axis 2 separated the 175m samples from the surface and DCM and axis 3 contained similar DCM-175m-15m sequences.
-
-```{r}
-rank_mat <- MT_peaks %>%
-  left_join(MT_feats) %>%
-  filter(polarity=="pos") %>%
-  group_by(feature) %>%
-  mutate(value=rank(norm_area)) %>%
-  # mutate(value=value/max(value)) %>%
-  # mutate(value=scale(value)[,1]) %>%
-  ungroup() %>%
-  select(filename, feature, value) %>%
-  pivot_wider(names_from = feature) %>%
-  column_to_rownames("filename")
-permadata <- filled_file_metadata %>%
-  filter(cruise=="MT") %>%
-  filter(samp_type=="Smp") %>%
-  mutate(disc_sla=case_when(
-    sla < -0.05~"cyclone",
-    sla > 0.05~"anticyclone",
-    TRUE~"neither"
-  )) %>%
-  mutate(disc_hour=case_when(
-    between(local_hour, 4, 9)~"morning",
-    between(local_hour, 10, 15)~"midday",
-    between(local_hour, 16, 21)~"evening",
-    TRUE~"night"
-  ))
-permanout_all_int <- vegan::adonis2(
-  rank_mat~depth*disc_sla*disc_hour, data = permadata, by = "margin", 
-  permutations = 999, method = "manhattan"
-)
-permanout_depthsla_int <- vegan::adonis2(
-  rank_mat~depth*disc_sla+disc_hour, data = permadata, by = "margin", 
-  permutations = 999, method = "manhattan"
-)
-permanout_depthtime_int <- vegan::adonis2(
-  rank_mat~depth*disc_hour+disc_sla, data = permadata, by = "margin", 
-  permutations = 999, method = "manhattan"
-)
-permanout_timesla_int <- vegan::adonis2(
-  rank_mat~depth+disc_sla*disc_hour, data = permadata, by = "margin", 
-  permutations = 999, method = "manhattan"
-)
-permanout_no_int <- vegan::adonis2(
-  rank_mat~depth+disc_sla+disc_hour, data = permadata, by = "margin", 
-  permutations = 9999, method = "manhattan"
-)
-```
-
-The less-subjective PERMANOVA confirmed that depth was a major factor in controlling the metabolome, explaining ~32% of the variation with a p-value <0.001.
-Time of day and SLA were both found to be significant factors as well but explained only 4-5% and 3-4% of the variance in the metabolome, respectively, with p-values in the 0.001-0.005 range.
-Testing for interactions revealed a significant three-way interaction with p-values in the 0.04-0.05 range explaining ~5% of the variance and a marginally significant two-way interaction between depth and SLA (p ~ 0.09), while time of day didn't interact significantly with either depth or SLA.
-
-### Metabolite response to depth
-
-```{r posthoc_classes and posthoc_table}
-kw_diff_feats <- MT_peaks %>%
-  left_join(filled_file_metadata, by="filename") %>% 
-  select(feature, depth, value=norm_area) %>%
-  nest(data=-feature) %>%
-  mutate(tidyaov=map(data, ~ broom::tidy(kruskal.test(.x$value~.x$depth)))) %>%
-  unnest(tidyaov) %>%
-  mutate(pval=p.adjust(p.value, method = "fdr")) %>%
-  filter(pval < 0.05) %>%
-  select(feature, data) %>%
-  unnest(data)
-nondiff_feats <- MT_peaks %>%
-  filter(!feature%in%kw_diff_feats$feature) %>%
-  group_by(feature) %>%
-  summarize(resp_class="All equal")
-
-resp_class_levels <- c("All equal", 
-    "15m > DCM = 175m", "15m > DCM > 175m", "175m < DCM = 15m",
-    "DCM max 15m mid", "DCM > 15m = 175m", "DCM max 175m mid", 
-    "15m < DCM = 175m", "15m < DCM < 175m", "175m > DCM = 15m", 
-    "DCM min 15m mid", "DCM < 15m = 175m", "DCM min 175m mid", 
-    "Hard to tell")
-
-posthoc_classes <- kw_diff_feats %>%
-  group_by(feature) %>%
-  rstatix::dunn_test(value~depth, p.adjust.method = "none") %>%
-  mutate(contrast=paste(group2, group1, sep = "_")) %>%
-  select(feature, contrast, est=statistic, pval=p.adj) %>%
-  pivot_wider(names_from = contrast, values_from = c(pval, est)) %>%
-  mutate(resp_class=case_when(
-    pval_DCM_15m<0.05 & pval_175m_15m<0.05 & pval_175m_DCM<0.05 & est_DCM_15m<0 & est_175m_15m<0 & est_175m_DCM<0 ~ "15m > DCM > 175m",
-    pval_DCM_15m<0.05 & pval_175m_15m<0.05 & pval_175m_DCM<0.05 & est_DCM_15m>0 & est_175m_15m<0 & est_175m_DCM<0 ~ "DCM max 15m mid",
-    pval_DCM_15m<0.05 & pval_175m_15m<0.05 & pval_175m_DCM<0.05 & est_DCM_15m<0 & est_175m_15m>0 & est_175m_DCM>0 ~ "DCM min 15m mid",
-    pval_DCM_15m<0.05 & pval_175m_15m<0.05 & pval_175m_DCM<0.05 & est_DCM_15m>0 & est_175m_15m>0 & est_175m_DCM>0 ~ "15m < DCM < 175m",
-    pval_DCM_15m<0.05 & pval_175m_15m<0.05 & pval_175m_DCM<0.05 & est_DCM_15m>0 & est_175m_15m>0 & est_175m_DCM<0 ~ "DCM max 175m mid",
-    pval_DCM_15m<0.05 & pval_175m_15m<0.05 & pval_175m_DCM<0.05 & est_DCM_15m<0 & est_175m_15m<0 & est_175m_DCM>0 ~ "DCM min 175m mid",
-    pval_DCM_15m<0.05 & pval_175m_15m<0.05 & pval_175m_DCM>0.05 & est_DCM_15m<0 & est_175m_15m<0 ~ "15m > DCM = 175m",
-    pval_DCM_15m<0.05 & pval_175m_15m<0.05 & pval_175m_DCM>0.05 & est_DCM_15m>0 & est_175m_15m>0 ~ "15m < DCM = 175m",
-    pval_DCM_15m>0.05 & pval_175m_15m<0.05 & pval_175m_DCM<0.05 & est_175m_15m<0 & est_175m_DCM<0 ~ "175m < DCM = 15m",
-    pval_DCM_15m>0.05 & pval_175m_15m<0.05 & pval_175m_DCM<0.05 & est_175m_15m>0 & est_175m_DCM>0 ~ "175m > DCM = 15m",
-    pval_DCM_15m<0.05 & pval_175m_15m>0.05 & pval_175m_DCM<0.05 & est_DCM_15m>0 & est_175m_DCM<0 ~ "DCM > 15m = 175m",
-    pval_DCM_15m<0.05 & pval_175m_15m>0.05 & pval_175m_DCM<0.05 & est_DCM_15m<0 & est_175m_DCM>0 ~ "DCM < 15m = 175m",
-    pval_DCM_15m>0.05 & pval_175m_15m>0.05 & pval_175m_DCM>0.05 ~ "All equal",
-    TRUE ~ "Hard to tell"
-  )) %>%
-  bind_rows(nondiff_feats) %>%
-  mutate(resp_class=factor(resp_class, levels = resp_class_levels)) %>%
-  arrange(resp_class) %>%
-  select(feature, resp_class)
-posthoc_table <- table(posthoc_classes$resp_class)
-posthoc_classes %>%
-  left_join(MT_feats, by="feature") %>%
-  mutate(feature=ifelse(is.na(compound_name), feature, compound_name)) %>%
-  arrange(startsWith(feature, "FT"), desc(avgarea)) %>%
-  group_by(resp_class) %>%
-  summarize(n=n(), example=unique(feature)[1]) %>%
-  setNames(c("Response type", "Number of metabolites", "Example metabolite"))
-```
-
-Approximately half (`r posthoc_table["All equal"]` / `r sum(posthoc_table)`) of all detected compounds showed no significant (p > 0.05 after FDR correction) response to depth.
-The metabolites in these different response classes is shown in the figure below and demonstrates a highly non-uniform distribution in this rank-ordered fold-change space.
-Most points are concentrated in the upper-right corner of the plot, indicating that many metabolites are most abundant in the surface or at the DCM.
-Compounds that decrease linearly with depth form a tight cluster at the very top of the plot, indicating very clear separation between the three depths, while transitioning more towards those compounds that have a maximum at the DCM causes separation to become less distinct.
-While more rare, compounds either not found or found at the lowest relative abundances at the surface are present in the data and shown in the bottom right corner of the plot.
-Notably, the left side of this plot is very sparse, with only three compounds falling into those categories that could be classified as having a minimum at the DCM.
-
-```{r clockplot, fig.height=6, fig.width=6}
-rank_peaks <- MT_peaks %>%
-  group_by(feature) %>%
-  mutate(norm_area=rank(norm_area)) %>%
-  # mutate(value=value/max(value)) %>%
-  # mutate(value=scale(value)[,1]) %>%
-  ungroup()
-clockplot_df <- rank_peaks %>%
-  left_join(filled_file_metadata, by="filename") %>%
-  select(feature, norm_area, depth, filename) %>%
-  group_by(feature) %>%
-  mutate(mean_area=mean(norm_area)) %>%
-  group_by(feature, depth) %>%
-  summarise(mean_fold_change=mean(norm_area/mean_area)) %>%
-  pivot_wider(names_from=depth, values_from=mean_fold_change) %>%
-  left_join(posthoc_classes, by="feature") %>%
-  mutate(x=DCM-1) %>%       # cursed but works??
-  mutate(y=`15m`+DCM/2-1.5) # super fucking cursed
-clock_scale <- 0.66
-axis_xends <- c(-1/2, 1.1, -1/2)*clock_scale
-axis_yends <- c(sqrt(3)/2, 0, -sqrt(3)/2)*clock_scale
-axis_y_dodge <- c(2, 2, -2)*clock_scale/25
-
-color_scale <- c("grey50", hcl(h = seq(15, 375, length=13), l=65, c=100)[1:12], "grey80")
-names(color_scale) <- levels(clockplot_df$resp_class)
-
-clock_gp <- ggplot(clockplot_df) +
-  annotate("segment", x=c(0,0,0), y=c(0,0,0), 
-           xend=axis_xends, yend=axis_yends,
-           arrow=arrow(length=unit(0.02, "npc"), type = "closed")) +
-  annotate("segment", x=c(0,0,0), y=c(0,0,0),
-           xend=-axis_xends, yend=-axis_yends,
-           lty = 2) +
-  annotate("label", x=axis_xends, y=axis_yends+axis_y_dodge,
-           label=c("15m axis", "DCM axis", "175m axis"), 
-           fill="#FFFFFFAA", label.size=0) +
-  geom_point(aes(x=x, y=y, color=resp_class), alpha=0.8) +
-  coord_fixed(xlim = c(-1.1, 1.3)*clock_scale, ylim = c(-1.1, 1.1)*clock_scale) +
-  scale_color_manual(values = color_scale, breaks = names(color_scale)) +
-  theme_void() +
-  theme(legend.position = "right", legend.title = element_blank())
-
-xampleMaker <- function(classes){
-  which_cmpds <- posthoc_classes %>%
-    filter(resp_class%in%classes) %>%
-    left_join(MT_feats, by="feature") %>%
-    select(feature, compound_name, resp_class, avgarea) %>%
-    arrange(is.na(compound_name), desc(avgarea)) %>%
-    group_by(resp_class) %>%
-    slice(1)
-  xamples_df <- which_cmpds %>%
-    left_join(rank_peaks, by="feature") %>%
-    group_by(resp_class) %>%
-    mutate(value=norm_area-min(norm_area)+1) %>%
-    ungroup() %>%
-    mutate(feature=ifelse(is.na(compound_name), feature, compound_name)) %>%
-    left_join(filled_file_metadata, by="filename") %>%
-    select(resp_class, feature, filename, depth, value) %>%
-    mutate(depth=factor(depth, levels = c("175m", "DCM", "15m")))
-  xamples_label_df <- xamples_df %>%
-    group_by(resp_class, feature) %>%
-    summarize(midpt=(max(value)-min(value))/2+min(value))
-  ggplot() +
-    geom_boxplot(aes(x=value, y=depth), data = xamples_df, width=0.8) +
-    geom_text(aes(x=midpt, label=resp_class, color=resp_class), y=4.8, data = xamples_label_df) +
-    geom_text(aes(x=midpt, label=feature, color=resp_class), y=4.1, data = xamples_label_df) +
-    annotate("segment", x=0, y=-Inf, xend=0, yend=3.5) +
-    annotate("segment", x=0, y=3.5, xend=Inf, yend=3.5) +
-    facet_wrap(~factor(resp_class, levels = classes), scales = "free_x") +
-    scale_color_manual(values = color_scale, breaks = names(color_scale)) +
-    theme_void() +
-    coord_cartesian(ylim=c(0.5, 4.5)) +
-    theme(strip.text = element_blank(),
-          legend.position = "none",
-          axis.text.y = element_text())
-}
-xample_gp_top <- xampleMaker(c("15m > DCM = 175m", "15m > DCM > 175m", "175m < DCM = 15m"))
-xample_gp_bot <- xampleMaker(c("175m > DCM = 15m", "15m < DCM < 175m", "DCM > 15m = 175m"))
-
-gridExtra::grid.arrange(xample_gp_top, clock_gp, xample_gp_bot,
-                        layout_matrix=matrix(c(1,2,2,2,3), ncol = 1))
-```
-
-*Figure XX: Plot of metabolite response to depth shown across three axes corresponding to the significance, sign, and magnitude of the post-hoc test results.
-Points aligning with the 15m axis correspond to compounds with most of their abundance found in the surface ocea; points far to the right side correspond to compounds that are found only at the deep chlorophyll maximum; points found at the bottom of the plot are those compounds that increased more or less linearly with depth.
-Distance from plot center corresponds to increasing fold-change estimate between the depth categories, with most compounds having insignificant fold change with depth and falling in the center of the plot.
-A few categories are shown as boxplots with selected example compounds above and below the central figure.*
-
-`r unname(posthoc_table["15m > DCM = 175m"])` compounds were found that had significantly higher values at 15 meters than the DCM or 175 meters.
-N6-Acetyl-L-lysine is the only known compound in this group and is a borderline inclusion, as its depth profile could also be characterized as linearly decreasing - however, the DCM and 175 meter samples were not statistically different.
-
-```{r MAA search, eval=FALSE}
-MAA_masses <- readxl::read_xlsx("../metadata/raw_data/MAA_masses.xlsx", sheet = 2) %>%
-  select(1, 3, 5, 7, 9) %>%
-  setNames(c("MAA", "formula", "mz", "rt", "MSMS")) %>%
-  filter(MAA!="Common")
-
-MAA_maybe <- MAA_masses %>%
-  pmap_dfr(function(...){
-    row_data <- data.frame(...)
-    MT_feats %>%
-      filter(mzmed%between%pmppm(row_data$mz, 10)) %>%
-      mutate(MAA_name=row_data$MAA, MAA_rt=row_data$rt, MAA_formula=row_data$formula)
-  }) %>%
-  left_join(posthoc_classes, by="feature") %>%
-  select(feature, mzmed, rtmed, avgarea, MAA_name, MAA_rt, resp_class)
-
-msms_files <- list.files("../mzMLs/pos/MSMS", pattern = "180821.*DDA", full.names = TRUE)
-msmsdata <- grabMSdata(msms_files)
-
-MAA_msms <- MAA_maybe %>%
-  mutate(rtmed=rtmed/60) %>%
-  pmap_dfr(function(...){
-    row_data <- data.frame(...)
-    msmsdata$MS2[rt%between%(row_data$rtmed+c(-1, 1))][premz%between%pmppm(row_data$mzmed, 50)] %>%
-      mutate(MAA=row_data$MAA_name)
-  }) %>%
-  group_by(MAA, filename, rt) %>%
-  arrange(desc(int)) %>%
-  # slice(5) %>%
-  summarize(MSMS=paste(fragmz, int, sep = ", ", collapse = "; "))
-
-MAA_msdata <- MAA_maybe %>%
-  mutate(rtmed=rtmed/60) %>%
-  pmap_dfr(function(...){
-    row_data <- data.frame(...)
-    v <- msmsdata$MS1[mz%between%pmppm(row_data$mzmed)]
-    v$MAA <- row_data$MAA_name
-    v
-  })
-# Plot all the peaks' MS1 and add lines at their MSMS scan RTs
-ggplot() +
-  geom_line(aes(x=rt, y=int, group=filename), data = MAA_msdata) +
-  geom_vline(aes(xintercept=rt, group=filename), data = MAA_msms, color="red") +
-  facet_wrap(~MAA, scales = "free_y") +
-  xlim(6, 12)
-
-aster_db_MSMS <- MAA_masses %>%
-  filter(MAA=="Asterina-330" & !is.na(MSMS)) %>%
-  separate_rows(MSMS, sep = "; ") %>%
-  separate(MSMS, into = c("fragmz", "int"), sep = ", ", convert = TRUE)
-MAA_msms %>%
-  filter(MAA=="Asterina-330") %>%
-  # mutate(which_peak=ifelse(rt>8, "tallskinnysecond", "shortfatfirst")) %>%
-  separate_rows(MSMS, sep = "; ") %>%
-  separate(MSMS, into = c("fragmz", "int"), sep = ", ", convert = TRUE) %>%
-  group_by(filename) %>%
-  mutate(int=int/max(int)) %>%
-  filter(int>0.05) %>%
-  ggplot(aes(x=fragmz, y=int)) +
-  geom_vline(aes(xintercept=fragmz), data=aster_db_MSMS, color="red") +
-  geom_point() +
-  geom_segment(aes(yend=0, xend=fragmz)) +
-  # facet_wrap(~which_peak) +
-  theme_bw()
-# plotly::ggplotly()
-```
-
-[Should I split off the MAA analysis into separate section to be discussed after walking through all the clockplot categories?]
-
-However, several compounds in this category were putatively identified as mycosporine-like amino acids (MAAs) using *m/z* matching and MS^2^ fragmentation information.
-Asterina-330, shinorine, and palythine were simply identified by checking the fragment masses against those provided in @parailloux2020 because they were the only peak in the chromatogram and were assigned a confidence level of 2 according to MSI reporting standards (@sumner2007).
-Porphyra-334 had two possible peaks, only the earlier eluting of which was was fragmented in the DDA scan and which matched MSMS spectra in online databases, but without fragmentation for the second peak we cannot be confident of this identification.
-Both peaks have the same type of response to depth (essentially zero abundance outside of the 15 meter samples) but have minimal correlation with each other in the surface samples, confirming that this is unlikely to be the same compound somehow split or poorly-integrated.
-
-```{r porphyra dupe check, eval=FALSE}
-porph_data <- MAA_maybe %>%
-  filter(MAA_name=="Porphyra-334") %>%
-  left_join(MT_peaks, by="feature") %>%
-  select(feature, norm_area, filename) %>%
-  pivot_wider(values_from = norm_area, names_from = feature) %>%
-  left_join(filled_file_metadata, by="filename")
-
- porph_data %>%
-  ggplot() +
-  geom_point(aes(x=FT1409, y=FT1410)) +
-  # geom_point(aes(x=log10(FT1409), y=log10(FT1410))) +
-  facet_wrap(~depth, scales = "free")
-
-porph_data %>%
- filter(depth=="15m") %>%
-  with(cor(FT1409, FT1410))
-```
-
-Mycosporine-glycine and mycosporine-2-glycine were more difficult to annotate with a specific confidence level.
-The peak tentatively identified as mycosporine-2-glycine shared no fragments with the list of common MAA fragments that they put forward and probably remains at confidence level 3.
-Mycosporine-glycine had two features in the chromatogram at an *m/z* ratio of 246.0972, one short and wide feature eluting earlier around 9.4 minutes and a second taller and sharper feature at 10.2 minutes.
-DDA data was collected for both features and both shared the common MAA fragment of *m/z* 122.0602 but no others.
-The earlier eluting short feature had a unique low-intensity fragment replicated across multiple collision energies at an *m/z* of 228.0861, likely representing a water loss.
-It therefore seems likely that one of these features is mycosporine-glycine but this information alone cannot be used to identify which of the two it is.
-
-It was also difficult to assign a specific confidence level to the isomer couple palythene/usujirene.
-Three features were found in the *m/z* window around 285.1444 +/- 10 ppm, one of these was unlikely because its retention time was high (12.4 minutes), the feature itself was 10-20x smaller in area than the other two, and no MSMS scans were obtained.
-We first considered whether our chromatographic conditions could have separated these isomers to produce the two remaining peaks, but closer investigation of the MSMS spectra revealed that several distinguishing fragments were produced upon fragmentation at multiple voltages.
-The earlier eluting feature with a retention time of 6.9 minutes had two fragments not found in the later eluting feature (*m/z*s of 241.0817 and 226.1310) and the later feature eluting around 8.5 minutes had fragments not seen in the earlier (*m/z*s of 252.1101, 239.1025, and 225.1234).
-Given that palythene and usujirene are isomers differing only in the cis/trans positioning of a double bond, these distinguishing fragments make it unlikely that one feature is palythene and the other is usujirene and more likely that one of the features is an unknown metabolite.
-Both features shared several fragments but none of them were MAA-diagnostic as named in @parailloux2020 (*m/z*s of 270.1208, 197.0921, and a curiously small fragment at 60.0451).
-
-Nearly all of the putative MAA peaks described above fell into this first category of compounds that were found abundantly at the surface but with very low abundances at the DCM indistinguishable from the abundances found in the 175 meter samples.
-For shinorine and the later-eluting mycosporine-glycine, the samples collected at the DCM contained significantly more abundant quantities of the compound than those at 175m, which is our second category of interest in Figure XX above.
-
-Compounds that decreased linearly with depth composed the second-largest category of compounds that responded to depth, with `r unname(posthoc_table["15m > DCM > 175m"])` compounds.
-These compounds had highest concentrations at the surface, middling concentrations at the DCM, and lowest concentrations at 175 meters.
-Several known compounds fall into this category, notably dimethylsulfoniopropionate (DMSP) and gonyol, the amino acid leucine, and several betaine molecules - proline betaine and carnitine.
-Three unknown molecules have also been tentatively identified as betaines in this category - beta-Alanine betaine, for which our lab obtained a standard after these samples were run and can match via retention time ordering, *m/z*, and MSMS for a confidence level very close to 1, and homoserine betaine and threonine betaine which have been identified by MS1 *m/z* matching and MSMS betaine fragments (58.0659, 59.0737, 60.0815) for a confidence level of 2.
-With retention time ordering, we were also able to annotate each peak uniquely because we have homoserine and threonine in our standards and know that threonine elutes before homoserine.
-
-```{r MSMS for betaines, eval=FALSE}
-msms_files <- list.files("../mzMLs/pos/MSMS", pattern = "180821.*DDA", full.names = TRUE)
-msmsdata <- grabMSdata(msms_files)
-
-mz_i <- 162.113019 #homoserine betaine, threonine betaine, carnitine
-# mz_i <- 132.1018 # Leucine, isoleucine, beta-alanine betaine, alanine betaine(?)
-
-msmsdata$MS1[mz%between%pmppm(mz_i)] %>%
-  ggplot() +
-  geom_line(aes(x=rt, y=int, group=filename)) +
-  geom_vline(aes(xintercept=rt), 
-             data = msmsdata$MS2[premz%between%pmppm(mz_i)],
-             color="red")
-
-obs_MSMS <- msmsdata$MS2[premz%between%pmppm(mz_i)] %>%
-  filter(rt%between%(8.2+c(-1, 1))) %>%
-  filter(!fragmz%between%pmppm(mz_i)) %>%
-  group_by(filename) %>%
-  mutate(int=int/max(int))
-theo_MSMS <- "https://raw.githubusercontent.com/IngallsLabUW/Ingalls_Standards/master/MSMS/data_processed/Ingalls_Lab_Standards_MSMS.csv" %>%
-  read_csv(show_col_types = FALSE) %>%
-  filter(compound_name=="Glycine betaine") %>% 
-  separate_rows(MS2, sep = "; ") %>%
-  separate(MS2, into = c("fragmz", "int"), sep = ", ", convert = TRUE) %>%
-  filter(!fragmz%between%pmppm(132.1018)) %>%
-  group_by(voltage) %>%
-  mutate(int=int/max(int))
-
-ggplot() +
-  geom_point(aes(x=fragmz, y=int), data = obs_MSMS) +
-  geom_segment(aes(x=fragmz, y=int, xend=fragmz, yend=0), data = obs_MSMS) +
-  geom_point(aes(x=fragmz, y=-int), data = theo_MSMS, col="blue") +
-  geom_segment(aes(x=fragmz, y=-int, xend=fragmz, yend=0), data = theo_MSMS, col="blue") +
-  geom_hline(yintercept = 0) +
-  facet_wrap(~voltage)
-```
-
-The group of compounds with equally high values at the surface and the DCM and significantly lower values at 175 meters contained the most compounds (`r unname(posthoc_table["175m < DCM = 15m"])`, or `r paste0(signif(unname(posthoc_table["175m < DCM = 15m"])/sum(posthoc_table), 3)*100, "%")`) other than the category of those that didn't respond to depth at all.
-This category was dominated in raw peak area [change to concentration w targeted data] by two known compounds, trigonelline and dimethylsulfonioacetate (DMS-Ac), together composing nearly 30% of the signal detected in this size class.
-Other known compounds in this class include the nucleobase cytosine, the nucleoside adenosine and its deoxy form, creatine, choline, and taurine as well as the amino acids alanine, glutamate, and glutamine.
-Betonicine and turicine fell into this category but could not be separated by HILIC chromatography as stereoisomers.
-Acetylcholine was also identified in this category at slightly lower confidence (FT0432, authentic standard purchased and run after this data set and tentative annotation done by *m/z* and retention time matching) as well as hypoxanthine (FT0357) in the same way.
-An additional large unknown was identified with the same mass as betonicine but eluting a minute earlier perhaps corresponding to 3-Dehydrocarnitine or N-Acetyl-valine based on *m/z* matching.
-
-Compounds abundant at the DCM were also common, with a total of `r sum(unname(posthoc_table[c("DCM max 15m mid", "DCM > 15m = 175m", "DCM max 175m mid")]))` features across the three depth categories that had maxima at the DCM.
-Many traditional osmolytes were found in this category, including glycine betaine, trimethylamine N-oxide, homarine, and ectoine.
-Several amino acids and their derivatives fell here - glycine, proline, arginine, tyrosine, asparagine, threonine (combined with homoserine as the two did not appear to separate in retention time [revisit this - I claim that we can tell the betaine versions apart because these DO separate]) and beta-alanine.
-The nucleobase guanine also had a maximum at the DCM.
-This collection of compounds was notable to us as containing several nitrogen-rich compounds, which inspired the followup analysis of C:N ratio in section XX below.
-Several unknown compounds with very large average peak areas were also categorized as having DCM maxima, most notably MT_POS_FT0483 (*m/z* of 151.1441, retention time about 7 minutes) and MT_POS_FT0510 ([need mz and rt info here]) were in the top 20 largest peak areas but neither belonged to our targeted list.
-
-Compounds most abundant at depth are unexpected given general trends of particulate matter decreasing with depth, but prior research has noted several compounds that accumulate with depth (@healarseno?).
-Here, we find significantly more compounds with this trend than we expected with a total of `r sum(unname(posthoc_table[c("175m > DCM = 15m", "15m < DCM < 175m")]))` compounds that had highest peak areas in samples taken from 175 meters.
-Arsenobetaine was the only identified compound in these two depth response classes and had a very linear response (Fig above).
-It's unknown whether these other compounds have a similar role as toxic element shunts or are simply some kind of degradation product, but some of them nonetheless have quite large peak areas.
-Features MT_POS_FT0617 and MT_POS_FT0618 are especially interesting because they share an *m/z* ratio (170.11757) and differ in retention time by less than a minute (8.3 vs 9 minutes), but both belong to this category of compounds that are most abundant at 175 meters.
-
-Compounds that have a minimum at the DCM are even rarer than those that accumulate with depth, requiring either rapid drawdown at the DCM (as might be the case for a chlorophyll or other DCM-max precursor) or increased synthesis above and below it.
-Nonetheless, a total of `r sum(unname(posthoc_table[c("DCM min 15m mid", "DCM < 15m = 175m", "DCM min 175m mid")]))` compounds were identified with this trend - none of which belonged to our list of targeted compounds and therefore none of which could be identified at high confidence.
-Despite the use of stringent false-discovery correction methods outlined above, it is entirely possible these compounds are simply statistical anomalies and it should be noted that the fold-change differences shown in the figure above are relatively weak.
-
-### Predicting metabolite response classes from molecular formulae
-
-[Need to update intro with ideas for this section, justify why we did it]
-
-[I don't know how to explain why I'm using the y-axis coordinates here but it's basically  proxy for whether a metabolite's more abundant at the surface or at depth.
-High values correspond to three groups where 15m values are super high, and low (negative) values correspond to groups where a metabolite is mostly found at 175 meters] When we regressed the vertical axis of Figure XX above against the derived C:N ratio of each metabolite, we expected to find a positive trend with those metabolites most abundant at the surface having the highest C:N ratio based on the increased abundance of light energy and the decreased availability of nitrate at the surface relative to depth.
-However, we found no significant trend between C:N and the y-coordinate of the metabolites.
-This absence of a trend was robust when using only known metabolites or when using all metabolites for which a molecular formula could be calculated (see Methods).
-
-```{r fig.height=3, fig.width=5}
-formula2elements <- function(formula_vec){
-  split_formulas <- formula_vec %>%
-    gregexpr(pattern = "[A-Z][a-z]*[0-9]*") %>%
-    regmatches(x = formula_vec)
-  elements_in <- lapply(split_formulas, gsub, pattern = "[0-9]", replacement = "")
-  element_counts <- lapply(split_formulas, gsub, pattern = "[A-z]", replacement = "")
-  element_counts <- lapply(element_counts, function(x){
-    x[!nchar(x)]<-1
-    return(as.numeric(x))
-  })
-  mapply(`names<-`, element_counts, elements_in, SIMPLIFY = FALSE)
-}
-
-# Targeted compounds
-targ_cn_ratios <- clockplot_df %>%
-  select(feature, y, resp_class) %>%
-  left_join(MT_feats, by="feature") %>%
-  filter(compound_type!="Internal Standard") %>%
-  mutate(elem_list=formula2elements(formula)) %>%
-  unnest_wider(elem_list) %>%
-  mutate(CN_ratio=C/N) %>%
-  arrange(CN_ratio) %>%
-  select(feature, y, CN_ratio, resp_class) %>%
-  mutate(method_type="Targeted")
-
-# Untargeted compounds
-untarg_cn_ratios <- clockplot_df %>%
-  select(feature, y, resp_class) %>%
-  left_join(MT_feats, by="feature") %>%
-  mutate(elem_list=formula2elements(calc_formula)) %>%
-  unnest_wider(elem_list) %>%
-  mutate(CN_ratio=C/N) %>%
-  arrange(CN_ratio) %>%
-  select(feature, y, CN_ratio, resp_class) %>%
-  mutate(method_type="Untargeted")
-
-all_cn_ratios <- targ_cn_ratios %>%
-  bind_rows(untarg_cn_ratios) %>%
-  filter(!is.na(CN_ratio))
-
-lm_output <- all_cn_ratios %>%
-  nest(data=-method_type) %>%
-  mutate(lm_pval=map_dbl(data, ~broom::tidy(lm(.x$CN_ratio~.x$y))$p.value[2])) %>%
-  mutate(lm_label=paste("Regression\np-value:", round(lm_pval, 3)))
-
-all_cn_ratios %>%
-  ggplot(aes(x=y, y=CN_ratio)) +
-  geom_point(aes(color=resp_class)) +
-  scale_color_manual(values = color_scale, breaks = names(color_scale), limits=force) +
-  geom_smooth(method = "lm", formula = "y~x") +
-  geom_label(aes(label=lm_label), data = lm_output, 
-             x=Inf, y=Inf, hjust=1, vjust=1, label.size = 0) +
-  facet_wrap(~method_type, nrow=1) +
-  xlim(-0.67, 0.67) +
-  ylab("C:N ratio") +
-  xlab("y-axis coordinate from Figure XX above") +
-  theme_bw() +
-  theme(legend.position = "top", legend.title = element_blank()) +
-  guides(color = guide_legend(nrow = 3))
-```
-
-However, while performing preliminary data exploration we did find that compounds that decrease monotonically with depth did have significantly higher C:N ratios than those that remain at comparable abundances between the surface and the DCM.
-
-```{r, fig.width=5}
-t_labels <- all_cn_ratios %>%
-  filter(resp_class%in%c("15m > DCM > 175m", "175m < DCM = 15m")) %>%
-  nest(data=-method_type) %>%
-  mutate(t_pval=map_dbl(data, ~broom::tidy(t.test(.x$CN_ratio~.x$resp_class))$p.value)) %>%
-  mutate(t_label=paste("t-test p-value:", round(t_pval, 5)))
-
-all_cn_ratios %>%
-  filter(resp_class%in%c("15m > DCM > 175m", "175m < DCM = 15m")) %>%
-  ggplot() +
-  geom_violin(aes(x=resp_class, y=CN_ratio, color=resp_class), lwd=0.8) +
-  geom_boxplot(aes(x=resp_class, y=CN_ratio, color=resp_class), lwd=0.8, 
-               width=0.2, outlier.shape = NA, fill="#FFFFFF33") +
-  geom_jitter(aes(x=resp_class, y=CN_ratio), color="grey40", width = 0.1) +
-  scale_color_manual(values = color_scale, breaks = names(color_scale)) +
-  geom_label(aes(label=t_label), data = t_labels, 
-           x=Inf, y=Inf, hjust=1, vjust=1, label.size = 0) +
-  facet_wrap(~method_type, nrow=1) +
-  ylab("C:N ratio") +
-  xlab(NULL) +
-  theme_bw() +
-  theme(legend.position = "none")
-```
-
-[Optional additional discussion could occur based on the compounds with a DCM max continuing this trend, implying that if anything it's the compounds that are most abundant at the DCM that have the lowest C:N ratios, but then goes back up again at 175m unexpectedly, see temporary fig below]
-
-```{r}
-all_cn_ratios %>%
-  ggplot() +
-  geom_boxplot(aes(x=resp_class, y=CN_ratio, color=resp_class), lwd=1.2, varwidth = TRUE) +
-  scale_color_manual(values = color_scale, breaks = names(color_scale)) +
-  facet_wrap(~method_type, nrow=1) +
-  ylab("C:N ratio") +
-  xlab(NULL) +
-  theme_bw() +
-  theme(legend.position = "none") +
-  coord_flip()
-```
-
-### Metabolite response to SLA
-
-```{r metabolite sla response}
-# For metabolites
-trend_w_sla <- MT_peaks %>%
-  left_join(filled_file_metadata %>% select(filename, depth, sla), by="filename") %>%
-  group_by(feature) %>%
-  mutate(value=rank(norm_area)) %>%
-  ungroup() %>%
-  nest(data=-c(feature, depth)) %>%
-  mutate(lmtest=map(data, ~lm(.x$value~.x$sla))) %>%
-  mutate(lmtidy=map(lmtest, broom::tidy)) %>%
-  suppressWarnings() %>%
-  unnest(lmtidy) %>%
-  mutate(term=ifelse(term=="(Intercept)", "intercept", "slope")) %>%
-  pivot_wider(names_from = term, values_from = estimate) %>%
-  fill(intercept) %>%
-  filter(!is.na(slope)) %>%
-  select(feature, depth, data, slope, intercept, p.value) %>%
-  mutate(p.value=p.adjust(p.value, method = "fdr")) %>%
-  mutate(p_type=case_when(
-    p.value<0.05 & slope>0 ~ "Positive trend",
-    p.value<0.05 & slope<0 ~ "Negative trend",
-    TRUE ~ "No trend"
-  )) %>%
-  select(feature, depth, p_type, slope, intercept) %>%
-  mutate(p_type=factor(p_type, levels = c(
-    "No trend", "Negative trend", "Positive trend")
-  ))
-alldepth_sla_response <- trend_w_sla %>%
-  group_by(feature) %>%
-  summarize(n_depths=sum(p_type!="No trend")) %>%
-  count(n_depths) %>%
-  pull(n, n_depths)
-sla_trend_table <- trend_w_sla %>%
-  count(depth, p_type)
-sla_resp_types <- trend_w_sla %>%
-  select(-slope, -intercept) %>%
-  pivot_wider(names_from = depth, values_from = p_type) %>%
-  mutate(sla_resp_type=paste(`15m`, `DCM`, `175m`)) %>%
-  group_by(sla_resp_type) %>%
-  # filter(`DCM`=="Negative trend" & `175m`=="Positive trend") %>% left_join(MT_feats)
-  summarize(`15m`=unique(`15m`), `DCM`=unique(`DCM`), `175m`=unique(`175m`), n=n()) %>%
-  select(-sla_resp_type) %>%
-  arrange(desc(n)) %>%
-  filter(n>1) %>%
-  mutate(across(-n, ~case_when(
-    .x=="No trend" ~ "---",
-    .x=="Positive trend" ~ "Pos",
-    .x=="Negative trend" ~ "Neg"
-  )))
-
-# Metabolite plot
-metab_gp <- MT_peaks %>%
-  group_by(feature) %>%
-  mutate(value=rank(norm_area)) %>%
-  ungroup() %>%
-  left_join(filled_file_metadata %>% select(filename, depth, sla), by="filename") %>%
-  right_join(trend_w_sla, by = c("feature", "depth")) %>%
-  mutate(plotcol=ifelse(p_type!="No trend", feature, NA)) %>%
-  ungroup() %>%
-  ggplot(aes(x=sla, y=value, group=feature, color=plotcol, label=feature)) +
-  geom_point() +
-  geom_abline(aes(slope=slope, intercept=intercept, color=plotcol), lwd=0.8) +
-  facet_grid(depth~p_type) +
-  theme_bw() +
-  theme(legend.position = "none") +
-  ggtitle("A) Metabolites")
-```
-
-As expected from the multivariate PERMANOVA results detailed above, we find that relatively few compounds are sensitive to sea level anomalies.
-`r unname(alldepth_sla_response["0"])` metabolites were independent of SLA at all three depths, `r unname(alldepth_sla_response["1"])` had a significant trend at exactly one depth, `r unname(alldepth_sla_response["2"])` responded to SLA at two depths, and only `r unname(alldepth_sla_response["3"])` metabolite (MT_POS_FT0709, an unknown with *m/z* of 180.1019 and retention time ~2 minutes) had a trend with SLA at all depths, always negative.
-Notably, compounds at the DCM are much more likely to have a negative trend with SLA, while compounds at 175 meters are likely to have an opposite trend.
-The distribution of metabolites across these categories is shown below in Figure XX.
-This effect was rarely due to a given compound switching from a positive to negative trend (just `r sla_resp_types$n[4]` of the `r nrow(MT_feats)` metabolites switched), but was instead largely driven by compounds with no trend at the DCM switching to a positive trend at 175 meters (`r sla_resp_types$n[2]` / `r nrow(MT_feats)`) and compounds with no trend at other depths trending negatively with SLA at the DCM (`r sla_resp_types$n[3]` / `r nrow(MT_feats)`).
-
-```{r metadata sla response}
-# Metadata calcs and plot
-metadata_sla_df <- filled_file_metadata %>%
-  filter(cruise=="MT") %>%
-  filter(samp_type=="Smp") %>%
-  select(untripl, depth, sla, no23_um:last_col()) %>%
-  pivot_longer(cols = no23_um:last_col()) %>%
-  group_by(name) %>%
-  mutate(value=rank(value)) %>%
-  ungroup() %>%
-  distinct()
-trend_w_sla_meta <- metadata_sla_df %>%
-  nest(data=-c(name, depth)) %>%
-  mutate(lmtest=map(data, ~lm(.x$value~.x$sla))) %>%
-  mutate(lmtidy=map(lmtest, broom::tidy)) %>%
-  suppressWarnings() %>%
-  unnest(lmtidy) %>%
-  mutate(term=ifelse(term=="(Intercept)", "intercept", "slope")) %>%
-  pivot_wider(names_from = term, values_from = estimate) %>%
-  fill(intercept) %>%
-  filter(!is.na(slope)) %>%
-  select(name, depth, data, slope, intercept, p.value) %>%
-  mutate(p.value=p.adjust(p.value, method = "fdr")) %>%
-  mutate(p_type=case_when(
-    p.value<0.05 & slope>0 ~ "Positive trend",
-    p.value<0.05 & slope<0 ~ "Negative trend",
-    TRUE ~ "No trend"
-  )) %>%
-  select(name, depth, p_type, slope, intercept) %>%
-  mutate(p_type=factor(p_type, levels = c(
-    "No trend", "Negative trend", "Positive trend")
-  ))
-meta_gp <- metadata_sla_df %>%
-  right_join(trend_w_sla_meta, by = c("name", "depth")) %>%
-  mutate(plotcol=ifelse(p_type!="No trend", name, NA)) %>%
-  ungroup() %>%
-  mutate(name=factor(name, levels = factor_levels, labels = factor_labels)) %>%
-  ggplot(aes(x=sla, y=value, group=name, color=name, label=name)) +
-  geom_vline(xintercept=0, color="grey75") +
-  geom_point() +
-  geom_abline(aes(slope=slope, intercept=intercept, color=name), lwd=0.8) +
-  facet_grid(depth~p_type) +
-  theme_bw() +
-  theme(axis.title.y = element_blank(), axis.text.y = element_blank()) +
-  ggtitle("B) Environmental parameters")
-
-gridExtra::grid.arrange(metab_gp, meta_gp, layout_matrix=matrix(c(1,1,2,2,2), nrow = 1))
-```
-
-The response to SLA was weakest at the surface when measured by the absolute number of compounds in each category.
-At a depth of 15 meters, only `r sla_trend_table$n[2]` compounds had a significant negative trend with SLA (more abundant in cyclone than anticyclone) while `r sla_trend_table$n[3]` had a positive trend, together representing `r paste0(round(sum(sla_trend_table$n[2:3])/(nrow(trend_w_sla)/3), 3)*100, "%")` of the total compounds.
-No known compounds had negative SLA trends despite particulate N and picoeukaryote abundance falling into this category, but known compounds with positive SLA trends were 5-Oxoproline and N6-Acetyl-L-lysine as well as the environmental parameters DOP/DON, nitrate + nitrite, and phosphate.
-
-At the DCM, the proportion of compounds that responded to SLA increased to `r paste0(round(sum(sla_trend_table$n[5:6])/(nrow(trend_w_sla)/3), 3)*100, "%")`, with `r sla_trend_table$n[5]` metabolites trending negatively and `r sla_trend_table$n[6]` in the opposite direction.
-Metabolites in this first category included the abundant osmolyte homarine and its [similar molecules] nicotinic acid and trigonelline.
-Carnitine and O-acetylcarnitine are also included as well as adenosine and its degradation product inosine, plus the other nucleoside cytidine.
-Other known molecules were the osmolyte DMS-Ac, the amino acids alanine and proline, the glycine derivative sarcosine, and the compound tentatively identified as threonine betaine (see above).
-The environmental parameters that were most strongly negative were particulate carbon and particulate nitrogen.
-Only one unidentified metabolite had the opposite trend of increased abundance in the anticyclone, as well as DON and DOP.
-
-Finally, at 175 meters, nearly 10% of all measured compounds (`r paste0(round(sum(sla_trend_table$n[8:9])/(nrow(trend_w_sla)/3), 3)*100, "%")`) changed with SLA.
-`r sla_trend_table$n[8]` were less abundant in the cyclone than the anticyclone, and `r sla_trend_table$n[9]` increased in abundance across the eddy dipole.
-Compounds with a negative SLA trend consisted of two unknown molecules as well as N + N, and phosphate.
-Compounds with a positive trend were much more numerous and included several betaines (glycine betaine, proline betaine, betonicine/turicine, and the pseudo-betaine arsenobetaine) plus guanine, gonyol, dexpanthenol, and the amino acids leucine and tyrosine.
-Alanine, O-acetylcarnitine, and the putatively identified threonine betaine also fell into this class, making them among the few compounds that actually switched the sign of their trend between the DCM and 175 meters.
-
-### Investigating the relative importance of sea level anomaly on a higher-resolution depth scale
-
-```{r MC diff stats}
-MC_feats <- all_feats %>%
-  filter(cruise=="MC") %>%
-  # filter(polarity=="pos") %>%
-  select(-classes, -CSI_name) %>%
-  rename("calc_formula"=formula) %>%
-  left_join(clean_stans, by=c("compound_name", "polarity"))
-MC_peaks <- MC_feats %>%
-  left_join(all_peaks, by="feature") %>%
-  left_join(filled_file_metadata, by="filename") %>%
-  filter(samp_type=="Smp") %>%
-  select(feature, norm_area, filename)
-
-most_stat_diff <- MC_peaks %>%
-  left_join(filled_file_metadata, by="filename") %>%
-  select(feature, filename, station, depth, norm_area) %>% 
-  nest(data=-feature) %>%
-  ### Classic t-test
-  # mutate(test=map(data, ~broom::tidy(t.test(.x$norm_area~.x$station)))) %>%
-  # unnest(test) %>%
-  ### Non-parametric rank-sum test
-  # mutate(test=map(data, ~broom::tidy(wilcox.test(.x$norm_area~.x$station)))) %>%
-  # unnest(test) %>%
-  # rename(estimate=statistic) %>%
-  ### Permutational t-test
-  mutate(test=map(data, ~broom::tidy(aovp(.x$norm_area~.x$station, seqs = FALSE,
-                                          settings=FALSE)))) %>%
-  unnest(test) %>%
-  rename(p.value=`Pr(Prob)`) %>%
-  mutate(estimate=1) %>%
-  filter(term==".x$station1") %>%
-  ### No matter what test, run the below
-  mutate(pval=p.adjust(p.value, method = "fdr")) %>%
-  arrange(pval) %>%
-  select(feature, data, estimate, pval) %>%
-  mutate(pval_cats = case_when(
-    pval < 0.001 ~ "Highly signif",
-    pval < 0.05 ~ "Signif",
-    pval >=0.05 ~ "Nonsig"
-  )) %>%
-  mutate(fold_change=map_dbl(data, ~mean(.x$norm_area[.x$station=="L1"])/
-                             mean(.x$norm_area[.x$station=="L2"])))
-
-pval_cat_counts <- most_stat_diff %>%
-  pull(pval_cats) %>% table()
-
-# most_stat_diff %>%
-#   left_join(all_feats) %>%
-#   filter(!is.na(compound_name)) %>%
-#   select(pval_cats, compound_name, fold_change) %>%
-#   arrange(desc(fold_change)) %>%
-#   group_by(pval_cats) %>%
-#   group_split()
-```
-
-We found `r sum(pval_cat_counts)` features in this new set of data using the untargeted methods described above.
-Of those, `r unname(pval_cat_counts["Highly signif"])` were significantly different between the two eddies with p-values less than 0.001 and `r unname(pval_cat_counts["Signif"])` were significantly different with p-values between 0.001 and 0.05 after controlling the false discovery rate.
-All known compounds found to be significantly different were elevated in the cyclonic samples, but several unknown molecules had the opposite trend and were enriched in the anticyclone.
-
-```{r MC targ concentrations, eval=FALSE}
-highly_sig_feats <- most_stat_diff %>%
-  left_join(all_feats, by="feature") %>%
-  filter(!is.na(compound_name)) %>%
-  filter(pval_cats=="Highly signif") %>%
-  pull(compound_name)
-MC_targ <- "../targeted/made_data/longdata.csv" %>%
-  read_csv(show_col_types = FALSE) %>%
-  filter(compound_name%in%highly_sig_feats) %>%
-  mutate(filename=paste0(filename, ".mzML")) %>%
-  left_join(filled_file_metadata, by="filename") %>%
-  filter(cruise=="MC") %>%
-  filter(samp_type=="Smp")
-MC_targ %>%
-  group_by(compound_name, station) %>%
-  summarize(mean_val=mean(nM)) %>%
-  pivot_wider(names_from = station, values_from = mean_val) %>%
-  mutate(fold_change=L1/L2) %>%
-  arrange(desc(fold_change))
-# MC_targ %>%
-#   filter(compound_name=="Isethionic acid") %>%
-#   group_by(station) %>%
-#   summarise(max_nM=max(nM), min_nM=min(nM))
-```
-
-Eleven known compounds were highly sensitive (permutational t-test p-value < 0.001 after false discovery rate correction) to sampling location.
-Isethionic acid increased from 0.03 nM in the anticyclone to 0.12 nM in the cyclone, a 4x fold change and the largest difference among the known compounds.
-This molecule separated completely between the eddies, with values less than 0.08 nM found exclusively in the anticyclonic feature and all larger values exclusively in the cyclone.
-
-Compounds that were significantly enrighed at the DCM in the cyclone relative to the anticyclone at a less-dramatic p-value cutoff of 0.05 included several amino acids (glycine as noted above, alanine, asparagine, and serine) as well as amino acid derivatives (sarcosine, choline, homoserine, theanine, and 4-aminobutyric acid).
-Homarine and its demethylated form nicotinic acid were also distinctive between the eddies, as well as a surprising number of sulfur-containing compounds (the aforementioned isethionic acid, plus DMSP, DHPS, and N-methyltaurine.) Other compounds included the betaines proline betaine, carnitine, and the putative threonine betaine, plus AMP, cytidine, and 2-O-alpha-D-Glucosylglycerol.
-
-```{r MC volcano plot}
-most_stat_diff %>%
-  left_join(MC_feats, by = "feature") %>%
-  filter(!is.na(compound_name)) %>%
-  slice(1:20) %>%
-  unnest(data) %>%
-  mutate(eddy_type=ifelse(station=="L1", "Cyclonic", "Anticyclonic")) %>%
-  ggplot() +
-  geom_boxplot(aes(x=norm_area, y=depth, color=eddy_type)) +
-  facet_wrap(~compound_name, scales = "free_x") +
-  scale_color_manual(breaks = c("Cyclonic", "Anticyclonic"), values = c(muted("blue"), muted("red")))
-
-MT_type <- trend_w_sla %>%
-  filter(depth=="DCM") %>%
-  left_join(all_feats, by="feature") %>%
-  select(feature, p_type, group_feat) %>%
-  mutate(p_type=as.character(p_type))
-
-most_stat_diff %>%
-  mutate(pval=ifelse(pval==0, 0.001, pval)) %>%
-  left_join(MC_feats, by = "feature") %>%
-  mutate(which_stat_bigger=ifelse(estimate>0, "L1", "L2")) %>%
-  mutate(fold_change=log2(fold_change)) %>%
-  mutate(cmpd_known=!is.na(compound_name)) %>%
-  mutate(compound_name=ifelse(is.na(compound_name), "", compound_name)) %>%
-  left_join(MT_type, by="group_feat") %>%
-  mutate(p_type=ifelse(is.na(p_type), "Not found in MT", p_type)) %>%
-  ggplot(aes(x=fold_change, y=-log10(pval), label=compound_name)) +
-  geom_vline(xintercept = 0) +
-  geom_hline(yintercept = 0) +
-  geom_hline(yintercept = -log10(0.05), lty=2) +
-  geom_point(aes(color=p_type)) +
-  geom_text_repel(max.overlaps = 10, min.segment.length = 0) +
-  annotate("label", x=Inf, y=Inf, label="Enriched in cyclone", 
-           hjust=1, vjust=1, label.r=unit(0, "in")) +
-  annotate("label", x=-Inf, y=Inf, label="Enriched in anticyclone", 
-           hjust=0, vjust=1, label.r=unit(0, "in")) +
-  theme_bw() +
-  xlim(-4.5, 4.5) +
-  scale_color_manual(breaks=c("Negative trend", "No trend", "Not found in MT", "Positive trend"),
-                     values = c(muted("red"), "grey50", "grey80", muted("blue")), name="Trend in eddy\ntransect")
-```
-
-*Figure XX: Volcano plot of metabolites, with the log2 fold-change difference between the cyclone and anticyclone shown along the x-axis and the log10-scaled p-value (Student's t-test followed by FDR correction) on the y-axis.
-Compounds that have the largest biological differences are found on the far left and far right sides of the graph while compounds with the largest statistical differences are found at the top of the plot.
-The dotted line denotes an alpha level of 0.05.*
-
-```{r check different unknowns, eval=FALSE}
-feature_msdata <- most_stat_diff %>%
-  left_join(MC_feats, by="feature") %>%
-  filter(is.na(compound_name)) %>%
-  select(feature, mzmed, rtmed, calc_formula) %>%
-  # filter(feature=="MC_POS_FT410")
-  slice(1)
-
-# msms_files <- list.files("../mzMLs/pos/MSMS", pattern = "180205.*dda", full.names = TRUE)
-# msmsdata <- grabMSdata(msms_files)
-
-mz_i <- feature_msdata$mzmed
-ggplot() +
-  geom_line(aes(x=rt, y=int, group=filename),
-            data = msmsdata$MS1[mz%between%pmppm(mz_i)]) +
-  geom_vline(aes(xintercept=rt), color="steelblue",
-             data = msmsdata$MS2[premz%between%pmppm(mz_i)] %>% distinct(rt)) +
-  geom_vline(xintercept = feature_msdata$rtmed/60+c(-0.5, 0.5), color="red")
-msmsdata$MS2[premz%between%pmppm(mz_i)] %>%
-  filter(!fragmz%between%pmppm(mz_i, 10)) %>%
-  filter(rt%between%(feature_msdata$rtmed/60+c(-0.5, 0.5))) %>%
-  ggplot(aes(x=fragmz, y=int)) +
-  geom_segment(aes(xend=fragmz, yend=0)) +
-  geom_point()
-
-mgf_maker <- function(feature_msdata, ms1, ms2, output_file){
-  ms1_to_write <- paste(
-    "BEGIN IONS", paste0("PEPMASS=", feature_msdata$mzmed),
-    "MSLEVEL=1", "CHARGE=1+", 
-    paste0(apply(ms1, 1, paste, collapse=" "), collapse = "\n"),
-    "END IONS", "", 
-    sep = "\n")
-  ms2_to_write <- lapply(unique(ms2$filename), function(volt_i){
-    paste0(c("BEGIN IONS", paste0("PEPMASS=", feature_msdata$mzmed),
-           "MSLEVEL=2", "CHARGE=1+",
-           paste0(apply(ms2[ms2$filename==volt_i, c("fragmz", "int")], 
-                 1, paste, collapse=" "), collapse="\n"),
-           "END IONS",
-           ""), collapse="\n")
-  })
-  outtext <- paste0(ms1_to_write, "\n", paste0(unlist(ms2_to_write), collapse = "\n"))
-  writeLines(outtext, con = output_file)
-}
-
-ms1 <- matrix(c(feature_msdata$mzmed, feature_msdata$avg_area), ncol = 2)
-ms2 <- msmsdata$MS2[premz%between%pmppm(mz_i)] %>%
-  filter(rt%between%c(7, 8))
-mgf_maker(feature_msdata, ms1, ms2, output_file = "~/../Desktop/temp.mgf")
-```
-
-We also did some preliminary searches for the identity of the low p-value compounds that were enriched in the anticyclone using CSI:FingerID software.
-MC_POS_FT410 (*m/z* 179.0047, retention time 7.3 minutes) had a fold-change enrichment of 4x in the anticyclonic eddy and we annotated it as arsenobetaine after purchasing the standard which was run at a later time and matched in MS^1^ and MS^2^ *m/z* values at a confidence level very close to 1.
-MC_POS_FT328 (*m/z* 160.0968, retention time 7.8 minutes, putative formula $C_7H_{14}NO_3$), another compound that separated completely between the two eddies and had a p-value of approximately 10^-7^ shares an exact mass with betonicine/turicine but is not likely to be this compound, as we later purchased this standard and its elution time corresponds to a separate peak in the same chromatogram.
-Instead, CSI:FingerID and CANOPUS provided top hits of acetylvaline and 5-acetamidopentanoate with a 90% posterior probability of containing an N-acyl amino acid group fora confidence of approximately 3.
-MC_POS_FT400 (*m/z* 176.1281, retention time 8.1 minutes, putative formula $C_8H_{17}NO_3$, confidence level 3) appears to be some unidentified betaine, with a 97+% chance of belonging to the quaternary ammonium salts class, and may in fact represent a methylated version of the MC_POS_FT328 described just above.
-
-### How robust are these results?
-
-We tested the robustness of our results by returning to the same approximate location one year later and again sampled two eddies of opposite polarities.
-Each eddy was sampled at 25 meters and the deep chlorophyll maximum (~105 meters in the cyclonic eddy, ~122 meters in the anticyclone) and once at 6 AM local time and again at 6 PM local time.
-This time, only nutrient data (nitrate + nitrite and soluble reactive phosphorus), particulate data (PC and PN), and flow cytometry data was collected.
-
-```{r metadata boxplots}
-filled_FK_metadata <- filled_file_metadata %>%
-  filter(cruise=="FK") %>%
-  filter(samp_type=="Smp")
-
-distinct_FK_metadata <- filled_FK_metadata %>%
-  mutate(basename=str_remove(untripl, "190715_Smp_")) %>%
-  mutate(cyc_type=ifelse(station%in%c("62", "64"), "cyc", "anti")) %>%
-  select(basename, cyc_type, depth, sla, ends_with("abund"), ends_with("um")) %>%
-  select(-oxy_um, -si_um, -DON_um, -DOP_um) %>%
-  distinct()
-
-# If we decide to use a figure (boxplots of values by depth)
-distinct_FK_metadata %>%
-  mutate(depth=factor(depth, levels = c("DCM", "25m"))) %>%
-  select(basename, cyc_type, depth, sla, ends_with("abund"), ends_with("um")) %>%
-  pivot_longer(-c(basename, cyc_type, depth, sla)) %>%
-  mutate(name=factor(name, levels = factor_levels, labels = twoline_labels)) %>%
-  filter(!is.na(value)) %>%
-  ggplot() +
-  # geom_boxplot(aes(x=value, y=depth)) +
-  geom_jitter(aes(x=value, y=depth, color=cyc_type), height = 0.1, width = 0) +
-  facet_wrap(~name, scales="free_x", nrow=2) +
-  theme(axis.title = element_blank(), legend.position = "none")
-```
-
-```{r metadata table}
-# If we decide to use a table
-distinct_FK_metadata %>%
-  select(basename, cyc_type, depth, sla, ends_with("abund"), ends_with("um")) %>%
-  pivot_longer(-c(basename, cyc_type, depth, sla)) %>%
-  group_by(depth, name) %>%
-  summarize(print_col=paste(
-    signif(mean(value, na.rm=TRUE), digits = 3), "±", 
-    signif(sd(value, na.rm=TRUE), digits = 1)
-  ), .groups = "drop") %>%
-  # Pivot into nice format
-  pivot_wider(names_from="depth", values_from = print_col) %>%
-  mutate(name=factor(name, levels = factor_levels, labels = factor_labels)) %>%
-  arrange(name)
-```
-
-```{r welch ANOVA FK}
-welch_aov <- distinct_FK_metadata %>%
-  select(basename, depth, sla, ends_with("abund"),
-         ends_with("um"), ends_with("ug")) %>%
-  pivot_longer(-c(basename, depth, sla)) %>%
-  # Perform the ANOVA
-  nest(data=-name) %>%
-  mutate(aovtest=map(data, ~broom::tidy(oneway.test(.x$value~.x$depth)))) %>%
-  suppressMessages() %>%
-  unnest(aovtest) %>%
-  # Format p-values nicely
-  mutate(p.value=p.adjust(p.value, method = "holm"))
-```
-
-The environmental data show many similar trends to the previous year and are well within the ranges documented elsewhere.
-Nitrate + nitrite and phosphate are both lower in the anticyclonic eddy than the cyclonic one, and PC and PN are both generally lower as well with a much larger difference at the DCM.
-Notably, heterotrophic bacteria and *Synechococcus* were consistently more abundant in the anticyclonic eddy.
-
-```{r FK v MT v MC distance correlation}
-nested_peaks <- all_feats %>%
-  group_by(group_feat) %>%
-  filter(n()==3) %>%
-  ungroup(group_feat) %>%
-  left_join(all_peaks, by="feature") %>%
-  left_join(filled_file_metadata, by = c("cruise", "filename")) %>%
-  filter(samp_type=="Smp") %>%
-  group_by(cruise, feature) %>%
-  mutate(norm_area=norm_area/max(norm_area)) %>%
-  ungroup()
-
-fk_dist <- nested_peaks %>%
-  filter(cruise=="FK") %>%
-  select(feature, filename, norm_area) %>%
-  pivot_wider(names_from = filename, values_from = norm_area) %>%
-  column_to_rownames("feature") %>%
-  data.matrix() %>%
-  dist()
-mt_dist <- nested_peaks %>%
-  filter(cruise=="MT") %>%
-  select(feature, filename, norm_area) %>%
-  pivot_wider(names_from = filename, values_from = norm_area) %>%
-  column_to_rownames("feature") %>%
-  data.matrix() %>%
-  dist()
-mc_dist <- nested_peaks %>%
-  filter(cruise=="MC") %>%
-  select(feature, filename, norm_area) %>%
-  pivot_wider(names_from = filename, values_from = norm_area) %>%
-  column_to_rownames("feature") %>%
-  data.matrix() %>%
-  dist()
-
-dist_mat <- cbind(as.numeric(fk_dist), as.numeric(mc_dist), as.numeric(mt_dist))
-colnames(dist_mat) <- c("FK", "MC", "MT")
-cor_mat <- round(cor(dist_mat), 3)
-```
-
-When we calculated the similarity of the three separate metabolomes, we obtained correlation coefficients of `r cor_mat["MC", "MT"]` for the eddy center and eddy transect data, `r cor_mat["MC", "FK"]` for the eddy center and Falkor data, and `r cor_mat["MT", "FK"]` for the transect and Falkor data.
-
-```{r SLA p-value crosscruise comparison, eval=FALSE}
-FK_sla_diff <- all_peaks %>%
-  left_join(all_feats, by = "feature") %>%
-  left_join(filled_file_metadata, by = c("filename", "cruise")) %>%
-  mutate(eddy=ifelse(station%in%c("62", "64"), "cyc", "anti")) %>%
-  filter(samp_type=="Smp") %>%
-  filter(cruise=="FK") %>%
-  nest(data=-c(group_feat, feature)) %>%
-  # mutate(pval=map_dbl(data, ~broom::tidy(t.test(.x$norm_area~.x$eddy))$p.value)) %>%
-  mutate(pval=map_dbl(data, ~broom::tidy(lmp(.x$norm_area~.x$eddy, seqs=FALSE, settings=FALSE))$statistic[2])) %>%
-  select(group_feat, feature, pval) %>%
-  mutate(pval=p.adjust(pval, method = "fdr"))
-
-MT_sla_diff <- all_peaks %>%
-  left_join(all_feats, by = "feature") %>%
-  left_join(filled_file_metadata, by = c("filename", "cruise")) %>%
-  filter(samp_type=="Smp") %>%
-  filter(cruise=="MT") %>%
-  nest(data=-c(group_feat, feature)) %>%
-  mutate(pval=map_dbl(data, ~broom::tidy(lmp(.x$norm_area~.x$sla, seqs=FALSE, settings=FALSE))$statistic[2])) %>%
-  select(group_feat, feature, pval) %>%
-  mutate(pval=p.adjust(pval, method = "fdr")) %>%
-  filter(pval<0.05)
-
-MC_sla_diff <- all_peaks %>%
-  left_join(all_feats, by = "feature") %>%
-  left_join(filled_file_metadata, by = c("filename", "cruise")) %>%
-  filter(samp_type=="Smp") %>%
-  filter(cruise=="MC") %>%
-  nest(data=-c(group_feat, feature)) %>%
-  mutate(pval=map_dbl(data, ~broom::tidy(lmp(.x$norm_area~.x$station, seqs=FALSE, settings=FALSE))$statistic[2])) %>%
-  select(group_feat, feature, pval) %>%
-  mutate(pval=p.adjust(pval, method = "fdr")) %>%
-  filter(pval<0.05)
-
-all_feats %>%
-  filter(group_feat%in%MT_sla_diff$group_feat) %>%
-  filter(group_feat%in%MC_sla_diff$group_feat) %>%
-  filter(group_feat%in%(FK_sla_diff %>% filter(pval<0.05) %>% pull(group_feat))) %>%
-  select(group_feat, feature, mzmed, rtmed, compound_name, formula) %>%
-  arrange(mzmed)
-
-all_feats %>%
-  filter(group_feat%in%MT_sla_diff$group_feat) %>%
-  filter(group_feat%in%MC_sla_diff$group_feat) %>%
-  left_join(FK_sla_diff %>% select(group_feat, pval), by="group_feat") %>%
-  select(group_feat, feature, mzmed, rtmed, compound_name, formula, pval) %>%
-  arrange(mzmed)
-```
-
-Although we did not find any compounds that differed significantly across all three data sets, we found several shared compounds between the eddy transect and eddy center samples.
-Sarcosine, inosine, nicotinic acid, and isethionic acid were all more abundant in the cyclone than the anticyclone, as well as seven unknown compounds.
-Of these, 4 were unfound in the Falkor data set, while the other 7 were simply not different (p-values ranging from 0.13 - 0.62).
-
-## Discussion
-
-### Environmental data
-
-The results obtained in this study from the environmental data are generally in good agreement with those previously described (e.g.
-@Barone2022, @HenderikxFreitas2020, and @harke2021) despite the different statistical methodology and the use here of interpolated values instead of raw data.
-
-### Overall metabolite response
-
-It is relatively unsurprising that we're able to separate the samples by depth given the results from the environmental data.
-Plotting the metabolome along NMDS axes 1 and 3 creates clear visual clusters corresponding to each depth (Fig.
-XX).
-NMDS axis 2, however, represents an unknown source of variation.
-Many of the triplicate triangles separate along this axis and often span zero.
-Due to the variance between triplicates here, it is unlikely that this variation is biological or environmental and instead may be accounting for differences in sample processing, extraction, or injection despite most compounds being normalized to a best-matched internal standard.
-
-The SLA-driven separation within the 175 meter samples is also quite informative.
-The deep samples from stations with a larger (anticyclonic) SLA all fall closer to the shallow samples along NMDS axis 1 than those from stations with a negative (cyclonic) SLA.
-This result makes sense given the context of anticyclones pushing surface water downward upon formation, deepening pycnoclines and causing the metabolome at 175m to look more like a metabolome sampled at a shallower depth.
-This trend is not visually present in either the DCM or 15m samples, however, indicating that the effect of SLA may actually be strongest at 175 meters overall.
-There was no clear separation by time of day in the NMDS plots that we could detect visually which was a little surprising given prior research showing the complex diel cycling of metabolites over the course of the day.
-
-The PERMANOVA results validate this subjective assessment with a very large R^2^ value (~30% variance explained) for depth, while time of day and eddy polarity had higher p-values (still significant at an alpha-level of 0.05) and lower R^2^ values, in total accounting for approximately 40% of the variance in the metabolome.
-A significant three-way interaction between depth, time, and SLA complicates this interpretation slightly, with each parameter potentially depending on the interaction of the other two, but with a low R^2^ and marginal p-value the interaction can probably be ignored for an exploratory analysis like this.
-The remaining ~60% of the variance in the data could be explained by true biological variation in the original seawater, differential treatment during the sample processing steps, an unexplored confounding variable, or most likely some combination of the above.
-
-### Metabolite response to depth
-
-The distinctly nonhomogenous distribution of the metabolome in Figure XX emphasizes the diversity of metabolite responses to depth.
-Rather than displaying a single, universal trend, compounds vary wildly in their abundance throughout the water column.
-Many appear to track biomass, correlating well with PC, PN and particle abundance with maximum concentrations somewhere between the surface and the DCM.
-Others have more complicated responses, whether found exclusively at the surface or *increasing* in concentration with depth, that illustrate the nuance missed by bulk particle measurements.
-Our conservative use of rank-order statistics here intentionally sacrifices some experimental power but guarantees that the signal observed raw peak-area space or z-scoring space would be at least as strong as the separations observed here.
-Compounds found mostly in the surface layer are synthesized in excess of biomass in the surface ocean [mixed layer], where light is highly abundant but nutrients are scarce.
-The presence of MAAs in this category provides one possible reason for excess synthesis in the surface ocean, as these compounds are known to be synthesized under UV light and act as "sunscreen" compounds in the environment [ref].
-At the deep chlorophyll maximum, we found several compounds again synthesized in excess of biomass - most obviously chlorophyll, but also several nitrogen-rich compounds like amino acids, osmolytes, and the nucleobase guanine.
-Despite this apparent trend, our followup analysis revealed no strong correlation between depth category and C:N content.
-[Anitra, can you talk about or add citations about a lot of these compounds seeming to come from euks?
-Would be a nice connection back to the metadata where picoeuks are found almost exclusively at the DCM].
-The absence of compounds that have a minimum at the DCM is relatively unsurprising.
-To have a DCM minimum requires either extensive drawdown at the DCM (which might occur for precursor molecules of DCM maximum compounds) or synthesis at the surface and depth which does not occur at the DCM itself (perhaps some general stress compound).
-Finally, compounds (both known and unknown) that did not respond to depth may do so for several reasons.
-First, it is possible but unlikely that a compound is being synthesized at a rate exactly inversely proportional to the decrease in biomass with depth - this would cause the observed concentration to be fixed across depths.
-Similarly, highly recalcitrant compounds could degrade at a rate too slow to be detected over the ~150 meters that we sampled across.
-Alternatively, these points could represent compounds that are highly sensitive to the sampling, extraction, or measurement process - because this classification scheme relies on the variance within each depth relative to the variance between depths, some highly sensitive compounds may have much higher noise levels and the depth signal is simply not visible.
-Our decision to use the nonparametric statistics here even when some data may be normally distributed exacerbates this problem, protecting strongly against false positives in the case of non-normality but undoubtedly causing some truly depth-sensitive compounds to be observed as uniform in abundance with depth.
-Finally, some of these compounds could be contaminants introduced during the sampling and measurement process but not all, because we know the identities of several molecules that belong to this category and do not expect them to be contaminants introduced by tubing, filters, or solvents.
-Most likely is that this null hypothesis category consists of a combination of the above, but we were not able to distinguish those potential explanations here.
-
-### Metabolite response to SLA, both alone and compared to depth
-
-Many of the trends with SLA observed here can be explained as a function of biomass.
-At the DCM, biomass is high in the cyclone as nutrient-rich water is upwelled into the euphotic zone and lower in the anticyclone.
-Conversely, in the 175 meter samples biomass is depleted from the cyclone and instead enriched at the DCM (?
-ask benedetto to explain this again).
-As expected, the response to SLA was weakest within the mixed layer at the surface where processes other than eddy upwelling/downwelling dominate.
-Despite this, we find slightly higher nitrate and nitrite concentrations in the anticyclone, as well as DON and DOP, so it's not completely unsurprising that some metabolites follow this trend.
-The increasing effect of SLA on metabolites and environmental variables is equally unsurprising, as density surfaces are displaced more at depth, though it's unclear how deep this effect is likely to propagate and future sampling at depths below 175 meters could improve our understanding of the depth at which eddies cease to affect the metabolome because that cannot be concluded from this study.
-Given that all of these samples were from the particulate fraction, it's also unsurprising that most metabolites track with PC/PN at the DCM and 175 meters.
-Conversely, those few metabolites that are either independent of PC/PN or negatively correlated with these variables deserve future study to determine whether they are real biological signal and, if so, what the biological functions they correspond to.
-
-We used a very conservative statistical method to compare the variance introduced by ~40 centimeters of SLA to the variance introduced by sampling across 50 meters of the water column and found a surprising number of compounds for which the difference between the two eddies was larger than the difference with depth.
-For the `r sum(pval_cat_counts["Highly signif"], pval_cat_counts["Signif"])` compounds that were statistically different despite not stratifying by depth in any way, it is clear that SLA is at least as important as sampling depth around the DCM and the care that is currently taken to sample at exact depths using the CTD should be the same care taken to account for differences in SLA.
-We expected from the earlier analysis on the eddy transect to find that most compounds were enriched in the cyclone, but when we used the two endpoints established at the center of each eddy we found significantly more compounds enriched in the anticyclone.
-This may be due to the increased power of the t-tests (compared to the ranked regressions run above) or due to the increased difference expected from sampling the endpoints alone.
-
-### Replicating/robustness of results
-
-We expected to find a strong correlation between the metabolome measured during the eddy transect and the eddy centers, as these two data sets sampled the same water mass, albeit separated by a few days.
-We expected a lower correlation coefficient for comparisons to the Falkor data sets, as that data was collected a full year after the original eddy center and eddy transect.
-Surprisingly, we learned that the two data sets that sampled the centers of an eddy (the eddy center and Falkor data sets).
-The eddy transect metabolome was actually the odd one out, with correlation coefficients around 0.45 compared to the (quite a bit higher) 0.55 obtained when comparing the two sets of eddy center samples.
-This could imply that the areas outside of an eddy contain a large number of relatively unique compounds, or could be due to differences like time of day that was controlled for the eddy center and Falkor samples but not for the eddy transect.
-
-The lack of any specific metabolite differing with SLA across all three data sets is more troubling.
-We would expect to find several such metabolites given the low p-values presented individually in each data set, but even at higher alpha values there was very little correlation between the p-value observed and its likelihood of differing with respect to SLA in the Falkor data.
-This may be due to the smaller sample size of the Falkor data set - with only 24 samples, the analysis is undoubtedly underpowered to detect the small differences observed in the eddy transect data with a sample size of 99.
-More powerful statistics such as multilevel modeling may also be able to better manage within-sample variation but also often require strong priors for the mechanisms by which compounds interact with each other and the environment or priors that are simply unavailable for these understudied molecules.
-That some compounds reproduced between the transect and eddy center data is heartening, but we are at this point unable to claim that we have discovered a particular biomarker for eddy state.
-
-## Conclusions
-
-TBD

### 2023-06-21 metadata/metadata_control.Rmd
Switched isotopes to correct option

@@
 clean_stans <- "https://raw.githubusercontent.com/IngallsLabUW/" %>%
-  paste0("Ingalls_Standards/master/Ingalls_Lab_Standards.csv") %>%
+  paste0("Ingalls_Standards/8fb6679505f692ec7d1cb4bf54ff1c5d4967c9a6/") %>%
+  paste0("Ingalls_Lab_Standards.csv") %>%
   read_csv() %>%

### 2023-06-26 metadata/metadata_control.Rmd
Adding file timestamp back to file metadata

@@
 
+if(!file.exists("raw_data/timestamp_filedata.csv")){
+  timestamp_filedata <- list.files("../mzMLs", pattern = "mzML", recursive = TRUE, full.names = TRUE) %>%
+    RaMS::grabMSdata(grab_what="metadata") %>%
+    pluck("metadata") %>%
+    distinct(filename, timestamp, polarity) %>%
+    mutate(polarity=str_extract(polarity, "..."))
+
+  pooled_patches <- timestamp_filedata %>%
+    filter(str_detect(filename, "180821_Poo")) %>%
+    distinct(filename, timestamp) %>%
+    drop_na()
+  timestamp_filedata <- timestamp_filedata %>%
+    rows_patch(pooled_patches)
+  write_csv(timestamp_filedata, "raw_data/timestamp_filedata.csv")
+} else {
+  timestamp_filedata <- read_csv("raw_data/timestamp_filedata.csv")
+}
+
 complete_file_data <- bind_rows(samp_filedata, extra_filedata) %>%
+  left_join(timestamp_filedata, by=c("filename", "polarity")) %>%
   left_join(samp_log)
+extra_filedata <- extra_filedata %>%
+  left_join(timestamp_filedata)
 ```

### 2023-06-27 metadata/metadata_control.Rmd
Adding MS run as a metadata column

@@
   bind_rows(extra_filedata) %>%
+  mutate(ms_run=cruise) %>%
   mutate(cruise=ifelse(cruise%in%c("MC", "MT"), "MS", cruise))

### 2023-08-29 targeted/targeted_control.Rmd
Complete rerun from MW file download

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = TRUE, purl = TRUE)
 options("readr.num_columns" = 0)
@@
   select(new=compound_name, old=name_old) %>%
+  distinct()
+date_filter <- clean_stans %>%
+  select(compound_name=name_old, date_added) %>%
   distinct()
@@
 ```{r falkorRFs}
-falkor_rundate <- strptime("190715", "%y%m%d")
+falkor_rundate <- 190715
 falkor_stans <- clean_stans %>%
@@
                            str_replace(filename, "Poo", "QC"),
-                           filename))
+                           filename)) %>%
+  anti_join(date_filter %>% filter(date_added>180821))
 
@@
   filter(compound_name!="Methyl 3-methylthiopropionate") %>%
+  filter(compound_name!="N(e)-Acetyl-Lysine") %>%
+  filter(compound_name!="Valine") %>%
   mutate(compound_name=str_replace(compound_name, "Monophosphate, 15N5", "monophosphate, 15N5")) %>%
@@
   anti_join(distinct(., compound_name, polarity) %>% filter(duplicated(compound_name)))
+
+# Fix sample 180821_Smp_MS6C315m_C which didn't get extraction standards added
+ext_stans_list <- c(
+  "Isethionic acid, 13C2",
+  "L-Cysteic acid, 2H3",
+  "Sulfoacetic acid, 13C2",
+  "3-Sulfolactate, 13C3",
+  "Taurine, 2H4"
+)
+update_df <- clean_ints_transect %>%
+  filter(str_detect(filename, "Smp")) %>%
+  filter(polarity=="neg") %>%
+  filter(compound_name%in%ext_stans_list) %>%
+  group_by(compound_name) %>%
+  summarise(area=median(area)) %>%
+  mutate(filename="180821_Smp_MS6C315m_C")
+clean_ints_transect <- clean_ints_transect %>%
+  rows_update(update_df, by = c("compound_name", "filename"))
 ```
@@
 transect_stans <- clean_stans %>%
-  filter(date_added<=transect_rundate)
+  filter(strptime(date_added, "%y%m%d")<=transect_rundate)
 
 rfs_transect <- transect_stans %>%
-  left_join(clean_ints_transect, by="compound_name") %>%
+  left_join(clean_ints_transect, by=c("compound_name", "polarity")) %>%
   filter(str_detect(filename, "_Std_")) %>% 
@@
 
+```{r transect fix sucrose/trehalose integrations, eval=FALSE}
+# Sucrose and trehalose got integrated together which is obnoxious because we 
+# have IS for each of those so it'd be really nice to get those good numbers
+# Re-integrating using RaMS
+
+library(RaMS)
+# msdata <- list.files("../tmzMLs/neg", full.names = TRUE, 
+#                      pattern = "170706|180821|180205") %>% 
+#   grabMSdata(verbosity = 0)
+# tresuc_msdata <- msdata$MS1[mz%between%pmppm(341.10839, 10)]
+# IS_msdata <- msdata$MS1[mz%between%pmppm(353.1497, 10)]
+
+suc_data <- tresuc_msdata[rt%between%c(11.2, 12.1)] %>%
+  group_by(filename) %>%
+  summarize(area=sum(int))
+suc_IS_data <- IS_msdata[rt%between%c(11.2, 12.1)] %>%
+  group_by(filename) %>%
+  summarize(IS_area=sum(int))
+# ggplot(suc_IS_data) + geom_col(aes(x=filename, y=IS_area))
+suc_nM <- left_join(suc_data, suc_IS_data) %>%
+  mutate(nM=area/IS_area*5000 * 0.0004/10 * 2) %>%
+  mutate(compound_name="Sucrose")
+
+tre_data <- tresuc_msdata[rt%between%c(12.1, 12.9)] %>%
+  group_by(filename) %>%
+  summarize(area=sum(int))
+tre_IS_data <- IS_msdata[rt%between%c(12.1, 12.9)] %>%
+  group_by(filename) %>%
+  summarize(IS_area=sum(int))
+# ggplot(suc_IS_data) + geom_col(aes(x=filename, y=IS_area))
+tre_nM <- left_join(tre_data, tre_IS_data) %>%
+  mutate(nM=area/IS_area*5000 * 0.0004/10 * 2) %>%
+  mutate(compound_name="Trehalose")
+
+update_df <- bind_rows(tre_nM, suc_nM) %>%
+  select(compound_name, nM, filename) %>%
+  mutate(polarity="neg") %>%
+  mutate(filename=str_remove(filename, ".tmzML"))
+mesotransect_data <- mesotransect_data %>%
+  rows_upsert(update_df, by=c("compound_name", "nM", "filename", "polarity"))
+```
+
 # Mesocenter things
@@
   select(-integrated, -in_samps) %>% 
-  mutate(area=as.numeric(area))
+  mutate(area=as.numeric(area)) %>%
+  filter(!compound_name%in%c("Turicine", "Hydroxyisoleucine"))
 ```
@@
 # Slightly weird because Mesocenter data doesn't have H2OinMatrix samples
-# Instead, use rf_ratio from Falkor to swap in
 center_rundate <- strptime("180205", "%y%m%d")
 center_stans <- clean_stans %>%
-  filter(date_added<=center_rundate)
+  filter(strptime(date_added, "%y%m%d")<=center_rundate)
+
+custom_h2o_mat <- clean_ints_center %>%
+  filter(str_detect(filename, "Poo_Full")) %>%
+  group_by(compound_name) %>%
+  summarise(newH2OInMatrix=mean(area)*(1-(62.94*100/80)/100))
 
 rfs_center <- center_stans %>% 
-  left_join(clean_ints_center, by="compound_name") %>% 
+  left_join(clean_ints_center, by=c("compound_name", "polarity")) %>%
   filter(str_detect(filename, "_Std_")) %>%
@@
   pivot_wider(names_from = type, values_from=mean_area) %>%
+  left_join(custom_h2o_mat) %>%
+  # left_join(rfs_falkor) %>%
   mutate(rf_h2o=InH2O/conc_um) %>%
-  left_join(rfs_falkor %>% select("rf_ratio", "compound_name"), by="compound_name")
+  mutate(rf_matrix=(InMatrix - newH2OInMatrix)/conc_um) %>%
+  mutate(rf_ratio=rf_matrix/rf_h2o)
 
@@
 ) %>%
-  left_join(clean_stans, by="compound_name")
+  left_join(clean_stans, by=c("compound_name", "polarity"))
 write.csv(raw_area_data, file = "made_data/raw_area_data.csv", row.names = FALSE)
@@
 quant_data <- bind_rows(falkor_data, mesotransect_data, mesocenter_data) %>%
-  filter(compound_name!="Choline")
+  filter(compound_name!="Choline") %>%
+  mutate(nM=ifelse(compound_name=="Homarine", nM/25, nM))
 write.csv(quant_data, file = "made_data/longdata.csv", row.names = FALSE)

### 2023-08-29 manuscript/manuscript.Rmd
All-new results section completed

@@
 ---
-title: "AllMeso manuscript"
+title: "MESOSCOPE manuscript results/figures"
 author: "William Kumler"
 date: "`r Sys.Date()`"
-output: html_document
+output: 
+  word_document:
+    reference_docx: template.docx
 ---
 
-```{r Run analysis}
+```{r Run analysis, include=FALSE}
+if(!file.exists("../mzMLs/neg/180205_Poo_TruePoo_Full1.mzML")){
+  init_timeout <- options("timeout")$timeout
+  options(timeout=3600) # Allow 1 hour for each FTP download
+
+  download.file("ftp://www.metabolomicsworkbench.org/Studies/ST002788_HILIC_NEG.zip",
+                destfile = "../mzMLs/ST002788_HILIC_NEG.zip")
+  unzip("../mzMLs/ST002788_HILIC_NEG.zip", exdir = "../mzMLs/neg")
+  file.remove("../mzMLs/ST002788_HILIC_NEG.zip")
+  download.file("ftp://www.metabolomicsworkbench.org/Studies/ST002789_HILIC_NEG.zip",
+                destfile = "../mzMLs/ST002789_HILIC_NEG.zip")
+  unzip("../mzMLs/ST002789_HILIC_NEG.zip", exdir = "../mzMLs/neg")
+  file.remove("../mzMLs/ST002789_HILIC_NEG.zip")
+  download.file("ftp://www.metabolomicsworkbench.org/Studies/ST002788_HILIC_POS.zip",
+                destfile = "../mzMLs/ST002788_HILIC_POS.zip")
+  unzip("../mzMLs/ST002788_HILIC_POS.zip", exdir = "../mzMLs/pos")
+  file.remove("../mzMLs/ST002788_HILIC_POS.zip")
+  download.file("ftp://www.metabolomicsworkbench.org/Studies/ST002789_HILIC_POS.zip",
+                destfile = "../mzMLs/ST002789_HILIC_POS.zip")
+  stop("File extraction has to be done manually - file too large for R's unzip")
+  file.remove("../mzMLs/ST002789_HILIC_POS.zip")
+  options(timeout=init_timeout)
+}
+
 if(!file.exists("../metadata/filled_file_metadata.csv")){
@@
 }
+
+if(!file.exists("../targeted/all_peaks.csv")){
+  setwd("../targeted")
+  knitr::purl("../targeted/targeted_control.Rmd")
+  source("targeted_control.R")
+  file.remove("targeted_control.R")
+  setwd("../manuscript")
+}
+
 if(!file.exists("../untargeted/all_peaks.csv")){
   setwd("../untargeted")
+  knitr::purl("untargeted_control.Rmd")
   data_iters <- c("FK_pos", "FK_neg", "MS_pos", "MS_neg")
-  knitr::purl("untargeted_control.Rmd")
   for(cruise in c("FK", "MS")){
@@
   }
-  file.remove("untargeted_control.R")
+  file.remove("../untargeted/untargeted_control.R")
   map(data_iters, function(output_dir){
@@
 }
-if(!file.exists("../targeted/all_peaks.csv")){
-  setwd("../targeted")
-  knitr::purl("../targeted/targeted_control.Rmd")
-  source("targeted_control.R")
-  file.remove("targeted_control.R")
-  setwd("../manuscript")
+```
+
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, eval = TRUE)
+library(tidyverse)
+
+filled_file_metadata <- read_csv("../metadata/filled_file_metadata.csv") %>%
+  filter(polarity=="pos") %>%
+  select(-polarity)
+clean_stans <- read_csv("../metadata/clean_stans.csv")
+untarg_feats <- read_csv("../untargeted/all_feats.csv")
+untarg_peaks <- read_csv("../untargeted/all_peaks.csv") %>%
+  filter(str_detect(filename, "_Smp_"))
+targ_peaks <- read_csv("../targeted/all_peaks.csv") %>%
+    mutate(filename=paste0(filename, ".mzML"))
+
+metadata_levels <- c(
+    "no23_um", "po4_um", "si_um", "PC_um", "PN_um", "DON_um", "DOP_um", "oxy_um",
+    "chl_ug", "phaeo_ug", "alk", "dic", "ph", "theta",
+    paste0(c("pro", "syn", "euk", "het"), "_abund"), "beam_atten",
+    "abs_depth", "time", "lat", "sla", "local_hour", 
+    "TDN", "TDP", "temp", "pressure", "sal", "sigma",
+    "ctd_chl"
+  )
+metadata_labels <- c(
+    "Nitrate + nitrite", "Phosphate", "Silicate", "Particulate C", "Particulate N",
+    "Dissolved organic N", "Dissolved organic P", "Oxygen", "Chlorophyll a",
+    "Phaeophytin", "Alkalinity", "Dissolved inorganic C", "pH", "Density",
+    paste(c("Pro.", "Syn.", "Picoeuk.", "Het. bact."), "abundance"),
+    "Beam attenuation",
+    "Depth", "Time", "Latitude", "Sea level anomaly", "Local hour", 
+    "Total dissolved nitrogen", "Total dissolved phosphorus",
+    "Temperature", "Pressure", "Salinity", "Sigma-theta",
+    "Chlorophyll (fluor.)"
+  )
+metadata_twoline_labels <- c(
+    "Nitrate + nitrite", "Phosphate", "Silicate", "Particulate C", "Particulate N",
+    "Dissolved\norganic N", "Dissolved\norganic P", "Oxygen", "Chlorophyll a",
+    "Phaeophytin", "Alkalinity", "Dissolved\ninorganic C", "pH", "Density",
+    paste(c("Prochlorococcus", "Synechococcus", "Picoeukaryote", 
+            "Heterotrophic bact."), "\nabundance"),
+    "Beam attenuation",
+    "Depth", "Time", "Latitude", "Sea level anomaly", "Local hour",
+    "Total dissolved\nnitrogen", "Total dissolved\nphosphorus",
+    "Temperature", "Pressure", "Salinity", "Sigma-theta", "Chlorophyll\n(fluorescence)"
+  )
+metadata_short_labels <- c("N+N", "PO4", "Si", "PC", "PN", "DON", "DOP", "O2", "Chl a",
+                  "Phaeophytin", "Alk", "DIC", "pH", "Density", 
+                  paste(c("Pro", "Syn", "Picoeuk", "Het. bact."), "abund"),
+                  "Beam atten.", "Depth", "Time", "Latitude", "SLA", "Local hour",
+                  "Tot. N", "Tot. P", "Temp", "dB", "Sal", "σ-θ", "CTD chl.")
+
+splividis <- function(n, ...){
+  vir <- viridis::viridis(n, ...)
+  if(n%%2==0){
+    vir[as.numeric(matrix(1:n, ncol=2, byrow = TRUE))]
+  } else {
+    index_vec <- suppressWarnings(as.numeric(matrix(1:n, ncol=2, byrow = TRUE)))
+    vir[head(index_vec, -1)]
+  }
 }
@@
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
-library(tidyverse)
+### Metabolome changes significantly across the eddy dipole at DCM and 175m
 
-filled_file_metadata <- read_csv("../metadata/filled_file_metadata.csv")
-clean_stans <- read_csv("../metadata/clean_stans.csv")
-untarg_feats <- read_csv("../untargeted/all_feats.csv")
-untarg_peaks <- read_csv("../untargeted/all_peaks.csv")
-targ_peaks <- read_csv("../targeted/all_peaks.csv")
+We found that the metabolome differed between the cyclone and anticyclone samples, with multivariate statistics detecting the largest differences in the DCM and 175 meter samples while the surface was more homogenous (Figure 1).
+At the DCM, approximately 16% of the variation in the metabolome could be explained by SLA alone according to the PERMANOVA (p = 0.003) and SLA was strongly correlated with the first principal component (r = .615, % variance explained by PC1 = 27.2), indicating that this was one of the largest sources of variation in the dataset.
+Notably, the samples taken from the exact center of the cyclonic eddy were highly distinct and likely drove much of the variance explained.
+In the deep samples, 12% of the variance was explained by SLA (p = 0.008) according to the PERMANOVA and SLA was highly correlated with the second PC of the dataset (r = .697, % variance explained by PC2 = 15.4), though the first PC did not seem to have any visible pattern with metadata and appeared to capture largely variation between biological triplicates.
+At the surface, SLA trends are less evident with a larger p-value (0.087) and a lower R$^2$ (0.09) (Figure 1).
+
+To characterize this difference, we plotted an "average" metabolite per station at each depth by taking the median value of all the detected mass features after z-score normalization (Figure 1).
+As expected, the average metabolite collected at the exact center of the cyclonic eddy (Station 12) was much more abundant than in any sample at this same depth in the rest of the dipole samples.
+Similarly, the 175 meter samples showed the largest metabolite abundances in the anticyclone relative to other stations at this depth.
+This method also allowed us to see the remarkably sinusoidal response of the average metabolite across the stations.
+This response correlated negatively with SLA at the DCM (lower SLA = higher average values, $r_{DCM}= -.716$) but positively with SLA at the surface and at depth ($r_{surface} = .211$, $r_{deep} = .830$).
+
+  - Discussion: note that the "neither" samples all group together, implying eddy effect rather than location ("outside" the eddy is more similar to each other than to the insides of either one)
+  - Discussion: it's kinda crazy that the exact center of the eddy is so wildly different at the DCM from stations that are nearby but not
+
+```{r multivar_analysis object creation - nmds/pca/permanova}
+set.seed(123)
+multivar_analysis <- untarg_peaks %>%
+  left_join(filled_file_metadata %>% distinct(ms_run, station, filename, depth), 
+          by="filename") %>%
+  mutate(depth=factor(depth, levels=c("Surface", "DCM", "Deep"), 
+                      labels=c("Surface  (15m)", "DCM  (110-120m)", "Deep  (175m)"))) %>%
+  filter(ms_run=="MT") %>%
+  nest(data=-depth) %>%
+  mutate(wide_peaks=map(data, function(metab_df){
+    metab_df %>%
+      select(feature, filename, norm_area) %>%
+      group_by(feature) %>%
+      filter(sum(norm_area==0)<n()*0.9) %>%
+      mutate(norm_area=scale(norm_area)[,1]) %>%
+      pivot_wider(names_from = "feature", values_from = norm_area) %>%
+      column_to_rownames("filename") %>%
+      data.matrix()
+  })) %>%
+  mutate(nmds_output=map(wide_peaks, function(metab_mat){
+    env_vec <- pull(filled_file_metadata, sla_corr, filename)[rownames(metab_mat)]
+    vegan::metaMDS(metab_mat, k = 2, distance = "manhattan", trace = 0, 
+                   autotransform = FALSE, noshare = FALSE, wascores = FALSE) %>%
+    vegan::MDSrotate(env_vec)
+  })) %>%
+  mutate(nmds_stress=map_dbl(nmds_output, pluck, "stress")) %>%
+  mutate(perm_output=map(wide_peaks, function(metab_mat){
+    adon_expl <- data.frame(filename=rownames(metab_mat)) %>%
+      left_join(filled_file_metadata %>% select(filename, sla_class), by="filename")
+    vegan::adonis2(
+      metab_mat~sla_class, data = adon_expl, by = "margin", 
+      permutations = 999, method = "manhattan")
+  })) %>%
+  mutate(perm_r2=map_dbl(perm_output, pluck, "R2", 1)) %>%
+  mutate(perm_pval=map_dbl(perm_output, pluck, "Pr(>F)", 1)) %>%
+  mutate(pca_output=map(wide_peaks, prcomp)) %>%
+  mutate(nmds_stress=paste("s =", round(nmds_stress, digits = 3))) %>%
+  mutate(perm_r2=round(perm_r2, digits = 2)) %>%
+  mutate(perm_pval=round(perm_pval, digits = 4)) %>%
+  mutate(perm_label=paste0("R2 = ", perm_r2, "\np = ", perm_pval))
 ```
+
+```{r figure nmds and median metabolite across eddy transect}
+nmds_sla_gp <- multivar_analysis %>%
+  mutate(nmds_points=map(nmds_output, function(nmds_output_i){
+    nmds_output_i %>% pluck("points") %>% as.data.frame()
+  })) %>%
+  select(depth, nmds_points) %>%
+  unnest_longer(nmds_points, indices_to = "filename") %>%
+  unnest_wider(nmds_points) %>%
+  left_join(filled_file_metadata %>% distinct(filename, sla_corr)) %>%
+  mutate(MDS2=ifelse(str_detect(depth,"Deep"), MDS2*-1, MDS2)) %>%
+  ggplot() +
+  geom_vline(xintercept = 0) +
+  geom_hline(yintercept = 0) +
+  geom_point(aes(x=MDS1, y=MDS2, fill=sla_corr), color="black", size=4, pch=21) +
+  geom_label(x=-Inf, y=-Inf, aes(label=nmds_stress), hjust=0, vjust=0, 
+             data = multivar_analysis, label.r = unit(0, "in")) +
+  geom_label(x=-Inf, y=Inf, aes(label=perm_label), hjust=0, vjust=1,
+             data = multivar_analysis, label.r = unit(0, "in"), lineheight = .7) +
+  facet_wrap(~depth, nrow=1) +
+  scale_fill_gradient2(
+    low = scales::muted("blue"), high = scales::muted("red"),  mid = "grey90",
+    aesthetics = c("color", "fill"),
+  ) +
+  labs(fill="Sea level\nanomaly\n(cm)\n") +
+  theme_bw() +
+  coord_fixed()
+
+med_metab_df <- untarg_peaks %>%
+  left_join(filled_file_metadata %>% distinct(filename, ms_run, station, depth, sla_corr),
+            by="filename") %>%
+  filter(ms_run=="MT") %>%
+  group_by(depth, feature) %>%
+  filter(sum(norm_area==0)<n()*0.9) %>%
+  mutate(norm_area=scale(norm_area)[,1]) %>%
+  ungroup() %>%
+  group_by(filename, depth, station, sla_corr) %>%
+  summarize(avg_amt=mean(norm_area)) %>%
+  ungroup()
+# Calculations for correlation between grand median metabolite and SLA
+# med_metab_df %>%
+#   group_by(station, depth, sla_corr) %>%
+#   summarise(grand_avg_amt=mean(avg_amt)) %>%
+#   group_by(depth) %>%
+#   summarise(sim=cor(sla_corr, grand_avg_amt))
+med_metab_sla_gp <- med_metab_df %>%
+  mutate(station=factor(station, levels=as.character(4:14))) %>%
+  mutate(depth=factor(depth, levels=c("Surface", "DCM", "Deep"))) %>%
+  ggplot() +
+  geom_hline(yintercept = 0) +
+  geom_boxplot(aes(x=station, y=avg_amt, color=sla_corr), linewidth=0.8) +
+  facet_wrap(~depth, nrow=1) +
+  scale_color_gradient2(low = scales::muted("blue"), high = scales::muted("red"),
+                        mid="grey90", aesthetics = c("color", "fill"), guide = NULL) +
+  theme_bw() +
+  theme(strip.background = element_blank(), strip.text = element_blank()) +
+  labs(x="Station", y="Median metabolite z-score")
+
+library(egg)
+nolegend_plot <- ggarrange(nmds_sla_gp+theme(legend.position = "none"), 
+                           med_metab_sla_gp, ncol = 1, heights = c(0.7, 0.3), 
+                           draw = FALSE)
+library(cowplot)
+nmds_and_med_metab <- plot_grid(nolegend_plot, get_legend(nmds_sla_gp), 
+                                nrow=1, rel_widths = c(0.9, 0.1))
+ggsave("nmds_and_med_metab.png", plot = nmds_and_med_metab, device = "png", 
+       path = "figures", width = 6.5, height = 4.5, units = "in", dpi = 300,
+       bg="white")
+```
+
+![](figures/nmds_and_med_metab.png)
+
+*Figure 1: Plots of the metabolome across the eddy dipole, broken down by depth.
+The top row of plots are non-metric multidimensional scaling (NMDS) plots, in which points correspond to individual samples and have been colored by their corrected sea level height anomaly.
+NMDS stress values (s) have been reported in the bottom left corner, while PERMANOVA R$^2$ and p-values are reported in the top left.
+SLA trends are visible in the DCM and deep samples, with dark blue circles consistently discriminating from the dark red circles along the first multidimensional axis.
+The bottom row of plots show the direction and magnitude of this effect by plotting the grand mean of the normalized metabolite peak areas in the stations taken across the eddy dipole.*
+
+```{r supp figure/table prcomp vs SLA}
+pcdata <- multivar_analysis %>%
+  select(depth, pca_output) %>%
+  mutate(pca_x=map(pca_output, function(pca_output_i){
+    pca_output_i$x[,1:3] %>% as.data.frame()
+  })) %>%
+  select(depth, pca_x) %>%
+  unnest_longer(pca_x, indices_to = "filename") %>%
+  unnest_wider(pca_x) %>%
+  left_join(filled_file_metadata %>% distinct(filename, sla_corr)) %>%
+  pivot_longer(starts_with("PC"))
+pcvars <- multivar_analysis %>%
+  select(depth, pca_output) %>%
+  mutate(perc_expl=map(pca_output, function(pca_output_i){
+    var_expl <- pca_output_i$sdev^2
+    data.frame(perc_expl=var_expl[1:3]/sum(var_expl), name=paste0("PC", 1:3))
+  })) %>%
+  unnest(perc_expl) %>%
+  mutate(perc_label=paste0(round(perc_expl, digits = 3)*100, "%"))
+
+# Correlation between PCs and SLA at each depth
+# pcdata %>%
+#   nest(data=-c(depth, name)) %>%
+#   mutate(cor_output=map_dbl(data, ~abs(cor(.x$value,.x$sla_corr))))
+
+# Linear model significance testing and parameter report
+# pcdata %>%
+#   nest(data=-c(depth, name)) %>%
+#   mutate(lm_output=map(data, ~broom::tidy(lm(.x$value~.x$sla_corr)))) %>%
+#   unnest(lm_output) %>%
+#   filter(term==".x$sla_corr") %>%
+#   arrange(name, depth) %>%
+#   select(depth, name, p.value) %>%
+#   split(.$name)
+
+lm_gp <- ggplot(pcdata, aes(x=sla_corr, y=value)) +
+  geom_smooth(method="lm", formula=y~x, color="black") +
+  geom_point(aes(fill=sla_corr), pch=21, color="black") +
+  geom_label(data = pcvars, aes(label=perc_label),
+             hjust=1, vjust=1, x=Inf, y=Inf, label.r = unit(0, "in")) +
+  facet_grid(depth~name, switch = "y") +
+  scale_y_continuous(position = "right") +
+  scale_fill_gradient2(
+    low = scales::muted("blue"), high = scales::muted("red"),  mid = "grey90",
+    aesthetics = c("color", "fill"), guide = NULL
+  ) +
+  labs(x="Corrected SLA, in cm", y="Principal component value",
+       color="Corrected\nSLA (cm)") +
+  theme_bw()
+deep_pc1_barplot_gp <- pcdata %>%
+  left_join(filled_file_metadata %>% distinct(filename, station, tripl)) %>%
+  mutate(station=factor(station, levels=as.character(4:14),
+                        labels=paste0("St.", 4:14))) %>%
+  arrange(station) %>%
+  filter(name=="PC1" & str_detect(depth, "Deep")) %>%
+  ggplot() +
+  geom_hline(yintercept = 0) +
+  geom_col(aes(x=tripl, y=value)) +
+  facet_wrap(~station, nrow = 1, scales="free_x") +
+  scale_y_continuous(position = "right") +
+  labs(x="Triplicate", y="PC1", title = paste(
+    "    Variance among deep (175 meter) samples in PC1"
+  ))
+pca_sla_cor_gp <- cowplot::plot_grid(lm_gp, deep_pc1_barplot_gp, 
+                                     rel_heights = c(0.6, 0.4), 
+                                     ncol = 1, labels = "AUTO")
+ggsave("pca_sla_cor_gp.png", pca_sla_cor_gp, device = "png", path = "supplement",
+       width = 6.5, height = 6.5, units = "in", dpi = 300)
+
+```
+
+![](supplement/pca_sla_cor_gp.png)
+
+*Supplemental figure: Results of a principal component analysis performed separately on each depth of the eddy transect metabolome.
+A) Regressions of SLA against various principal components, broken down by depth.
+% variance explained by each PC is noted in the upper right corner of the plot.
+Colors have been added corresponding to the SLA value for clarity but are redundant with the x-axis values.
+Lines of best fit and the associated standard error have been added behind the points in black and translucent grey, respectively.
+Strong associations exist between PC1 and SLA in the DCM samples and PC2 and SLA in the deep samples.
+B) Principal component 1 values for the 175 meter samples at each station showing large variance between triplicates rather than any external factor.*
+
+A large part of this metabolic response to SLA is due to the increased biomass in the cyclone DCM and the parallel increase in the anticyclone deep samples.
+We used a subset of the prior dataset, only those metabolites that we were able to quantify, and measured their in-situ concentrations in the particulate matter (Figure 2).
+
+```{r figure stacked barplot MT eddy transect w PC and SLA}
+avg_targ_concs <- targ_peaks %>%
+  left_join(filled_file_metadata, by="filename") %>%
+  filter(ms_run=="MT") %>%
+  filter(samp_type=="Smp") %>%
+  group_by(compound_name) %>%
+  mutate(mean_cmpd_nm=mean(nM)) %>%
+  ungroup()
+base_targ_gp <- avg_targ_concs %>%
+  group_by(compound_name, untripl) %>%
+  summarise(nM=mean(nM), mean_cmpd_nm=unique(mean_cmpd_nm)) %>%
+  ungroup() %>%
+  arrange(desc(mean_cmpd_nm)) %>%
+  mutate(compound_name=ifelse(compound_name=="Dimethylsulfoniopropionate", "DMSP", compound_name)) %>%
+  mutate(compound_name=ifelse(
+    compound_name%in%head(unique(compound_name), 11), 
+    compound_name, "Other")) %>%
+  mutate(compound_name=fct_inorder(compound_name)) %>%
+  left_join(filled_file_metadata %>% distinct(untripl, depth, station)) %>%
+  group_by(depth, station, compound_name) %>%
+  summarise(nM=sum(nM), .groups = "keep") %>%
+  summarise(mean_nM=mean(nM)) %>%
+  mutate(depth=factor(depth, levels=c("Surface", "DCM", "Deep"), 
+                      labels=c("Surface  (15m)", "DCM  (110-120m)", "Deep  (175m)"))) %>%
+  mutate(station=factor(station, levels=as.character(4:14))) %>%
+  ggplot() +
+  geom_hline(yintercept = 0) +
+  geom_col(aes(x=station, y=mean_nM, fill=compound_name), width = 0.9) +
+  scale_fill_manual(values = c(hcl.colors(12)[c(1,3,5,7,9,11,2,4,6,8,10)], "grey50")) +
+  scale_y_continuous(expand = c(0.005, 0.1)) +
+  labs(x="Station", y="Mean nM metabolite concentration", fill=NULL) +
+  facet_grid(~depth, scales="free", space = "free") +
+  theme_bw() +
+  theme(legend.position = "bottom")
+
+sla_loess_points <- filled_file_metadata %>%
+  filter(samp_type=="Smp") %>%
+  filter(ms_run=="MT") %>%
+  distinct(station, sla_corr) %>%
+  mutate(station=factor(station, levels=as.character(4:14))) %>%
+  group_by(station) %>%
+  summarise(sla_corr=mean(sla_corr)) %>%
+  mutate(station=as.numeric(station)) %>%
+  loess(formula = sla_corr~station) %>%
+  predict(., newdata=data.frame(station=seq(1, 11, 0.01))) %>%
+  data.frame(station=seq(1, 11, 0.01), sla_corr=.)
+arrow_df <- data.frame(
+  cyc_type=c("Anticyclone\ncenter", "Cyclone\ncenter"),
+  x=c("6", "12"), xend=c("6", "12"),
+  y=c(5, 4), yend=c(7.5, 6.5),
+  color=c(scales::muted("red"), scales::muted("blue")),
+  depth=factor("Deep", levels=c("Surface", "DCM", "Deep"), 
+               labels=c("Surface  (15m)", "DCM  (110-120m)", "Deep  (175m)")),
+  lab_x=c("5", "11")
+)
+targ_gp_w_sla <- base_targ_gp +
+  ggnewscale::new_scale_fill() +
+  geom_tile(aes(x=station, y=-0.5, fill=sla_corr),
+            color=NA, width=0.11, height=0.5, data = sla_loess_points) +
+  scale_fill_gradient2(
+    low = scales::muted("blue"), high = scales::muted("red"),  mid = "grey90",
+    aesthetics = c("fill"), guide = "none"
+  ) +
+  geom_segment(aes(x=x, y=y, xend=xend, yend=yend, color=color), data=arrow_df,
+               arrow=arrow(ends = "first", length = unit(0.1, "npc"), type = "closed"),
+               linewidth=0, linejoin = "bevel") +
+  geom_segment(aes(x=x, y=y+0.5, xend=xend, yend=yend, color=color), data=arrow_df, linewidth=2) +
+  geom_text(aes(x=lab_x, y=yend+0.5, color=color, label=cyc_type), 
+            data=arrow_df, vjust=0, hjust=0, lineheight=0.8) +
+  scale_color_identity()
+
+PC_data <- filled_file_metadata %>%
+  filter(samp_type=="Smp") %>%
+  filter(ms_run=="MT") %>%
+  distinct(station, PC_um, depth) %>%
+  mutate(depth=factor(depth, levels=c("Surface", "DCM", "Deep"), 
+                      labels=c("Surface  (15m)", "DCM  (110-120m)", "Deep  (175m)"))) %>%
+  mutate(station=factor(station, levels=as.character(4:14)))
+targ_gp_w_sla_PC <- targ_gp_w_sla +
+  geom_point(aes(x=station, y=PC_um*5), data=PC_data, pch=21, fill="white", color="black") +
+  scale_y_continuous(sec.axis = sec_axis(~./5, name = "Particulate carbon (o, µM)"),
+                     expand = c(0.005, 0.1))
+
+ggsave("targ_gp_w_sla_PC.png", plot = targ_gp_w_sla_PC, device = "png", 
+       path = "figures", width = 6.5, height = 4.5, units = "in", dpi = 300)
+```
+
+![](figures/targ_gp_w_sla_PC.png)
+
+*Figure 2: Concentrations of known compounds across the eddy dipole, broken down by depth.
+Stacked bars correspond to known compounds named in the legend underneath, with compounds not in the top 11 by absolute concentration grouped into the "Other" category.
+DMSP = dimethylsulfoniopropionate.
+Sea level anomaly (SLA) is shown by the red-blue color scale underneath the barplot with red corresponding to a positive SLA (anticyclone) and blue corresponding to negative SLA (cyclone).
+The eddy centers (stations 6 and 12) have been labeled with arrows of the same color scheme in the final panel.
+Particulate carbon concentrations in µmol/L are shown as white dots and use the y-axis scale on the right side (200x the metabolite scale).*
+
+```{r avg targ conc values and numbers for text, eval=FALSE}
+avg_targ_concs %>%
+  group_by(filename) %>%
+  summarise(nM=sum(nM)) %>%
+  left_join(filled_file_metadata %>% distinct(filename, untripl, PC_um, depth)) %>%
+  group_by(untripl, PC_um, depth) %>%
+  summarise(nM=mean(nM), .groups = "drop") %>%
+  group_by(depth) %>%
+  summarise(PC_vs_nM=broom::tidy(cor.test(PC_um, nM)))
+  # summarise(PC_vs_nM=broom::tidy(cor.test(PC_um, nM, method = "spearman")))
+  # ggplot(aes(x=PC_um, y=nM, color=depth)) + geom_point() + geom_smooth(method="lm")
+avg_targ_concs %>%
+  mutate(compound_name=fct_inorder(compound_name)) %>%
+  mutate(depth=factor(depth, levels=c("Surface", "DCM", "Deep"))) %>%
+  group_by(depth, compound_name) %>%
+  summarise(nM=mean(nM)) %>%
+  arrange(compound_name) %>%
+  pivot_wider(names_from = depth, values_from = nM) %>%
+  write_csv("supplement/avg_targ_conc_by_depth.csv")
+avg_targ_concs %>%
+  mutate(depth=factor(depth, levels=rev(c("Surface", "DCM", "Deep")))) %>%
+  group_by(compound_name, depth) %>%
+  summarise(nM=mean(nM)) %>%
+  mutate(nM=log10(nM)) %>%
+  ggplot(aes(x=nM, y=depth, group=compound_name)) +
+  geom_path()
+avg_targ_concs %>%
+  mutate(depth=factor(depth, levels=rev(c("Surface", "DCM", "Deep")))) %>%
+  group_by(compound_name, depth) %>%
+  summarise(nM=mean(nM)) %>%
+  mutate(nM=rank(nM)) %>%
+  ungroup() %>%
+  pivot_wider(names_from = depth, values_from = nM) %>%
+  mutate(depth_class=paste(Surface, DCM, Deep)) %>%
+  count(depth_class)
+```
+
+We found that the overall concentration of known molecules was highest at the DCM of the cyclone, in line with our expectations and prior results.
+The most abundant quantified molecule overall was trimethylamine N-oxide (TMAO), representing an average concentration of 1.38 nM.
+At the surface, however, hydroxyisoleucine was the most concentrated metabolite at 1.53 nM due to its monotonic decrease with depth as opposed to a majority of the osmolytes (37/53) which reached maximum concentrations at the DCM instead (Figure 2).
+
+Bulk particulate carbon (PC) measurements correlated well with overall metabolite concentrations ($r= .554$, p-value < 0.001) but this was largely driven by the lower biomass and metabolite abundance at depth.
+When analyzed at each depth separately, the DCM ($r_{DCM}= .790$, p-value = 0.003) and 175 meter samples ($r_{Deep}= .554$, p-value = 0.077) had weaker but still good correlations, but no such association was detected at the surface ($r_{Surface}= -.125$, p-value = 0.714).
+
+```{r supp figure fraction of PC in metabolites}
+pc_comp_df <- targ_peaks %>%
+  filter(str_detect(filename, "180821_Smp")) %>%
+  left_join(clean_stans %>% select(compound_name, formula, polarity)) %>%
+  mutate(n_C=as.numeric(str_extract(formula, "(?<=C)\\d+"))) %>%
+  mutate(nM_C=n_C*nM) %>%
+  left_join(filled_file_metadata, by = join_by(filename)) %>%
+  filter(samp_type=="Smp") %>%
+  group_by(filename) %>%
+  summarize(total_nM=sum(nM_C), PC_um=unique(PC_um)) %>%
+  mutate(frac_in_metab=total_nM/1000/PC_um) %>%
+  left_join(filled_file_metadata, by = join_by(filename)) %>%
+  mutate(depth=str_extract(depth, "DCM|Deep|Surface")) %>%
+  mutate(depth=factor(depth, levels=c("Surface", "DCM", "Deep"))) %>%
+  mutate(station=factor(station, levels=c(as.character(4:14))))
+
+# Significance testing between depths and SLA classes for % PC in metab
+# pc_comp_df %>% aov(formula=frac_in_metab~depth) %>% TukeyHSD()
+# pc_comp_df %>% aov(formula=frac_in_metab~sla_class) %>% TukeyHSD()
+
+# Significance testing for differences in % in metabs as a function of SLA by depth
+# pc_comp_df %>%
+#   nest(data=-depth) %>%
+#   mutate(aov_out=map(data, ~broom::tidy(aov(.x$frac_in_metab~.x$sla_class)))) %>%
+#   mutate(kw_out=map(data, ~broom::tidy(kruskal.test(.x$frac_in_metab~.x$sla_class)))) %>%
+#   pivot_longer(c(aov_out, kw_out)) %>%
+#   unnest(value) %>%
+#   drop_na(p.value) %>%
+#   select(depth, name, p.value) %>%
+#   group_by(name) %>%
+#   mutate(p.value=p.adjust(p.value, method="fdr")) %>%
+#   ungroup() %>%
+#   arrange(name, depth)
+# pc_comp_df %>%
+#   ggplot(aes(x=sla_class, y=frac_in_metab)) +
+#   geom_boxplot() +
+#   facet_wrap(~depth, nrow=1)
+
+# Permutation testing for effect size test on Deep SLA differences
+# set.seed(123)
+# map_dbl(seq(1, 2, 0.1), function(n_mult){
+#   map_dbl(1:100, function(h){
+#     pc_comp_df %>%
+#       filter(depth=="Deep") %>%
+#       group_by(sla_class) %>%
+#       summarize(mean=mean(frac_in_metab), sd=sd(frac_in_metab), n=n()) %>%
+#       mutate(sim_vals=map(1:3, function(i){rnorm(.$n[i]*n_mult, .$mean[i], .$sd[i])})) %>%
+#       unnest_longer(sim_vals) %>%
+#       aov(formula=sim_vals~sla_class) %>%
+#       broom::tidy() %>%
+#       filter(term=="sla_class") %>%
+#       pull(p.value)
+#   }) %>% mean()
+# })
+
+pc_frac_metab_gp <- ggplot(pc_comp_df, aes(x=station, y=frac_in_metab, color=sla_corr)) +
+  geom_hline(yintercept = 0) +
+  geom_boxplot(width = 0.9, linewidth=0.8) +
+  geom_point(color="black", size=2) +
+  scale_y_continuous(labels = scales::label_percent()) +
+  facet_grid(~depth, scales="free") +
+  scale_color_gradient2(
+    low = scales::muted("blue"), high = scales::muted("red"),  mid = "grey90",
+    aesthetics = c("color", "fill")
+  ) +
+  guides(color=guide_colorbar(direction = "horizontal", title.position = "top", 
+                              title.hjust = 0.5)) +
+  labs(color="Sea level\nanomaly (cm)", x="Station",
+       y="Fraction of PC represented by known metabolites") +
+  theme_bw() +
+  theme(legend.position = c(0.85, 0.8), legend.background = element_rect(color="grey90"))
+ggsave("pc_frac_metab_gp.png", pc_frac_metab_gp, device = "png", path = "supplement",
+       width = 6.5, height = 3.5, units = "in", dpi = 300)
+```
+
+![](supplement/pc_frac_metab_gp.png)
+
+*Supplemental figure: Fraction of particulate carbon represented by known metabolites.
+Boxplots set behind points are colored according to sea level anomaly (SLA) with red corresponding to a positive SLA (anticyclone) and blue corresponding to negative SLA (cyclone).*
+
+This correlation between PC and overall metabolite abundance implied that the fraction of PC represented by known metabolites was likely to be fixed across the eddy transect and this expectation was validated by followup analyses.
+Known metabolites contributed between approximately 1.5% and 5% of the carbon biomass to the total PC load, but this variance was not clearly due to either depth or SLA.
+The only difference we observed was a slightly reduced metabolite contribution in the 175 meter samples (0.3-0.4 percentage points lower at depth, p-value=0.018) while no difference was detected between the surface and DCM samples or between the cyclone, anticyclone, and outside the eddy samples (all p-values > 0.1).
+However, this lack of difference observed may have been due to the small sample sizes (18 anticyclonic samples, 9 cyclonic, and 6 neither).
+A followup permutational sensitivity analysis indicated that a difference in means of 0.4 percentage points would have been obtained at the α < 0.05 level for the 175 meter samples if 1.5x as many samples had been collected with mean and variances remaining similar.
+
+  - Discussion: why are my % PC in metabs lower than Angie's? Because her PC values were way higher but metab concentrations were about the same. Wait, are they even?
+
+```{r comparison to Angies PC values way higher than mine, eval=FALSE}
+angie_pcdata <- "../metadata/raw_data/Karl_HL2_WC_PCPN_Final_Data.xlsx" %>%
+  openxlsx::read.xlsx() %>%
+  slice(-1) %>%
+  select(depth=`Depth.(m)`, PC_um=`C.value.(µmol/L)`) %>%
+  filter(depth<30) %>%
+  mutate(cruise="HOE-Legacy")
+
+my_pcdata <- filled_file_metadata %>%
+  filter(depth=="Surface") %>%
+  distinct(depth, PC_um) %>%
+  mutate(cruise="MESOSCOPE")
+
+rbind(angie_pcdata, my_pcdata) %>%
+  ggplot() +
+  geom_boxplot(aes(x=cruise, y=PC_um)) +
+  scale_y_continuous(limits = c(0, NA))
+```
+
+```{r lm_line_rank_order supplemental figure}
+reg_coefs <- untarg_peaks %>%
+  left_join(filled_file_metadata %>% distinct(filename, ms_run, station, depth, sla_corr)) %>%
+  filter(ms_run=="MT") %>%
+  group_by(feature) %>%
+  filter(sum(norm_area==0)<n()*0.9) %>%
+  mutate(norm_area=rank(norm_area)) %>%
+  ungroup() %>%
+  nest(data=-c(depth, feature)) %>%
+  mutate(test=map(data, ~broom::tidy(lm(.x$norm_area~.x$sla_corr)))) %>%
+  unnest(test)
+reg_coefs %>%
+  filter(str_detect(term, "sla_corr")) %>%
+  mutate(p.value=p.adjust(p.value, method="fdr")) %>%
+  arrange(p.value) %>%
+  left_join(untarg_feats) %>%
+  select(feature, depth, estimate, p.value, mzmed, rtmed, compound_name) %>%
+  filter(p.value<0.05) %>%
+  mutate(depth=factor(depth, levels=c("Surface", "DCM", "Deep"))) %>%
+  mutate(rtmed=round(rtmed, digits = 2)) %>%
+  mutate(mzmed=round(mzmed, digits = 5)) %>%
+  arrange(depth, p.value) %>%
+  group_by(depth) %>%
+  group_modify(~ add_row(.x, .after = nrow(.))) %>%
+  mutate(across(everything(), ~ifelse(is.na(.x), "", .x))) %>%
+  write_csv("supplement/lm_est_rank_order.csv")
+
+clean_slopes <- reg_coefs %>%
+  select(feature, data, depth, term, estimate, p.value) %>%
+  mutate(term=ifelse(term==".x$sla_corr", "slope", "intercept")) %>%
+  pivot_wider(names_from = term, values_from=c(p.value, estimate)) %>%
+  mutate(p.value=p.adjust(p.value_slope, method="fdr")) %>%
+  select(-p.value_intercept, -p.value_slope) %>%
+  filter(p.value<0.05) %>%
+  mutate(pointcol=ifelse(sign(estimate_slope)==1, "Anticyclone enriched", "Cyclone enriched")) %>%
+  mutate(depth=factor(depth, levels=c("Surface", "DCM", "Deep")))
+metadata_slopes <- filled_file_metadata %>%
+  filter(ms_run=="MT") %>%
+  filter(samp_type=="Smp") %>%
+  select(depth, pressure:beam_atten) %>%
+  distinct() %>%
+  pivot_longer(-c(sla_corr, depth)) %>%
+  group_by(name) %>%
+  mutate(value=rank(value)*3) %>%
+  nest(data=-c(name, depth)) %>%
+  mutate(lmout=map(data, ~broom::tidy(lm(.x$value~.x$sla_corr)))) %>%
+  unnest(lmout) %>%
+  filter(name!="dcm_depth") %>%
+  select(depth, name, data, term, estimate, p.value) %>%
+  pivot_wider(names_from = term, values_from = c(estimate, p.value)) %>%
+  mutate(p.value=p.adjust(`p.value_.x$sla_corr`)) %>%
+  filter(p.value<0.05) %>%
+  select(-c(`p.value_(Intercept)`, `p.value_.x$sla_corr`)) %>%
+  mutate(intercept=`estimate_(Intercept)`, slope=`estimate_.x$sla_corr`) %>%
+  mutate(pointcol=ifelse(sign(slope)==1, "Anticyclone enriched", "Cyclone enriched")) %>%
+  mutate(depth=factor(depth, levels=c("Surface", "DCM", "Deep"))) %>%
+  mutate(name=factor(name, levels=metadata_levels, labels=metadata_short_labels))
+ggplot() +
+  geom_abline(aes(intercept=estimate_intercept, slope=estimate_slope), data = clean_slopes) +
+  geom_point(aes(x=sla_corr, y=norm_area), data=unnest(clean_slopes, data)) +
+  geom_abline(data = metadata_slopes, aes(slope=slope, intercept=intercept, color=name)) +
+  geom_point(aes(x=sla_corr, y=value, color=name), data=unnest(metadata_slopes, data)) +
+  facet_grid(pointcol~depth) +
+  ggtitle("Todo: fix this graph")
+```
+
+*Supplemental figure: Linear regressions of rank-order peak area against SLA showing strong enrichment at the cyclone DCM and the anticyclone 175 meter samples.*
+
+### Compositional shifts in metabolome as well as bulk biomass
+
+In addition to the absolute shifts in metabolite concentration due to biomass differences, we also explored the relative composition of the particulate matter.
+
+```{r regression relative abundance coefficients, eval=FALSE}
+reg_coefs <- untarg_peaks %>%
+  # filter(feature!="MS_NEG_0155") %>%
+  left_join(filled_file_metadata %>% distinct(filename, ms_run, station, depth, sla_corr)) %>%
+  filter(ms_run=="MT") %>%
+  filter(depth%in%c("Surface", "DCM", "Deep")) %>%
+  filter(feature%in%unique(feature)) %>%
+  group_by(filename) %>%
+  mutate(norm_area=norm_area/sum(norm_area)) %>%
+  ungroup() %>%
+  nest(data=-c(depth, feature)) %>%
+  mutate(test=map(data, ~broom::tidy(lm(.x$norm_area~.x$sla_corr)), .progress = TRUE)) %>%
+  unnest(test)
+reg_coefs %>%
+  filter(str_detect(term, "sla_corr")) %>%
+  mutate(p.value=p.adjust(p.value, method="fdr")) %>%
+  arrange(p.value) %>%
+  left_join(untarg_feats) %>%
+  select(feature, depth, estimate, p.value, mzmed, rtmed, compound_name) %>%
+  filter(p.value<0.05) %>%
+  # as.data.frame() %>%
+  filter(depth=="Deep") %>% 
+  print(n=Inf)
+reg_coefs %>%
+  filter(str_detect(term, "sla_corr")) %>%
+  left_join(untarg_feats) %>%
+  filter(compound_name=="Acetylcholine?") %>%
+  # filter(feature=="MS_POS_1395") %>%
+  filter(depth=="Deep") %>%
+  select(compound_name, depth, data, estimate) %>%
+  unnest(data) %>%
+  group_by(sla_corr, station) %>%
+  ggplot(aes(x=sla_corr, y=norm_area*100)) +
+  geom_smooth(method="lm", formula="y~x") +
+  geom_boxplot(aes(group=sla_corr)) +
+  scale_y_continuous(limits = c(0, NA))
+```
+
+In the 15m surface samples, only seven compounds had significant changes in relative peak area across the eddy transect, two of which were known: 5-oxoproline, which decreased from 0.2-0.3% of the total peak area at the center of the anticyclone to 0.05-0.1% in the cyclone; and a combined peak of sarcosine & beta-alanine which increased from 0.01% to 0.02-0.03%.
+The largest shift significant at an α = 0.05 level was an unknown mass feature with *m/z* = 189.12338 and a retention time around 8.6 minutes, the relative contribution of which increased from approximately 0.2-0.3% in the cyclone to ~1.5% in the anticyclone.
+
+At the DCM, twenty-two compounds changed significantly in relative peak area across the eddy dipole.
+Seven of these were known metabolites, of which trigonelline and homarine were most enriched in the cyclone while arsenobetaine was the only metabolite with a larger peak area fraction in the anticyclone.
+Homarine showed a very large and highly significant shift, representing about 2.5% of the total peak area in the anticyclone but 7-8% in the cyclone center.
+Trigonelline, an isomer structurally very similar but biologically entirely distinct, had a surprisingly strong correlation with homarine (Pearson's $r=.855$) with peak areas consistently around one third of homarine's but large enough to have the second-largest shift in relative peak area of the significantly different compounds.
+Similarly, arsenobetaine represented about 0.1% of the peak area in the cyclone and 0.5-0.6% of the total peak area in the anticyclone.
+The most significantly different compounds at the DCM in each direction, however, were both unknowns.
+The lowest p-value ($6.4*10^{-6}$ after FDR correction) in the DCM data was a mass feature with *m/z* of 173.09211, a retention time also around 8.6 minutes, and enrichment in the anticyclone with a putative chemical formula of [M+H] = C$_7$H$_{13}$N$_2$O$_3$, possibly glyclproline or prolylglycine.
+The next-lowest (p = $2.5*10^{-4}$ after FDR correction) was enriched in the cyclone and had an *m/z* of 275.07119 and a retention time of 11.4 minutes with a putative formula of C$_{18}$H$_{11}$O$_3$.
+
+Finally, in the 175 meter samples we detected 32 mass features with a significant association with SLA.
+Notably, all detected nucleobases (guanine, adenine, and cytosine) were positively associated with SLA (higher values in the anticyclone than the cyclone), though the strongest associations were found in O-acetylcarnitine, betonicine, and tyrosine.
+Both O-acetylcarnitine and betonicine effectively quintupled their contribution to total peak area, shifting from approximately 0.1% to ~0.5% over the transect.
+Acetylcholine was the only known compound more relatively abundant in the cyclone than the anticyclone along with one unknown of *m/z* 131.034.
+The unknown at *m/z* 173.09211 with the strongest trend at the DCM again had the strongest trend in the 175 meter samples, here with an FDR-corrected p-value of $2.0*10^{-7}$.
+
+  - Discussion: surprised that so many compounds are still different after normalizing within each file - expected that to remove biomass signal
+- Discussion: why are homarine and trigonelline so correlated?
+- Discussion: homarine/trigonelline has a larger effect size but lower p-value because of nonlinear fit (exponential rise into eddy center rather than linear)
+- Discussion: how can most compounds be relatively enriched in the anticyclone at 175 meters? Where is the remainder of the effect size going?
+- Discussion: all the nucleobases upregulated in the 175m anticyclone, why? Bacteria?
+
+```{r possible figure stacked barplot ranked space}
+top10_stacked_area_station_gp <- map(c("Surface", "DCM", "Deep"), function(depth_i){
+  reg_coefs %>%
+    filter(str_detect(term, "sla_corr")) %>%
+    mutate(p.value=p.adjust(p.value, method="fdr")) %>%
+    filter(depth==depth_i) %>%
+    arrange(p.value) %>%
+    group_by(estimate>0) %>%
+    slice(1:5) %>%
+    ungroup() %>%
+    left_join(untarg_feats %>% select(feature, compound_name, mzmed, rtmed)) %>%
+    mutate(feature=ifelse(!is.na(compound_name), compound_name, 
+                          paste0(round(mzmed, 3), "_", round(rtmed, 1)))) %>%
+    mutate(feature=fct_inorder(feature)) %>%
+    unnest(data) %>%
+    select(feature, depth, norm_area, filename, station) %>%
+    group_by(feature) %>%
+    # mutate(norm_area=rank(norm_area)) %>%
+    mutate(norm_area=norm_area/sum(norm_area)) %>%
+    mutate(station=factor(station, levels=as.character(4:14))) %>%
+    ggplot() +
+    geom_col(aes(x=station, y=norm_area, fill=feature), position = "fill") +
+    scale_y_continuous(expand = c(0,0)) +
+    labs(x="Station", y="Relative peak area", fill="Mass feature") +
+    theme_bw()
+}) %>% gridExtra::arrangeGrob(grobs = ., ncol = 1)
+ggsave("top10_stacked_area_station_gp.png", plot = top10_stacked_area_station_gp, 
+       device = "png", path = "supplement", width = 6.5, height = 9, 
+       units = "in", dpi = 300)
+```
+
+![](supplement/top10_stacked_area_station_gp.png)
+
+*Supplemental figure?
+Plots of the 5 most significantly enriched metabolites in both the cyclone (top 5) and anticyclone (bottom 5) after normalizing to the total peak area in each mass feature.
+Nicely shows a stronger trend with increasing depth but feels a little bit like lying because there's 100 other compounds that don't change in between these extremes.
+Tyrosine is also misannotated here.*
+
+```{r homarine/trigonelline correlation, eval=FALSE}
+reg_coefs %>%
+  filter(str_detect(term, "sla_corr")) %>%
+  left_join(untarg_feats) %>%
+  filter(compound_name%in%c("Homarine", "Trigonelline")) %>%
+  filter(depth=="DCM") %>%
+  select(compound_name, depth, data, estimate) %>%
+  unnest(data) %>%
+  select(compound_name, filename, norm_area) %>%
+  pivot_wider(names_from=compound_name, values_from=norm_area) %>%
+  column_to_rownames("filename") %>%
+  cor()
+```
+
+### Eddy center DCM samples show differences between depths and eddy polarities
+
+```{r MC NMDS and PERMANOVA calculations}
+set.seed(123)
+multivar_analysis_MC <- untarg_peaks %>%
+  left_join(filled_file_metadata %>% distinct(ms_run, station, filename, depth), 
+          by="filename") %>% filter(ms_run=="MC") %>% 
+  mutate(depth=factor(depth, levels=c("DCMless20m", "DCMless10m", "DCM", 
+                                      "DCMplus10m", "DCMplus20m"))) %>%
+  nest(data=everything()) %>%
+  mutate(wide_peaks=map(data, function(metab_df){
+    metab_df %>%
+      select(feature, filename, norm_area) %>%
+      group_by(feature) %>%
+      filter(sum(norm_area==0)<n()*0.9) %>%
+      mutate(norm_area=scale(norm_area)[,1]) %>%
+      pivot_wider(names_from = "feature", values_from = norm_area) %>%
+      column_to_rownames("filename") %>%
+      data.matrix()
+  })) %>%
+  mutate(nmds_output=map(wide_peaks, function(metab_mat){
+    env_vec <- pull(filled_file_metadata, sla_corr, filename)[rownames(metab_mat)]
+    vegan::metaMDS(metab_mat, k = 2, distance = "manhattan", trace = 0, 
+                   autotransform = FALSE, noshare = FALSE, wascores = FALSE) %>%
+    vegan::MDSrotate(env_vec)
+  })) %>%
+  mutate(nmds_stress=map_dbl(nmds_output, pluck, "stress")) %>%
+  mutate(perm_output=map(wide_peaks, function(metab_mat){
+    adon_expl <- data.frame(filename=rownames(metab_mat)) %>%
+      left_join(filled_file_metadata %>% select(filename, sla_class, depth), by="filename")
+    vegan::adonis2(
+      metab_mat~sla_class+depth, data = adon_expl, by = "margin", 
+      permutations = 999, method = "manhattan")
+  })) %>%
+  mutate(perm_r2_sla=map_dbl(perm_output, pluck, "R2", 1)) %>%
+  mutate(perm_r2_depth=map_dbl(perm_output, pluck, "R2", 2)) %>%
+  mutate(perm_pval=map_dbl(perm_output, pluck, "Pr(>F)", 1)) %>%
+  mutate(pca_output=map(wide_peaks, prcomp)) %>%
+  mutate(nmds_stress=paste("s =", round(nmds_stress, digits = 3)))
+  # mutate(perm_r2=round(perm_r2, digits = 2)) %>%
+  # mutate(perm_pval=round(perm_pval, digits = 4)) %>%
+  # mutate(perm_label=paste0("R2 = ", perm_r2, "\np = ", perm_pval))
+```
+
+To provide context for the sea level anomaly responses detected above, we additionally performed high-resolution depth sampling around the DCM at the center of both eddy poles.
+Here, we found that the ~50 meter difference in depth and the ~40 centimeter difference in SLA explained similar amounts of variation in the data (R$^2_{depth}=0.253$, R$^2_{SLA}=0.238$) and were both highly significant factors (permutational p-values << 0.001).
+Additionally, both sample depth and SLA were well-correlated with the first principal component of the metabolite matrix (r$_{depth}=.766$, r$_{SLA}=.684$, fraction of variance explained by PC$_1$ = 30.8%) (Figure 3).
+
+```{r figure MC NMDS plot}
+MC_nmds_gp <- multivar_analysis_MC$nmds_output %>%
+  pluck(1, "points") %>% 
+  as.data.frame() %>%
+  rownames_to_column("filename") %>%
+  left_join(filled_file_metadata %>% distinct(filename, sla_class, depth)) %>%
+  mutate(depth=factor(depth, levels=c("DCMless20m", "DCMless10m", "DCM", 
+                                      "DCMplus10m", "DCMplus20m"),
+                      labels=c("-20m", "-10m", "DCM", "+10m", "+20m"))) %>%
+  ggplot() +
+  geom_vline(xintercept = 0) +
+  geom_hline(yintercept = 0) +
+  geom_point(aes(x=MDS1, y=MDS2, color=sla_class), size=4, pch=19) +
+  scale_color_manual(
+    values=c(scales::muted("blue"), scales::muted("red")),
+    breaks=c("Cyclone", "Anticyclone"),
+  ) +
+  annotate("label", x=-200, y=100, label="Cyclone", color=scales::muted("blue"),
+           label.r = unit(0, "in"), size=6) +
+  annotate("label", x=100, y=120, label="Anticyclone", color=scales::muted("red"),
+           label.r = unit(0, "in"), size=6) +
+  annotate("label", x=-Inf, y=Inf, label=multivar_analysis_MC$nmds_stress,
+           hjust=0, vjust=1, label.r = unit(0, "in")) +
+  geom_point(aes(x=MDS1, y=MDS2, fill=depth, alpha=depth), size=4, pch=21, color="black") +
+  scale_fill_manual(
+    values = c("grey90", "grey80", "grey60", "grey40", "grey20"),
+    breaks = c("-20m", "-10m", "DCM", "+10m", "+20m")
+  ) +
+  scale_alpha_manual(
+    values = c(0.4, 0.2, 0, 0.2, 0.4),
+    breaks = c("-20m", "-10m", "DCM", "+10m", "+20m")
+  ) +
+  coord_fixed() +
+  guides(alpha="none", color="none", fill=guide_legend(
+    title = "Depth", title.position = "top", title.hjust = 0.5, direction = "horizontal"
+  )) +
+  theme_bw() +
+  theme(legend.position = "bottom")
+ggsave("MC_nmds_gp.png", plot = MC_nmds_gp, device = "png", 
+       path = "figures", width = 6.5, height = 4.5, units = "in", dpi = 300)
+```
+
+![](figures/MC_nmds_gp.png)
+
+*Figure 3: NMDS plot of high-resolution depth sampling around the deep chlorophyll maximum (DCM, ~110 meters) at the eddy center of each polarity.
+Points are colored by the eddy from which the associated sample was collected and shaded by the depth.
+Visible separation between the cyclonic and anticyclonic eddy is evident along the first NMDS axis.*
+
+```{r PCA checks for MC axis regressions, eval=FALSE}
+multivar_analysis_MC$pca_output[[1]]$x %>%
+  as.data.frame() %>%
+  rownames_to_column("filename") %>%
+  left_join(filled_file_metadata) %>%
+  select(abs_depth, sla_corr, PC1:PC3) %>%
+  # cor() %>% round(3)
+  ggplot(aes(x=abs_depth, y=PC1, color=sla_corr)) +
+  geom_point()
+
+head(multivar_analysis_MC$pca_output[[1]]$sdev^2/sum(multivar_analysis_MC$pca_output[[1]]$sdev^2)*100)
+```
+
+  - Discussion: discuss how the cyclone+20m is pretty similar to the anticyclone samples and why
+  - Discussion: less of a clear depth trend despite the stats saying equal?
+  - Discussion: two possible models, one where samples from the same absolute depth are similar, one where samples from the same distance from DCM are similar.
+Can I argue for which one is better here?
+
+To characterize the observed differences in multivariate space we used k-means clustering as an unsupervised way to identify dominant trends within the data (Figure 4).
+We found that a majority of the metabolites (Clusters 1 and 2 in Figure 4, 63% of the total) fell into clusters with larger peak areas in the cyclone, as expected, while also detecting 32 metabolites that clustered such that the mean metabolite was enriched in the anticyclone (Cluster 4, Figure 4).
+Cluster #2 notably had a distinct decrease in relative peak area with depth, while the other clusters had much less clear depth trends.
+Cluster #1 also showed a neat DCM maximum for the samples from the anticyclone, correlating well with flow cytometry counts of picoeukaryotes (r=.851) that was not present in the cyclone (r=.013).
+
+```{r k-means clustering}
+set.seed(123)
+kclust <- untarg_peaks %>%
+  left_join(filled_file_metadata, by="filename") %>%
+  filter(ms_run=="MC") %>%
+  select(feature, filename, norm_area) %>%
+  group_by(feature) %>%
+  filter(sum(norm_area==0)<n()*0.9) %>%
+  mutate(norm_area=scale(norm_area)[,1]) %>%
+  pivot_wider(names_from = "filename", values_from = norm_area) %>%
+  column_to_rownames("feature") %>%
+  data.matrix() %>%
+  kmeans(centers = 4)
+
+kclust_nms <- names(kclust$cluster)
+kclust$cluster <- order(kclust$size, decreasing = TRUE)[kclust$cluster]
+names(kclust$cluster) <- kclust_nms
+kclust$size <- sort(kclust$size, decreasing = TRUE)
+
+kclust_df <- kclust %>%
+  pluck("cluster") %>%
+  data.frame(cluster=.) %>%
+  rownames_to_column("feature") %>%
+  left_join(untarg_peaks, multiple = "all", by="feature") %>%
+  left_join(filled_file_metadata, by="filename") %>%
+  filter(ms_run=="MC") %>%
+  arrange(desc(abs_depth)) %>%
+  mutate(depth=fct_inorder(depth)) %>%
+  group_by(feature) %>%
+  mutate(norm_area=scale(norm_area)[,1]) %>%
+  group_by(cluster, depth, sla_class) %>%
+  summarise(mean_rank=mean(norm_area), iqr_value=sd(norm_area), .groups = "drop")
+kclust_gp <- kclust_df %>%
+  left_join(data.frame(size=kclust$size) %>% mutate(cluster=row_number()), by="cluster") %>%
+  mutate(cluster=paste0("Cluster #", cluster, ": n=", size)) %>%
+  mutate(depth=factor(depth, labels=rev(c("-20m", "-10m", "DCM", "+10m", "+20m")),
+                      levels=rev(c("DCMless20m", "DCMless10m", "DCM", "DCMplus10m", "DCMplus20m")))) %>%
+  ggplot(aes(y=depth, color=sla_class)) +
+  geom_vline(xintercept = 0) +
+  # geom_ribbon(aes(xmin=mean_rank-iqr_value, xmax=mean_rank+iqr_value, group=sla_class),
+  #             linewidth=1, fill="#FFFFFF22") +
+  geom_path(aes(x=mean_rank, group=sla_class), linewidth=1) +
+  geom_point(aes(x=mean_rank), size=2) +
+  geom_label(data = data.frame(cluster="Cluster #1: n=73", x=c(0, 0.9), depth=c(5.5, 1),
+                               sla_class=c("Anticyclone", "Cyclone")),
+             aes(x=x, label=sla_class)) +
+  facet_wrap(~cluster, nrow = 1) +
+  scale_color_manual(values = c(scales::muted("blue"), scales::muted("red"), "grey50"), 
+                     breaks = c("Cyclone", "Anticyclone", "Neither"),
+                     guide="none") +
+  labs(x="Mean z-scaled metabolite area", y="Sampling depth", color=NULL) +
+  coord_cartesian(ylim = c(1, 5.5)) +
+  theme_bw()
+
+ggsave("kclust_gp.png", plot = kclust_gp, device = "png", 
+       path = "figures", width = 6.5, height = 2.5, units = "in", dpi = 300)
+```
+
+![](figures/kclust_gp.png)
+
+*Figure 4: k-means clustering performed on the high-resolution depth samples from the centers of each eddy.
+Points denote the average z-scored peak area for both known and unknown metabolites across the samples and are colored by the eddy from which they were taken.
+Clusters have been ordered by number of metabolites in each group and the total is denoted in the panel titles.
+Both depth trends (mostly a net decrease in metabolites with depth) and eddy effects (cyclonic enrichment in clusters 1 and 2, anticyclone enrichment in cluster 4) are observable.*
+
+  - Discussion: what does the existence of cluster 3 mean? Tracking with biomass exactly or just too much noise?
+
+```{r cluster cor w metadata, eval=FALSE}
+filled_file_metadata %>% 
+  filter(ms_run=="MC") %>% 
+  distinct(depth, euk_abund, sla_class) %>% 
+  left_join(kclust_df) %>%
+  filter(cluster==1) %>%
+  filter(sla_class=="Cyclone") %>%
+  # with(cor(mean_rank, euk_abund))
+  ggplot() +
+  geom_point(aes(x=euk_abund, y=mean_rank, color=sla_class))
+```
+
+We also investigated the response of individual metabolites in these high-resolution depth profile samples, expecting to find the strongest signal in this data given the extreme SLA values sampled and the large number of compounds with an SLA separation in the k-means analysis.
+This expectation was substantiated by a majority of the metabolites differing significantly between the two eddies (121/228, α = 0.05).
+Of those, 15 were enriched in the anticyclone and 106 were more abundant in the cyclone.
+As expected, all compounds enriched in the anticyclone were part of Cluster 4 and all those enriched in the cyclone belonged to either cluster 1 or 2 (Figure 5).
+
+```{r pairwise_tests and fold_change_df}
+MC_simpledata <- untarg_peaks %>%
+  left_join(filled_file_metadata %>% distinct(filename, ms_run, station), by="filename") %>%
+  filter(ms_run=="MC") %>%
+  select(feature, filename, norm_area, station) %>%
+  group_by(feature) %>%
+  filter(sum(norm_area==0)<n()*0.9) %>%
+  group_by(filename) %>%
+  # mutate(norm_area=norm_area/sum(norm_area)) %>%
+  ungroup()
+pairwise_tests <- MC_simpledata %>%
+  nest(data=c(filename, norm_area, station)) %>%
+  mutate(stattest=map(data, ~broom::tidy(wilcox.test(.x$norm_area~.x$station)))) %>%
+  unnest(stattest) %>%
+  select(feature, statistic, p.value) %>%
+  mutate(p.value=p.adjust(p.value, method="fdr"))
+
+fold_change_df <- MC_simpledata %>%
+  group_by(feature, station) %>%
+  summarise(mean_area=mean(norm_area, na.rm=TRUE)) %>%
+  pivot_wider(names_from=station, values_from=mean_area) %>%
+  mutate(fold_change=L1/L2) %>%
+  select(feature, fold_change)
+
+```
+
+```{r eddy center volcano plot}
+volcanoplot_gp <- pairwise_tests %>%
+  left_join(fold_change_df, by = join_by(feature)) %>%
+  left_join(rownames_to_column(data.frame(cluster=kclust$cluster), "feature"), 
+            by = join_by(feature)) %>%
+  left_join(untarg_feats %>% select(feature, compound_name), by = join_by(feature)) %>%
+  mutate(known=ifelse(is.na(compound_name), "Unknown", "Known")) %>%
+  mutate(compound_name=ifelse(is.na(compound_name), feature, compound_name)) %>%
+  ggplot(aes(x=log2(fold_change), y=-log10(p.value), color=factor(cluster), 
+             shape=known, label=compound_name)) +
+  geom_vline(xintercept = 0) +
+  geom_hline(yintercept = 0) +
+  geom_hline(yintercept = 2, lty=2) +
+  geom_point(size=2, stroke=1) +
+  scale_y_continuous(breaks = 0:6, labels = paste0("10<sup>-", 0:6, "</sup>")) +
+  scale_shape_manual(values=c(4, 16), breaks = c("Unknown", "Known")) +
+  scale_color_manual(
+    values = c("#414487FF", "#7AD151FF", "#22A884FF", "#2A788EFF"),
+    breaks = 1:4
+  ) +
+  theme_bw() +
+  labs(x="Log-2 fold-change cyclone enrichment", y="Mann-Whitney FDR-corrected p-value",
+       shape="Metabolite\nstatus", color="k-means\ncluster") +
+  guides(color=guide_legend(order = 2), shape=guide_legend(order=1)) +
+  theme(axis.text.y = ggtext::element_markdown())
+
+ggsave("volcanoplot_gp.png", plot = volcanoplot_gp, device = "png", 
+       path = "figures", width = 6.5, height = 3, units = "in", dpi = 300)
+```
+
+![](figures/volcanoplot_gp.png)
+
+*Figure 5: Volcano plot of known and unknown metabolites from high-resolution depth sampling at the deep chlorophyll maximum (DCM) of each eddy pole.
+Points correspond to the FDR-corrected p-value estimated by the nonparametric Mann-Whitney U test and the log$_2$ fold-change calculated with the average peak area in the cyclone divided by the average peak area in the anticyclone.
+Colors have been assigned using the k-means clusters from Figure 4 above and shapes have been assigned based on the status of the mass feature as either a known metabolite that was run with an authentic standard or an unidentified metabolite.*
+
+Many of the known metabolites enriched in the cyclone are expected given the overall higher biomass, particularly as many have functions as osmolytes in the cell.
+Metabolites enriched in excess of biomass were taurine and isethionic acid, the reduced sulfur osmolytes dimethylsulfoniopropionate (DMSP) and dimethylsulfonioacetate (DMS-Ac), and the isomers homarine and trigonelline.
+A few metabolites more abundant in the anticyclone were also given putative identifications based on RT and *m/z* matching with internal standards later added.
+Of these, the putative arsenobetaine was the most significantly different among these with a peak area at the anticyclone DCM nearly quadruple that of the cyclone.
+Of note, the 173.09211 mass feature noted above was among the most significantly different between the two eddies, along with two isomers at an *m/z* of 170.11756 and retention times around 8-9 minutes that increased by a factor of 1.5 in the anticyclone (putative formula C$_7$H$_{14}$N$_4$O).
+
+  - Discussion: implications of the mostly diatom/dinoflagellate metabolites upregulated in cyclone likely reflecting the fact that those are the species most abundant.
+
+```{r putative 170.11756 formula, eval=FALSE}
+# Two possible formula: C9H16NO2 and C7H14N4O
+library(arrow)
+dset <- open_dataset("G:/My Drive/AllMeso/tmzMLs/MC_all_file_arr_pqds")
+
+# Too low-intensity for reliable isotope info
+dset %>%
+  filter(mz%between%pmppm(170.11757, 5)) %>%
+  filter(rt%between%c(7, 10)) %>%
+  collect() %>%
+  left_join(filled_file_metadata) %>%
+  ggplot() +
+  geom_line(aes(x=rt, y=int, group=filename, color=station))
+rtr <- c(8.7, 9.2)
+gly_data <- dset %>%
+  filter(rt%between%rtr) %>%
+  collect()
+mzr <- pmppm(170.11756, 10)
+reduce(list(
+  gly_data[mz%between%mzr] %>%
+    group_by(filename) %>%
+    summarise(M=sum(int)),
+  gly_data[mz%between%(mzr+1.003355)] %>%
+    group_by(filename) %>%
+    summarise(C13=sum(int))
+), merge, all.x=TRUE) %>%
+  pivot_longer(C13) %>%
+  mutate(prop=value/M) %>%
+  print(n=Inf)
+
+# But there are useful MS2 fragments
+library(RaMS)
+mz_group <- function(mz_vals, ppm){
+  group_vec <- numeric(length(mz_vals))
+  group_num <- 1L
+  init_vec <- mz_vals
+  names(init_vec) <- seq_along(init_vec)
+  while(length(init_vec)>0){
+    mz_i <- init_vec[1]
+    err <- mz_i*ppm/1000000
+    mz_idxs <- init_vec>mz_i-err & init_vec<mz_i+err
+    group_vec[as.numeric(names(mz_idxs)[mz_idxs])] <- group_num
+    init_vec <- init_vec[!mz_idxs]
+    group_num <- group_num+1L
+  }
+  group_vec
+}
+msmsdata_MC <- "../mzMLs/pos/MSMS" %>%
+  list.files(, full.names = TRUE, pattern = "180205") %>%
+  grabMSdata()
+msmsdata_MC$MS2[premz%between%pmppm(170.11757, 10)] %>%
+  mutate(mz_group=mz_group(fragmz, 5)) %>%
+  group_by(mz_group, filename) %>%
+  summarize(fragmz=mean(fragmz), int=mean(int)) %>%
+  filter(n()>1) %>%
+  group_by(mz_group) %>%
+  summarise(fragmz=mean(fragmz), int=mean(int)) %>%
+  filter(fragmz<170) %>%
+  arrange(desc(int))
+# Top two frags have formulae of 125.1075 = C7H13N2
+# and 89.07148 = C3H9N2O, both requiring 2 N atoms
+```
+
+```{r comparison to Katherines culture data}
+cyc_enriched <- pairwise_tests %>%
+  filter(p.value<0.05) %>%
+  left_join(untarg_feats) %>%
+  filter(!is.na(compound_name)) %>%
+  filter(statistic>110)
+
+enriched_metabs <- c("Homarine", "Trigonelline", "Taurine", "DMSP",
+                     "Dimethylsulfonioacetate (DMS-Ac)", "Isethionic Acid")
+culture_metadata <- "../metadata/raw_data/Heal_Metabs_SuppTabs_corrected20230327.xlsx" %>%
+  readxl::read_excel(sheet = "S7_Culturedescriptions_full")
+culture_areas <- "../metadata/raw_data/Heal_Metabs_SuppTabs_corrected20230327.xlsx" %>%
+  readxl::read_excel(sheet = "S5_FullEnviroCulture_PeakAreas") %>%
+  filter(Identification%in%enriched_metabs)
+katherine_comp_mostcycenriched_gp <- culture_areas %>%
+  select(Identification, any_of(culture_metadata$`Sample ID`)) %>%
+  mutate(Identification=ifelse(Identification=="Dimethylsulfonioacetate (DMS-Ac)", 
+                               "DMS-Ac", Identification)) %>%
+  mutate(Identification=factor(Identification, levels=c(
+    "Homarine", "Trigonelline", "Isethionic Acid", "Taurine", "DMSP","DMS-Ac"
+  ))) %>%
+  pivot_longer(-Identification, names_to = "Sample ID") %>%
+  mutate(value=as.numeric(value)) %>%
+  left_join(culture_metadata, by=join_by(`Sample ID`)) %>%
+  arrange(`Broad taxon`) %>%
+  mutate(`Short ID`=fct_inorder(`Short ID`)) %>%
+  mutate(`Species`=fct_inorder(`Species`)) %>%
+  drop_na() %>%
+  ggplot() +
+  geom_boxplot(aes(x=Species, y=value, color=`Broad taxon`)) +
+  facet_wrap(~Identification, ncol = 2, scales="free_y") +
+  scale_y_log10(name="Peak area") +
+  theme_bw() +
+  theme(axis.text.x = element_text(angle=90, hjust=1, vjust=0.5))
+
+ggsave("katherine_comp_mostcycenriched_gp.png", plot = katherine_comp_mostcycenriched_gp, 
+       device = "png", path = "supplement", width = 6.5, height = 5, units = "in", dpi = 300)
+```
+
+![](supplement/katherine_comp_mostcycenriched_gp.png)
+
+*Supplemental figure? Culture data from Katherine's manuscript showing the presence of the DCM cyclone-enriched metabolites mostly present in diatoms and dinoflagellates.*
+
+```{r enrichment plot DCM comparison, eval=FALSE}
+MC_stacked_mostdiff_gp <- pairwise_tests %>%
+  arrange(p.value) %>%
+  group_by(statistic>110) %>%
+  slice(1:5) %>%
+  ungroup() %>%
+  # filter(p.value<0.00005) %>%
+  left_join(untarg_feats %>% select(feature, compound_name), by = join_by(feature)) %>%
+  mutate(compound_name=ifelse(is.na(compound_name), feature, compound_name)) %>%
+  mutate(compound_name=ifelse(compound_name=="Dimethylsulfoniopropionate", "DMSP", compound_name)) %>%
+  arrange(statistic) %>%
+  mutate(compound_name=fct_inorder(compound_name)) %>%
+  select(feature, compound_name) %>%
+  left_join(untarg_peaks %>% select(filename, norm_area, feature), by="feature") %>%
+  left_join(filled_file_metadata %>% distinct(filename, station, abs_depth), by="filename") %>%
+  filter(station%in%c("L1", "L2")) %>%
+  arrange(station) %>%
+  mutate(station=ifelse(station=="L1", "Cyclone", "Anticyclone")) %>%
+  mutate(station=fct_inorder(station)) %>%
+  arrange(abs_depth) %>%
+  mutate(filename=fct_inorder(filename)) %>%
+  group_by(feature) %>%
+  # mutate(norm_area=rank(norm_area)) %>%
+  mutate(norm_area=norm_area/sum(norm_area)) %>%
+  ungroup() %>%
+  ggplot() +
+  geom_col(aes(x=filename, y=norm_area, fill=compound_name, group=filename), 
+           position = "fill", color="grey50", width=1) +
+  geom_rect(aes(xmin=-Inf, xmax=Inf, ymin=1, ymax=1.1), 
+            fill=c("#8277b6", "#b05d56"), color="black",
+            data = data.frame(station=fct_inorder(c("Cyclone", "Anticyclone")))) +
+  geom_text(aes(x=8, y=1.05, label=station), hjust=0.5,
+            data = data.frame(station=fct_inorder(c("Cyclone", "Anticyclone")))) +
+  facet_grid(~station, scales="free_x") +
+  scale_fill_manual(values = splividis(10, begin = 0.1, end = 0.95)) +
+  guides(fill=guide_legend(reverse = TRUE)) +
+  scale_y_discrete(expand = c(0,0)) +
+  theme_bw() +
+  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        strip.background = element_blank(), strip.text = element_blank()) +
+  labs(x=NULL, y="Relative peak area", fill="Mass feature")
+ggsave("MC_stacked_mostdiff_gp.png", plot = MC_stacked_mostdiff_gp, device = "png", 
+       path = "supplement", width = 6.5, height = 3, units = "in", dpi = 300)
+```
+
+![](supplement/MC_stacked_mostdiff_gp.png)
+
+*Supplemental figure?
+Particulate matter composition differences between the cyclone and anticyclone when considering only the top 5 most different metabolites enriched in each eddy separately.
+Each sample corresponds to one multicolored vertical bar while metabolites correspond to the colors provided in the legend and have been scaled to the sum of the total peak areas for that metabolite in the sample set.*
+
+### Falkor data reveals different community and response to SLA
+
+Given the strong signals detected in the samples from the MESOSCOPE cruise both across the eddy transect as well as in the high-resolution DCM sampling, we expected to find some of the same results in a new dataset.
+Data was collected during a 2018 cruise on the R/V *Falkor* that again targeted both a cyclonic and anticyclonic eddy in the North Pacific Subtropical Gyre near Station ALOHA.
+
+```{r figure falkor nmds and permanova}
+set.seed(123)
+scaled_fk_data <- untarg_peaks %>%
+  # filter(filename!="190715_Smp_FK180310S62C1-DCM_B.mzML") %>%
+  left_join(filled_file_metadata %>% distinct(ms_run, station, filename, depth), 
+          by="filename") %>%
+  filter(ms_run=="FK") %>%
+  mutate(depth=factor(depth, levels=c("Surface", "DCM"))) %>%
+  select(feature, filename, norm_area) %>%
+  group_by(feature) %>%
+  filter(sum(norm_area==0)<n()*0.9)
+metab_mat_fk <- scaled_fk_data %>%
+  mutate(norm_area=scale(norm_area)[,1]) %>%
+  # mutate(norm_area=rank(norm_area)) %>%
+  pivot_wider(names_from = "feature", values_from = norm_area) %>%
+  column_to_rownames("filename") %>%
+  data.matrix()
+env_vec <- pull(filled_file_metadata, sla_corr, filename)[rownames(metab_mat_fk)]
+nmds_fk <- metab_mat_fk %>%
+  vegan::metaMDS(k = 2, distance = "manhattan", trace = 0, autotransform = FALSE,
+                 noshare = FALSE, wascores = FALSE) %>%
+  vegan::MDSrotate(env_vec)
+adon_expl <- data.frame(filename=rownames(metab_mat_fk)) %>%
+  left_join(filled_file_metadata %>% select(filename, sla_class, depth), by="filename")
+adon_fk <- vegan::adonis2(
+  metab_mat_fk~sla_class+depth, data = adon_expl, by = "margin", 
+  permutations = 999, method = "manhattan")
+
+fk_nmdsplot <- nmds_fk$points %>%
+  as.data.frame() %>%
+  rownames_to_column("filename") %>%
+  left_join(filled_file_metadata %>% distinct(filename, sla_corr, depth)) %>%
+  mutate(MDS2=MDS2*-1) %>%
+  ggplot(aes(label=filename)) +
+  geom_vline(xintercept = 0) +
+  geom_hline(yintercept = 0) +
+  geom_point(aes(x=MDS1, y=MDS2, color=sla_corr, shape=depth), size=4, stroke=2) +
+  scale_color_gradient2(
+    low = scales::muted("blue"), high = scales::muted("red"),  
+    mid = "grey90", aesthetics = "color"
+  ) +
+  scale_shape_manual(values = c(2, 6), breaks = c("Surface", "DCM")) +
+  labs(color="Sea level\nanomaly\n(cm)\n", shape="Depth") +
+  theme_bw() +
+  coord_fixed()
+
+fk_nmds_stress <-  paste("Stress =", round(nmds_fk$stress, digits = 3))
+fk_perm_label <- expression(atop(paste("R"[sla]^2, "=.090, p=0.017"), 
+                                 paste("R"[depth]^2, "=.251, p<0.001")))
+fk_nmdsplot <- fk_nmdsplot +
+  annotate("label", x=-Inf, y=-Inf, label=fk_nmds_stress, hjust=0, vjust=0, 
+           label.r = unit(0, "in")) +
+  annotate("label", x=-Inf, y=Inf, label=fk_perm_label, hjust=0, vjust=1, 
+           label.r = unit(0, "in"))
+
+ggsave("fk_nmdsplot.png", plot = fk_nmdsplot, device = "png",
+       path = "figures", width = 6.5, height = 4, units = "in", dpi = 300,
+       bg="white")
+```
+
+The Falkor data was characterized by high inter-replicate variability relative to the samples from the MESOSCOPE cruise, with surface samples in particular highly variable in multivariate space (Figure 6).
+Despite this, we still saw the importance of sea level anomaly (SLA) as a significant explanatory factor in the dataset (R$^2_{SLA} = 0.090$, p-value = 0.025), after the dominant depth differences between the surface (25m) and DCM (110-120m) (R$^2_{depth} = 0.251$, p-value << 0.001).
+
+![](figures/fk_nmdsplot.png)
+
+*Figure 6: Non-metric multidimensional scaling (NMDS) plot from the Falkor cruise data, in which points correspond to individual samples and have been colored by their corrected sea level height anomaly and shaped by the depth from which they were collected.
+NMDS stress values (s) have been reported in the bottom left corner, while PERMANOVA R$^2$ and p-values are reported in the top left.
+Surface samples are clearly distinct from the DCM samples and an SLA signal is visible at the surface.*
+
+```{r top20compounds stacked barplot, eval=FALSE}
+scaled_fk_data %>%
+  ungroup() %>%
+  arrange(desc(norm_area)) %>%
+  filter(feature%in%head(unique(feature), 20)) %>%
+  left_join(filled_file_metadata) %>%
+  left_join(untarg_feats) %>%
+  mutate(feature=ifelse(is.na(compound_name), feature, compound_name)) %>%
+  mutate(feature=fct_inorder(feature)) %>%
+  arrange(depth, filename) %>%
+  mutate(filename=str_remove(filename, "190715_Smp_FK180310")) %>%
+  mutate(filename=str_remove(filename, "-25m_|-DCM_")) %>%
+  mutate(filename=fct_inorder(filename)) %>%
+  ggplot() +
+  geom_col(aes(x=filename, y=norm_area, fill=feature)) +
+  theme(axis.text.x = element_text(angle=90, hjust = 1, vjust=0.5)) +
+  facet_wrap(~depth, scales="free_y", ncol=1)
+```
+
+When analyzed separately as distinct depths, the surface samples had higher R$^2$ values (R$^2$=.26) and a lower p-value (0.013), while the DCM samples were no longer likely to be distinct between the two eddies (p-value = 0.171) (Supplemental figure XX).
+
+```{r falkor nmdsanova by depth}
+set.seed(123)
+multivar_analysis_FK <- untarg_peaks %>%
+  filter(filename!="190715_Smp_FK180310S62C1-DCM_B.mzML") %>%
+  left_join(filled_file_metadata %>% distinct(ms_run, station, filename, depth), 
+          by="filename") %>%
+  filter(ms_run=="FK") %>%
+  mutate(depth=factor(depth, levels=c("Surface", "DCM"))) %>%
+  nest(data=-depth) %>%
+  mutate(wide_peaks=map(data, function(metab_df){
+    metab_df %>%
+      select(feature, filename, norm_area) %>%
+      group_by(feature) %>%
+      filter(sum(norm_area==0)<n()*0.9) %>%
+      mutate(norm_area=scale(norm_area)[,1]) %>%
+      pivot_wider(names_from = "feature", values_from = norm_area) %>%
+      column_to_rownames("filename") %>%
+      data.matrix()
+  })) %>%
+  mutate(nmds_output=map(wide_peaks, function(metab_mat){
+    env_vec <- pull(filled_file_metadata, sla_corr, filename)[rownames(metab_mat)]
+    vegan::metaMDS(metab_mat, k = 2, distance = "manhattan", trace = 0, 
+                   autotransform = FALSE, noshare = FALSE, wascores = FALSE) %>%
+    vegan::MDSrotate(env_vec)
+  })) %>%
+  mutate(nmds_stress=map_dbl(nmds_output, pluck, "stress")) %>%
+  mutate(perm_output=map(wide_peaks, function(metab_mat){
+    adon_expl <- data.frame(filename=rownames(metab_mat)) %>%
+      left_join(filled_file_metadata %>% select(filename, sla_class), by="filename")
+    vegan::adonis2(
+      metab_mat~sla_class, data = adon_expl, by = "margin", 
+      permutations = 999, method = "manhattan")
+  })) %>%
+  mutate(perm_r2=map_dbl(perm_output, pluck, "R2", 1)) %>%
+  mutate(perm_pval=map_dbl(perm_output, pluck, "Pr(>F)", 1)) %>%
+  mutate(pca_output=map(wide_peaks, prcomp)) %>%
+  mutate(nmds_stress=paste("s =", round(nmds_stress, digits = 3))) %>%
+  mutate(perm_r2=round(perm_r2, digits = 2)) %>%
+  mutate(perm_pval=round(perm_pval, digits = 4)) %>%
+  mutate(perm_label=paste0("R2 = ", perm_r2, "\np = ", perm_pval))
+nmds_sla_gp <- multivar_analysis_FK %>%
+  mutate(nmds_points=map(nmds_output, function(nmds_output_i){
+    nmds_output_i %>% pluck("points") %>% as.data.frame()
+  })) %>%
+  select(depth, nmds_points) %>%
+  unnest_longer(nmds_points, indices_to = "filename") %>%
+  unnest_wider(nmds_points) %>%
+  left_join(filled_file_metadata %>% distinct(filename, sla_corr)) %>%
+  ggplot() +
+  geom_vline(xintercept = 0) +
+  geom_hline(yintercept = 0) +
+  geom_point(aes(x=MDS1, y=MDS2, fill=sla_corr), color="black", size=4, pch=21) +
+  facet_wrap(~depth, nrow=1) +
+  scale_fill_gradient2(
+    low = scales::muted("blue"), high = scales::muted("red"),  mid = "grey90",
+    aesthetics = c("color", "fill"),
+  ) +
+  labs(fill="Sea level\nanomaly\n(cm)\n") +
+  theme_bw() +
+  coord_fixed()
+
+med_metab_df <- untarg_peaks %>%
+  left_join(filled_file_metadata %>% distinct(filename, ms_run, station, depth, sla_corr),
+            by="filename") %>%
+  filter(ms_run=="FK") %>%
+  group_by(depth, feature) %>%
+  filter(sum(norm_area==0)<n()*0.9) %>%
+  mutate(norm_area=scale(norm_area)[,1]) %>%
+  ungroup() %>%
+  group_by(filename, depth, station, sla_corr) %>%
+  summarize(avg_amt=mean(norm_area)) %>%
+  ungroup()
+# Calculations for correlation between grand median metabolite and SLA
+# med_metab_df %>%
+#   group_by(station, depth, sla_corr) %>%
+#   summarise(grand_avg_amt=mean(avg_amt)) %>%
+#   group_by(depth) %>%
+#   summarise(sim=cor(sla_corr, grand_avg_amt))
+med_metab_sla_gp <- med_metab_df %>%
+  mutate(station=factor(station, levels=c("62", "64", "77", "80"))) %>%
+  mutate(depth=factor(depth, levels=c("Surface", "DCM"))) %>%
+  ggplot() +
+  geom_hline(yintercept = 0) +
+  geom_boxplot(aes(x=station, y=avg_amt, color=sla_corr), linewidth=0.8) +
+  facet_wrap(~depth, nrow=1) +
+  scale_color_gradient2(low = scales::muted("blue"), high = scales::muted("red"),
+                        mid="grey90", aesthetics = c("color", "fill"), guide = NULL) +
+  theme_bw() +
+  theme(strip.background = element_blank(), strip.text = element_blank()) +
+  labs(x="Station", y="Median metabolite z-score")
+
+library(egg)
+nolegend_plot <- ggarrange(nmds_sla_gp+theme(legend.position = "none"), 
+                           med_metab_sla_gp, ncol = 1, heights = c(0.7, 0.3), 
+                           draw = FALSE)
+library(cowplot)
+fk_nmds_and_med_metab <- plot_grid(nolegend_plot, get_legend(nmds_sla_gp), 
+                                nrow=1, rel_widths = c(0.85, 0.15))
+
+ggsave("fk_nmds_and_med_metab.png", plot = fk_nmds_and_med_metab, device = "png",
+       path = "supplement", width = 6.5, height = 5.5, units = "in", dpi = 300,
+       bg="white")
+```
+
+![](supplement/fk_nmds_and_med_metab.png)
+
+*Supplemental figure(?): Plots of the metabolome in the Falkor data, broken down by depth.
+The top row of plots are non-metric multidimensional scaling (NMDS) plots, in which points correspond to individual samples and have been colored by their corrected sea level height anomaly.
+SLA trends are visible in the surface samples, with dark blue circles consistently discriminating from the dark red circles.
+The bottom row of plots show the direction and magnitude of this effect by plotting the grand mean of the normalized metabolite peak areas in the stations taken across the eddy dipole.*
+
+```{r falkor concentrations comparison to MESOSCOPE, eval=FALSE}
+fk_targ <- targ_peaks %>% 
+  left_join(filled_file_metadata) %>%
+  filter(ms_run%in%c("FK", "MT")) %>%
+  filter(samp_type=="Smp") %>%
+  select(compound_name, filename, nM, depth, ms_run)
+fk_targ %>%
+  group_by(depth, compound_name, ms_run) %>%
+  summarise(nM=mean(nM)) %>%
+  arrange(desc(nM)) %>%
+  pivot_wider(names_from = c(ms_run, depth), values_from = nM) %>%
+  print(n=Inf)
+cmpd_comp_gp <- fk_targ %>%
+  filter(depth!="Deep") %>%
+  ggplot() +
+  geom_boxplot(aes(x=ms_run, y=nM)) +
+  facet_grid(compound_name~depth, scales="free_y")
+ggsave("cmpd_comp_gp.pdf", cmpd_comp_gp, device = "pdf", width = 6, height = 50, limitsize = FALSE)
+# Visual assessment of boxplots from the above output:
+# Compounds more abundant in Falkor: citrulline, aspartic acid, ornithine, threonine, tyrosine, malic acid
+# Compounds more abundant in MESO: 4-aminobutyric, adenine, gonyol, 4-hydroxyisoleucine, alanine, leucine, nicotinic acid
+```
+
+Overall, we detected several compounds that were distinctly more or less abundant in the Falkor dataset than the MESOSCOPE one.
+Compounds more abundant in Falkor were the amino acids aspartate, threonine, and tyrosine as well as the components of the urea cycle citrulline and ornithine and the TCA intermediate malic acid.
+Compounds less abundant in Falkor were adenine, gonyol, hydroxyisoleucine, leucine, alanine, and nicotinic acid.
+
+  - Discussion: is there a common thread between the up/down regulated compounds?
+Urea cycle for Falkor cruise w aspartate/citrulline/ornithine/malic acid?
+Connect back to the differing community composition of the two cruises?
+Any chance we have urea numbers for Falkor at all?
+Ask about those transcripts?
+I want to connect this to the diatom urea cycle paper (10.1038/nature10074) and imply that during Falkor there was an ornithine-urea cycle disruption (10.1038/s41598-020-73715-8) but there's very few diatoms in Falkor???
+Does the same thing happen in haptophytes?
+It's not like this is different within the cruise but across cruises still seems interesting.
+
+We were also able to use the Falkor dataset to test whether the compounds detected as significantly different in the MESOSCOPE dataset were also different in this eddy pair.
+We expected to find abundant TMAO and more hydroxyisoleucine at the surface than the DCM in addition to most compounds enriched in the cyclone DCM with biomass, especially the six compounds mentioned above in the results of Figure 5 (trigonelline, homarine, DMSAc, DMSP, taurine, isethionic acid).
+
+```{r abundant TMAO and hydroxyisoleucine depth trends, eval=FALSE}
+fk_targ <- targ_peaks %>% 
+  left_join(filled_file_metadata) %>%
+  filter(ms_run%in%c("FK")) %>%
+  filter(samp_type=="Smp") %>%
+  select(compound_name, filename, nM, depth, ms_run)
+
+fk_targ %>%
+  filter(compound_name=="Hydroxyisoleucine") %>%
+  left_join(filled_file_metadata) %>%
+  aov(formula=nM~depth) %>% summary()
+  # ggplot(aes(x=depth, y=nM, color=station)) +
+  # geom_jitter(width=0.2)
+  # geom_boxplot()
+
+fk_targ %>%
+  filter(compound_name%in%c("Homarine", "Trigonelline", "Dimethylsulfonioacetate", "Dimethylsulfoniopropionate", "Taurine", "Isethionic acid", "Arsenobetaine")) %>%
+  left_join(filled_file_metadata) %>%
+  group_by(filename) %>%
+  # mutate(nM=nM/sum(nM)) %>%
+  ungroup() %>%
+  filter(depth=="DCM") %>%
+  # nest(data=-compound_name) %>%
+  # mutate(aovout=map(data, ~broom::tidy(aov(.x$nM~.x$sla_class)))) %>%
+  # unnest(aovout) %>% filter(term==".x$sla_class") %>%
+  # mutate(p.value=p.adjust(p.value, method="fdr"))
+  ggplot(aes(x=sla_class, y=nM)) +
+  # geom_boxplot() +
+  geom_point() +
+  facet_wrap(~compound_name, scales="free_y")
+
+untarg_feats %>%
+  filter(mzmed%between%pmppm(170.11756, 5) | mzmed%between%pmppm(173.09211, 5)) %>%
+  filter(startsWith(feature, "FK")) %>% 
+  left_join(untarg_peaks) %>%
+  left_join(filled_file_metadata) %>%
+  nest(data=-feature) %>%
+  mutate(aovout=map(data, ~broom::tidy(aov(.x$norm_area~.x$sla_class)))) %>%
+  unnest(aovout) %>% filter(term==".x$sla_class") %>%
+  mutate(p.value=p.adjust(p.value, method="fdr"))
+  select(feature, sla_class, norm_area) %>% 
+  pivot_wider(names_from = sla_class, values_from = norm_area, values_fn = mean)
+  ggplot(aes(x=sla_class, y=norm_area)) +
+  # geom_boxplot() +
+  geom_point() +
+  facet_wrap(~feature, scales="free_y")
+```
+
+As expected, TMAO was again one of the most abundant metabolites detected in the particulate matter with concentrations around 1.2 nM at the surface and 0.4 nM at the DCM, second only to the surprisingly high levels of the amino acid glycine (2 nM surface concentrations, 0.9 nM DCM concentrations), though the differences between the surface and DCM were not significant for either compound.
+We did detect a significant difference between depths for hydroxyisoleucine, once again expected, despite the aforementioned lower concentrations in this sample set (surface mean = 0.32 nM, DCM mean = 0.21 nM, t-test p-value=0.022 with n=12 samples at each depth).
+
+The eddy effects, on the other hand, were much less strong during this cruise.
+Of the six metabolites with the strongest differences across the eddy dipole at the DCM in the MESOSCOPE cruise, only isethionic acid was significantly different in Falkor (isethionic acid t-test p-value$_{FDR}$ = 0.043, other p-values > 0.25).
+Notably, this difference in isethionic acid concentration was only found after normalizing each sample to the sum of all metabolites in the sample to control for biomass.
+Arsenobetaine, found in the MESOSCOPE samples to be strongly enriched in the anticyclone, was again found to have the same pattern here even at the DCM both when normalizing to biomass and when not.
+Surprisingly, the 173.09211 *m/z* mass feature was also slightly enriched but this time in the *cyclone* DCM with peak areas approximately 1.4 times the values in the anticyclone (t-test p-value = 0.067, Mann-Whitney p-value = 0.045).
+The 170.11756 *m/z* mass feature was not detected in the Falkor samples.
+
+  - Discussion: differences between this and MESOSCOPE could be due in large part to the different community composition here - MESOSCOPE looked like a big diatom bloom (according to Benedetto paper) while Sam's results show that this was more dinoflagellate/haptophyte and diatoms had "no trend".
+Maybe also eddy age?
+  - Discussion: Surface more variable than DCM in Falkor (kinda also for MESOSCOPE?)
+  - Discussion: What the hell is with these samples.
+There's an insane amount of variance within triplicates and the SLA signal is highest at the *surface* despite Sam's data and the MESOSCOPE dataset strongly implying the opposite should be true - although if a lot of the metabolism is driven by haptophytes then those are much more abundant at the surface?

### 2023-09-01 manuscript/manuscript.Rmd
Adding additional sections
Added most of an introduction
Feels maybe worse than my previous one but I feel more ownership over this and it's more SLA-specific

@@
 
+## Abstract
+
+## Introduction
+
+Until recently, the ocean's subtropical gyres were considered homogenous and static [@Karl1999].
+In the past 25 years, however, regular sampling and advances in measurement technology have revealed its variability on time scales from the diel to the seasonal [@Karl2017; @Karl2021].
+A major source of this variability comes in the form of mesoscale eddies in which water is entrained into circular surface currents tens to hundreds of kilometers in diameter [@McGillicuddy2016].
+Prior research has established the importance of these mesoscale features in introducing nutrients into the euphotic zone and altering the community composition [@Sweeney2003; @Barone2019].
+In particular, the deep chlorophyll maximum (DCM) is known to correspond to the top of the nitricline and differs dramatically between cyclonic and anticyclonic eddies in depth and intensity.
+@Barone2022 reported a shift in the microbial DCM community with eukaryotic phytoplankton enriched and prokaryotic phytoplankton depleted in cyclonic eddies with the converse true for anticyclonic ones, while @Hawco2021 and @Dugenne2023 showed that some of this shift was due to alterations in dissolved iron chemistry and nitrogen fixation.
+
+These recent works have highlighted the effect of eddies on both bulk oceanographic parameters such as particulate carbon and nitrogen and the composition of the community, but we still lack direct links between the genomic and transcriptomic assays occurring regularly around the world and the particulate organic matter produced by those organisms [@Kharbush2020].
+This missing link is bridged by metabolomics, a nascent field studying the structure and function of the cell on a molecular level.
+Of particular interest within the metabolome are those small, polar molecules that represent the fundamental unit of cell biology and function as building blocks for macromolecules [@Moran2022].
+These small molecules also have roles of their own within the cell, acting as osmolytes, antioxidants, nutrient storage compounds, and more [@Geraldes2021; @VanMooy2009; @Mojzes2020; @Kharbush2020].
+Despite these critical functions, they are difficult to study in the marine environment due to their low concentrations and similarity to the abundant salt ions, resulting in challenges in detection and quantification [@Boysen2018; @Moran2022; @Sacks2022].
+
+  - Important to do because the actual composition of particles determines fate
+  - Reference my own paper somewhere too
+
+Here, we use spatially resolved measurements of metabolites in particles across an eddy dipole in the North Pacific Subtropical Gyre (NPSG) to categorize unidentified molecular features based on their co occurrence with a set of simultaneously measured physical, chemical, and biological parameters.
+The gyre in particular represents an intriguing opportunity for metabolomic research because it has been shown to contain the highest proportion of chemical features with unknown structure and function and have highly variable responses changes in depth and across diel cycles [@Heal2021; @Boysen2021].
+This study connects biogeochemical measurements to a molecular-scale perspective of particulate organic matter, linking the chemical currencies that fuel metabolism and interactions among marine microbes to the environment in which they’re found.
+
+  - We identify hundreds of known and unknown molecular features of interest that may relate to the biogeochemical functioning of ocean gyres.
+
+## Methods
+## Results
 ### Metabolome changes significantly across the eddy dipole at DCM and 175m
@@
 There's an insane amount of variance within triplicates and the SLA signal is highest at the *surface* despite Sam's data and the MESOSCOPE dataset strongly implying the opposite should be true - although if a lot of the metabolism is driven by haptophytes then those are much more abundant at the surface?
+
+## Discussion

### 2023-09-06 manuscript/manuscript.Rmd
Fixing up intro, adding section headers, creating supplement, ce

--- 2023-08-29 manuscript/manuscript.Rmd
@@
 ---
-title: "MESOSCOPE manuscript results/figures"
+title: "MESOSCOPE manuscript"
 author: "William Kumler"
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, eval = TRUE)
+knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, eval = FALSE)
 library(tidyverse)
@@
 
-### Metabolome changes significantly across the eddy dipole at DCM and 175m
+# Abstract
+
+# Introduction
+
+Until recently, the ocean's subtropical gyres were considered homogenous and static [@Karl1999].
+In the past 25 years, however, regular sampling and advances in measurement technology have revealed its variability on time scales from the diel to the seasonal [@Karl2017; @Karl2021].
+A major source of this variability comes in the form of mesoscale eddies in which water is entrained into circular surface currents tens to hundreds of kilometers in diameter [@Karl2017; @McGillicuddy2016].
+Prior research has established the importance of these mesoscale features in introducing nutrients into the euphotic zone and altering the community composition [@Sweeney2003; @Barone2019].
+In particular, the deep chlorophyll maximum (DCM) is known to correspond to the top of the nitricline and differs dramatically between cyclonic and anticyclonic eddies in depth and intensity.
+@Barone2022 reported a shift in the microbial DCM community with eukaryotic phytoplankton enriched and prokaryotic phytoplankton depleted in cyclonic eddies with the converse true for anticyclonic ones, while @Hawco2021 and @Dugenne2023 showed that some of this shift was due to alterations in dissolved iron chemistry and nitrogen fixation, especially at the surface.
+
+These recent works have highlighted the effect of eddies on both bulk oceanographic parameters such as particulate carbon and nitrogen and the composition of the microbial community, but we still lack direct links between the genomic and transcriptomic assays occurring regularly around the world and the particulate organic matter produced by those organisms [@Kharbush2020].
+This missing link is bridged by metabolomics, a nascent field studying the structure and function of the cell on a molecular level.
+Of particular interest within the metabolome are those small, polar molecules that represent the fundamental unit of cell biology and function as building blocks for macromolecules [@Moran2022].
+These small molecules also have roles of their own within the cell, acting as osmolytes, antioxidants, nutrient storage compounds, and more [@Geraldes2021; @Mojzes2020; @Kharbush2020].
+Despite these critical functions, they are difficult to study in the marine environment due to their low concentrations and similarity to the abundant salt ions, resulting in challenges in detection and quantification [@Boysen2018; @Moran2022; @Sacks2022].
+
+Here, we use spatially resolved measurements of metabolites in particles across an eddy dipole in the North Pacific Subtropical Gyre (NPSG) to categorize unidentified molecular features based on their co occurrence with a set of simultaneously measured physical, chemical, and biological parameters.
+The gyre in particular represents an intriguing opportunity for metabolomic research because it has been shown to contain the highest proportion of chemical features with unknown structure and function and have highly variable responses to changes in latitude, depth, and across diel cycles [@Heal2021; @Boysen2021; @Kumler2023].
+This study connects biogeochemical measurements to a molecular-scale perspective of particulate organic matter, linking the chemical currencies that fuel metabolism and interactions among marine microbes to the mesoscale environment in which they’re found.
+
+# Results
+## Metabolome changes significantly across the eddy dipole at DCM and 175m
 
@@
 Notably, the samples taken from the exact center of the cyclonic eddy were highly distinct and likely drove much of the variance explained.
-In the deep samples, 12% of the variance was explained by SLA (p = 0.008) according to the PERMANOVA and SLA was highly correlated with the second PC of the dataset (r = .697, % variance explained by PC2 = 15.4), though the first PC did not seem to have any visible pattern with metadata and appeared to capture largely variation between biological triplicates.
+In the deep samples, 12% of the variance was explained by SLA (p = 0.008) according to the PERMANOVA and SLA was highly correlated with the second PC of the dataset (r = .697, % variance explained by PC2 = 15.4), though the first PC did not seem to have any visible pattern with metadata and appeared to capture largely variation between biological triplicates (Supplemental figure 1).
 At the surface, SLA trends are less evident with a larger p-value (0.087) and a lower R$^2$ (0.09) (Figure 1).
@@
 
-```{r supp figure/table prcomp vs SLA}
-pcdata <- multivar_analysis %>%
-  select(depth, pca_output) %>%
-  mutate(pca_x=map(pca_output, function(pca_output_i){
-    pca_output_i$x[,1:3] %>% as.data.frame()
-  })) %>%
-  select(depth, pca_x) %>%
-  unnest_longer(pca_x, indices_to = "filename") %>%
-  unnest_wider(pca_x) %>%
-  left_join(filled_file_metadata %>% distinct(filename, sla_corr)) %>%
-  pivot_longer(starts_with("PC"))
-pcvars <- multivar_analysis %>%
-  select(depth, pca_output) %>%
-  mutate(perc_expl=map(pca_output, function(pca_output_i){
-    var_expl <- pca_output_i$sdev^2
-    data.frame(perc_expl=var_expl[1:3]/sum(var_expl), name=paste0("PC", 1:3))
-  })) %>%
-  unnest(perc_expl) %>%
-  mutate(perc_label=paste0(round(perc_expl, digits = 3)*100, "%"))
-
-# Correlation between PCs and SLA at each depth
-# pcdata %>%
-#   nest(data=-c(depth, name)) %>%
-#   mutate(cor_output=map_dbl(data, ~abs(cor(.x$value,.x$sla_corr))))
-
-# Linear model significance testing and parameter report
-# pcdata %>%
-#   nest(data=-c(depth, name)) %>%
-#   mutate(lm_output=map(data, ~broom::tidy(lm(.x$value~.x$sla_corr)))) %>%
-#   unnest(lm_output) %>%
-#   filter(term==".x$sla_corr") %>%
-#   arrange(name, depth) %>%
-#   select(depth, name, p.value) %>%
-#   split(.$name)
-
-lm_gp <- ggplot(pcdata, aes(x=sla_corr, y=value)) +
-  geom_smooth(method="lm", formula=y~x, color="black") +
-  geom_point(aes(fill=sla_corr), pch=21, color="black") +
-  geom_label(data = pcvars, aes(label=perc_label),
-             hjust=1, vjust=1, x=Inf, y=Inf, label.r = unit(0, "in")) +
-  facet_grid(depth~name, switch = "y") +
-  scale_y_continuous(position = "right") +
-  scale_fill_gradient2(
-    low = scales::muted("blue"), high = scales::muted("red"),  mid = "grey90",
-    aesthetics = c("color", "fill"), guide = NULL
-  ) +
-  labs(x="Corrected SLA, in cm", y="Principal component value",
-       color="Corrected\nSLA (cm)") +
-  theme_bw()
-deep_pc1_barplot_gp <- pcdata %>%
-  left_join(filled_file_metadata %>% distinct(filename, station, tripl)) %>%
-  mutate(station=factor(station, levels=as.character(4:14),
-                        labels=paste0("St.", 4:14))) %>%
-  arrange(station) %>%
-  filter(name=="PC1" & str_detect(depth, "Deep")) %>%
-  ggplot() +
-  geom_hline(yintercept = 0) +
-  geom_col(aes(x=tripl, y=value)) +
-  facet_wrap(~station, nrow = 1, scales="free_x") +
-  scale_y_continuous(position = "right") +
-  labs(x="Triplicate", y="PC1", title = paste(
-    "    Variance among deep (175 meter) samples in PC1"
-  ))
-pca_sla_cor_gp <- cowplot::plot_grid(lm_gp, deep_pc1_barplot_gp, 
-                                     rel_heights = c(0.6, 0.4), 
-                                     ncol = 1, labels = "AUTO")
-ggsave("pca_sla_cor_gp.png", pca_sla_cor_gp, device = "png", path = "supplement",
-       width = 6.5, height = 6.5, units = "in", dpi = 300)
-
-```
-
-![](supplement/pca_sla_cor_gp.png)
-
-*Supplemental figure: Results of a principal component analysis performed separately on each depth of the eddy transect metabolome.
-A) Regressions of SLA against various principal components, broken down by depth.
-% variance explained by each PC is noted in the upper right corner of the plot.
-Colors have been added corresponding to the SLA value for clarity but are redundant with the x-axis values.
-Lines of best fit and the associated standard error have been added behind the points in black and translucent grey, respectively.
-Strong associations exist between PC1 and SLA in the DCM samples and PC2 and SLA in the deep samples.
-B) Principal component 1 values for the 175 meter samples at each station showing large variance between triplicates rather than any external factor.*
-
 A large part of this metabolic response to SLA is due to the increased biomass in the cyclone DCM and the parallel increase in the anticyclone deep samples.
@@
 
-```{r supp figure fraction of PC in metabolites}
+```{r PC vs metab abund anovas and perm test, eval=FALSE}
 pc_comp_df <- targ_peaks %>%
@@
 # Significance testing between depths and SLA classes for % PC in metab
-# pc_comp_df %>% aov(formula=frac_in_metab~depth) %>% TukeyHSD()
-# pc_comp_df %>% aov(formula=frac_in_metab~sla_class) %>% TukeyHSD()
+pc_comp_df %>% aov(formula=frac_in_metab~depth) %>% TukeyHSD()
+pc_comp_df %>% aov(formula=frac_in_metab~sla_class) %>% TukeyHSD()
 
 # Significance testing for differences in % in metabs as a function of SLA by depth
-# pc_comp_df %>%
-#   nest(data=-depth) %>%
-#   mutate(aov_out=map(data, ~broom::tidy(aov(.x$frac_in_metab~.x$sla_class)))) %>%
-#   mutate(kw_out=map(data, ~broom::tidy(kruskal.test(.x$frac_in_metab~.x$sla_class)))) %>%
-#   pivot_longer(c(aov_out, kw_out)) %>%
-#   unnest(value) %>%
-#   drop_na(p.value) %>%
-#   select(depth, name, p.value) %>%
-#   group_by(name) %>%
-#   mutate(p.value=p.adjust(p.value, method="fdr")) %>%
-#   ungroup() %>%
-#   arrange(name, depth)
-# pc_comp_df %>%
-#   ggplot(aes(x=sla_class, y=frac_in_metab)) +
-#   geom_boxplot() +
-#   facet_wrap(~depth, nrow=1)
+pc_comp_df %>%
+  nest(data=-depth) %>%
+  mutate(aov_out=map(data, ~broom::tidy(aov(.x$frac_in_metab~.x$sla_class)))) %>%
+  mutate(kw_out=map(data, ~broom::tidy(kruskal.test(.x$frac_in_metab~.x$sla_class)))) %>%
+  pivot_longer(c(aov_out, kw_out)) %>%
+  unnest(value) %>%
+  drop_na(p.value) %>%
+  select(depth, name, p.value) %>%
+  group_by(name) %>%
+  mutate(p.value=p.adjust(p.value, method="fdr")) %>%
+  ungroup() %>%
+  arrange(name, depth)
+pc_comp_df %>%
+  ggplot(aes(x=sla_class, y=frac_in_metab)) +
+  geom_boxplot() +
+  facet_wrap(~depth, nrow=1)
 
 # Permutation testing for effect size test on Deep SLA differences
-# set.seed(123)
-# map_dbl(seq(1, 2, 0.1), function(n_mult){
-#   map_dbl(1:100, function(h){
-#     pc_comp_df %>%
-#       filter(depth=="Deep") %>%
-#       group_by(sla_class) %>%
-#       summarize(mean=mean(frac_in_metab), sd=sd(frac_in_metab), n=n()) %>%
-#       mutate(sim_vals=map(1:3, function(i){rnorm(.$n[i]*n_mult, .$mean[i], .$sd[i])})) %>%
-#       unnest_longer(sim_vals) %>%
-#       aov(formula=sim_vals~sla_class) %>%
-#       broom::tidy() %>%
-#       filter(term=="sla_class") %>%
-#       pull(p.value)
-#   }) %>% mean()
-# })
-
-pc_frac_metab_gp <- ggplot(pc_comp_df, aes(x=station, y=frac_in_metab, color=sla_corr)) +
-  geom_hline(yintercept = 0) +
-  geom_boxplot(width = 0.9, linewidth=0.8) +
-  geom_point(color="black", size=2) +
-  scale_y_continuous(labels = scales::label_percent()) +
-  facet_grid(~depth, scales="free") +
-  scale_color_gradient2(
-    low = scales::muted("blue"), high = scales::muted("red"),  mid = "grey90",
-    aesthetics = c("color", "fill")
-  ) +
-  guides(color=guide_colorbar(direction = "horizontal", title.position = "top", 
-                              title.hjust = 0.5)) +
-  labs(color="Sea level\nanomaly (cm)", x="Station",
-       y="Fraction of PC represented by known metabolites") +
-  theme_bw() +
-  theme(legend.position = c(0.85, 0.8), legend.background = element_rect(color="grey90"))
-ggsave("pc_frac_metab_gp.png", pc_frac_metab_gp, device = "png", path = "supplement",
-       width = 6.5, height = 3.5, units = "in", dpi = 300)
-```
-
-![](supplement/pc_frac_metab_gp.png)
-
-*Supplemental figure: Fraction of particulate carbon represented by known metabolites.
-Boxplots set behind points are colored according to sea level anomaly (SLA) with red corresponding to a positive SLA (anticyclone) and blue corresponding to negative SLA (cyclone).*
+set.seed(123)
+map_dbl(seq(1, 2, 0.1), function(n_mult){
+  map_dbl(1:100, function(h){
+    pc_comp_df %>%
+      filter(depth=="Deep") %>%
+      group_by(sla_class) %>%
+      summarize(mean=mean(frac_in_metab), sd=sd(frac_in_metab), n=n()) %>%
+      mutate(sim_vals=map(1:3, function(i){rnorm(.$n[i]*n_mult, .$mean[i], .$sd[i])})) %>%
+      unnest_longer(sim_vals) %>%
+      aov(formula=sim_vals~sla_class) %>%
+      broom::tidy() %>%
+      filter(term=="sla_class") %>%
+      pull(p.value)
+  }) %>% mean()
+})
+```
 
 This correlation between PC and overall metabolite abundance implied that the fraction of PC represented by known metabolites was likely to be fixed across the eddy transect and this expectation was validated by followup analyses.
-Known metabolites contributed between approximately 1.5% and 5% of the carbon biomass to the total PC load, but this variance was not clearly due to either depth or SLA.
+Known metabolites contributed between approximately 1.5% and 5% of the carbon biomass to the total PC load, but this variance was not clearly due to either depth or SLA (Supplemental figure 2).
 The only difference we observed was a slightly reduced metabolite contribution in the 175 meter samples (0.3-0.4 percentage points lower at depth, p-value=0.018) while no difference was detected between the surface and DCM samples or between the cyclone, anticyclone, and outside the eddy samples (all p-values > 0.1).
@@
 
-```{r lm_line_rank_order supplemental figure}
-reg_coefs <- untarg_peaks %>%
-  left_join(filled_file_metadata %>% distinct(filename, ms_run, station, depth, sla_corr)) %>%
-  filter(ms_run=="MT") %>%
-  group_by(feature) %>%
-  filter(sum(norm_area==0)<n()*0.9) %>%
-  mutate(norm_area=rank(norm_area)) %>%
-  ungroup() %>%
-  nest(data=-c(depth, feature)) %>%
-  mutate(test=map(data, ~broom::tidy(lm(.x$norm_area~.x$sla_corr)))) %>%
-  unnest(test)
-reg_coefs %>%
-  filter(str_detect(term, "sla_corr")) %>%
-  mutate(p.value=p.adjust(p.value, method="fdr")) %>%
-  arrange(p.value) %>%
-  left_join(untarg_feats) %>%
-  select(feature, depth, estimate, p.value, mzmed, rtmed, compound_name) %>%
-  filter(p.value<0.05) %>%
-  mutate(depth=factor(depth, levels=c("Surface", "DCM", "Deep"))) %>%
-  mutate(rtmed=round(rtmed, digits = 2)) %>%
-  mutate(mzmed=round(mzmed, digits = 5)) %>%
-  arrange(depth, p.value) %>%
-  group_by(depth) %>%
-  group_modify(~ add_row(.x, .after = nrow(.))) %>%
-  mutate(across(everything(), ~ifelse(is.na(.x), "", .x))) %>%
-  write_csv("supplement/lm_est_rank_order.csv")
-
-clean_slopes <- reg_coefs %>%
-  select(feature, data, depth, term, estimate, p.value) %>%
-  mutate(term=ifelse(term==".x$sla_corr", "slope", "intercept")) %>%
-  pivot_wider(names_from = term, values_from=c(p.value, estimate)) %>%
-  mutate(p.value=p.adjust(p.value_slope, method="fdr")) %>%
-  select(-p.value_intercept, -p.value_slope) %>%
-  filter(p.value<0.05) %>%
-  mutate(pointcol=ifelse(sign(estimate_slope)==1, "Anticyclone enriched", "Cyclone enriched")) %>%
-  mutate(depth=factor(depth, levels=c("Surface", "DCM", "Deep")))
-metadata_slopes <- filled_file_metadata %>%
-  filter(ms_run=="MT") %>%
-  filter(samp_type=="Smp") %>%
-  select(depth, pressure:beam_atten) %>%
-  distinct() %>%
-  pivot_longer(-c(sla_corr, depth)) %>%
-  group_by(name) %>%
-  mutate(value=rank(value)*3) %>%
-  nest(data=-c(name, depth)) %>%
-  mutate(lmout=map(data, ~broom::tidy(lm(.x$value~.x$sla_corr)))) %>%
-  unnest(lmout) %>%
-  filter(name!="dcm_depth") %>%
-  select(depth, name, data, term, estimate, p.value) %>%
-  pivot_wider(names_from = term, values_from = c(estimate, p.value)) %>%
-  mutate(p.value=p.adjust(`p.value_.x$sla_corr`)) %>%
-  filter(p.value<0.05) %>%
-  select(-c(`p.value_(Intercept)`, `p.value_.x$sla_corr`)) %>%
-  mutate(intercept=`estimate_(Intercept)`, slope=`estimate_.x$sla_corr`) %>%
-  mutate(pointcol=ifelse(sign(slope)==1, "Anticyclone enriched", "Cyclone enriched")) %>%
-  mutate(depth=factor(depth, levels=c("Surface", "DCM", "Deep"))) %>%
-  mutate(name=factor(name, levels=metadata_levels, labels=metadata_short_labels))
-ggplot() +
-  geom_abline(aes(intercept=estimate_intercept, slope=estimate_slope), data = clean_slopes) +
-  geom_point(aes(x=sla_corr, y=norm_area), data=unnest(clean_slopes, data)) +
-  geom_abline(data = metadata_slopes, aes(slope=slope, intercept=intercept, color=name)) +
-  geom_point(aes(x=sla_corr, y=value, color=name), data=unnest(metadata_slopes, data)) +
-  facet_grid(pointcol~depth) +
-  ggtitle("Todo: fix this graph")
-```
-
-*Supplemental figure: Linear regressions of rank-order peak area against SLA showing strong enrichment at the cyclone DCM and the anticyclone 175 meter samples.*
-
-### Compositional shifts in metabolome as well as bulk biomass
+## Compositional shifts in metabolome as well as bulk biomass
 
@@
 The most significantly different compounds at the DCM in each direction, however, were both unknowns.
-The lowest p-value ($6.4*10^{-6}$ after FDR correction) in the DCM data was a mass feature with *m/z* of 173.09211, a retention time also around 8.6 minutes, and enrichment in the anticyclone with a putative chemical formula of [M+H] = C$_7$H$_{13}$N$_2$O$_3$, possibly glyclproline or prolylglycine.
+The lowest p-value ($6.4*10^{-6}$ after FDR correction) in the DCM data was a mass feature with *m/z* of 173.09211, a retention time also around 8.6 minutes, and enrichment in the anticyclone with a putative chemical formula of [M+H] = C$_7$H$_{13}$N$_2$O$_3$, possibly glycylproline or prolylglycine.
 The next-lowest (p = $2.5*10^{-4}$ after FDR correction) was enriched in the cyclone and had an *m/z* of 275.07119 and a retention time of 11.4 minutes with a putative formula of C$_{18}$H$_{11}$O$_3$.
@@
 - Discussion: all the nucleobases upregulated in the 175m anticyclone, why? Bacteria?
-
-```{r possible figure stacked barplot ranked space}
-top10_stacked_area_station_gp <- map(c("Surface", "DCM", "Deep"), function(depth_i){
-  reg_coefs %>%
-    filter(str_detect(term, "sla_corr")) %>%
-    mutate(p.value=p.adjust(p.value, method="fdr")) %>%
-    filter(depth==depth_i) %>%
-    arrange(p.value) %>%
-    group_by(estimate>0) %>%
-    slice(1:5) %>%
-    ungroup() %>%
-    left_join(untarg_feats %>% select(feature, compound_name, mzmed, rtmed)) %>%
-    mutate(feature=ifelse(!is.na(compound_name), compound_name, 
-                          paste0(round(mzmed, 3), "_", round(rtmed, 1)))) %>%
-    mutate(feature=fct_inorder(feature)) %>%
-    unnest(data) %>%
-    select(feature, depth, norm_area, filename, station) %>%
-    group_by(feature) %>%
-    # mutate(norm_area=rank(norm_area)) %>%
-    mutate(norm_area=norm_area/sum(norm_area)) %>%
-    mutate(station=factor(station, levels=as.character(4:14))) %>%
-    ggplot() +
-    geom_col(aes(x=station, y=norm_area, fill=feature), position = "fill") +
-    scale_y_continuous(expand = c(0,0)) +
-    labs(x="Station", y="Relative peak area", fill="Mass feature") +
-    theme_bw()
-}) %>% gridExtra::arrangeGrob(grobs = ., ncol = 1)
-ggsave("top10_stacked_area_station_gp.png", plot = top10_stacked_area_station_gp, 
-       device = "png", path = "supplement", width = 6.5, height = 9, 
-       units = "in", dpi = 300)
-```
-
-![](supplement/top10_stacked_area_station_gp.png)
-
-*Supplemental figure?
-Plots of the 5 most significantly enriched metabolites in both the cyclone (top 5) and anticyclone (bottom 5) after normalizing to the total peak area in each mass feature.
-Nicely shows a stronger trend with increasing depth but feels a little bit like lying because there's 100 other compounds that don't change in between these extremes.
-Tyrosine is also misannotated here.*
 
@@
 
-### Eddy center DCM samples show differences between depths and eddy polarities
+## Eddy center DCM samples show differences between depths and eddy polarities
 
@@
 We found that a majority of the metabolites (Clusters 1 and 2 in Figure 4, 63% of the total) fell into clusters with larger peak areas in the cyclone, as expected, while also detecting 32 metabolites that clustered such that the mean metabolite was enriched in the anticyclone (Cluster 4, Figure 4).
-Cluster #2 notably had a distinct decrease in relative peak area with depth, while the other clusters had much less clear depth trends.
-Cluster #1 also showed a neat DCM maximum for the samples from the anticyclone, correlating well with flow cytometry counts of picoeukaryotes (r=.851) that was not present in the cyclone (r=.013).
+Cluster 2 notably had a distinct decrease in relative peak area with depth, while the other clusters had much less clear depth trends.
+Cluster 1 also showed a neat DCM maximum for the samples from the anticyclone, correlating well with flow cytometry counts of picoeukaryotes (r=.851) that was not present in the cyclone (r=.013).
 
@@
 
-```{r comparison to Katherines culture data}
-cyc_enriched <- pairwise_tests %>%
-  filter(p.value<0.05) %>%
-  left_join(untarg_feats) %>%
-  filter(!is.na(compound_name)) %>%
-  filter(statistic>110)
-
-enriched_metabs <- c("Homarine", "Trigonelline", "Taurine", "DMSP",
-                     "Dimethylsulfonioacetate (DMS-Ac)", "Isethionic Acid")
-culture_metadata <- "../metadata/raw_data/Heal_Metabs_SuppTabs_corrected20230327.xlsx" %>%
-  readxl::read_excel(sheet = "S7_Culturedescriptions_full")
-culture_areas <- "../metadata/raw_data/Heal_Metabs_SuppTabs_corrected20230327.xlsx" %>%
-  readxl::read_excel(sheet = "S5_FullEnviroCulture_PeakAreas") %>%
-  filter(Identification%in%enriched_metabs)
-katherine_comp_mostcycenriched_gp <- culture_areas %>%
-  select(Identification, any_of(culture_metadata$`Sample ID`)) %>%
-  mutate(Identification=ifelse(Identification=="Dimethylsulfonioacetate (DMS-Ac)", 
-                               "DMS-Ac", Identification)) %>%
-  mutate(Identification=factor(Identification, levels=c(
-    "Homarine", "Trigonelline", "Isethionic Acid", "Taurine", "DMSP","DMS-Ac"
-  ))) %>%
-  pivot_longer(-Identification, names_to = "Sample ID") %>%
-  mutate(value=as.numeric(value)) %>%
-  left_join(culture_metadata, by=join_by(`Sample ID`)) %>%
-  arrange(`Broad taxon`) %>%
-  mutate(`Short ID`=fct_inorder(`Short ID`)) %>%
-  mutate(`Species`=fct_inorder(`Species`)) %>%
-  drop_na() %>%
-  ggplot() +
-  geom_boxplot(aes(x=Species, y=value, color=`Broad taxon`)) +
-  facet_wrap(~Identification, ncol = 2, scales="free_y") +
-  scale_y_log10(name="Peak area") +
-  theme_bw() +
-  theme(axis.text.x = element_text(angle=90, hjust=1, vjust=0.5))
-
-ggsave("katherine_comp_mostcycenriched_gp.png", plot = katherine_comp_mostcycenriched_gp, 
-       device = "png", path = "supplement", width = 6.5, height = 5, units = "in", dpi = 300)
-```
-
-![](supplement/katherine_comp_mostcycenriched_gp.png)
-
-*Supplemental figure? Culture data from Katherine's manuscript showing the presence of the DCM cyclone-enriched metabolites mostly present in diatoms and dinoflagellates.*
-
-```{r enrichment plot DCM comparison, eval=FALSE}
-MC_stacked_mostdiff_gp <- pairwise_tests %>%
-  arrange(p.value) %>%
-  group_by(statistic>110) %>%
-  slice(1:5) %>%
-  ungroup() %>%
-  # filter(p.value<0.00005) %>%
-  left_join(untarg_feats %>% select(feature, compound_name), by = join_by(feature)) %>%
-  mutate(compound_name=ifelse(is.na(compound_name), feature, compound_name)) %>%
-  mutate(compound_name=ifelse(compound_name=="Dimethylsulfoniopropionate", "DMSP", compound_name)) %>%
-  arrange(statistic) %>%
-  mutate(compound_name=fct_inorder(compound_name)) %>%
-  select(feature, compound_name) %>%
-  left_join(untarg_peaks %>% select(filename, norm_area, feature), by="feature") %>%
-  left_join(filled_file_metadata %>% distinct(filename, station, abs_depth), by="filename") %>%
-  filter(station%in%c("L1", "L2")) %>%
-  arrange(station) %>%
-  mutate(station=ifelse(station=="L1", "Cyclone", "Anticyclone")) %>%
-  mutate(station=fct_inorder(station)) %>%
-  arrange(abs_depth) %>%
-  mutate(filename=fct_inorder(filename)) %>%
-  group_by(feature) %>%
-  # mutate(norm_area=rank(norm_area)) %>%
-  mutate(norm_area=norm_area/sum(norm_area)) %>%
-  ungroup() %>%
-  ggplot() +
-  geom_col(aes(x=filename, y=norm_area, fill=compound_name, group=filename), 
-           position = "fill", color="grey50", width=1) +
-  geom_rect(aes(xmin=-Inf, xmax=Inf, ymin=1, ymax=1.1), 
-            fill=c("#8277b6", "#b05d56"), color="black",
-            data = data.frame(station=fct_inorder(c("Cyclone", "Anticyclone")))) +
-  geom_text(aes(x=8, y=1.05, label=station), hjust=0.5,
-            data = data.frame(station=fct_inorder(c("Cyclone", "Anticyclone")))) +
-  facet_grid(~station, scales="free_x") +
-  scale_fill_manual(values = splividis(10, begin = 0.1, end = 0.95)) +
-  guides(fill=guide_legend(reverse = TRUE)) +
-  scale_y_discrete(expand = c(0,0)) +
-  theme_bw() +
-  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        strip.background = element_blank(), strip.text = element_blank()) +
-  labs(x=NULL, y="Relative peak area", fill="Mass feature")
-ggsave("MC_stacked_mostdiff_gp.png", plot = MC_stacked_mostdiff_gp, device = "png", 
-       path = "supplement", width = 6.5, height = 3, units = "in", dpi = 300)
-```
-
-![](supplement/MC_stacked_mostdiff_gp.png)
-
-*Supplemental figure?
-Particulate matter composition differences between the cyclone and anticyclone when considering only the top 5 most different metabolites enriched in each eddy separately.
-Each sample corresponds to one multicolored vertical bar while metabolites correspond to the colors provided in the legend and have been scaled to the sum of the total peak areas for that metabolite in the sample set.*
-
-### Falkor data reveals different community and response to SLA
+## Falkor data reveals different community and response to SLA
 
@@
 
-When analyzed separately as distinct depths, the surface samples had higher R$^2$ values (R$^2$=.26) and a lower p-value (0.013), while the DCM samples were no longer likely to be distinct between the two eddies (p-value = 0.171) (Supplemental figure XX).
+When analyzed separately as distinct depths, the surface samples had higher R$^2$ values (R$^2$=.26) and a lower p-value (0.013), while the DCM samples were no longer likely to be distinct between the two eddies (p-value = 0.171) (Supplemental figure 3).
+
+```{r falkor concentrations comparison to MESOSCOPE, eval=FALSE}
+fk_targ <- targ_peaks %>% 
+  left_join(filled_file_metadata) %>%
+  filter(ms_run%in%c("FK", "MT")) %>%
+  filter(samp_type=="Smp") %>%
+  select(compound_name, filename, nM, depth, ms_run)
+fk_targ %>%
+  group_by(depth, compound_name, ms_run) %>%
+  summarise(nM=mean(nM)) %>%
+  arrange(desc(nM)) %>%
+  pivot_wider(names_from = c(ms_run, depth), values_from = nM) %>%
+  print(n=Inf)
+cmpd_comp_gp <- fk_targ %>%
+  filter(depth!="Deep") %>%
+  ggplot() +
+  geom_boxplot(aes(x=ms_run, y=nM)) +
+  facet_grid(compound_name~depth, scales="free_y")
+ggsave("cmpd_comp_gp.pdf", cmpd_comp_gp, device = "pdf", width = 6, height = 50, limitsize = FALSE)
+# Visual assessment of boxplots from the above output:
+# Compounds more abundant in Falkor: citrulline, aspartic acid, ornithine, threonine, tyrosine, malic acid
+# Compounds more abundant in MESO: 4-aminobutyric, adenine, gonyol, 4-hydroxyisoleucine, alanine, leucine, nicotinic acid
+```
+
+Overall, we detected several compounds that were distinctly more or less abundant in the Falkor dataset than the MESOSCOPE one.
+Compounds more abundant in Falkor were the amino acids aspartate, threonine, and tyrosine as well as the components of the urea cycle citrulline and ornithine and the TCA intermediate malic acid.
+Compounds less abundant in Falkor were adenine, gonyol, hydroxyisoleucine, leucine, alanine, and nicotinic acid.
+
+  - Discussion: is there a common thread between the up/down regulated compounds?
+Urea cycle for Falkor cruise w aspartate/citrulline/ornithine/malic acid?
+Connect back to the differing community composition of the two cruises?
+Any chance we have urea numbers for Falkor at all?
+Ask about those transcripts?
+I want to connect this to the diatom urea cycle paper (10.1038/nature10074) and imply that during Falkor there was an ornithine-urea cycle disruption (10.1038/s41598-020-73715-8) but there's very few diatoms in Falkor???
+Does the same thing happen in haptophytes?
+It's not like this is different within the cruise but across cruises still seems interesting.
+
+We were also able to use the Falkor dataset to test whether the compounds detected as significantly different in the MESOSCOPE dataset were also different in this eddy pair.
+We expected to find abundant TMAO and more hydroxyisoleucine at the surface than the DCM in addition to most compounds enriched in the cyclone DCM with biomass, especially the six compounds mentioned above in the results of Figure 5 (trigonelline, homarine, DMSAc, DMSP, taurine, isethionic acid).
+
+```{r abundant TMAO and hydroxyisoleucine depth trends, eval=FALSE}
+fk_targ <- targ_peaks %>% 
+  left_join(filled_file_metadata) %>%
+  filter(ms_run%in%c("FK")) %>%
+  filter(samp_type=="Smp") %>%
+  select(compound_name, filename, nM, depth, ms_run)
+
+fk_targ %>%
+  filter(compound_name=="Hydroxyisoleucine") %>%
+  left_join(filled_file_metadata) %>%
+  aov(formula=nM~depth) %>% summary()
+  # ggplot(aes(x=depth, y=nM, color=station)) +
+  # geom_jitter(width=0.2)
+  # geom_boxplot()
+
+fk_targ %>%
+  filter(compound_name%in%c("Homarine", "Trigonelline", "Dimethylsulfonioacetate", "Dimethylsulfoniopropionate", "Taurine", "Isethionic acid", "Arsenobetaine")) %>%
+  left_join(filled_file_metadata) %>%
+  group_by(filename) %>%
+  # mutate(nM=nM/sum(nM)) %>%
+  ungroup() %>%
+  filter(depth=="DCM") %>%
+  # nest(data=-compound_name) %>%
+  # mutate(aovout=map(data, ~broom::tidy(aov(.x$nM~.x$sla_class)))) %>%
+  # unnest(aovout) %>% filter(term==".x$sla_class") %>%
+  # mutate(p.value=p.adjust(p.value, method="fdr"))
+  ggplot(aes(x=sla_class, y=nM)) +
+  # geom_boxplot() +
+  geom_point() +
+  facet_wrap(~compound_name, scales="free_y")
+
+untarg_feats %>%
+  filter(mzmed%between%pmppm(170.11756, 5) | mzmed%between%pmppm(173.09211, 5)) %>%
+  filter(startsWith(feature, "FK")) %>% 
+  left_join(untarg_peaks) %>%
+  left_join(filled_file_metadata) %>%
+  nest(data=-feature) %>%
+  mutate(aovout=map(data, ~broom::tidy(aov(.x$norm_area~.x$sla_class)))) %>%
+  unnest(aovout) %>% filter(term==".x$sla_class") %>%
+  mutate(p.value=p.adjust(p.value, method="fdr"))
+  select(feature, sla_class, norm_area) %>% 
+  pivot_wider(names_from = sla_class, values_from = norm_area, values_fn = mean)
+  ggplot(aes(x=sla_class, y=norm_area)) +
+  # geom_boxplot() +
+  geom_point() +
+  facet_wrap(~feature, scales="free_y")
+```
+
+As expected, TMAO was again one of the most abundant metabolites detected in the particulate matter with concentrations around 1.2 nM at the surface and 0.4 nM at the DCM, second only to the surprisingly high levels of the amino acid glycine (2 nM surface concentrations, 0.9 nM DCM concentrations), though the differences between the surface and DCM were not significant for either compound.
+We did detect a significant difference between depths for hydroxyisoleucine, once again expected, despite the aforementioned lower concentrations in this sample set (surface mean = 0.32 nM, DCM mean = 0.21 nM, t-test p-value=0.022 with n=12 samples at each depth).
+
+The eddy effects, on the other hand, were much less strong during this cruise.
+Of the six metabolites with the strongest differences across the eddy dipole at the DCM in the MESOSCOPE cruise, only isethionic acid was significantly different in Falkor (isethionic acid t-test p-value$_{FDR}$ = 0.043, other p-values > 0.25).
+Notably, this difference in isethionic acid concentration was only found after normalizing each sample to the sum of all metabolites in the sample to control for biomass.
+Arsenobetaine, found in the MESOSCOPE samples to be strongly enriched in the anticyclone, was again found to have the same pattern here even at the DCM both when normalizing to biomass and when not.
+Surprisingly, the 173.09211 *m/z* mass feature was also slightly enriched but this time in the *cyclone* DCM with peak areas approximately 1.4 times the values in the anticyclone (t-test p-value = 0.067, Mann-Whitney p-value = 0.045).
+The 170.11756 *m/z* mass feature was not detected in the Falkor samples.
+
+  - Discussion: differences between this and MESOSCOPE could be due in large part to the different community composition here - MESOSCOPE looked like a big diatom bloom (according to Benedetto paper) while Sam's results show that this was more dinoflagellate/haptophyte and diatoms had "no trend".
+Maybe also eddy age?
+  - Discussion: Surface more variable than DCM in Falkor (kinda also for MESOSCOPE?)
+  - Discussion: What the hell is with these samples.
+There's an insane amount of variance within triplicates and the SLA signal is highest at the *surface* despite Sam's data and the MESOSCOPE dataset strongly implying the opposite should be true - although if a lot of the metabolism is driven by haptophytes then those are much more abundant at the surface?
+
+# Discussion
+
+# Methods
+# Conclusions
+# References
+
+<div id="refs"></div>
+
+# Supplement
+## Supplemental figure 1
+
+```{r supp figure/table prcomp vs SLA}
+pcdata <- multivar_analysis %>%
+  select(depth, pca_output) %>%
+  mutate(pca_x=map(pca_output, function(pca_output_i){
+    pca_output_i$x[,1:3] %>% as.data.frame()
+  })) %>%
+  select(depth, pca_x) %>%
+  unnest_longer(pca_x, indices_to = "filename") %>%
+  unnest_wider(pca_x) %>%
+  left_join(filled_file_metadata %>% distinct(filename, sla_corr)) %>%
+  pivot_longer(starts_with("PC"))
+pcvars <- multivar_analysis %>%
+  select(depth, pca_output) %>%
+  mutate(perc_expl=map(pca_output, function(pca_output_i){
+    var_expl <- pca_output_i$sdev^2
+    data.frame(perc_expl=var_expl[1:3]/sum(var_expl), name=paste0("PC", 1:3))
+  })) %>%
+  unnest(perc_expl) %>%
+  mutate(perc_label=paste0(round(perc_expl, digits = 3)*100, "%"))
+
+# Correlation between PCs and SLA at each depth
+# pcdata %>%
+#   nest(data=-c(depth, name)) %>%
+#   mutate(cor_output=map_dbl(data, ~abs(cor(.x$value,.x$sla_corr))))
+
+# Linear model significance testing and parameter report
+# pcdata %>%
+#   nest(data=-c(depth, name)) %>%
+#   mutate(lm_output=map(data, ~broom::tidy(lm(.x$value~.x$sla_corr)))) %>%
+#   unnest(lm_output) %>%
+#   filter(term==".x$sla_corr") %>%
+#   arrange(name, depth) %>%
+#   select(depth, name, p.value) %>%
+#   split(.$name)
+
+lm_gp <- ggplot(pcdata, aes(x=sla_corr, y=value)) +
+  geom_smooth(method="lm", formula=y~x, color="black") +
+  geom_point(aes(fill=sla_corr), pch=21, color="black") +
+  geom_label(data = pcvars, aes(label=perc_label),
+             hjust=1, vjust=1, x=Inf, y=Inf, label.r = unit(0, "in")) +
+  facet_grid(depth~name, switch = "y") +
+  scale_y_continuous(position = "right") +
+  scale_fill_gradient2(
+    low = scales::muted("blue"), high = scales::muted("red"),  mid = "grey90",
+    aesthetics = c("color", "fill"), guide = NULL
+  ) +
+  labs(x="Corrected SLA, in cm", y="Principal component value",
+       color="Corrected\nSLA (cm)") +
+  theme_bw()
+deep_pc1_barplot_gp <- pcdata %>%
+  left_join(filled_file_metadata %>% distinct(filename, station, tripl)) %>%
+  mutate(station=factor(station, levels=as.character(4:14),
+                        labels=paste0("St.", 4:14))) %>%
+  arrange(station) %>%
+  filter(name=="PC1" & str_detect(depth, "Deep")) %>%
+  ggplot() +
+  geom_hline(yintercept = 0) +
+  geom_col(aes(x=tripl, y=value)) +
+  facet_wrap(~station, nrow = 1, scales="free_x") +
+  scale_y_continuous(position = "right") +
+  labs(x="Triplicate", y="PC1", title = paste(
+    "    Variance among deep (175 meter) samples in PC1"
+  ))
+pca_sla_cor_gp <- cowplot::plot_grid(lm_gp, deep_pc1_barplot_gp, 
+                                     rel_heights = c(0.6, 0.4), 
+                                     ncol = 1, labels = "AUTO")
+ggsave("pca_sla_cor_gp.png", pca_sla_cor_gp, device = "png", path = "supplement",
+       width = 6.5, height = 6.5, units = "in", dpi = 300)
+
+```
+
+![](supplement/pca_sla_cor_gp.png)
+
+*Supplemental figure 1: Results of a principal component analysis performed separately on each depth of the eddy transect metabolome.
+A) Regressions of SLA against various principal components, broken down by depth.
+% variance explained by each PC is noted in the upper right corner of the plot.
+Colors have been added corresponding to the SLA value for clarity but are redundant with the x-axis values.
+Lines of best fit and the associated standard error have been added behind the points in black and translucent grey, respectively.
+Strong associations exist between PC1 and SLA in the DCM samples and PC2 and SLA in the deep samples.
+B) Principal component 1 values for the 175 meter samples at each station showing large variance between triplicates rather than any external factor.*
+
+## Supplemental figure 2
+
+```{r supp figure fraction of PC in metabolites}
+pc_comp_df <- targ_peaks %>%
+  filter(str_detect(filename, "180821_Smp")) %>%
+  left_join(clean_stans %>% select(compound_name, formula, polarity)) %>%
+  mutate(n_C=as.numeric(str_extract(formula, "(?<=C)\\d+"))) %>%
+  mutate(nM_C=n_C*nM) %>%
+  left_join(filled_file_metadata, by = join_by(filename)) %>%
+  filter(samp_type=="Smp") %>%
+  group_by(filename) %>%
+  summarize(total_nM=sum(nM_C), PC_um=unique(PC_um)) %>%
+  mutate(frac_in_metab=total_nM/1000/PC_um) %>%
+  left_join(filled_file_metadata, by = join_by(filename)) %>%
+  mutate(depth=str_extract(depth, "DCM|Deep|Surface")) %>%
+  mutate(depth=factor(depth, levels=c("Surface", "DCM", "Deep"))) %>%
+  mutate(station=factor(station, levels=c(as.character(4:14))))
+
+pc_frac_metab_gp <- ggplot(pc_comp_df, aes(x=station, y=frac_in_metab, color=sla_corr)) +
+  geom_hline(yintercept = 0) +
+  geom_boxplot(width = 0.9, linewidth=0.8) +
+  geom_point(color="black", size=2) +
+  scale_y_continuous(labels = scales::label_percent()) +
+  facet_grid(~depth, scales="free") +
+  scale_color_gradient2(
+    low = scales::muted("blue"), high = scales::muted("red"),  mid = "grey90",
+    aesthetics = c("color", "fill")
+  ) +
+  guides(color=guide_colorbar(direction = "horizontal", title.position = "top", 
+                              title.hjust = 0.5)) +
+  labs(color="Sea level\nanomaly (cm)", x="Station",
+       y="Fraction of PC represented by known metabolites") +
+  theme_bw() +
+  theme(legend.position = c(0.85, 0.8), legend.background = element_rect(color="grey90"))
+ggsave("pc_frac_metab_gp.png", pc_frac_metab_gp, device = "png", path = "supplement",
+       width = 6.5, height = 3.5, units = "in", dpi = 300)
+```
+
+![](supplement/pc_frac_metab_gp.png)
+
+*Supplemental figure 2: Fraction of particulate carbon represented by known metabolites.
+Boxplots set behind points are colored according to sea level anomaly (SLA) with red corresponding to a positive SLA (anticyclone) and blue corresponding to negative SLA (cyclone).*
+
+## Supplemental figure 3
 
@@
 
-*Supplemental figure(?): Plots of the metabolome in the Falkor data, broken down by depth.
+*Supplemental figure 3: Plots of the metabolome in the Falkor data, broken down by depth.
 The top row of plots are non-metric multidimensional scaling (NMDS) plots, in which points correspond to individual samples and have been colored by their corrected sea level height anomaly.
@@
 
-```{r falkor concentrations comparison to MESOSCOPE, eval=FALSE}
-fk_targ <- targ_peaks %>% 
-  left_join(filled_file_metadata) %>%
-  filter(ms_run%in%c("FK", "MT")) %>%
+## Unnumbered figures, location/inclusion TBD
+
+```{r lm_line_rank_order supplemental figure}
+reg_coefs <- untarg_peaks %>%
+  left_join(filled_file_metadata %>% distinct(filename, ms_run, station, depth, sla_corr)) %>%
+  filter(ms_run=="MT") %>%
+  group_by(feature) %>%
+  filter(sum(norm_area==0)<n()*0.9) %>%
+  mutate(norm_area=rank(norm_area)) %>%
+  ungroup() %>%
+  nest(data=-c(depth, feature)) %>%
+  mutate(test=map(data, ~broom::tidy(lm(.x$norm_area~.x$sla_corr)))) %>%
+  unnest(test)
+reg_coefs %>%
+  filter(str_detect(term, "sla_corr")) %>%
+  mutate(p.value=p.adjust(p.value, method="fdr")) %>%
+  arrange(p.value) %>%
+  left_join(untarg_feats) %>%
+  select(feature, depth, estimate, p.value, mzmed, rtmed, compound_name) %>%
+  filter(p.value<0.05) %>%
+  mutate(depth=factor(depth, levels=c("Surface", "DCM", "Deep"))) %>%
+  mutate(rtmed=round(rtmed, digits = 2)) %>%
+  mutate(mzmed=round(mzmed, digits = 5)) %>%
+  arrange(depth, p.value) %>%
+  group_by(depth) %>%
+  group_modify(~ add_row(.x, .after = nrow(.))) %>%
+  mutate(across(everything(), ~ifelse(is.na(.x), "", .x))) %>%
+  write_csv("supplement/lm_est_rank_order.csv")
+
+clean_slopes <- reg_coefs %>%
+  select(feature, data, depth, term, estimate, p.value) %>%
+  mutate(term=ifelse(term==".x$sla_corr", "slope", "intercept")) %>%
+  pivot_wider(names_from = term, values_from=c(p.value, estimate)) %>%
+  mutate(p.value=p.adjust(p.value_slope, method="fdr")) %>%
+  select(-p.value_intercept, -p.value_slope) %>%
+  filter(p.value<0.05) %>%
+  mutate(pointcol=ifelse(sign(estimate_slope)==1, "Anticyclone enriched", "Cyclone enriched")) %>%
+  mutate(depth=factor(depth, levels=c("Surface", "DCM", "Deep")))
+metadata_slopes <- filled_file_metadata %>%
+  filter(ms_run=="MT") %>%
   filter(samp_type=="Smp") %>%
-  select(compound_name, filename, nM, depth, ms_run)
-fk_targ %>%
-  group_by(depth, compound_name, ms_run) %>%
-  summarise(nM=mean(nM)) %>%
-  arrange(desc(nM)) %>%
-  pivot_wider(names_from = c(ms_run, depth), values_from = nM) %>%
-  print(n=Inf)
-cmpd_comp_gp <- fk_targ %>%
-  filter(depth!="Deep") %>%
+  select(depth, pressure:beam_atten) %>%
+  distinct() %>%
+  pivot_longer(-c(sla_corr, depth)) %>%
+  group_by(name) %>%
+  mutate(value=rank(value)*3) %>%
+  nest(data=-c(name, depth)) %>%
+  mutate(lmout=map(data, ~broom::tidy(lm(.x$value~.x$sla_corr)))) %>%
+  unnest(lmout) %>%
+  filter(name!="dcm_depth") %>%
+  select(depth, name, data, term, estimate, p.value) %>%
+  pivot_wider(names_from = term, values_from = c(estimate, p.value)) %>%
+  mutate(p.value=p.adjust(`p.value_.x$sla_corr`)) %>%
+  filter(p.value<0.05) %>%
+  select(-c(`p.value_(Intercept)`, `p.value_.x$sla_corr`)) %>%
+  mutate(intercept=`estimate_(Intercept)`, slope=`estimate_.x$sla_corr`) %>%
+  mutate(pointcol=ifelse(sign(slope)==1, "Anticyclone enriched", "Cyclone enriched")) %>%
+  mutate(depth=factor(depth, levels=c("Surface", "DCM", "Deep"))) %>%
+  mutate(name=factor(name, levels=metadata_levels, labels=metadata_short_labels))
 ggplot() +
-  geom_boxplot(aes(x=ms_run, y=nM)) +
-  facet_grid(compound_name~depth, scales="free_y")
-ggsave("cmpd_comp_gp.pdf", cmpd_comp_gp, device = "pdf", width = 6, height = 50, limitsize = FALSE)
-# Visual assessment of boxplots from the above output:
-# Compounds more abundant in Falkor: citrulline, aspartic acid, ornithine, threonine, tyrosine, malic acid
-# Compounds more abundant in MESO: 4-aminobutyric, adenine, gonyol, 4-hydroxyisoleucine, alanine, leucine, nicotinic acid
-```
-
-Overall, we detected several compounds that were distinctly more or less abundant in the Falkor dataset than the MESOSCOPE one.
-Compounds more abundant in Falkor were the amino acids aspartate, threonine, and tyrosine as well as the components of the urea cycle citrulline and ornithine and the TCA intermediate malic acid.
-Compounds less abundant in Falkor were adenine, gonyol, hydroxyisoleucine, leucine, alanine, and nicotinic acid.
-
-  - Discussion: is there a common thread between the up/down regulated compounds?
-Urea cycle for Falkor cruise w aspartate/citrulline/ornithine/malic acid?
-Connect back to the differing community composition of the two cruises?
-Any chance we have urea numbers for Falkor at all?
-Ask about those transcripts?
-I want to connect this to the diatom urea cycle paper (10.1038/nature10074) and imply that during Falkor there was an ornithine-urea cycle disruption (10.1038/s41598-020-73715-8) but there's very few diatoms in Falkor???
-Does the same thing happen in haptophytes?
-It's not like this is different within the cruise but across cruises still seems interesting.
-
-We were also able to use the Falkor dataset to test whether the compounds detected as significantly different in the MESOSCOPE dataset were also different in this eddy pair.
-We expected to find abundant TMAO and more hydroxyisoleucine at the surface than the DCM in addition to most compounds enriched in the cyclone DCM with biomass, especially the six compounds mentioned above in the results of Figure 5 (trigonelline, homarine, DMSAc, DMSP, taurine, isethionic acid).
-
-```{r abundant TMAO and hydroxyisoleucine depth trends, eval=FALSE}
-fk_targ <- targ_peaks %>% 
-  left_join(filled_file_metadata) %>%
-  filter(ms_run%in%c("FK")) %>%
-  filter(samp_type=="Smp") %>%
-  select(compound_name, filename, nM, depth, ms_run)
-
-fk_targ %>%
-  filter(compound_name=="Hydroxyisoleucine") %>%
-  left_join(filled_file_metadata) %>%
-  aov(formula=nM~depth) %>% summary()
-  # ggplot(aes(x=depth, y=nM, color=station)) +
-  # geom_jitter(width=0.2)
-  # geom_boxplot()
-
-fk_targ %>%
-  filter(compound_name%in%c("Homarine", "Trigonelline", "Dimethylsulfonioacetate", "Dimethylsulfoniopropionate", "Taurine", "Isethionic acid", "Arsenobetaine")) %>%
-  left_join(filled_file_metadata) %>%
-  group_by(filename) %>%
-  # mutate(nM=nM/sum(nM)) %>%
+  geom_abline(aes(intercept=estimate_intercept, slope=estimate_slope), data = clean_slopes) +
+  geom_point(aes(x=sla_corr, y=norm_area), data=unnest(clean_slopes, data)) +
+  geom_abline(data = metadata_slopes, aes(slope=slope, intercept=intercept, color=name)) +
+  geom_point(aes(x=sla_corr, y=value, color=name), data=unnest(metadata_slopes, data)) +
+  facet_grid(pointcol~depth) +
+  ggtitle("Todo: fix this graph")
+```
+
+*Supplemental figure? Linear regressions of rank-order peak area against SLA showing strong enrichment at the cyclone DCM and the anticyclone 175 meter samples.*
+
+```{r possible figure stacked barplot ranked space}
+top10_stacked_area_station_gp <- map(c("Surface", "DCM", "Deep"), function(depth_i){
+  reg_coefs %>%
+    filter(str_detect(term, "sla_corr")) %>%
+    mutate(p.value=p.adjust(p.value, method="fdr")) %>%
+    filter(depth==depth_i) %>%
+    arrange(p.value) %>%
+    group_by(estimate>0) %>%
+    slice(1:5) %>%
     ungroup() %>%
-  filter(depth=="DCM") %>%
-  # nest(data=-compound_name) %>%
-  # mutate(aovout=map(data, ~broom::tidy(aov(.x$nM~.x$sla_class)))) %>%
-  # unnest(aovout) %>% filter(term==".x$sla_class") %>%
-  # mutate(p.value=p.adjust(p.value, method="fdr"))
-  ggplot(aes(x=sla_class, y=nM)) +
-  # geom_boxplot() +
-  geom_point() +
-  facet_wrap(~compound_name, scales="free_y")
-
-untarg_feats %>%
-  filter(mzmed%between%pmppm(170.11756, 5) | mzmed%between%pmppm(173.09211, 5)) %>%
-  filter(startsWith(feature, "FK")) %>% 
-  left_join(untarg_peaks) %>%
-  left_join(filled_file_metadata) %>%
-  nest(data=-feature) %>%
-  mutate(aovout=map(data, ~broom::tidy(aov(.x$norm_area~.x$sla_class)))) %>%
-  unnest(aovout) %>% filter(term==".x$sla_class") %>%
-  mutate(p.value=p.adjust(p.value, method="fdr"))
-  select(feature, sla_class, norm_area) %>% 
-  pivot_wider(names_from = sla_class, values_from = norm_area, values_fn = mean)
-  ggplot(aes(x=sla_class, y=norm_area)) +
-  # geom_boxplot() +
-  geom_point() +
-  facet_wrap(~feature, scales="free_y")
-```
-
-As expected, TMAO was again one of the most abundant metabolites detected in the particulate matter with concentrations around 1.2 nM at the surface and 0.4 nM at the DCM, second only to the surprisingly high levels of the amino acid glycine (2 nM surface concentrations, 0.9 nM DCM concentrations), though the differences between the surface and DCM were not significant for either compound.
-We did detect a significant difference between depths for hydroxyisoleucine, once again expected, despite the aforementioned lower concentrations in this sample set (surface mean = 0.32 nM, DCM mean = 0.21 nM, t-test p-value=0.022 with n=12 samples at each depth).
-
-The eddy effects, on the other hand, were much less strong during this cruise.
-Of the six metabolites with the strongest differences across the eddy dipole at the DCM in the MESOSCOPE cruise, only isethionic acid was significantly different in Falkor (isethionic acid t-test p-value$_{FDR}$ = 0.043, other p-values > 0.25).
-Notably, this difference in isethionic acid concentration was only found after normalizing each sample to the sum of all metabolites in the sample to control for biomass.
-Arsenobetaine, found in the MESOSCOPE samples to be strongly enriched in the anticyclone, was again found to have the same pattern here even at the DCM both when normalizing to biomass and when not.
-Surprisingly, the 173.09211 *m/z* mass feature was also slightly enriched but this time in the *cyclone* DCM with peak areas approximately 1.4 times the values in the anticyclone (t-test p-value = 0.067, Mann-Whitney p-value = 0.045).
-The 170.11756 *m/z* mass feature was not detected in the Falkor samples.
-
-  - Discussion: differences between this and MESOSCOPE could be due in large part to the different community composition here - MESOSCOPE looked like a big diatom bloom (according to Benedetto paper) while Sam's results show that this was more dinoflagellate/haptophyte and diatoms had "no trend".
-Maybe also eddy age?
-  - Discussion: Surface more variable than DCM in Falkor (kinda also for MESOSCOPE?)
-  - Discussion: What the hell is with these samples.
-There's an insane amount of variance within triplicates and the SLA signal is highest at the *surface* despite Sam's data and the MESOSCOPE dataset strongly implying the opposite should be true - although if a lot of the metabolism is driven by haptophytes then those are much more abundant at the surface?
+    left_join(untarg_feats %>% select(feature, compound_name, mzmed, rtmed)) %>%
+    mutate(feature=ifelse(!is.na(compound_name), compound_name, 
+                          paste0(round(mzmed, 3), "_", round(rtmed, 1)))) %>%
+    mutate(feature=fct_inorder(feature)) %>%
+    unnest(data) %>%
+    select(feature, depth, norm_area, filename, station) %>%
+    group_by(feature) %>%
+    # mutate(norm_area=rank(norm_area)) %>%
+    mutate(norm_area=norm_area/sum(norm_area)) %>%
+    mutate(station=factor(station, levels=as.character(4:14))) %>%
+    ggplot() +
+    geom_col(aes(x=station, y=norm_area, fill=feature), position = "fill") +
+    scale_y_continuous(expand = c(0,0)) +
+    labs(x="Station", y="Relative peak area", fill="Mass feature") +
+    theme_bw()
+}) %>% gridExtra::arrangeGrob(grobs = ., ncol = 1)
+ggsave("top10_stacked_area_station_gp.png", plot = top10_stacked_area_station_gp, 
+       device = "png", path = "supplement", width = 6.5, height = 9, 
+       units = "in", dpi = 300)
+```
+
+![](supplement/top10_stacked_area_station_gp.png)
+
+*Supplemental figure?
+Plots of the 5 most significantly enriched metabolites in both the cyclone (top 5) and anticyclone (bottom 5) after normalizing to the total peak area in each mass feature.
+Nicely shows a stronger trend with increasing depth but feels a little bit like lying because there's 100 other compounds that don't change in between these extremes.
+Tyrosine is also misannotated here.*
+
+```{r comparison to Katherines culture data}
+cyc_enriched <- pairwise_tests %>%
+  filter(p.value<0.05) %>%
+  left_join(untarg_feats) %>%
+  filter(!is.na(compound_name)) %>%
+  filter(statistic>110)
+
+enriched_metabs <- c("Homarine", "Trigonelline", "Taurine", "DMSP",
+                     "Dimethylsulfonioacetate (DMS-Ac)", "Isethionic Acid")
+culture_metadata <- "../metadata/raw_data/Heal_Metabs_SuppTabs_corrected20230327.xlsx" %>%
+  readxl::read_excel(sheet = "S7_Culturedescriptions_full")
+culture_areas <- "../metadata/raw_data/Heal_Metabs_SuppTabs_corrected20230327.xlsx" %>%
+  readxl::read_excel(sheet = "S5_FullEnviroCulture_PeakAreas") %>%
+  filter(Identification%in%enriched_metabs)
+katherine_comp_mostcycenriched_gp <- culture_areas %>%
+  select(Identification, any_of(culture_metadata$`Sample ID`)) %>%
+  mutate(Identification=ifelse(Identification=="Dimethylsulfonioacetate (DMS-Ac)", 
+                               "DMS-Ac", Identification)) %>%
+  mutate(Identification=factor(Identification, levels=c(
+    "Homarine", "Trigonelline", "Isethionic Acid", "Taurine", "DMSP","DMS-Ac"
+  ))) %>%
+  pivot_longer(-Identification, names_to = "Sample ID") %>%
+  mutate(value=as.numeric(value)) %>%
+  left_join(culture_metadata, by=join_by(`Sample ID`)) %>%
+  arrange(`Broad taxon`) %>%
+  mutate(`Short ID`=fct_inorder(`Short ID`)) %>%
+  mutate(`Species`=fct_inorder(`Species`)) %>%
+  drop_na() %>%
+  ggplot() +
+  geom_boxplot(aes(x=Species, y=value, color=`Broad taxon`)) +
+  facet_wrap(~Identification, ncol = 2, scales="free_y") +
+  scale_y_log10(name="Peak area") +
+  theme_bw() +
+  theme(axis.text.x = element_text(angle=90, hjust=1, vjust=0.5))
+
+ggsave("katherine_comp_mostcycenriched_gp.png", plot = katherine_comp_mostcycenriched_gp, 
+       device = "png", path = "supplement", width = 6.5, height = 5, units = "in", dpi = 300)
+```
+
+![](supplement/katherine_comp_mostcycenriched_gp.png)
+
+*Supplemental figure? Culture data from Katherine's manuscript showing the presence of the DCM cyclone-enriched metabolites mostly present in diatoms and dinoflagellates.*
+
+```{r enrichment plot DCM comparison, eval=FALSE}
+MC_stacked_mostdiff_gp <- pairwise_tests %>%
+  arrange(p.value) %>%
+  group_by(statistic>110) %>%
+  slice(1:5) %>%
+  ungroup() %>%
+  # filter(p.value<0.00005) %>%
+  left_join(untarg_feats %>% select(feature, compound_name), by = join_by(feature)) %>%
+  mutate(compound_name=ifelse(is.na(compound_name), feature, compound_name)) %>%
+  mutate(compound_name=ifelse(compound_name=="Dimethylsulfoniopropionate", "DMSP", compound_name)) %>%
+  arrange(statistic) %>%
+  mutate(compound_name=fct_inorder(compound_name)) %>%
+  select(feature, compound_name) %>%
+  left_join(untarg_peaks %>% select(filename, norm_area, feature), by="feature") %>%
+  left_join(filled_file_metadata %>% distinct(filename, station, abs_depth), by="filename") %>%
+  filter(station%in%c("L1", "L2")) %>%
+  arrange(station) %>%
+  mutate(station=ifelse(station=="L1", "Cyclone", "Anticyclone")) %>%
+  mutate(station=fct_inorder(station)) %>%
+  arrange(abs_depth) %>%
+  mutate(filename=fct_inorder(filename)) %>%
+  group_by(feature) %>%
+  # mutate(norm_area=rank(norm_area)) %>%
+  mutate(norm_area=norm_area/sum(norm_area)) %>%
+  ungroup() %>%
+  ggplot() +
+  geom_col(aes(x=filename, y=norm_area, fill=compound_name, group=filename), 
+           position = "fill", color="grey50", width=1) +
+  geom_rect(aes(xmin=-Inf, xmax=Inf, ymin=1, ymax=1.1), 
+            fill=c("#8277b6", "#b05d56"), color="black",
+            data = data.frame(station=fct_inorder(c("Cyclone", "Anticyclone")))) +
+  geom_text(aes(x=8, y=1.05, label=station), hjust=0.5,
+            data = data.frame(station=fct_inorder(c("Cyclone", "Anticyclone")))) +
+  facet_grid(~station, scales="free_x") +
+  scale_fill_manual(values = splividis(10, begin = 0.1, end = 0.95)) +
+  guides(fill=guide_legend(reverse = TRUE)) +
+  scale_y_discrete(expand = c(0,0)) +
+  theme_bw() +
+  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        strip.background = element_blank(), strip.text = element_blank()) +
+  labs(x=NULL, y="Relative peak area", fill="Mass feature")
+ggsave("MC_stacked_mostdiff_gp.png", plot = MC_stacked_mostdiff_gp, device = "png", 
+       path = "supplement", width = 6.5, height = 3, units = "in", dpi = 300)
+```
+
+![](supplement/MC_stacked_mostdiff_gp.png)
+
+*Supplemental figure?
+Particulate matter composition differences between the cyclone and anticyclone when considering only the top 5 most different metabolites enriched in each eddy separately.
+Each sample corresponds to one multicolored vertical bar while metabolites correspond to the colors provided in the legend and have been scaled to the sum of the total peak areas for that metabolite in the sample set.*

### 2023-09-06 manuscript/manuscript.Rmd
Adding Methods section from masters manuscript

@@
 
+Despite its historical reputation as a static and unchanging environment, the North Pacific Subtropical Gyre shows significant variation in its particulate composition due to mesoscale eddies.
+
 # Methods
+
+## Cruise information
+Samples were collected from two cruises in the North Pacific Subtropical Gyre near Station ALOHA that targeted strong mesoscale eddy features as described in @Dugenne2023 and @Gleich2023.
+Briefly, the 2017 MESOSCOPE cruise (KM1709 on the *R/V Kilo Moana*) consisted of a transect across the cyclonic and anticyclonic poles of the eddy dipole as well as two long-term Lagrangian stations at the center of each eddy.
+The cyclonic pole of the eddy had a maximum negative SLA anomaly of -20 centimeters and the anticyclonic center reached +24 centimeters.
+The transect samples were taken at various times of day as the ship transected the dipole, and the eddy center samples were all collected between 5 and 8 pm.
+In 2018, the Falkor cruise (FK180310 on the *R/V Falkor*) targeted a new eddy dipole in approximately the same location with samples taken at the center of each eddy (maximum negative anomaly in the cyclone = -15 cm, maximum positive anomaly in the anticyclone = 26 cm).
+
+## Biogeochemical data
+
+Biogeochemical measurements were collected as described in @Barone2022 and @Dugenne2023 mostly following protocols used by the HOT program (http://hahana.soest.hawaii.edu/hot/methods/results.html).
+Briefly, all environmental data was collected via CTD Niskin bottles except beam attenuation which was measured via Hyperpro and sea level anomaly (SLA) which was measured via satellite.
+Chlorophyll and pheophytin were measured fluorometrically, particulate carbon (PC) and particulate nitrogen (PN) were measured using an elemental analyzer, and nitrate + nitrite (N+N) and soluble reactive phosphorus (phosphate, PO$_4^{3-}$) were measured on an autoanalyzer except where concentrations were below 100nM in which case they were measured using chemiluminescence.
+Values were interpolated linearly to the locations at which metabolite samples were taken using all data collected at the given station.
+
+## Metabolite sample collection
+Samples were obtained using the onboard CTD rosette to collect water from 15 meters, the deep chlorophyll maximum (DCM), and 175 meters during the eddy transect and from the DCM ±10 and ±20 meters during the eddy center sampling in the MESOSCOPE cruise and from 25 meters and the DCM during the Falkor cruise.
+The DCM was determined visually from fluorometer data during the CTD downcast and Niskin bottles were tripped during the return trip to the surface.
+Seawater from each depth was sampled in triplicate by firing one Niskin bottle for each sample.
+Samples were brought to the surface and decanted into prewashed (3x with DI, 3x with sampled seawater) polycarbonate bottles for filtration.
+Samples were filtered by peristaltic pump onto 142mm 0.2 µm Durapore filters held by polycarbonate filter holders on a Masterflex tubing line.
+Pressures were kept as low as possible while still producing a reasonable rate of flow through the filter, approximately 250-500 mL per minute.
+Samples were then removed from the filter holder using solvent-washed tweezers and placed into pre-combusted aluminum foil packets that were then flash-frozen in liquid nitrogen before being stored at -80 °C until extraction.
+A methodological blank was also collected by running filtrate through a new filter and then treated identically to the samples.
+
+## Metabolite sample extraction
+Extraction followed a modified Bligh & Dyer approach as detailed in @Boysen2018.
+Briefly, filters were added to PTFE centrifuge tubes with a 1:1 mix of 100 µm and 400 µm silica beads, approximately 2mL -20 °C Optima-grade DCM, and approximately 3mL -20 °C 1:1 methanol/water solution (both also Optima-grade).
+Extraction standards were added during this step.
+The samples were then bead-beaten three times, followed by triplicate washes with fresh methanol/water mixture.
+Samples were then dried down under clean nitrogen gas and warmed using a Fisher-Scientific Reacti-Therm module.
+Dried aqueous fractions were re-dissolved in 380 µL of Optima-grade water and amended with 20 µL isotope-labeled injection standards.
+Additional internal standards were added at this point to measure the variability introduced by chromatography and ionization, and the reconstituted fraction was syringe-filtered to remove any potential clogging material.
+This aqueous fraction was then aliquoted into an HPLC vial for injection on the HILIC column and diluted 1:1 with Optima-grade water.
+A pooled sample was created by combining 20 µL of each sample into the same HPLC vial, and a 1:1 dilution with water half-strength sample was aliquot from that to assess matrix effects and obscuring variation [@Boysen2018].
+Also run alongside the environmental samples were two mixes of authentic standards in water and in an aliquot of the pooled sample at a variety of concentrations for quality control, annotation, and absolute concentration calculations.
+HPLC vials containing the samples were frozen at -80 °C until thawing shortly before injection.
+
+## HPLC-MS methods
+Separate HPLC-MS runs were used for the MESOSCOPE eddy transect, the MESOSCOPE eddy center, and the Falkor cruise data.
+For each, a SeQuant ZIC-pHILIC column (5 um particle size, 2.1 mm x 150 mm, from Millipore) was used with 10 mM ammonium carbonate in 85:15 acetonitrile to water (Solvent A) and 10 mM ammonium carbonate in 85:15 water to acetonitrile (Solvent B) at a flow rate of 0.15 mL/min.
+The column was held at 100% A for 2 minutes, ramped to 64% B over 18 minutes, ramped to 100% B over 1 minute, held at 100% B for 5 minutes, and equilibrated at 100% A for 25 minutes (50 minutes total).
+The column was maintained at 30 °C.
+The injection volume was 2 µL for samples and standard mixes.
+When starting a batch, the column was equilibrated at the starting conditions for at least 30 minutes.
+To improve the performance of the HILIC column, we maintained the same injection volume, kept the instrument running water blanks between samples as necessary, and injected standards in a representative matrix (the pooled sample) in addition to standards in water.
+After each batch, the column was flushed with 10 mM ammonium carbonate in 85:15 water to acetonitrile for 20 to 30 minutes.
+
+Metabolomic data was collected on a Thermo Q Exactive HF hybrid Orbitrap (QE) mass spectrometer.
+The capillary and auxiliary gas heater temperatures were maintained at 320°C and 100°C, respectively.
+The S-lens RF level was kept at 65, the H-ESI voltage was set to 3.3 kV and sheath gas, auxiliary gas, and sweep gas flow rates were set at 16, 3, and 1, respectively.
+Polarity switching was used with a scan range of 60 to 900 *m/z* and a resolution of 60,000.
+Calibration was performed every 3-4 days at a target mass of 200 *m/z*.
+DDA data was collected from the pooled samples for high-confidence annotation of knowns and unknowns.
+All files were then converted to an open-source mzML format and centroided via Proteowizard’s msConvert tool.
+
+## Targeted metabolomic methods
+
+Targeted compounds were manually integrated using Skyline and MSDIAL.
+Skyline was used for both the eddy center samples and the Falkor cruise data while MSDIAL was used for the eddy transect data due to the larger number of files in that dataset.
+Raw peak areas were then normalized to their best-matched internal standard per @Boysen2018 except that all compounds were normalized to their best match no matter how much or how little improvement was detected.
+Normalized peaks were then calibrated via a single-point standard curve as determined uniquely for each compound from the authentic standards and used to convert normalized peak area into moles per µL injected and scaled to the amount of seawater filtered to provide a final estimate of environmental concentration for each compound in moles per liter of seawater filtered.
+
+## Untargeted metabolomic methods
+
+Samples were also processed via the untargeted workflow detailed in @Kumler2023.
+In short, XCMS was used for peak detection and followed by manual inspection of each mass feature for quality with only those peaks classified as "Good" used in the analysis here.
+Features were then matched with an internal standard as described above.
+
+## Statistics
+
+All statistics were run in R version `r paste(version$major, version$minor, sep=".")`.
+We used multivariate statistics provided by the vegan package [@R-vegan] to assess the overall impact of various metrics across the metabolome followed by univariate statistics to investigate individually significant responses.
+Non-parametric tests were used when data were unlikely to obey parametric assumptions and permutational statistics were preferred wherever possible.
+Multiple comparison problems were controlled where necessary using a false-discovery rate correction.
+
 # Conclusions

### 2023-09-20 manuscript/manuscript.Rmd
Finished discussion and polished up a few todo's
Ready to send to collaborators imo

@@
 ---
-title: "MESOSCOPE manuscript"
+title: "Metabolites reflect the mesoscale variability in biogeochemistry introduced by sea-surface eddies in the North Pacific Subtropical Gyre"
 author: "William Kumler"
@@
     reference_docx: template.docx
+bibliography:
+- Exported Items.bib
+- r-packages.bib
 ---
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, eval = FALSE)
+knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, eval = TRUE)
 library(tidyverse)
+library(RaMS)
 
@@
 
+TBD
+
 # Introduction
@@
 Of particular interest within the metabolome are those small, polar molecules that represent the fundamental unit of cell biology and function as building blocks for macromolecules [@Moran2022].
-These small molecules also have roles of their own within the cell, acting as osmolytes, antioxidants, nutrient storage compounds, and more [@Geraldes2021; @Mojzes2020; @Kharbush2020].
+These small molecules also have roles of their own within the cell, acting as osmolytes, antioxidants, nutrient storage compounds, and more [@Mojzes2020; @Kharbush2020].
 Despite these critical functions, they are difficult to study in the marine environment due to their low concentrations and similarity to the abundant salt ions, resulting in challenges in detection and quantification [@Boysen2018; @Moran2022; @Sacks2022].
@@
 This response correlated negatively with SLA at the DCM (lower SLA = higher average values, $r_{DCM}= -.716$) but positively with SLA at the surface and at depth ($r_{surface} = .211$, $r_{deep} = .830$).
-
-  - Discussion: note that the "neither" samples all group together, implying eddy effect rather than location ("outside" the eddy is more similar to each other than to the insides of either one)
-  - Discussion: it's kinda crazy that the exact center of the eddy is so wildly different at the DCM from stations that are nearby but not
 
@@
   }) %>% mean()
-})
+}, .progress = TRUE)
 ```
@@
 Known metabolites contributed between approximately 1.5% and 5% of the carbon biomass to the total PC load, but this variance was not clearly due to either depth or SLA (Supplemental figure 2).
-The only difference we observed was a slightly reduced metabolite contribution in the 175 meter samples (0.3-0.4 percentage points lower at depth, p-value=0.018) while no difference was detected between the surface and DCM samples or between the cyclone, anticyclone, and outside the eddy samples (all p-values > 0.1).
+The only difference we observed was a slightly reduced metabolite contribution in the 175 meter samples (0.6 percentage points lower at 175 meters, p-value=0.002) while no difference was detected between the surface and DCM samples or between the cyclone, anticyclone, and outside the eddy samples (all p-values > 0.1).
 However, this lack of difference observed may have been due to the small sample sizes (18 anticyclonic samples, 9 cyclonic, and 6 neither).
-A followup permutational sensitivity analysis indicated that a difference in means of 0.4 percentage points would have been obtained at the α < 0.05 level for the 175 meter samples if 1.5x as many samples had been collected with mean and variances remaining similar.
-
-  - Discussion: why are my % PC in metabs lower than Angie's? Because her PC values were way higher but metab concentrations were about the same. Wait, are they even?
+A followup permutational sensitivity analysis indicated that a difference in means of 0.4 percentage points would have been obtained at the α < 0.05 level for SLA in the 175 meter samples if ~1.5x as many samples had been collected with mean and variances remaining similar.
+
+  - Discussion: My % PC in metabs lower than Angie's? Because her PC values were way higher but metab concentrations were about the same?
 
@@
   scale_y_continuous(limits = c(0, NA))
+```
+
+### A single sample highlights a network of related metabolites
+
+One sample, taken from the surface at Station 11 (MS11C215m_A), was highly distinct from all others (including its paired triplicates) with several metabolites measuring two or more orders of magnitude larger in both peak area and concentration.
+Heuristic exploration of nearby masses identified a total of 14 compounds highly abundant in this one sample alone.
+The only identified compounds with this trend were taurine and acetyltaurine, though three others with *m/z* 146.1176, 148.0968, and 206.1388 were discovered to be likely isomers of known compounds (acetylcholine/(3-carboxypropyl)trimethylammonium, hydroxyisoleucine, dexpanthenol, respectively).
+One feature (*m/z* = 184.0637, RT = 7.0 minutes) was identified as choline sulfate using a second run of standards, while an additional mass feature at 168.0689 and RT = 8.5 was putatively flagged as taurine betaine (same mass as hydroxyisoleucine).
+This mass was also one of the most highly enriched in the MS11C215m_A sample after taurine itself, with a peak area nearly 30x that of the next largest sample.
+No clear explanation emerged explaining which compounds were hyperabundant in this surface sample, though one possible explanation was the *Trichodesmium* bloom occurring at the same time.
+
+```{r MS11A offenders table}
+ms11_offenders <- untarg_peaks %>%
+  filter(str_detect(filename, "180821")) %>%
+  group_by(feature) %>%
+  filter(sum(norm_area==0)<3) %>%
+  mutate(cursed_area=norm_area[filename=="180821_Smp_MS11C215m_A.mzML"]) %>%
+  filter(cursed_area>0) %>%
+  mutate(rel_area=norm_area/cursed_area) %>%
+  filter(filename!="180821_Smp_MS11C215m_A.mzML") %>%
+  group_by(feature) %>%
+  summarise(mean_rel=mean(rel_area), max_rel=max(rel_area)) %>%
+  # ggplot(aes(x=mean_rel, y=max_rel, label=feature)) + geom_point() + 
+  # scale_y_log10() + scale_x_log10() +
+  # geom_vline(xintercept = 0.25) + geom_hline(yintercept = 0.5)
+  filter(mean_rel<0.25 & max_rel < 0.5) %>%
+  arrange(mean_rel) %>% 
+  add_row(feature="MS_NEG_1123") %>%
+  left_join(untarg_feats %>% select(feature:compound_name)) %>%
+  mutate(mzmed=ifelse(str_detect(feature, "NEG"), mzmed+1.007276*2, mzmed)) %>%
+  filter(feature!="MS_NEG_0768") %>% # Remove taurine duplicate
+  add_row(mzmed=c(198.0794, 184.0637), rtmed=c(5, 7))
+
+# ms11_offenders$mzmed %>% sort() %>% `names<-`(., round(., digits=3)) %>% dist(diag = TRUE) %>% round(digits=4)
+
+stans <- read_csv("https://raw.githubusercontent.com/IngallsLabUW/Ingalls_Standards/master/Ingalls_Lab_Standards.csv") %>%
+  filter(Column=="HILIC" & z>0) %>%
+  select(cmpd_name=Compound_Name, formula=Empirical_Formula) %>%
+  group_by(formula) %>%
+  summarise(cmpd_name=paste(cmpd_name, collapse = ", "))
+base_mz <- 85.089149355+1.007276
+iter_mzs <- expand_grid(CH2=0:6, O=2:6) %>%
+  mutate(mz=base_mz+CH2*14.01565+O*15.994915) %>%
+  mutate(mz=round(mz, digits=5)) %>%
+  mutate(cmd=paste0("mz%between%pmppm(", mz, ", 10)")) %>%
+  mutate(formula=paste0("C", 5+CH2, "H", 11+CH2*2, "NO", O)) %>%
+  left_join(stans) %>%
+  mutate(label=ifelse(is.na(cmpd_name), formula, paste0(formula, " (", cmpd_name, ")"))) %>%
+  mutate(casewhen_arg=paste0(cmd, "~'", label, "'"))
+ms11_df <- c("C11H23NO6", "C11H23NO6", "C9H19NO5", "C8H17NO5", "C6H13NO6", "C8H17NO4") %>%
+  data.frame(formula=., rtmed=c(8.5, 10, 8.5, 10.2, 5.2, 9.2)) %>%
+  left_join(iter_mzs) %>%
+  select(mzmed=mz, rtmed, formula, cmpd_name) %>%
+  bind_rows(ms11_offenders) %>%
+  arrange(mzmed) %>%
+  select(mzmed, rtmed, formula, compound_name, feature) %>%
+  mutate(formula=case_when(
+    compound_name=="Taurine"~"C2H8NO3S",
+    feature=="MS_POS_1050"~"C7H16NO2",
+    compound_name=="Hydroxyisoleucine"~"C6H14NO3",
+    compound_name=="N-Acetyltaurine"~"C4H10NO4S",
+    feature=="MS_POS_1340"~"C5H14NO3S",
+    feature=="MS_POS_1419"~"C8H18NO3",
+    mzmed%between%pmppm(184.0637, 10)~"C5H14NO4S",
+    mzmed%between%pmppm(198.0794, 10)~"C6H16NO4S",
+    compound_name=="Dexpanthenol"~"C9H20NO4",
+    TRUE~formula
+  )) %>%
+  mutate(compound_name=case_when(
+    compound_name=="(3-Carboxypropyl)trimethylammonium?"~NA,
+    compound_name=="Hydroxyisoleucine"~NA,
+    compound_name=="Dexpanthenol"~NA,
+    TRUE~compound_name
+  ))
+knitr::kable(ms11_df)
+```
+
+We explored this *Trichodesmium* hypothesis further by analyzing additional samples taken from the bloom to see whether the compounds abundant in *Trichodesmium* were also abundant in samples taken directly from the bloom itself.
+Of the 14 masses identified in the environmental sample, 7 had good-quality peaks when viewed in the bloom samples as extracted ion chromatograms, though only 4 of these matched closely in retention time.
+
+```{r trichobloomsamples comparison, eval=FALSE}
+library(RaMS)
+trichodata <- list.files("../mzMLs/tricho_pos", pattern=".mzML", full.names = TRUE) %>%
+  grabMSdata()
+tricho_eics <- ms11_df %>%
+  pmap(function(...){
+  row_data <- list(...)
+  peak_data <- trichodata$MS1[mz%between%pmppm(row_data$mzmed, 10)]
+  cbind(peak_data, formula=row_data$formula)
+}) %>%
+  bind_rows()
+
+tricho_eics %>%
+  filter(rt%between%c(2, 15)) %>%
+  arrange(mz) %>%
+  mutate(formula=fct_inorder(formula)) %>%
+  ggplot() +
+  geom_line(aes(x=rt, y=int, group=filename)) +
+  facet_wrap(~formula, scales="free", ncol=2) +
+  geom_vline(aes(xintercept = rtmed), color="red", data=ms11_df)
+
+trichodata$MS1[mz%between%pmppm(162.1125, 10)] %>%
+  mutate(samp_type=str_extract(filename, "Smp|Poo|Mix1|Mix2|Mat")) %>%
+  filter(samp_type!="Mat") %>%
+  ggplot() +
+  geom_line(aes(x=rt, y=int, group=filename, color=samp_type)) +
+  ggtitle("Homoserine betaine in Tricho bloom samples")
+```
+
+Finally, we performed a search of the untargeted data for homoserine betaine, a known *Trichodesmium* osmolyte [@Pade2016].
+Although this compound was not in our standard library due to a lack of purchasable authentic standards, we identified three mass features in the 162.1125 *m/z* window with retention times of 7.1, 8.4, and 9.9 minutes.
+The feature at RT 9.9 matched our carnitine standard's retention time, leaving the two earlier peaks with putative identifications of threonine betaine at 7.1 and homoserine betaine at 8.4 given the elution order of threonine before homoserine in our authentic standards.
+This order is the reverse of that in @Pade2016, which showed homoserine betaine eluting prior to carnitine as expected given our use of a polar HILIC column and their use of a nonpolar Hypercarb column.
+They did not report an additional peak detected at this mass, leaving our putative identification of the threonine betaine as low-confidence.
+None of the three 162.1125 features showed the same degree of enrichment in MS11C215m_A as taurine or the other 13 compounds detected with that pattern.
+Although MS11C215m_A was the largest, second-largest, and third-largest peak for carnitine, threonine betaine, and homoserine betaine, respectively, it was never more than triple the average surface value or double the next-largest value.
+
+```{r homoserinebetaine search, eval=FALSE}
+library(arrow)
+dset <- open_dataset("../tmzMLs/MT_all_file_arr_pqds/")
+hmb_eic <- dset %>%
+  filter(mz%between%pmppm(162.1125, 10)) %>%
+  # filter(mz%between%pmppm(120.0655, 10)) %>%
+  collect()
+hmb_eic %>%
+  filter(rt%between%c(6, 11)) %>%
+  # filter(rt%between%c(9, 13)) %>%
+  mutate(samp_type=str_extract(filename, "Poo|Smp|InH2O|Mat")) %>%
+  # distinct(filename, samp_type) %>%
+  # filter(samp_type!="Smp")
+  mutate(plotcol=ifelse(str_detect(filename, "180821_Smp_MS11C215m_A.mzML"), 
+                          "MS11C215m_A", samp_type)) %>%
+  ggplot() +
+  geom_line(aes(x=rt, y=int, group=filename, color=plotcol)) +
+  facet_wrap(~samp_type, ncol=1, scales="free_y")
+
+untarg_feats %>% 
+  filter(cruise=="MS") %>%
+  filter(mzmed%between%pmppm(162.1125))
+untarg_peaks %>%
+  filter(feature%in%c("MS_POS_1246", "MS_POS_1247", "MS_POS_1248")) %>%
+  mutate(feature=factor(feature, levels=c("MS_POS_1248", "MS_POS_1246", "MS_POS_1247"),
+                        labels=c("Threonine betaine (RT=7.1)", 
+                                 "Homoserine betaine (RT=8.4)",
+                                 "Carnitine (RT=9.9)"))) %>%
+  left_join(filled_file_metadata) %>%
+  filter(ms_run=="MT") %>%
+  mutate(station=factor(station, levels=as.character(4:14))) %>%
+  mutate(depth=factor(depth, levels=c("Surface", "DCM", "Deep"))) %>%
+  mutate(plotcol=ifelse(filename=="180821_Smp_MS11C215m_A.mzML", "red", "black")) %>%
+  ggplot(aes(x=station, y=norm_area, label=filename, color=plotcol)) +
+  geom_point() +
+  facet_grid(feature~depth, scales="free_y") +
+  geom_hline(yintercept = 0) +
+  scale_color_identity()
+
+untarg_peaks %>%
+  filter(feature%in%c("MS_POS_1246", "MS_POS_1247", "MS_POS_1248")) %>%
+  left_join(filled_file_metadata) %>%
+  filter(ms_run=="MT") %>%
+  group_by(feature) %>%
+  mutate(norm_area=norm_area/norm_area[filename=="180821_Smp_MS11C215m_A.mzML"]) %>%
+  group_by(feature, depth) %>%
+  summarise(mean_area=max(norm_area)) %>%
+  pivot_wider(names_from = "depth", values_from = mean_area)
 ```
@@
   left_join(untarg_feats) %>%
-  filter(compound_name=="Acetylcholine?") %>%
+  filter(compound_name=="Homarine") %>%
   # filter(feature=="MS_POS_1395") %>%
-  filter(depth=="Deep") %>%
+  filter(depth=="DCM") %>%
   select(compound_name, depth, data, estimate) %>%
@@
 
-Finally, in the 175 meter samples we detected 32 mass features with a significant association with SLA.
+Finally, in the 175 meter samples we detected 43 mass features with a significant association with SLA.
 Notably, all detected nucleobases (guanine, adenine, and cytosine) were positively associated with SLA (higher values in the anticyclone than the cyclone), though the strongest associations were found in O-acetylcarnitine, betonicine, and tyrosine.
@@
 The unknown at *m/z* 173.09211 with the strongest trend at the DCM again had the strongest trend in the 175 meter samples, here with an FDR-corrected p-value of $2.0*10^{-7}$.
-
-  - Discussion: surprised that so many compounds are still different after normalizing within each file - expected that to remove biomass signal
-- Discussion: why are homarine and trigonelline so correlated?
-- Discussion: homarine/trigonelline has a larger effect size but lower p-value because of nonlinear fit (exponential rise into eddy center rather than linear)
-- Discussion: how can most compounds be relatively enriched in the anticyclone at 175 meters? Where is the remainder of the effect size going?
-- Discussion: all the nucleobases upregulated in the 175m anticyclone, why? Bacteria?
 
@@
 To provide context for the sea level anomaly responses detected above, we additionally performed high-resolution depth sampling around the DCM at the center of both eddy poles.
-Here, we found that the ~50 meter difference in depth and the ~40 centimeter difference in SLA explained similar amounts of variation in the data (R$^2_{depth}=0.253$, R$^2_{SLA}=0.238$) and were both highly significant factors (permutational p-values << 0.001).
+Here, we found that the ~40 meter difference in depth and the ~40 centimeter difference in SLA explained similar amounts of variation in the data (R$^2_{depth}=0.253$, R$^2_{SLA}=0.238$) and were both highly significant factors (permutational p-values << 0.001).
 Additionally, both sample depth and SLA were well-correlated with the first principal component of the metabolite matrix (r$_{depth}=.766$, r$_{SLA}=.684$, fraction of variance explained by PC$_1$ = 30.8%) (Figure 3).
+A 2D NMDS plot of the samples showed good separation between the eddies, with cyclonic samples largely on the left side of Figure 3 while anticyclonic samples fell almost entirely to the right.
+Notably, the cyclonic samples show a much larger spread than the anticyclonic ones, indicating larger sample variability at the surface than the DCM.
+A depth gradient is visible along MDS 2, with the deepest samples generally at the top right of the plot and the shallowest ones closer to the bottom and left.
 
@@
   rownames_to_column("filename") %>%
-  left_join(filled_file_metadata %>% distinct(filename, sla_class, depth)) %>%
+  left_join(filled_file_metadata %>% distinct(filename, sla_class, depth, tripl)) %>%
   mutate(depth=factor(depth, levels=c("DCMless20m", "DCMless10m", "DCM", 
@@
                       labels=c("-20m", "-10m", "DCM", "+10m", "+20m"))) %>%
-  ggplot() +
+  ggplot(aes(label=tripl)) +
   geom_vline(xintercept = 0) +
@@
 
-  - Discussion: discuss how the cyclone+20m is pretty similar to the anticyclone samples and why
-  - Discussion: less of a clear depth trend despite the stats saying equal?
-  - Discussion: two possible models, one where samples from the same absolute depth are similar, one where samples from the same distance from DCM are similar.
-Can I argue for which one is better here?
+  - Discussion: I still really like the idea of modeling things according to the slope of their change across the eddy dipole.
+Light-sensitive or pressure-sensitive (e.g.
+sunscreen molecules, piezolytes) metabolites wouldn't change at all, DCM-specific metabolites (e.g.
+chlorophyll) would change by 20m or so as the DCM does, nutrient-sensitive metabolites (guanine?
+glutamate/glutamine?) would track with the isonutrials, and the salinity or temperature sensitive metabolites (aka pure osmolytes) would track with isopycnals.
+Do I have enough data to show that here?
+I forget what my early attempts to quantify these gradients looked like.
 
 To characterize the observed differences in multivariate space we used k-means clustering as an unsupervised way to identify dominant trends within the data (Figure 4).
-We found that a majority of the metabolites (Clusters 1 and 2 in Figure 4, 63% of the total) fell into clusters with larger peak areas in the cyclone, as expected, while also detecting 32 metabolites that clustered such that the mean metabolite was enriched in the anticyclone (Cluster 4, Figure 4).
+We found that a majority of the metabolites (Clusters 1 and 2, 63% of the total) fell into clusters with larger peak areas in the cyclone, as expected, while also detecting 32 metabolites that clustered such that the mean metabolite was enriched in the anticyclone (Cluster 4, Figure 4).
 Cluster 2 notably had a distinct decrease in relative peak area with depth, while the other clusters had much less clear depth trends.
@@
 
-```{r k-means clustering}
+```{r k-means clusterplot}
 set.seed(123)
@@
   coord_cartesian(ylim = c(1, 5.5)) +
-  theme_bw()
-
-ggsave("kclust_gp.png", plot = kclust_gp, device = "png", 
-       path = "figures", width = 6.5, height = 2.5, units = "in", dpi = 300)
-```
-
-![](figures/kclust_gp.png)
-
-*Figure 4: k-means clustering performed on the high-resolution depth samples from the centers of each eddy.
-Points denote the average z-scored peak area for both known and unknown metabolites across the samples and are colored by the eddy from which they were taken.
-Clusters have been ordered by number of metabolites in each group and the total is denoted in the panel titles.
-Both depth trends (mostly a net decrease in metabolites with depth) and eddy effects (cyclonic enrichment in clusters 1 and 2, anticyclone enrichment in cluster 4) are observable.*
-
-  - Discussion: what does the existence of cluster 3 mean? Tracking with biomass exactly or just too much noise?
-
-```{r cluster cor w metadata, eval=FALSE}
-filled_file_metadata %>% 
-  filter(ms_run=="MC") %>% 
-  distinct(depth, euk_abund, sla_class) %>% 
-  left_join(kclust_df) %>%
-  filter(cluster==1) %>%
-  filter(sla_class=="Cyclone") %>%
-  # with(cor(mean_rank, euk_abund))
-  ggplot() +
-  geom_point(aes(x=euk_abund, y=mean_rank, color=sla_class))
-```
-
-We also investigated the response of individual metabolites in these high-resolution depth profile samples, expecting to find the strongest signal in this data given the extreme SLA values sampled and the large number of compounds with an SLA separation in the k-means analysis.
-This expectation was substantiated by a majority of the metabolites differing significantly between the two eddies (121/228, α = 0.05).
-Of those, 15 were enriched in the anticyclone and 106 were more abundant in the cyclone.
-As expected, all compounds enriched in the anticyclone were part of Cluster 4 and all those enriched in the cyclone belonged to either cluster 1 or 2 (Figure 5).
+  theme_bw() +
+  theme(plot.background = element_rect(fill = "white"))
+```
 
@@
   geom_hline(yintercept = 0) +
-  geom_hline(yintercept = 2, lty=2) +
+  geom_hline(yintercept = 1.3, lty=2) +
   geom_point(size=2, stroke=1) +
@@
   theme_bw() +
-  labs(x="Log-2 fold-change cyclone enrichment", y="Mann-Whitney FDR-corrected p-value",
+  labs(x="Log-2 fold-change cyclone enrichment", y="Mann-Whitney p-value\n(FDR-corrected)",
        shape="Metabolite\nstatus", color="k-means\ncluster") +
   guides(color=guide_legend(order = 2), shape=guide_legend(order=1)) +
-  theme(axis.text.y = ggtext::element_markdown())
-
-ggsave("volcanoplot_gp.png", plot = volcanoplot_gp, device = "png", 
-       path = "figures", width = 6.5, height = 3, units = "in", dpi = 300)
-```
-
-![](figures/volcanoplot_gp.png)
-
-*Figure 5: Volcano plot of known and unknown metabolites from high-resolution depth sampling at the deep chlorophyll maximum (DCM) of each eddy pole.
-Points correspond to the FDR-corrected p-value estimated by the nonparametric Mann-Whitney U test and the log$_2$ fold-change calculated with the average peak area in the cyclone divided by the average peak area in the anticyclone.
-Colors have been assigned using the k-means clusters from Figure 4 above and shapes have been assigned based on the status of the mass feature as either a known metabolite that was run with an authentic standard or an unidentified metabolite.*
+  theme(axis.text.y = ggtext::element_markdown(),
+        plot.background = element_rect(fill = "white"))
+```
+
+```{r kclust volcano plot combine}
+kclust_volcano_gp <- egg::ggarrange(
+  kclust_gp, volcanoplot_gp + theme(legend.position = "none"), 
+  ncol=1, heights = c(2, 1.5)) %>%
+  cowplot::plot_grid(
+    cowplot::get_legend(volcanoplot_gp), rel_widths = c(0.85, 0.15)
+  )
+ggsave("kclust_volcano_gp.png", plot = kclust_volcano_gp, device = "png", 
+       path = "figures", width = 6.5, height = 4, units = "in", dpi = 300,
+       bg="white")
+```
+
+![](figures/kclust_volcano_gp.png)
+
+*Figure 4: Distribution of metabolites in the high-resolution depth samples from the centers of each eddy.
+The upper row of plots shows k-means clusters where points denote the average z-scored peak area for both known and unknown metabolites across the samples and are colored by the eddy from which they were taken.
+Clusters have been ordered by number of metabolites in each group and the total is denoted in the panel titles.
+Both depth trends (mostly a net decrease in metabolites with depth) and eddy effects (cyclonic enrichment in clusters 1 and 2, anticyclone enrichment in cluster 4) are observable.
+The lower plot shows the individual known and unknown metabolites where points correspond to the FDR-corrected p-value estimated by the nonparametric Mann-Whitney U test and the log$_2$ fold-change calculated with the average peak area in the cyclone divided by the average peak area in the anticyclone.
+Colors have been assigned using the k-means clusters and shapes have been assigned based on the status of the mass feature as either a known metabolite that was run with an authentic standard or an unidentified metabolite.*
+
+```{r cluster cor w metadata, eval=FALSE}
+filled_file_metadata %>% 
+  filter(ms_run=="MC") %>% 
+  distinct(depth, euk_abund, sla_class) %>% 
+  left_join(kclust_df) %>%
+  filter(cluster==1) %>%
+  filter(sla_class=="Cyclone") %>%
+  # with(cor(mean_rank, euk_abund))
+  ggplot() +
+  geom_point(aes(x=euk_abund, y=mean_rank, color=sla_class))
+```
+
+We also investigated the response of individual metabolites in these high-resolution depth profile samples, expecting to find the strongest signal in this data given the extreme SLA values sampled and the large number of compounds with an SLA separation in the k-means analysis.
+This expectation was substantiated by a majority of the metabolites differing significantly between the two eddies (121/228, α = 0.05).
+Of those, 15 were enriched in the anticyclone and 106 were more abundant in the cyclone.
+As expected, all compounds enriched in the anticyclone were part of Cluster 4 and all those enriched in the cyclone belonged to either cluster 1 or 2 (Figure 4).
 
@@
 Of note, the 173.09211 mass feature noted above was among the most significantly different between the two eddies, along with two isomers at an *m/z* of 170.11756 and retention times around 8-9 minutes that increased by a factor of 1.5 in the anticyclone (putative formula C$_7$H$_{14}$N$_4$O).
-
-  - Discussion: implications of the mostly diatom/dinoflagellate metabolites upregulated in cyclone likely reflecting the fact that those are the species most abundant.
 
@@
 
-## Falkor data reveals different community and response to SLA
+## Falkor data reveals a different community and response to SLA
 
@@
 
-The Falkor data was characterized by high inter-replicate variability relative to the samples from the MESOSCOPE cruise, with surface samples in particular highly variable in multivariate space (Figure 6).
+The Falkor data was characterized by high inter-replicate variability relative to the samples from the MESOSCOPE cruise, with surface samples in particular highly variable in multivariate space (Figure 5).
 Despite this, we still saw the importance of sea level anomaly (SLA) as a significant explanatory factor in the dataset (R$^2_{SLA} = 0.090$, p-value = 0.025), after the dominant depth differences between the surface (25m) and DCM (110-120m) (R$^2_{depth} = 0.251$, p-value << 0.001).
@@
 
-*Figure 6: Non-metric multidimensional scaling (NMDS) plot from the Falkor cruise data, in which points correspond to individual samples and have been colored by their corrected sea level height anomaly and shaped by the depth from which they were collected.
+*Figure 5: Non-metric multidimensional scaling (NMDS) plot from the Falkor cruise data, in which points correspond to individual samples and have been colored by their corrected sea level height anomaly and shaped by the depth from which they were collected.
 NMDS stress values (s) have been reported in the bottom left corner, while PERMANOVA R$^2$ and p-values are reported in the top left.
@@
 
-  - Discussion: is there a common thread between the up/down regulated compounds?
-Urea cycle for Falkor cruise w aspartate/citrulline/ornithine/malic acid?
-Connect back to the differing community composition of the two cruises?
-Any chance we have urea numbers for Falkor at all?
-Ask about those transcripts?
-I want to connect this to the diatom urea cycle paper (10.1038/nature10074) and imply that during Falkor there was an ornithine-urea cycle disruption (10.1038/s41598-020-73715-8) but there's very few diatoms in Falkor???
-Does the same thing happen in haptophytes?
-It's not like this is different within the cruise but across cruises still seems interesting.
-
 We were also able to use the Falkor dataset to test whether the compounds detected as significantly different in the MESOSCOPE dataset were also different in this eddy pair.
-We expected to find abundant TMAO and more hydroxyisoleucine at the surface than the DCM in addition to most compounds enriched in the cyclone DCM with biomass, especially the six compounds mentioned above in the results of Figure 5 (trigonelline, homarine, DMSAc, DMSP, taurine, isethionic acid).
+We expected to find abundant TMAO and more hydroxyisoleucine at the surface than the DCM in addition to most compounds enriched in the cyclone DCM with biomass, especially the six compounds mentioned above in the results of Figure 4 (trigonelline, homarine, DMSAc, DMSP, taurine, isethionic acid).
 
@@
 
-  - Discussion: differences between this and MESOSCOPE could be due in large part to the different community composition here - MESOSCOPE looked like a big diatom bloom (according to Benedetto paper) while Sam's results show that this was more dinoflagellate/haptophyte and diatoms had "no trend".
-Maybe also eddy age?
-  - Discussion: Surface more variable than DCM in Falkor (kinda also for MESOSCOPE?)
-  - Discussion: What the hell is with these samples.
-There's an insane amount of variance within triplicates and the SLA signal is highest at the *surface* despite Sam's data and the MESOSCOPE dataset strongly implying the opposite should be true - although if a lot of the metabolism is driven by haptophytes then those are much more abundant at the surface?
-
 # Discussion
 
+## Multivariate approaches reveal metabolome-wide shifts across eddy dipoles
+
 Despite its historical reputation as a static and unchanging environment, the North Pacific Subtropical Gyre shows significant variation in its particulate composition due to mesoscale eddies.
+We sampled the metabolome across two sets of eddy dipoles to explore the total effect of sea level anomaly (SLA).
+
+In all of the datasets analyzed here, we detected a significant difference in the composition of the metabolome.
+This effect was strongest in the samples taken during the Lagrangian stations at the center of each eddy in the 2017 MESOSCOPE cruise, with nearly a quarter of the total variance explained by the eddy from which the samples were taken.
+In the samples taken along the eddy transect, the largest effect was detected in the deepest samples taken from 175 meters and from the DCM, with much less of a response at the surface.
+In the Falkor samples, however, the opposite was true with a larger SLA response at the surface than that at the DCM.
+In each case, the differences introduced by SLA were smaller than those of depth when compared directly, with even the high-resolution sampling around the DCM finding slightly more variance explained by the 40 meter difference in sampling depth than the 40 centimeter difference in SLA.
+This result agrees well with prior research by @Heal2021 and @Kumler2023 showing that depth has a large effect on the metabolome.
+
+Although possible that these differences are due to latitudinal shifts or simply background variation in the gyre environment, the transect data in particular implies that this is more due to eddies given the relative similarity of samples taken outside of the eddies.
+The transect sampled both outside of each edge and and between the two eddies across a large spatial gradient (~4° latitude and 2° longitude), yet the non-eddy samples from all three locations grouped together, showed similar median metabolite concentrations, and were visually similar in the most abundant metabolite composition.
+These results imply that strong cyclonic and anticyclonic eddies represent endpoints in NPSG composition, in agreement with results from @Barone2019 which found the most extreme values in the Hawaii Ocean Time series data to be usually detected when eddies passed over Station ALOHA.
+We also detected the largest differences in the DCM metabolome at the exact center of the eddy, with Station 12 highly distinct from even the nearest stations, while no such stark difference was found in the anticyclone (Figures 1 and 2).
+This perhaps indicates that the exact center of the eddy at the DCM is a unique metabolome relative to the rest of the transect rather than a smooth continuation of the existing trends.
+
+A large portion of this effect was due to differences in biomass, with earlier work in the same location by @Barone2022 showing that particulate carbon (PC), chlorophyll, and beam attenuation were all 20-80% higher in the DCM of the cyclone relative to the anticyclone.
+We see this reflected particularly well in the targeted metabolites measured here, with clear depth trends and a strong correlation between total metabolite concentration and PC values.
+However, even after controlling for biomass effects by normalizing each sample to the sum of signal measured within it we still found a significant SLA effect, likely due to shifts in the community composition beyond bulk biomass as species are reduced or replaced entirely by the increased eukaryote presence in the cyclone DCM.
+
+High-resolution depth sampling around the DCM at the eddy centers allowed us to explore the multivariate interactions between depth and SLA.
+Given that the difference in DCM depth between the two eddies was approximately 20 meters, our 10-meter sampling resolution allowed us to explore two potential models for metabolites.
+The first model, one driven largely by pressure or light availability, expects the absolute depth to be a stronger predictor than the difference in depth from the DCM center.
+The second model is more sensitive to dissolved water column parameters such as salinity or temperature, and predicts that samples equal distances from the DCM will be the most similar.
+The strong correlation of both depth and SLA with the first principal component of the PCA indicates that the largest variance in the system is best captured by a combined model with largest PC1 values in the deep anticyclone and smallest in the shallow cyclone.
+
+```{r multimodel testing maybe, eval=FALSE}
+
+dist_mat <- untarg_peaks %>%
+  left_join(filled_file_metadata %>% distinct(ms_run, station, filename, depth), 
+            by="filename") %>% filter(ms_run=="MC") %>% 
+  mutate(depth=factor(depth, levels=c("DCMless20m", "DCMless10m", "DCM", 
+                                      "DCMplus10m", "DCMplus20m"))) %>%
+  arrange(feature, station, depth) %>%
+  mutate(filename=fct_inorder(filename)) %>%
+  group_by(feature) %>%
+  mutate(norm_area=rank(norm_area)) %>%
+  select(feature, filename, norm_area) %>%
+  pivot_wider(names_from=feature, values_from="norm_area") %>%
+  mutate(filename=str_remove_all(filename, "180205_Smp_|.mzML")) %>%
+  column_to_rownames("filename") %>%
+  dist(method = "manhattan") %>%
+  as.matrix()
+dist_mat %>%
+  as.data.frame() %>%
+  rownames_to_column("filename_a") %>%
+  pivot_longer(-filename_a, names_to="filename_b", values_to = "dist") %>%
+  mutate(filename_a=fct_inorder(filename_a)) %>%
+  mutate(filename_b=fct_inorder(filename_b)) %>%
+  filter(as.numeric(filename_b)>=as.numeric(filename_a)) %>%
+  ggplot() +
+  geom_raster(aes(x=filename_a, y=filename_b, fill=dist)) +
+  coord_fixed() +
+  scale_fill_gradient2(midpoint = 2200) +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle=90, hjust=1, vjust=0.5))
+
+expl_df <- filled_file_metadata %>% 
+  distinct(ms_run, station, filename, depth, samp_type) %>%
+  left_join(untarg_peaks, by="filename") %>%
+  filter(ms_run=="MC") %>%
+  filter(samp_type=="Smp") %>%
+  mutate(model1_groups=as.numeric(factor(depth))) %>%
+  mutate(model2_groups=case_when(
+    station=="L1" & depth %in% c("DCMless20", "DCMless10") ~ 1,
+    station=="L1" & depth=="DCM" | station=="L2" & depth=="DCMless20" ~ 2,
+    station=="L1" & depth=="DCMplus10" | station=="L2" & depth=="DCMless10" ~ 3,
+    station=="L1" & depth=="DCMplus20" | station=="L2" & depth=="DCM" ~ 4,
+    TRUE ~ 5
+  )) %>%
+  distinct(filename, model1_groups, model2_groups)
+
+vegan::adonis2(
+  dist_mat~model1_groups+model2_groups, data = expl_df, by = "margin", 
+  permutations = 999, method = "manhattan")
+```
+
+The NMDS plots of the high-resolution depth samples were also illustrative of sample similarity, with clear SLA and depth trends.
+The anticyclonic samples grouped tightly and showed little variance with depth or triplicate when compared to the cyclonic samples with the exception of DCM triplicate B, which was highly distinct for unknown reasons.
+The cyclonic samples, on the other hand, were much more variable.
+One notable aspect of their clustering was the way in which the deepest samples (DCM plus 20m) grouped most closely to the anticyclonic samples, perhaps indicating that below the DCM the metabolome rapidly approaches a uniform deep-water signal as seen in the 175 meter samples of the eddy transect in @Kumler2023, which showed that the samples from the deep euphotic zone were more similar to each other than those taken from the surface or DCM were to other surface or DCM samples.
+
+Given the confluence of both depth and SLA signals and the way that the boxplots of Figure 1 and Supplemental figure 3 would confound the SLA signal with the depth variance, we used k-means clustering to group similar compounds and provide a reduced dimensionality space for visualization.
+This revealed four major patterns of metabolite response, with a majority of compounds responding to eddy state (Clusters 1, 2, and 4 in Figure 4).
+It is interesting to note that the sole cluster in which abundances in the anticyclone are greater than those of the cyclone is also the only cluster to generally increase in abundance with depth (Cluster 4), perhaps indicating that compounds more concentrated in the anticyclone are the same kind of degradation products and recalcitrant carbon that is found at depth.
+Cluster 3, which shows no clear SLA difference and only a slight change with depth, includes compounds which are either unaffected by the environment or simply lack an environmental signal due to large inter-sample variability.
+
+The Falkor cruise upset many of the expectations we developed during the MESOSCOPE cruise in the previous year.
+Most surprising was the large difference between the eddies detected at the *surface*, while the DCM samples were functionally indistinguishable.
+This contrasts directly with the results from MESOSCOPE and the results in @Dugenne2023 and @Gleish2023, in which the largest differences were found at the DCM or below.
+This was especially surprising given the large inter-replicate differences between the Falkor surface samples, while the DCM samples tended to be much more visually consistent.
+
+## Univariate approaches highlight individual metabolites responding to lifted and depressed isopycnals
+
+Several molecule- or pathway-specific narratives emerged from the metabolomic data described above as well.
+The untargeted approach used here allowed us to describe and chatacterize signals from small molecules whose identity was unknown, a particularly promising approach in open ocean gyres where the largest fraction of unknowns is found [@Heal2021].
+We found that this was especially true in anticyclonic eddies at all depths, with the diversity of the cyclone much better characterized by our list of authentic standards.
+Many of the trends detected below persisted even after normalizing within each sample, indicating that the shifts discussed have implications beyond simple scaling with biomass.
+
+Of those molecules whose identity was known, the clearest response to SLA was that of the enigmatic osmolyte homarine at the DCM.
+This abundant compound increased approximately threefold in concentration from < 50 pM in the anticyclone to around 150 pM in the center of the cyclone in both the eddy transect and eddy center datasets.
+The pattern was weaker in the noisier Falkor data but the largest concentrations were still detected in the cyclone and the median cyclone measurement was greater than the maximum value for the anticyclone.
+This molecule has a well-established role as an osmolyte in eukaryotic phytoplankton and cyanobacteria [@Gebser2013; @Dawson2020; @Heal2021; @Durham2022] as well as a documented decrease in abundance with depth [@Heal2021], indicating that this shift is potentially due to any combination of differences in salinity, DCM depth, total biomass, and community composition.
+Homarine's tight correlation with trigonelline, however, is entirely mystifying as the two molecules are not well-known to have a strong biological relationship beyond their structural similarity.
+
+```{r homarine by SLA plot, eval=FALSE}
+targ_hom <- targ_peaks %>%
+  filter(compound_name=="Homarine") %>%
+  left_join(filled_file_metadata) %>%
+  filter(samp_type=="Smp")
+mt_hom_gp <- targ_hom %>%
+  filter(ms_run=="MT") %>%
+  filter(depth=="DCM") %>%
+  ggplot(aes(x=sla_corr, y=nM)) +
+  geom_hline(yintercept = 0) +
+  geom_smooth(method = "lm", formula="y~x", color="black") +
+  geom_point(aes(fill=sla_corr), pch=21, size=4) +
+  scale_fill_gradient2(
+    low = scales::muted("blue"), high = scales::muted("red"),  mid = "grey90",
+    aesthetics = "fill", guide = NULL
+  ) +
+  coord_cartesian(ylim = c(0, 0.24), xlim=c(-15.77, 24.99), expand = FALSE) +
+  labs(x="Sea level anomaly (cm)", y="Concentration of homarine (nM)") +
+  theme_bw()
+mc_hom_gp <- targ_hom %>%
+  filter(ms_run=="MC") %>%
+  ggplot(aes(x=station, y=nM, color=station)) +
+  geom_hline(yintercept = 0) +
+  geom_boxplot(linewidth=0.75) +
+  coord_cartesian(ylim = c(0, 0.24), xlim=c(0.5, 2.5), expand = FALSE) + 
+  scale_x_discrete(breaks = c("L1", "L2"), labels = c("Cyclone\ncenter", "Anticyclone\ncenter")) +
+  scale_color_manual(values = c(scales::muted("blue"), scales::muted("red")),
+                     breaks = c("L1", "L2")) +
+  theme_bw() +
+  theme(axis.text.y = element_blank(), axis.title.y = element_blank(),
+        legend.position = "none", axis.title.x = element_blank())
+egg::ggarrange(mt_hom_gp, mc_hom_gp, nrow=1, widths = c(2, 1))
+
+targ_hom %>%
+  filter(ms_run=="FK") %>%
+  filter(depth=="Surface") %>%
+  # with(t.test(formula=nM~sla_class))
+  # with(wilcox.test(formula=nM~sla_class))
+  ggplot(aes(x=sla_class, y=nM, color=sla_class)) +
+  geom_hline(yintercept = 0) +
+  geom_boxplot(linewidth=0.75) +
+  coord_cartesian(ylim = c(0, 0.5), xlim=c(0.5, 2.5), expand = FALSE) +
+  scale_x_discrete(breaks = c("Cyclone", "Anticyclone"), labels = c("Cyclone\ncenter", "Anticyclone\ncenter")) +
+  scale_color_manual(values = c(scales::muted("blue"), scales::muted("red")),
+                     breaks = c("Cyclone", "Anticyclone")) +
+  theme_bw() +
+  theme(axis.text.y = element_blank(), axis.title.y = element_blank(),
+        legend.position = "none", axis.title.x = element_blank())
+```
+
+Similarly to homarine, isethionate is a known osmolyte found exclusively in eukaryotes according to a previous study [@Durham2022] and strongly associated with a few diatoms in particular [@ref or culture data].
+This compound was enriched in the cyclone of the DCM in both the MESOSCOPE and Falkor cruises along with its precursor taurine, validating earlier results from @Barone2022 that showed strong enrichment of most eukaryotes at the DCM and *Pseudo-nitzschia* in particular.
+[Expand on this with after additional meetings with Bryn and following up with Sam - TPA (K03851 should definitely have been picked up, poly-A selected for euks?, see if iseJ (the 1.1.2-) can still be searched for without a KO number (yes according to Frank bc the sequence exists but complicated by catabolism/anabolism))]
+
+```{r isethionate and taurine, eval=FALSE}
+targ_peaks %>%
+  filter(compound_name%in%c("Isethionic acid", "Taurine")) %>%
+  filter(filename!="180821_Smp_MS11C215m_A.mzML") %>%
+  left_join(filled_file_metadata) %>%
+  filter(samp_type=="Smp") %>%
+  mutate(depth=fct_inorder(depth)) %>%
+  mutate(ms_run=factor(ms_run, levels=c("MT", "MC", "FK"))) %>%
+  mutate(sla_class=factor(sla_class, levels=c("Anticyclone", "Neither", "Cyclone"))) %>%
+  # group_by(compound_name, ms_run, sla_class, depth) %>%
+  # summarise(nM=mean(nM)) %>%
+  # pivot_wider(names_from = sla_class, values_from = nM)
+  ggplot() +
+  geom_boxplot(aes(x=depth, y=nM, color=sla_class)) +
+  facet_grid(compound_name~ms_run, scales="free")
+```
+
+Taurine was also among the compounds highly enriched in the anomalous surface sample from Station 11 (MS11C215m_A), likely along with its trimethylated form taurine betaine.
+While putative in our analysis, trimethyltaurine has been previously reported in corals and red algae, where it probably has a destabilizing influence on proteins [@Impellizzeri1975; @Hill2017; @Bruzdziak2018].
+We have a few hypotheses for the striking pattern in these compounds and the other 13 metabolites detailed above in Results.
+Most simply, the signal could simply be contamination during the sampling process, with some unknown material dominated by these compounds added to the filter during collection or processing.
+More interestingly, the signal could represent a real biological difference, perhaps indicating a pathway or organism highly abundant in this sample alone.
+As noted above, the MS11C215m_A sample was collected during a *Trichodesmium* bloom, an environment characterized by large heterogeneity even between replicate samples [@Rouco2014; @Rouco2016].
+10 liters of seawater were filtered for each sample but it is conceivable that a tuft or puff of *Trichodesmium* was captured in only this one as previously noted by @Rouco2014.
+The presence of a predictable trend across multiple chemically similar metabolites is the strongest bit of evidence in favor of the *Trichodesium* hypothesis, but the fact that these metabolites do not appear robustly in the bloom samples and that homoserine betaine did not appear to be especially abundant challenges it.
+One final note is that homoserine betaine was only documented as the sole major compatible solute in the Clade III *T.
+erythraeum* strain IMS101, while the bloom was more likely due to Clade I species given their relative abundance in the NPSG [@Rouco2016].
+[Email Sonya about the presence of compatible solute genes in Clade I tricho?
+Especially taurine/TMT?
+Ask about IFCB samples (they're on the dashboard http://ifcbdb.soest.hawaii.edu/timeline?dataset=MESOSCOPE&bin=D20170708T043509_IFCB000 I just don't know what to ask)]
+
+The strong SLA signal detected among the 175 meter samples was partially surprising to us, having expected the effect to be strongest at the DCM where isopycnals are interacting strongly with light from the surface.
+Instead, we found more compounds overall to be significantly different across the eddy transect in the 175 meter samples than we did at the DCM (α=0.05, 43 compounds at 175 meters vs 22 at the DCM).
+The enrichment of nucleobases in the 175 meter samples may hint at bacterial biomass enrichment, as nucleobases tend to be highly abundant in bacterial cultures [@ref or culture data].
+The abundance of a mass feature putatively identified as acetylcholine in the 175 meter cyclone samples followed the same general trend as arsenobetaine, both enriched at depth and in the anticyclone, and may be another compound that's the result of heterotrophic degradation of phytoplankton metabolites [@Heal2021].
+These results raise important questions about the depth at which the community is isolated from SLA effects, with additional data from @Barone2022 and @Gleish2023 indicating that even at 250 meters eddy effects are still discernable.
+
+A final pattern worth noting is the apparent upregulation of metabolites in the urea-ornithine cycle (OUC) among the samples taken during the Falkor cruise.
+Aspartate, citrulline, ornithine, and malic acid were all found to be more abundant in the Falkor samples relative to MESOSCOPE.
+The OUC is a staple of metazoan metabolism and was recently revealed to be present in stramenopiles as well, where it enables rapid response to increases in nitrogen supply [@Allen2011].
+The enrichment of these metabolites in the Falkor cruise may indicate that there was still rapid uptake and turnover of nitrogen in the 2018 eddy pair while the 2017 eddies were more stable, despite both eddies being in a similar state of stable anticyclone and weakening cyclone [@Dugenne2023].
+
+More broadly, differences between the two cruises are most likely due to differences in eddy age and the earlier sampling time of the 2018 cruise (March/April instead of the June/July sampling in 2017) resulting in shifts in community composition and response.
+@Barone2022 documents the 2017 cruise community composition, reporting abundant Prymnesiophyceae at both the surface and DCM (largely), with other significant contributions from dinoflagellates, diatoms, and green algae.
+In the 2018 cruise both haptophytes (likely prymnesiophytes?) and dinoflagellates again dominated the eukaryotic fraction, though notably very little was contributed by diatoms [@Gleish2023].
+Overall we are inclined to agree with @Dugenne2023 and @Harke2021 in their concerns about generalizing eddy effects due to uncontrolled differences in these crucial factors and will note that many of the difficulties documented there are exacerbated when working with highly multidimensional metabolite data.
 
@@
 ## Cruise information
-Samples were collected from two cruises in the North Pacific Subtropical Gyre near Station ALOHA that targeted strong mesoscale eddy features as described in @Dugenne2023 and @Gleich2023.
+Samples were collected from two cruises in the North Pacific Subtropical Gyre near Station ALOHA that targeted strong mesoscale eddy features as described in @Dugenne2023 and @Gleish2023.
 Briefly, the 2017 MESOSCOPE cruise (KM1709 on the *R/V Kilo Moana*) consisted of a transect across the cyclonic and anticyclonic poles of the eddy dipole as well as two long-term Lagrangian stations at the center of each eddy.
@@
 Samples were also processed via the untargeted workflow detailed in @Kumler2023.
-In short, XCMS was used for peak detection and followed by manual inspection of each mass feature for quality with only those peaks classified as "Good" used in the analysis here.
-Features were then matched with an internal standard as described above.
+In short, XCMS [@R-xcms] was used for peak detection and followed by manual inspection of each mass feature for quality with only those peaks classified as "Good" used in the analysis here.
+Features were then matched with and normalized to an internal standard as described above.
 
@@
 Non-parametric tests were used when data were unlikely to obey parametric assumptions and permutational statistics were preferred wherever possible.
-Multiple comparison problems were controlled where necessary using a false-discovery rate correction.
+Multiple comparison problems were controlled where necessary using a false-discovery rate correction [@Benjamini1995].
+For the analyses involving the eddy transect samples, sea level anomaly was treated as a continuous variable with the exception of the PERMANOVA, for which SLA was categorized as cyclonic if less than -5 cm and anticyclonic if greater than 5 cm, with a third category for stations in between [@Anderson2001].
+Both the eddy center and Falkor data treated SLA as a categorical variable.
 
 # Conclusions
+
+Here, we measure the effect of mesoscale eddies on the metabolome of the North Pacific Subtropical Gyre in both absolute and relative space. [TBD].
+
 # References
+
+```{r packages, eval=FALSE}
+packages_used <- list.files("..", pattern = "*\\.R$|*\\.Rmd$", recursive = TRUE) %>%
+  str_subset("old|scripts", negate=TRUE) %>%
+  map(function(file_name){
+    v <- readLines(paste0("../", file_name))
+    libs <- unique(str_extract(v, "(?<=library\\().*?(?=\\)|,)"))
+    refs <- unique(str_extract(v, "[[:alnum:]]*(?=::)"))
+    pkgs <- unique(c(na.omit(libs), na.omit(refs)))
+  }) %>% unlist() %>% unique()
+knitr::write_bib(packages_used[nchar(packages_used)>0], file = "r-packages.bib")
+```
 

### 2023-11-11 manuscript/manuscript.Rmd
Easy edits to intro

@@
 ---
-title: "Metabolites reflect the mesoscale variability in biogeochemistry introduced by sea-surface eddies in the North Pacific Subtropical Gyre"
+title: "Metabolites reflect variability introduced by mesoscale eddies in the North Pacific Subtropical Gyre"
 author: "William Kumler"
@@
 
-Until recently, the ocean's subtropical gyres were considered homogenous and static [@Karl1999].
-In the past 25 years, however, regular sampling and advances in measurement technology have revealed its variability on time scales from the diel to the seasonal [@Karl2017; @Karl2021].
-A major source of this variability comes in the form of mesoscale eddies in which water is entrained into circular surface currents tens to hundreds of kilometers in diameter [@Karl2017; @McGillicuddy2016].
+High frequency, long-term observations at Station ALOHA in the North Pacific Subtropical Gyre (NPSG) over the past 25 years have revealed temporal and spatial variability in what had prevously been considered a homogenous environment.[@Karl1999; @Karl2017; @Karl2021] A major source of this variability comes in the form of mesoscale eddies in which water is entrained into circular surface currents tens to hundreds of kilometers in diameter [@Karl2017; @McGillicuddy2016].
 Prior research has established the importance of these mesoscale features in introducing nutrients into the euphotic zone and altering the community composition [@Sweeney2003; @Barone2019].
-In particular, the deep chlorophyll maximum (DCM) is known to correspond to the top of the nitricline and differs dramatically between cyclonic and anticyclonic eddies in depth and intensity.
-@Barone2022 reported a shift in the microbial DCM community with eukaryotic phytoplankton enriched and prokaryotic phytoplankton depleted in cyclonic eddies with the converse true for anticyclonic ones, while @Hawco2021 and @Dugenne2023 showed that some of this shift was due to alterations in dissolved iron chemistry and nitrogen fixation, especially at the surface.
-
-These recent works have highlighted the effect of eddies on both bulk oceanographic parameters such as particulate carbon and nitrogen and the composition of the microbial community, but we still lack direct links between the genomic and transcriptomic assays occurring regularly around the world and the particulate organic matter produced by those organisms [@Kharbush2020].
-This missing link is bridged by metabolomics, a nascent field studying the structure and function of the cell on a molecular level.
-Of particular interest within the metabolome are those small, polar molecules that represent the fundamental unit of cell biology and function as building blocks for macromolecules [@Moran2022].
-These small molecules also have roles of their own within the cell, acting as osmolytes, antioxidants, nutrient storage compounds, and more [@Mojzes2020; @Kharbush2020].
-Despite these critical functions, they are difficult to study in the marine environment due to their low concentrations and similarity to the abundant salt ions, resulting in challenges in detection and quantification [@Boysen2018; @Moran2022; @Sacks2022].
-
-Here, we use spatially resolved measurements of metabolites in particles across an eddy dipole in the North Pacific Subtropical Gyre (NPSG) to categorize unidentified molecular features based on their co occurrence with a set of simultaneously measured physical, chemical, and biological parameters.
+In particular, the depth and intensity of the deep chlorophyll maximum (DCM) at the top of the nitricline differs dramatically between cyclonic and anticyclonic eddies.
+@Barone2022 reported a shift in the microbial community at the DCM with eukaryotic phytoplankton enriched and prokaryotic phytoplankton depleted in cyclonic eddies.
+@Hawco2021 and @Dugenne2023 showed that some of this difference was due to alterations in dissolved iron chemistry and nitrogen fixation, especially at the surface.
+
+Here, we use measurements of metabolites in particles across an eddy dipole in the North Pacific Subtropical Gyre (NPSG) to categorize both known compounds and unidentified mass features based on their co-occurrence with a set of simultaneously measured physical, chemical, and biological parameters.
 The gyre in particular represents an intriguing opportunity for metabolomic research because it has been shown to contain the highest proportion of chemical features with unknown structure and function and have highly variable responses to changes in latitude, depth, and across diel cycles [@Heal2021; @Boysen2021; @Kumler2023].

### 2023-11-11 manuscript/manuscript.Rmd
Adding comments to manuscript text
Finished adding notes and easy edits to results section

@@
 @Hawco2021 and @Dugenne2023 showed that some of this difference was due to alterations in dissolved iron chemistry and nitrogen fixation, especially at the surface.
+[Anitra: so the surface is driving changes at the DCM?]
 
@@
 # Results
-## Metabolome changes significantly across the eddy dipole at DCM and 175m
-
-We found that the metabolome differed between the cyclone and anticyclone samples, with multivariate statistics detecting the largest differences in the DCM and 175 meter samples while the surface was more homogenous (Figure 1).
+## Metabolome variability across the eddy dipole
+
+Using data for all molecular features, we found that the metabolome differed between the cyclone and anticyclone samples, with multivariate statistics detecting the largest differences in the DCM and 175 meter samples (Figure 1).
 At the DCM, approximately 16% of the variation in the metabolome could be explained by SLA alone according to the PERMANOVA (p = 0.003) and SLA was strongly correlated with the first principal component (r = .615, % variance explained by PC1 = 27.2), indicating that this was one of the largest sources of variation in the dataset.
 Notably, the samples taken from the exact center of the cyclonic eddy were highly distinct and likely drove much of the variance explained.
-In the deep samples, 12% of the variance was explained by SLA (p = 0.008) according to the PERMANOVA and SLA was highly correlated with the second PC of the dataset (r = .697, % variance explained by PC2 = 15.4), though the first PC did not seem to have any visible pattern with metadata and appeared to capture largely variation between biological triplicates (Supplemental figure 1).
+In the 175m samples, 12% of the variance was explained by SLA (p = 0.008) according to the PERMANOVA and SLA was highly correlated with the second PC of the dataset (r = .697, % variance explained by PC2 = 15.4), though the first PC did not seem to have any visible pattern with metadata and appeared to capture largely variation between biological triplicates (Supplemental figure 1).
 At the surface, SLA trends are less evident with a larger p-value (0.087) and a lower R$^2$ (0.09) (Figure 1).
 
-To characterize this difference, we plotted an "average" metabolite per station at each depth by taking the median value of all the detected mass features after z-score normalization (Figure 1).
+To characterize this difference, we plotted an "average" metabolite peak area per station at each depth by taking the median value of all the detected mass features after z-score normalization (Figure 1).
 As expected, the average metabolite collected at the exact center of the cyclonic eddy (Station 12) was much more abundant than in any sample at this same depth in the rest of the dipole samples.
@@
 
-A large part of this metabolic response to SLA is due to the increased biomass in the cyclone DCM and the parallel increase in the anticyclone deep samples.
-We used a subset of the prior dataset, only those metabolites that we were able to quantify, and measured their in-situ concentrations in the particulate matter (Figure 2).
+We expected a large part of the difference in metabolomes with respect to SLA to be due to the increased biomass in the cyclone DCM and the parallel increase in the anticyclone deep samples.
+Within our molecular feature population, a subset are known metabolites that we quantified using standards (Figure 2).
 
@@
 *Figure 2: Concentrations of known compounds across the eddy dipole, broken down by depth.
-Stacked bars correspond to known compounds named in the legend underneath, with compounds not in the top 11 by absolute concentration grouped into the "Other" category.
+Stacked bars correspond to known compounds, with the top 11 shown and all other metabolites summed in grey (n=[TBD]).
 DMSP = dimethylsulfoniopropionate.
@@
 The eddy centers (stations 6 and 12) have been labeled with arrows of the same color scheme in the final panel.
-Particulate carbon concentrations in µmol/L are shown as white dots and use the y-axis scale on the right side (200x the metabolite scale).*
+Particulate carbon concentrations in µmol/L are shown as white dots and use the y-axis scale on the right side (200x the metabolite scale) [Anitra: This implies that the metabolites are 1/200 of the POC.
+But you have plotted compound concentration not carbon for metabolites.
+Assuming each compound has an average of 4 carbon atoms, you would have to multiply by 4 so that is 1/50 (2%) or so of POC.
+I think you need to make this clear somehow.
+Or you could make an additional plot with C for each compound.
+I see supp fig 2 has the %POC so maybe that is enough but maybe call out that figure here.][Will: I agree, let's brainstorm].*
 
@@
 We found that the overall concentration of known molecules was highest at the DCM of the cyclone, in line with our expectations and prior results.
-The most abundant quantified molecule overall was trimethylamine N-oxide (TMAO), representing an average concentration of 1.38 nM.
-At the surface, however, hydroxyisoleucine was the most concentrated metabolite at 1.53 nM due to its monotonic decrease with depth as opposed to a majority of the osmolytes (37/53) which reached maximum concentrations at the DCM instead (Figure 2).
-
-Bulk particulate carbon (PC) measurements correlated well with overall metabolite concentrations ($r= .554$, p-value < 0.001) but this was largely driven by the lower biomass and metabolite abundance at depth.
-When analyzed at each depth separately, the DCM ($r_{DCM}= .790$, p-value = 0.003) and 175 meter samples ($r_{Deep}= .554$, p-value = 0.077) had weaker but still good correlations, but no such association was detected at the surface ($r_{Surface}= -.125$, p-value = 0.714).
+The most abundant quantified molecule overall was trimethylamine N-oxide (TMAO), with an average concentration of 1.38 nM.
+At the surface hydroxyisoleucine was the most abundant metabolite at 1.53 nM due to its monotonic decrease with depth as opposed to a majority of the osmolytes (37/53) which reached maximum concentrations at the DCM instead (Figure 2).
+
+[Anitra: comments on the paragraph above: Maybe instead of this somewhat convoluted description you could just break it down by depth and say which compound was most abundant at each depth?
+Or which compound was most abundant in the cyclone and anticyclone.
+And then in a separate sentence just simply state that most compounds reached a ma at the DCM, but hydroxyisoleucine was an exception.
+Also, are you implying that hydroxyisoleucine is an osmolyte?
+Is that true?
+Also, what do you mean by "average"?]
+
+Bulk particulate organic carbon (POC) measurements correlated well with overall [Anitra: what do you mean by overall here?] metabolite concentrations ($r= .554$, p-value < 0.001) but this was largely driven by both parameters having higher values in the surface than at depth.
+When analyzed at each depth individually, the DCM ($r_{DCM}= .790$, p-value = 0.003) and 175 meter samples ($r_{Deep}= .554$, p-value = 0.077) had weaker but still good correlations, but no such association was detected at the surface ($r_{Surface}= -.125$, p-value = 0.714).
 
@@
 
-This correlation between PC and overall metabolite abundance implied that the fraction of PC represented by known metabolites was likely to be fixed across the eddy transect and this expectation was validated by followup analyses.
-Known metabolites contributed between approximately 1.5% and 5% of the carbon biomass to the total PC load, but this variance was not clearly due to either depth or SLA (Supplemental figure 2).
+This correlation between PC [Anitra: Is there ATP data?
+Might be interesting to compare.
+Will: instinct says no, I think that ATP was really only collected during the diel phase and from a few spots during lagrangian.
+TODO: Double check the dashboard] and overall metabolite abundance implied that the fraction of PC represented by known metabolites was likely to be fixed across the eddy transect and this expectation was validated by followup analyses.
+[Anitra: I do not see this at all.
+I see spikes in metabolite concentration that are not matched with spikes in PC.
+Mostly at the eddy centers.
+Do you not think this is suggestive of a change in the contribution of biomass to PC in the eddy centers?
+I would assume if you had biomass instead of PC you would be able to see biomass increase at the centers where metabolites increase.
+TODO: Check the same values using beam attenuation (interpolate if necessary)] Known metabolites contributed between approximately 1.5% and 5% of the carbon to the total PC load, but this variance was not clearly due to either depth or SLA (Supplemental figure 2).
 The only difference we observed was a slightly reduced metabolite contribution in the 175 meter samples (0.6 percentage points lower at 175 meters, p-value=0.002) while no difference was detected between the surface and DCM samples or between the cyclone, anticyclone, and outside the eddy samples (all p-values > 0.1).
-However, this lack of difference observed may have been due to the small sample sizes (18 anticyclonic samples, 9 cyclonic, and 6 neither).
+A higher proportion of PC in living biomass would translate to a higher proportion of metabolites in POC.
+This lack of difference may have been due to the small sample sizes (6 anticyclonic stations, 3 cyclonic, and 2 neither, each sampled in triplicate).
 A followup permutational sensitivity analysis indicated that a difference in means of 0.4 percentage points would have been obtained at the α < 0.05 level for SLA in the 175 meter samples if ~1.5x as many samples had been collected with mean and variances remaining similar.
+
+[TODO: switch to POC instead of PC after confirming it's only the organics]
 
@@
 
-One sample, taken from the surface at Station 11 (MS11C215m_A), was highly distinct from all others (including its paired triplicates) with several metabolites measuring two or more orders of magnitude larger in both peak area and concentration.
-Heuristic exploration of nearby masses identified a total of 14 compounds highly abundant in this one sample alone.
+One sample, taken from the surface at Station 11 (MS11C215m_A), was highly distinct from all others (including its paired triplicates) with several metabolites measuring two or more orders of magnitude larger in peak area and concentration both before and after normalization to internal standards.
+Heuristic exploration of potentially related masses identified a total of 14 compounds highly abundant in this one sample alone.
 The only identified compounds with this trend were taurine and acetyltaurine, though three others with *m/z* 146.1176, 148.0968, and 206.1388 were discovered to be likely isomers of known compounds (acetylcholine/(3-carboxypropyl)trimethylammonium, hydroxyisoleucine, dexpanthenol, respectively).
@@
 This mass was also one of the most highly enriched in the MS11C215m_A sample after taurine itself, with a peak area nearly 30x that of the next largest sample.
-No clear explanation emerged explaining which compounds were hyperabundant in this surface sample, though one possible explanation was the *Trichodesmium* bloom occurring at the same time.
+No clear explanation emerged explaining why some compounds were hyperabundant in this surface sample, though one possible explanation was the *Trichodesmium* bloom occurring at the same time.
 
@@
 
+[TODO: Clean up this table. Make sure the RT values only 2-3 sigfigs, make column names legible, etc.]
+
 We explored this *Trichodesmium* hypothesis further by analyzing additional samples taken from the bloom to see whether the compounds abundant in *Trichodesmium* were also abundant in samples taken directly from the bloom itself.
@@
 Finally, we performed a search of the untargeted data for homoserine betaine, a known *Trichodesmium* osmolyte [@Pade2016].
-Although this compound was not in our standard library due to a lack of purchasable authentic standards, we identified three mass features in the 162.1125 *m/z* window with retention times of 7.1, 8.4, and 9.9 minutes.
+Although this compound was not in our standard library due to a lack of authentic standards, we identified three mass features in the 162.1125 *m/z* window with retention times of 7.1, 8.4, and 9.9 minutes.
 The feature at RT 9.9 matched our carnitine standard's retention time, leaving the two earlier peaks with putative identifications of threonine betaine at 7.1 and homoserine betaine at 8.4 given the elution order of threonine before homoserine in our authentic standards.
@@
 
-In addition to the absolute shifts in metabolite concentration due to biomass differences, we also explored the relative composition of the particulate matter.
+In addition to absolute shifts in metabolite concentration due to biomass differences, we also explored compositional shifts in the metabolomes.
 
@@
 At the DCM, twenty-two compounds changed significantly in relative peak area across the eddy dipole.
-Seven of these were known metabolites, of which trigonelline and homarine were most enriched in the cyclone while arsenobetaine was the only metabolite with a larger peak area fraction in the anticyclone.
+Seven of these were known metabolites, of which trigonelline and homarine were most enriched in the cyclone while arsenobetaine was the only metabolite with a larger fraction of the total peak area in the anticyclone.
 Homarine showed a very large and highly significant shift, representing about 2.5% of the total peak area in the anticyclone but 7-8% in the cyclone center.
@@
 To provide context for the sea level anomaly responses detected above, we additionally performed high-resolution depth sampling around the DCM at the center of both eddy poles.
-Here, we found that the ~40 meter difference in depth and the ~40 centimeter difference in SLA explained similar amounts of variation in the data (R$^2_{depth}=0.253$, R$^2_{SLA}=0.238$) and were both highly significant factors (permutational p-values << 0.001).
+Here, we found that the ~40 meter difference in depth and the transition between the eddies explained similar amounts of variation in the data (R$^2_{depth}=0.253$, R$^2_{SLA}=0.238$) and were both highly significant factors (permutational p-values << 0.001).
 Additionally, both sample depth and SLA were well-correlated with the first principal component of the metabolite matrix (r$_{depth}=.766$, r$_{SLA}=.684$, fraction of variance explained by PC$_1$ = 30.8%) (Figure 3).
@@
 Cluster 2 notably had a distinct decrease in relative peak area with depth, while the other clusters had much less clear depth trends.
-Cluster 1 also showed a neat DCM maximum for the samples from the anticyclone, correlating well with flow cytometry counts of picoeukaryotes (r=.851) that was not present in the cyclone (r=.013).
+Cluster 1 also showed a DCM maximum for the samples from the anticyclone, correlating well with flow cytometry counts of picoeukaryotes (r=.851) that was not present in the cyclone (r=.013).
 
@@
 Of these, the putative arsenobetaine was the most significantly different among these with a peak area at the anticyclone DCM nearly quadruple that of the cyclone.
-Of note, the 173.09211 mass feature noted above was among the most significantly different between the two eddies, along with two isomers at an *m/z* of 170.11756 and retention times around 8-9 minutes that increased by a factor of 1.5 in the anticyclone (putative formula C$_7$H$_{14}$N$_4$O).
+Of note, the 173.09211 *m/z* mass feature noted above was among the most significantly different between the two eddies, along with two isomers at an *m/z* of 170.11756 and retention times around 8-9 minutes that increased by a factor of 1.5 in the anticyclone (putative formula C$_7$H$_{14}$N$_4$O).
 
@@
 
-As expected, TMAO was again one of the most abundant metabolites detected in the particulate matter with concentrations around 1.2 nM at the surface and 0.4 nM at the DCM, second only to the surprisingly high levels of the amino acid glycine (2 nM surface concentrations, 0.9 nM DCM concentrations), though the differences between the surface and DCM were not significant for either compound.
+As expected, TMAO was again one of the most abundant metabolites detected in the particulate matter with concentrations around 1.2 nM at the surface and 0.4 nM at the DCM, second only to the high levels of the amino acid glycine (2 nM surface, 0.9 nM DCM), though the differences between the surface and DCM were not significant for either compound.
 We did detect a significant difference between depths for hydroxyisoleucine, once again expected, despite the aforementioned lower concentrations in this sample set (surface mean = 0.32 nM, DCM mean = 0.21 nM, t-test p-value=0.022 with n=12 samples at each depth).

### 2023-11-11 manuscript/manuscript.Rmd
Finished adding notes and comments from Anitra

@@
 
-Despite its historical reputation as a static and unchanging environment, the North Pacific Subtropical Gyre shows significant variation in its particulate composition due to mesoscale eddies.
-We sampled the metabolome across two sets of eddy dipoles to explore the total effect of sea level anomaly (SLA).
-
-In all of the datasets analyzed here, we detected a significant difference in the composition of the metabolome.
+The North Pacific Subtropical Gyre shows significant variation in its particulate composition due to mesoscale eddies.
+We sampled the metabolome across two sets of eddy dipoles to explore the effect of sea level anomaly (SLA).
+
+In all of the datasets analyzed here, we detected a significant difference in the composition of the metabolome between eddies.
 This effect was strongest in the samples taken during the Lagrangian stations at the center of each eddy in the 2017 MESOSCOPE cruise, with nearly a quarter of the total variance explained by the eddy from which the samples were taken.
@@
 ```{r multimodel testing maybe, eval=FALSE}
-
 dist_mat <- untarg_peaks %>%
@@
 The anticyclonic samples grouped tightly and showed little variance with depth or triplicate when compared to the cyclonic samples with the exception of DCM triplicate B, which was highly distinct for unknown reasons.
-The cyclonic samples, on the other hand, were much more variable.
+In contrast, the cyclonic samples were much more variable as is perhaps expected when the biomass is in the form of larger phytoplankton that are less homogenous in the environment.
 One notable aspect of their clustering was the way in which the deepest samples (DCM plus 20m) grouped most closely to the anticyclonic samples, perhaps indicating that below the DCM the metabolome rapidly approaches a uniform deep-water signal as seen in the 175 meter samples of the eddy transect in @Kumler2023, which showed that the samples from the deep euphotic zone were more similar to each other than those taken from the surface or DCM were to other surface or DCM samples.
@@
 This contrasts directly with the results from MESOSCOPE and the results in @Dugenne2023 and @Gleish2023, in which the largest differences were found at the DCM or below.
-This was especially surprising given the large inter-replicate differences between the Falkor surface samples, while the DCM samples tended to be much more visually consistent.
+This was especially surprising given the large inter-replicate differences between the Falkor surface samples, while the DCM samples tended to be much more consistent.
 
@@
 The untargeted approach used here allowed us to describe and chatacterize signals from small molecules whose identity was unknown, a particularly promising approach in open ocean gyres where the largest fraction of unknowns is found [@Heal2021].
-We found that this was especially true in anticyclonic eddies at all depths, with the diversity of the cyclone much better characterized by our list of authentic standards.
+[Anitra: how many of your unknowns were also unknowns in Katherine's paper?] We found that this was especially true in anticyclonic eddies at all depths, with the diversity of the cyclone much better characterized by our list of authentic standards.
 Many of the trends detected below persisted even after normalizing within each sample, indicating that the shifts discussed have implications beyond simple scaling with biomass.
@@
 The pattern was weaker in the noisier Falkor data but the largest concentrations were still detected in the cyclone and the median cyclone measurement was greater than the maximum value for the anticyclone.
-This molecule has a well-established role as an osmolyte in eukaryotic phytoplankton and cyanobacteria [@Gebser2013; @Dawson2020; @Heal2021; @Durham2022] as well as a documented decrease in abundance with depth [@Heal2021], indicating that this shift is potentially due to any combination of differences in salinity, DCM depth, total biomass, and community composition.
-Homarine's tight correlation with trigonelline, however, is entirely mystifying as the two molecules are not well-known to have a strong biological relationship beyond their structural similarity.
+This molecule has a well-established role as an osmolyte in eukaryotic phytoplankton and cyanobacteria [@Gebser2013; @Dawson2020; @Heal2021; @Durham2022] as well as a documented decrease in abundance with depth [@Heal2021], indicating that this shift is potentially due to any combination of differences in salinity [Anitra: Really?
+Did salinity change much?
+Will: Half a PSU at 175m, 0.3 at DCM.
+Idk if that's enough - check Hannah's paper?], DCM depth, total biomass, and community composition.
+Homarine was tightly correlated with trigonelline, and although they have structural similarity they are not known to have a relationship beyond their shared function as osmolytes.
 
@@
 
-Similarly to homarine, isethionate is a known osmolyte found exclusively in eukaryotes according to a previous study [@Durham2022] and strongly associated with a few diatoms in particular [@ref or culture data].
+Similarly to homarine, isethionate is a known osmolyte thus far found exclusively in eukaryotes [@Durham2022] and strongly associated with a few diatoms in particular [@ref or culture data].
 This compound was enriched in the cyclone of the DCM in both the MESOSCOPE and Falkor cruises along with its precursor taurine, validating earlier results from @Barone2022 that showed strong enrichment of most eukaryotes at the DCM and *Pseudo-nitzschia* in particular.
@@
 
-Taurine was also among the compounds highly enriched in the anomalous surface sample from Station 11 (MS11C215m_A), likely along with its trimethylated form taurine betaine.
+[Anitra: OK to comment on this but needs to be 3-4 sentences] Taurine was also among the compounds highly enriched in the anomalous surface sample from Station 11 (MS11C215m_A), likely along with its trimethylated form taurine betaine.
 While putative in our analysis, trimethyltaurine has been previously reported in corals and red algae, where it probably has a destabilizing influence on proteins [@Impellizzeri1975; @Hill2017; @Bruzdziak2018].
 We have a few hypotheses for the striking pattern in these compounds and the other 13 metabolites detailed above in Results.
-Most simply, the signal could simply be contamination during the sampling process, with some unknown material dominated by these compounds added to the filter during collection or processing.
-More interestingly, the signal could represent a real biological difference, perhaps indicating a pathway or organism highly abundant in this sample alone.
+Most simply, the signal could be contamination during the sampling process, with some unknown material dominated by these compounds added to the filter during collection or processing.
+Alternatively, the signal could represent a real biological difference, perhaps indicating a pathway or organism highly abundant in this sample alone.
 As noted above, the MS11C215m_A sample was collected during a *Trichodesmium* bloom, an environment characterized by large heterogeneity even between replicate samples [@Rouco2014; @Rouco2016].
@@
 Aspartate, citrulline, ornithine, and malic acid were all found to be more abundant in the Falkor samples relative to MESOSCOPE.
-The OUC is a staple of metazoan metabolism and was recently revealed to be present in stramenopiles as well, where it enables rapid response to increases in nitrogen supply [@Allen2011].
+The OUC is a staple of metazoan metabolism and is present in stramenopiles as well where it enables rapid response to increases in nitrogen supply [@Allen2011; @Bender et al?].
 The enrichment of these metabolites in the Falkor cruise may indicate that there was still rapid uptake and turnover of nitrogen in the 2018 eddy pair while the 2017 eddies were more stable, despite both eddies being in a similar state of stable anticyclone and weakening cyclone [@Dugenne2023].

### 2023-12-01 metadata/metadata_control.Rmd
Saving some cleaned data products for use elsewhere

@@
 complete_ctddata <- bind_rows(meso_ctddata, falkor_ctddata)
+write_csv(complete_ctddata, "complete_ctddata.csv")
 ```
@@
 all_core <- bind_rows(meso_core, falkor_core)
+write_csv(all_core, "all_core_gof.csv")
 ```

### 2023-12-01 manuscript/manuscript.Rmd
Adding putative author list

@@
 title: "Metabolites reflect variability introduced by mesoscale eddies in the North Pacific Subtropical Gyre"
-author: "William Kumler"
+author: "William Kumler, Laura T. Carlson, Wei Qin, Rachel A. Lundeen, Benedetto Barone, and Anitra E. Ingalls"
 date: "`r Sys.Date()`"

### 2023-12-01 manuscript/manuscript.Rmd
Switching to msystems fig width of 6.875 inches

@@
 ggsave("nmds_and_med_metab.png", plot = nmds_and_med_metab, device = "png", 
-       path = "figures", width = 6.5, height = 4.5, units = "in", dpi = 300,
+       path = "figures", width = 6.875, height = 4.5, units = "in", dpi = 300,
        bg="white")
@@
 ggsave("MC_nmds_gp.png", plot = MC_nmds_gp, device = "png", 
-       path = "figures", width = 6.5, height = 4.5, units = "in", dpi = 300)
+       path = "figures", width = 6.875, height = 4.5, units = "in", dpi = 300)
 ```
@@
 ggsave("kclust_volcano_gp.png", plot = kclust_volcano_gp, device = "png", 
-       path = "figures", width = 6.5, height = 4, units = "in", dpi = 300,
+       path = "figures", width = 6.875, height = 4, units = "in", dpi = 300,
        bg="white")
@@
 ggsave("fk_nmdsplot.png", plot = fk_nmdsplot, device = "png",
-       path = "figures", width = 6.5, height = 4, units = "in", dpi = 300,
+       path = "figures", width = 6.875, height = 4, units = "in", dpi = 300,
        bg="white")
@@
 ggsave("pca_sla_cor_gp.png", pca_sla_cor_gp, device = "png", path = "supplement",
-       width = 6.5, height = 6.5, units = "in", dpi = 300)
+       width = 6.875, height = 6.5, units = "in", dpi = 300)
 
@@
 ## Supplemental figure 2
-
-```{r supp figure fraction of PC in metabolites}
-pc_comp_df <- targ_peaks %>%
-  filter(str_detect(filename, "180821_Smp")) %>%
-  left_join(clean_stans %>% select(compound_name, formula, polarity)) %>%
-  mutate(n_C=as.numeric(str_extract(formula, "(?<=C)\\d+"))) %>%
-  mutate(nM_C=n_C*nM) %>%
-  left_join(filled_file_metadata, by = join_by(filename)) %>%
-  filter(samp_type=="Smp") %>%
-  group_by(filename) %>%
-  summarize(total_nM=sum(nM_C), PC_um=unique(PC_um)) %>%
-  mutate(frac_in_metab=total_nM/1000/PC_um) %>%
-  left_join(filled_file_metadata, by = join_by(filename)) %>%
-  mutate(depth=str_extract(depth, "DCM|Deep|Surface")) %>%
-  mutate(depth=factor(depth, levels=c("Surface", "DCM", "Deep"))) %>%
-  mutate(station=factor(station, levels=c(as.character(4:14))))
-
-pc_frac_metab_gp <- ggplot(pc_comp_df, aes(x=station, y=frac_in_metab, color=sla_corr)) +
-  geom_hline(yintercept = 0) +
-  geom_boxplot(width = 0.9, linewidth=0.8) +
-  geom_point(color="black", size=2) +
-  scale_y_continuous(labels = scales::label_percent()) +
-  facet_grid(~depth, scales="free") +
-  scale_color_gradient2(
-    low = scales::muted("blue"), high = scales::muted("red"),  mid = "grey90",
-    aesthetics = c("color", "fill")
-  ) +
-  guides(color=guide_colorbar(direction = "horizontal", title.position = "top", 
-                              title.hjust = 0.5)) +
-  labs(color="Sea level\nanomaly (cm)", x="Station",
-       y="Fraction of PC represented by known metabolites") +
-  theme_bw() +
-  theme(legend.position = c(0.85, 0.8), legend.background = element_rect(color="grey90"))
-ggsave("pc_frac_metab_gp.png", pc_frac_metab_gp, device = "png", path = "supplement",
-       width = 6.5, height = 3.5, units = "in", dpi = 300)
-```
-
-![](supplement/pc_frac_metab_gp.png)
 
@@
 ggsave("fk_nmds_and_med_metab.png", plot = fk_nmds_and_med_metab, device = "png",
-       path = "supplement", width = 6.5, height = 5.5, units = "in", dpi = 300,
+       path = "supplement", width = 6.875, height = 5.5, units = "in", dpi = 300,
        bg="white")
@@
 ggsave("top10_stacked_area_station_gp.png", plot = top10_stacked_area_station_gp, 
-       device = "png", path = "supplement", width = 6.5, height = 9, 
+       device = "png", path = "supplement", width = 6.875, height = 9, 
        units = "in", dpi = 300)
@@
 ggsave("katherine_comp_mostcycenriched_gp.png", plot = katherine_comp_mostcycenriched_gp, 
-       device = "png", path = "supplement", width = 6.5, height = 5, units = "in", dpi = 300)
+       device = "png", path = "supplement", width = 6.875, height = 5, units = "in", dpi = 300)
 ```
@@
 ggsave("MC_stacked_mostdiff_gp.png", plot = MC_stacked_mostdiff_gp, device = "png", 
-       path = "supplement", width = 6.5, height = 3, units = "in", dpi = 300)
+       path = "supplement", width = 6.875, height = 3, units = "in", dpi = 300)
 ```

### 2023-12-08 manuscript/manuscript.Rmd
Finished implementing Anitra edits to results and intro

@@
 
-High frequency, long-term observations at Station ALOHA in the North Pacific Subtropical Gyre (NPSG) over the past 25 years have revealed temporal and spatial variability in what had prevously been considered a homogenous environment.[@Karl1999; @Karl2017; @Karl2021] A major source of this variability comes in the form of mesoscale eddies in which water is entrained into circular surface currents tens to hundreds of kilometers in diameter [@Karl2017; @McGillicuddy2016].
+High frequency, long-term observations at Station ALOHA in the North Pacific Subtropical Gyre (NPSG) over the past 25 years have revealed temporal and spatial variability in what had prevously been considered a homogenous environment [@Karl1999; @Karl2017; @Karl2021].
+A major source of this variability comes in the form of mesoscale eddies in which water is entrained into circular surface currents tens to hundreds of kilometers in diameter [@Karl2017; @McGillicuddy2016].
 Prior research has established the importance of these mesoscale features in introducing nutrients into the euphotic zone and altering the community composition [@Sweeney2003; @Barone2019].
 In particular, the depth and intensity of the deep chlorophyll maximum (DCM) at the top of the nitricline differs dramatically between cyclonic and anticyclonic eddies.
-@Barone2022 reported a shift in the microbial community at the DCM with eukaryotic phytoplankton enriched and prokaryotic phytoplankton depleted in cyclonic eddies.
-@Hawco2021 and @Dugenne2023 showed that some of this difference was due to alterations in dissolved iron chemistry and nitrogen fixation, especially at the surface.
-[Anitra: so the surface is driving changes at the DCM?]
+@Barone2022 reported a shift in the microbial community at the DCM with eukaryotic phytoplankton enriched and prokaryotic phytoplankton depleted in cyclonic eddies, and @Hawco2021 and @Dugenne2023 showed that some of this difference was due to alterations in dissolved iron chemistry and nitrogen fixation.
 
@@
 SLA trends are visible in the DCM and deep samples, with dark blue circles consistently discriminating from the dark red circles along the first multidimensional axis.
-The bottom row of plots show the direction and magnitude of this effect by plotting the grand mean of the normalized metabolite peak areas in the stations taken across the eddy dipole.*
+The bottom row of plots show the direction and magnitude of this effect by plotting the grand median of the z-score normalized metabolite peak areas in the stations taken across the eddy dipole with the upper whisker extent, median bar, and lower whisker extent corresponding to the values measured in three biological triplicates.*
 
@@
   ungroup()
-base_targ_gp <- avg_targ_concs %>%
+arrow_df <- data.frame(
+  cyc_type=c("Anticyclone\ncenter", "Cyclone\ncenter"),
+  x=c("6", "12"), xend=c("6", "12"),
+  y=c(5, 4), yend=c(9.5, 6.5),
+  color=c(scales::muted("red"), scales::muted("blue")),
+  depth=factor("Deep", levels=c("Surface", "DCM", "Deep"), 
+               labels=c("Surface  (15m)", "DCM  (110-120m)", "Deep  (175m)")),
+  lab_x=c("5", "10")
+)
+barplot_targ_gp <- avg_targ_concs %>%
   group_by(compound_name, untripl) %>%
-  summarise(nM=mean(nM), mean_cmpd_nm=unique(mean_cmpd_nm)) %>%
+  summarise(nM=median(nM), mean_cmpd_nm=unique(mean_cmpd_nm)) %>%
   ungroup() %>%
   arrange(desc(mean_cmpd_nm)) %>%
-  mutate(compound_name=ifelse(compound_name=="Dimethylsulfoniopropionate", "DMSP", compound_name)) %>%
+  mutate(compound_name=ifelse(compound_name=="Trimethylamine N-oxide", "TMAO", compound_name)) %>%
+  mutate(compound_name=ifelse(compound_name=="Hydroxyisoleucine", "HO-Ile", compound_name)) %>%
+  mutate(compound_name=ifelse(compound_name=="Glycine betaine", "GBT", compound_name)) %>%
+  mutate(compound_name=str_remove(compound_name, "^L-")) %>%
   mutate(compound_name=ifelse(
-    compound_name%in%head(unique(compound_name), 11), 
+    compound_name%in%head(unique(compound_name), 9), 
     compound_name, "Other")) %>%
@@
   geom_col(aes(x=station, y=mean_nM, fill=compound_name), width = 0.9) +
-  scale_fill_manual(values = c(hcl.colors(12)[c(1,3,5,7,9,11,2,4,6,8,10)], "grey50")) +
+  scale_fill_manual(values = c(splividis(9), "grey50")) +
   scale_y_continuous(expand = c(0.005, 0.1)) +
-  labs(x="Station", y="Mean nM metabolite concentration", fill=NULL) +
+  labs(x=NULL, y="Median nM\nmetabolite concentration", fill=NULL) +
   facet_grid(~depth, scales="free", space = "free") +
-  theme_bw() +
-  theme(legend.position = "bottom")
-
-sla_loess_points <- filled_file_metadata %>%
-  filter(samp_type=="Smp") %>%
-  filter(ms_run=="MT") %>%
-  distinct(station, sla_corr) %>%
-  mutate(station=factor(station, levels=as.character(4:14))) %>%
-  group_by(station) %>%
-  summarise(sla_corr=mean(sla_corr)) %>%
-  mutate(station=as.numeric(station)) %>%
-  loess(formula = sla_corr~station) %>%
-  predict(., newdata=data.frame(station=seq(1, 11, 0.01))) %>%
-  data.frame(station=seq(1, 11, 0.01), sla_corr=.)
-arrow_df <- data.frame(
-  cyc_type=c("Anticyclone\ncenter", "Cyclone\ncenter"),
-  x=c("6", "12"), xend=c("6", "12"),
-  y=c(5, 4), yend=c(7.5, 6.5),
-  color=c(scales::muted("red"), scales::muted("blue")),
-  depth=factor("Deep", levels=c("Surface", "DCM", "Deep"), 
-               labels=c("Surface  (15m)", "DCM  (110-120m)", "Deep  (175m)")),
-  lab_x=c("5", "11")
-)
-targ_gp_w_sla <- base_targ_gp +
+  guides(fill=guide_legend(nrow=2, byrow = TRUE)) +
   ggnewscale::new_scale_fill() +
-  geom_tile(aes(x=station, y=-0.5, fill=sla_corr),
-            color=NA, width=0.11, height=0.5, data = sla_loess_points) +
+  geom_segment(aes(x=x, y=y, xend=xend, yend=yend, color=color), data=arrow_df,
+               arrow=arrow(ends = "first", length = unit(0.1, "npc"), type = "closed"),
+               linewidth=0, linejoin = "bevel") +
+  geom_segment(aes(x=x, y=y+0.5, xend=xend, yend=yend, color=color), 
+               data=arrow_df, linewidth=2) +
+  geom_text(aes(x=lab_x, y=yend+0.5, color=color, label=cyc_type), 
+            data=arrow_df, vjust=0, hjust=0, lineheight=0.8) +
+  scale_x_discrete(breaks=as.character(seq(4, 14, 2))) +
   scale_fill_gradient2(
@@
   ) +
-  geom_segment(aes(x=x, y=y, xend=xend, yend=yend, color=color), data=arrow_df,
-               arrow=arrow(ends = "first", length = unit(0.1, "npc"), type = "closed"),
-               linewidth=0, linejoin = "bevel") +
-  geom_segment(aes(x=x, y=y+0.5, xend=xend, yend=yend, color=color), data=arrow_df, linewidth=2) +
-  geom_text(aes(x=lab_x, y=yend+0.5, color=color, label=cyc_type), 
-            data=arrow_df, vjust=0, hjust=0, lineheight=0.8) +
-  scale_color_identity()
-
-PC_data <- filled_file_metadata %>%
+  scale_color_identity() +
+  theme_bw() +
+  theme(legend.position = "top", axis.text.x = element_blank(), 
+        legend.margin = margin(t = 10,0,0,0), 
+        legend.box.margin = margin(-10,-10,-10,-10),
+        plot.margin = unit(numeric(4), "cm"),
+        strip.text = element_text(face = "bold"))
+
+frac_pc_in_metab_df <- targ_peaks %>%
+  filter(str_detect(filename, "180821_Smp")) %>%
+  left_join(clean_stans %>% select(compound_name, formula, polarity)) %>%
+  mutate(n_C=as.numeric(str_extract(formula, "(?<=C)\\d+"))) %>%
+  mutate(nM_C=n_C*nM) %>%
+  left_join(filled_file_metadata, by = join_by(filename)) %>%
   filter(samp_type=="Smp") %>%
-  filter(ms_run=="MT") %>%
-  distinct(station, PC_um, depth) %>%
-  mutate(depth=factor(depth, levels=c("Surface", "DCM", "Deep"), 
-                      labels=c("Surface  (15m)", "DCM  (110-120m)", "Deep  (175m)"))) %>%
-  mutate(station=factor(station, levels=as.character(4:14)))
-targ_gp_w_sla_PC <- targ_gp_w_sla +
-  geom_point(aes(x=station, y=PC_um*5), data=PC_data, pch=21, fill="white", color="black") +
-  scale_y_continuous(sec.axis = sec_axis(~./5, name = "Particulate carbon (o, µM)"),
-                     expand = c(0.005, 0.1))
-
-ggsave("targ_gp_w_sla_PC.png", plot = targ_gp_w_sla_PC, device = "png", 
-       path = "figures", width = 6.5, height = 4.5, units = "in", dpi = 300)
-```
-
-![](figures/targ_gp_w_sla_PC.png)
-
-*Figure 2: Concentrations of known compounds across the eddy dipole, broken down by depth.
-Stacked bars correspond to known compounds, with the top 11 shown and all other metabolites summed in grey (n=[TBD]).
-DMSP = dimethylsulfoniopropionate.
-Sea level anomaly (SLA) is shown by the red-blue color scale underneath the barplot with red corresponding to a positive SLA (anticyclone) and blue corresponding to negative SLA (cyclone).
-The eddy centers (stations 6 and 12) have been labeled with arrows of the same color scheme in the final panel.
-Particulate carbon concentrations in µmol/L are shown as white dots and use the y-axis scale on the right side (200x the metabolite scale) [Anitra: This implies that the metabolites are 1/200 of the POC.
-But you have plotted compound concentration not carbon for metabolites.
-Assuming each compound has an average of 4 carbon atoms, you would have to multiply by 4 so that is 1/50 (2%) or so of POC.
-I think you need to make this clear somehow.
-Or you could make an additional plot with C for each compound.
-I see supp fig 2 has the %POC so maybe that is enough but maybe call out that figure here.][Will: I agree, let's brainstorm].*
+  group_by(filename) %>%
+  summarize(total_nM=sum(nM_C), PC_um=unique(PC_um)) %>%
+  mutate(frac_in_metab=total_nM/1000/PC_um) %>%
+  left_join(filled_file_metadata, by = join_by(filename, PC_um)) %>%
+  mutate(depth=str_extract(depth, "DCM|Deep|Surface")) %>%
+  mutate(depth=factor(depth, levels=c("Surface", "DCM", "Deep"))) %>%
+  mutate(station=factor(station, levels=c(as.character(4:14))))
+pc_frac_metab_gp <- frac_pc_in_metab_df %>%
+  ggplot(aes(x=station, y=frac_in_metab, color=sla_corr)) +
+  geom_hline(yintercept = 0) +
+  geom_boxplot(width = 0.9, linewidth=0.8) +
+  scale_x_discrete(breaks=as.character(seq(4, 14, 2))) +
+  scale_y_continuous(labels = scales::label_percent()) +
+  facet_grid(~depth, scales="free") +
+  scale_color_gradient2(
+    low = scales::muted("blue"), high = scales::muted("red"),  mid = "grey90",
+    aesthetics = c("color", "fill")
+  ) +
+  guides(color=guide_colorbar(direction = "horizontal", title.position = "top", 
+                              title.hjust = 0.5, barheight = 0.2, reverse = TRUE)) +
+  labs(color="Corr. SLA (cm)", x="Station",
+       y="Fraction of PC represented\nby known metabolites") +
+  theme_bw() +
+  theme(legend.position = c(0.84, 0.75), legend.background = element_rect(color="grey90"),
+        strip.background = element_blank(), strip.text = element_blank(),
+        plot.margin = margin(0,1,0,0))
+
+targ_gp_w_sla_frac <- egg::ggarrange(barplot_targ_gp, pc_frac_metab_gp, draw = FALSE)
+
+no_outlier_frac_pc_in_metab_df <- frac_pc_in_metab_df %>%
+  filter(!(station=="11" & tripl=="A" & depth=="Surface")) %>%
+  filter(!(station=="6" & tripl=="B" & depth=="Surface")) %>%
+  filter(!(station=="7" & tripl=="C" & depth=="Surface")) %>%
+  filter(!(station=="11" & tripl=="B" & depth=="DCM"))
+pc_totalnm_plot <- no_outlier_frac_pc_in_metab_df %>%
+  ggplot() +
+  aes(x=PC_um, y=total_nM) +
+  geom_smooth(method="lm", color="black") +
+  geom_point(aes(fill=sla_corr), pch=21, color="black") +
+  facet_wrap(~depth, ncol=1, scales="free") +
+  labs(x="Particulate\ncarbon (uM)", y="Total nM carbon in known metabolites") +
+  scale_fill_gradient2(
+    low = scales::muted("blue"), high = scales::muted("red"),  mid = "grey90",
+    aesthetics = c("fill"), guide = "none"
+  ) +
+  theme_bw() +
+  theme(strip.text = element_text(face = "bold"))
+
+targ_gp_w_sla_pc <- cowplot::plot_grid(
+  targ_gp_w_sla_frac, pc_totalnm_plot, rel_widths = c(3, 1), labels = "AUTO"
+) +
+  theme(plot.background = element_rect(fill="white", color=NA))
+
+plot(targ_gp_w_sla_pc)
+
+ggsave("targ_gp_w_sla_pc.png", plot = targ_gp_w_sla_pc, device = "png", 
+       path = "figures", width = 6.875, height = 4.5, units = "in", dpi = 300)
+```
+
+![](figures/targ_gp_w_sla_pc.png)
+
+*Figure 2: A) Upper row of stacked barplots shows known compounds measured across the eddy dipole and separated by depth.
+Bar height corresponds to median triplicate concentration for each metabolite, with the top 9 shown and the 44 other identified metabolites summed in grey.
+TMAO = trimethylamine N-oxide, HO-Ile = hydroxyisoleucine, GBT = glycine betaine, DCM = deep chlorophyll maximum.
+Lower row of boxplots shows the fraction of total particulate carbon (PC) in the known metabolites across the eddy dipole, with the upper whisker extent, median bar, and lower whisker extent corresponding to the values measured in three biological triplicates.
+Colors correspond to corrected sea level anomaly (Corr.
+SLA) in centimeters, with dark red indicating anticyclonic (positive SLA) and dark blue indicating cyclonic (negative SLA) eddy state.
+B) Type I linear regressions of particulate carbon against total nM carbon in known metabolites.
+Points have been colored according to the SLA scheme in A) and have been placed on independent axes.
+Outlier points have been removed from the surface at stations 6, 10, and 11 as well as a DCM sample from station 12.
+For the full regression, see Supplemental Figure XX.*
 
 ```{r avg targ conc values and numbers for text, eval=FALSE}
-avg_targ_concs %>%
-  group_by(filename) %>%
-  summarise(nM=sum(nM)) %>%
-  left_join(filled_file_metadata %>% distinct(filename, untripl, PC_um, depth)) %>%
-  group_by(untripl, PC_um, depth) %>%
-  summarise(nM=mean(nM), .groups = "drop") %>%
-  group_by(depth) %>%
-  summarise(PC_vs_nM=broom::tidy(cor.test(PC_um, nM)))
-  # summarise(PC_vs_nM=broom::tidy(cor.test(PC_um, nM, method = "spearman")))
-  # ggplot(aes(x=PC_um, y=nM, color=depth)) + geom_point() + geom_smooth(method="lm")
 avg_targ_concs %>%
@@
 We found that the overall concentration of known molecules was highest at the DCM of the cyclone, in line with our expectations and prior results.
-The most abundant quantified molecule overall was trimethylamine N-oxide (TMAO), with an average concentration of 1.38 nM.
-At the surface hydroxyisoleucine was the most abundant metabolite at 1.53 nM due to its monotonic decrease with depth as opposed to a majority of the osmolytes (37/53) which reached maximum concentrations at the DCM instead (Figure 2).
-
-[Anitra: comments on the paragraph above: Maybe instead of this somewhat convoluted description you could just break it down by depth and say which compound was most abundant at each depth?
-Or which compound was most abundant in the cyclone and anticyclone.
-And then in a separate sentence just simply state that most compounds reached a ma at the DCM, but hydroxyisoleucine was an exception.
-Also, are you implying that hydroxyisoleucine is an osmolyte?
-Is that true?
-Also, what do you mean by "average"?]
-
-Bulk particulate organic carbon (POC) measurements correlated well with overall [Anitra: what do you mean by overall here?] metabolite concentrations ($r= .554$, p-value < 0.001) but this was largely driven by both parameters having higher values in the surface than at depth.
-When analyzed at each depth individually, the DCM ($r_{DCM}= .790$, p-value = 0.003) and 175 meter samples ($r_{Deep}= .554$, p-value = 0.077) had weaker but still good correlations, but no such association was detected at the surface ($r_{Surface}= -.125$, p-value = 0.714).
-
-```{r PC vs metab abund anovas and perm test, eval=FALSE}
-pc_comp_df <- targ_peaks %>%
-  filter(str_detect(filename, "180821_Smp")) %>%
-  left_join(clean_stans %>% select(compound_name, formula, polarity)) %>%
-  mutate(n_C=as.numeric(str_extract(formula, "(?<=C)\\d+"))) %>%
-  mutate(nM_C=n_C*nM) %>%
-  left_join(filled_file_metadata, by = join_by(filename)) %>%
-  filter(samp_type=="Smp") %>%
-  group_by(filename) %>%
-  summarize(total_nM=sum(nM_C), PC_um=unique(PC_um)) %>%
-  mutate(frac_in_metab=total_nM/1000/PC_um) %>%
-  left_join(filled_file_metadata, by = join_by(filename)) %>%
-  mutate(depth=str_extract(depth, "DCM|Deep|Surface")) %>%
-  mutate(depth=factor(depth, levels=c("Surface", "DCM", "Deep"))) %>%
-  mutate(station=factor(station, levels=c(as.character(4:14))))
-
-# Significance testing between depths and SLA classes for % PC in metab
-pc_comp_df %>% aov(formula=frac_in_metab~depth) %>% TukeyHSD()
-pc_comp_df %>% aov(formula=frac_in_metab~sla_class) %>% TukeyHSD()
-
-# Significance testing for differences in % in metabs as a function of SLA by depth
-pc_comp_df %>%
-  nest(data=-depth) %>%
-  mutate(aov_out=map(data, ~broom::tidy(aov(.x$frac_in_metab~.x$sla_class)))) %>%
-  mutate(kw_out=map(data, ~broom::tidy(kruskal.test(.x$frac_in_metab~.x$sla_class)))) %>%
-  pivot_longer(c(aov_out, kw_out)) %>%
-  unnest(value) %>%
-  drop_na(p.value) %>%
-  select(depth, name, p.value) %>%
-  group_by(name) %>%
-  mutate(p.value=p.adjust(p.value, method="fdr")) %>%
-  ungroup() %>%
-  arrange(name, depth)
-pc_comp_df %>%
-  ggplot(aes(x=sla_class, y=frac_in_metab)) +
-  geom_boxplot() +
-  facet_wrap(~depth, nrow=1)
-
-# Permutation testing for effect size test on Deep SLA differences
-set.seed(123)
-map_dbl(seq(1, 2, 0.1), function(n_mult){
-  map_dbl(1:100, function(h){
-    pc_comp_df %>%
-      filter(depth=="Deep") %>%
-      group_by(sla_class) %>%
-      summarize(mean=mean(frac_in_metab), sd=sd(frac_in_metab), n=n()) %>%
-      mutate(sim_vals=map(1:3, function(i){rnorm(.$n[i]*n_mult, .$mean[i], .$sd[i])})) %>%
-      unnest_longer(sim_vals) %>%
-      aov(formula=sim_vals~sla_class) %>%
-      broom::tidy() %>%
-      filter(term=="sla_class") %>%
-      pull(p.value)
-  }) %>% mean()
-}, .progress = TRUE)
-```
-
-This correlation between PC [Anitra: Is there ATP data?
-Might be interesting to compare.
-Will: instinct says no, I think that ATP was really only collected during the diel phase and from a few spots during lagrangian.
-TODO: Double check the dashboard] and overall metabolite abundance implied that the fraction of PC represented by known metabolites was likely to be fixed across the eddy transect and this expectation was validated by followup analyses.
-[Anitra: I do not see this at all.
-I see spikes in metabolite concentration that are not matched with spikes in PC.
-Mostly at the eddy centers.
-Do you not think this is suggestive of a change in the contribution of biomass to PC in the eddy centers?
-I would assume if you had biomass instead of PC you would be able to see biomass increase at the centers where metabolites increase.
-TODO: Check the same values using beam attenuation (interpolate if necessary)] Known metabolites contributed between approximately 1.5% and 5% of the carbon to the total PC load, but this variance was not clearly due to either depth or SLA (Supplemental figure 2).
-The only difference we observed was a slightly reduced metabolite contribution in the 175 meter samples (0.6 percentage points lower at 175 meters, p-value=0.002) while no difference was detected between the surface and DCM samples or between the cyclone, anticyclone, and outside the eddy samples (all p-values > 0.1).
-A higher proportion of PC in living biomass would translate to a higher proportion of metabolites in POC.
-This lack of difference may have been due to the small sample sizes (6 anticyclonic stations, 3 cyclonic, and 2 neither, each sampled in triplicate).
-A followup permutational sensitivity analysis indicated that a difference in means of 0.4 percentage points would have been obtained at the α < 0.05 level for SLA in the 175 meter samples if ~1.5x as many samples had been collected with mean and variances remaining similar.
-
-[TODO: switch to POC instead of PC after confirming it's only the organics]
-
-  - Discussion: My % PC in metabs lower than Angie's? Because her PC values were way higher but metab concentrations were about the same?
+The most abundant quantified molecule overall was trimethylamine N-oxide (TMAO), with an average concentration of 1.38 nM and a distinct DCM maximum (Figure 2).
+This pattern of a subsurface maximum at the DCM was mirrored by a majority of the known molecules (37/53).
+Other abundant molecules decreased monotonically with depth, with highest values at the surface (Figure 2).
+
+```{r PC_vs_totnM regression values, eval=FALSE}
+no_outlier_frac_pc_in_metab_df %>%
+  ggplot(aes(x=PC_um, y=total_nM)) + 
+  geom_smooth(method="lm", color="grey40") +
+  geom_point(aes(color=depth)) + 
+  geom_smooth(aes(color=depth), method="lm") +
+  theme_bw()
+
+# Overall trend between PC and total nM metabolites (all depths)
+broom::tidy(lm(total_nM~PC_um, data=no_outlier_frac_pc_in_metab_df))[2,]
+with(no_outlier_frac_pc_in_metab_df, cor(PC_um, total_nM))
+lme4::lmer(total_nM~PC_um + (1|depth), data=no_outlier_frac_pc_in_metab_df)
+
+# By-depth trend between PC and total nM metabolites
+no_outlier_frac_pc_in_metab_df %>%
+  group_by(depth) %>%
+  summarise(overall=broom::tidy(lm(total_nM~PC_um))[2,], 
+            cor_vals=cor(PC_um, total_nM))
+```
+
+As expected, bulk particulate carbon (PC) measurements showed a strong trend with summed metabolite concentrations ($\beta = 23.7$, Pearson's $r$ = .876, p-value < 0.001).
+This overall trend was largely driven by both higher values in the surface than at depth but when analyzed at each depth individually the general correspondence held.
+The DCM had the strongest trend ($\beta$ = 22.1, $r= .689$, p-value < 0.001) while Surface ($\beta$ = 16.0, $r = .375$, p-value = 0.04) and Deep ($\beta$ = 23.9, $r = 0.503$, p-value = 0.003) values were weaker but still compelling relationships.
+However, this association required the removal of three samples in the surface and one in the DCM with anomalously high metabolite loads.
+The full regression is reported in Supplemental Figure XX.
+
+```{r PC vs total_nM ANOVA by depth, eval=FALSE}
+# More carbon in metabolites at DCM when outliers removed?
+no_outlier_frac_pc_in_metab_df %>%
+  mutate(pred_total_nM=predict(lm(formula=total_nM~PC_um, data=.), newdata=.)) %>%
+  select(filename, depth, station, sla_class, sla_corr, total_nM, PC_um, pred_total_nM) %>%
+  mutate(total_nM_resid=total_nM-pred_total_nM) %>%
+  aov(formula=total_nM_resid~depth) %>% TukeyHSD() %>% broom::tidy()
+  # ggplot() + geom_boxplot(aes(x=depth, y=total_nM_resid)) + aes(color=sla_class)
+
+# More carbon in metabolites at DCM when outliers handled with rank testing?
+frac_pc_in_metab_df %>%
+  mutate(pred_total_nM=predict(lm(formula=total_nM~PC_um, data=.), newdata=.)) %>%
+  select(filename, depth, total_nM, PC_um, pred_total_nM) %>%
+  mutate(total_nM_resid=rank(total_nM-pred_total_nM)) %>%
+  kruskal.test(formula=total_nM_resid~depth) %>% broom::tidy()
+  # ggplot() + geom_boxplot(aes(x=depth, y=total_nM_resid))
+
+# More carbon in metabolites at DCM when measured by % in metabs too
+no_outlier_frac_pc_in_metab_df %>%
+  aov(formula=frac_in_metab~depth) %>% TukeyHSD() %>% broom::tidy()
+  # ggplot() + geom_boxplot(aes(x=depth, y=frac_in_metab))
+no_outlier_frac_pc_in_metab_df %>%
+  group_by(depth) %>%
+  summarise(median(frac_in_metab))
+frac_pc_in_metab_df %>%
+  kruskal.test(formula=frac_in_metab~depth, data=.) %>% broom::tidy()
+  # ggplot() + geom_boxplot(aes(x=depth, y=rank(frac_in_metab)))
+```
+
+Known metabolites contributed between 1.5% and 5% of the carbon to the total PC load with higher values at the DCM than the surface or 175m samples (Figure 2, Supplemental Figure XX).
+DCM samples contained approximately 25% (0.5 percentage points) more carbon in the known metabolite pool (One-way ANOVA with outliers removed plus Tukey post-hoc, p<0.001, n=30-33) but there was no clear trend with SLA.
 
@@
   slice(-1) %>%
-  select(depth=`Depth.(m)`, PC_um=`C.value.(µmol/L)`) %>%
-  filter(depth<30) %>%
+  select(abs_depth=`Depth.(m)`, PC_um=`C.value.(µmol/L)`) %>%
+  # filter(depth<30) %>%
   mutate(cruise="HOE-Legacy")
@@
 my_pcdata <- filled_file_metadata %>%
-  filter(depth=="Surface") %>%
-  distinct(depth, PC_um) %>%
+  # filter(depth=="Surface") %>%
+  distinct(abs_depth, PC_um) %>%
   mutate(cruise="MESOSCOPE")
@@
   ggplot() +
-  geom_boxplot(aes(x=cruise, y=PC_um)) +
-  scale_y_continuous(limits = c(0, NA))
+  geom_point(aes(x=PC_um, y=-abs_depth, color=cruise))
 ```
@@
     TRUE~compound_name
-  ))
+  )) %>%
+  mutate(rtmed=round(rtmed, 1)) %>%
+  mutate(compound_name=ifelse(is.na(compound_name), "", compound_name)) %>%
+  mutate(feature=ifelse(is.na(feature), "", feature)) %>%
+  dplyr::rename(`Median m/z`=mzmed, `Median RT`=rtmed, `Putative formula`=formula,
+                `Compound name`=compound_name, `Feature ID`=feature)
 knitr::kable(ms11_df)
 ```
-
-[TODO: Clean up this table. Make sure the RT values only 2-3 sigfigs, make column names legible, etc.]
 
@@
 
-  - Discussion: I still really like the idea of modeling things according to the slope of their change across the eddy dipole.
-Light-sensitive or pressure-sensitive (e.g.
-sunscreen molecules, piezolytes) metabolites wouldn't change at all, DCM-specific metabolites (e.g.
-chlorophyll) would change by 20m or so as the DCM does, nutrient-sensitive metabolites (guanine?
-glutamate/glutamine?) would track with the isonutrials, and the salinity or temperature sensitive metabolites (aka pure osmolytes) would track with isopycnals.
-Do I have enough data to show that here?
-I forget what my early attempts to quantify these gradients looked like.
-
 To characterize the observed differences in multivariate space we used k-means clustering as an unsupervised way to identify dominant trends within the data (Figure 4).
@@
 # Supplement
-## Supplemental figure 1
+## Supp fig: PCA axes 1-3 vs SLA by depth
 
@@
 
-## Supplemental figure 2
-
-*Supplemental figure 2: Fraction of particulate carbon represented by known metabolites.
-Boxplots set behind points are colored according to sea level anomaly (SLA) with red corresponding to a positive SLA (anticyclone) and blue corresponding to negative SLA (cyclone).*
+## Supp fig: Stacked barplot of top metabs by triplicate
+
+```{r targ barplot relative and absolute}
+targ_barplot_df <- targ_peaks %>%
+  left_join(filled_file_metadata, by="filename") %>%
+  filter(ms_run=="MT") %>%
+  filter(samp_type=="Smp") %>%
+  group_by(compound_name) %>%
+  mutate(mean_cmpd_nm=mean(nM)) %>%
+  ungroup() %>%
+  arrange(desc(mean_cmpd_nm)) %>%
+  mutate(compound_name=ifelse(compound_name=="Trimethylamine N-oxide", "TMAO", compound_name)) %>%
+  mutate(compound_name=ifelse(compound_name=="Hydroxyisoleucine", "HO-Ile", compound_name)) %>%
+  mutate(compound_name=ifelse(compound_name=="Glycine betaine", "GBT", compound_name)) %>%
+  mutate(compound_name=str_remove(compound_name, "^L-")) %>%
+  mutate(compound_name=ifelse(
+    compound_name%in%head(unique(compound_name), 9), 
+    compound_name, "Other")) %>%
+  mutate(compound_name=fct_inorder(compound_name)) %>%
+  group_by(depth, station, compound_name, tripl) %>%
+  summarise(nM=sum(nM), .groups = "keep") %>%
+  mutate(depth=factor(depth, levels=c("Surface", "DCM", "Deep"), 
+                      labels=c("Surface  (15m)", "DCM  (110-120m)", "Deep  (175m)"))) %>%
+  mutate(station=factor(station, levels=as.character(4:14)))
+base_tripl_barplot <- targ_barplot_df %>%
+  ggplot() +
+  aes(x=tripl, y=nM, fill=compound_name) +
+  geom_hline(yintercept = 0) +
+  facet_grid(depth~station, scales="free_y", switch = "x") +
+  scale_fill_manual(values = c(splividis(9), "grey50")) +
+  labs(x="Station", y="Metabolite concentration (nM)", fill=NULL) +
+  theme_bw() +
+  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(),
+        strip.background.x = element_blank(), 
+        panel.border = element_blank())
+abs_targ_bar_gp <- base_tripl_barplot +
+  geom_col(width=1) +
+  ggtitle("A) Absolute concentrations of most abundant metabolites") +
+  scale_y_continuous(expand = expansion()) +
+  theme(legend.position="none", strip.text.x = element_blank(), axis.title.x = element_blank())
+rel_targ_bar_gp <- base_tripl_barplot +
+  geom_col(width=1, position = "fill") +
+  scale_y_continuous(labels = scales::label_percent(), expand = expansion()) +
+  theme(legend.position = "bottom") +
+  guides(fill=guide_legend(nrow=2, byrow = TRUE)) +
+  ggtitle("B) Relative composition of the metabolome")
+egg::ggarrange(abs_targ_bar_gp, rel_targ_bar_gp, ncol=1)
+```
+
+## Supp fig: Regressions of PC vs total nM metabolites with outliers
+
+```{r}
+comb_pc_reg_df <- bind_rows(
+  frac_pc_in_metab_df %>% mutate(status="All files"),
+  no_outlier_frac_pc_in_metab_df %>%
+    mutate(status="Dropped outliers")
+)
+
+PC_comb_gp <- gridExtra::arrangeGrob(
+  comb_pc_reg_df %>%
+    ggplot(aes(x=PC_um, y=total_nM)) +
+    geom_point(aes(fill=depth), pch=21) +
+    geom_smooth(method="lm", color="black") +
+    geom_smooth(aes(color=depth), method="lm") +
+    facet_wrap(~status, ncol=1, scales="free_y") +
+    theme_bw() +
+    theme(legend.position = "top"),
+  comb_pc_reg_df %>%
+    ggplot() +
+    aes(x=PC_um, y=total_nM) +
+    geom_smooth(method="lm", color="black") +
+    geom_point(aes(fill=sla_corr), pch=21, color="black", size=3) +
+    ggh4x::facet_grid2(depth~status, scales="free", independent="all") +
+    labs(x="Particulate carbon (uM)", 
+         y="Total nM carbon in known metabolites",
+         fill="Corr.\nSLA\n(cm)") +
+    scale_fill_gradient2(
+      low = scales::muted("blue"), high = scales::muted("red"),  mid = "grey90"
+    ) +
+    theme_bw() +
+    theme(legend.position = "top"),
+  nrow = 1
+)
+plot(PC_comb_gp)
+comb_pc_reg_df %>%
+  nest(data=-c(status, depth)) %>%
+  mutate(reg_vals=map(data, ~broom::tidy(lm(.x$total_nM~.x$PC_um)))) %>%
+  unnest(reg_vals) %>%
+  filter(term!="(Intercept)")
+
+# xmdl <- lm(total_nM~PC_um, data=all_targ_pc_reg_df)
+# plot(fitted(xmdl), residuals(xmdl))
+# dfbeta(xmdl) %>% 
+#   as.data.frame() %>%
+#   cbind(all_targ_pc_reg_df$filename)
+```
+
+*Supplemental figure 2: [TBD something about triplicates]*
 

### 2023-12-14 metadata/metadata_control.Rmd
Adding uncorrected PC values to metadata

@@
 interp_df <- interp_df %>%
+  mutate(PC_uncor=PC_um) %>%
   mutate(pred_PC=interp_df %>% 

### 2024-03-07 manuscript/manuscript.Rmd
Pre-Anitra edits v2

@@
 High frequency, long-term observations at Station ALOHA in the North Pacific Subtropical Gyre (NPSG) over the past 25 years have revealed temporal and spatial variability in what had prevously been considered a homogenous environment [@Karl1999; @Karl2017; @Karl2021].
-A major source of this variability comes in the form of mesoscale eddies in which water is entrained into circular surface currents tens to hundreds of kilometers in diameter [@Karl2017; @McGillicuddy2016].
+A major source of variability comes in the form of mesoscale eddies in which water is entrained into circular surface currents tens to hundreds of kilometers in diameter [@Karl2017; @McGillicuddy2016].
 Prior research has established the importance of these mesoscale features in introducing nutrients into the euphotic zone and altering the community composition [@Sweeney2003; @Barone2019].
@@
         plot.margin = unit(numeric(4), "cm"),
-        strip.text = element_text(face = "bold"))
+        strip.text = element_text(face = "bold"),
+        plot.background = element_rect(fill="white"))
 
@@
         strip.background = element_blank(), strip.text = element_blank(),
-        plot.margin = margin(0,1,0,0))
+        plot.margin = margin(0,1,0,0), plot.background = element_rect(fill="white"))
 
@@
 
-no_outlier_frac_pc_in_metab_df <- frac_pc_in_metab_df %>%
-  filter(!(station=="11" & tripl=="A" & depth=="Surface")) %>%
-  filter(!(station=="6" & tripl=="B" & depth=="Surface")) %>%
-  filter(!(station=="7" & tripl=="C" & depth=="Surface")) %>%
-  filter(!(station=="11" & tripl=="B" & depth=="DCM"))
-pc_totalnm_plot <- no_outlier_frac_pc_in_metab_df %>%
-  ggplot() +
-  aes(x=PC_um, y=total_nM) +
-  geom_smooth(method="lm", color="black") +
-  geom_point(aes(fill=sla_corr), pch=21, color="black") +
-  facet_wrap(~depth, ncol=1, scales="free") +
-  labs(x="Particulate\ncarbon (uM)", y="Total nM carbon in known metabolites") +
-  scale_fill_gradient2(
-    low = scales::muted("blue"), high = scales::muted("red"),  mid = "grey90",
-    aesthetics = c("fill"), guide = "none"
-  ) +
-  theme_bw() +
-  theme(strip.text = element_text(face = "bold"))
-
-targ_gp_w_sla_pc <- cowplot::plot_grid(
-  targ_gp_w_sla_frac, pc_totalnm_plot, rel_widths = c(3, 1), labels = "AUTO"
-) +
-  theme(plot.background = element_rect(fill="white", color=NA))
-
-plot(targ_gp_w_sla_pc)
-
-ggsave("targ_gp_w_sla_pc.png", plot = targ_gp_w_sla_pc, device = "png", 
-       path = "figures", width = 6.875, height = 4.5, units = "in", dpi = 300)
-```
-
-![](figures/targ_gp_w_sla_pc.png)
-
-*Figure 2: A) Upper row of stacked barplots shows known compounds measured across the eddy dipole and separated by depth.
+plot(targ_gp_w_sla_frac)
+
+ggsave("targ_gp_w_sla_frac.png", plot = targ_gp_w_sla_frac, device = "png", 
+       path = "figures", width = 6.875, height = 4.5, units = "in", dpi = 300, 
+       bg = "white")
+```
+
+![](figures/targ_gp_w_sla_frac.png)
+
+*Figure 2: Upper row of stacked barplots shows known compounds measured across the eddy dipole and separated by depth.
 Bar height corresponds to median triplicate concentration for each metabolite, with the top 9 shown and the 44 other identified metabolites summed in grey.
@@
 Colors correspond to corrected sea level anomaly (Corr.
-SLA) in centimeters, with dark red indicating anticyclonic (positive SLA) and dark blue indicating cyclonic (negative SLA) eddy state.
-B) Type I linear regressions of particulate carbon against total nM carbon in known metabolites.
-Points have been colored according to the SLA scheme in A) and have been placed on independent axes.
-Outlier points have been removed from the surface at stations 6, 10, and 11 as well as a DCM sample from station 12.
-For the full regression, see Supplemental Figure XX.*
+SLA) in centimeters, with dark red indicating anticyclonic (positive SLA) and dark blue indicating cyclonic (negative SLA) eddy state.*
 
@@
 This pattern of a subsurface maximum at the DCM was mirrored by a majority of the known molecules (37/53).
-Other abundant molecules decreased monotonically with depth, with highest values at the surface (Figure 2).
+Most other abundant molecules decreased monotonically with depth, with highest values at the surface (Figure 2).
 
 ```{r PC_vs_totnM regression values, eval=FALSE}
-no_outlier_frac_pc_in_metab_df %>%
-  ggplot(aes(x=PC_um, y=total_nM)) + 
-  geom_smooth(method="lm", color="grey40") +
-  geom_point(aes(color=depth)) + 
-  geom_smooth(aes(color=depth), method="lm") +
-  theme_bw()
+no_outlier_frac_pc_in_metab_df <- frac_pc_in_metab_df %>%
+  filter(!(station=="11" & tripl=="A" & depth=="Surface")) %>%
+  filter(!(station=="6" & tripl=="B" & depth=="Surface")) %>%
+  filter(!(station=="7" & tripl=="C" & depth=="Surface")) %>%
+  filter(!(station=="11" & tripl=="B" & depth=="DCM"))
 
@@
 
-As expected, bulk particulate carbon (PC) measurements showed a strong trend with summed metabolite concentrations ($\beta = 23.7$, Pearson's $r$ = .876, p-value < 0.001).
-This overall trend was largely driven by both higher values in the surface than at depth but when analyzed at each depth individually the general correspondence held.
-The DCM had the strongest trend ($\beta$ = 22.1, $r= .689$, p-value < 0.001) while Surface ($\beta$ = 16.0, $r = .375$, p-value = 0.04) and Deep ($\beta$ = 23.9, $r = 0.503$, p-value = 0.003) values were weaker but still compelling relationships.
-However, this association required the removal of three samples in the surface and one in the DCM with anomalously high metabolite loads.
-The full regression is reported in Supplemental Figure XX.
+The small-scale oceanographic variability introduced by mesoscale eddy features allowed us to also test the robustness of a model expecting a linear relationship between the total carbon contained within metabolites and bulk particulate carbon (PC) measurements.
+As expected, PC values showed a strong trend with summed metabolite concentrations ($\beta = 23.7\pm1.35$ (SE), Pearson's $r$ = .876, p-value < 0.001, Figure 2.5).
+This overall trend was largely driven by higher values in the surface than at depth, but when analyzed at each depth individually the general correspondence held.
+The DCM had the strongest trend ($\beta = 22.1\pm2.24$, $r= .689$, p-value < 0.001) while Surface ($\beta = 16.0\pm7.45$, $r = .375$, p-value = 0.04) and Deep ($\beta = 23.9\pm7.36$, $r = 0.503$, p-value = 0.003) values were weaker but still compelling relationships (Figure 2.5).
+However, this association required the removal of three samples in the surface and one in the DCM with anomalously high metabolite loads as well as two surface PC samples with anomalously high values from Stations 10 and 11 for which PC was interpolated by linear regression against beam attenuation.
+The full regression is reported in Supplemental Figure 3.
+
+```{r Fig 2.5 PC vs tot nM plot}
+pc_totalnm_plot <- no_outlier_frac_pc_in_metab_df %>%
+  ggplot(aes(x=PC_um, y=total_nM)) +
+  geom_point(aes(fill=depth), pch=21, size=3) +
+  geom_smooth(method="lm", color="white", lwd=2.5, se=FALSE) +
+  geom_smooth(method="lm", color="black", lwd=1.5) +
+  geom_smooth(aes(group=depth), method="lm", color="white", lwd=2.5, se = FALSE) +
+  geom_smooth(aes(color=depth), method="lm", lwd=1.5) +
+  theme_bw() +
+  theme(legend.position = "top") +
+  labs(color="Sampling depth",
+       fill="Sampling depth",
+       x="Particulate carbon (uM)", 
+       y="Total nM carbon in known metabolites") +
+  scale_color_manual(breaks=c("Surface", "DCM", "Deep"), 
+                     values = c("#F4BB23", "#028E34", "#0B505C"), aesthetics = c("fill", "color"))
+pc_totalnm_plot_bydepth <- no_outlier_frac_pc_in_metab_df %>%
+  ggplot() +
+  aes(x=PC_um, y=total_nM) +
+  geom_smooth(method="lm", color="black") +
+  geom_point(aes(fill=sla_corr), pch=21, color="black") +
+  facet_wrap(~depth, ncol=1, scales="free") +
+  labs(x="Particulate\ncarbon (uM)", y="Total nM carbon in known metabolites") +
+  scale_fill_gradient2(
+    low = scales::muted("blue"), high = scales::muted("red"),  mid = "grey90",
+    aesthetics = c("fill"), name="Corr.\nSLA\n(cm)"
+  ) +
+  theme_bw() +
+  theme(strip.text = element_text(face = "bold"))
+
+comb_pcnm_plot <- cowplot::plot_grid(pc_totalnm_plot, pc_totalnm_plot_bydepth, 
+                                     rel_widths = c(5, 3))
+
+ggsave("comb_pcnm_plot.png", plot = comb_pcnm_plot, device = "png", 
+       path = "figures", width = 6.875, height = 4.5, units = "in", dpi = 300, 
+       bg = "white")
+```
+
+![](figures/comb_pcnm_plot.png)
+
+*Figure 2.5: Type I linear regressions of particulate carbon against total nM carbon in known metabolites both as a whole and divided by depth.
+Points have been colored according to the SLA and have been placed on independent axes.
+Outlier metabolite points have been removed from the surface at stations 6, 10, and 11 as well as a DCM sample from station 12, while two outlier PC points have been interpolated from beam attenuation for surface values at stations 10 and 11.
+For the full regression, see Supplemental Figure 3.*
 
@@
 
-Known metabolites contributed between 1.5% and 5% of the carbon to the total PC load with higher values at the DCM than the surface or 175m samples (Figure 2, Supplemental Figure XX).
+Known metabolites contributed between 1.5% and 5% of the carbon to the total PC load with higher values at the DCM than the surface or 175m samples (Figure 2, Supplemental Figure 3).
 DCM samples contained approximately 25% (0.5 percentage points) more carbon in the known metabolite pool (One-way ANOVA with outliers removed plus Tukey post-hoc, p<0.001, n=30-33) but there was no clear trend with SLA.
@@
 Heuristic exploration of potentially related masses identified a total of 14 compounds highly abundant in this one sample alone.
-The only identified compounds with this trend were taurine and acetyltaurine, though three others with *m/z* 146.1176, 148.0968, and 206.1388 were discovered to be likely isomers of known compounds (acetylcholine/(3-carboxypropyl)trimethylammonium, hydroxyisoleucine, dexpanthenol, respectively).
-One feature (*m/z* = 184.0637, RT = 7.0 minutes) was identified as choline sulfate using a second run of standards, while an additional mass feature at 168.0689 and RT = 8.5 was putatively flagged as taurine betaine (same mass as hydroxyisoleucine).
+The only identified compounds with this trend were taurine and acetyltaurine, though three others with *m/z* 146.1176, 148.0968, and 206.1388 were discovered to be likely isomers of known compounds (acetylcholine/(3-carboxypropyl)trimethylammonium, hydroxyisoleucine, and dexpanthenol, respectively).
+One feature (*m/z* = 184.0637, RT = 7.0 minutes) was identified as choline sulfate using a second run of standards, while an additional mass feature at 168.0689 and RT = 8.5 was putatively flagged as taurine betaine.
 This mass was also one of the most highly enriched in the MS11C215m_A sample after taurine itself, with a peak area nearly 30x that of the next largest sample.
-No clear explanation emerged explaining why some compounds were hyperabundant in this surface sample, though one possible explanation was the *Trichodesmium* bloom occurring at the same time.
+No clear explanation emerged explaining why some compounds were hyperabundant in this surface sample, though one possible explanation was the *Trichodesmium* bloom occurring at the same time [@Dugenne2023].
 
@@
     compound_name=="Dexpanthenol"~NA,
+    formula=="C5H14NO4S"~"Choline sulfate",
+    formula=="C5H14NO3S"~"Taurine betaine?",
     TRUE~compound_name
@@
 
-We explored this *Trichodesmium* hypothesis further by analyzing additional samples taken from the bloom to see whether the compounds abundant in *Trichodesmium* were also abundant in samples taken directly from the bloom itself.
-Of the 14 masses identified in the environmental sample, 7 had good-quality peaks when viewed in the bloom samples as extracted ion chromatograms, though only 4 of these matched closely in retention time.
+| Median m/z| Median RT|Putative formula |Compound name    |Feature ID  |
+|----------:|---------:|:----------------|:----------------|:-----------|
+|   126.0221|      11.1|C2H8NO3S         |Taurine          |MS_POS_0677 |
+|   146.1176|       8.5|C7H16NO2         |                 |MS_POS_1050 |
+|   148.0968|       8.5|C6H14NO3         |                 |MS_POS_1080 |
+|   168.0317|       4.9|C4H10NO4S        |N-Acetyltaurine  |MS_NEG_1123 |
+|   168.0689|       8.5|C5H14NO3S        |Taurine betaine? |MS_POS_1340 |
+|   176.1282|       8.1|C8H18NO3         |                 |MS_POS_1419 |
+|   184.0637|       7.0|C5H14NO4S        |Choline sulfate  |            |
+|   192.1230|       9.2|C8H17NO4         |                 |            |
+|   196.0816|       5.2|C6H13NO6         |                 |            |
+|   198.0794|       5.0|C6H16NO4S        |                 |            |
+|   206.1388|       7.1|C9H20NO4         |                 |MS_POS_1683 |
+|   208.1180|      10.2|C8H17NO5         |                 |            |
+|   222.1336|       8.5|C9H19NO5         |                 |            |
+|   266.1598|       8.5|C11H23NO6        |                 |            |
+|   266.1598|      10.0|C11H23NO6        |                 |            |
+
+*Table 1: Detected mass features with peak areas in sample MS11C215m_A much larger than the sample with the next-largest peak area.
+RT = retention time in minutes.
+Taurine betaine is followed by a ?
+to denote that this identification is putative.
+Feature ID is a disambiguating key produced by XCMS processing.*
 
@@
 
-Finally, we performed a search of the untargeted data for homoserine betaine, a known *Trichodesmium* osmolyte [@Pade2016].
+We explored this *Trichodesmium* hypothesis further by analyzing additional samples taken from the bloom to see whether compounds abundant in hand-picked *Trichodesmium* colonies were also abundant in samples taken directly from the bloom itself.
+After searching for the 14 mass features above in the environmental sample, 7 had good-quality peaks when viewed in the bloom samples as extracted ion chromatograms, though only 4 of these matched closely in retention time.
+We also performed a search of the untargeted data for homoserine betaine, a known *T.
+erythraeum* (strain IMS101) osmolyte [@Pade2016].
 Although this compound was not in our standard library due to a lack of authentic standards, we identified three mass features in the 162.1125 *m/z* window with retention times of 7.1, 8.4, and 9.9 minutes.
@@
 They did not report an additional peak detected at this mass, leaving our putative identification of the threonine betaine as low-confidence.
-None of the three 162.1125 features showed the same degree of enrichment in MS11C215m_A as taurine or the other 13 compounds detected with that pattern.
-Although MS11C215m_A was the largest, second-largest, and third-largest peak for carnitine, threonine betaine, and homoserine betaine, respectively, it was never more than triple the average surface value or double the next-largest value.
+None of the three 162.1125 features showed the same degree of enrichment in MS11C215m_A as taurine or the other 13 compounds detected, though MS11C215m_A was the largest, second-largest, and third-largest peak for carnitine, threonine betaine, and homoserine betaine, respectively.
 
@@
   scale_shape_manual(values = c(2, 6), breaks = c("Surface", "DCM")) +
-  labs(color="Sea level\nanomaly\n(cm)\n", shape="Depth") +
+  labs(color="Corr.\nSLA\n(cm)", shape="Depth") +
   theme_bw() +
@@
 In each case, the differences introduced by SLA were smaller than those of depth when compared directly, with even the high-resolution sampling around the DCM finding slightly more variance explained by the 40 meter difference in sampling depth than the 40 centimeter difference in SLA.
-This result agrees well with prior research by @Heal2021 and @Kumler2023 showing that depth has a large effect on the metabolome.
+This result agrees well with prior research by @Heal2021 and @Kumler2023 showing the large effect of sampling depth on the metabolome.
 
@@
 
-Given the confluence of both depth and SLA signals and the way that the boxplots of Figure 1 and Supplemental figure 3 would confound the SLA signal with the depth variance, we used k-means clustering to group similar compounds and provide a reduced dimensionality space for visualization.
+Given the confluence of both depth and SLA signals and the way that the boxplots of Figure 1 and Supplemental figure 1 would confound the SLA signal with the depth variance, we used k-means clustering to group similar compounds and provide a reduced dimensionality space for visualization.
 This revealed four major patterns of metabolite response, with a majority of compounds responding to eddy state (Clusters 1, 2, and 4 in Figure 4).
@@
 
+```{r}
+raw_kheal <- read_csv("https://raw.githubusercontent.com/kheal/Gradients1_SemiTargeted3/master/Tables/Manuscript_tables/SuppTables/Full_EnviroCulture_PeakAreas.csv")
+kheal_envirosamp <- raw_kheal %>%
+  filter(Column=="HILIC") %>%
+  mutate(polarity=ifelse(z>0, "pos", "neg")) %>%
+  distinct(MassFeature_Column, compound_name_kath=Identification, Confidence, mz, rt, polarity) %>%
+  mutate(rt=rt/60)
+
+untarg_feats %>%
+  filter(str_detect(feature, "FK")) %>%
+  select(feature, polarity, mzmed, rtmed, compound_name_will=compound_name) %>%
+  mutate(mz_lower=mzmed*(1-5/1e6)) %>%
+  mutate(mz_upper=mzmed*(1+5/1e6)) %>%
+  full_join(kheal_envirosamp, by=join_by(polarity, mz_lower<mz, mz_upper>mz)) %>%
+  filter(compound_name_kath=="Histidine") %>%
+  distinct(feature, MassFeature_Column, mzmed, rtmed, mz, rt, compound_name_will, compound_name_kath) %>%
+  print(n=Inf)
+
+```
+
 Several molecule- or pathway-specific narratives emerged from the metabolomic data described above as well.
 The untargeted approach used here allowed us to describe and chatacterize signals from small molecules whose identity was unknown, a particularly promising approach in open ocean gyres where the largest fraction of unknowns is found [@Heal2021].
-[Anitra: how many of your unknowns were also unknowns in Katherine's paper?] We found that this was especially true in anticyclonic eddies at all depths, with the diversity of the cyclone much better characterized by our list of authentic standards.
+We found that this was especially true in anticyclonic eddies at all depths, with the diversity of the cyclone much better characterized by our list of authentic standards.
 Many of the trends detected below persisted even after normalizing within each sample, indicating that the shifts discussed have implications beyond simple scaling with biomass.
@@
 The pattern was weaker in the noisier Falkor data but the largest concentrations were still detected in the cyclone and the median cyclone measurement was greater than the maximum value for the anticyclone.
-This molecule has a well-established role as an osmolyte in eukaryotic phytoplankton and cyanobacteria [@Gebser2013; @Dawson2020; @Heal2021; @Durham2022] as well as a documented decrease in abundance with depth [@Heal2021], indicating that this shift is potentially due to any combination of differences in salinity [Anitra: Really?
-Did salinity change much?
-Will: Half a PSU at 175m, 0.3 at DCM.
-Idk if that's enough - check Hannah's paper?], DCM depth, total biomass, and community composition.
+This molecule has a well-established role as an osmolyte in eukaryotic phytoplankton and cyanobacteria [@Gebser2013; @Dawson2020; @Heal2021; @Durham2022] as well as a documented decrease in abundance with depth [@Heal2021], indicating that this shift is potentially due to any combination of differences in location, DCM depth, total biomass, and community composition.
 Homarine was tightly correlated with trigonelline, and although they have structural similarity they are not known to have a relationship beyond their shared function as osmolytes.
@@
 
-Similarly to homarine, isethionate is a known osmolyte thus far found exclusively in eukaryotes [@Durham2022] and strongly associated with a few diatoms in particular [@ref or culture data].
+Similarly to homarine, isethionate is a known osmolyte thus far found exclusively in eukaryotes [@Durham2022] and strongly associated with a few diatoms in particular [@Heal2021].
 This compound was enriched in the cyclone of the DCM in both the MESOSCOPE and Falkor cruises along with its precursor taurine, validating earlier results from @Barone2022 that showed strong enrichment of most eukaryotes at the DCM and *Pseudo-nitzschia* in particular.
-[Expand on this with after additional meetings with Bryn and following up with Sam - TPA (K03851 should definitely have been picked up, poly-A selected for euks?, see if iseJ (the 1.1.2-) can still be searched for without a KO number (yes according to Frank bc the sequence exists but complicated by catabolism/anabolism))]
+Taurine was also among the compounds highly enriched in the anomalous surface sample from Station 11 (MS11C215m_A), likely along with its trimethylated form taurine betaine (N-trimethyltaurine).
+While putative in our analysis, taurine betaine has been previously reported in corals and red algae, where it probably has a destabilizing influence on proteins [@Impellizzeri1975; @Hill2017; @Bruzdziak2018].
+Although the distinctive taurine pattern could have been due to a real biological difference, it is difficult to assign much confidence to this hypothesis.
+While *Trichodesmium* are notoriously patchy [@Rouco2014; @Rouco2016], we did not see a strong match to the cultured cyanobacteria in any of the other metabolites with this pattern and the distinctive homoserine betaine osmolyte was missing [@Pade2016].
 
@@
 
-[Anitra: OK to comment on this but needs to be 3-4 sentences] Taurine was also among the compounds highly enriched in the anomalous surface sample from Station 11 (MS11C215m_A), likely along with its trimethylated form taurine betaine.
-While putative in our analysis, trimethyltaurine has been previously reported in corals and red algae, where it probably has a destabilizing influence on proteins [@Impellizzeri1975; @Hill2017; @Bruzdziak2018].
-We have a few hypotheses for the striking pattern in these compounds and the other 13 metabolites detailed above in Results.
-Most simply, the signal could be contamination during the sampling process, with some unknown material dominated by these compounds added to the filter during collection or processing.
-Alternatively, the signal could represent a real biological difference, perhaps indicating a pathway or organism highly abundant in this sample alone.
-As noted above, the MS11C215m_A sample was collected during a *Trichodesmium* bloom, an environment characterized by large heterogeneity even between replicate samples [@Rouco2014; @Rouco2016].
-10 liters of seawater were filtered for each sample but it is conceivable that a tuft or puff of *Trichodesmium* was captured in only this one as previously noted by @Rouco2014.
-The presence of a predictable trend across multiple chemically similar metabolites is the strongest bit of evidence in favor of the *Trichodesium* hypothesis, but the fact that these metabolites do not appear robustly in the bloom samples and that homoserine betaine did not appear to be especially abundant challenges it.
-One final note is that homoserine betaine was only documented as the sole major compatible solute in the Clade III *T.
-erythraeum* strain IMS101, while the bloom was more likely due to Clade I species given their relative abundance in the NPSG [@Rouco2016].
-[Email Sonya about the presence of compatible solute genes in Clade I tricho?
-Especially taurine/TMT?
-Ask about IFCB samples (they're on the dashboard http://ifcbdb.soest.hawaii.edu/timeline?dataset=MESOSCOPE&bin=D20170708T043509_IFCB000 I just don't know what to ask)]
-
 The strong SLA signal detected among the 175 meter samples was partially surprising to us, having expected the effect to be strongest at the DCM where isopycnals are interacting strongly with light from the surface.
 Instead, we found more compounds overall to be significantly different across the eddy transect in the 175 meter samples than we did at the DCM (α=0.05, 43 compounds at 175 meters vs 22 at the DCM).
-The enrichment of nucleobases in the 175 meter samples may hint at bacterial biomass enrichment, as nucleobases tend to be highly abundant in bacterial cultures [@ref or culture data].
+The enrichment of nucleobases in the 175 meter samples may hint at bacterial biomass enrichment, as nucleobases tend to be highly abundant in bacterial cultures [@Heal2021].
 The abundance of a mass feature putatively identified as acetylcholine in the 175 meter cyclone samples followed the same general trend as arsenobetaine, both enriched at depth and in the anticyclone, and may be another compound that's the result of heterotrophic degradation of phytoplankton metabolites [@Heal2021].
@@
 Aspartate, citrulline, ornithine, and malic acid were all found to be more abundant in the Falkor samples relative to MESOSCOPE.
-The OUC is a staple of metazoan metabolism and is present in stramenopiles as well where it enables rapid response to increases in nitrogen supply [@Allen2011; @Bender et al?].
+The OUC is a staple of metazoan metabolism and is present in stramenopiles as well where it enables rapid response to increases in nitrogen supply [@Allen2011; @Bender2014].
 The enrichment of these metabolites in the Falkor cruise may indicate that there was still rapid uptake and turnover of nitrogen in the 2018 eddy pair while the 2017 eddies were more stable, despite both eddies being in a similar state of stable anticyclone and weakening cyclone [@Dugenne2023].
@@
   ggtitle("B) Relative composition of the metabolome")
-egg::ggarrange(abs_targ_bar_gp, rel_targ_bar_gp, ncol=1)
-```
+comb_targ_rel_gp <- egg::ggarrange(abs_targ_bar_gp, rel_targ_bar_gp, ncol=1)
+
+ggsave("supplement/barplot_comb_targ_rel_gp.png", comb_targ_rel_gp, device = "png", width = 6.875, height = 5.5, units = "in", dpi = 300, bg="white")
+```
+
+![](supplement/barplot_comb_targ_rel_gp.png)
+
+*Supplemental figure 2: Stacked barplots of known metabolite concentration as shown in main text Figure 2 but broken down by triplicate.
+The same data is shown in A) as B) but B) is rendered in relative contribution space instead of absolute concentration.*
 
@@
 comb_pc_reg_df <- bind_rows(
-  frac_pc_in_metab_df %>% mutate(status="All files"),
+  frac_pc_in_metab_df %>% 
+    mutate(status="All files") %>%
+    mutate(PC_um=PC_uncor),
   no_outlier_frac_pc_in_metab_df %>%
@@
     theme_bw() +
-    theme(legend.position = "top"),
+    theme(legend.position = "top") +
+    labs(color="Sampling depth",
+         fill="Sampling depth",
+         x="Particulate carbon (uM)", 
+         y="Total nM carbon in known metabolites") +
+    scale_color_manual(breaks=c("Surface", "DCM", "Deep"), 
+                     values = c("#F4BB23", "#028E34", "#0B505C"), 
+                     aesthetics = c("fill", "color")),
   comb_pc_reg_df %>%
@@
     geom_smooth(method="lm", color="black") +
-    geom_point(aes(fill=sla_corr), pch=21, color="black", size=3) +
+    geom_point(aes(fill=sla_corr), pch=21, color="black", size=2) +
     ggh4x::facet_grid2(depth~status, scales="free", independent="all") +
@@
 
+ggsave("supplement/PC_reg_w_wo_outliers.png", PC_comb_gp, device = "png", width = 6.875, height = 4, units = "in", dpi = 300, bg="white")
+
+# filter(!(station=="11" & tripl=="A" & depth=="Surface")) %>%
+# filter(!(station=="6" & tripl=="B" & depth=="Surface")) %>%
+# filter(!(station=="7" & tripl=="C" & depth=="Surface")) %>%
+# filter(!(station=="11" & tripl=="B" & depth=="DCM"))
+
 # xmdl <- lm(total_nM~PC_um, data=all_targ_pc_reg_df)
@@
 
-*Supplemental figure 2: [TBD something about triplicates]*
-
-## Supplemental figure 3
+![](supplement/PC_reg_w_wo_outliers.png)
+
+*Supplemental figure 3: Regressions of particulate carbon against summed nM concentrations of known metabolites.
+The full regression with data from all sampling depths is shown on the left, while individual depths are shown on the right.
+For both the combined and depth-divided plots, Type I regression lines are shown both with and without outliers.
+Outlier data points were removed from the surface samples in Station 6, triplicate B; Station 7, triplicate C; and Station 11, triplicate A and from the DCM samples at Station 11, triplicate B.*
+
+## Supp fig: Falkor NMDS and PERMANOVA broken down by depth
 
@@
 
-*Supplemental figure 3: Plots of the metabolome in the Falkor data, broken down by depth.
+*Supplemental figure 4: Plots of the metabolome in the Falkor data, broken down by depth.
 The top row of plots are non-metric multidimensional scaling (NMDS) plots, in which points correspond to individual samples and have been colored by their corrected sea level height anomaly.
@@
   ggtitle("Todo: fix this graph")
-```
+ggsave("supplement/many_lin_reg_by_SLA.png", device = "png", width = 6.875, height = 5.5, units = "in", dpi = 300, bg="white")
+
+```
+
+![](supplement/many_lin_reg_by_SLA.png)
 

### 2024-03-07 manuscript/manuscript.Rmd
Intro edits done

@@
 A major source of variability comes in the form of mesoscale eddies in which water is entrained into circular surface currents tens to hundreds of kilometers in diameter [@Karl2017; @McGillicuddy2016].
-Prior research has established the importance of these mesoscale features in introducing nutrients into the euphotic zone and altering the community composition [@Sweeney2003; @Barone2019].
+These mesoscale features introduce nutrients into the euphotic zone and alter the community composition [@Sweeney2003; @Barone2019].
 In particular, the depth and intensity of the deep chlorophyll maximum (DCM) at the top of the nitricline differs dramatically between cyclonic and anticyclonic eddies.
@@
 Here, we use measurements of metabolites in particles across an eddy dipole in the North Pacific Subtropical Gyre (NPSG) to categorize both known compounds and unidentified mass features based on their co-occurrence with a set of simultaneously measured physical, chemical, and biological parameters.
-The gyre in particular represents an intriguing opportunity for metabolomic research because it has been shown to contain the highest proportion of chemical features with unknown structure and function and have highly variable responses to changes in latitude, depth, and across diel cycles [@Heal2021; @Boysen2021; @Kumler2023].
-This study connects biogeochemical measurements to a molecular-scale perspective of particulate organic matter, linking the chemical currencies that fuel metabolism and interactions among marine microbes to the mesoscale environment in which they’re found.
+The gyre in particular represents an intriguing opportunity for metabolomic research because it has been shown to contain the highest proportion of chemical features with unknown structure and function, many of which have highly variable responses to changes in latitude, depth, and across diel cycles [@Heal2021; @Boysen2021; @Kumler2023].
+This study connects biogeochemical measurements to a molecular-scale perspective of particulate organic matter, linking the chemical currencies that fuel metabolism and interactions among marine microbes to the mesoscale environment in which the currencies are found.
 

### 2024-03-12 manuscript/manuscript.Rmd
Manuscript as sent to coauthors

@@
 title: "Metabolites reflect variability introduced by mesoscale eddies in the North Pacific Subtropical Gyre"
-author: "William Kumler, Laura T. Carlson, Wei Qin, Rachel A. Lundeen, Benedetto Barone, and Anitra E. Ingalls"
+author: "William Kumler, Wei Qin, Rachel A. Lundeen, Benedetto Barone, Laura T. Carlson, and Anitra E. Ingalls"
 date: "`r Sys.Date()`"
@@
 - r-packages.bib
+csl: limnology-and-oceanography.csl
 ---
@@
 
-TBD
+Mesoscale eddies significantly alter open ocean environments such as the North Pacific Subtropical Gyre.
+Previous studies have explored eddy effects on biogeochemical measurements and community composition but not on the molecular composition of particulate organic matter.
+This study reports the exact concentration of 67 metabolites and relative abundances for 640 molecular features to understand how mesoscale eddies impact the NPSG's metabolome during two cruises in 2017 and 2018.
+We find that many metabolites track biomass trends, but biosignals like isethionic acid, homarine, and trigonelline linked to picoeukaryotes were enriched at the deep chlorophyll maximum of the cyclonic features, while degradation products such as arsenobetaine were enriched in anticyclones.
+Additionally, we identify taurine betaine (N,N,N-trimethyltaurine) in open ocean samples for the first time.
+By analyzing depth variability (accounting for 20-40% of metabolomic variability across ~150 meters) and isopycnal disruption (explaining 10-20% of variability across a sea level anomaly range of 40 centimeters), this analysis constrains the importance of mesoscale eddies in modeling the chemical composition of particulate matter in the largest biomes on the planet.
 
@@
 High frequency, long-term observations at Station ALOHA in the North Pacific Subtropical Gyre (NPSG) over the past 25 years have revealed temporal and spatial variability in what had prevously been considered a homogenous environment [@Karl1999; @Karl2017; @Karl2021].
-A major source of variability comes in the form of mesoscale eddies in which water is entrained into circular surface currents tens to hundreds of kilometers in diameter [@Karl2017; @McGillicuddy2016].
+A major source of variability comes in the form of mesoscale eddies in which water is entrained into circular surface currents tens to hundreds of kilometers in diameter [@Karl2017; @McGillicuddy2016; @Rii2022].
 These mesoscale features introduce nutrients into the euphotic zone and alter the community composition [@Sweeney2003; @Barone2019].
@@
 
-Using data for all molecular features, we found that the metabolome differed between the cyclone and anticyclone samples, with multivariate statistics detecting the largest differences in the DCM and 175 meter samples (Figure 1).
-At the DCM, approximately 16% of the variation in the metabolome could be explained by SLA alone according to the PERMANOVA (p = 0.003) and SLA was strongly correlated with the first principal component (r = .615, % variance explained by PC1 = 27.2), indicating that this was one of the largest sources of variation in the dataset.
+We found that the metabolome clearly differed between the cyclonic and anticyclonic samples when all molecular features were analyzed.
+Multivariate statistics detected the largest differences in particulate matter composition at the DCM and 175 meter samples (Figure 1).
+At the DCM, approximately 16% of the variation in the metabolome could be explained by sea level anomaly (SLA) alone (PERMANOVA, p = 0.003) and SLA was strongly correlated with the first principal component (r = .615, % variance explained by PC1 = 27.2), indicating that this was one of the largest sources of variation in the dataset.
 Notably, the samples taken from the exact center of the cyclonic eddy were highly distinct and likely drove much of the variance explained.
-In the 175m samples, 12% of the variance was explained by SLA (p = 0.008) according to the PERMANOVA and SLA was highly correlated with the second PC of the dataset (r = .697, % variance explained by PC2 = 15.4), though the first PC did not seem to have any visible pattern with metadata and appeared to capture largely variation between biological triplicates (Supplemental figure 1).
-At the surface, SLA trends are less evident with a larger p-value (0.087) and a lower R$^2$ (0.09) (Figure 1).
-
-To characterize this difference, we plotted an "average" metabolite peak area per station at each depth by taking the median value of all the detected mass features after z-score normalization (Figure 1).
-As expected, the average metabolite collected at the exact center of the cyclonic eddy (Station 12) was much more abundant than in any sample at this same depth in the rest of the dipole samples.
+In the 175m samples, 12% of the variance was explained by SLA (PERMANOVA, p = 0.008) and SLA was highly correlated with the second PC of the dataset (r = .697, % variance explained by PC2 = 15.4), though the first PC did not seem to have any visible pattern with metadata and appeared to largely capture variation between biological triplicates (Supplemental figure 1).
+At the surface, SLA trends are less evident with a larger p-value (PERMANOVA, p=0.087) and a lower R$^2$ (0.09) (Figure 1).
+
+To characterize the difference between eddies in a lower-dimensional space, we plotted an "average" metabolite peak area per station at each depth by taking the median value of each of the detected mass features after z-score normalization (Figure 1).
+As expected, the average metabolite collected at the exact center of the cyclonic eddy (Station 12) was more abundant than in other samples at this same depth.
 Similarly, the 175 meter samples showed the largest metabolite abundances in the anticyclone relative to other stations at this depth.
@@
 
-We expected a large part of the difference in metabolomes with respect to SLA to be due to the increased biomass in the cyclone DCM and the parallel increase in the anticyclone deep samples.
-Within our molecular feature population, a subset are known metabolites that we quantified using standards (Figure 2).
+We expected a large part of the SLA trends to be due to the elevated biomass in the cyclone DCM relative to the anticyclone and the inverse pattern in the 175 meter samples.
+While we were unable to quantify a majority of the molecular features in the dataset for direct comparison to biomass metrics, a subset of those were known metabolites that we quantified using standards (Figure 2).
 
@@
 
-We found that the overall concentration of known molecules was highest at the DCM of the cyclone, in line with our expectations and prior results.
-The most abundant quantified molecule overall was trimethylamine N-oxide (TMAO), with an average concentration of 1.38 nM and a distinct DCM maximum (Figure 2).
+We found that the overall concentration of known molecules was highest at the DCM of the cyclone.
+The most abundant quantified molecule in the eddy transect samples was trimethylamine N-oxide (TMAO), with an average concentration of 1.38 nM and a distinct DCM maximum (Figure 2).
 This pattern of a subsurface maximum at the DCM was mirrored by a majority of the known molecules (37/53).
-Most other abundant molecules decreased monotonically with depth, with highest values at the surface (Figure 2).
-
-```{r PC_vs_totnM regression values, eval=FALSE}
+Most other abundant molecules decreased monotonically with depth, with highest values at the surface.
+
+```{r PC_vs_totnM regression values}
 no_outlier_frac_pc_in_metab_df <- frac_pc_in_metab_df %>%
@@
 Known metabolites contributed between 1.5% and 5% of the carbon to the total PC load with higher values at the DCM than the surface or 175m samples (Figure 2, Supplemental Figure 3).
-DCM samples contained approximately 25% (0.5 percentage points) more carbon in the known metabolite pool (One-way ANOVA with outliers removed plus Tukey post-hoc, p<0.001, n=30-33) but there was no clear trend with SLA.
+DCM samples contained approximately 25% more carbon in the known metabolite pool (One-way ANOVA with outliers removed plus Tukey post-hoc, p<0.001, n=30-33) but there was no clear trend with SLA.
 
@@
 This mass was also one of the most highly enriched in the MS11C215m_A sample after taurine itself, with a peak area nearly 30x that of the next largest sample.
-No clear explanation emerged explaining why some compounds were hyperabundant in this surface sample, though one possible explanation was the *Trichodesmium* bloom occurring at the same time [@Dugenne2023].
+No clear explanation emerged for why some compounds were hyperabundant in this surface sample, though one possible explanation was the *Trichodesmium* bloom that was also sampled at this time [@Dugenne2023].
 
@@
 
-*Table 1: Detected mass features with peak areas in sample MS11C215m_A much larger than the sample with the next-largest peak area.
+*Table 1: Detected mass features that were enriched in the 15m sample from Station 11 (MS11C215m_A).
 RT = retention time in minutes.
@@
 
-We explored this *Trichodesmium* hypothesis further by analyzing additional samples taken from the bloom to see whether compounds abundant in hand-picked *Trichodesmium* colonies were also abundant in samples taken directly from the bloom itself.
+We explored the possibility that a nearby *Trichodesmium* bloom was responsible for the unusual composition of this one sample by analyzing additional samples taken directly from the bloom to see whether compounds abundant in hand-picked *Trichodesmium* colonies were also abundant the Station 11 samples.
 After searching for the 14 mass features above in the environmental sample, 7 had good-quality peaks when viewed in the bloom samples as extracted ion chromatograms, though only 4 of these matched closely in retention time.
@@
 The feature at RT 9.9 matched our carnitine standard's retention time, leaving the two earlier peaks with putative identifications of threonine betaine at 7.1 and homoserine betaine at 8.4 given the elution order of threonine before homoserine in our authentic standards.
-This order is the reverse of that in @Pade2016, which showed homoserine betaine eluting prior to carnitine as expected given our use of a polar HILIC column and their use of a nonpolar Hypercarb column.
+This order is the reverse of that in @Pade2016, which showed homoserine betaine eluting prior to carnitine.
+This elution order is expected given our use of a normal phse HILIC column and their use of a reversed phase Hypercarb column.
 They did not report an additional peak detected at this mass, leaving our putative identification of the threonine betaine as low-confidence.
@@
 
-In addition to absolute shifts in metabolite concentration due to biomass differences, we also explored compositional shifts in the metabolomes.
+In addition to absolute shifts in metabolite concentration due to biomass differences, we also explored compositional shifts in the metabolomes by assessing the fraction of the total metabolite pool contributed by each compound across the eddy transect.
 
@@
 To provide context for the sea level anomaly responses detected above, we additionally performed high-resolution depth sampling around the DCM at the center of both eddy poles.
-Here, we found that the ~40 meter difference in depth and the transition between the eddies explained similar amounts of variation in the data (R$^2_{depth}=0.253$, R$^2_{SLA}=0.238$) and were both highly significant factors (permutational p-values << 0.001).
-Additionally, both sample depth and SLA were well-correlated with the first principal component of the metabolite matrix (r$_{depth}=.766$, r$_{SLA}=.684$, fraction of variance explained by PC$_1$ = 30.8%) (Figure 3).
-A 2D NMDS plot of the samples showed good separation between the eddies, with cyclonic samples largely on the left side of Figure 3 while anticyclonic samples fell almost entirely to the right.
+Here, we found that the ~40 meter depth range and the transition between the eddies explained similar amounts of variation in the data (R$^2_{depth}=0.253$, R$^2_{SLA}=0.238$) and were both highly significant factors (permutational p-values << 0.001).
+Additionally, both sample depth and SLA were correlated with the first principal component of the metabolite matrix (r$_{depth}=.766$, r$_{SLA}=.684$, fraction of variance explained by PC$_1$ = 30.8%) (Figure 3).
+A 2D NMDS plot of the samples showed clear separation between the eddies, with cyclonic samples largely on the left side of Figure 3 while anticyclonic samples fell almost entirely to the right.
 Notably, the cyclonic samples show a much larger spread than the anticyclonic ones, indicating larger sample variability at the surface than the DCM.
@@
   geom_hline(yintercept = 0) +
-  geom_point(aes(x=MDS1, y=MDS2, color=sla_class), size=4, pch=19) +
+  geom_point(aes(x=MDS1, y=MDS2, color=sla_class), size=5, pch=19) +
   scale_color_manual(
@@
            hjust=0, vjust=1, label.r = unit(0, "in")) +
-  geom_point(aes(x=MDS1, y=MDS2, fill=depth, alpha=depth), size=4, pch=21, color="black") +
+  geom_point(aes(x=MDS1, y=MDS2, fill=depth, alpha=depth), size=5, pch=21, color="black") +
   scale_fill_manual(
-    values = c("grey90", "grey80", "grey60", "grey40", "grey20"),
+    values = c("white", "white", "white", "black", "black"),
     breaks = c("-20m", "-10m", "DCM", "+10m", "+20m")
@@
   scale_alpha_manual(
-    values = c(0.4, 0.2, 0, 0.2, 0.4),
+    values = c(0.7, 0.4, 0.2, 0.1, 0.4),
     breaks = c("-20m", "-10m", "DCM", "+10m", "+20m")
@@
   guides(alpha="none", color="none", fill=guide_legend(
-    title = "Depth", title.position = "top", title.hjust = 0.5, direction = "horizontal"
+    title = "Depth", title.position = "top", title.hjust = 0.5, direction = "horizontal",
+    override.aes = list(fill=c("grey90", "grey80", "grey60", "grey40", "grey20"))
   )) +
@@
   theme(legend.position = "bottom")
+MC_nmds_gp
 ggsave("MC_nmds_gp.png", plot = MC_nmds_gp, device = "png", 
@@
   summarise(mean_rank=mean(norm_area), iqr_value=sd(norm_area), .groups = "drop")
+library(ggh4x)
+strip_theme <- strip_themed(background_x = elem_list_rect(fill= c(
+  "#414487FF", "#7AD151FF", "#22A884FF", "#2A788EFF")))
 kclust_gp <- kclust_df %>%
   left_join(data.frame(size=kclust$size) %>% mutate(cluster=row_number()), by="cluster") %>%
-  mutate(cluster=paste0("Cluster #", cluster, ": n=", size)) %>%
+  mutate(clusterlab=paste0("Cluster #", cluster, ": n=", size)) %>%
   mutate(depth=factor(depth, labels=rev(c("-20m", "-10m", "DCM", "+10m", "+20m")),
@@
   geom_vline(xintercept = 0) +
-  # geom_ribbon(aes(xmin=mean_rank-iqr_value, xmax=mean_rank+iqr_value, group=sla_class),
-  #             linewidth=1, fill="#FFFFFF22") +
   geom_path(aes(x=mean_rank, group=sla_class), linewidth=1) +
-  geom_point(aes(x=mean_rank), size=2) +
-  geom_label(data = data.frame(cluster="Cluster #1: n=73", x=c(0, 0.9), depth=c(5.5, 1),
-                               sla_class=c("Anticyclone", "Cyclone")),
-             aes(x=x, label=sla_class)) +
-  facet_wrap(~cluster, nrow = 1) +
+  geom_point(aes(x=mean_rank), size=3) +
+  ggh4x::facet_wrap2(~clusterlab, nrow = 1, strip = strip_theme) +
   scale_color_manual(values = c(scales::muted("blue"), scales::muted("red"), "grey50"), 
@@
   theme_bw() +
-  theme(plot.background = element_rect(fill = "white"))
+  theme(plot.background = element_rect(fill = "white"),
+        strip.background = element_rect(fill = "white"),
+        strip.text = element_text(face = "bold", color="white"))
 ```
@@
   select(feature, fold_change)
-
 ```
@@
   labs(x="Log-2 fold-change cyclone enrichment", y="Mann-Whitney p-value\n(FDR-corrected)",
-       shape="Metabolite\nstatus", color="k-means\ncluster") +
-  guides(color=guide_legend(order = 2), shape=guide_legend(order=1)) +
+       shape="Metabolite status", color="k-means\ncluster") +
+  guides(color="none", shape=guide_legend(order=1)) +
   theme(axis.text.y = ggtext::element_markdown(),
-        plot.background = element_rect(fill = "white"))
+        plot.background = element_rect(fill = "white"),
+        legend.position = c(0, 1), legend.justification = c(0, 1), 
+        legend.background = element_rect(fill="#FFFFFF88", color=NA),
+        legend.spacing.y=unit(0, "in"),
+        legend.key.size = unit(0, "cm"))
 ```
@@
 kclust_volcano_gp <- egg::ggarrange(
-  kclust_gp, volcanoplot_gp + theme(legend.position = "none"), 
-  ncol=1, heights = c(2, 1.5)) %>%
-  cowplot::plot_grid(
-    cowplot::get_legend(volcanoplot_gp), rel_widths = c(0.85, 0.15)
-  )
+  kclust_gp, volcanoplot_gp, 
+  ncol=1, heights = c(2, 1.5))
 ggsave("kclust_volcano_gp.png", plot = kclust_volcano_gp, device = "png", 
@@
 The lower plot shows the individual known and unknown metabolites where points correspond to the FDR-corrected p-value estimated by the nonparametric Mann-Whitney U test and the log$_2$ fold-change calculated with the average peak area in the cyclone divided by the average peak area in the anticyclone.
-Colors have been assigned using the k-means clusters and shapes have been assigned based on the status of the mass feature as either a known metabolite that was run with an authentic standard or an unidentified metabolite.*
+Colors have been assigned using the k-means clusters and shapes have been assigned based on the status of the mass feature as either a known metabolite that was matched to an authentic standard or an unidentified metabolite.*
 
@@
 
-Many of the known metabolites enriched in the cyclone are expected given the overall higher biomass, particularly as many have functions as osmolytes in the cell.
+Many of the known metabolites enriched in the cyclone are expected given the overall higher biomass, particularly as many function as osmolytes in the cell.
 Metabolites enriched in excess of biomass were taurine and isethionic acid, the reduced sulfur osmolytes dimethylsulfoniopropionate (DMSP) and dimethylsulfonioacetate (DMS-Ac), and the isomers homarine and trigonelline.
@@
 
-The North Pacific Subtropical Gyre shows significant variation in its particulate composition due to mesoscale eddies.
-We sampled the metabolome across two sets of eddy dipoles to explore the effect of sea level anomaly (SLA).
-
-In all of the datasets analyzed here, we detected a significant difference in the composition of the metabolome between eddies.
+We measured the metabolome of samples collected across two sets of eddy dipoles to explore the effect of sea level anomaly (SLA) on metabolite composition and found that mesoscale eddies explain a significant portion of observed variations in particulate metabolomes.
+
+In all of the datasets analyzed here, we detected a significant difference in the composition of the metabolome between eddy dipoles.
 This effect was strongest in the samples taken during the Lagrangian stations at the center of each eddy in the 2017 MESOSCOPE cruise, with nearly a quarter of the total variance explained by the eddy from which the samples were taken.
@@
 
-Although possible that these differences are due to latitudinal shifts or simply background variation in the gyre environment, the transect data in particular implies that this is more due to eddies given the relative similarity of samples taken outside of the eddies.
-The transect sampled both outside of each edge and and between the two eddies across a large spatial gradient (~4° latitude and 2° longitude), yet the non-eddy samples from all three locations grouped together, showed similar median metabolite concentrations, and were visually similar in the most abundant metabolite composition.
-These results imply that strong cyclonic and anticyclonic eddies represent endpoints in NPSG composition, in agreement with results from @Barone2019 which found the most extreme values in the Hawaii Ocean Time series data to be usually detected when eddies passed over Station ALOHA.
+Although possible that the observed differences in metabolomes across eddy dipoles are due to latitudinal shifts or simply background variation in the gyre environment, the transect data in particular implies that this is not the case.
+Samples taken when the absolute value of the SLA was less than 5 cm were more similar to each other than to those samples taken within the eddies despite larger differences in latitude.
+The transect sampled both outside of each eddy and between the two across a large spatial gradient (~4° latitude and 2° longitude), yet the non-eddy (absolute value of SLA < 5 cm) samples from all three locations grouped together, showed similar median metabolite concentrations, and were visually similar in the most abundant metabolite composition.
+These results imply that strong cyclonic and anticyclonic eddies represent endpoints in NPSG composition, in agreement with results from @Barone2019 which found the most extreme values in the Hawaii Ocean Time series data typically detected when eddies passed over Station ALOHA.
 We also detected the largest differences in the DCM metabolome at the exact center of the eddy, with Station 12 highly distinct from even the nearest stations, while no such stark difference was found in the anticyclone (Figures 1 and 2).
@@
 
-A large portion of this effect was due to differences in biomass, with earlier work in the same location by @Barone2022 showing that particulate carbon (PC), chlorophyll, and beam attenuation were all 20-80% higher in the DCM of the cyclone relative to the anticyclone.
+A large driver of metabolomic differences was shifts in total biomass, with earlier work in the same location by @Barone2022 showing that particulate carbon (PC), chlorophyll, and beam attenuation were all 20-80% higher in the DCM of the cyclone relative to the anticyclone.
 We see this reflected particularly well in the targeted metabolites measured here, with clear depth trends and a strong correlation between total metabolite concentration and PC values.
-However, even after controlling for biomass effects by normalizing each sample to the sum of signal measured within it we still found a significant SLA effect, likely due to shifts in the community composition beyond bulk biomass as species are reduced or replaced entirely by the increased eukaryote presence in the cyclone DCM.
+However, even after controlling for biomass effects by normalizing each sample to the sum of signal measured within it we still found a significant SLA effect, likely due to shifts in the community composition and in particular the increased eukaryote presence in the cyclone DCM.
 
@@
 In contrast, the cyclonic samples were much more variable as is perhaps expected when the biomass is in the form of larger phytoplankton that are less homogenous in the environment.
-One notable aspect of their clustering was the way in which the deepest samples (DCM plus 20m) grouped most closely to the anticyclonic samples, perhaps indicating that below the DCM the metabolome rapidly approaches a uniform deep-water signal as seen in the 175 meter samples of the eddy transect in @Kumler2023, which showed that the samples from the deep euphotic zone were more similar to each other than those taken from the surface or DCM were to other surface or DCM samples.
-
-Given the confluence of both depth and SLA signals and the way that the boxplots of Figure 1 and Supplemental figure 1 would confound the SLA signal with the depth variance, we used k-means clustering to group similar compounds and provide a reduced dimensionality space for visualization.
+One notable aspect of their clustering was the way in which the deepest samples (DCM plus 20m) grouped most closely to the anticyclonic samples, perhaps indicating that below the DCM the metabolome rapidly approaches a uniform deep-water signal.
+This result was noted previously in the 175 meter samples of the eddy transect in @Kumler2023, where samples from the deep euphotic zone showed greater intra-depth similarity than samples from the DCM or surface.
+
+Given the confluence of both depth and SLA signals and the way that the boxplots of Figure 1 and Supplemental Figure 1 would confound the SLA signal with the depth variance, we used k-means clustering to group similar compounds and provide a reduced dimensionality space for visualization.
 This revealed four major patterns of metabolite response, with a majority of compounds responding to eddy state (Clusters 1, 2, and 4 in Figure 4).
@@
 Most surprising was the large difference between the eddies detected at the *surface*, while the DCM samples were functionally indistinguishable.
-This contrasts directly with the results from MESOSCOPE and the results in @Dugenne2023 and @Gleish2023, in which the largest differences were found at the DCM or below.
-This was especially surprising given the large inter-replicate differences between the Falkor surface samples, while the DCM samples tended to be much more consistent.
+This contrasts directly with the results from MESOSCOPE and the results in @Gleich2024, who found eddy-driven shifts in protistan community composition to be larger at depth than at the surface.
+However, @Dugenne2023 found differences in nitrogen fixation rate and nitrogen fixer composition varying widely throughout the upper water column.
+The large surface differences were especially surprising given the large inter-replicate differences between the Falkor surface samples in this study, while the DCM samples tended to be much more consistent.
 
@@
 
-Several molecule- or pathway-specific narratives emerged from the metabolomic data described above as well.
-The untargeted approach used here allowed us to describe and chatacterize signals from small molecules whose identity was unknown, a particularly promising approach in open ocean gyres where the largest fraction of unknowns is found [@Heal2021].
-We found that this was especially true in anticyclonic eddies at all depths, with the diversity of the cyclone much better characterized by our list of authentic standards.
+Several molecule- or pathway-specific narratives emerged from the metabolomic data.
+The untargeted approach used here allowed us to describe and characterize signals from small molecules whose identity was unknown, a particularly promising approach in open ocean gyres where the largest fraction of unknowns is found [@Heal2021].
+We found that our list of authentic standards covered fewer metabolites in anticyclonic eddies at all depths, with the molecular diversity of the cyclone much better characterized.
 Many of the trends detected below persisted even after normalizing within each sample, indicating that the shifts discussed have implications beyond simple scaling with biomass.
@@
 The pattern was weaker in the noisier Falkor data but the largest concentrations were still detected in the cyclone and the median cyclone measurement was greater than the maximum value for the anticyclone.
-This molecule has a well-established role as an osmolyte in eukaryotic phytoplankton and cyanobacteria [@Gebser2013; @Dawson2020; @Heal2021; @Durham2022] as well as a documented decrease in abundance with depth [@Heal2021], indicating that this shift is potentially due to any combination of differences in location, DCM depth, total biomass, and community composition.
-Homarine was tightly correlated with trigonelline, and although they have structural similarity they are not known to have a relationship beyond their shared function as osmolytes.
+This molecule has a well-established role as an osmolyte in eukaryotic phytoplankton and cyanobacteria [@Gebser2013; @Dawson2020; @Heal2021; @Durham2022] as well as a documented decrease in abundance with depth [@Heal2021], indicating that its response to SLA is potentially due to differences in physical and chemical attributes that in turn control biomass, community composition, and recycling rates.
+Homarine was also tightly correlated with trigonelline, and although they have structural similarity they are not known to have a relationship beyond their shared function as osmolytes.
 
@@
   coord_cartesian(ylim = c(0, 0.5), xlim=c(0.5, 2.5), expand = FALSE) +
-  scale_x_discrete(breaks = c("Cyclone", "Anticyclone"), labels = c("Cyclone\ncenter", "Anticyclone\ncenter")) +
   scale_color_manual(values = c(scales::muted("blue"), scales::muted("red")),
                      breaks = c("Cyclone", "Anticyclone")) +
-  theme_bw() +
-  theme(axis.text.y = element_blank(), axis.title.y = element_blank(),
-        legend.position = "none", axis.title.x = element_blank())
-```
-
-Similarly to homarine, isethionate is a known osmolyte thus far found exclusively in eukaryotes [@Durham2022] and strongly associated with a few diatoms in particular [@Heal2021].
+  theme_bw()
+
+```
+
+Isethionate is a known osmolyte thus far found exclusively in eukaryotes [@Durham2022] and strongly associated with a few diatoms in particular [@Heal2021].
 This compound was enriched in the cyclone of the DCM in both the MESOSCOPE and Falkor cruises along with its precursor taurine, validating earlier results from @Barone2022 that showed strong enrichment of most eukaryotes at the DCM and *Pseudo-nitzschia* in particular.
@@
 While putative in our analysis, taurine betaine has been previously reported in corals and red algae, where it probably has a destabilizing influence on proteins [@Impellizzeri1975; @Hill2017; @Bruzdziak2018].
-Although the distinctive taurine pattern could have been due to a real biological difference, it is difficult to assign much confidence to this hypothesis.
-While *Trichodesmium* are notoriously patchy [@Rouco2014; @Rouco2016], we did not see a strong match to the cultured cyanobacteria in any of the other metabolites with this pattern and the distinctive homoserine betaine osmolyte was missing [@Pade2016].
+While *Trichodesmium* are notoriously patchy [@Rouco2014; @Rouco2016], we did not see a strong match to the cultured cyanobacteria in any of the other metabolites with this pattern and the osmolyte homoserine betaine shown by @Pade2016 to be characteristic of *T.
+erythraeum* strain IMS101 was missing.
 
@@
 The enrichment of nucleobases in the 175 meter samples may hint at bacterial biomass enrichment, as nucleobases tend to be highly abundant in bacterial cultures [@Heal2021].
-The abundance of a mass feature putatively identified as acetylcholine in the 175 meter cyclone samples followed the same general trend as arsenobetaine, both enriched at depth and in the anticyclone, and may be another compound that's the result of heterotrophic degradation of phytoplankton metabolites [@Heal2021].
-These results raise important questions about the depth at which the community is isolated from SLA effects, with additional data from @Barone2022 and @Gleish2023 indicating that even at 250 meters eddy effects are still discernable.
-
-A final pattern worth noting is the apparent upregulation of metabolites in the urea-ornithine cycle (OUC) among the samples taken during the Falkor cruise.
+The abundance of a mass feature putatively identified as acetylcholine in the 175 meter cyclone samples followed the same general trend as arsenobetaine, both enriched at depth and in the anticyclone, and may be another compound that's the result of heterotrophic degradation of phytoplankton metabolites [@Durham2022].
+These results raise important questions about the depth at which the community is isolated from SLA effects, with additional data from @Barone2022 and @Gleich2024 indicating that even at 250 meters eddy effects are still discernable.
+
+A final pattern worth noting is the enrichment of metabolites in the urea-ornithine cycle (OUC) among the samples taken during the Falkor cruise.
 Aspartate, citrulline, ornithine, and malic acid were all found to be more abundant in the Falkor samples relative to MESOSCOPE.
@@
 More broadly, differences between the two cruises are most likely due to differences in eddy age and the earlier sampling time of the 2018 cruise (March/April instead of the June/July sampling in 2017) resulting in shifts in community composition and response.
-@Barone2022 documents the 2017 cruise community composition, reporting abundant Prymnesiophyceae at both the surface and DCM (largely), with other significant contributions from dinoflagellates, diatoms, and green algae.
-In the 2018 cruise both haptophytes (likely prymnesiophytes?) and dinoflagellates again dominated the eukaryotic fraction, though notably very little was contributed by diatoms [@Gleish2023].
-Overall we are inclined to agree with @Dugenne2023 and @Harke2021 in their concerns about generalizing eddy effects due to uncontrolled differences in these crucial factors and will note that many of the difficulties documented there are exacerbated when working with highly multidimensional metabolite data.
+@Barone2022 documents the 2017 cruise community composition, reporting abundant Prymnesiophyceae at both the surface and DCM, with other significant contributions from dinoflagellates, diatoms, and green algae.
+In the 2018 cruise both haptophytes and dinoflagellates again dominated the eukaryotic fraction, though notably very little was contributed by diatoms [@Gleich2024].
+@Dugenne2023 and @Harke2021 also cite uncontrolled differences between eddies based on their origin and age.
+
+Missing links between the environment with its associated community and the composition of particulate matter make it difficult to estimate this important control on the marine carbon cycle.
+This is especially true for dynamic and short-lived environments such as eddies.
+Metabolites offer a way to bridge this gap in knowledge but the relative impact of eddies on their contribution and fate was not previously estimated.
+In this manuscript, we have shown the sensitivity of the metabolome to various environmental and community factors and contrasted it to the well-established effect of depth to constrain the importance of including eddy effects in carbon cycle models.
 
@@
 ## Cruise information
-Samples were collected from two cruises in the North Pacific Subtropical Gyre near Station ALOHA that targeted strong mesoscale eddy features as described in @Dugenne2023 and @Gleish2023.
-Briefly, the 2017 MESOSCOPE cruise (KM1709 on the *R/V Kilo Moana*) consisted of a transect across the cyclonic and anticyclonic poles of the eddy dipole as well as two long-term Lagrangian stations at the center of each eddy.
+Samples were collected from two cruises in the North Pacific Subtropical Gyre near Station ALOHA that targeted strong mesoscale eddy features as described in @Dugenne2023 and @Gleich2024.
+Briefly, the 2017 MESOSCOPE cruise (Microbial Ecology of the Surface Ocean-Simons Collaboration on Ocean Processes and Ecology, KM1709 on the *R/V Kilo Moana*) consisted of a transect across the cyclonic and anticyclonic poles of the eddy dipole as well as two long-term Lagrangian stations at the center of each eddy.
 The cyclonic pole of the eddy had a maximum negative SLA anomaly of -20 centimeters and the anticyclonic center reached +24 centimeters.
@@
 
-Here, we measure the effect of mesoscale eddies on the metabolome of the North Pacific Subtropical Gyre in both absolute and relative space. [TBD].
+This study reports how changes in the biogeochemical environment and its associated community due to mesoscale eddy-driven mechanisms affect the metabolome and thereby the composition of particulate organic matter.
+The transect across an oligotrophic eddy dipole shows that many metabolites track closely with metrics of overall biomass, with eddy effects stronger at 175 meters than at the surface or deep chlorophyll maximum (DCM).
+High-resolution depth sampling of the DCM at the center of each eddy elucidated several known and unknown biomarkers likely corresponding to the well-documented increase in picoeukaryotes in the cyclone.
+Metabolite clusters aligned well with expected trends in abundance with depth and between the two eddies, with rapid convergence below the DCM towards a general deep-sea metabolite signal.
+By contrasting these effects with a followup analysis in the centers of a separate dipole the following year, we also learned that the importance of eddies may be an unpredictable interaction between origin and age.
+Finally, the metabolites with the strongest responses to eddy effects were unknown and a relatively smaller proportion of known metabolites in the anticyclone demonstrates the utility of untargeted metabolomics in exploring environmental variation and the large amount of information that can still be gleaned from such datasets.
 

### 2024-03-12 targeted/targeted_control.Rmd
Manuscript as sent to coauthors

@@
   coord_fixed()
+
+rfs <- read_csv("made_data/rf_ratios.csv")
+rfs %>%
+  select(compound_name, cruise, rf_matrix) %>%
+  pivot_wider(names_from = cruise, values_from = rf_matrix) %>%
+  ggplot(aes(x=Falkor, y=Mesotransect, label=compound_name)) +
+  geom_point() +
+  geom_abline(slope = 1, intercept = 0) +
+  lims(x=c(0, NA), y=c(0, NA)) +
+  coord_fixed() +
+  scale_y_log10() +
+  scale_x_log10()
+plotly::ggplotly()
 ```

### 2024-04-09 untargeted/untargeted_control.Rmd
Patching reproducibility gaps

---
title: "output_control_v2"
author: "William Kumler"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE, purl=TRUE}
knitr::opts_chunk$set(echo = TRUE, purl=FALSE)
options(pillar.sigfig=7)
options(readr.num_columns = 0)
options(readr.show_col_types = FALSE)
library(tidyverse)
library(data.table)
library(xcms)
library(RaMS)

# polarity <- "pos"
# polarity <- "neg"
# cruise <- "FK"
# cruise <- "MS"

output_folder <- paste0(c(cruise, polarity, "output/"), collapse = "_")
if(!dir.exists(output_folder))dir.create(output_folder)

mzML_files <- if(cruise=="FK"){
  list.files(paste0("../mzMLs/", polarity), full.names=TRUE, 
             pattern = "190715")
} else if(cruise=="MS"){
  list.files(paste0("../mzMLs/", polarity), full.names=TRUE, 
             pattern = "170706|180821|180205")
} else {
  stop("Unsupported cruise!")
}

parallel_param <- SnowParam(
  workers = parallel::detectCores()-1, 
  tasks = length(mzML_files), 
  progressbar = TRUE
)
output_csv_write <- function(object){
  obj_name <- deparse(substitute(object))
  write_csv(object, file = paste0(output_folder, obj_name, ".csv"))
}
filled_file_metadata <- read_csv("../metadata/filled_file_metadata.csv") %>% filter(polarity==!!polarity)
```

```{r Make metadata}
sample_type_vec <- c("Blk", "25mSmp", "DCMSmp", "175mSmp", "15mSmp", "Std", "Poo")
file_data <- data.frame(filename=mzML_files) %>%
  mutate(samp_type=str_extract(filename, "Blk|Smp|Std|Poo")) %>%
  mutate(depth=str_extract(filename, "25m|DCM|175m|15m")) %>%
  mutate(depth=ifelse(is.na(depth), "", depth)) %>%
  mutate(samp_group=factor(paste0(depth, samp_type), levels=sample_type_vec))
output_csv_write(file_data)
```

```{r Read in MS files}
register(BPPARAM = SerialParam(progressbar = TRUE))
msnexp <- readMSData(
  files = file_data$filename, 
  pdata = new("NAnnotatedDataFrame", file_data), 
  msLevel. = 1, 
  mode = "onDisk"
)
```

```{r Perform peakpicking}
register(BPPARAM = parallel_param)
cwp <- CentWaveParam(
  ppm = 5, 
  peakwidth = c(20, 80), 
  prefilter = c(3, 1e6), 
  snthresh = 0, 
  verboseColumns = TRUE, 
  extendLengthMSW = TRUE, 
  integrate = 2
)
msnexp_withpeaks <- findChromPeaks(msnexp, cwp)
```

```{r Perform RT correction}
register(BPPARAM = SerialParam(progressbar = TRUE))
obp <- ObiwarpParam(
  binSize = 0.1, 
  centerSample = round(nrow(file_data)/2), 
  response = 1, 
  distFun = "cor_opt"
)
msnexp_rtcor <- adjustRtime(msnexp_withpeaks, obp)

# Extract corrected RTs for later use in feature extraction
init_rts <- msnexp_rtcor %>%
  dropAdjustedRtime() %>%
  rtime() %>%
  as.data.frame() %>%
  rownames_to_column() %>%
  set_names(c("scanid", "init_rt"))
rt_corrections <- msnexp_rtcor %>%
  rtime() %>%
  as.data.frame() %>%
  rownames_to_column() %>%
  set_names(c("scanid", "new_rt")) %>%
  mutate(filename=basename(fileNames(msnexp_rtcor)[as.numeric(str_extract(scanid, "\\d+"))])) %>%
  left_join(init_rts, by="scanid") %>% 
  mutate(new_rt=round(new_rt/60, 10), rt=round(init_rt/60, 10)) %>%
  select(-init_rt, -scanid)
output_csv_write(rt_corrections)
```

```{r Correspondence}
register(BPPARAM = SerialParam(progressbar = TRUE))
pdp <- PeakDensityParam(
  sampleGroups = file_data$samp_group, 
  bw = 12, 
  minFraction = 0.1, 
  binSize = 0.001, 
  minSamples = 2
)
msnexp_grouped <- groupChromPeaks(msnexp_rtcor, pdp)
```

```{r Peak filling}
register(BPPARAM = parallel_param)
fpp <- FillChromPeaksParam(ppm = 2.5)
msnexp_filled <- fillChromPeaks(msnexp_grouped, fpp)

saveRDS(msnexp_filled, paste0(output_folder, "msnexp_filled.rds"))
```

```{r Organize into useful format}
msnexp_filled <- readRDS(paste0(output_folder, "msnexp_filled.rds"))
peak_data_long <- msnexp_filled %>%
  chromPeaks() %>%
  as.data.frame() %>%
  rownames_to_column() %>%
  mutate(peakidx=row_number())
peak_data <- msnexp_filled %>%
  featureDefinitions() %>%
  as.data.frame() %>%
  select(mzmed, rtmed, npeaks, peakidx) %>%
  rownames_to_column("id") %>%
  unnest_longer(peakidx) %>%
  rename_with(~paste0("feat_", .x), .cols = -peakidx) %>%
  dplyr::rename(feature="feat_id") %>%
  left_join(peak_data_long, by = join_by(peakidx)) %>%
  mutate(filename=basename(fileNames(msnexp_filled))[sample])

peak_bounds <- peak_data %>%
  select(feature, filename, mzmin, mzmax, rtmin, rtmax) %>%
  mutate(rtmin=rtmin/60, rtmax=rtmax/60)

output_csv_write(peak_bounds)
output_csv_write(peak_data)
```

```{r Raw data feature extraction}
file_data <- read_csv(paste0(output_folder, "file_data.csv"))
peak_bounds <- read_csv(paste0(output_folder, "peak_bounds.csv"))
rt_corrections <- read_csv(paste0(output_folder, "rt_corrections.csv"))
qscoreCalculator <- function(dt){
  rt <- dt$rt
  int <- dt$int
  #Check for bogus EICs
  if(length(rt)<5){
    return(data.frame(SNR=NA, peak_cor=NA))
  }
  #Calculate where each rt would fall on a beta dist (accounts for missed scans)
  scaled_rts <- (rt-min(rt))/(max(rt)-min(rt))

  # Create a couple different skews and test fit
  maybe_skews <- c(2.5,3,4,5) #Add 7 to catch more multipeaks and more noise
  #Add 2 to catch very slopey peaks and more noise
  best_skew <- maybe_skews[which.max(sapply(maybe_skews, function(x){
    cor(dbeta(scaled_rts, shape1 = x, shape2 = 5), int)
  }))]
  perf_peak <- dbeta(scaled_rts, shape1 = best_skew, shape2 = 5)
  peak_cor <- cor(perf_peak, int)

  #Calculate the normalized residuals
  residuals <- int/max(int)-perf_peak/max(perf_peak)
  #Calculate the minimum SD, after normalizing for any shape discrepancy
  old_res_sd <- sd(residuals)
  norm_residuals <- diff(residuals)
  new_res_sd <- sd(norm_residuals)
  while(new_res_sd<old_res_sd){
    old_res_sd <- new_res_sd
    norm_residuals <- diff(residuals)
    new_res_sd <- sd(residuals)
  }
  #Calculate SNR
  SNR <- (max(int)-min(int))/sd(norm_residuals*max(int))
  #Return the quality score
  return(data.frame(SNR=SNR, peak_cor=peak_cor))
}

library(furrr)
plan(multisession, workers = parallel::detectCores()-1)

peakshape_rawdata <- future_map(file_data$filename, function(filename_i){
  msdata_i <- grabMSdata(filename_i, grab_what="MS1", verbosity = 0) %>%
    pluck("MS1") %>%
    mutate(rt=round(rt, 10)) %>%
    left_join(rt_corrections, by=c("filename", "rt")) %>%
    select(rt=new_rt, mz, int, filename)
  peak_bounds %>%
    filter(filename==basename(filename_i)) %>%
    pmap(function(...){
      with(list(...), {
        msdata_i[mz%between%c(mzmin, mzmax)][rt%between%c(rtmin, rtmax)] %>%
          qscoreCalculator() %>%
          mutate(feature)
      })
    }) %>%
    bind_rows() %>%
    mutate(filename=basename(filename_i))
}, .progress = TRUE) %>%
  bind_rows()
peakshape_feature_summary <- peakshape_rawdata %>%
  group_by(feature) %>%
  summarise(med_SNR=median(SNR, na.rm=TRUE), med_cor=median(peak_cor, na.rm=TRUE))

output_csv_write(peakshape_rawdata)
output_csv_write(peakshape_feature_summary)
```

```{r Choose only quality features and remove blanks and summarize neatly}
peak_data <- read_csv(paste0(output_folder, "peak_data.csv"))
peakshape_feature_summary <- read_csv(paste0(output_folder, "peakshape_feature_summary.csv"))

blk_features <- peak_data %>%
  group_by(feature) %>%
  summarize(blk_area=mean(into[str_detect(filename, "_Blk_")]),
            smp_area=mean(into[str_detect(filename, "_Smp_")])) %>%
  filter(smp_area/blk_area<3)

clean_feat_data <- peak_data %>%
  filter(!is.na(feat_mzmed)) %>%
  group_by(feature) %>%
  summarise(mean_mz=unique(feat_mzmed), mean_rt=unique(feat_rtmed), 
            mean_area=mean(into)) %>%
  filter(mean_rt%between%c(30, 1200)) %>%
  left_join(peakshape_feature_summary, by = join_by(feature))
if(polarity=="pos"){
  classified_feats <- read_csv(paste0(output_folder, "classified_feats.csv")) %>%
    select(feature, feat_class)
  quality_feat_data <- clean_feat_data %>%
    left_join(classified_feats, by = join_by(feature))
} else {
  quality_model <- read_csv(paste0(cruise, "_pos_output/peakshape_feature_summary.csv")) %>%
    left_join(read_csv(paste0(cruise, "_pos_output/classified_feats.csv"))) %>%    
      filter(feat_class%in%c("Good", "Bad")) %>%
    mutate(feat_class=ifelse(feat_class=="Good", 1, 0)) %>%
    glm(formula=feat_class~med_cor+med_SNR, family = binomial)
  quality_feat_data <- clean_feat_data %>%
    mutate(pred_prob=predict(quality_model, ., type="response")) %>%
    mutate(feat_class=ifelse(pred_prob>0.5, "Good", "Bad"))
}

quality_peaks <- quality_feat_data %>%
  filter(feat_class%in%c("Good", "Stans only")) %>%
  select(feature) %>%
  anti_join(blk_features) %>%
  left_join(peak_data, by=join_by(feature), multiple="all") %>%
  select(feature, feat_mzmed, feat_rtmed, into, filename) %>%
  mutate(filename=basename(filename)) %>%
  arrange(into) %>%
  group_by(feature, filename) %>%
  slice(1) %>%
  ungroup() %>%
  complete(feature, filename, fill = list(into=0)) %>%
  group_by(feature, filename) %>%
  mutate(feat_mzmed=mean(feat_mzmed, na.rm=TRUE),
         feat_rtmed=mean(feat_rtmed, na.rm=TRUE))
output_csv_write(quality_peaks)
```

```{r De-adduct and de-isotope, purl=TRUE}
quality_peaks <- read_csv(paste0(output_folder, "quality_peaks.csv"))

addiso_df <- tribble(
  ~polarity, ~addiso_type, ~mz_diff,
  "both", "C13", 1.003355,
  "both", "N15", 0.997035,
  "both", "C13N15", 2.00039,
  "both", "O18", 2.004244,
  "pos", "Na", 22.98922-1.007276,
  "pos", "K", 38.96315-1.007276,
  "neg", "Cl", 34.96940+1.007276,
  "neg", "Ac", 59.01385+1.007276
) %>%
  filter(polarity%in%c(!!polarity, "both")) %>%
  select(-polarity)

possible_addisos <- quality_peaks %>%
  pivot_wider(names_from = filename, values_from = into) %>%
  select(feature, feat_mzmed, feat_rtmed) %>%
  cross_join(x=., y=., suffix=c("", "_match")) %>%
  filter(feature!=feature_match) %>%
  # filter(feature=="FT0340") %>%
  filter(feat_rtmed>feat_rtmed_match-10 & feat_rtmed<feat_rtmed_match+10) %>%
  select(-feat_rtmed, -feat_rtmed_match) %>%
  expand_grid(addiso_df) %>%
  mutate(addiso_mz=feat_mzmed+mz_diff) %>%
  # select(-addiso_type, -mz_diff) %>%
  group_by(feature) %>%
  rowwise() %>%
  filter(feat_mzmed_match%between%pmppm(addiso_mz, 20)) %>%
  select(feature, addiso=feature_match, addiso_type)

actual_addisos <- possible_addisos %>%
  left_join(quality_peaks, by=join_by(feature), multiple="all") %>%
  left_join(quality_peaks, by=join_by(filename, addiso==feature), suffix=c("", "_addiso")) %>%
  group_by(feature, addiso, addiso_type) %>%
  summarize(addiso_cor=cor(into, into_addiso, use="complete"), .groups="drop") %>%
  filter(addiso_cor>0.9) %>%
  select(feature, addiso, addiso_type)
deaddiso_peaks <- quality_peaks %>%
  filter(!feature%in%actual_addisos$addiso)
output_csv_write(deaddiso_peaks)
```

```{r Get manual IS integrations from Skyline, eval=!file.exists("IS_integrations/all_is.csv"), purl=TRUE}
# Check that SkylineRunner.exe is installed
if(system("IS_integrations/SkylineRunner.exe", show.output.on.console = FALSE)){
  # Website = https://skyline.ms/wiki/home/software/Skyline/page.view?name=SkylineInstall_64_21-1&submit=false
  stop("Please grab a copy of SkylineRunner from the website")
}

skyline_files <- list.files("IS_integrations", pattern = "sky$", full.names = TRUE)

success_vec <- sapply(skyline_files, function(skyfile){
  polarity <- str_extract(skyfile, "POSITIVE|NEGATIVE")
  polarity <- switch(polarity, "POSITIVE"="pos", "NEGATIVE"="neg")
  cruise <- str_extract(skyfile, "Falkor|Mesocenter|Mesotransect")
  cruise <- tolower(cruise)

  sky_cmd <- paste0(
    "IS_integrations/SkylineRunner.exe",
    " --in=", skyfile,
    ' --report-name="Ingalls_Lab_QE_Transition Results"', 
    ' --report-file=IS_integrations/', cruise, "_", polarity, ".csv"
  )
  failed <- system(sky_cmd)
  if(failed){
    stop(paste("Extraction failed for file", skyfile))
  }

  new_csv_name <- paste0("IS_integrations/", cruise, "_", polarity, ".csv")
  new_csv <- read.csv(new_csv_name)
  new_csv$polarity <- polarity
  new_csv$cruise <- cruise
  write.csv(new_csv, new_csv_name, row.names = FALSE)
})

csv_pattern <- "falkor|mesotransect|mesocenter_pos|neg.csv"
new_csvs <- list.files("IS_integrations/", pattern = csv_pattern, full.names = TRUE)

csv_list <- lapply(new_csvs, read.csv)
csv_master <- do.call(what = "rbind", csv_list) %>%
  filter(str_detect(Replicate.Name, "Poo|Smp|Blk")) %>%
  mutate(Area=as.numeric(Area))

# Fix sample 180821_Smp_MS6C315m_C which didn't get extraction standards added
# Current fix is to replace with median Area of all other sample files
ext_stans_list <- c(
  "Isethionic Acid, 13C2",
  "L-Cysteic Acid, D3",
  "Sulfoacetic Acid, 13C2",
  "Sulfolactic Acid, 13C3",
  "Taurine, D4"
)
update_df <- csv_master %>%
  filter(str_detect(Replicate.Name, "Smp")) %>%
  filter(polarity=="neg") %>%
  filter(Precursor.Ion.Name%in%ext_stans_list) %>%
  mutate(Area=as.numeric(Area)) %>%
  group_by(Precursor.Ion.Name) %>%
  summarise(Area=median(Area)) %>%
  mutate(Replicate.Name="180821_Smp_MS6C315m_C")
csv_master <- csv_master %>%
  rows_update(update_df, by = c("Precursor.Ion.Name", "Replicate.Name")) %>%
  mutate(cruise=case_when(
    cruise=="falkor" ~"FK", 
    cruise=="mesotransect" ~ "MT",
    cruise=="mesocenter" ~ "MC",
    TRUE ~ "Unknown")) %>%
  # Drop GMP since the peaks look terrible
  filter(Precursor.Ion.Name!="Guanosine Monophosphate, 15N5")

write.csv(csv_master, "IS_integrations/all_is.csv", row.names = FALSE)
```

```{r correct MC triplicate letter shifts, purl=TRUE}
deaddiso_peaks <- read_csv(paste0(output_folder, "deaddiso_peaks.csv"))
if(cruise=="MS"){
  tripl5A_factor <- deaddiso_peaks %>%
    left_join(filled_file_metadata) %>%
    filter(ms_run=="MC") %>%
    arrange(timestamp) %>%
    filter(tripl=="A") %>%
    select(feature, filename, into, timestamp) %>%
    mutate(scale_cat=ifelse(timestamp<as.POSIXct("2018-02-06 18:00:00", tz = "UTC"), "pre", "post")) %>%
    group_by(feature, scale_cat) %>%
    summarise(med_area=median(into, na.rm=TRUE)) %>%
    pivot_wider(names_from = scale_cat, values_from = med_area) %>%
    mutate(scale_factor=pre/post) %>%
    select(feature, scale_factor) %>%
    mutate(scale_factor=ifelse(is.finite(scale_factor), scale_factor, 1))
  deaddiso_peaks <- deaddiso_peaks %>%
    left_join(filled_file_metadata %>% select(filename, timestamp)) %>%
    mutate(scale_cat=ifelse(timestamp<as.POSIXct("2018-02-06 18:00:00", tz = "UTC"), "pre", "post")) %>%
    left_join(tripl5A_factor) %>%
    mutate(into=ifelse(scale_cat=="pre", into/scale_factor, into))

  triplAgBgC_factor <- deaddiso_peaks %>%
    left_join(filled_file_metadata) %>%
    filter(ms_run=="MC") %>%
    filter(samp_type=="Smp") %>%
    group_by(feature, tripl) %>%
    summarise(med_area=median(into, na.rm=TRUE)) %>%
    pivot_wider(names_from = tripl, values_from = med_area) %>%
    mutate(across(c(A, B, C), ~.x/mean(c(A, B, C)))) %>%
    # plotly::plot_ly(x=~A, y=~B, z=~C, type="scatter3d", mode="markers", opacity=0.2)
    pivot_longer(c(A, B, C), names_to = "tripl", values_to = "AgBgC_factor") %>%
    mutate(AgBgC_factor=ifelse(is.finite(AgBgC_factor), AgBgC_factor, 1)) %>%
    left_join(filled_file_metadata %>% filter(ms_run=="MC") %>% distinct(filename, tripl),
              multiple = "all") %>%
    select(-tripl)
  deaddiso_peaks <- deaddiso_peaks %>%
    left_join(filled_file_metadata %>% select(filename, tripl)) %>%
    left_join(triplAgBgC_factor, by=c("feature", "filename")) %>%
    mutate(AgBgC_factor=ifelse(is.finite(AgBgC_factor), AgBgC_factor, 1)) %>%
    mutate(into=into/AgBgC_factor)

}

IS_info <- "https://raw.githubusercontent.com/IngallsLabUW/Ingalls_Standards" %>%
  paste0("/8fb6679505f692ec7d1cb4bf54ff1c5d4967c9a6/Ingalls_Lab_Standards.csv") %>%
  read_csv() %>%
  filter(Column=="HILIC") %>%
  filter(Compound_Type=="Internal Standard") %>%
  select(name_new=Compound_Name, name_old=Compound_Name_Original)
IS_areas <- read_csv("IS_integrations/all_is.csv") %>%
  select(filename=Replicate.Name, compound_name=Precursor.Ion.Name, area=Area) %>%
  mutate(filename=paste0(filename, ".mzML")) %>%
  left_join(IS_info, by=join_by(compound_name==name_old)) %>%
  select(-compound_name) %>%
  dplyr::rename(IS_name=name_new, IS_area=area) %>%
  pivot_wider(names_from=IS_name, values_from = "IS_area") %>%
  mutate("Inj_vol"=ifelse(str_detect(filename, "Poo.*Half"), 0.5, 1)) %>%
  pivot_longer(-filename, names_to = "IS_name", values_to = "IS_area") %>%
  filter(IS_name!="Guanosine monophosphate, 15N5")
if(cruise=="MS"){
  tripl5A_factor_IS <- IS_areas %>%
    left_join(filled_file_metadata) %>%
    filter(ms_run=="MC") %>%
    arrange(timestamp) %>%
    filter(tripl=="A") %>%
    select(IS_name, filename, IS_area, timestamp) %>%
    mutate(scale_cat=ifelse(timestamp<as.POSIXct("2018-02-06 18:00:00", tz = "UTC"), "pre", "post")) %>%
    group_by(IS_name, scale_cat) %>%
    summarise(med_area=median(IS_area, na.rm=TRUE)) %>%
    pivot_wider(names_from = scale_cat, values_from = med_area) %>%
    mutate(scale_factor=pre/post) %>%
    select(IS_name, scale_factor) %>%
    mutate(scale_factor=ifelse(is.finite(scale_factor), scale_factor, 1))
  IS_areas <- IS_areas %>%
    left_join(filled_file_metadata %>% select(filename, timestamp)) %>%
    mutate(scale_cat=ifelse(timestamp<as.POSIXct("2018-02-06 18:00:00", tz = "UTC"), "pre", "post")) %>%
    left_join(tripl5A_factor_IS) %>%
    mutate(IS_area=ifelse(scale_cat=="pre", IS_area/scale_factor, IS_area))
  triplAgBgC_factor_IS <- IS_areas %>%
    left_join(filled_file_metadata) %>%
    filter(ms_run=="MC") %>%
    filter(samp_type=="Smp") %>%
    group_by(IS_name, tripl) %>%
    summarise(med_area=median(IS_area, na.rm=TRUE)) %>%
    pivot_wider(names_from = tripl, values_from = med_area) %>%
    mutate(across(c(A, B, C), ~.x/mean(c(A, B, C)))) %>%
    # plotly::plot_ly(x=~A, y=~B, z=~C, type="scatter3d", mode="markers", opacity=0.2)
    pivot_longer(c(A, B, C), names_to = "tripl", values_to = "AgBgC_factor") %>%
    mutate(AgBgC_factor=ifelse(is.finite(AgBgC_factor), AgBgC_factor, 1)) %>%
    left_join(filled_file_metadata %>% filter(ms_run=="MC") %>% distinct(filename, tripl),
              multiple = "all") %>%
    select(-tripl)
  IS_areas <- IS_areas %>%
    left_join(filled_file_metadata %>% select(filename, tripl)) %>%
    left_join(triplAgBgC_factor_IS, by=c("IS_name", "filename")) %>%
    mutate(AgBgC_factor=ifelse(is.finite(AgBgC_factor), AgBgC_factor, 1)) %>%
    mutate(IS_area=IS_area/AgBgC_factor)
}
```

```{r BMIS, purl=TRUE}
pooled_IS_areas <- IS_areas %>%
  filter(str_detect(filename, "_Poo_"))

BMIS_choices <- deaddiso_peaks %>%
  filter(str_detect(filename, "_Poo_")) %>%
  full_join(pooled_IS_areas, multiple = "all", by = join_by(filename)) %>%
  group_by(feature, IS_name) %>%
  mutate(plot_area=into/IS_area*mean(IS_area)) %>%
  summarise(cv=sd(plot_area)/mean(plot_area)) %>% 
  arrange(cv) %>%
  slice(1) %>%
  select(feature, IS_name) %>%
  filter(!is.na(feature))

clean_peak_data <- deaddiso_peaks %>%
  group_by(feature) %>%
  filter(var(into)>0) %>% # Remove all the stans-only for which all samps are zero
  ungroup() %>%
  left_join(BMIS_choices, by = join_by(feature)) %>%
  left_join(IS_areas, by = join_by(filename, IS_name)) %>%
  group_by(feature) %>%
  mutate(norm_area=into/IS_area*mean(IS_area)) %>%
  select(feature, filename, norm_area)

output_csv_write(clean_peak_data)
output_csv_write(BMIS_choices)
```

Additional feature-based data below

```{r Annotate standards}
deaddiso_peaks <- read_csv(paste0(output_folder, "deaddiso_peaks.csv"))
deaddiso_feats <- deaddiso_peaks %>%
  group_by(feature) %>%
  filter(!is.na(feat_mzmed)) %>%
  summarise(mzmed=unique(feat_mzmed), rtmed=unique(feat_rtmed)/60)
pol_stans <- read_csv("../metadata/clean_stans.csv") %>%
  mutate(compound_name=case_when(
    date_added>180821 & cruise=="MS" ~ paste0(compound_name, "?"),
    date_added>190715 & cruise=="FK" ~ paste0(compound_name, "?"),
    TRUE~compound_name
  )) %>%
  select(-date_added) %>%
  filter(polarity==!!polarity)

mass_matched_stans <- pol_stans %>%
  pmap(function(...){
    with(list(...), {
      deaddiso_feats %>%
        filter(mzmed%between%pmppm(stan_mz, 10)) %>%
        distinct(feature) %>%
        mutate(compound_name)
    })
  }, .progress=TRUE) %>%
  bind_rows()

mix_matched_stans <- deaddiso_peaks %>%
  # filter(feature%in%c("FT0264")) %>%
  filter(str_detect(filename, "_Std_.*Mix.InH2O")) %>%
  mutate(mix=str_extract(filename, "Mix(1|2)")) %>%
  # ggplot() + geom_boxplot(aes(x=mix, y=into, color=mix)) +
  # ggplot() + geom_col(aes(x=filename, y=into, fill=mix)) +
  # facet_wrap(~feature, ncol = 1, scales = "free_y")
  select(feature, mix, into) %>%
  pivot_wider(names_from = mix, values_from = into, values_fn = mean) %>%
  left_join(mass_matched_stans, by="feature", multiple="all") %>%
  left_join(pol_stans, by="compound_name") %>%
  mutate(mix_pick=log10(Mix1/Mix2)*ifelse(mix=="Mix1", 1, -1)) %>%
  arrange(desc(mix_pick)) %>%
  filter(!is.na(compound_name)) %>%
  group_by(compound_name) %>%
  slice(1) %>%
  ungroup() %>%
  arrange(mix_pick)
mix_matched_stans %>% 
  filter(compound_type=="Internal Standard") %>%
  select(feature, compound_name, stan_mz, stan_rt) %>%
  mutate(compound_name=str_remove(compound_name, ", .*"))

# mix_matched_stans %>%
#   group_by(compound_name) %>%
#   filter(n()>1)
# mix_matched_stans %>%
#   arrange(feature) %>%
#   group_by(feature) %>%
#   filter(n()>1)
# mix_matched_stans %>%
#   select(feature, compound_name, stan_rt) %>%
#   left_join(deaddiso_feats) %>%
#   ggplot(aes(label=compound_name)) +
#   geom_abline(intercept = c(-2:2), slope = 1, color=paste0("grey", c(80, 50, 20, 50, 80))) +
#   geom_point(aes(x=stan_rt, y=rtmed)) +
#   coord_fixed() + theme_bw()
# plotly::ggplotly()
```

```{r Manual annotation fixes}
rmAnno <- function(){

}
reAnno <- function(){

}

if(cruise=="FK") {
  if(polarity=="pos"){
    stan_annotations <- deaddiso_feats %>%
      left_join(mix_matched_stans %>% select(feature, compound_name), 
                by="feature", multiple="all") %>%
      mutate(compound_name=case_when(
        feature=="FT0502"~"Pyro from glutamate",
        feature=="FT0503"~"Pyro from glutamine",
        compound_name=="5-Oxoproline"~NA,
        compound_name=="L-Valine"~NA,
        compound_name=="L-Isoleucine"~NA,
        feature=="FT0180"~"beta-Alanine",
        feature=="FT0181"~"Sarcosine",
        feature=="FT0179"~"L-Alanine",
        feature=="FT0264"~"Dimethylglycine",
        feature=="FT0260"~"4-Aminobutyric acid",
        feature=="FT0357"~"L-Threonine",
        feature=="FT0356"~"L-Homoserine",
        compound_name=="Acetylcholine"~NA,
        feature=="FT0551"~"Allopurinol",
        feature=="FT0552"~"Hypoxanthine",
        feature=="FT0568"~"Adenine, 15N2",
        feature=="FT0565"~NA,
        TRUE~compound_name
      )) %>%
      arrange(compound_name) %>%
      group_by(compound_name) %>%
      slice(1)
    stan_annotations %>%
      group_by(feature) %>%
      filter(n()>1) %>%
      ungroup() %>%
      arrange(feature)
  } else if(polarity=="neg") {
    stan_annotations <- deaddiso_feats %>%
      left_join(mix_matched_stans %>% select(feature, compound_name), 
                by="feature", multiple="all")
  }
} else if(cruise=="MS") {
  if(polarity=="pos"){
    stan_annotations <- deaddiso_feats %>%
      left_join(mix_matched_stans %>% select(feature, compound_name), 
                by="feature", multiple="all")
  } else if(polarity=="neg") {
    stan_annotations <- deaddiso_feats %>%
      left_join(mix_matched_stans %>% select(feature, compound_name), 
                by="feature", multiple="all")
  }
}
output_csv_write(stan_annotations)
```

```{r Annotate formulae}
formula_annotations <- distinct(deaddiso_feats, feature, formula=NA)
output_csv_write(formula_annotations)
```

```{r Annotate structures and compound classes}
struct_annotations <- distinct(deaddiso_feats, feature, structure=NA)
class_annotations <- distinct(deaddiso_feats, feature, class=NA)

output_csv_write(struct_annotations)
output_csv_write(class_annotations)
```

```{r Clean up and write out, purl=TRUE}
peak_data <- read_csv(paste0(output_folder, "peak_data.csv"))
clean_peak_data <- read_csv(paste0(output_folder, "clean_peak_data.csv"))
stan_annotations <- read_csv(paste0(output_folder, "stan_annotations.csv")) %>%
  group_by(feature) %>%
  summarise(compound_name=paste(compound_name, collapse = "/"))
formula_annotations <- read_csv(paste0(output_folder, "formula_annotations.csv"))
class_annotations <- read_csv(paste0(output_folder, "class_annotations.csv"))
struct_annotations <- read_csv(paste0(output_folder, "struct_annotations.csv"))

clean_feat_data <- peak_data %>%
  filter(!is.na(feat_mzmed)) %>%
  group_by(feature) %>%
  summarise(mzmed=unique(feat_mzmed), rtmed=unique(feat_rtmed)/60) %>%
  inner_join(clean_peak_data %>% distinct(feature)) %>%
  left_join(stan_annotations) %>%
  left_join(struct_annotations) %>%
  left_join(class_annotations)

output_csv_write(clean_feat_data)
```

### 2024-04-09 manuscript/manuscript.Rmd
Patching reproducibility gaps

@@
                 destfile = "../mzMLs/ST002789_HILIC_POS.zip")
-  stop("File extraction has to be done manually - file too large for R's unzip")
+  unzip("../mzMLs/ST002789_HILIC_POS.zip", exdir = "../mzMLs/pos", unzip="unzip")
   file.remove("../mzMLs/ST002789_HILIC_POS.zip")
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, eval = FALSE)
+knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, eval = TRUE)
 library(tidyverse)
 library(RaMS)
+library(vegan)
+library(cowplot)
+library(viridis)
+library(scales)
+library(ggnewscale)
+library(egg)
+library(broom)
+library(lme4)
+library(openxlsx)
+library(readxl)
+library(ggh4x)
+library(ggtext)
+library(gridExtra)
+
+if(!dir.exists("supplement"))dir.create("supplement")
 

### 2024-05-02 manuscript/manuscript.Rmd
Head reset issues

@@
 library(gridExtra)
+library(openxlsx)
 
@@
   select(-polarity)
-clean_stans <- read_csv("../metadata/clean_stans.csv")
+clean_stans <- read_csv("../metadata/clean_stans.csv") %>%
+  mutate(n_C=as.numeric(str_extract(formula, "(?<=C)\\d+"))) %>%
+  mutate(n_N=as.numeric(str_extract(formula, "(?<=N)\\d+"))) %>%
+  mutate(n_N=case_when(
+    !is.na(n_N)~n_N,
+    str_detect(formula, "N")~1,
+    TRUE~0
+  ))
+
 untarg_feats <- read_csv("../untargeted/all_feats.csv")
@@
 
-Mesoscale eddies significantly alter open ocean environments such as the North Pacific Subtropical Gyre.
-Previous studies have explored eddy effects on biogeochemical measurements and community composition but not on the molecular composition of particulate organic matter.
-This study reports the exact concentration of 67 metabolites and relative abundances for 640 molecular features to understand how mesoscale eddies impact the NPSG's metabolome during two cruises in 2017 and 2018.
-We find that many metabolites track biomass trends, but biosignals like isethionic acid, homarine, and trigonelline linked to picoeukaryotes were enriched at the deep chlorophyll maximum of the cyclonic features, while degradation products such as arsenobetaine were enriched in anticyclones.
-Additionally, we identify taurine betaine (N,N,N-trimethyltaurine) in open ocean samples for the first time.
-By analyzing depth variability (accounting for 20-40% of metabolomic variability across ~150 meters) and isopycnal disruption (explaining 10-20% of variability across a sea level anomaly range of 40 centimeters), this analysis constrains the importance of mesoscale eddies in modeling the chemical composition of particulate matter in the largest biomes on the planet.
+Mesoscale eddies significantly alter open ocean environments such as those found in the subtropical gyres that cover a large fraction of the global ocean.
+Previous studies have explored eddy effects on biogeochemistry and microbial community composition but not on the molecular composition of particulate organic matter.
+This study reports the exact concentration of 67 metabolites and relative abundances for 640 molecular features to understand how mesoscale eddies impact the metabolome of the North Pacific Subtropical Gyre during two cruises in 2017 and 2018.
+We find that many metabolites track biomass trends, but biomarker metabolites like isethionic acid, homarine, and trigonelline linked to eukaryotic phytoplankton and enriched at the deep chlorophyll maximum of the cyclonic features, while degradation products such as arsenobetaine were enriched in anticyclones.
+By analyzing depth variability (accounting for 20-40% of metabolomic variability across ~150 meters) and eddy pumping (explaining 10-20% of variability across a sea level anomaly range of 40 centimeters), this analysis constrains the importance of mesoscale eddies in modeling the chemical composition of particulate matter in the largest biomes on the planet.
 
@@
 
-High frequency, long-term observations at Station ALOHA in the North Pacific Subtropical Gyre (NPSG) over the past 25 years have revealed temporal and spatial variability in what had prevously been considered a homogenous environment [@Karl1999; @Karl2017; @Karl2021].
-A major source of variability comes in the form of mesoscale eddies in which water is entrained into circular surface currents tens to hundreds of kilometers in diameter [@Karl2017; @McGillicuddy2016; @Rii2022].
-These mesoscale features introduce nutrients into the euphotic zone and alter the community composition [@Sweeney2003; @Barone2019].
-In particular, the depth and intensity of the deep chlorophyll maximum (DCM) at the top of the nitricline differs dramatically between cyclonic and anticyclonic eddies.
-@Barone2022 reported a shift in the microbial community at the DCM with eukaryotic phytoplankton enriched and prokaryotic phytoplankton depleted in cyclonic eddies, and @Hawco2021 and @Dugenne2023 showed that some of this difference was due to alterations in dissolved iron chemistry and nitrogen fixation.
-
-Here, we use measurements of metabolites in particles across an eddy dipole in the North Pacific Subtropical Gyre (NPSG) to categorize both known compounds and unidentified mass features based on their co-occurrence with a set of simultaneously measured physical, chemical, and biological parameters.
-The gyre in particular represents an intriguing opportunity for metabolomic research because it has been shown to contain the highest proportion of chemical features with unknown structure and function, many of which have highly variable responses to changes in latitude, depth, and across diel cycles [@Heal2021; @Boysen2021; @Kumler2023].
-This study connects biogeochemical measurements to a molecular-scale perspective of particulate organic matter, linking the chemical currencies that fuel metabolism and interactions among marine microbes to the mesoscale environment in which the currencies are found.
+High frequency, long-term observations at Station ALOHA in the North Pacific Subtropical Gyre (NPSG) over the past 25 years have revealed temporal and spatial variability in what had previously been considered a homogenous environment [@Karl1999; @Karl2017; @Karl2021].
+A major source of variability comes in the form of mesoscale eddies in which water is entrained into circular surface currents tens to hundreds of kilometers in diameter [@Karl2017; @McGillicuddy2016].
+
+Eddies are observed via satellite altimetry, which measures anomalies in sea surface height.
+Those that have a positive sea level anomaly (SLA) typically indicate regions of downwelling as water converges and are associated with clockwise rotation in the Northern Hemisphere.
+In contrast, a negative SLA usually corresponds to counterclockwise rotation in the Northern Hemisphere upwelling as deep water is lifted into the sunlit region of the ocean to replace water forced away from the eddy center [@McGillicuddy2016].
+While mode-water eddies have been observed and are an exception to the conventions established above [@Sweeney2003; @McGillicuddy2007], here we focus on the more common cyclonic (negative SLA) and anticyclonic (positive SLA) mesoscale features near station ALOHA [@Barone2019].
+
+The uplift of deep, nutrient-rich seawater into the euphotic zone disrupts microbial communities [@Rii2022].
+Measurements of chlorophyll in cyclones reveal a shallower and more intense maximum [@Barone2022; @Cornec2021] as a result of eukaryotic phytoplankton growing in the increased nutrient concentrations while cyanobacterial biomass is reduced [@Hawco2021].
+This growth of large eukaryotes corresponds to a net increase in biomass and primary productivity [@Benitez-Nelson2007], especially near the deep chlorophyll maximum [@Barone2022].
+In contrast, anticyclones produce conditions favorable to nitrogen fixation by accumulating organisms such as Crocosphaera and Trichodesmium [@Fong2008; @Olson2015; @Dugenne2020; @Dugenne2023].
+
+This shift in community structure and productivity should result in a corresponding shift in the composition of the particulate organic matter produced.
+Eukaryotic organisms have distinct chemical fingerprints from those of cyanobacteria and heterotrophic bacteria [@Heal2021; @Durham2022] and the relative contribution of these different taxa to total biomass should reflect their chemical composition.
+However, recent work from @Harke2021 showed that overall community function may be robust across eddy types at the surface while @Gleich2024 detected significant differences in protistan community composition and metabolic potential down to 250 meters, with heterotrophy-associated transcripts found at higher abundances in the cyclone.
+These results raise questions about the effect of mesoscale features on particulate organic matter composition.
 
 # Results
-## Metabolome variability across the eddy dipole
-
-We found that the metabolome clearly differed between the cyclonic and anticyclonic samples when all molecular features were analyzed.
-Multivariate statistics detected the largest differences in particulate matter composition at the DCM and 175 meter samples (Figure 1).
-At the DCM, approximately 16% of the variation in the metabolome could be explained by sea level anomaly (SLA) alone (PERMANOVA, p = 0.003) and SLA was strongly correlated with the first principal component (r = .615, % variance explained by PC1 = 27.2), indicating that this was one of the largest sources of variation in the dataset.
-Notably, the samples taken from the exact center of the cyclonic eddy were highly distinct and likely drove much of the variance explained.
-In the 175m samples, 12% of the variance was explained by SLA (PERMANOVA, p = 0.008) and SLA was highly correlated with the second PC of the dataset (r = .697, % variance explained by PC2 = 15.4), though the first PC did not seem to have any visible pattern with metadata and appeared to largely capture variation between biological triplicates (Supplemental figure 1).
-At the surface, SLA trends are less evident with a larger p-value (PERMANOVA, p=0.087) and a lower R$^2$ (0.09) (Figure 1).
-
-To characterize the difference between eddies in a lower-dimensional space, we plotted an "average" metabolite peak area per station at each depth by taking the median value of each of the detected mass features after z-score normalization (Figure 1).
-As expected, the average metabolite collected at the exact center of the cyclonic eddy (Station 12) was more abundant than in other samples at this same depth.
-Similarly, the 175 meter samples showed the largest metabolite abundances in the anticyclone relative to other stations at this depth.
-This method also allowed us to see the remarkably sinusoidal response of the average metabolite across the stations.
-This response correlated negatively with SLA at the DCM (lower SLA = higher average values, $r_{DCM}= -.716$) but positively with SLA at the surface and at depth ($r_{surface} = .211$, $r_{deep} = .830$).
+
+We explored the impacts of isopycnal uplift and depression on the composition of particulate organic matter across pairs of mesoscale eddies of opposite polarity in the North Pacific Subtropical Gyre (NPSG).
+The first eddy pair was the focus of the July 2017 Microbial Ecology of the Surface Ocean-Simons Collaboration on Ocean Processes and Ecology (MESO-SCOPE, KM1709 aboard the R/V *Kilo Moana*) cruise and sampling consisted of two phases (Figure 1).
+In the first phase, a transect was taken from the north to the south consisting of 11 stations corresponding to the sampling sites in @Barone2022 and @Dugenne2023.
+Metabolomic samples of seawater from 15 meters, the deep chlorophyll maximum (DCM), and 175 meters were collected at each station via 0.2 µm filter.
+The next phase of the MESO-SCOPE cruise focused on Lagrangian sampling of the eddy centers, starting with the cyclone (SLA = -14 cm) before progressing to the anticyclone (SLA = 24 cm).
+There, metabolomics samples were taken from the DCM as well as 10 and 20 meters above and below it.
+The second eddy pair was targeted during the March 2018 Hawaiian Eddy Experiment (HEE, FK180310 aboard the R/V *Falkor*).
+During this followup cruise, a strong anticyclone (SLA = 21 cm) and a nearby cyclone (SLA = -13 cm) were targeted for a similar set of biogeochemical measurements described in @Dugenne2023 and @Gleich2024 (Figure 1).
+Metabolomics samples were taken from the center of each eddy as described above at both 25 meters and the DCM.
+We analyzed these three different datasets separately and discuss the findings from each in sequence below.
+
+![](figures/MapForWill_v3.jpg)
+
+*Figure 1: Sampling during MESO-SCOPE and the Hawaiian Eddy Experiment (HEE).
+Cruise bounds are shown in the large central map with Station ALOHA colored as a point in red near the Hawaiian Islands.
+Yellow stars denote sampling locations and station numbers relative to the sea level anomaly contours in the background during both the June 2017 MESO-SCOPE cruise (left) and the April 2018 HEE cruise (right).*
+
+Metabolomics samples from all three sample sets were extracted via a modified Bligh & Dyer [@Boysen2018], with the aqueous fraction run on a HILIC column paired with a Q-Exactive HF mass spectrometer.
+Metabolites for which an authentic standard was available and run alongside were integrated in Skyline (HEE and MESO-SCOPE eddy centers) and MSDIAL (MESO-SCOPE eddy transect).
+Untargeted analysis was performed using the R package XCMS with each feature manually inspected for quality assurance.
+
+## Metabolome variability across adjacent mesoscale eddies of opposing polarity during MESO-SCOPE explained by variations in sea level anomaly
+
+The metabolome clearly differed between cyclonic and anticyclonic samples, though the magnitude of this difference varied by depth.
+The largest differences in particulate matter composition were detected at the DCM and 175 meter samples, as shown by the NMDS in Figure 2.
+At the DCM, approximately 16% of the variation in the metabolome could be explained by sea level anomaly (SLA) alone (PERMANOVA, p = 0.003).
+Additionally, SLA was strongly correlated with the first principal component (PC) of the metabolome  with a Pearson's *r* of .615 and 27% of the variance explained by PC1 (Supplemental figure 1A), indicating that this was one of the largest sources of variation in the dataset.
+Notably, the samples taken from the exact center of the cyclonic eddy at Station 12 were highly distinct and likely drove much of the variance explained.
+In the 175 meter samples, 12% of the variance was explained by SLA (PERMANOVA, p = 0.008) and SLA was highly correlated with the second PC of the dataset (r = .697, % variance explained by PC2 = 15.4, Supplemental figure 1A), though the first PC did not seem to have any visible pattern with metadata and appeared to largely capture variation between biological triplicates (Supplemental figure 1B).
+At 15 meters, SLA trends are less evident with a larger p-value (PERMANOVA, p=0.087) and a lower R$^2$ (0.09) (Figure 2).
 
@@
   mutate(depth=factor(depth, levels=c("Surface", "DCM", "Deep"), 
-                      labels=c("Surface  (15m)", "DCM  (110-120m)", "Deep  (175m)"))) %>%
+                      labels=c("Surface  (15 m)", "DCM  (110-120 m)", "Deep  (175 m)"))) %>%
   filter(ms_run=="MT") %>%
@@
   mutate(perm_pval=round(perm_pval, digits = 4)) %>%
-  mutate(perm_label=paste0("R2 = ", perm_r2, "\np = ", perm_pval))
+  mutate(perm_label=paste0("R^2^ = ", perm_r2, "<br>p = ", perm_pval))
 ```
@@
   geom_point(aes(x=MDS1, y=MDS2, fill=sla_corr), color="black", size=4, pch=21) +
-  geom_label(x=-Inf, y=-Inf, aes(label=nmds_stress), hjust=0, vjust=0, 
+  geom_richtext(x=-Inf, y=Inf, aes(label=perm_label), hjust=0, vjust=1, fill="#FFFFFF33",
+             data = multivar_analysis, label.r = unit(0, "in"), lineheight = .7) +
+  geom_richtext(x=-Inf, y=-Inf, aes(label=nmds_stress), hjust=0, vjust=0, fill="#FFFFFF33",
              data = multivar_analysis, label.r = unit(0, "in")) +
-  geom_label(x=-Inf, y=Inf, aes(label=perm_label), hjust=0, vjust=1,
-             data = multivar_analysis, label.r = unit(0, "in"), lineheight = .7) +
   facet_wrap(~depth, nrow=1) +
@@
   ) +
-  labs(fill="Sea level\nanomaly\n(cm)\n") +
+  labs(fill="Corr.\nSLA\n(cm)\n", x="NMDS axis 1", y="NMDS axis 2") +
   theme_bw() +
+  theme(strip.text = element_text(size = 12)) +
   coord_fixed()
@@
 #   summarise(sim=cor(sla_corr, grand_avg_amt))
+mean_med_metab_df <- med_metab_df %>%
+  mutate(station=factor(station, levels=as.character(4:14))) %>%
+  mutate(depth=factor(depth, levels=c("Surface", "DCM", "Deep"))) %>%
+  group_by(depth, station, sla_corr) %>%
+  summarise(mean_val=mean(avg_amt))
 med_metab_sla_gp <- med_metab_df %>%
@@
   geom_hline(yintercept = 0) +
-  geom_boxplot(aes(x=station, y=avg_amt, color=sla_corr), linewidth=0.8) +
+  # geom_col(aes(x=station, y=mean_val, fill=sla_corr), color="black", 
+  #            data=mean_med_metab_df) +
+  geom_point(aes(x=station, y=avg_amt), color="black", size=2) +
+  geom_point(aes(x=station, y=mean_val, fill=sla_corr), color="black",
+             size=4, pch=21, data=mean_med_metab_df, alpha=0.7) +
   facet_wrap(~depth, nrow=1) +
@@
   labs(x="Station", y="Median metabolite z-score")
+
+# med_metab_sla_gp <- med_metab_df %>%
+#   mutate(station=factor(station, levels=as.character(4:14))) %>%
+#   mutate(depth=factor(depth, levels=c("Surface", "DCM", "Deep"))) %>%
+#   ggplot() +
+#   geom_hline(yintercept = 0) +
+#   geom_point(aes(x=station, y=avg_amt, fill=sla_corr), color="black", size=4, pch=21) +
+#   facet_wrap(~depth, nrow=1) +
+#   scale_color_gradient2(low = scales::muted("blue"), high = scales::muted("red"),
+#                         mid="grey90", aesthetics = c("color", "fill"), guide = NULL) +
+#   theme_bw() +
+#   theme(strip.background = element_blank(), strip.text = element_blank()) +
+#   labs(x="Station", y="Median metabolite z-score")
 
@@
 
-*Figure 1: Plots of the metabolome across the eddy dipole, broken down by depth.
-The top row of plots are non-metric multidimensional scaling (NMDS) plots, in which points correspond to individual samples and have been colored by their corrected sea level height anomaly.
+*Figure 2: Distribution of metabolites in multivariate space across adjacent eddies of opposite polarity during the MESO-SCOPE transect, broken down by depth.
+Top panels depict non-metric multidimensional scaling (NMDS) plots with individual samples colored based on their corrected sea level anomaly.
 NMDS stress values (s) have been reported in the bottom left corner, while PERMANOVA R$^2$ and p-values are reported in the top left.
-SLA trends are visible in the DCM and deep samples, with dark blue circles consistently discriminating from the dark red circles along the first multidimensional axis.
-The bottom row of plots show the direction and magnitude of this effect by plotting the grand median of the z-score normalized metabolite peak areas in the stations taken across the eddy dipole with the upper whisker extent, median bar, and lower whisker extent corresponding to the values measured in three biological triplicates.*
-
-We expected a large part of the SLA trends to be due to the elevated biomass in the cyclone DCM relative to the anticyclone and the inverse pattern in the 175 meter samples.
-While we were unable to quantify a majority of the molecular features in the dataset for direct comparison to biomass metrics, a subset of those were known metabolites that we quantified using standards (Figure 2).
+SLA trends are visible in the DCM and 175 meter samples, with dark blue circles consistently discriminating from the dark red circles along the first multidimensional axis.
+Bottom panels depict the direction and magnitude of this effect by plotting the grand mean value of three biological triplicates in color and the raw values in black behind.*
+
+To characterize the difference between eddies in a lower-dimensional space, we plotted an average metabolite peak area per station at each depth.
+We performed z-score normalization on each mass feature to give them all equal weight while preserving the variance between samples then calculated the median z-score for each sample, resulting in large positive values when metabolites are more abundant in a given sample than in the median sample (Figure 2).
+The average metabolite collected at the exact center of the cyclonic eddy (Station 12) had higher values than other samples at this same depth, indicating that most metabolites were more abundant at the cyclone's DCM than elsewhere in the transect.
+The 175 meter samples, on the other hand, showed the largest metabolite abundances in the anticyclone while the cyclone had consistently lower median z-scores.
+This method highlighted the negative correlation between the average metabolite and SLA at the DCM ($r_{DCM}= -.716$) while the 15 meter and 175 meter samples showed the opposite trend ($r_{15m} = .211$, $r_{175m} = .830$).
+
+```{r perc increase at cyclone DCM center, eval=FALSE}
+filled_file_metadata %>%
+  filter(ms_run=="MT") %>%
+  filter(depth=="DCM") %>%
+  distinct(cruise, station, cast, depth, PC_um) %>%
+  group_by(MT_DCM_cyc_center=ifelse(station==12, "St12", "Other")) %>%
+  summarise(mean_val=mean(PC_um)) %>%
+  pivot_wider(names_from=MT_DCM_cyc_center, values_from = mean_val) %>%
+  mutate(perc_increase=St12/Other)
+
+targ_peaks %>%
+  left_join(filled_file_metadata, by="filename") %>%
+  filter(ms_run=="MT") %>%
+  filter(samp_type=="Smp") %>%
+  filter(depth=="DCM") %>%
+  group_by(filename, station, tripl, PC_um) %>%
+  summarize(tot_nM=sum(nM)) %>%
+  group_by(MT_DCM_cyc_center=ifelse(station==12, "St12", "Other")) %>%
+  # group_by(MT_DCM_cyc_center=ifelse(station==12, "St12", "Other"), tripl) %>%
+  summarise(mean_val=mean(tot_nM)) %>%
+  pivot_wider(names_from=MT_DCM_cyc_center, values_from = mean_val) %>%
+  mutate(perc_increase=St12/Other)
+```
+
+```{r PC_vs_totnM regression values, include=FALSE}
+frac_pc_in_metab_df <- targ_peaks %>%
+  filter(str_detect(filename, "180821_Smp")) %>%
+  left_join(clean_stans %>% select(compound_name, formula, polarity, n_C, n_N)) %>%
+  mutate(nM_C=n_C*nM, nM_N=n_N*nM) %>%
+  group_by(filename) %>%
+  summarize(total_nM_C=sum(nM_C), total_nM_N=sum(nM_N)) %>%
+  left_join(filled_file_metadata, by = join_by(filename)) %>%
+  mutate(frac_C_in_metab=total_nM_C/1000/PC_um,
+         frac_N_in_metab=total_nM_N/1000/PN_um) %>%
+  mutate(depth=str_extract(depth, "DCM|Deep|Surface")) %>%
+  mutate(depth=factor(depth, levels=c("Surface", "DCM", "Deep"))) %>%
+  mutate(station=factor(station, levels=c(as.character(4:14))))
+
+no_outlier_frac_pc_in_metab_df <- frac_pc_in_metab_df %>%
+  filter(!(station=="11" & tripl=="A" & depth=="Surface")) %>%
+  ungroup()
+
+# Overall trend between PC and total nM metabolites (all depths)
+broom::tidy(lm(total_nM_C~PC_um, data=no_outlier_frac_pc_in_metab_df))[2,]
+with(no_outlier_frac_pc_in_metab_df, cor(PC_um, total_nM_C))
+lme4::lmer(total_nM_C~PC_um + (1|depth), data=no_outlier_frac_pc_in_metab_df)
+
+# By-depth trend between PC and total nM metabolites
+no_outlier_frac_pc_in_metab_df %>%
+  group_by(depth) %>%
+  summarise(overall=broom::tidy(lm(total_nM_C~PC_um))[2,], 
+            cor_vals=cor(PC_um, total_nM_C))
+
+# Same for PN
+broom::tidy(lm(total_nM_N~PN_um, data=no_outlier_frac_pc_in_metab_df))[2,]
+with(no_outlier_frac_pc_in_metab_df, cor(PN_um, total_nM_N))
+lme4::lmer(total_nM_N~PN_um + (1|depth), data=no_outlier_frac_pc_in_metab_df)
+no_outlier_frac_pc_in_metab_df %>%
+  group_by(depth) %>%
+  summarise(overall=broom::tidy(lm(total_nM_N~PN_um))[2,], 
+            cor_vals=cor(PN_um, total_nM_N))
+
+filled_file_metadata %>%
+  distinct(PC_um, PN_um) %>%
+  drop_na() %>%
+  cor()
+```
+
+This shift in metabolite abundance was largely driven by a corresponding shift in biomass.
+Particulate carbon and summed metabolite concentration were tightly correlated across all samples and depths ($\beta = 24.9\pm1.82$ (SE), Pearson's $r$ = .813, p-value < 0.001, Supplemental figure 2).
+This overall trend was largely driven by higher values at 15 meters and the DCM than at depth, but when analyzed at each depth individually the general correspondence held (Supplemental figure 2).
+The DCM and 175 meter samples had the stronger trends ($\beta_{DCM} = 22.6\pm5.53$, $r= .593$, p-value < 0.001; $\beta_{175m} = 23.9\pm7.36$, $r= .503$, p-value = 0.003) while 15 meter samples showed no significant relationship ($\beta_{15m} = 3.4\pm12.9$, $r= .049$, p-value = 0.792).
+This meant that metabolites contributed a fixed fraction of carbon to the particulate pool with our 53 quantified metabolites representing between 1.5% and 5% of the carbon in the system (Figure 2).
+Particulate nitrogen was tightly correlated with particulate carbon ($r=.933$) but the fraction of particulate nitrogen represented by the quantified metabolites was slightly higher across the transect with contributions typically between 2% and 6%.
+
+Samples taken from the DCM of the cyclone center (Station 12, Figure 2) showed especially high particulate carbon (67% increase, rising from 1.71 μM across all other DCM samples to 2.85 μM at Station 12) and the highest measured total metabolite concentrations (73% higher, increasing from 10.0 nM to 17.2 nM).
+This explains the large separation observed between these samples and the rest of the data in the NMDS plot of Figure 2.
 
 ```{r figure stacked barplot MT eddy transect w PC and SLA}
+arrow_df <- data.frame(
+  cyc_type=c("Anticyclone\ncenter", "Cyclone\ncenter"),
+  x=c("6", "12"), xend=c("6", "12"),
+  y=c(5, 4), yend=c(9.5, 6.5),
+  color=c(scales::muted("red"), scales::muted("blue")),
+  depth=factor("Deep", levels=c("Surface", "DCM", "Deep"), 
+               labels=c("Surface  (15 m)", "DCM  (110-120 m)", "Deep  (175 m)")),
+  lab_x=c("5", "10")
+)
 avg_targ_concs <- targ_peaks %>%
@@
   ungroup()
-arrow_df <- data.frame(
-  cyc_type=c("Anticyclone\ncenter", "Cyclone\ncenter"),
-  x=c("6", "12"), xend=c("6", "12"),
-  y=c(5, 4), yend=c(9.5, 6.5),
-  color=c(scales::muted("red"), scales::muted("blue")),
-  depth=factor("Deep", levels=c("Surface", "DCM", "Deep"), 
-               labels=c("Surface  (15m)", "DCM  (110-120m)", "Deep  (175m)")),
-  lab_x=c("5", "10")
-)
 barplot_targ_gp <- avg_targ_concs %>%
@@
   mutate(depth=factor(depth, levels=c("Surface", "DCM", "Deep"), 
-                      labels=c("Surface  (15m)", "DCM  (110-120m)", "Deep  (175m)"))) %>%
+                      labels=c("Surface  (15 m)", "DCM  (110-120 m)", "Deep  (175 m)"))) %>%
   mutate(station=factor(station, levels=as.character(4:14))) %>%
@@
   scale_y_continuous(expand = c(0.005, 0.1)) +
-  labs(x=NULL, y="Median nM\nmetabolite concentration", fill=NULL) +
+  labs(x=NULL, y="Median metabolite\n concentration (nM)", fill=NULL) +
   facet_grid(~depth, scales="free", space = "free") +
@@
 
-frac_pc_in_metab_df <- targ_peaks %>%
-  filter(str_detect(filename, "180821_Smp")) %>%
-  left_join(clean_stans %>% select(compound_name, formula, polarity)) %>%
-  mutate(n_C=as.numeric(str_extract(formula, "(?<=C)\\d+"))) %>%
-  mutate(nM_C=n_C*nM) %>%
-  left_join(filled_file_metadata, by = join_by(filename)) %>%
-  filter(samp_type=="Smp") %>%
-  group_by(filename) %>%
-  summarize(total_nM=sum(nM_C), PC_um=unique(PC_um)) %>%
-  mutate(frac_in_metab=total_nM/1000/PC_um) %>%
-  left_join(filled_file_metadata, by = join_by(filename, PC_um)) %>%
-  mutate(depth=str_extract(depth, "DCM|Deep|Surface")) %>%
-  mutate(depth=factor(depth, levels=c("Surface", "DCM", "Deep"))) %>%
-  mutate(station=factor(station, levels=c(as.character(4:14))))
+med_frac_df <- frac_pc_in_metab_df %>%
+  group_by(depth, station, sla_corr) %>%
+  summarize(frac_C_in_metab=mean(frac_C_in_metab)) %>%
+  ungroup()
 pc_frac_metab_gp <- frac_pc_in_metab_df %>%
-  ggplot(aes(x=station, y=frac_in_metab, color=sla_corr)) +
+  ggplot(aes(x=station, y=frac_C_in_metab)) +
   geom_hline(yintercept = 0) +
-  geom_boxplot(width = 0.9, linewidth=0.8) +
+  # geom_boxplot(aes(color=sla_corr), width = 0.9, linewidth=0.8) +
+  geom_point(color="black", size=2) +
+  geom_point(aes(fill=sla_corr), data=med_frac_df, size=4, pch=21, alpha=0.7) +
   scale_x_discrete(breaks=as.character(seq(4, 14, 2))) +
@@
   ) +
-  guides(color=guide_colorbar(direction = "horizontal", title.position = "top", 
+  guides(fill=guide_colorbar(direction = "horizontal", title.position = "top", 
                               title.hjust = 0.5, barheight = 0.2, reverse = TRUE)) +
-  labs(color="Corr. SLA (cm)", x="Station",
+  labs(fill="Corr. SLA (cm)", x="Station",
        y="Fraction of PC represented\nby known metabolites") +
@@
 
-plot(targ_gp_w_sla_frac)
-
 ggsave("targ_gp_w_sla_frac.png", plot = targ_gp_w_sla_frac, device = "png", 
@@
 
-*Figure 2: Upper row of stacked barplots shows known compounds measured across the eddy dipole and separated by depth.
-Bar height corresponds to median triplicate concentration for each metabolite, with the top 9 shown and the 44 other identified metabolites summed in grey.
+*Figure 3: Differences in known metabolite concentration across the pair of adjacent eddies in the MESO-SCOPE transect separated by depth.
+Top panels depict concentrations of known compounds measured, where bar height corresponds to median triplicate concentration for each metabolite with the top 9 shown and the 44 other identified metabolites summed in grey.
 TMAO = trimethylamine N-oxide, HO-Ile = hydroxyisoleucine, GBT = glycine betaine, DCM = deep chlorophyll maximum.
-Lower row of boxplots shows the fraction of total particulate carbon (PC) in the known metabolites across the eddy dipole, with the upper whisker extent, median bar, and lower whisker extent corresponding to the values measured in three biological triplicates.
+Lower panels depict the fraction of total particulate carbon in the known metabolites, with the mean value of three biological triplicates in color and the raw values in black behind.
 Colors correspond to corrected sea level anomaly (Corr.
-SLA) in centimeters, with dark red indicating anticyclonic (positive SLA) and dark blue indicating cyclonic (negative SLA) eddy state.*
+SLA), with dark red indicating anticyclonic (positive SLA) and dark blue indicating cyclonic (negative SLA) eddy state.*
 
@@
   summarise(nM=mean(nM)) %>%
-  mutate(nM=log10(nM)) %>%
   ggplot(aes(x=nM, y=depth, group=compound_name)) +
-  geom_path()
+  geom_path() +
+  scale_x_log10()
 avg_targ_concs %>%
@@
 
-We found that the overall concentration of known molecules was highest at the DCM of the cyclone.
-The most abundant quantified molecule in the eddy transect samples was trimethylamine N-oxide (TMAO), with an average concentration of 1.38 nM and a distinct DCM maximum (Figure 2).
-This pattern of a subsurface maximum at the DCM was mirrored by a majority of the known molecules (37/53).
-Most other abundant molecules decreased monotonically with depth, with highest values at the surface.
-
-```{r PC_vs_totnM regression values}
-no_outlier_frac_pc_in_metab_df <- frac_pc_in_metab_df %>%
-  filter(!(station=="11" & tripl=="A" & depth=="Surface")) %>%
-  filter(!(station=="6" & tripl=="B" & depth=="Surface")) %>%
-  filter(!(station=="7" & tripl=="C" & depth=="Surface")) %>%
-  filter(!(station=="11" & tripl=="B" & depth=="DCM"))
-
-# Overall trend between PC and total nM metabolites (all depths)
-broom::tidy(lm(total_nM~PC_um, data=no_outlier_frac_pc_in_metab_df))[2,]
-with(no_outlier_frac_pc_in_metab_df, cor(PC_um, total_nM))
-lme4::lmer(total_nM~PC_um + (1|depth), data=no_outlier_frac_pc_in_metab_df)
-
-# By-depth trend between PC and total nM metabolites
-no_outlier_frac_pc_in_metab_df %>%
-  group_by(depth) %>%
-  summarise(overall=broom::tidy(lm(total_nM~PC_um))[2,], 
-            cor_vals=cor(PC_um, total_nM))
-```
-
-The small-scale oceanographic variability introduced by mesoscale eddy features allowed us to also test the robustness of a model expecting a linear relationship between the total carbon contained within metabolites and bulk particulate carbon (PC) measurements.
-As expected, PC values showed a strong trend with summed metabolite concentrations ($\beta = 23.7\pm1.35$ (SE), Pearson's $r$ = .876, p-value < 0.001, Figure 2.5).
-This overall trend was largely driven by higher values in the surface than at depth, but when analyzed at each depth individually the general correspondence held.
-The DCM had the strongest trend ($\beta = 22.1\pm2.24$, $r= .689$, p-value < 0.001) while Surface ($\beta = 16.0\pm7.45$, $r = .375$, p-value = 0.04) and Deep ($\beta = 23.9\pm7.36$, $r = 0.503$, p-value = 0.003) values were weaker but still compelling relationships (Figure 2.5).
-However, this association required the removal of three samples in the surface and one in the DCM with anomalously high metabolite loads as well as two surface PC samples with anomalously high values from Stations 10 and 11 for which PC was interpolated by linear regression against beam attenuation.
-The full regression is reported in Supplemental Figure 3.
-
-```{r Fig 2.5 PC vs tot nM plot}
-pc_totalnm_plot <- no_outlier_frac_pc_in_metab_df %>%
-  ggplot(aes(x=PC_um, y=total_nM)) +
-  geom_point(aes(fill=depth), pch=21, size=3) +
-  geom_smooth(method="lm", color="white", lwd=2.5, se=FALSE) +
-  geom_smooth(method="lm", color="black", lwd=1.5) +
-  geom_smooth(aes(group=depth), method="lm", color="white", lwd=2.5, se = FALSE) +
-  geom_smooth(aes(color=depth), method="lm", lwd=1.5) +
-  theme_bw() +
-  theme(legend.position = "top") +
-  labs(color="Sampling depth",
-       fill="Sampling depth",
-       x="Particulate carbon (uM)", 
-       y="Total nM carbon in known metabolites") +
-  scale_color_manual(breaks=c("Surface", "DCM", "Deep"), 
-                     values = c("#F4BB23", "#028E34", "#0B505C"), aesthetics = c("fill", "color"))
-pc_totalnm_plot_bydepth <- no_outlier_frac_pc_in_metab_df %>%
-  ggplot() +
-  aes(x=PC_um, y=total_nM) +
-  geom_smooth(method="lm", color="black") +
-  geom_point(aes(fill=sla_corr), pch=21, color="black") +
-  facet_wrap(~depth, ncol=1, scales="free") +
-  labs(x="Particulate\ncarbon (uM)", y="Total nM carbon in known metabolites") +
-  scale_fill_gradient2(
-    low = scales::muted("blue"), high = scales::muted("red"),  mid = "grey90",
-    aesthetics = c("fill"), name="Corr.\nSLA\n(cm)"
-  ) +
-  theme_bw() +
-  theme(strip.text = element_text(face = "bold"))
-
-comb_pcnm_plot <- cowplot::plot_grid(pc_totalnm_plot, pc_totalnm_plot_bydepth, 
-                                     rel_widths = c(5, 3))
-
-ggsave("comb_pcnm_plot.png", plot = comb_pcnm_plot, device = "png", 
-       path = "figures", width = 6.875, height = 4.5, units = "in", dpi = 300, 
-       bg = "white")
-```
-
-![](figures/comb_pcnm_plot.png)
-
-*Figure 2.5: Type I linear regressions of particulate carbon against total nM carbon in known metabolites both as a whole and divided by depth.
-Points have been colored according to the SLA and have been placed on independent axes.
-Outlier metabolite points have been removed from the surface at stations 6, 10, and 11 as well as a DCM sample from station 12, while two outlier PC points have been interpolated from beam attenuation for surface values at stations 10 and 11.
-For the full regression, see Supplemental Figure 3.*
+The most abundant intracellular metabolite quantified in the eddy transect samples was trimethylamine N-oxide, with an average concentration of 1.38 nM and a distinct DCM maximum (Figure 2), though a few samples had concentrations an order of magnitude higher for unknown reasons (Supplemental figure 3).
+A majority of the known molecules (37/53) had a similar pattern with a subsurface maximum at the DCM, including glycine betaine, glutamate, and guanine.
+13 molecules obeyed a different pattern and decreased monotonically with depth, including the molecules hydroxyisoleucine, gonyol, and dimethylsulfoniopropionate.
+Two known molecules (arsenobetaine and O-acetylcarnitine) increased with depth, potentially representing particle degradation during export.
 
@@
 
-Known metabolites contributed between 1.5% and 5% of the carbon to the total PC load with higher values at the DCM than the surface or 175m samples (Figure 2, Supplemental Figure 3).
-DCM samples contained approximately 25% more carbon in the known metabolite pool (One-way ANOVA with outliers removed plus Tukey post-hoc, p<0.001, n=30-33) but there was no clear trend with SLA.
-
-```{r comparison to Angies PC values way higher than mine, eval=FALSE}
+```{r comparison to Angies PC values, eval=FALSE}
 angie_pcdata <- "../metadata/raw_data/Karl_HL2_WC_PCPN_Final_Data.xlsx" %>%
@@
   select(abs_depth=`Depth.(m)`, PC_um=`C.value.(µmol/L)`) %>%
-  # filter(depth<30) %>%
+  # filter(abs_depth<30) %>%
   mutate(cruise="HOE-Legacy")
@@
   distinct(abs_depth, PC_um) %>%
-  mutate(cruise="MESOSCOPE")
+  mutate(cruise="MESO-SCOPE")
 
@@
 
-One sample, taken from the surface at Station 11 (MS11C215m_A), was highly distinct from all others (including its paired triplicates) with several metabolites measuring two or more orders of magnitude larger in peak area and concentration both before and after normalization to internal standards.
+One sample, taken from 15 meters at Station 11, was highly distinct from all others (including its paired triplicates) with several metabolites measuring two or more orders of magnitude larger in peak area and concentration both before and after normalization to internal standards (Supplemental figure 3).
 Heuristic exploration of potentially related masses identified a total of 14 compounds highly abundant in this one sample alone.
-The only identified compounds with this trend were taurine and acetyltaurine, though three others with *m/z* 146.1176, 148.0968, and 206.1388 were discovered to be likely isomers of known compounds (acetylcholine/(3-carboxypropyl)trimethylammonium, hydroxyisoleucine, and dexpanthenol, respectively).
-One feature (*m/z* = 184.0637, RT = 7.0 minutes) was identified as choline sulfate using a second run of standards, while an additional mass feature at 168.0689 and RT = 8.5 was putatively flagged as taurine betaine.
-This mass was also one of the most highly enriched in the MS11C215m_A sample after taurine itself, with a peak area nearly 30x that of the next largest sample.
-No clear explanation emerged for why some compounds were hyperabundant in this surface sample, though one possible explanation was the *Trichodesmium* bloom that was also sampled at this time [@Dugenne2023].
+The only identified compounds with this trend were taurine and acetyltaurine, though three others with *m/z* ratios of 146.1176, 148.0968, and 206.1388 were discovered to be likely isomers of known compounds (acetylcholine, hydroxyisoleucine, and dexpanthenol, respectively).
+One feature (*m/z* = 184.0637, RT = 7.0 minutes) was identified as choline sulfate using a second run of standards, while an additional mass feature at 168.0689 and eluting at 8.5 minutes was putatively flagged as taurine betaine.
+This mass was also one of the most highly enriched in the sample after taurine itself, with a peak area nearly 30 times that of the next largest sample.
+No clear explanation emerged for why some compounds were hyperabundant in this surface sample, though one possibility was the *Trichodesmium* bloom that was also sampled at this time [@Dugenne2023].
 
@@
   left_join(stans) %>%
-  mutate(label=ifelse(is.na(cmpd_name), formula, paste0(formula, " (", cmpd_name, ")"))) %>%
-  mutate(casewhen_arg=paste0(cmd, "~'", label, "'"))
+  mutate(label=ifelse(is.na(cmpd_name), formula, paste0(formula, " (", cmpd_name, ")")))
 ms11_df <- c("C11H23NO6", "C11H23NO6", "C9H19NO5", "C8H17NO5", "C6H13NO6", "C8H17NO4") %>%
@@
   mutate(rtmed=round(rtmed, 1)) %>%
+  mutate(exact_mass=str_replace(formula, "N", "N1")) %>%
+  mutate(exact_mass=str_replace(exact_mass, "S", "S1")) %>%
+  separate(exact_mass, into = c("drop", "C", "H", "N", "O", "S"), sep = "[A-Z]") %>%
+  mutate(across(C:S, as.numeric)) %>%
+  mutate(S=ifelse(is.na(S), 0, 1)) %>%
+  mutate(calc_mass=C*12+H*1.007825+N*14.003074+O*15.994915+S*31.972072) %>%
+  mutate(calc_mass=ifelse(abs(calc_mass-mzmed)>1, calc_mass+1.007276, calc_mass)) %>%
+  mutate(delta_ppm=ifelse(is.na(feature), 0, round((mzmed-calc_mass)*1e6/mzmed, 2))) %>%
   mutate(compound_name=ifelse(is.na(compound_name), "", compound_name)) %>%
-  mutate(feature=ifelse(is.na(feature), "", feature)) %>%
-  dplyr::rename(`Median m/z`=mzmed, `Median RT`=rtmed, `Putative formula`=formula,
-                `Compound name`=compound_name, `Feature ID`=feature)
-knitr::kable(ms11_df)
-```
-
-| Median m/z| Median RT|Putative formula |Compound name    |Feature ID  |
-|----------:|---------:|:----------------|:----------------|:-----------|
-|   126.0221|      11.1|C2H8NO3S         |Taurine          |MS_POS_0677 |
-|   146.1176|       8.5|C7H16NO2         |                 |MS_POS_1050 |
-|   148.0968|       8.5|C6H14NO3         |                 |MS_POS_1080 |
-|   168.0317|       4.9|C4H10NO4S        |N-Acetyltaurine  |MS_NEG_1123 |
-|   168.0689|       8.5|C5H14NO3S        |Taurine betaine? |MS_POS_1340 |
-|   176.1282|       8.1|C8H18NO3         |                 |MS_POS_1419 |
-|   184.0637|       7.0|C5H14NO4S        |Choline sulfate  |            |
-|   192.1230|       9.2|C8H17NO4         |                 |            |
-|   196.0816|       5.2|C6H13NO6         |                 |            |
-|   198.0794|       5.0|C6H16NO4S        |                 |            |
-|   206.1388|       7.1|C9H20NO4         |                 |MS_POS_1683 |
-|   208.1180|      10.2|C8H17NO5         |                 |            |
-|   222.1336|       8.5|C9H19NO5         |                 |            |
-|   266.1598|       8.5|C11H23NO6        |                 |            |
-|   266.1598|      10.0|C11H23NO6        |                 |            |
-
-*Table 1: Detected mass features that were enriched in the 15m sample from Station 11 (MS11C215m_A).
+  select(-feature) %>%
+  select(`Median m/z`=mzmed, `Median RT (min)`=rtmed, `Putative formula`=formula,
+         `Delta mz (ppm)`=delta_ppm, `Compound name`=compound_name)
+
+# knitr::kable(ms11_df)
+```
+
+| Median m/z| Median RT (min)|Putative formula | Delta mz (ppm)|Compound name    |
+|----------:|---------------:|:----------------|--------------:|:----------------|
+|   126.0221|            11.1|C2H8NO3S         |          -3.35|Taurine          |
+|   146.1176|             8.5|C7H16NO2         |          -3.63|                 |
+|   148.0968|             8.5|C6H14NO3         |          -3.88|                 |
+|   168.0317|             4.9|C4H10NO4S        |          -7.87|N-Acetyltaurine  |
+|   168.0689|             8.5|C5H14NO3S        |          -3.33|Taurine betaine? |
+|   176.1282|             8.1|C8H18NO3         |          -2.88|                 |
+|   184.0637|             7.0|C5H14NO4S        |           0.00|Choline sulfate  |
+|   192.1230|             9.2|C8H17NO4         |           0.00|                 |
+|   196.0816|             5.2|C6H13NO6         |           0.00|                 |
+|   198.0794|             5.0|C6H16NO4S        |           0.00|                 |
+|   206.1388|             7.1|C9H20NO4         |          -1.93|                 |
+|   208.1180|            10.2|C8H17NO5         |           0.00|                 |
+|   222.1336|             8.5|C9H19NO5         |           0.00|                 |
+|   266.1598|             8.5|C11H23NO6        |           0.00|                 |
+|   266.1598|            10.0|C11H23NO6        |           0.00|                 |
+
+*Table 1: Detected mass features that were enriched in the 15 meter surface sample from MESO-SCOPE Station 11.
 RT = retention time in minutes.
@@
 to denote that this identification is putative.
-Feature ID is a disambiguating key produced by XCMS processing.*
+Mass difference between the median m/z and putative formula is reported in parts-per-million.*
 
@@
 
-We explored the possibility that a nearby *Trichodesmium* bloom was responsible for the unusual composition of this one sample by analyzing additional samples taken directly from the bloom to see whether compounds abundant in hand-picked *Trichodesmium* colonies were also abundant the Station 11 samples.
+We explored the possibility that a nearby *Trichodesmium* bloom was responsible for the unusual composition of this one sample by analyzing additional samples taken directly from the bloom to see whether compounds abundant in hand-picked *Trichodesmium* colonies were also abundant in the Station 11 sample.
 After searching for the 14 mass features above in the environmental sample, 7 had good-quality peaks when viewed in the bloom samples as extracted ion chromatograms, though only 4 of these matched closely in retention time.
@@
 erythraeum* (strain IMS101) osmolyte [@Pade2016].
-Although this compound was not in our standard library due to a lack of authentic standards, we identified three mass features in the 162.1125 *m/z* window with retention times of 7.1, 8.4, and 9.9 minutes.
+Although this compound was not run in water alongside the sample to confirm a retention time match due to a lack of authentic standards, we identified three mass features in the 162.1125 *m/z* window with retention times of 7.1, 8.4, and 9.9 minutes.
 The feature at RT 9.9 matched our carnitine standard's retention time, leaving the two earlier peaks with putative identifications of threonine betaine at 7.1 and homoserine betaine at 8.4 given the elution order of threonine before homoserine in our authentic standards.
 This order is the reverse of that in @Pade2016, which showed homoserine betaine eluting prior to carnitine.
-This elution order is expected given our use of a normal phse HILIC column and their use of a reversed phase Hypercarb column.
-They did not report an additional peak detected at this mass, leaving our putative identification of the threonine betaine as low-confidence.
-None of the three 162.1125 features showed the same degree of enrichment in MS11C215m_A as taurine or the other 13 compounds detected, though MS11C215m_A was the largest, second-largest, and third-largest peak for carnitine, threonine betaine, and homoserine betaine, respectively.
+This elution order is expected given our use of a normal phase HILIC column and their use of a reversed phase Hypercarb column.
+They did not report an additional peak detected at this mass, leaving our putative identification of the homoserine and threonine betaines as low-confidence.
+None of the three 162.1125 features showed the same degree of enrichment in the anomalous sample as taurine or the other 13 compounds listed in Table 1, though the anomalous sample contained the largest, second-largest, and third-largest peak for carnitine, threonine betaine, and homoserine betaine, respectively.
+These results are fairly inconclusive but we are inclined to believe that a *Trichodesmium* colony is not likely to be the cause of the inflated values in this particular sample.
 
@@
 
-## Compositional shifts in metabolome as well as bulk biomass
+## Particulate organic matter compositional shifts across eddy transect
 
 In addition to absolute shifts in metabolite concentration due to biomass differences, we also explored compositional shifts in the metabolomes by assessing the fraction of the total metabolite pool contributed by each compound across the eddy transect.
+This allowed us to isolate the signal due to shifts in community composition and organismal response from those introduced by net growth of the community as a whole.
 
@@
 
-In the 15m surface samples, only seven compounds had significant changes in relative peak area across the eddy transect, two of which were known: 5-oxoproline, which decreased from 0.2-0.3% of the total peak area at the center of the anticyclone to 0.05-0.1% in the cyclone; and a combined peak of sarcosine & beta-alanine which increased from 0.01% to 0.02-0.03%.
+In the 15 meter surface samples, only seven compounds had significant changes in relative peak area across the eddy transect, two of which were known: 5-oxoproline, which decreased from 0.2-0.3% of the total peak area at the center of the anticyclone to 0.05-0.1% in the cyclone; and a combined peak of sarcosine and beta-alanine which increased from 0.01% to 0.02-0.03%.
 The largest shift significant at an α = 0.05 level was an unknown mass feature with *m/z* = 189.12338 and a retention time around 8.6 minutes, the relative contribution of which increased from approximately 0.2-0.3% in the cyclone to ~1.5% in the anticyclone.
 
-At the DCM, twenty-two compounds changed significantly in relative peak area across the eddy dipole.
+At the DCM, twenty-two compounds changed significantly in relative peak area across the transect.
 Seven of these were known metabolites, of which trigonelline and homarine were most enriched in the cyclone while arsenobetaine was the only metabolite with a larger fraction of the total peak area in the anticyclone.
 Homarine showed a very large and highly significant shift, representing about 2.5% of the total peak area in the anticyclone but 7-8% in the cyclone center.
-Trigonelline, an isomer structurally very similar but biologically entirely distinct, had a surprisingly strong correlation with homarine (Pearson's $r=.855$) with peak areas consistently around one third of homarine's but large enough to have the second-largest shift in relative peak area of the significantly different compounds.
+Trigonelline (N-methyl niacin), an isomer structurally very similar to homarine but biologically distinct, had a surprisingly strong correlation with homarine (Pearson's $r=.855$).
+Its peak areas were consistently around one third of homarine's but still large enough to have the second-largest shift in relative peak area of the significantly different compounds.
 Similarly, arsenobetaine represented about 0.1% of the peak area in the cyclone and 0.5-0.6% of the total peak area in the anticyclone.
 The most significantly different compounds at the DCM in each direction, however, were both unknowns.
-The lowest p-value ($6.4*10^{-6}$ after FDR correction) in the DCM data was a mass feature with *m/z* of 173.09211, a retention time also around 8.6 minutes, and enrichment in the anticyclone with a putative chemical formula of [M+H] = C$_7$H$_{13}$N$_2$O$_3$, possibly glycylproline or prolylglycine.
-The next-lowest (p = $2.5*10^{-4}$ after FDR correction) was enriched in the cyclone and had an *m/z* of 275.07119 and a retention time of 11.4 minutes with a putative formula of C$_{18}$H$_{11}$O$_3$.
+The lowest p-value ($6.4*10^{-6}$ after FDR correction) in the DCM data was a mass feature with an *m/z* of 173.09211, a retention time also around 8.6 minutes, and enrichment in the anticyclone with a putative chemical formula of [M+H] = C$_7$H$_{13}$N$_2$O$_3$, possibly glycylproline or prolylglycine.
+The next-lowest (p = $2.5*10^{-4}$ after FDR correction) was enriched in the cyclone and had an *m/z* of 275.0712 and a retention time of 11.4 minutes with a putative formula of C$_{18}$H$_{11}$O$_3$.
 
@@
 Both O-acetylcarnitine and betonicine effectively quintupled their contribution to total peak area, shifting from approximately 0.1% to ~0.5% over the transect.
-Acetylcholine was the only known compound more relatively abundant in the cyclone than the anticyclone along with one unknown of *m/z* 131.034.
-The unknown at *m/z* 173.09211 with the strongest trend at the DCM again had the strongest trend in the 175 meter samples, here with an FDR-corrected p-value of $2.0*10^{-7}$.
+Acetylcholine was the only known compound more relatively abundant in the cyclone than the anticyclone along with one unknown of *m/z* 131.0340.
+The unknown at *m/z* 173.0921 with the strongest trend at the DCM again had the strongest trend in the 175 meter samples, here with an FDR-corrected p-value of $2.0*10^{-7}$.
 
@@
 
-## Eddy center DCM samples show differences between depths and eddy polarities
+## High vertical resolution sampling near eddy center DCM reveals distinct metabolome between cyclone and anticyclone
 
@@
   mutate(pca_output=map(wide_peaks, prcomp)) %>%
-  mutate(nmds_stress=paste("s =", round(nmds_stress, digits = 3)))
+  mutate(nmds_stress=paste("NMDS stress =", round(nmds_stress, digits = 3)))
   # mutate(perm_r2=round(perm_r2, digits = 2)) %>%
@@
 
-To provide context for the sea level anomaly responses detected above, we additionally performed high-resolution depth sampling around the DCM at the center of both eddy poles.
-Here, we found that the ~40 meter depth range and the transition between the eddies explained similar amounts of variation in the data (R$^2_{depth}=0.253$, R$^2_{SLA}=0.238$) and were both highly significant factors (permutational p-values << 0.001).
-Additionally, both sample depth and SLA were correlated with the first principal component of the metabolite matrix (r$_{depth}=.766$, r$_{SLA}=.684$, fraction of variance explained by PC$_1$ = 30.8%) (Figure 3).
-A 2D NMDS plot of the samples showed clear separation between the eddies, with cyclonic samples largely on the left side of Figure 3 while anticyclonic samples fell almost entirely to the right.
-Notably, the cyclonic samples show a much larger spread than the anticyclonic ones, indicating larger sample variability at the surface than the DCM.
-A depth gradient is visible along MDS 2, with the deepest samples generally at the top right of the plot and the shallowest ones closer to the bottom and left.
+Given the strength of the metabolomic response to SLA at the DCM, we additionally performed high-resolution depth sampling around the DCM at the center of both eddy poles.
+Here, we found that the ~40 meter depth range and the transition between the eddies explained similar amounts of variation in the data (R$^2_{depth}=0.253$, R$^2_{SLA}=0.238$) and were both highly significant factors (permutational p-values << 0.001, Figure 4).
+Additionally, both sample depth and SLA were correlated with the first principal component of the metabolite matrix ($r_{depth}=.766$, $r_{SLA}=.684$, fraction of variance explained by PC1 = 30.8%).
+The cyclonic samples in particular show a much larger spread than the anticyclonic ones, indicating larger sample variability in the cyclone DCM relative to the anticyclone.
+A depth gradient is visible along NMDS 2, with the deepest samples generally at the top right of the plot and the shallowest ones closer to the bottom (Figure 4).
 
@@
                       labels=c("-20m", "-10m", "DCM", "+10m", "+20m"))) %>%
-  ggplot(aes(label=tripl)) +
+  mutate(fill_col=case_when(
+    depth=="-20m" & sla_class=="Cyclone"~"#c4c4e0",
+    depth=="-10m" & sla_class=="Cyclone"~"#8989c1",
+    depth=="DCM" & sla_class=="Cyclone"~"#6161ad",
+    depth=="+10m" & sla_class=="Cyclone"~"#343489",
+    depth=="+20m" & sla_class=="Cyclone"~"#23235b",
+    depth=="-20m" & sla_class=="Anticyclone"~"#dabdbd",
+    depth=="-10m" & sla_class=="Anticyclone"~"#b57c7c",
+    depth=="DCM" & sla_class=="Anticyclone"~"#9c5050",
+    depth=="+10m" & sla_class=="Anticyclone"~"#762020",
+    depth=="+20m" & sla_class=="Anticyclone"~"#4f1616"
+  )) %>%
+  ggplot() +
   geom_vline(xintercept = 0) +
   geom_hline(yintercept = 0) +
-  geom_point(aes(x=MDS1, y=MDS2, color=sla_class), size=5, pch=19) +
-  scale_color_manual(
-    values=c(scales::muted("blue"), scales::muted("red")),
-    breaks=c("Cyclone", "Anticyclone"),
-  ) +
+  geom_point(aes(x=MDS1, y=MDS2, fill=fill_col, shape=sla_class, color=depth), size=6) +
+  scale_fill_identity() +
+  scale_color_manual(values = rep("white", 5)) +
+  scale_shape_manual(breaks=c("Cyclone", "Anticyclone"), values=c(25, 24)) +
   annotate("label", x=-200, y=100, label="Cyclone", color=scales::muted("blue"),
@@
            label.r = unit(0, "in"), size=6) +
-  annotate("label", x=-Inf, y=Inf, label=multivar_analysis_MC$nmds_stress,
-           hjust=0, vjust=1, label.r = unit(0, "in")) +
-  geom_point(aes(x=MDS1, y=MDS2, fill=depth, alpha=depth), size=5, pch=21, color="black") +
-  scale_fill_manual(
-    values = c("white", "white", "white", "black", "black"),
-    breaks = c("-20m", "-10m", "DCM", "+10m", "+20m")
-  ) +
-  scale_alpha_manual(
-    values = c(0.7, 0.4, 0.2, 0.1, 0.4),
-    breaks = c("-20m", "-10m", "DCM", "+10m", "+20m")
-  ) +
-  coord_fixed() +
-  guides(alpha="none", color="none", fill=guide_legend(
-    title = "Depth", title.position = "top", title.hjust = 0.5, direction = "horizontal",
-    override.aes = list(fill=c("grey90", "grey80", "grey60", "grey40", "grey20"))
+  guides(alpha="none", fill="none", shape="none", color=guide_legend(
+    title = "Depth:    ", title.position = "left", title.hjust = 0.5, direction = "horizontal",
+    keywidth=0.2, keyheight=0.1,
+    override.aes = list(fill=c("grey90", "grey80", "grey60", "grey40", "grey20"), 
+                        pch=21, color="black")
   )) +
+  labs(x="NMDS axis 1", y="NMDS axis 2") +
   theme_bw() +
-  theme(legend.position = "bottom")
-MC_nmds_gp
+  theme(legend.position = "top", legend.text = element_text(margin = margin(r = 15, l=2, unit = "pt"))) +
+  annotate("label", x=-Inf, y=Inf, hjust=0, vjust=1, label.r = unit(0, "in"),
+           label=expression(atop(paste("R"[SLA]^2, "=.238"), paste("R"[depth]^2, "=.253")))) +
+  annotate("label", x=-Inf, y=-Inf, label=multivar_analysis_MC$nmds_stress, 
+           hjust=0, vjust=-2, label.r = unit(0, "in"))
+
 ggsave("MC_nmds_gp.png", plot = MC_nmds_gp, device = "png", 
@@
 
-*Figure 3: NMDS plot of high-resolution depth sampling around the deep chlorophyll maximum (DCM, ~110 meters) at the eddy center of each polarity.
-Points are colored by the eddy from which the associated sample was collected and shaded by the depth.
-Visible separation between the cyclonic and anticyclonic eddy is evident along the first NMDS axis.*
+*Figure 4: NMDS plot of high-resolution depth sampling around the deep chlorophyll maximum (DCM, ~115 meters) at the two eddy centers during MESO-SCOPE.
+Red upward-pointed triangles are from the anticyclone and blue downward-pointed ones are from the cyclone.
+Shading intensity reflects the depth above or below the DCM.
+PERMANOVA estimates of the variance explained by depth and sea level anomaly (SLA) are noted in the upper left corner and the NMDS stress value is reported in the bottom left.*
 
@@
 
-To characterize the observed differences in multivariate space we used k-means clustering as an unsupervised way to identify dominant trends within the data (Figure 4).
-We found that a majority of the metabolites (Clusters 1 and 2, 63% of the total) fell into clusters with larger peak areas in the cyclone, as expected, while also detecting 32 metabolites that clustered such that the mean metabolite was enriched in the anticyclone (Cluster 4, Figure 4).
-Cluster 2 notably had a distinct decrease in relative peak area with depth, while the other clusters had much less clear depth trends.
-Cluster 1 also showed a DCM maximum for the samples from the anticyclone, correlating well with flow cytometry counts of picoeukaryotes (r=.851) that was not present in the cyclone (r=.013).
+To characterize the observed differences in multivariate space we used k-means clustering as an unsupervised way to identify dominant trends within the data (Figure 5).
+We found that a majority of the metabolites (Clusters 1 and 2, 63% of the total) fell into clusters with larger peak areas in the cyclone, as expected, while also detecting 32 metabolites that clustered such that the mean metabolite was enriched in the anticyclone (Cluster 4, Figure 5).
+Cluster 2 had a distinct decrease in relative peak area with depth, while the other clusters had much less clear depth trends.
+Cluster 1 also showed a DCM maximum for the samples from the anticyclone, correlating well with flow cytometry counts of picoeukaryotes (Pearson's $r=.851$) that was not present in the cyclone ($r=.013$).
 
@@
   summarise(mean_rank=mean(norm_area), iqr_value=sd(norm_area), .groups = "drop")
+
+lab_df <- tribble(
+  ~xlab, ~ylab, ~sla_class, ~clusterlab,
+  0, 5.5, "Anticyclone", "Cluster #1: n=73",
+  0.8, 1, "Cyclone", "Cluster #1: n=73"
+)
+
 library(ggh4x)
@@
   geom_point(aes(x=mean_rank), size=3) +
+  geom_label(aes(x=xlab, y=ylab, label=sla_class, color=sla_class), 
+             data=lab_df, label.r = unit(0, "in")) +
   ggh4x::facet_wrap2(~clusterlab, nrow = 1, strip = strip_theme) +
@@
 ```{r eddy center volcano plot}
+label_df <- pairwise_tests %>%
+  left_join(untarg_feats) %>%
+  filter(compound_name%in%c("Isethionic acid", "Trigonelline", "Arsenobetaine",
+                            "Homarine", "Dimethylsulfonioacetate", "Dimethylsulfoniopropionate")) %>%
+  left_join(fold_change_df, by = join_by(feature)) %>%
+  mutate(xspot = log2(fold_change), yspot=-log10(p.value))
+
 volcanoplot_gp <- pairwise_tests %>%
@@
   geom_hline(yintercept = 1.3, lty=2) +
+  annotate("segment", x=label_df$xspot[5], y=6.8, xend=label_df$xspot[5], yend=label_df$yspot[5], 
+           # arrow=arrow(type="closed", angle = 25, length = unit(0.05, "npc")), 
+           lwd=0.7, color="grey50") +
+  annotate("label", x=-2, y=6.8, label="Arsenobetaine", hjust=0, label.r=unit(0, "in")) +
+  annotate("segment", x=1.5, y=6.8, xend=label_df$xspot[3], yend=label_df$yspot[3], 
+           lwd=0.7, color="grey50") +
+  annotate("label", x=1.5, y=6.8, label="Homarine", hjust=1, label.r=unit(0, "in")) +
+  annotate("segment", x=2, y=6.8, xend=label_df$xspot[2], yend=label_df$yspot[2], 
+           lwd=0.7, color="grey50") +
+  annotate("label", x=2, y=6.8, label="DMSP", hjust=0, label.r=unit(0, "in")) +
+  annotate("segment", x=Inf, y=label_df$yspot[1], xend=label_df$xspot[1], yend=label_df$yspot[1], 
+           lwd=0.7, color="grey50") +
+  annotate("label", x=Inf, y=label_df$yspot[1], label="DMS-Ac", hjust=1, vjust=0.3, label.r=unit(0, "in")) +
+  annotate("segment", x=3.5, y=5, xend=1.9, yend=5, 
+           lwd=0.7, color="grey50") +
+  annotate("segment", x=1.9, y=5, xend=label_df$xspot[4], yend=label_df$yspot[4], 
+           lwd=0.7, color="grey50") +
+  annotate("label", x=Inf, y=5, label="Trigonelline", hjust=1, label.r=unit(0, "in")) +
+  annotate("segment", x=3.6, y=3, xend=2.2, yend=3, 
+           lwd=0.7, color="grey50") +
+  annotate("segment", x=2.2, y=3, xend=label_df$xspot[6], yend=label_df$yspot[6], 
+           lwd=0.7, color="grey50") +
+  annotate("label", x=Inf, y=3, label="Isethionate", hjust=1, label.r=unit(0, "in")) +
   geom_point(size=2, stroke=1) +
-  scale_y_continuous(breaks = 0:6, labels = paste0("10<sup>-", 0:6, "</sup>")) +
+  scale_x_continuous(breaks=c(-4:4)) +
+  scale_y_continuous(breaks = 0:6, labels = paste0("10<sup>-", 0:6, "</sup>"), expand = expansion(c(0.01, 0.05))) +
   scale_shape_manual(values=c(4, 16), breaks = c("Unknown", "Known")) +
@@
         legend.position = c(0, 1), legend.justification = c(0, 1), 
-        legend.background = element_rect(fill="#FFFFFF88", color=NA),
+        legend.background = element_rect(fill="#FFFFFF88", color="black"),
         legend.spacing.y=unit(0, "in"),
@@
   kclust_gp, volcanoplot_gp, 
-  ncol=1, heights = c(2, 1.5))
+  ncol=1, heights = c(1.5, 2), draw = FALSE)
 ggsave("kclust_volcano_gp.png", plot = kclust_volcano_gp, device = "png", 
-       path = "figures", width = 6.875, height = 4, units = "in", dpi = 300,
+       path = "figures", width = 6.875, height = 5, units = "in", dpi = 300,
        bg="white")
@@
 
-*Figure 4: Distribution of metabolites in the high-resolution depth samples from the centers of each eddy.
+*Figure 5: Distribution of metabolites in the high-resolution depth samples from the centers of each MESO-SCOPE eddy.
 The upper row of plots shows k-means clusters where points denote the average z-scored peak area for both known and unknown metabolites across the samples and are colored by the eddy from which they were taken.
@@
 The lower plot shows the individual known and unknown metabolites where points correspond to the FDR-corrected p-value estimated by the nonparametric Mann-Whitney U test and the log$_2$ fold-change calculated with the average peak area in the cyclone divided by the average peak area in the anticyclone.
-Colors have been assigned using the k-means clusters and shapes have been assigned based on the status of the mass feature as either a known metabolite that was matched to an authentic standard or an unidentified metabolite.*
+Colors have been assigned using the k-means clusters and shapes have been assigned based on the status of the mass feature as either a known metabolite that was matched to an authentic standard or an unidentified metabolite.
+The dashed line across the figure represents the 0.05 level of significance as a visual cue for metabolites above which the differences between the eddies are unlikely to be due to chance.
+DMSP = dimethylsulfoniopropionate, DMS-Ac = dimethylsulfonioacetate*
 
@@
 
-We also investigated the response of individual metabolites in these high-resolution depth profile samples, expecting to find the strongest signal in this data given the extreme SLA values sampled and the large number of compounds with an SLA separation in the k-means analysis.
-This expectation was substantiated by a majority of the metabolites differing significantly between the two eddies (121/228, α = 0.05).
+The majority of the individual metabolites in these high-resolution depth profiles differed significantly between the two eddies (121/228, α = 0.05).
 Of those, 15 were enriched in the anticyclone and 106 were more abundant in the cyclone.
-As expected, all compounds enriched in the anticyclone were part of Cluster 4 and all those enriched in the cyclone belonged to either cluster 1 or 2 (Figure 4).
-
-Many of the known metabolites enriched in the cyclone are expected given the overall higher biomass, particularly as many function as osmolytes in the cell.
-Metabolites enriched in excess of biomass were taurine and isethionic acid, the reduced sulfur osmolytes dimethylsulfoniopropionate (DMSP) and dimethylsulfonioacetate (DMS-Ac), and the isomers homarine and trigonelline.
-A few metabolites more abundant in the anticyclone were also given putative identifications based on RT and *m/z* matching with internal standards later added.
-Of these, the putative arsenobetaine was the most significantly different among these with a peak area at the anticyclone DCM nearly quadruple that of the cyclone.
-Of note, the 173.09211 *m/z* mass feature noted above was among the most significantly different between the two eddies, along with two isomers at an *m/z* of 170.11756 and retention times around 8-9 minutes that increased by a factor of 1.5 in the anticyclone (putative formula C$_7$H$_{14}$N$_4$O).
+As expected, all compounds enriched in the anticyclone were part of Cluster 4 and all those enriched in the cyclone belonged to either Cluster 1 or 2 (Figure 5).
+
+Many of the known metabolites enriched in the cyclone function as osmolytes in the cell and might be enriched due to the increased eukaryotic phytoplankton biomass.
+However, some metabolites such as isethionate, the reduced sulfur osmolytes dimethylsulfoniopropionate (DMSP) and dimethylsulfonioacetate (DMS-Ac), and the isomers homarine and trigonelline, were enriched in excess of biomass (Figure 5).
+A few metabolites more abundant in the anticyclone were also given putative identifications based on RT and *m/z* matching with internal standards run at a later time on the mass spectrometer.
+Of these, the putative arsenobetaine was the most significantly different among these with a peak area at the anticyclone DCM nearly quadruple that of the cyclone (Figure 5).
+Of note, the 173.0921 *m/z* mass feature noted above as strongly enriched in the anticyclone was among the most significantly different between the two eddies, along with two isomers at an *m/z* of 170.1176 and retention times around 8-9 minutes that increased by a factor of 1.5 in the anticyclone (putative formula C$_7$H$_{14}$N$_4$O).
 
@@
 
-## Falkor data reveals a different community and response to SLA
-
-Given the strong signals detected in the samples from the MESOSCOPE cruise both across the eddy transect as well as in the high-resolution DCM sampling, we expected to find some of the same results in a new dataset.
-Data was collected during a 2018 cruise on the R/V *Falkor* that again targeted both a cyclonic and anticyclonic eddy in the North Pacific Subtropical Gyre near Station ALOHA.
-
-```{r figure falkor nmds and permanova}
+## Hawaiian Eddy Experiment data reveals a different community and response to mesoscale eddies
+
+Given the strong signals detected in the samples from the MESO-SCOPE cruise both across the eddy transect as well as in the high-resolution DCM sampling, we expected to find some of the same results in an analogous dataset.
+Samples were collected during a 2018 cruise on the R/V *Falkor* that again targeted both a cyclonic and anticyclonic eddy in the North Pacific Subtropical Gyre near Station ALOHA as part of the Hawaiian Eddy Experiment (HEE).
+
+```{r falkor nmds and permanova}
 set.seed(123)
@@
   permutations = 999, method = "manhattan")
-
-fk_nmdsplot <- nmds_fk$points %>%
+```
+
+```{r figure falkor nmds}
+rot_mat <- matrix(c(sqrt(2)/2, sqrt(2)/2, -sqrt(2)/2, sqrt(2)/2), ncol=2)
+
+fk_nmdsplot <- nmds_fk$points%*%rot_mat %>%
   as.data.frame() %>%
+  setNames(c("MDS1", "MDS2")) %>%
   rownames_to_column("filename") %>%
-  left_join(filled_file_metadata %>% distinct(filename, sla_corr, depth)) %>%
+  left_join(filled_file_metadata %>% distinct(filename, sla_class, depth)) %>%
+  mutate(MDS1=MDS1*-1) %>%
   mutate(MDS2=MDS2*-1) %>%
+  mutate(fill_col=case_when(
+    depth=="Surface" & sla_class=="Cyclone"~"25 m Cyclone",
+    depth=="DCM" & sla_class=="Cyclone"~"25 m Anticyclone",
+    depth=="Surface" & sla_class=="Anticyclone"~"DCM Cyclone",
+    depth=="DCM" & sla_class=="Anticyclone"~"DCM Anticyclone"
+  )) %>%
   ggplot(aes(label=filename)) +
@@
   geom_hline(yintercept = 0) +
-  geom_point(aes(x=MDS1, y=MDS2, color=sla_corr, shape=depth), size=4, stroke=2) +
-  scale_color_gradient2(
-    low = scales::muted("blue"), high = scales::muted("red"),  
-    mid = "grey90", aesthetics = "color"
-  ) +
-  scale_shape_manual(values = c(2, 6), breaks = c("Surface", "DCM")) +
-  labs(color="Corr.\nSLA\n(cm)", shape="Depth") +
+  geom_point(aes(x=MDS1, y=MDS2, fill=fill_col, shape=sla_class), size=4, color="white") +
+  scale_fill_manual(breaks=c("25 m Cyclone", "25 m Anticyclone", "DCM Cyclone", "DCM Anticyclone"),
+                    values=c("#8989c1", "#343489", "#b57c7c", "#762020")) +
+  scale_shape_manual(breaks=c("Cyclone", "Anticyclone"), values=c(25, 24)) +
+  labs(color="Corr.\nSLA\n(cm)", shape="Depth", x="NMDS axis 1", y="NMDS axis 2") +
   theme_bw() +
+  annotate("label", x=-Inf, y=-Inf, hjust=0, vjust=0, label.r = unit(0, "in"),
+           label=paste("Stress =", round(nmds_fk$stress, digits = 3))) +
+  annotate("label", x=-Inf, y=Inf, hjust=0, vjust=1, label.r = unit(0, "in"),
+           label=expression(atop(paste("R"[sla]^2, "=.090, p=0.017"),
+                                 paste("R"[depth]^2, "=.251, p<0.001")))) +
+  guides(fill=guide_legend(title = NULL, override.aes = list(
+    fill=c("#8989c1", "#b57c7c", "#343489", "#762020"), color="white",
+    shape=c(25, 24, 25, 24)
+  )), shape="none") +
   coord_fixed()
-
-fk_nmds_stress <-  paste("Stress =", round(nmds_fk$stress, digits = 3))
-fk_perm_label <- expression(atop(paste("R"[sla]^2, "=.090, p=0.017"), 
-                                 paste("R"[depth]^2, "=.251, p<0.001")))
-fk_nmdsplot <- fk_nmdsplot +
-  annotate("label", x=-Inf, y=-Inf, label=fk_nmds_stress, hjust=0, vjust=0, 
-           label.r = unit(0, "in")) +
-  annotate("label", x=-Inf, y=Inf, label=fk_perm_label, hjust=0, vjust=1, 
-           label.r = unit(0, "in"))
 
@@
 
-The Falkor data was characterized by high inter-replicate variability relative to the samples from the MESOSCOPE cruise, with surface samples in particular highly variable in multivariate space (Figure 5).
-Despite this, we still saw the importance of sea level anomaly (SLA) as a significant explanatory factor in the dataset (R$^2_{SLA} = 0.090$, p-value = 0.025), after the dominant depth differences between the surface (25m) and DCM (110-120m) (R$^2_{depth} = 0.251$, p-value << 0.001).
+The HEE data was characterized by high inter-replicate variability relative to the samples from the MESO-SCOPE cruise, with 25 meter samples in particular highly variable in multivariate space (Figure 6).
+Despite this, we still saw the importance of sea level anomaly as a significant explanatory factor in the dataset (R$^2_{SLA} = 0.090$, p-value = 0.017) as well as the larger depth differences between the surface (25 meters) and DCM (110 - 120 meters) (R$^2_{depth} = 0.251$, p-value << 0.001).
 
@@
 
-*Figure 5: Non-metric multidimensional scaling (NMDS) plot from the Falkor cruise data, in which points correspond to individual samples and have been colored by their corrected sea level height anomaly and shaped by the depth from which they were collected.
-NMDS stress values (s) have been reported in the bottom left corner, while PERMANOVA R$^2$ and p-values are reported in the top left.
-Surface samples are clearly distinct from the DCM samples and an SLA signal is visible at the surface.*
+*Figure 6: Non-metric multidimensional scaling (NMDS) plot from the Hawaiian Eddy Experiment cruise data, in which points correspond to individual samples and have been colored and shaped by their source eddy status and shaded by the depth from which they were collected.
+The NMDS stress value has been reported in the bottom left corner, while PERMANOVA R$^2$ and p-values are reported in the top left.
+Samples from 25 meters deep are visibly distinct from the deep chlorophyll maximum (DCM) samples and an SLA signal is visible in the 25 meter samples only.*
 
@@
 
-When analyzed separately as distinct depths, the surface samples had higher R$^2$ values (R$^2$=.26) and a lower p-value (0.013), while the DCM samples were no longer likely to be distinct between the two eddies (p-value = 0.171) (Supplemental figure 3).
-
-```{r falkor concentrations comparison to MESOSCOPE, eval=FALSE}
+When analyzed separately as distinct depths, the 25 meter samples had higher variances explained by eddy (R$^2$=.26) and a lower p-value (0.013), while the DCM samples were no longer likely to be distinct between the two eddies (p-value = 0.171) (Figure 6, Supplemental figure 4).
+
+```{r falkor concentrations comparison to MESO-SCOPE, eval=FALSE}
 fk_targ <- targ_peaks %>% 
@@
 
-Overall, we detected several compounds that were distinctly more or less abundant in the Falkor dataset than the MESOSCOPE one.
-Compounds more abundant in Falkor were the amino acids aspartate, threonine, and tyrosine as well as the components of the urea cycle citrulline and ornithine and the TCA intermediate malic acid.
-Compounds less abundant in Falkor were adenine, gonyol, hydroxyisoleucine, leucine, alanine, and nicotinic acid.
-
-We were also able to use the Falkor dataset to test whether the compounds detected as significantly different in the MESOSCOPE dataset were also different in this eddy pair.
-We expected to find abundant TMAO and more hydroxyisoleucine at the surface than the DCM in addition to most compounds enriched in the cyclone DCM with biomass, especially the six compounds mentioned above in the results of Figure 4 (trigonelline, homarine, DMSAc, DMSP, taurine, isethionic acid).
+We were also able to use the HEE dataset to test whether the compounds detected as significantly different in the MESO-SCOPE dataset were also different in this eddy pair.
+We expected to find abundant TMAO and more hydroxyisoleucine at the surface than the DCM in addition to most compounds enriched in the cyclone DCM with biomass, especially the six compounds mentioned above in the results of Figure 5 (trigonelline, homarine, DMS-Ac, DMSP, taurine, and isethionic acid).
 
@@
 
-As expected, TMAO was again one of the most abundant metabolites detected in the particulate matter with concentrations around 1.2 nM at the surface and 0.4 nM at the DCM, second only to the high levels of the amino acid glycine (2 nM surface, 0.9 nM DCM), though the differences between the surface and DCM were not significant for either compound.
-We did detect a significant difference between depths for hydroxyisoleucine, once again expected, despite the aforementioned lower concentrations in this sample set (surface mean = 0.32 nM, DCM mean = 0.21 nM, t-test p-value=0.022 with n=12 samples at each depth).
+As expected, TMAO was again one of the most abundant metabolites detected in the particulate matter with concentrations around 1.2 nM at 25 meters and 0.4 nM at the DCM, second only to the high levels of the amino acid glycine (2 nM at 25 meters, 0.9 nM DCM), though the differences between 25 meters and DCM were not significant for either compound.
+We also detected a significant difference between depths for hydroxyisoleucine (25 meter mean = 0.32 nM, DCM mean = 0.21 nM, t-test p-value=0.022 with n=12 samples at each depth).
+These results imply that the overall depth structure of the metabolome of the gyre is relatively fixed for the most abundant compounds.
 
 The eddy effects, on the other hand, were much less strong during this cruise.
-Of the six metabolites with the strongest differences across the eddy dipole at the DCM in the MESOSCOPE cruise, only isethionic acid was significantly different in Falkor (isethionic acid t-test p-value$_{FDR}$ = 0.043, other p-values > 0.25).
-Notably, this difference in isethionic acid concentration was only found after normalizing each sample to the sum of all metabolites in the sample to control for biomass.
-Arsenobetaine, found in the MESOSCOPE samples to be strongly enriched in the anticyclone, was again found to have the same pattern here even at the DCM both when normalizing to biomass and when not.
-Surprisingly, the 173.09211 *m/z* mass feature was also slightly enriched but this time in the *cyclone* DCM with peak areas approximately 1.4 times the values in the anticyclone (t-test p-value = 0.067, Mann-Whitney p-value = 0.045).
-The 170.11756 *m/z* mass feature was not detected in the Falkor samples.
+Of the six known metabolites with the strongest differences across the eddy transect at the DCM in the MESO-SCOPE cruise, only isethionic acid was significantly different in the HEE dataset (isethionic acid t-test p-value$_{FDR}$ = 0.043, other five p-values > 0.25).
+This difference in isethionic acid concentration was only found after normalizing each sample to the sum of all metabolites in the sample to control for biomass.
+Arsenobetaine was again found to strongly enriched in the anticyclone at the DCM both when normalizing to biomass and when not, just as in MESO-SCOPE.
+Surprisingly, the 173.0921 *m/z* mass feature was also slightly enriched but this time in the *cyclone* DCM with peak areas approximately 1.4 times the values in the anticyclone (t-test p-value = 0.067, Mann-Whitney p-value = 0.045).
+The 170.1176 *m/z* mass feature was not detected in the HEE samples.
 
@@
 
-## Multivariate approaches reveal metabolome-wide shifts across eddy dipoles
-
-We measured the metabolome of samples collected across two sets of eddy dipoles to explore the effect of sea level anomaly (SLA) on metabolite composition and found that mesoscale eddies explain a significant portion of observed variations in particulate metabolomes.
-
-In all of the datasets analyzed here, we detected a significant difference in the composition of the metabolome between eddy dipoles.
-This effect was strongest in the samples taken during the Lagrangian stations at the center of each eddy in the 2017 MESOSCOPE cruise, with nearly a quarter of the total variance explained by the eddy from which the samples were taken.
-In the samples taken along the eddy transect, the largest effect was detected in the deepest samples taken from 175 meters and from the DCM, with much less of a response at the surface.
-In the Falkor samples, however, the opposite was true with a larger SLA response at the surface than that at the DCM.
+## Multivariate approaches reveal metabolome-wide shifts across pairs of eddies of opposite polarity
+
+We measured the metabolome of samples collected across two sets of adjacent eddies of opposite polarity to explore the effect of sea level anomaly (SLA) on metabolite composition and found that the altered biogeochemistry and microbial community composition between cyclonic and anticyclonic eddies explain a significant portion of the observed variations in their particulate metabolomes.
+
+In all of the datasets analyzed here, we detected a significant difference in the composition of the metabolome between the adjacent eddies.
+This effect was strongest in the samples taken during the Lagrangian stations at the center of each eddy in the 2017 MESO-SCOPE cruise, with nearly a quarter of the total variance explained by the eddy from which the samples were taken.
+In the samples taken along the eddy transect, the largest effect was detected in the deepest samples taken from 175 meters and from the DCM, with much less of a response at 25 meters.
+In the HEE samples, however, the opposite was true with a larger SLA response at 25 meters than that at the DCM.
 In each case, the differences introduced by SLA were smaller than those of depth when compared directly, with even the high-resolution sampling around the DCM finding slightly more variance explained by the 40 meter difference in sampling depth than the 40 centimeter difference in SLA.
@@
 
-Although possible that the observed differences in metabolomes across eddy dipoles are due to latitudinal shifts or simply background variation in the gyre environment, the transect data in particular implies that this is not the case.
+Although possible that the observed differences in metabolomes across adjacent eddies differing in polarity are due to latitudinal shifts or simply background variation in the gyre environment, the transect data in particular implies that this is unlikely to be the case.
 Samples taken when the absolute value of the SLA was less than 5 cm were more similar to each other than to those samples taken within the eddies despite larger differences in latitude.
@@
 We also detected the largest differences in the DCM metabolome at the exact center of the eddy, with Station 12 highly distinct from even the nearest stations, while no such stark difference was found in the anticyclone (Figures 1 and 2).
-This perhaps indicates that the exact center of the eddy at the DCM is a unique metabolome relative to the rest of the transect rather than a smooth continuation of the existing trends.
-
-A large driver of metabolomic differences was shifts in total biomass, with earlier work in the same location by @Barone2022 showing that particulate carbon (PC), chlorophyll, and beam attenuation were all 20-80% higher in the DCM of the cyclone relative to the anticyclone.
-We see this reflected particularly well in the targeted metabolites measured here, with clear depth trends and a strong correlation between total metabolite concentration and PC values.
+This perhaps indicates that the exact center of the cyclone has a unique metabolome relative at the DCM relative to the rest of the transect rather than existing as a smooth continuation.
+
+Shifts in total biomass were a large driver of metabolomic differences, with earlier work in the same location by @Barone2022 showing that particulate carbon, chlorophyll, and beam attenuation were all 20-80% higher in the DCM of the cyclone relative to the anticyclone.
+We see this reflected particularly well in the targeted metabolites measured here, with clear depth trends and a strong correlation between total metabolite concentration and particulate carbon values.
 However, even after controlling for biomass effects by normalizing each sample to the sum of signal measured within it we still found a significant SLA effect, likely due to shifts in the community composition and in particular the increased eukaryote presence in the cyclone DCM.
-
-High-resolution depth sampling around the DCM at the eddy centers allowed us to explore the multivariate interactions between depth and SLA.
-Given that the difference in DCM depth between the two eddies was approximately 20 meters, our 10-meter sampling resolution allowed us to explore two potential models for metabolites.
-The first model, one driven largely by pressure or light availability, expects the absolute depth to be a stronger predictor than the difference in depth from the DCM center.
-The second model is more sensitive to dissolved water column parameters such as salinity or temperature, and predicts that samples equal distances from the DCM will be the most similar.
-The strong correlation of both depth and SLA with the first principal component of the PCA indicates that the largest variance in the system is best captured by a combined model with largest PC1 values in the deep anticyclone and smallest in the shallow cyclone.
 
@@
 
-The NMDS plots of the high-resolution depth samples were also illustrative of sample similarity, with clear SLA and depth trends.
+The NMDS plots of the high-resolution depth samples were also illustrative of sample dissimilarity with clear SLA and depth trends.
 The anticyclonic samples grouped tightly and showed little variance with depth or triplicate when compared to the cyclonic samples with the exception of DCM triplicate B, which was highly distinct for unknown reasons.
@@
 One notable aspect of their clustering was the way in which the deepest samples (DCM plus 20m) grouped most closely to the anticyclonic samples, perhaps indicating that below the DCM the metabolome rapidly approaches a uniform deep-water signal.
-This result was noted previously in the 175 meter samples of the eddy transect in @Kumler2023, where samples from the deep euphotic zone showed greater intra-depth similarity than samples from the DCM or surface.
-
-Given the confluence of both depth and SLA signals and the way that the boxplots of Figure 1 and Supplemental Figure 1 would confound the SLA signal with the depth variance, we used k-means clustering to group similar compounds and provide a reduced dimensionality space for visualization.
-This revealed four major patterns of metabolite response, with a majority of compounds responding to eddy state (Clusters 1, 2, and 4 in Figure 4).
+This result was noted previously in the 175 meter samples of the eddy transect in @Kumler2023, where samples from the deep euphotic zone showed greater intra-depth similarity than samples from the DCM or 25 meters.
+
+Given the confluence of both depth and SLA signals and the way that the boxplots of Figure 2 would confound the SLA signal with the depth variance, we used k-means clustering to group similar compounds and provide a reduced dimensionality space for visualization.
+This revealed four major patterns of metabolite response, with a majority of compounds responding to eddy state (Clusters 1, 2, and 4 in Figure 5).
 It is interesting to note that the sole cluster in which abundances in the anticyclone are greater than those of the cyclone is also the only cluster to generally increase in abundance with depth (Cluster 4), perhaps indicating that compounds more concentrated in the anticyclone are the same kind of degradation products and recalcitrant carbon that is found at depth.
@@
 
-The Falkor cruise upset many of the expectations we developed during the MESOSCOPE cruise in the previous year.
+The Hawaiian Eddy Experiment (HEE) data upset many of the expectations we developed during the MESO-SCOPE cruise in the previous year.
 Most surprising was the large difference between the eddies detected at the *surface*, while the DCM samples were functionally indistinguishable.
-This contrasts directly with the results from MESOSCOPE and the results in @Gleich2024, who found eddy-driven shifts in protistan community composition to be larger at depth than at the surface.
+This contrasts directly with the results from MESO-SCOPE and the results in @Gleich2024, who found eddy-driven shifts in protistan community composition to be larger at depth than at the surface.
 However, @Dugenne2023 found differences in nitrogen fixation rate and nitrogen fixer composition varying widely throughout the upper water column.
-The large surface differences were especially surprising given the large inter-replicate differences between the Falkor surface samples in this study, while the DCM samples tended to be much more consistent.
+The large surface differences were especially surprising given the large inter-replicate differences between the HEE surface samples in this study, while the DCM samples tended to be much more consistent.
 
@@
 
-```{r}
+```{r kheal comparison, include=FALSE}
 raw_kheal <- read_csv("https://raw.githubusercontent.com/kheal/Gradients1_SemiTargeted3/master/Tables/Manuscript_tables/SuppTables/Full_EnviroCulture_PeakAreas.csv")
@@
   filter(compound_name_kath=="Histidine") %>%
-  distinct(feature, MassFeature_Column, mzmed, rtmed, mz, rt, compound_name_will, compound_name_kath) %>%
-  print(n=Inf)
+  distinct(feature, MassFeature_Column, mzmed, rtmed, mz, rt, compound_name_will, compound_name_kath)
 
@@
 We found that our list of authentic standards covered fewer metabolites in anticyclonic eddies at all depths, with the molecular diversity of the cyclone much better characterized.
-Many of the trends detected below persisted even after normalizing within each sample, indicating that the shifts discussed have implications beyond simple scaling with biomass.
+Many of the trends detected persisted even after normalizing within each sample, indicating that the shifts discussed have implications beyond simple scaling with biomass.
 
@@
 This abundant compound increased approximately threefold in concentration from < 50 pM in the anticyclone to around 150 pM in the center of the cyclone in both the eddy transect and eddy center datasets.
-The pattern was weaker in the noisier Falkor data but the largest concentrations were still detected in the cyclone and the median cyclone measurement was greater than the maximum value for the anticyclone.
+The pattern was weaker in the noisier HEE data but the largest concentrations were still detected in the cyclone and the median cyclone measurement was greater than the maximum value for the anticyclone.
 This molecule has a well-established role as an osmolyte in eukaryotic phytoplankton and cyanobacteria [@Gebser2013; @Dawson2020; @Heal2021; @Durham2022] as well as a documented decrease in abundance with depth [@Heal2021], indicating that its response to SLA is potentially due to differences in physical and chemical attributes that in turn control biomass, community composition, and recycling rates.
@@
 Isethionate is a known osmolyte thus far found exclusively in eukaryotes [@Durham2022] and strongly associated with a few diatoms in particular [@Heal2021].
-This compound was enriched in the cyclone of the DCM in both the MESOSCOPE and Falkor cruises along with its precursor taurine, validating earlier results from @Barone2022 that showed strong enrichment of most eukaryotes at the DCM and *Pseudo-nitzschia* in particular.
-Taurine was also among the compounds highly enriched in the anomalous surface sample from Station 11 (MS11C215m_A), likely along with its trimethylated form taurine betaine (N-trimethyltaurine).
+This compound was enriched in the cyclone of the DCM in both the MESO-SCOPE and HEE cruises along with its precursor taurine, validating earlier results from @Barone2022 that showed strong enrichment of eukaryotic phytoplankton at the DCM and *Pseudo-nitzschia* in particular.
+Taurine was also among the compounds highly enriched in the anomalous 15 meter sample from Station 11, likely along with its N-trimethylated form taurine betaine (N-trimethyltaurine).
 While putative in our analysis, taurine betaine has been previously reported in corals and red algae, where it probably has a destabilizing influence on proteins [@Impellizzeri1975; @Hill2017; @Bruzdziak2018].
@@
 
-The strong SLA signal detected among the 175 meter samples was partially surprising to us, having expected the effect to be strongest at the DCM where isopycnals are interacting strongly with light from the surface.
+The strong SLA signal detected among the 175 meter samples was partially surprising to us, having expected the effect to be strongest at the DCM where eddy effects lift large concentrations of nutrients above the 1% light level.
 Instead, we found more compounds overall to be significantly different across the eddy transect in the 175 meter samples than we did at the DCM (α=0.05, 43 compounds at 175 meters vs 22 at the DCM).
-The enrichment of nucleobases in the 175 meter samples may hint at bacterial biomass enrichment, as nucleobases tend to be highly abundant in bacterial cultures [@Heal2021].
-The abundance of a mass feature putatively identified as acetylcholine in the 175 meter cyclone samples followed the same general trend as arsenobetaine, both enriched at depth and in the anticyclone, and may be another compound that's the result of heterotrophic degradation of phytoplankton metabolites [@Durham2022].
+The enrichment of nucleobases in the 175 meter samples may hint at bacterial biomass, as nucleobases tend to be highly abundant in bacterial cultures [@Heal2021].
+The abundance of a mass feature putatively identified as acetylcholine in the 175 meter cyclone samples followed the same general trend as arsenobetaine, both enriched at depth and in the anticyclone, and may be another compound that results from heterotrophic degradation of phytoplankton metabolites [@Durham2022].
 These results raise important questions about the depth at which the community is isolated from SLA effects, with additional data from @Barone2022 and @Gleich2024 indicating that even at 250 meters eddy effects are still discernable.
 
-A final pattern worth noting is the enrichment of metabolites in the urea-ornithine cycle (OUC) among the samples taken during the Falkor cruise.
-Aspartate, citrulline, ornithine, and malic acid were all found to be more abundant in the Falkor samples relative to MESOSCOPE.
+A final pattern worth noting is the enrichment of metabolites in the urea-ornithine cycle (OUC) among the samples taken during the HEE cruise.
+Aspartate, citrulline, ornithine, and malic acid were all found to be more abundant in the HEE samples relative to MESO-SCOPE.
 The OUC is a staple of metazoan metabolism and is present in stramenopiles as well where it enables rapid response to increases in nitrogen supply [@Allen2011; @Bender2014].
-The enrichment of these metabolites in the Falkor cruise may indicate that there was still rapid uptake and turnover of nitrogen in the 2018 eddy pair while the 2017 eddies were more stable, despite both eddies being in a similar state of stable anticyclone and weakening cyclone [@Dugenne2023].
-
-More broadly, differences between the two cruises are most likely due to differences in eddy age and the earlier sampling time of the 2018 cruise (March/April instead of the June/July sampling in 2017) resulting in shifts in community composition and response.
-@Barone2022 documents the 2017 cruise community composition, reporting abundant Prymnesiophyceae at both the surface and DCM, with other significant contributions from dinoflagellates, diatoms, and green algae.
-In the 2018 cruise both haptophytes and dinoflagellates again dominated the eukaryotic fraction, though notably very little was contributed by diatoms [@Gleich2024].
-@Dugenne2023 and @Harke2021 also cite uncontrolled differences between eddies based on their origin and age.
+The enrichment of these metabolites in the HEE cruise may indicate that there was still rapid uptake and turnover of nitrogen in the 2018 eddy pair while the 2017 eddies were more stable, despite both eddies being in a similar state of stable anticyclone and weakening cyclone [@Dugenne2023].
+
+Differences between the two cruises could also be due to differences in eddy age and/or the different period of the year when sampling took place during the two cruises (March/April in 2018 instead of the June/July sampling in 2017) resulting in shifts in community composition and response.
+@Barone2022 documents the 2017 cruise community composition, reporting abundant Prymnesiophyceae at both 15 meters and the DCM, with other significant contributions from dinoflagellates, diatoms, and green algae.
+In the 2018 cruise both haptophytes and dinoflagellates again dominated the eukaryotic fraction, though surprisingly little was contributed by diatoms [@Gleich2024].
+@Dugenne2023 and @Harke2021 also highlight uncontrolled differences between the two sets of eddies based on their origin and age.
 
@@
 
+# Conclusions
+
+This study reports how changes in the biogeochemistry and community composition due to mesoscale eddies affect the metabolome and thereby the composition of particulate organic matter.
+A transect across adjacent mesoscale eddies of opposite polarity showed that many metabolites track closely with metrics of overall biomass, with eddy effects stronger at 175 meters than at 15 meters or the deep chlorophyll maximum (DCM).
+High-resolution depth sampling of the DCM at the center of each eddy elucidated several known and unknown biomarkers likely corresponding to the previously documented increase in eukaryotic phytoplankton in the cyclone.
+Metabolite clusters aligned well with expected trends in abundance with depth and between the two eddies, with rapid convergence below the DCM towards a general deep-sea metabolite signal.
+By contrasting these effects with a follow-up analysis in the centers of a separate pair of mesoscale eddies of opposite polarity the following year, we learned that the impact of eddies may be influenced by their origin, age, and period of the year.
+Finally, the metabolites with the strongest responses to eddy effects were unknown and a relatively smaller proportion of known metabolites in the anticyclone demonstrates the utility of untargeted metabolomics in exploring environmental variation and the large amount of information that can still be gleaned from such datasets.
+
 # Methods
@@
 Samples were collected from two cruises in the North Pacific Subtropical Gyre near Station ALOHA that targeted strong mesoscale eddy features as described in @Dugenne2023 and @Gleich2024.
-Briefly, the 2017 MESOSCOPE cruise (Microbial Ecology of the Surface Ocean-Simons Collaboration on Ocean Processes and Ecology, KM1709 on the *R/V Kilo Moana*) consisted of a transect across the cyclonic and anticyclonic poles of the eddy dipole as well as two long-term Lagrangian stations at the center of each eddy.
-The cyclonic pole of the eddy had a maximum negative SLA anomaly of -20 centimeters and the anticyclonic center reached +24 centimeters.
-The transect samples were taken at various times of day as the ship transected the dipole, and the eddy center samples were all collected between 5 and 8 pm.
-In 2018, the Falkor cruise (FK180310 on the *R/V Falkor*) targeted a new eddy dipole in approximately the same location with samples taken at the center of each eddy (maximum negative anomaly in the cyclone = -15 cm, maximum positive anomaly in the anticyclone = 26 cm).
+Briefly, the 2017 MESO-SCOPE cruise (Microbial Ecology of the Surface Ocean-Simons Collaboration on Ocean Processes and Ecology, KM1709 on the *R/V Kilo Moana*) consisted of a transect across adjacent cyclonic and anticyclonic eddies as well as two long-term Lagrangian stations at the center of each eddy.
+The cyclonic eddy had a maximum negative sea level anomaly (SLA) of -20 centimeters and the anticyclonic eddy reached +24 centimeters.
+The transect samples were taken at various times of day as the ship transected the adjacent eddies while the eddy center samples were all collected between 5 and 8 pm.
+In 2018, the Hawaiian Eddy Experiment (HEE, FK180310 on the *R/V Falkor*) targeted a new cyclonic and anticyclonic eddy in approximately the same location with samples taken at the center of each eddy (maximum negative anomaly in the cyclone = -15 cm, maximum positive anomaly in the anticyclone = +26 cm) (Figure 1).
 
@@
 Biogeochemical measurements were collected as described in @Barone2022 and @Dugenne2023 mostly following protocols used by the HOT program (http://hahana.soest.hawaii.edu/hot/methods/results.html).
-Briefly, all environmental data was collected via CTD Niskin bottles except beam attenuation which was measured via Hyperpro and sea level anomaly (SLA) which was measured via satellite.
-Chlorophyll and pheophytin were measured fluorometrically, particulate carbon (PC) and particulate nitrogen (PN) were measured using an elemental analyzer, and nitrate + nitrite (N+N) and soluble reactive phosphorus (phosphate, PO$_4^{3-}$) were measured on an autoanalyzer except where concentrations were below 100nM in which case they were measured using chemiluminescence.
+Briefly, all environmental data was collected via CTD rosette except SLA which was measured via satellite.
+Particulate carbon and particulate nitrogen were measured using an elemental analyzer, and nitrate + nitrite (N+N) and soluble reactive phosphorus (phosphate, PO$_4^{3-}$) were measured on an autoanalyzer except where concentrations were below 100 nM in which case they were measured using chemiluminescence.
 Values were interpolated linearly to the locations at which metabolite samples were taken using all data collected at the given station.
-
-## Metabolite sample collection
-Samples were obtained using the onboard CTD rosette to collect water from 15 meters, the deep chlorophyll maximum (DCM), and 175 meters during the eddy transect and from the DCM ±10 and ±20 meters during the eddy center sampling in the MESOSCOPE cruise and from 25 meters and the DCM during the Falkor cruise.
+Two particulate carbon values were determined to be spurious, with concentrations 2-3 times higher than typical Station ALOHA values, and were instead estimated using a best-fit linear regression against beam attenuation.
+
+## Sample collection for particulate metabolites
+Samples were obtained using the onboard CTD rosette to collect water from 15 meters, the deep chlorophyll maximum (DCM), and 175 meters during the MESO-SCOPE eddy transect and from the DCM ±10 and ±20 meters during the eddy center sampling in the MESO-SCOPE cruise and from 25 meters and the DCM during the HEE cruise.
 The DCM was determined visually from fluorometer data during the CTD downcast and Niskin bottles were tripped during the return trip to the surface.
@@
 Samples were brought to the surface and decanted into prewashed (3x with DI, 3x with sampled seawater) polycarbonate bottles for filtration.
-Samples were filtered by peristaltic pump onto 142mm 0.2 µm Durapore filters held by polycarbonate filter holders on a Masterflex tubing line.
+10L samples were filtered by peristaltic pump onto 142mm 0.2 µm Durapore filters held by polycarbonate filter holders on a Masterflex tubing line.
 Pressures were kept as low as possible while still producing a reasonable rate of flow through the filter, approximately 250-500 mL per minute.
@@
 Extraction followed a modified Bligh & Dyer approach as detailed in @Boysen2018.
-Briefly, filters were added to PTFE centrifuge tubes with a 1:1 mix of 100 µm and 400 µm silica beads, approximately 2mL -20 °C Optima-grade DCM, and approximately 3mL -20 °C 1:1 methanol/water solution (both also Optima-grade).
-Extraction standards were added during this step.
-The samples were then bead-beaten three times, followed by triplicate washes with fresh methanol/water mixture.
+Briefly, filters were randomized and added to PTFE centrifuge tubes with a 1:1 mix of 100 µm and 400 µm silica beads, approximately 2 mL -20 °C Optima-grade dichloromethane, and approximately 3 mL -20 °C 1:1 methanol/water solution (both also Optima-grade).
+Isotope labeled extraction standards were also added during this step (Supplemental table 1).
+The samples were then bead-beaten three times, followed by triplicate extraction of the aqueous layer and addition with fresh methanol/water mixture with additional bead-beatings in between.
 Samples were then dried down under clean nitrogen gas and warmed using a Fisher-Scientific Reacti-Therm module.
-Dried aqueous fractions were re-dissolved in 380 µL of Optima-grade water and amended with 20 µL isotope-labeled injection standards.
-Additional internal standards were added at this point to measure the variability introduced by chromatography and ionization, and the reconstituted fraction was syringe-filtered to remove any potential clogging material.
-This aqueous fraction was then aliquoted into an HPLC vial for injection on the HILIC column and diluted 1:1 with Optima-grade water.
-A pooled sample was created by combining 20 µL of each sample into the same HPLC vial, and a 1:1 dilution with water half-strength sample was aliquot from that to assess matrix effects and obscuring variation [@Boysen2018].
-Also run alongside the environmental samples were two mixes of authentic standards in water and in an aliquot of the pooled sample at a variety of concentrations for quality control, annotation, and absolute concentration calculations.
+Dried aqueous fractions were reconstituted in 380 µL Optima-grade water and amended with 20 µL isotope-labeled injection standards (Supplemental table 1) to measure the variability introduced by chromatography and ionization.
+The reconstituted fraction was syringe-filtered into precombusted glass inserts in HPLC vials to remove any potential clogging material.
+Samples were then additionally diluted 1:1 with Optima-grade water to prevent overloading on the column and reduce salt effects.
+
+A pooled sample was created by combining 20 µL of each sample into the same HPLC vial, and a 1:1 dilution with water created a half-strength pooled sample from that to assess matrix effects and obscuring variation [@Boysen2018].
+Also run alongside the environmental samples were samples of authentic standards split into two mixes (Supplemental table 2).
+Standards were dissolved in Optima-grade water to help with identification and into an aliquot of the pooled sample to quantify matrix response factors.
 HPLC vials containing the samples were frozen at -80 °C until thawing shortly before injection.
@@
 ## HPLC-MS methods
-Separate HPLC-MS runs were used for the MESOSCOPE eddy transect, the MESOSCOPE eddy center, and the Falkor cruise data.
-For each, a SeQuant ZIC-pHILIC column (5 um particle size, 2.1 mm x 150 mm, from Millipore) was used with 10 mM ammonium carbonate in 85:15 acetonitrile to water (Solvent A) and 10 mM ammonium carbonate in 85:15 water to acetonitrile (Solvent B) at a flow rate of 0.15 mL/min.
+Separate HPLC-MS runs were used for the MESO-SCOPE eddy transect, the MESO-SCOPE eddy center, and the HEE cruise data.
+Eddy center samples were run in February 2018, eddy transect samples were run later during August of that year, and the HEE samples were run in July 2019.
+Each run was treated as a single batch and injected in sequence while maintaining the randomization that had occurred during extraction to minimize chromatographic shifts from solvent or column switching.
+
+For each run, a Waters Acquity I-Class UPLC with a SeQuant ZIC-pHILIC column (5 µm particle size, 2.1 mm x 150 mm, from Millipore) was used with 10 mM ammonium carbonate in 85:15 acetonitrile to water (Solvent A) and 10 mM ammonium carbonate in 85:15 water to acetonitrile (Solvent B) at a flow rate of 0.15 mL/min.
 The column was held at 100% A for 2 minutes, ramped to 64% B over 18 minutes, ramped to 100% B over 1 minute, held at 100% B for 5 minutes, and equilibrated at 100% A for 25 minutes (50 minutes total).
@@
 
-Metabolomic data was collected on a Thermo Q Exactive HF hybrid Orbitrap (QE) mass spectrometer.
+The Waters Acquity UPLC was coupled to a Thermo Q Exactive HF hybrid Orbitrap high resolution mass spectrometer equipped with a heated electrospray ionization source (H-ESI).
+The H-ESI voltage was set to 3.3 kV and sheath gas, auxiliary gas, and sweep gas flow rates were set at 16, 3, and 1, respectively.
 The capillary and auxiliary gas heater temperatures were maintained at 320°C and 100°C, respectively.
-The S-lens RF level was kept at 65, the H-ESI voltage was set to 3.3 kV and sheath gas, auxiliary gas, and sweep gas flow rates were set at 16, 3, and 1, respectively.
-Polarity switching was used with a scan range of 60 to 900 *m/z* and a resolution of 60,000.
-Calibration was performed every 3-4 days at a target mass of 200 *m/z*.
-DDA data was collected from the pooled samples for high-confidence annotation of knowns and unknowns.
-All files were then converted to an open-source mzML format and centroided via Proteowizard’s msConvert tool.
+Full scan analyses were performed with polarity switching and a scan range of 60 to 900 *m/z* at a resolution of 60,000.
+The instrument was mass calibrated at 200 *m/z* before each run began and once again during the eddy transect sample set to ensure calibrations were always performed within 3-4 days.
+All data files were then converted to an open-source mzML format and vendor centroided via Proteowizard’s msConvert tool.
+All mzML files have been uploaded to Metabolomics Workbench under Project ID PR001738.
 
@@
 
-Targeted compounds were manually integrated using Skyline and MSDIAL.
-Skyline was used for both the eddy center samples and the Falkor cruise data while MSDIAL was used for the eddy transect data due to the larger number of files in that dataset.
-Raw peak areas were then normalized to their best-matched internal standard per @Boysen2018 except that all compounds were normalized to their best match no matter how much or how little improvement was detected.
-Normalized peaks were then calibrated via a single-point standard curve as determined uniquely for each compound from the authentic standards and used to convert normalized peak area into moles per µL injected and scaled to the amount of seawater filtered to provide a final estimate of environmental concentration for each compound in moles per liter of seawater filtered.
+Compounds for which we had an authentic standard (Supplemental table 2) were manually integrated using Skyline and MSDIAL with the standard mixes used to ensure the correct peak was integrated by matching retention time.
+Compounds that were visually determined to have poor peak quality in the samples were removed from the downstream analysis.
+Skyline was used for both the eddy center samples and the HEE cruise data while MSDIAL was used for the eddy transect data due to the larger number of files in that dataset.
+Raw peak areas were then normalized to their best-matched internal standard (Supplemental table 1) per @Boysen2018 except that all compounds were normalized to their best match no matter how minimal the improvement.
+Normalized peaks were then calibrated via a single-point standard curve as determined uniquely for each compound from the authentic standard mixes and used to convert normalized peak area into moles per µL injected and scaled to the amount of seawater filtered to provide a final estimate of environmental concentration for each compound in moles per liter of seawater filtered.
+
+```{r supp table 1 standards list}
+clean_stans <- read_csv("../metadata/clean_stans.csv")
+raafay_notes <- read_csv("../targeted/raw_data/Raafay_HILIC_notes.csv") %>%
+  select(compound_name, in_samps)
+
+IS_table <- clean_stans %>%
+  filter(compound_type=="Internal Standard") %>%
+  select(compound_name, formula, stan_mz, stan_rt, conc_um, date_added, polarity) %>%
+  mutate(ms_run=case_when(
+    date_added<180205~"All",
+    date_added<=180821~"MESO-SCOPE Eddy Transect and HEE",
+    date_added<=190715~"HEE only",
+    TRUE~"None"
+  )) %>%
+  filter(ms_run!="None") %>%
+  mutate(polarity=ifelse(polarity=="pos", "Positive", "Negative")) %>% 
+  mutate(stan_type=case_when(
+    compound_name%in%c("L-Cysteic acid, 2H3", "3-Sulfolactate, 13C3", "Sulfoacetic acid, 13C2", "Isethionic acid, 13C2", "Taurine, 2H4")~"Extraction",
+    TRUE~"Internal"
+  )) %>%
+  arrange(stan_mz) %>%
+  select(`Compound name`=compound_name, `Molecular formula`=formula,
+         `m/z ratio`=stan_mz, `Retention time (min)`=stan_rt,
+         `Concentration (uM)`=conc_um, `Polarity integrated`=polarity,
+         `Which runs`=ms_run, `Extraction or internal`=stan_type)
+
+Auth_table <- clean_stans %>%
+  mutate(ms_run=case_when(
+    date_added<180205~"All",
+    date_added<=180821~"MESO-SCOPE Eddy Transect and HEE",
+    date_added<=190715~"HEE only",
+    TRUE~"None"
+  )) %>%
+  filter(ms_run!="None") %>%
+  left_join(raafay_notes) %>%
+  mutate(polarity=ifelse(polarity=="pos", "Positive", "Negative")) %>%
+  arrange(stan_mz) %>%
+  select(`Compound name`=compound_name, `Molecular formula`=formula,
+         `m/z ratio`=stan_mz, `Retention time (min)`=stan_rt,
+         `Concentration (uM)`=conc_um, `Polarity integrated`=polarity,
+         `Standard mix`=mix, `Which runs`=ms_run, `Good quality`=in_samps)
+
+wb <- createWorkbook(creator = "William Kumler, wkumler@uw.edu")
+addWorksheet(wb, sheetName = "Table 1 Internal Standards")
+writeData(x=IS_table, wb = wb, sheet = "Table 1 Internal Standards")
+
+addWorksheet(wb, sheetName = "Table 2 Authentic Standards")
+writeData(x=Auth_table, wb = wb, sheet = "Table 2 Authentic Standards")
+
+saveWorkbook(wb, file = "supplement/Supplemental Tables.xlsx", overwrite = TRUE)
+```
 
@@
 Samples were also processed via the untargeted workflow detailed in @Kumler2023.
-In short, XCMS [@R-xcms] was used for peak detection and followed by manual inspection of each mass feature for quality with only those peaks classified as "Good" used in the analysis here.
-Features were then matched with and normalized to an internal standard as described above.
+In short, XCMS [@R-xcms] was used for feature detection with CentWave peakpicking, Obiwarp alignment, and peak density correspondence/grouping.
+This was followed by manual inspection of the extracted ion chromatogram for each mass feature by an expert for quality with only those peaks classified as "Good" used in the analysis here.
+Features were then matched with and normalized to an internal standard (Supplemental table 1) as described above.
 
@@
 
-All statistics were run in R version `r paste(version$major, version$minor, sep=".")`.
+All statistics were run in R version 4.3.3.
 We used multivariate statistics provided by the vegan package [@R-vegan] to assess the overall impact of various metrics across the metabolome followed by univariate statistics to investigate individually significant responses.
@@
 For the analyses involving the eddy transect samples, sea level anomaly was treated as a continuous variable with the exception of the PERMANOVA, for which SLA was categorized as cyclonic if less than -5 cm and anticyclonic if greater than 5 cm, with a third category for stations in between [@Anderson2001].
-Both the eddy center and Falkor data treated SLA as a categorical variable.
-
-# Conclusions
-
-This study reports how changes in the biogeochemical environment and its associated community due to mesoscale eddy-driven mechanisms affect the metabolome and thereby the composition of particulate organic matter.
-The transect across an oligotrophic eddy dipole shows that many metabolites track closely with metrics of overall biomass, with eddy effects stronger at 175 meters than at the surface or deep chlorophyll maximum (DCM).
-High-resolution depth sampling of the DCM at the center of each eddy elucidated several known and unknown biomarkers likely corresponding to the well-documented increase in picoeukaryotes in the cyclone.
-Metabolite clusters aligned well with expected trends in abundance with depth and between the two eddies, with rapid convergence below the DCM towards a general deep-sea metabolite signal.
-By contrasting these effects with a followup analysis in the centers of a separate dipole the following year, we also learned that the importance of eddies may be an unpredictable interaction between origin and age.
-Finally, the metabolites with the strongest responses to eddy effects were unknown and a relatively smaller proportion of known metabolites in the anticyclone demonstrates the utility of untargeted metabolomics in exploring environmental variation and the large amount of information that can still be gleaned from such datasets.
+Both the MESO-SCOPE eddy center and HEE data treated SLA as a categorical variable.
 
@@
 # Supplement
-## Supp fig: PCA axes 1-3 vs SLA by depth
+## Supp fig 1: PCA axes 1-3 vs SLA by depth
 
@@
 Lines of best fit and the associated standard error have been added behind the points in black and translucent grey, respectively.
-Strong associations exist between PC1 and SLA in the DCM samples and PC2 and SLA in the deep samples.
+Strong associations exist between PC1 and SLA in the DCM samples and PC2 and SLA in the 175 meter samples.
 B) Principal component 1 values for the 175 meter samples at each station showing large variance between triplicates rather than any external factor.*
 
-## Supp fig: Stacked barplot of top metabs by triplicate
+## Supp fig 2: Biomass regression
+
+```{r Fig 2.5 PC vs tot nM plot}
+pc_totalnm_plot <- no_outlier_frac_pc_in_metab_df %>%
+  ggplot(aes(x=PC_um, y=total_nM_C)) +
+  geom_smooth(method="lm", color="white", lwd=2.5, se=FALSE) +
+  geom_smooth(method="lm", color="black", lwd=1.5) +
+  geom_point(aes(fill=depth), pch=21, size=3) +
+  geom_smooth(aes(group=depth), method="lm", color="white", lwd=2.5, se = FALSE) +
+  geom_smooth(aes(color=depth), method="lm", lwd=1.5) +
+  geom_vline(xintercept = 0) +
+  geom_hline(yintercept = 0) +
+  theme_bw() +
+  theme(legend.position = c(0,1), legend.justification = c(0, 1), legend.background = element_rect(color="black")) +
+  labs(color="Sampling depth",
+       fill="Sampling depth",
+       x="Particulate carbon (μM)", 
+       y="Total nM carbon in known metabolites") +
+  scale_color_manual(breaks=c("Surface", "DCM", "Deep"), 
+                     labels=c("15 m", "DCM", "175 m"),
+                     values = c("#F4BB23", "#028E34", "#0B505C"), aesthetics = c("fill", "color"))
+
+pn_totalnm_plot <- no_outlier_frac_pc_in_metab_df %>%
+  ggplot(aes(x=PN_um, y=total_nM_N)) +
+  geom_point(aes(fill=depth), pch=21, size=3, alpha=0.8) +
+  geom_smooth(method="lm", color="white", lwd=2.5, se=FALSE) +
+  geom_smooth(method="lm", color="black", lwd=1.5) +
+  geom_smooth(aes(group=depth), method="lm", color="white", lwd=2.5, se = FALSE) +
+  geom_smooth(aes(color=depth), method="lm", lwd=1.5) +
+  geom_vline(xintercept = 0) +
+  geom_hline(yintercept = 0) +
+  theme_bw() +
+  theme(legend.position = "none") +
+  labs(color="Sampling depth",
+       fill="Sampling depth",
+       x="Particulate nitrogen (μM)", 
+       y="Total nM nitrogen in known metabolites") +
+  scale_color_manual(breaks=c("Surface", "DCM", "Deep"), 
+                     labels=c("15 m", "DCM", "175 m"),
+                     values = c("#F4BB23", "#028E34", "#0B505C"), aesthetics = c("fill", "color"))
+
+library(ggh4x)
+strip_theme <- strip_themed(background_x = elem_list_rect(fill= c("#F4BB23", "#028E34", "#0B505C")))
+
+pc_totalnm_plot_bydepth <- no_outlier_frac_pc_in_metab_df %>%
+  mutate(depth=factor(depth, levels=c("Surface", "DCM", "Deep"),
+                      labels=c("15 m", "DCM", "175 m"))) %>%
+  ggplot() +
+  aes(x=PC_um, y=total_nM_C) +
+  geom_smooth(method="lm", color="black") +
+  geom_point(aes(fill=sla_corr), pch=21, color="black", size=3, alpha=0.8) +
+  ggh4x::facet_wrap2(~depth, nrow=1, scales="free", strip = strip_theme) +
+  labs(x="Particulate carbon (μM)", y="Total nM carbon in\nknown metabolites") +
+  scale_fill_gradient2(
+    low = scales::muted("blue"), high = scales::muted("red"),  mid = "grey90",
+    aesthetics = c("fill"), name="Corr.\nSLA\n(cm)"
+  ) +
+  theme_bw() +
+  theme(strip.text = element_text(face = "bold", color="white")) +
+  guides(fill=guide_colorbar(title.hjust = 0.5, barheight = 5, title.position = "bottom"))
+
+comb_pcpn_plot <- egg::ggarrange(pc_totalnm_plot, pn_totalnm_plot, nrow=1, draw = FALSE)
+
+comb_pcpnnm_plot <- cowplot::plot_grid(comb_pcpn_plot, pc_totalnm_plot_bydepth, ncol = 1,
+                                     rel_heights = c(5, 3))
+
+ggsave("comb_pcpnnm_plot.png", plot = comb_pcpnnm_plot, device = "png", 
+       path = "supplement", width = 6.875, height = 5, units = "in", dpi = 300, 
+       bg = "white")
+```
+
+![](supplement/comb_pcpnnm_plot.png)
+
+*Supplemental figure 2: Type I linear regressions of particulate carbon and nitrogen against total nM carbon and nitrogen in known metabolites both as a whole and divided by depth.
+Points have been colored according to depth (top row of plots) or SLA (bottom row of plots) and have been placed on independent axes.
+Two outlier particulate carbon points have been interpolated from beam attenuation for 15 meter values at stations 10 and 11.*
+
+## Supp fig 3: Stacked barplot of top metabs by triplicate
 
@@
   mutate(depth=factor(depth, levels=c("Surface", "DCM", "Deep"), 
-                      labels=c("Surface  (15m)", "DCM  (110-120m)", "Deep  (175m)"))) %>%
+                      labels=c("15 m", "DCM", "175 m"))) %>%
   mutate(station=factor(station, levels=as.character(4:14)))
@@
   geom_col(width=1, position = "fill") +
-  scale_y_continuous(labels = scales::label_percent(), expand = expansion()) +
+  scale_y_continuous(labels = scales::label_percent(), expand = expansion(0.1)) +
   theme(legend.position = "bottom") +
@@
   ggtitle("B) Relative composition of the metabolome")
-comb_targ_rel_gp <- egg::ggarrange(abs_targ_bar_gp, rel_targ_bar_gp, ncol=1)
-
-ggsave("supplement/barplot_comb_targ_rel_gp.png", comb_targ_rel_gp, device = "png", width = 6.875, height = 5.5, units = "in", dpi = 300, bg="white")
+comb_targ_rel_gp <- egg::ggarrange(abs_targ_bar_gp, rel_targ_bar_gp, ncol=1, draw=FALSE)
+
+ggsave("supplement/barplot_comb_targ_rel_gp.png", comb_targ_rel_gp, device = "png", width = 6.875, height = 6.5, units = "in", dpi = 300, bg="white")
 ```
@@
 
-*Supplemental figure 2: Stacked barplots of known metabolite concentration as shown in main text Figure 2 but broken down by triplicate.
-The same data is shown in A) as B) but B) is rendered in relative contribution space instead of absolute concentration.*
-
-## Supp fig: Regressions of PC vs total nM metabolites with outliers
-
-```{r}
-comb_pc_reg_df <- bind_rows(
-  frac_pc_in_metab_df %>% 
-    mutate(status="All files") %>%
-    mutate(PC_um=PC_uncor),
-  no_outlier_frac_pc_in_metab_df %>%
-    mutate(status="Dropped outliers")
-)
-
-PC_comb_gp <- gridExtra::arrangeGrob(
-  comb_pc_reg_df %>%
-    ggplot(aes(x=PC_um, y=total_nM)) +
-    geom_point(aes(fill=depth), pch=21) +
-    geom_smooth(method="lm", color="black") +
-    geom_smooth(aes(color=depth), method="lm") +
-    facet_wrap(~status, ncol=1, scales="free_y") +
-    theme_bw() +
-    theme(legend.position = "top") +
-    labs(color="Sampling depth",
-         fill="Sampling depth",
-         x="Particulate carbon (uM)", 
-         y="Total nM carbon in known metabolites") +
-    scale_color_manual(breaks=c("Surface", "DCM", "Deep"), 
-                     values = c("#F4BB23", "#028E34", "#0B505C"), 
-                     aesthetics = c("fill", "color")),
-  comb_pc_reg_df %>%
-    ggplot() +
-    aes(x=PC_um, y=total_nM) +
-    geom_smooth(method="lm", color="black") +
-    geom_point(aes(fill=sla_corr), pch=21, color="black", size=2) +
-    ggh4x::facet_grid2(depth~status, scales="free", independent="all") +
-    labs(x="Particulate carbon (uM)", 
-         y="Total nM carbon in known metabolites",
-         fill="Corr.\nSLA\n(cm)") +
-    scale_fill_gradient2(
-      low = scales::muted("blue"), high = scales::muted("red"),  mid = "grey90"
-    ) +
-    theme_bw() +
-    theme(legend.position = "top"),
-  nrow = 1
-)
-plot(PC_comb_gp)
-comb_pc_reg_df %>%
-  nest(data=-c(status, depth)) %>%
-  mutate(reg_vals=map(data, ~broom::tidy(lm(.x$total_nM~.x$PC_um)))) %>%
-  unnest(reg_vals) %>%
-  filter(term!="(Intercept)")
-
-ggsave("supplement/PC_reg_w_wo_outliers.png", PC_comb_gp, device = "png", width = 6.875, height = 4, units = "in", dpi = 300, bg="white")
-
-# filter(!(station=="11" & tripl=="A" & depth=="Surface")) %>%
-# filter(!(station=="6" & tripl=="B" & depth=="Surface")) %>%
-# filter(!(station=="7" & tripl=="C" & depth=="Surface")) %>%
-# filter(!(station=="11" & tripl=="B" & depth=="DCM"))
-
-# xmdl <- lm(total_nM~PC_um, data=all_targ_pc_reg_df)
-# plot(fitted(xmdl), residuals(xmdl))
-# dfbeta(xmdl) %>% 
-#   as.data.frame() %>%
-#   cbind(all_targ_pc_reg_df$filename)
-```
-
-![](supplement/PC_reg_w_wo_outliers.png)
-
-*Supplemental figure 3: Regressions of particulate carbon against summed nM concentrations of known metabolites.
-The full regression with data from all sampling depths is shown on the left, while individual depths are shown on the right.
-For both the combined and depth-divided plots, Type I regression lines are shown both with and without outliers.
-Outlier data points were removed from the surface samples in Station 6, triplicate B; Station 7, triplicate C; and Station 11, triplicate A and from the DCM samples at Station 11, triplicate B.*
-
-## Supp fig: Falkor NMDS and PERMANOVA broken down by depth
+*Supplemental figure 3: Stacked barplots of known metabolite concentration as shown in main text Figure 3 but with individual triplicates shown.
+The same data is shown in A) as B) but B) is rendered in relative contribution space instead of absolute concentration.
+TMAO = trimethylamine N-oxide, HO-Ile = hydroxyisoleucine, GBT = glycine betaine, DCM = deep chlorophyll maximum.*
+
+## Supp fig 4: HEE NMDS and PERMANOVA broken down by depth
 
@@
 
-*Supplemental figure 4: Plots of the metabolome in the Falkor data, broken down by depth.
+*Supplemental figure 4: Plots of the metabolome during the Hawaiian Eddy Experiment, broken down by depth.
 The top row of plots are non-metric multidimensional scaling (NMDS) plots, in which points correspond to individual samples and have been colored by their corrected sea level height anomaly.
-SLA trends are visible in the surface samples, with dark blue circles consistently discriminating from the dark red circles.
-The bottom row of plots show the direction and magnitude of this effect by plotting the grand mean of the normalized metabolite peak areas in the stations taken across the eddy dipole.*
-
-## Unnumbered figures, location/inclusion TBD
-
-```{r lm_line_rank_order supplemental figure}
-reg_coefs <- untarg_peaks %>%
-  left_join(filled_file_metadata %>% distinct(filename, ms_run, station, depth, sla_corr)) %>%
-  filter(ms_run=="MT") %>%
-  group_by(feature) %>%
-  filter(sum(norm_area==0)<n()*0.9) %>%
-  mutate(norm_area=rank(norm_area)) %>%
-  ungroup() %>%
-  nest(data=-c(depth, feature)) %>%
-  mutate(test=map(data, ~broom::tidy(lm(.x$norm_area~.x$sla_corr)))) %>%
-  unnest(test)
-reg_coefs %>%
-  filter(str_detect(term, "sla_corr")) %>%
-  mutate(p.value=p.adjust(p.value, method="fdr")) %>%
-  arrange(p.value) %>%
-  left_join(untarg_feats) %>%
-  select(feature, depth, estimate, p.value, mzmed, rtmed, compound_name) %>%
-  filter(p.value<0.05) %>%
-  mutate(depth=factor(depth, levels=c("Surface", "DCM", "Deep"))) %>%
-  mutate(rtmed=round(rtmed, digits = 2)) %>%
-  mutate(mzmed=round(mzmed, digits = 5)) %>%
-  arrange(depth, p.value) %>%
-  group_by(depth) %>%
-  group_modify(~ add_row(.x, .after = nrow(.))) %>%
-  mutate(across(everything(), ~ifelse(is.na(.x), "", .x))) %>%
-  write_csv("supplement/lm_est_rank_order.csv")
-
-clean_slopes <- reg_coefs %>%
-  select(feature, data, depth, term, estimate, p.value) %>%
-  mutate(term=ifelse(term==".x$sla_corr", "slope", "intercept")) %>%
-  pivot_wider(names_from = term, values_from=c(p.value, estimate)) %>%
-  mutate(p.value=p.adjust(p.value_slope, method="fdr")) %>%
-  select(-p.value_intercept, -p.value_slope) %>%
-  filter(p.value<0.05) %>%
-  mutate(pointcol=ifelse(sign(estimate_slope)==1, "Anticyclone enriched", "Cyclone enriched")) %>%
-  mutate(depth=factor(depth, levels=c("Surface", "DCM", "Deep")))
-metadata_slopes <- filled_file_metadata %>%
-  filter(ms_run=="MT") %>%
-  filter(samp_type=="Smp") %>%
-  select(depth, pressure:beam_atten) %>%
-  distinct() %>%
-  pivot_longer(-c(sla_corr, depth)) %>%
-  group_by(name) %>%
-  mutate(value=rank(value)*3) %>%
-  nest(data=-c(name, depth)) %>%
-  mutate(lmout=map(data, ~broom::tidy(lm(.x$value~.x$sla_corr)))) %>%
-  unnest(lmout) %>%
-  filter(name!="dcm_depth") %>%
-  select(depth, name, data, term, estimate, p.value) %>%
-  pivot_wider(names_from = term, values_from = c(estimate, p.value)) %>%
-  mutate(p.value=p.adjust(`p.value_.x$sla_corr`)) %>%
-  filter(p.value<0.05) %>%
-  select(-c(`p.value_(Intercept)`, `p.value_.x$sla_corr`)) %>%
-  mutate(intercept=`estimate_(Intercept)`, slope=`estimate_.x$sla_corr`) %>%
-  mutate(pointcol=ifelse(sign(slope)==1, "Anticyclone enriched", "Cyclone enriched")) %>%
-  mutate(depth=factor(depth, levels=c("Surface", "DCM", "Deep"))) %>%
-  mutate(name=factor(name, levels=metadata_levels, labels=metadata_short_labels))
-ggplot() +
-  geom_abline(aes(intercept=estimate_intercept, slope=estimate_slope), data = clean_slopes) +
-  geom_point(aes(x=sla_corr, y=norm_area), data=unnest(clean_slopes, data)) +
-  geom_abline(data = metadata_slopes, aes(slope=slope, intercept=intercept, color=name)) +
-  geom_point(aes(x=sla_corr, y=value, color=name), data=unnest(metadata_slopes, data)) +
-  facet_grid(pointcol~depth) +
-  ggtitle("Todo: fix this graph")
-ggsave("supplement/many_lin_reg_by_SLA.png", device = "png", width = 6.875, height = 5.5, units = "in", dpi = 300, bg="white")
-
-```
-
-![](supplement/many_lin_reg_by_SLA.png)
-
-*Supplemental figure? Linear regressions of rank-order peak area against SLA showing strong enrichment at the cyclone DCM and the anticyclone 175 meter samples.*
-
-```{r possible figure stacked barplot ranked space}
-top10_stacked_area_station_gp <- map(c("Surface", "DCM", "Deep"), function(depth_i){
-  reg_coefs %>%
-    filter(str_detect(term, "sla_corr")) %>%
-    mutate(p.value=p.adjust(p.value, method="fdr")) %>%
-    filter(depth==depth_i) %>%
-    arrange(p.value) %>%
-    group_by(estimate>0) %>%
-    slice(1:5) %>%
-    ungroup() %>%
-    left_join(untarg_feats %>% select(feature, compound_name, mzmed, rtmed)) %>%
-    mutate(feature=ifelse(!is.na(compound_name), compound_name, 
-                          paste0(round(mzmed, 3), "_", round(rtmed, 1)))) %>%
-    mutate(feature=fct_inorder(feature)) %>%
-    unnest(data) %>%
-    select(feature, depth, norm_area, filename, station) %>%
-    group_by(feature) %>%
-    # mutate(norm_area=rank(norm_area)) %>%
-    mutate(norm_area=norm_area/sum(norm_area)) %>%
-    mutate(station=factor(station, levels=as.character(4:14))) %>%
-    ggplot() +
-    geom_col(aes(x=station, y=norm_area, fill=feature), position = "fill") +
-    scale_y_continuous(expand = c(0,0)) +
-    labs(x="Station", y="Relative peak area", fill="Mass feature") +
-    theme_bw()
-}) %>% gridExtra::arrangeGrob(grobs = ., ncol = 1)
-ggsave("top10_stacked_area_station_gp.png", plot = top10_stacked_area_station_gp, 
-       device = "png", path = "supplement", width = 6.875, height = 9, 
-       units = "in", dpi = 300)
-```
-
-![](supplement/top10_stacked_area_station_gp.png)
-
-*Supplemental figure?
-Plots of the 5 most significantly enriched metabolites in both the cyclone (top 5) and anticyclone (bottom 5) after normalizing to the total peak area in each mass feature.
-Nicely shows a stronger trend with increasing depth but feels a little bit like lying because there's 100 other compounds that don't change in between these extremes.
-Tyrosine is also misannotated here.*
-
-```{r comparison to Katherines culture data}
-cyc_enriched <- pairwise_tests %>%
-  filter(p.value<0.05) %>%
-  left_join(untarg_feats) %>%
-  filter(!is.na(compound_name)) %>%
-  filter(statistic>110)
-
-enriched_metabs <- c("Homarine", "Trigonelline", "Taurine", "DMSP",
-                     "Dimethylsulfonioacetate (DMS-Ac)", "Isethionic Acid")
-culture_metadata <- "../metadata/raw_data/Heal_Metabs_SuppTabs_corrected20230327.xlsx" %>%
-  readxl::read_excel(sheet = "S7_Culturedescriptions_full")
-culture_areas <- "../metadata/raw_data/Heal_Metabs_SuppTabs_corrected20230327.xlsx" %>%
-  readxl::read_excel(sheet = "S5_FullEnviroCulture_PeakAreas") %>%
-  filter(Identification%in%enriched_metabs)
-katherine_comp_mostcycenriched_gp <- culture_areas %>%
-  select(Identification, any_of(culture_metadata$`Sample ID`)) %>%
-  mutate(Identification=ifelse(Identification=="Dimethylsulfonioacetate (DMS-Ac)", 
-                               "DMS-Ac", Identification)) %>%
-  mutate(Identification=factor(Identification, levels=c(
-    "Homarine", "Trigonelline", "Isethionic Acid", "Taurine", "DMSP","DMS-Ac"
-  ))) %>%
-  pivot_longer(-Identification, names_to = "Sample ID") %>%
-  mutate(value=as.numeric(value)) %>%
-  left_join(culture_metadata, by=join_by(`Sample ID`)) %>%
-  arrange(`Broad taxon`) %>%
-  mutate(`Short ID`=fct_inorder(`Short ID`)) %>%
-  mutate(`Species`=fct_inorder(`Species`)) %>%
-  drop_na() %>%
-  ggplot() +
-  geom_boxplot(aes(x=Species, y=value, color=`Broad taxon`)) +
-  facet_wrap(~Identification, ncol = 2, scales="free_y") +
-  scale_y_log10(name="Peak area") +
-  theme_bw() +
-  theme(axis.text.x = element_text(angle=90, hjust=1, vjust=0.5))
-
-ggsave("katherine_comp_mostcycenriched_gp.png", plot = katherine_comp_mostcycenriched_gp, 
-       device = "png", path = "supplement", width = 6.875, height = 5, units = "in", dpi = 300)
-```
-
-![](supplement/katherine_comp_mostcycenriched_gp.png)
-
-*Supplemental figure? Culture data from Katherine's manuscript showing the presence of the DCM cyclone-enriched metabolites mostly present in diatoms and dinoflagellates.*
-
-```{r enrichment plot DCM comparison, eval=FALSE}
-MC_stacked_mostdiff_gp <- pairwise_tests %>%
-  arrange(p.value) %>%
-  group_by(statistic>110) %>%
-  slice(1:5) %>%
-  ungroup() %>%
-  # filter(p.value<0.00005) %>%
-  left_join(untarg_feats %>% select(feature, compound_name), by = join_by(feature)) %>%
-  mutate(compound_name=ifelse(is.na(compound_name), feature, compound_name)) %>%
-  mutate(compound_name=ifelse(compound_name=="Dimethylsulfoniopropionate", "DMSP", compound_name)) %>%
-  arrange(statistic) %>%
-  mutate(compound_name=fct_inorder(compound_name)) %>%
-  select(feature, compound_name) %>%
-  left_join(untarg_peaks %>% select(filename, norm_area, feature), by="feature") %>%
-  left_join(filled_file_metadata %>% distinct(filename, station, abs_depth), by="filename") %>%
-  filter(station%in%c("L1", "L2")) %>%
-  arrange(station) %>%
-  mutate(station=ifelse(station=="L1", "Cyclone", "Anticyclone")) %>%
-  mutate(station=fct_inorder(station)) %>%
-  arrange(abs_depth) %>%
-  mutate(filename=fct_inorder(filename)) %>%
-  group_by(feature) %>%
-  # mutate(norm_area=rank(norm_area)) %>%
-  mutate(norm_area=norm_area/sum(norm_area)) %>%
-  ungroup() %>%
-  ggplot() +
-  geom_col(aes(x=filename, y=norm_area, fill=compound_name, group=filename), 
-           position = "fill", color="grey50", width=1) +
-  geom_rect(aes(xmin=-Inf, xmax=Inf, ymin=1, ymax=1.1), 
-            fill=c("#8277b6", "#b05d56"), color="black",
-            data = data.frame(station=fct_inorder(c("Cyclone", "Anticyclone")))) +
-  geom_text(aes(x=8, y=1.05, label=station), hjust=0.5,
-            data = data.frame(station=fct_inorder(c("Cyclone", "Anticyclone")))) +
-  facet_grid(~station, scales="free_x") +
-  scale_fill_manual(values = splividis(10, begin = 0.1, end = 0.95)) +
-  guides(fill=guide_legend(reverse = TRUE)) +
-  scale_y_discrete(expand = c(0,0)) +
-  theme_bw() +
-  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(),
-        strip.background = element_blank(), strip.text = element_blank()) +
-  labs(x=NULL, y="Relative peak area", fill="Mass feature")
-ggsave("MC_stacked_mostdiff_gp.png", plot = MC_stacked_mostdiff_gp, device = "png", 
-       path = "supplement", width = 6.875, height = 3, units = "in", dpi = 300)
-```
-
-![](supplement/MC_stacked_mostdiff_gp.png)
-
-*Supplemental figure?
-Particulate matter composition differences between the cyclone and anticyclone when considering only the top 5 most different metabolites enriched in each eddy separately.
-Each sample corresponds to one multicolored vertical bar while metabolites correspond to the colors provided in the legend and have been scaled to the sum of the total peak areas for that metabolite in the sample set.*
+SLA trends are visible in the 25 meter samples, with dark blue circles consistently discriminating from the dark red circles.
+The bottom row of plots show the direction and magnitude of this effect by plotting the grand mean of the normalized metabolite peak areas in the stations taken between the adjacent eddies of opposite polarity.*

### 2024-05-27 manuscript/manuscript.Rmd
Making trivial edits, noting where additional thought needed.

@@
 This study reports the exact concentration of 67 metabolites and relative abundances for 640 molecular features to understand how mesoscale eddies impact the metabolome of the North Pacific Subtropical Gyre during two cruises in 2017 and 2018.
-We find that many metabolites track biomass trends, but biomarker metabolites like isethionic acid, homarine, and trigonelline linked to eukaryotic phytoplankton and enriched at the deep chlorophyll maximum of the cyclonic features, while degradation products such as arsenobetaine were enriched in anticyclones.
-By analyzing depth variability (accounting for 20-40% of metabolomic variability across ~150 meters) and eddy pumping (explaining 10-20% of variability across a sea level anomaly range of 40 centimeters), this analysis constrains the importance of mesoscale eddies in modeling the chemical composition of particulate matter in the largest biomes on the planet.
+We find that many metabolites track biomass trends, but biomarker metabolites like isethionic acid, homarine, and trigonelline linked to eukaryotic phytoplankton were enriched at the deep chlorophyll maximum of the cyclonic features, while degradation products such as arsenobetaine were enriched in anticyclones.
+In every analysis, the metabolites with the strongest responses were detected via untargeted mass spectrometry, indicating that the molecules most sensitive to environmental perturbation were not among the core metabolites of the cell.
+By analyzing depth variability (accounting for 20-40% of metabolomic variability across ~150 meters) and the vertical displacement of isopycnal surfaces (explaining 10-20% of variability across a sea level anomaly range of 40 centimeters), this analysis constrains the importance of mesoscale eddies in shaping the chemical composition of particulate matter in the largest biomes on the planet.
 
@@
 
-High frequency, long-term observations at Station ALOHA in the North Pacific Subtropical Gyre (NPSG) over the past 25 years have revealed temporal and spatial variability in what had previously been considered a homogenous environment [@Karl1999; @Karl2017; @Karl2021].
+High frequency, long-term observations at Station ALOHA in the North Pacific Subtropical Gyre (NPSG) over the past 25 years have revealed temporal and spatial variability in what had previously been considered a relatively homogenous environment [@Karl1999; @Karl2017; @Karl2021].
 A major source of variability comes in the form of mesoscale eddies in which water is entrained into circular surface currents tens to hundreds of kilometers in diameter [@Karl2017; @McGillicuddy2016].
 
-Eddies are observed via satellite altimetry, which measures anomalies in sea surface height.
-Those that have a positive sea level anomaly (SLA) typically indicate regions of downwelling as water converges and are associated with clockwise rotation in the Northern Hemisphere.
-In contrast, a negative SLA usually corresponds to counterclockwise rotation in the Northern Hemisphere upwelling as deep water is lifted into the sunlit region of the ocean to replace water forced away from the eddy center [@McGillicuddy2016].
-While mode-water eddies have been observed and are an exception to the conventions established above [@Sweeney2003; @McGillicuddy2007], here we focus on the more common cyclonic (negative SLA) and anticyclonic (positive SLA) mesoscale features near station ALOHA [@Barone2019].
-
-The uplift of deep, nutrient-rich seawater into the euphotic zone disrupts microbial communities [@Rii2022].
-Measurements of chlorophyll in cyclones reveal a shallower and more intense maximum [@Barone2022; @Cornec2021] as a result of eukaryotic phytoplankton growing in the increased nutrient concentrations while cyanobacterial biomass is reduced [@Hawco2021].
+Eddies can be observed via satellite altimetry, which measures anomalies in sea surface height.
+Those that have a positive sea level anomaly (SLA) typically indicate regions where deep water layers and isopycnal surfaces are depressed and are associated with clockwise rotation in the Northern Hemisphere.
+In contrast, a negative SLA corresponds to counterclockwise rotation in the Northern Hemisphere and the uplift of deep water layers into the sunlit region of the ocean [@McGillicuddy2016].
+Mode-water eddies are an exception to the conventions established above [@Sweeney2003; @McGillicuddy2007] but here we only focus on the cyclonic (negative SLA) and anticyclonic (positive SLA) mesoscale features that are commonly observed near station ALOHA [@Barone2019].
+
+The uplift of deep, nutrient-rich seawater into the euphotic zone alters microbial communities [@Rii2022].
+Measurements of chlorophyll in cyclones reveal a shallower and more intense maximum [@Barone2022; @Cornec2021] as a result of eukaryotic phytoplankton thriving in the increased nutrient concentrations while cyanobacterial biomass is reduced [@Hawco2021].
 This growth of large eukaryotes corresponds to a net increase in biomass and primary productivity [@Benitez-Nelson2007], especially near the deep chlorophyll maximum [@Barone2022].
-In contrast, anticyclones produce conditions favorable to nitrogen fixation by accumulating organisms such as Crocosphaera and Trichodesmium [@Fong2008; @Olson2015; @Dugenne2020; @Dugenne2023].
+In contrast, anticyclones produce conditions favorable to nitrogen fixation by accumulating diazotrophs such as *Crocosphaera* and *Trichodesmium* [@Fong2008; @Olson2015; @Dugenne2020; @Dugenne2023].
 
@@
 These results raise questions about the effect of mesoscale features on particulate organic matter composition.
+[@Wei: You probably want to add several sentences here to briefly summarize what you did in this work to fill the gap.
+Often, you can propose a major hypothesis and mention the methods you used to analyze the samples to test your hypothesis.]
 
@@
 The first eddy pair was the focus of the July 2017 Microbial Ecology of the Surface Ocean-Simons Collaboration on Ocean Processes and Ecology (MESO-SCOPE, KM1709 aboard the R/V *Kilo Moana*) cruise and sampling consisted of two phases (Figure 1).
-In the first phase, a transect was taken from the north to the south consisting of 11 stations corresponding to the sampling sites in @Barone2022 and @Dugenne2023.
+In the first phase, a transect was taken from the north to the south consisting of 11 stations across a pair of eddies, corresponding to the sampling sites in @Barone2022 and @Dugenne2023.
 Metabolomic samples of seawater from 15 meters, the deep chlorophyll maximum (DCM), and 175 meters were collected at each station via 0.2 µm filter.
-The next phase of the MESO-SCOPE cruise focused on Lagrangian sampling of the eddy centers, starting with the cyclone (SLA = -14 cm) before progressing to the anticyclone (SLA = 24 cm).
+The next phase of the MESO-SCOPE cruise focused on Lagrangian sampling in the two eddy centers, starting with the cyclone (SLA = -14 cm) before progressing to the anticyclone (SLA = 24 cm).
 There, metabolomics samples were taken from the DCM as well as 10 and 20 meters above and below it.
@@
 Cruise bounds are shown in the large central map with Station ALOHA colored as a point in red near the Hawaiian Islands.
-Yellow stars denote sampling locations and station numbers relative to the sea level anomaly contours in the background during both the June 2017 MESO-SCOPE cruise (left) and the April 2018 HEE cruise (right).*
+Yellow stars denote sampling locations and station numbers relative to the sea level anomaly contours in the background during 28 June 2017 for the MESO-SCOPE cruise (left) and the 6 April 2018 for the HEE cruise (right).
+Lagrangian sampling near the eddy centers of the MESO-SCOPE cruise was performed by following drifters deployed in proximity of Stations 6 and 12.*
 
@@
 Untargeted analysis was performed using the R package XCMS with each feature manually inspected for quality assurance.
-
-## Metabolome variability across adjacent mesoscale eddies of opposing polarity during MESO-SCOPE explained by variations in sea level anomaly
+[@Wei: I think this paragraph belongs to Methods section.
+@Will: Maybe just shorten it extensively?
+Ask Anitra for advice since Rachel wanted this here.]
+
+## Metabolome variability across adjacent mesoscale eddies of opposing polarity during MESO-SCOPE explained by sea level anomaly variations
 
 The metabolome clearly differed between cyclonic and anticyclonic samples, though the magnitude of this difference varied by depth.
-The largest differences in particulate matter composition were detected at the DCM and 175 meter samples, as shown by the NMDS in Figure 2.
+The largest differences in particulate matter composition were detected in the DCM and 175 meter samples, as shown by the NMDS in Figure 2.
 At the DCM, approximately 16% of the variation in the metabolome could be explained by sea level anomaly (SLA) alone (PERMANOVA, p = 0.003).
 Additionally, SLA was strongly correlated with the first principal component (PC) of the metabolome  with a Pearson's *r* of .615 and 27% of the variance explained by PC1 (Supplemental figure 1A), indicating that this was one of the largest sources of variation in the dataset.
-Notably, the samples taken from the exact center of the cyclonic eddy at Station 12 were highly distinct and likely drove much of the variance explained.
+Notably, the samples taken from the exact center of the cyclonic eddy at Station 12 were highly distinct and likely drove much of the explained variance.
 In the 175 meter samples, 12% of the variance was explained by SLA (PERMANOVA, p = 0.008) and SLA was highly correlated with the second PC of the dataset (r = .697, % variance explained by PC2 = 15.4, Supplemental figure 1A), though the first PC did not seem to have any visible pattern with metadata and appeared to largely capture variation between biological triplicates (Supplemental figure 1B).
@@
 The DCM and 175 meter samples had the stronger trends ($\beta_{DCM} = 22.6\pm5.53$, $r= .593$, p-value < 0.001; $\beta_{175m} = 23.9\pm7.36$, $r= .503$, p-value = 0.003) while 15 meter samples showed no significant relationship ($\beta_{15m} = 3.4\pm12.9$, $r= .049$, p-value = 0.792).
-This meant that metabolites contributed a fixed fraction of carbon to the particulate pool with our 53 quantified metabolites representing between 1.5% and 5% of the carbon in the system (Figure 2).
+This meant that metabolites contributed a fixed fraction of carbon to the particulate pool with our 53 quantified metabolites representing between 1.5% and 5% of the carbon in the system (Figure 2) [@Benedetto: I find this very informative.
+It would be great to get a ballpark sense of the percentages including the unknown metabolites, but I understand that it might be too uncertain considering the required assumptions on molecular formulas and concentration.
+I would like to know your wild guess, though].
 Particulate nitrogen was tightly correlated with particulate carbon ($r=.933$) but the fraction of particulate nitrogen represented by the quantified metabolites was slightly higher across the transect with contributions typically between 2% and 6%.
@@
 A majority of the known molecules (37/53) had a similar pattern with a subsurface maximum at the DCM, including glycine betaine, glutamate, and guanine.
-13 molecules obeyed a different pattern and decreased monotonically with depth, including the molecules hydroxyisoleucine, gonyol, and dimethylsulfoniopropionate.
+Thirteen molecules obeyed a different pattern and decreased monotonically with depth, including the molecules hydroxyisoleucine, gonyol, and dimethylsulfoniopropionate.
 Two known molecules (arsenobetaine and O-acetylcarnitine) increased with depth, potentially representing particle degradation during export.
@@
 In addition to absolute shifts in metabolite concentration due to biomass differences, we also explored compositional shifts in the metabolomes by assessing the fraction of the total metabolite pool contributed by each compound across the eddy transect.
-This allowed us to isolate the signal due to shifts in community composition and organismal response from those introduced by net growth of the community as a whole.
+This allowed us to isolate the signal due to shifts in community composition and organismal response from those introduced by changes in biomass of the community as a whole.
 
@@
 Both O-acetylcarnitine and betonicine effectively quintupled their contribution to total peak area, shifting from approximately 0.1% to ~0.5% over the transect.
-Acetylcholine was the only known compound more relatively abundant in the cyclone than the anticyclone along with one unknown of *m/z* 131.0340.
-The unknown at *m/z* 173.0921 with the strongest trend at the DCM again had the strongest trend in the 175 meter samples, here with an FDR-corrected p-value of $2.0*10^{-7}$.
+Acetylcholine was the only known compound more relatively abundant in the cyclone than in the anticyclone along with one unknown of *m/z* 131.0340.
+The unknown at *m/z* 173.0921 with the strongest trend at the DCM again had the strongest trend in the 175 meter samples, here with an FDR-corrected p-value of $2.0 × 10^{-7}$.
 
@@
 
-Given the strength of the metabolomic response to SLA at the DCM, we additionally performed high-resolution depth sampling around the DCM at the center of both eddy poles.
+Given the strength of the metabolomic response to SLA at the DCM [@Benedetto: This seems to imply that you knew this before sampling], we additionally performed high-resolution depth sampling around the DCM at the center of both eddy poles.
 Here, we found that the ~40 meter depth range and the transition between the eddies explained similar amounts of variation in the data (R$^2_{depth}=0.253$, R$^2_{SLA}=0.238$) and were both highly significant factors (permutational p-values << 0.001, Figure 4).
@@
 
-Given the strong signals detected in the samples from the MESO-SCOPE cruise both across the eddy transect as well as in the high-resolution DCM sampling, we expected to find some of the same results in an analogous dataset.
-Samples were collected during a 2018 cruise on the R/V *Falkor* that again targeted both a cyclonic and anticyclonic eddy in the North Pacific Subtropical Gyre near Station ALOHA as part of the Hawaiian Eddy Experiment (HEE).
+Given the strong signals detected in the samples from the MESO-SCOPE cruise both across the eddy transect as well as in the high-resolution DCM sampling, we expected to find similar results in an analogous dataset.
+Samples were collected during a 2018 cruise on the R/V *Falkor* that again targeted both a cyclonic and an anticyclonic eddy in the North Pacific Subtropical Gyre near Station ALOHA as part of the Hawaiian Eddy Experiment (HEE).
 
@@
 The eddy effects, on the other hand, were much less strong during this cruise.
-Of the six known metabolites with the strongest differences across the eddy transect at the DCM in the MESO-SCOPE cruise, only isethionic acid was significantly different in the HEE dataset (isethionic acid t-test p-value$_{FDR}$ = 0.043, other five p-values > 0.25).
+Of the six known metabolites with the strongest differences across the eddy transect at the DCM in the MESO-SCOPE cruise, only isethionic acid was significantly different in the HEE dataset (isethionic acid t-test p-value$_{FDR}$ = 0.043, other five p-values > 0.25) [@Wei: If P values are not significant, you can calculate effect size.].
 This difference in isethionic acid concentration was only found after normalizing each sample to the sum of all metabolites in the sample to control for biomass.
@@
 We also detected the largest differences in the DCM metabolome at the exact center of the eddy, with Station 12 highly distinct from even the nearest stations, while no such stark difference was found in the anticyclone (Figures 1 and 2).
-This perhaps indicates that the exact center of the cyclone has a unique metabolome relative at the DCM relative to the rest of the transect rather than existing as a smooth continuation.
+This perhaps indicates that the exact center of the cyclone has a unique metabolome at the DCM relative to the rest of the transect, rather than existing as a smooth continuation.
 
@@
 This revealed four major patterns of metabolite response, with a majority of compounds responding to eddy state (Clusters 1, 2, and 4 in Figure 5).
-It is interesting to note that the sole cluster in which abundances in the anticyclone are greater than those of the cyclone is also the only cluster to generally increase in abundance with depth (Cluster 4), perhaps indicating that compounds more concentrated in the anticyclone are the same kind of degradation products and recalcitrant carbon that is found at depth.
+It is interesting to note that the sole cluster in which abundances in the anticyclone are greater than those of the cyclone is also the only cluster to generally increase in abundance with depth (Cluster 4), perhaps indicating that compounds more concentrated in the anticyclone are the same kind of degradation products and recalcitrant carbon typically found at depth.
 Cluster 3, which shows no clear SLA difference and only a slight change with depth, includes compounds which are either unaffected by the environment or simply lack an environmental signal due to large inter-sample variability.
 
-The Hawaiian Eddy Experiment (HEE) data upset many of the expectations we developed during the MESO-SCOPE cruise in the previous year.
+The Hawaiian Eddy Experiment (HEE) data upset many of the expectations [@Wei: consider calculating the effect size, as mentioned above.
+@Will: Not sure what Wei means by this.
+Should I do effect size calculations here?
+Or just above when we didn't see an effect?] we developed during the MESO-SCOPE cruise in the previous year.
 Most surprising was the large difference between the eddies detected at the *surface*, while the DCM samples were functionally indistinguishable.
@@
 This molecule has a well-established role as an osmolyte in eukaryotic phytoplankton and cyanobacteria [@Gebser2013; @Dawson2020; @Heal2021; @Durham2022] as well as a documented decrease in abundance with depth [@Heal2021], indicating that its response to SLA is potentially due to differences in physical and chemical attributes that in turn control biomass, community composition, and recycling rates.
-Homarine was also tightly correlated with trigonelline, and although they have structural similarity they are not known to have a relationship beyond their shared function as osmolytes.
+Curiously, the distribution of homarine was also tightly correlated with that of trigonelline.
+Although they have structural similarity they are not known to have a relationship beyond their shared function as osmolytes.
 
@@
 The strong SLA signal detected among the 175 meter samples was partially surprising to us, having expected the effect to be strongest at the DCM where eddy effects lift large concentrations of nutrients above the 1% light level.
-Instead, we found more compounds overall to be significantly different across the eddy transect in the 175 meter samples than we did at the DCM (α=0.05, 43 compounds at 175 meters vs 22 at the DCM).
+Instead, we found more compounds overall to be significantly different across the eddy transect in the 175 meter samples than we did at the DCM (α=0.05, 43 compounds at 175 meters vs 22 at the DCM) [@Benedetto: Notice that the analyses of SLA at Station ALOHA in Barone et al.
+(2019) clearly shows enhancement of organic matter and heterotrophic picoplankton below the DCM].
 The enrichment of nucleobases in the 175 meter samples may hint at bacterial biomass, as nucleobases tend to be highly abundant in bacterial cultures [@Heal2021].
@@
 Metabolites offer a way to bridge this gap in knowledge but the relative impact of eddies on their contribution and fate was not previously estimated.
-In this manuscript, we have shown the sensitivity of the metabolome to various environmental and community factors and contrasted it to the well-established effect of depth to constrain the importance of including eddy effects in carbon cycle models.
+In this work, we have shown the sensitivity of the metabolome to various environmental and community factors and contrasted it to the well-established effect of depth to constrain the importance of including eddy effects in models of elemental cycling.
 
@@
 High-resolution depth sampling of the DCM at the center of each eddy elucidated several known and unknown biomarkers likely corresponding to the previously documented increase in eukaryotic phytoplankton in the cyclone.
-Metabolite clusters aligned well with expected trends in abundance with depth and between the two eddies, with rapid convergence below the DCM towards a general deep-sea metabolite signal.
+Metabolite clusters aligned well with expected trends in abundance with depth and between the two eddies, with rapid convergence below the DCM towards a general deep-sea metabolite signal [@Benedetto: with rapid convergence below the DCM towards a general deep-sea metabolite signal.
+@Will: Yeah, you right.
+I'll have to rephrase this.].
 By contrasting these effects with a follow-up analysis in the centers of a separate pair of mesoscale eddies of opposite polarity the following year, we learned that the impact of eddies may be influenced by their origin, age, and period of the year.
-Finally, the metabolites with the strongest responses to eddy effects were unknown and a relatively smaller proportion of known metabolites in the anticyclone demonstrates the utility of untargeted metabolomics in exploring environmental variation and the large amount of information that can still be gleaned from such datasets.
+Finally, the metabolites with the strongest responses to eddy effects were unknown, and a relatively smaller proportion of known metabolites in the anticyclone demonstrates the utility of untargeted metabolomics in exploring environmental variation and the large amount of information that can still be gleaned from such datasets.
 
@@
 The transect samples were taken at various times of day as the ship transected the adjacent eddies while the eddy center samples were all collected between 5 and 8 pm.
-In 2018, the Hawaiian Eddy Experiment (HEE, FK180310 on the *R/V Falkor*) targeted a new cyclonic and anticyclonic eddy in approximately the same location with samples taken at the center of each eddy (maximum negative anomaly in the cyclone = -15 cm, maximum positive anomaly in the anticyclone = +26 cm) (Figure 1).
+In 2018, the Hawaiian Eddy Experiment (HEE, FK180310 on the *R/V Falkor*) targeted new cyclonic and anticyclonic eddies in approximately the same location with samples taken at the center of each eddy (maximum negative anomaly in the cyclone = -15 cm, maximum positive anomaly in the anticyclone = +26 cm) (Figure 1).
 
@@
 Briefly, all environmental data was collected via CTD rosette except SLA which was measured via satellite.
-Particulate carbon and particulate nitrogen were measured using an elemental analyzer, and nitrate + nitrite (N+N) and soluble reactive phosphorus (phosphate, PO$_4^{3-}$) were measured on an autoanalyzer except where concentrations were below 100 nM in which case they were measured using chemiluminescence.
-Values were interpolated linearly to the locations at which metabolite samples were taken using all data collected at the given station.
+Particulate carbon and particulate nitrogen were measured using an elemental analyzer.
+Nitrate + nitrite (N+N) was measured on an autoanalyzer except where concentrations were below 100 nM in which case they were measured using chemiluminescence.
+Soluble reactive phosphorus (phosphate, PO$_4^{3-}$) was measured on an autoanalyzer or following the magnesium induced coprecipitation method.
+Values were interpolated linearly to the depths at which metabolite samples were taken using all data collected at the given station.
 Two particulate carbon values were determined to be spurious, with concentrations 2-3 times higher than typical Station ALOHA values, and were instead estimated using a best-fit linear regression against beam attenuation.

### 2024-05-30 manuscript/manuscript.Rmd
Making some coauthor edits before Anitra's review

@@
 These results raise questions about the effect of mesoscale features on particulate organic matter composition.
-[@Wei: You probably want to add several sentences here to briefly summarize what you did in this work to fill the gap.
-Often, you can propose a major hypothesis and mention the methods you used to analyze the samples to test your hypothesis.]
+
+In this work, we directly tested for differences in the chemical composition of particulate matter due to changes in eddy state in the NPSG.
+We sampled from cyclonic and anticyclonic eddies spatially near each other and used targeted and untargeted mass spectrometry to explore changes in the composition of small, polar molecules.
+We expected to find that overall metabolite abundance would reflect shifts in biomass over multiple depths.
+We also expected to see enrichment in the deep chlorophyll maximum of cyclonic features for those metabolites especially abundant in eukaryotic organisms.
+Finally, we predicted that these patterns would be robust across years and sampling regimes for a reliable way to link these ocean features with the chemical composition of organic matter in the open ocean.
 
@@
 Untargeted analysis was performed using the R package XCMS with each feature manually inspected for quality assurance.
-[@Wei: I think this paragraph belongs to Methods section.
-@Will: Maybe just shorten it extensively?
+[Wei: I think this paragraph belongs to Methods section.
+Will: Maybe just shorten it extensively?
 Ask Anitra for advice since Rachel wanted this here.]
@@
 The DCM and 175 meter samples had the stronger trends ($\beta_{DCM} = 22.6\pm5.53$, $r= .593$, p-value < 0.001; $\beta_{175m} = 23.9\pm7.36$, $r= .503$, p-value = 0.003) while 15 meter samples showed no significant relationship ($\beta_{15m} = 3.4\pm12.9$, $r= .049$, p-value = 0.792).
-This meant that metabolites contributed a fixed fraction of carbon to the particulate pool with our 53 quantified metabolites representing between 1.5% and 5% of the carbon in the system (Figure 2) [@Benedetto: I find this very informative.
+This meant that metabolites contributed a fixed fraction of carbon to the particulate pool with our 53 quantified metabolites representing between 1.5% and 5% of the carbon in the system (Figure 2) [Benedetto: I find this very informative.
 It would be great to get a ballpark sense of the percentages including the unknown metabolites, but I understand that it might be too uncertain considering the required assumptions on molecular formulas and concentration.
@@
 
-Given the strength of the metabolomic response to SLA at the DCM [@Benedetto: This seems to imply that you knew this before sampling], we additionally performed high-resolution depth sampling around the DCM at the center of both eddy poles.
+Given the strength of the metabolomic response to SLA at the DCM , we additionally analyzed samples taken with high-resolution depth around the DCM at the center of both eddy poles.
 Here, we found that the ~40 meter depth range and the transition between the eddies explained similar amounts of variation in the data (R$^2_{depth}=0.253$, R$^2_{SLA}=0.238$) and were both highly significant factors (permutational p-values << 0.001, Figure 4).
@@
   group_by(filename) %>%
-  # mutate(nM=nM/sum(nM)) %>%
+  mutate(nM=nM/sum(nM)) %>%
   ungroup() %>%
   filter(depth=="DCM") %>%
-  # nest(data=-compound_name) %>%
-  # mutate(aovout=map(data, ~broom::tidy(aov(.x$nM~.x$sla_class)))) %>%
-  # unnest(aovout) %>% filter(term==".x$sla_class") %>%
-  # mutate(p.value=p.adjust(p.value, method="fdr"))
+  nest(data=-compound_name) %>%
+  mutate(aovout=map(data, ~broom::tidy(aov(.x$nM~.x$sla_class)))) %>%
+  unnest(aovout) %>% filter(term==".x$sla_class") %>%
+  mutate(p.value=p.adjust(p.value, method="fdr"))
   ggplot(aes(x=sla_class, y=nM)) +
@@
 The eddy effects, on the other hand, were much less strong during this cruise.
-Of the six known metabolites with the strongest differences across the eddy transect at the DCM in the MESO-SCOPE cruise, only isethionic acid was significantly different in the HEE dataset (isethionic acid t-test p-value$_{FDR}$ = 0.043, other five p-values > 0.25) [@Wei: If P values are not significant, you can calculate effect size.].
+Of the six known metabolites with the strong enrichment in the cyclone at the DCM in the MESO-SCOPE cruise, only isethionic acid was significantly different in the HEE dataset (isethionic acid t-test p-value$_{FDR}$ = 0.043, other five p-values > 0.25).
 This difference in isethionic acid concentration was only found after normalizing each sample to the sum of all metabolites in the sample to control for biomass.
@@
 
-The Hawaiian Eddy Experiment (HEE) data upset many of the expectations [@Wei: consider calculating the effect size, as mentioned above.
-@Will: Not sure what Wei means by this.
+The Hawaiian Eddy Experiment (HEE) data upset many of the expectations [Wei: consider calculating the effect size, as mentioned above.
+Will: Not sure what Wei means by this.
 Should I do effect size calculations here?
@@
 
-The strong SLA signal detected among the 175 meter samples was partially surprising to us, having expected the effect to be strongest at the DCM where eddy effects lift large concentrations of nutrients above the 1% light level.
-Instead, we found more compounds overall to be significantly different across the eddy transect in the 175 meter samples than we did at the DCM (α=0.05, 43 compounds at 175 meters vs 22 at the DCM) [@Benedetto: Notice that the analyses of SLA at Station ALOHA in Barone et al.
-(2019) clearly shows enhancement of organic matter and heterotrophic picoplankton below the DCM].
-The enrichment of nucleobases in the 175 meter samples may hint at bacterial biomass, as nucleobases tend to be highly abundant in bacterial cultures [@Heal2021].
+The strong SLA signal detected among the 175 meter samples was partially surprising to us given our expectation about the strongest effect at the DCM where eddy effects lift large concentrations of nutrients above the 1% light level.
+Instead, we found more compounds overall to be significantly different across the eddy transect in the 175 meter samples than we did at the DCM (α=0.05, 43 compounds at 175 meters vs 22 at the DCM).
+This was likely driven by the enhancement of organic matter and heterotrophic picoplankton below the DCM as seen in @Barone2019, with the enrichment of nucleobases in the 175 meter samples additionally hint at increases in bacterial biomass given that nucleobases tend to be highly abundant in bacterial cultures [@Heal2021].
 The abundance of a mass feature putatively identified as acetylcholine in the 175 meter cyclone samples followed the same general trend as arsenobetaine, both enriched at depth and in the anticyclone, and may be another compound that results from heterotrophic degradation of phytoplankton metabolites [@Durham2022].
@@
 High-resolution depth sampling of the DCM at the center of each eddy elucidated several known and unknown biomarkers likely corresponding to the previously documented increase in eukaryotic phytoplankton in the cyclone.
-Metabolite clusters aligned well with expected trends in abundance with depth and between the two eddies, with rapid convergence below the DCM towards a general deep-sea metabolite signal [@Benedetto: with rapid convergence below the DCM towards a general deep-sea metabolite signal.
-@Will: Yeah, you right.
-I'll have to rephrase this.].
+Metabolite clusters aligned well with expected trends in abundance with depth and between the two eddies, with convergence below the DCM towards a general deep-sea metabolite signal for many metabolites.
 By contrasting these effects with a follow-up analysis in the centers of a separate pair of mesoscale eddies of opposite polarity the following year, we learned that the impact of eddies may be influenced by their origin, age, and period of the year.
-Finally, the metabolites with the strongest responses to eddy effects were unknown, and a relatively smaller proportion of known metabolites in the anticyclone demonstrates the utility of untargeted metabolomics in exploring environmental variation and the large amount of information that can still be gleaned from such datasets.
+Finally, the metabolites with the strongest responses to eddy effects were unknown, and a relatively smaller proportion of known metabolites in the anticyclone demonstrates the utility of untargeted metabolomics in exploring environmental variation and the large amount of information that is missed by only investigating known compounds.
 
@@
   guides(fill=guide_legend(nrow=2, byrow = TRUE)) +
+  labs(y=NULL) +
   ggtitle("B) Relative composition of the metabolome")

### 2024-06-26 manuscript/manuscript.Rmd
Making initial changes from Anitra

@@
 Previous studies have explored eddy effects on biogeochemistry and microbial community composition but not on the molecular composition of particulate organic matter.
-This study reports the exact concentration of 67 metabolites and relative abundances for 640 molecular features to understand how mesoscale eddies impact the metabolome of the North Pacific Subtropical Gyre during two cruises in 2017 and 2018.
+This study reports the absolute concentration of 67 metabolites and relative abundances for 640 molecular features to understand how mesoscale eddies impact the metabolome of the North Pacific Subtropical Gyre during two cruises in 2017 and 2018.
 We find that many metabolites track biomass trends, but biomarker metabolites like isethionic acid, homarine, and trigonelline linked to eukaryotic phytoplankton were enriched at the deep chlorophyll maximum of the cyclonic features, while degradation products such as arsenobetaine were enriched in anticyclones.
-In every analysis, the metabolites with the strongest responses were detected via untargeted mass spectrometry, indicating that the molecules most sensitive to environmental perturbation were not among the core metabolites of the cell.
+In every analysis, metabolites with the strongest responses were detected via untargeted mass spectrometry, indicating that the molecules most sensitive to environmental perturbation were not among the characterized metabolome.
 By analyzing depth variability (accounting for 20-40% of metabolomic variability across ~150 meters) and the vertical displacement of isopycnal surfaces (explaining 10-20% of variability across a sea level anomaly range of 40 centimeters), this analysis constrains the importance of mesoscale eddies in shaping the chemical composition of particulate matter in the largest biomes on the planet.
@@
 
-High frequency, long-term observations at Station ALOHA in the North Pacific Subtropical Gyre (NPSG) over the past 25 years have revealed temporal and spatial variability in what had previously been considered a relatively homogenous environment [@Karl1999; @Karl2017; @Karl2021].
+High frequency observations at Station ALOHA in the North Pacific Subtropical Gyre (NPSG) over the past 25 years have revealed temporal and spatial variability in what had previously been considered a relatively homogenous environment [@Karl1999; @Karl2017; @Karl2021].
 A major source of variability comes in the form of mesoscale eddies in which water is entrained into circular surface currents tens to hundreds of kilometers in diameter [@Karl2017; @McGillicuddy2016].
@@
 The uplift of deep, nutrient-rich seawater into the euphotic zone alters microbial communities [@Rii2022].
-Measurements of chlorophyll in cyclones reveal a shallower and more intense maximum [@Barone2022; @Cornec2021] as a result of eukaryotic phytoplankton thriving in the increased nutrient concentrations while cyanobacterial biomass is reduced [@Hawco2021].
+Measurements of chlorophyll in cyclones reveal a shallower and more intense maximum [@Barone2022; @Cornec2021] as a result of eukaryotic phytoplankton thriving in the higher nutrient concentrations while cyanobacterial biomass is reduced [@Hawco2021].
 This growth of large eukaryotes corresponds to a net increase in biomass and primary productivity [@Benitez-Nelson2007], especially near the deep chlorophyll maximum [@Barone2022].
@@
 Eukaryotic organisms have distinct chemical fingerprints from those of cyanobacteria and heterotrophic bacteria [@Heal2021; @Durham2022] and the relative contribution of these different taxa to total biomass should reflect their chemical composition.
-However, recent work from @Harke2021 showed that overall community function may be robust across eddy types at the surface while @Gleich2024 detected significant differences in protistan community composition and metabolic potential down to 250 meters, with heterotrophy-associated transcripts found at higher abundances in the cyclone.
-These results raise questions about the effect of mesoscale features on particulate organic matter composition.
+However, recent work from @Harke2021 showed that overall community function may be robust across eddy types at the surface while @Gleich2024 detected significant differences in protistan community composition and metabolic potential down to 250 meters, with heterotrophy-associated protistan transcripts enriched in the cyclone.
 
@@
 We explored the impacts of isopycnal uplift and depression on the composition of particulate organic matter across pairs of mesoscale eddies of opposite polarity in the North Pacific Subtropical Gyre (NPSG).
-The first eddy pair was the focus of the July 2017 Microbial Ecology of the Surface Ocean-Simons Collaboration on Ocean Processes and Ecology (MESO-SCOPE, KM1709 aboard the R/V *Kilo Moana*) cruise and sampling consisted of two phases (Figure 1).
-In the first phase, a transect was taken from the north to the south consisting of 11 stations across a pair of eddies, corresponding to the sampling sites in @Barone2022 and @Dugenne2023.
-Metabolomic samples of seawater from 15 meters, the deep chlorophyll maximum (DCM), and 175 meters were collected at each station via 0.2 µm filter.
+The first eddy pair was the focus of the July 2017 cruise, Microbial Ecology of the Surface Ocean-Simons Collaboration on Ocean Processes and Ecology (MESO-SCOPE, KM1709 aboard the R/V *Kilo Moana*).
+During the first phase of our two-phase expedition, a transect was taken from the north to the south consisting of 11 stations across a pair of eddies (the same sampling sites as in @Barone2022 and @Dugenne2023, Figure 1).
+Samples for particulate metabolomic analysis were collected from 15 meters, the deep chlorophyll maximum (DCM), and 175 meters at each station onto 0.2 µm filters.
 The next phase of the MESO-SCOPE cruise focused on Lagrangian sampling in the two eddy centers, starting with the cyclone (SLA = -14 cm) before progressing to the anticyclone (SLA = 24 cm).
@@
 At the DCM, approximately 16% of the variation in the metabolome could be explained by sea level anomaly (SLA) alone (PERMANOVA, p = 0.003).
-Additionally, SLA was strongly correlated with the first principal component (PC) of the metabolome  with a Pearson's *r* of .615 and 27% of the variance explained by PC1 (Supplemental figure 1A), indicating that this was one of the largest sources of variation in the dataset.
+Additionally, SLA was strongly correlated with the first principal component (PC) of the metabolome  with a Pearson's *r* of .615 and 27% of the variance explained by PC1 (Supplemental Figure 1A), indicating that this was one of the largest sources of variation in the dataset.
 Notably, the samples taken from the exact center of the cyclonic eddy at Station 12 were highly distinct and likely drove much of the explained variance.
-In the 175 meter samples, 12% of the variance was explained by SLA (PERMANOVA, p = 0.008) and SLA was highly correlated with the second PC of the dataset (r = .697, % variance explained by PC2 = 15.4, Supplemental figure 1A), though the first PC did not seem to have any visible pattern with metadata and appeared to largely capture variation between biological triplicates (Supplemental figure 1B).
+In the 175 meter samples, 12% of the variance was explained by SLA (PERMANOVA, p = 0.008) and SLA was highly correlated with the second PC of the dataset (r = .697, % variance explained by PC2 = 15.4, Supplemental Figure 1A), though the first PC did not seem to have any visible pattern with metadata and appeared to largely capture variation between biological triplicates (Supplemental Figure 1B).
 At 15 meters, SLA trends are less evident with a larger p-value (PERMANOVA, p=0.087) and a lower R$^2$ (0.09) (Figure 2).
@@
 This shift in metabolite abundance was largely driven by a corresponding shift in biomass.
-Particulate carbon and summed metabolite concentration were tightly correlated across all samples and depths ($\beta = 24.9\pm1.82$ (SE), Pearson's $r$ = .813, p-value < 0.001, Supplemental figure 2).
-This overall trend was largely driven by higher values at 15 meters and the DCM than at depth, but when analyzed at each depth individually the general correspondence held (Supplemental figure 2).
+Particulate carbon and summed metabolite concentration were tightly correlated across all samples and depths ($\beta = 24.9\pm1.82$ (SE), Pearson's $r$ = .813, p-value < 0.001, Supplemental Figure 2).
+This overall trend was largely driven by higher values at 15 meters and the DCM than at depth, but when analyzed at each depth individually the general correspondence held (Supplemental Figure 2).
 The DCM and 175 meter samples had the stronger trends ($\beta_{DCM} = 22.6\pm5.53$, $r= .593$, p-value < 0.001; $\beta_{175m} = 23.9\pm7.36$, $r= .503$, p-value = 0.003) while 15 meter samples showed no significant relationship ($\beta_{15m} = 3.4\pm12.9$, $r= .049$, p-value = 0.792).
@@
 
-The most abundant intracellular metabolite quantified in the eddy transect samples was trimethylamine N-oxide, with an average concentration of 1.38 nM and a distinct DCM maximum (Figure 2), though a few samples had concentrations an order of magnitude higher for unknown reasons (Supplemental figure 3).
+The most abundant intracellular metabolite quantified in the eddy transect samples was trimethylamine N-oxide, with an average concentration of 1.38 nM and a distinct DCM maximum (Figure 2), though a few samples had concentrations an order of magnitude higher for unknown reasons (Supplemental Figure 3).
 A majority of the known molecules (37/53) had a similar pattern with a subsurface maximum at the DCM, including glycine betaine, glutamate, and guanine.
@@
 
-One sample, taken from 15 meters at Station 11, was highly distinct from all others (including its paired triplicates) with several metabolites measuring two or more orders of magnitude larger in peak area and concentration both before and after normalization to internal standards (Supplemental figure 3).
+One sample, taken from 15 meters at Station 11, was highly distinct from all others (including its paired triplicates) with several metabolites measuring two or more orders of magnitude larger in peak area and concentration both before and after normalization to internal standards (Supplemental Figure 3).
 Heuristic exploration of potentially related masses identified a total of 14 compounds highly abundant in this one sample alone.
@@
 
-Given the strength of the metabolomic response to SLA at the DCM , we additionally analyzed samples taken with high-resolution depth around the DCM at the center of both eddy poles.
-Here, we found that the ~40 meter depth range and the transition between the eddies explained similar amounts of variation in the data (R$^2_{depth}=0.253$, R$^2_{SLA}=0.238$) and were both highly significant factors (permutational p-values << 0.001, Figure 4).
+We further investigated the metabolomic response to SLA at the DCM by collecting samples taken at high-resolution depth intervals around the DCM at the center of both eddy poles.
+Here, we found that this ~40 meter depth range and the transition between the eddies explained similar amounts of variation in the data (R$^2_{depth}=0.253$, R$^2_{SLA}=0.238$) and were both highly significant factors (permutational p-values << 0.001, Figure 4).
 Additionally, both sample depth and SLA were correlated with the first principal component of the metabolite matrix ($r_{depth}=.766$, $r_{SLA}=.684$, fraction of variance explained by PC1 = 30.8%).
@@
   geom_hline(yintercept = 0) +
-  geom_point(aes(x=MDS1, y=MDS2, fill=fill_col, shape=sla_class), size=4, color="white") +
+  geom_point(aes(x=MDS1, y=MDS2, fill=fill_col, shape=sla_class), size=6, color="white") +
   scale_fill_manual(breaks=c("25 m Cyclone", "25 m Anticyclone", "DCM Cyclone", "DCM Anticyclone"),
@@
   )), shape="none") +
+  theme(legend.key.spacing.y = unit(6, "pt")) +
   coord_fixed()
@@
 
-When analyzed separately as distinct depths, the 25 meter samples had higher variances explained by eddy (R$^2$=.26) and a lower p-value (0.013), while the DCM samples were no longer likely to be distinct between the two eddies (p-value = 0.171) (Figure 6, Supplemental figure 4).
+When analyzed separately as distinct depths, the 25 meter samples had higher variances explained by eddy (R$^2$=.26) and a lower p-value (0.013), while the DCM samples were no longer likely to be distinct between the two eddies (p-value = 0.171) (Figure 6, Supplemental Figure 4).
 
@@
 This difference in isethionic acid concentration was only found after normalizing each sample to the sum of all metabolites in the sample to control for biomass.
-Arsenobetaine was again found to strongly enriched in the anticyclone at the DCM both when normalizing to biomass and when not, just as in MESO-SCOPE.
+Arsenobetaine was again found to be strongly enriched in the anticyclone at the DCM both when normalizing to biomass and when not, just as in MESO-SCOPE.
 Surprisingly, the 173.0921 *m/z* mass feature was also slightly enriched but this time in the *cyclone* DCM with peak areas approximately 1.4 times the values in the anticyclone (t-test p-value = 0.067, Mann-Whitney p-value = 0.045).
@@
 The OUC is a staple of metazoan metabolism and is present in stramenopiles as well where it enables rapid response to increases in nitrogen supply [@Allen2011; @Bender2014].
-The enrichment of these metabolites in the HEE cruise may indicate that there was still rapid uptake and turnover of nitrogen in the 2018 eddy pair while the 2017 eddies were more stable, despite both eddies being in a similar state of stable anticyclone and weakening cyclone [@Dugenne2023].
+The enrichment of these metabolites in the HEE cruise may indicate that there was still rapid uptake and turnover of nitrogen in the 2018 eddy pair while the 2017 eddies were less active, despite both eddy pairs being in a similar state of stable anticyclone and weakening cyclone [@Dugenne2023].
 
@@
 Briefly, filters were randomized and added to PTFE centrifuge tubes with a 1:1 mix of 100 µm and 400 µm silica beads, approximately 2 mL -20 °C Optima-grade dichloromethane, and approximately 3 mL -20 °C 1:1 methanol/water solution (both also Optima-grade).
-Isotope labeled extraction standards were also added during this step (Supplemental table 1).
+Isotope labeled extraction standards were also added during this step (Supplemental Table 1).
 The samples were then bead-beaten three times, followed by triplicate extraction of the aqueous layer and addition with fresh methanol/water mixture with additional bead-beatings in between.
-Samples were then dried down under clean nitrogen gas and warmed using a Fisher-Scientific Reacti-Therm module.
-Dried aqueous fractions were reconstituted in 380 µL Optima-grade water and amended with 20 µL isotope-labeled injection standards (Supplemental table 1) to measure the variability introduced by chromatography and ionization.
+Samples were then dried down under ultrapure nitrogen gas and warmed using a Fisher-Scientific Reacti-Therm module.
+Dried aqueous fractions were reconstituted in 380 µL Optima-grade water and amended with 20 µL isotope-labeled injection standards (Supplemental Table 1) to measure the variability introduced by chromatography and ionization.
 The reconstituted fraction was syringe-filtered into precombusted glass inserts in HPLC vials to remove any potential clogging material.
@@
 A pooled sample was created by combining 20 µL of each sample into the same HPLC vial, and a 1:1 dilution with water created a half-strength pooled sample from that to assess matrix effects and obscuring variation [@Boysen2018].
-Also run alongside the environmental samples were samples of authentic standards split into two mixes (Supplemental table 2).
+Also run alongside the environmental samples were samples of authentic standards split into two mixes (Supplemental Table 2).
 Standards were dissolved in Optima-grade water to help with identification and into an aliquot of the pooled sample to quantify matrix response factors.
@@
 
-Compounds for which we had an authentic standard (Supplemental table 2) were manually integrated using Skyline and MSDIAL with the standard mixes used to ensure the correct peak was integrated by matching retention time.
+Compounds for which we had an authentic standard (Supplemental Table 2) were manually integrated using Skyline and MSDIAL with the standard mixes used to ensure the correct peak was integrated by matching retention time.
 Compounds that were visually determined to have poor peak quality in the samples were removed from the downstream analysis.
 Skyline was used for both the eddy center samples and the HEE cruise data while MSDIAL was used for the eddy transect data due to the larger number of files in that dataset.
-Raw peak areas were then normalized to their best-matched internal standard (Supplemental table 1) per @Boysen2018 except that all compounds were normalized to their best match no matter how minimal the improvement.
+Raw peak areas were then normalized to their best-matched internal standard (Supplemental Table 1) per @Boysen2018 except that all compounds were normalized to their best match no matter how minimal the improvement.
 Normalized peaks were then calibrated via a single-point standard curve as determined uniquely for each compound from the authentic standard mixes and used to convert normalized peak area into moles per µL injected and scaled to the amount of seawater filtered to provide a final estimate of environmental concentration for each compound in moles per liter of seawater filtered.
@@
 This was followed by manual inspection of the extracted ion chromatogram for each mass feature by an expert for quality with only those peaks classified as "Good" used in the analysis here.
-Features were then matched with and normalized to an internal standard (Supplemental table 1) as described above.
+Features were then matched with and normalized to an internal standard (Supplemental Table 1) as described above.
 
@@
 
-*Supplemental figure 1: Results of a principal component analysis performed separately on each depth of the eddy transect metabolome.
+*Supplemental Figure 1: Results of a principal component analysis performed separately on each depth of the eddy transect metabolome.
 A) Regressions of SLA against various principal components, broken down by depth.
@@
 
-*Supplemental figure 2: Type I linear regressions of particulate carbon and nitrogen against total nM carbon and nitrogen in known metabolites both as a whole and divided by depth.
+*Supplemental Figure 2: Type I linear regressions of particulate carbon and nitrogen against total nM carbon and nitrogen in known metabolites both as a whole and divided by depth.
 Points have been colored according to depth (top row of plots) or SLA (bottom row of plots) and have been placed on independent axes.
@@
 
-*Supplemental figure 3: Stacked barplots of known metabolite concentration as shown in main text Figure 3 but with individual triplicates shown.
+*Supplemental Figure 3: Stacked barplots of known metabolite concentration as shown in main text Figure 3 but with individual triplicates shown.
 The same data is shown in A) as B) but B) is rendered in relative contribution space instead of absolute concentration.
@@
 
-*Supplemental figure 4: Plots of the metabolome during the Hawaiian Eddy Experiment, broken down by depth.
+*Supplemental Figure 4: Plots of the metabolome during the Hawaiian Eddy Experiment, broken down by depth.
 The top row of plots are non-metric multidimensional scaling (NMDS) plots, in which points correspond to individual samples and have been colored by their corrected sea level height anomaly.

### 2024-06-28 manuscript/manuscript.Rmd
Simple edits from meeting w Anitra

@@
 This study reports the absolute concentration of 67 metabolites and relative abundances for 640 molecular features to understand how mesoscale eddies impact the metabolome of the North Pacific Subtropical Gyre during two cruises in 2017 and 2018.
-We find that many metabolites track biomass trends, but biomarker metabolites like isethionic acid, homarine, and trigonelline linked to eukaryotic phytoplankton were enriched at the deep chlorophyll maximum of the cyclonic features, while degradation products such as arsenobetaine were enriched in anticyclones.
+We find that many metabolites track biomass trends, but metabolites like isethionic acid, homarine, and trigonelline linked to eukaryotic phytoplankton were enriched at the deep chlorophyll maximum of the cyclonic features, while degradation products such as arsenobetaine were enriched in anticyclones.
 In every analysis, metabolites with the strongest responses were detected via untargeted mass spectrometry, indicating that the molecules most sensitive to environmental perturbation were not among the characterized metabolome.
-By analyzing depth variability (accounting for 20-40% of metabolomic variability across ~150 meters) and the vertical displacement of isopycnal surfaces (explaining 10-20% of variability across a sea level anomaly range of 40 centimeters), this analysis constrains the importance of mesoscale eddies in shaping the chemical composition of particulate matter in the largest biomes on the planet.
+By analyzing depth variability (accounting for 20-40% of metabolomic variability across ~150 meters) and the vertical displacement of isopycnal surfaces (explaining 10-20% of variability across a sea level anomaly range of 40 centimeters and a spatial distance of 300 kilometers), this analysis constrains the importance of mesoscale eddies in shaping the chemical composition of particulate matter in the largest biomes on the planet.
 
@@
 However, recent work from @Harke2021 showed that overall community function may be robust across eddy types at the surface while @Gleich2024 detected significant differences in protistan community composition and metabolic potential down to 250 meters, with heterotrophy-associated protistan transcripts enriched in the cyclone.
+The chemical composition of organic matter is one metric of community function since community metabolism and interactions are in part mediated through organic molecules, making their measurement useful for determining shifts in community dynamics.
 
@@
 Lagrangian sampling near the eddy centers of the MESO-SCOPE cruise was performed by following drifters deployed in proximity of Stations 6 and 12.*
-
-Metabolomics samples from all three sample sets were extracted via a modified Bligh & Dyer [@Boysen2018], with the aqueous fraction run on a HILIC column paired with a Q-Exactive HF mass spectrometer.
-Metabolites for which an authentic standard was available and run alongside were integrated in Skyline (HEE and MESO-SCOPE eddy centers) and MSDIAL (MESO-SCOPE eddy transect).
-Untargeted analysis was performed using the R package XCMS with each feature manually inspected for quality assurance.
-[Wei: I think this paragraph belongs to Methods section.
-Will: Maybe just shorten it extensively?
-Ask Anitra for advice since Rachel wanted this here.]
 
@@
 SLA trends are visible in the DCM and 175 meter samples, with dark blue circles consistently discriminating from the dark red circles along the first multidimensional axis.
-Bottom panels depict the direction and magnitude of this effect by plotting the grand mean value of three biological triplicates in color and the raw values in black behind.*
+Bottom panels depict the direction and magnitude of this effect by plotting the mean value of all z-scored metabolites across three biological triplicates in color with the raw values in black behind.*
 

### 2024-07-01 manuscript/manuscript.Rmd
Restructuring for word counts and adding untargeted area discussion

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, eval = FALSE)
+knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, eval = TRUE)
 library(tidyverse)
@@
 The DCM and 175 meter samples had the stronger trends ($\beta_{DCM} = 22.6\pm5.53$, $r= .593$, p-value < 0.001; $\beta_{175m} = 23.9\pm7.36$, $r= .503$, p-value = 0.003) while 15 meter samples showed no significant relationship ($\beta_{15m} = 3.4\pm12.9$, $r= .049$, p-value = 0.792).
-This meant that metabolites contributed a fixed fraction of carbon to the particulate pool with our 53 quantified metabolites representing between 1.5% and 5% of the carbon in the system (Figure 2) [Benedetto: I find this very informative.
-It would be great to get a ballpark sense of the percentages including the unknown metabolites, but I understand that it might be too uncertain considering the required assumptions on molecular formulas and concentration.
-I would like to know your wild guess, though].
+This meant that metabolites contributed a relatively fixed fraction of carbon to the particulate pool with our 53 quantified metabolites representing between 1.5% and 5% of the carbon in the system (Figure 2).
 Particulate nitrogen was tightly correlated with particulate carbon ($r=.933$) but the fraction of particulate nitrogen represented by the quantified metabolites was slightly higher across the transect with contributions typically between 2% and 6%.
+
+Although the untargeted pool contains more molecular features, we expect that a majority of the signal has been captured by our 53 targeted compounds.
+Of the total peak area that passed quality control, approximately 67% was captured by these 53 molecules.
+An additional 15% of the total peak area was putatively annotated as various inorganic ions for which no authentic standard was available but could be matched by mass and isotope pattern.
+Additionally, our targeted list accounted for 13 out of the top 20 largest molecular features by total peak area with an additional 4 features annotated as the inorganic ions for 17/20 of the largest features known.
+However, these calculations should be taken as estimates because peak area does not correspond directly to environmental concentration and it is entirely possible for abundant environmental compounds to have small peak areas if they ionize poorly on the mass spectrometer or degrade during the sampling and extraction process.
+
+```{r untarg targ comparison, eval=FALSE}
+untarg_peaks %>%
+  select(feature, filename, norm_area) %>%
+  left_join(untarg_feats %>% select(feature, compound_name)) %>%
+  group_by(is.na(compound_name)) %>%
+  summarise(s=sum(norm_area))
+
+summed_labels <- untarg_peaks %>%
+  filter(str_detect(feature, "MS")) %>%
+  group_by(feature) %>%
+  summarise(s=sum(norm_area)) %>%
+  arrange(desc(s)) %>%
+  left_join(untarg_feats %>% select(feature, mzmed, rtmed, compound_name, polarity))
+  mutate(compound_name=ifelse(feature=="MS_NEG_0155", "Bromine79", compound_name)) %>%
+  mutate(compound_name=ifelse(feature=="MS_NEG_0778", "Iodine", compound_name)) %>%
+  mutate(compound_name=ifelse(feature=="MS_NEG_0161", "Sulfite", compound_name)) %>%
+  mutate(compound_name=ifelse(feature=="MS_NEG_0039", "Nitrate", compound_name)) %>%
+  mutate(compound_name=ifelse(feature=="MS_POS_1122", "TMAO [2M+H]", compound_name))
+
+summed_labels %>%
+  group_by(is.na(compound_name)) %>%
+  count()
+summed_labels %>%
+  group_by(is.na(compound_name)) %>%
+  summarize(s=sum(s)/1e9)
+
+summed_labels %>%
+  filter(feature!="MS_NEG_0055") %>%
+  filter(feature!="MS_NEG_0068") %>%
+  head(20) %>%
+  print(n=Inf)
+
+msdata <- grabMSdata(list.files("../tmzMLs/neg/", full.names=TRUE, pattern="180821_Poo"))
+v <- msdata$MS1[mz%between%pmppm(146.0451, 5)]
+qplotMS1data(v)
+```
 
@@
 
+pcbarplot <- frac_pc_in_metab_df %>% 
+  distinct(station, depth, PC_um) %>%
+  ggplot(aes(x=station, y=PC_um)) +
+  geom_col(fill="grey30") +
+  facet_grid(~depth, scales="free") +
+  labs(y="Particulate\ncarbon (μM)") +
+  scale_x_discrete(breaks=as.character(seq(4, 14, 2))) +
+  scale_y_continuous(expand = c(0.005, 0.1)) +
+  theme_bw() +
+  theme(strip.background = element_blank(), strip.text = element_blank(),
+        axis.text.x = element_blank(), axis.title.x = element_blank(),
+        plot.margin = margin(0,1,0,0), plot.background = element_rect(fill="white"))
+
 med_frac_df <- frac_pc_in_metab_df %>%
@@
 
-targ_gp_w_sla_frac <- egg::ggarrange(barplot_targ_gp, pc_frac_metab_gp, draw = FALSE)
-
+# targ_gp_w_sla_frac <- egg::ggarrange(barplot_targ_gp, pc_frac_metab_gp, draw = FALSE)
+# ggsave("targ_gp_w_sla_frac.png", plot = targ_gp_w_sla_frac, device = "png", 
+#        path = "figures", width = 6.875, height = 4.5, units = "in", dpi = 300, 
+#        bg = "white")
+targ_gp_w_sla_frac <- egg::ggarrange(barplot_targ_gp, pcbarplot, pc_frac_metab_gp,
+                                     draw = FALSE, heights = c(0.4, 0.2, 0.4))
 ggsave("targ_gp_w_sla_frac.png", plot = targ_gp_w_sla_frac, device = "png",
-       path = "figures", width = 6.875, height = 4.5, units = "in", dpi = 300, 
+       path = "figures", width = 6.875, height = 5.5, units = "in", dpi = 300,
        bg = "white")
@@
 TMAO = trimethylamine N-oxide, HO-Ile = hydroxyisoleucine, GBT = glycine betaine, DCM = deep chlorophyll maximum.
-Lower panels depict the fraction of total particulate carbon in the known metabolites, with the mean value of three biological triplicates in color and the raw values in black behind.
+Center panels show the corresponding measurements of particulate carbon (PC), while the lower panels depict the fraction of total particulate carbon in the known metabolites.
+The mean value of three biological triplicates is shown in color and the raw values in black behind.
 Colors correspond to corrected sea level anomaly (Corr.
@@
 
-### A single sample highlights a network of related metabolites
+## Particulate organic matter compositional shifts across eddy transect
+
+In addition to absolute shifts in metabolite concentration due to biomass differences, we also explored compositional shifts in the metabolomes by assessing the fraction of the total metabolite pool contributed by each compound across the eddy transect.
+This allowed us to isolate the signal due to shifts in community composition and organismal response from those introduced by changes in biomass of the community as a whole.
+
+```{r regression relative abundance coefficients, eval=FALSE}
+reg_coefs <- untarg_peaks %>%
+  # filter(feature!="MS_NEG_0155") %>%
+  left_join(filled_file_metadata %>% distinct(filename, ms_run, station, depth, sla_corr)) %>%
+  filter(ms_run=="MT") %>%
+  filter(depth%in%c("Surface", "DCM", "Deep")) %>%
+  filter(feature%in%unique(feature)) %>%
+  group_by(filename) %>%
+  mutate(norm_area=norm_area/sum(norm_area)) %>%
+  ungroup() %>%
+  nest(data=-c(depth, feature)) %>%
+  mutate(test=map(data, ~broom::tidy(lm(.x$norm_area~.x$sla_corr)), .progress = TRUE)) %>%
+  unnest(test)
+reg_coefs %>%
+  filter(str_detect(term, "sla_corr")) %>%
+  mutate(p.value=p.adjust(p.value, method="fdr")) %>%
+  arrange(p.value) %>%
+  left_join(untarg_feats) %>%
+  select(feature, depth, estimate, p.value, mzmed, rtmed, compound_name) %>%
+  filter(p.value<0.05) %>%
+  # as.data.frame() %>%
+  filter(depth=="Deep") %>% 
+  print(n=Inf)
+reg_coefs %>%
+  filter(str_detect(term, "sla_corr")) %>%
+  left_join(untarg_feats) %>%
+  filter(compound_name=="Homarine") %>%
+  # filter(feature=="MS_POS_1395") %>%
+  filter(depth=="DCM") %>%
+  select(compound_name, depth, data, estimate) %>%
+  unnest(data) %>%
+  group_by(sla_corr, station) %>%
+  ggplot(aes(x=sla_corr, y=norm_area*100)) +
+  geom_smooth(method="lm", formula="y~x") +
+  geom_boxplot(aes(group=sla_corr)) +
+  scale_y_continuous(limits = c(0, NA))
+```
+
+In the 15 meter surface samples, only seven compounds had significant changes in relative peak area across the eddy transect, two of which were known: 5-oxoproline, which decreased from 0.2-0.3% of the total peak area at the center of the anticyclone to 0.05-0.1% in the cyclone; and a combined peak of sarcosine and beta-alanine which increased from 0.01% to 0.02-0.03%.
+The largest shift significant at an α = 0.05 level was an unknown mass feature with *m/z* = 189.12338 and a retention time around 8.6 minutes, the relative contribution of which increased from approximately 0.2-0.3% in the cyclone to ~1.5% in the anticyclone.
+
+At the DCM, twenty-two compounds changed significantly in relative peak area across the transect.
+Seven of these were known metabolites, of which trigonelline and homarine were most enriched in the cyclone while arsenobetaine was the only metabolite with a larger fraction of the total peak area in the anticyclone.
+Homarine showed a very large and highly significant shift, representing about 2.5% of the total peak area in the anticyclone but 7-8% in the cyclone center.
+Trigonelline (N-methyl niacin), an isomer structurally very similar to homarine but biologically distinct, had a surprisingly strong correlation with homarine (Pearson's $r=.855$).
+Its peak areas were consistently around one third of homarine's but still large enough to have the second-largest shift in relative peak area of the significantly different compounds.
+Similarly, arsenobetaine represented about 0.1% of the peak area in the cyclone and 0.5-0.6% of the total peak area in the anticyclone.
+The most significantly different compounds at the DCM in each direction, however, were both unknowns.
+The lowest p-value ($6.4*10^{-6}$ after FDR correction) in the DCM data was a mass feature with an *m/z* of 173.09211, a retention time also around 8.6 minutes, and enrichment in the anticyclone with a putative chemical formula of [M+H] = C$_7$H$_{13}$N$_2$O$_3$, possibly glycylproline or prolylglycine.
+The next-lowest (p = $2.5*10^{-4}$ after FDR correction) was enriched in the cyclone and had an *m/z* of 275.0712 and a retention time of 11.4 minutes with a putative formula of C$_{18}$H$_{11}$O$_3$.
+
+Finally, in the 175 meter samples we detected 43 mass features with a significant association with SLA.
+Notably, all detected nucleobases (guanine, adenine, and cytosine) were positively associated with SLA (higher values in the anticyclone than the cyclone), though the strongest associations were found in O-acetylcarnitine, betonicine, and tyrosine.
+Both O-acetylcarnitine and betonicine effectively quintupled their contribution to total peak area, shifting from approximately 0.1% to ~0.5% over the transect.
+Acetylcholine was the only known compound more relatively abundant in the cyclone than in the anticyclone along with one unknown of *m/z* 131.0340.
+The unknown at *m/z* 173.0921 with the strongest trend at the DCM again had the strongest trend in the 175 meter samples, here with an FDR-corrected p-value of $2.0 × 10^{-7}$.
+
+```{r homarine/trigonelline correlation, eval=FALSE}
+reg_coefs %>%
+  filter(str_detect(term, "sla_corr")) %>%
+  left_join(untarg_feats) %>%
+  filter(compound_name%in%c("Homarine", "Trigonelline")) %>%
+  filter(depth=="DCM") %>%
+  select(compound_name, depth, data, estimate) %>%
+  unnest(data) %>%
+  select(compound_name, filename, norm_area) %>%
+  pivot_wider(names_from=compound_name, values_from=norm_area) %>%
+  column_to_rownames("filename") %>%
+  cor()
+```
+
+## High vertical resolution sampling near eddy center DCM reveals distinct metabolome between cyclone and anticyclone
+
+```{r MC NMDS and PERMANOVA calculations}
+set.seed(123)
+multivar_analysis_MC <- untarg_peaks %>%
+  left_join(filled_file_metadata %>% distinct(ms_run, station, filename, depth), 
+          by="filename") %>% filter(ms_run=="MC") %>% 
+  mutate(depth=factor(depth, levels=c("DCMless20m", "DCMless10m", "DCM", 
+                                      "DCMplus10m", "DCMplus20m"))) %>%
+  nest(data=everything()) %>%
+  mutate(wide_peaks=map(data, function(metab_df){
+    metab_df %>%
+      select(feature, filename, norm_area) %>%
+      group_by(feature) %>%
+      filter(sum(norm_area==0)<n()*0.9) %>%
+      mutate(norm_area=scale(norm_area)[,1]) %>%
+      pivot_wider(names_from = "feature", values_from = norm_area) %>%
+      column_to_rownames("filename") %>%
+      data.matrix()
+  })) %>%
+  mutate(nmds_output=map(wide_peaks, function(metab_mat){
+    env_vec <- pull(filled_file_metadata, sla_corr, filename)[rownames(metab_mat)]
+    vegan::metaMDS(metab_mat, k = 2, distance = "manhattan", trace = 0, 
+                   autotransform = FALSE, noshare = FALSE, wascores = FALSE) %>%
+    vegan::MDSrotate(env_vec)
+  })) %>%
+  mutate(nmds_stress=map_dbl(nmds_output, pluck, "stress")) %>%
+  mutate(perm_output=map(wide_peaks, function(metab_mat){
+    adon_expl <- data.frame(filename=rownames(metab_mat)) %>%
+      left_join(filled_file_metadata %>% select(filename, sla_class, depth), by="filename")
+    vegan::adonis2(
+      metab_mat~sla_class+depth, data = adon_expl, by = "margin", 
+      permutations = 999, method = "manhattan")
+  })) %>%
+  mutate(perm_r2_sla=map_dbl(perm_output, pluck, "R2", 1)) %>%
+  mutate(perm_r2_depth=map_dbl(perm_output, pluck, "R2", 2)) %>%
+  mutate(perm_pval=map_dbl(perm_output, pluck, "Pr(>F)", 1)) %>%
+  mutate(pca_output=map(wide_peaks, prcomp)) %>%
+  mutate(nmds_stress=paste("NMDS stress =", round(nmds_stress, digits = 3)))
+  # mutate(perm_r2=round(perm_r2, digits = 2)) %>%
+  # mutate(perm_pval=round(perm_pval, digits = 4)) %>%
+  # mutate(perm_label=paste0("R2 = ", perm_r2, "\np = ", perm_pval))
+```
+
+We further investigated the metabolomic response to SLA at the DCM by collecting samples taken at high-resolution depth intervals around the DCM at the center of both eddy poles.
+Here, we found that this ~40 meter depth range and the transition between the eddies explained similar amounts of variation in the data (R$^2_{depth}=0.253$, R$^2_{SLA}=0.238$) and were both highly significant factors (permutational p-values << 0.001, Figure 4).
+Additionally, both sample depth and SLA were correlated with the first principal component of the metabolite matrix ($r_{depth}=.766$, $r_{SLA}=.684$, fraction of variance explained by PC1 = 30.8%).
+The cyclonic samples in particular show a much larger spread than the anticyclonic ones, indicating larger sample variability in the cyclone DCM relative to the anticyclone.
+A depth gradient is visible along NMDS 2, with the deepest samples generally at the top right of the plot and the shallowest ones closer to the bottom (Figure 4).
+
+```{r figure MC NMDS plot}
+MC_nmds_gp <- multivar_analysis_MC$nmds_output %>%
+  pluck(1, "points") %>% 
+  as.data.frame() %>%
+  rownames_to_column("filename") %>%
+  left_join(filled_file_metadata %>% distinct(filename, sla_class, depth, tripl)) %>%
+  mutate(depth=factor(depth, levels=c("DCMless20m", "DCMless10m", "DCM", 
+                                      "DCMplus10m", "DCMplus20m"),
+                      labels=c("-20m", "-10m", "DCM", "+10m", "+20m"))) %>%
+  mutate(fill_col=case_when(
+    depth=="-20m" & sla_class=="Cyclone"~"#c4c4e0",
+    depth=="-10m" & sla_class=="Cyclone"~"#8989c1",
+    depth=="DCM" & sla_class=="Cyclone"~"#6161ad",
+    depth=="+10m" & sla_class=="Cyclone"~"#343489",
+    depth=="+20m" & sla_class=="Cyclone"~"#23235b",
+    depth=="-20m" & sla_class=="Anticyclone"~"#dabdbd",
+    depth=="-10m" & sla_class=="Anticyclone"~"#b57c7c",
+    depth=="DCM" & sla_class=="Anticyclone"~"#9c5050",
+    depth=="+10m" & sla_class=="Anticyclone"~"#762020",
+    depth=="+20m" & sla_class=="Anticyclone"~"#4f1616"
+  )) %>%
+  ggplot() +
+  geom_vline(xintercept = 0) +
+  geom_hline(yintercept = 0) +
+  geom_point(aes(x=MDS1, y=MDS2, fill=fill_col, shape=sla_class, color=depth), size=6) +
+  scale_fill_identity() +
+  scale_color_manual(values = rep("white", 5)) +
+  scale_shape_manual(breaks=c("Cyclone", "Anticyclone"), values=c(25, 24)) +
+  annotate("label", x=-200, y=100, label="Cyclone", color=scales::muted("blue"),
+           label.r = unit(0, "in"), size=6) +
+  annotate("label", x=100, y=120, label="Anticyclone", color=scales::muted("red"),
+           label.r = unit(0, "in"), size=6) +
+  guides(alpha="none", fill="none", shape="none", color=guide_legend(
+    title = "Depth:    ", title.position = "left", title.hjust = 0.5, direction = "horizontal",
+    keywidth=0.2, keyheight=0.1,
+    override.aes = list(fill=c("grey90", "grey80", "grey60", "grey40", "grey20"), 
+                        pch=21, color="black")
+  )) +
+  labs(x="NMDS axis 1", y="NMDS axis 2") +
+  theme_bw() +
+  theme(legend.position = "top", legend.text = element_text(margin = margin(r = 15, l=2, unit = "pt"))) +
+  annotate("label", x=-Inf, y=Inf, hjust=0, vjust=1, label.r = unit(0, "in"),
+           label=expression(atop(paste("R"[SLA]^2, "=.238"), paste("R"[depth]^2, "=.253")))) +
+  annotate("label", x=-Inf, y=-Inf, label=multivar_analysis_MC$nmds_stress, 
+           hjust=0, vjust=-2, label.r = unit(0, "in"))
+
+ggsave("MC_nmds_gp.png", plot = MC_nmds_gp, device = "png", 
+       path = "figures", width = 6.875, height = 4.5, units = "in", dpi = 300)
+```
+
+![](figures/MC_nmds_gp.png)
+
+*Figure 4: NMDS plot of high-resolution depth sampling around the deep chlorophyll maximum (DCM, ~115 meters) at the two eddy centers during MESO-SCOPE.
+Red upward-pointed triangles are from the anticyclone and blue downward-pointed ones are from the cyclone.
+Shading intensity reflects the depth above or below the DCM.
+PERMANOVA estimates of the variance explained by depth and sea level anomaly (SLA) are noted in the upper left corner and the NMDS stress value is reported in the bottom left.*
+
+```{r PCA checks for MC axis regressions, eval=FALSE}
+multivar_analysis_MC$pca_output[[1]]$x %>%
+  as.data.frame() %>%
+  rownames_to_column("filename") %>%
+  left_join(filled_file_metadata) %>%
+  select(abs_depth, sla_corr, PC1:PC3) %>%
+  # cor() %>% round(3)
+  ggplot(aes(x=abs_depth, y=PC1, color=sla_corr)) +
+  geom_point()
+
+head(multivar_analysis_MC$pca_output[[1]]$sdev^2/sum(multivar_analysis_MC$pca_output[[1]]$sdev^2)*100)
+```
+
+To characterize the observed differences in multivariate space we used k-means clustering as an unsupervised way to identify dominant trends within the data (Figure 5).
+We found that a majority of the metabolites (Clusters 1 and 2, 63% of the total) fell into clusters with larger peak areas in the cyclone, as expected, while also detecting 32 metabolites that clustered such that the mean metabolite was enriched in the anticyclone (Cluster 4, Figure 5).
+Cluster 2 had a distinct decrease in relative peak area with depth, while the other clusters had much less clear depth trends.
+Cluster 1 also showed a DCM maximum for the samples from the anticyclone, correlating well with flow cytometry counts of picoeukaryotes (Pearson's $r=.851$) that was not present in the cyclone ($r=.013$).
+
+```{r k-means clusterplot}
+set.seed(123)
+kclust <- untarg_peaks %>%
+  left_join(filled_file_metadata, by="filename") %>%
+  filter(ms_run=="MC") %>%
+  select(feature, filename, norm_area) %>%
+  group_by(feature) %>%
+  filter(sum(norm_area==0)<n()*0.9) %>%
+  mutate(norm_area=scale(norm_area)[,1]) %>%
+  pivot_wider(names_from = "filename", values_from = norm_area) %>%
+  column_to_rownames("feature") %>%
+  data.matrix() %>%
+  kmeans(centers = 4)
+
+kclust_nms <- names(kclust$cluster)
+kclust$cluster <- order(kclust$size, decreasing = TRUE)[kclust$cluster]
+names(kclust$cluster) <- kclust_nms
+kclust$size <- sort(kclust$size, decreasing = TRUE)
+
+kclust_df <- kclust %>%
+  pluck("cluster") %>%
+  data.frame(cluster=.) %>%
+  rownames_to_column("feature") %>%
+  left_join(untarg_peaks, multiple = "all", by="feature") %>%
+  left_join(filled_file_metadata, by="filename") %>%
+  filter(ms_run=="MC") %>%
+  arrange(desc(abs_depth)) %>%
+  mutate(depth=fct_inorder(depth)) %>%
+  group_by(feature) %>%
+  mutate(norm_area=scale(norm_area)[,1]) %>%
+  group_by(cluster, depth, sla_class) %>%
+  summarise(mean_rank=mean(norm_area), iqr_value=sd(norm_area), .groups = "drop")
+
+lab_df <- tribble(
+  ~xlab, ~ylab, ~sla_class, ~clusterlab,
+  0, 5.5, "Anticyclone", "Cluster #1: n=73",
+  0.8, 1, "Cyclone", "Cluster #1: n=73"
+)
+
+library(ggh4x)
+strip_theme <- strip_themed(background_x = elem_list_rect(fill= c(
+  "#414487FF", "#7AD151FF", "#22A884FF", "#2A788EFF")))
+kclust_gp <- kclust_df %>%
+  left_join(data.frame(size=kclust$size) %>% mutate(cluster=row_number()), by="cluster") %>%
+  mutate(clusterlab=paste0("Cluster #", cluster, ": n=", size)) %>%
+  mutate(depth=factor(depth, labels=rev(c("-20m", "-10m", "DCM", "+10m", "+20m")),
+                      levels=rev(c("DCMless20m", "DCMless10m", "DCM", "DCMplus10m", "DCMplus20m")))) %>%
+  ggplot(aes(y=depth, color=sla_class)) +
+  geom_vline(xintercept = 0) +
+  geom_path(aes(x=mean_rank, group=sla_class), linewidth=1) +
+  geom_point(aes(x=mean_rank), size=3) +
+  geom_label(aes(x=xlab, y=ylab, label=sla_class, color=sla_class), 
+             data=lab_df, label.r = unit(0, "in")) +
+  ggh4x::facet_wrap2(~clusterlab, nrow = 1, strip = strip_theme) +
+  scale_color_manual(values = c(scales::muted("blue"), scales::muted("red"), "grey50"), 
+                     breaks = c("Cyclone", "Anticyclone", "Neither"),
+                     guide="none") +
+  labs(x="Mean z-scaled metabolite area", y="Sampling depth", color=NULL) +
+  coord_cartesian(ylim = c(1, 5.5)) +
+  theme_bw() +
+  theme(plot.background = element_rect(fill = "white"),
+        strip.background = element_rect(fill = "white"),
+        strip.text = element_text(face = "bold", color="white"))
+```
+
+```{r pairwise_tests and fold_change_df}
+MC_simpledata <- untarg_peaks %>%
+  left_join(filled_file_metadata %>% distinct(filename, ms_run, station), by="filename") %>%
+  filter(ms_run=="MC") %>%
+  select(feature, filename, norm_area, station) %>%
+  group_by(feature) %>%
+  filter(sum(norm_area==0)<n()*0.9) %>%
+  group_by(filename) %>%
+  # mutate(norm_area=norm_area/sum(norm_area)) %>%
+  ungroup()
+pairwise_tests <- MC_simpledata %>%
+  nest(data=c(filename, norm_area, station)) %>%
+  mutate(stattest=map(data, ~broom::tidy(wilcox.test(.x$norm_area~.x$station)))) %>%
+  unnest(stattest) %>%
+  select(feature, statistic, p.value) %>%
+  mutate(p.value=p.adjust(p.value, method="fdr"))
+
+fold_change_df <- MC_simpledata %>%
+  group_by(feature, station) %>%
+  summarise(mean_area=mean(norm_area, na.rm=TRUE)) %>%
+  pivot_wider(names_from=station, values_from=mean_area) %>%
+  mutate(fold_change=L1/L2) %>%
+  select(feature, fold_change)
+```
+
+```{r eddy center volcano plot}
+label_df <- pairwise_tests %>%
+  left_join(untarg_feats) %>%
+  filter(compound_name%in%c("Isethionic acid", "Trigonelline", "Arsenobetaine",
+                            "Homarine", "Dimethylsulfonioacetate", "Dimethylsulfoniopropionate")) %>%
+  left_join(fold_change_df, by = join_by(feature)) %>%
+  mutate(xspot = log2(fold_change), yspot=-log10(p.value))
+
+volcanoplot_gp <- pairwise_tests %>%
+  left_join(fold_change_df, by = join_by(feature)) %>%
+  left_join(rownames_to_column(data.frame(cluster=kclust$cluster), "feature"), 
+            by = join_by(feature)) %>%
+  left_join(untarg_feats %>% select(feature, compound_name), by = join_by(feature)) %>%
+  mutate(known=ifelse(is.na(compound_name), "Unknown", "Known")) %>%
+  mutate(compound_name=ifelse(is.na(compound_name), feature, compound_name)) %>%
+  ggplot(aes(x=log2(fold_change), y=-log10(p.value), color=factor(cluster), 
+             shape=known, label=compound_name)) +
+  geom_vline(xintercept = 0) +
+  geom_hline(yintercept = 0) +
+  geom_hline(yintercept = 1.3, lty=2) +
+  annotate("segment", x=label_df$xspot[5], y=6.8, xend=label_df$xspot[5], yend=label_df$yspot[5], 
+           # arrow=arrow(type="closed", angle = 25, length = unit(0.05, "npc")), 
+           lwd=0.7, color="grey50") +
+  annotate("label", x=-2, y=6.8, label="Arsenobetaine", hjust=0, label.r=unit(0, "in")) +
+  annotate("segment", x=1.5, y=6.8, xend=label_df$xspot[3], yend=label_df$yspot[3], 
+           lwd=0.7, color="grey50") +
+  annotate("label", x=1.5, y=6.8, label="Homarine", hjust=1, label.r=unit(0, "in")) +
+  annotate("segment", x=2, y=6.8, xend=label_df$xspot[2], yend=label_df$yspot[2], 
+           lwd=0.7, color="grey50") +
+  annotate("label", x=2, y=6.8, label="DMSP", hjust=0, label.r=unit(0, "in")) +
+  annotate("segment", x=Inf, y=label_df$yspot[1], xend=label_df$xspot[1], yend=label_df$yspot[1], 
+           lwd=0.7, color="grey50") +
+  annotate("label", x=Inf, y=label_df$yspot[1], label="DMS-Ac", hjust=1, vjust=0.3, label.r=unit(0, "in")) +
+  annotate("segment", x=3.5, y=5, xend=1.9, yend=5, 
+           lwd=0.7, color="grey50") +
+  annotate("segment", x=1.9, y=5, xend=label_df$xspot[4], yend=label_df$yspot[4], 
+           lwd=0.7, color="grey50") +
+  annotate("label", x=Inf, y=5, label="Trigonelline", hjust=1, label.r=unit(0, "in")) +
+  annotate("segment", x=3.6, y=3, xend=2.2, yend=3, 
+           lwd=0.7, color="grey50") +
+  annotate("segment", x=2.2, y=3, xend=label_df$xspot[6], yend=label_df$yspot[6], 
+           lwd=0.7, color="grey50") +
+  annotate("label", x=Inf, y=3, label="Isethionate", hjust=1, label.r=unit(0, "in")) +
+  geom_point(size=2, stroke=1) +
+  scale_x_continuous(breaks=c(-4:4)) +
+  scale_y_continuous(breaks = 0:6, labels = paste0("10<sup>-", 0:6, "</sup>"), expand = expansion(c(0.01, 0.05))) +
+  scale_shape_manual(values=c(4, 16), breaks = c("Unknown", "Known")) +
+  scale_color_manual(
+    values = c("#414487FF", "#7AD151FF", "#22A884FF", "#2A788EFF"),
+    breaks = 1:4
+  ) +
+  theme_bw() +
+  labs(x="Log-2 fold-change cyclone enrichment", y="Mann-Whitney p-value\n(FDR-corrected)",
+       shape="Metabolite status", color="k-means\ncluster") +
+  guides(color="none", shape=guide_legend(order=1)) +
+  theme(axis.text.y = ggtext::element_markdown(),
+        plot.background = element_rect(fill = "white"),
+        legend.position = c(0, 1), legend.justification = c(0, 1), 
+        legend.background = element_rect(fill="#FFFFFF88", color="black"),
+        legend.spacing.y=unit(0, "in"),
+        legend.key.size = unit(0, "cm"))
+```
+
+```{r kclust volcano plot combine}
+kclust_volcano_gp <- egg::ggarrange(
+  kclust_gp, volcanoplot_gp, 
+  ncol=1, heights = c(1.5, 2), draw = FALSE)
+ggsave("kclust_volcano_gp.png", plot = kclust_volcano_gp, device = "png", 
+       path = "figures", width = 6.875, height = 5, units = "in", dpi = 300,
+       bg="white")
+```
+
+![](figures/kclust_volcano_gp.png)
+
+*Figure 5: Distribution of metabolites in the high-resolution depth samples from the centers of each MESO-SCOPE eddy.
+The upper row of plots shows k-means clusters where points denote the average z-scored peak area for both known and unknown metabolites across the samples and are colored by the eddy from which they were taken.
+Clusters have been ordered by number of metabolites in each group and the total is denoted in the panel titles.
+Both depth trends (mostly a net decrease in metabolites with depth) and eddy effects (cyclonic enrichment in clusters 1 and 2, anticyclone enrichment in cluster 4) are observable.
+The lower plot shows the individual known and unknown metabolites where points correspond to the FDR-corrected p-value estimated by the nonparametric Mann-Whitney U test and the log$_2$ fold-change calculated with the average peak area in the cyclone divided by the average peak area in the anticyclone.
+Colors have been assigned using the k-means clusters and shapes have been assigned based on the status of the mass feature as either a known metabolite that was matched to an authentic standard or an unidentified metabolite.
+The dashed line across the figure represents the 0.05 level of significance as a visual cue for metabolites above which the differences between the eddies are unlikely to be due to chance.
+DMSP = dimethylsulfoniopropionate, DMS-Ac = dimethylsulfonioacetate*
+
+```{r cluster cor w metadata, eval=FALSE}
+filled_file_metadata %>% 
+  filter(ms_run=="MC") %>% 
+  distinct(depth, euk_abund, sla_class) %>% 
+  left_join(kclust_df) %>%
+  filter(cluster==1) %>%
+  filter(sla_class=="Cyclone") %>%
+  # with(cor(mean_rank, euk_abund))
+  ggplot() +
+  geom_point(aes(x=euk_abund, y=mean_rank, color=sla_class))
+```
+
+The majority of the individual metabolites in these high-resolution depth profiles differed significantly between the two eddies (121/228, α = 0.05).
+Of those, 15 were enriched in the anticyclone and 106 were more abundant in the cyclone.
+As expected, all compounds enriched in the anticyclone were part of Cluster 4 and all those enriched in the cyclone belonged to either Cluster 1 or 2 (Figure 5).
+
+Many of the known metabolites enriched in the cyclone function as osmolytes in the cell and might be enriched due to the increased eukaryotic phytoplankton biomass.
+However, some metabolites such as isethionate, the reduced sulfur osmolytes dimethylsulfoniopropionate (DMSP) and dimethylsulfonioacetate (DMS-Ac), and the isomers homarine and trigonelline, were enriched in excess of biomass (Figure 5).
+A few metabolites more abundant in the anticyclone were also given putative identifications based on RT and *m/z* matching with internal standards run at a later time on the mass spectrometer.
+Of these, the putative arsenobetaine was the most significantly different among these with a peak area at the anticyclone DCM nearly quadruple that of the cyclone (Figure 5).
+Of note, the 173.0921 *m/z* mass feature noted above as strongly enriched in the anticyclone was among the most significantly different between the two eddies, along with two isomers at an *m/z* of 170.1176 and retention times around 8-9 minutes that increased by a factor of 1.5 in the anticyclone (putative formula C$_7$H$_{14}$N$_4$O).
+
+```{r putative 170.11756 formula, eval=FALSE}
+# Two possible formula: C9H16NO2 and C7H14N4O
+library(arrow)
+dset <- open_dataset("G:/My Drive/AllMeso/tmzMLs/MC_all_file_arr_pqds")
+
+# Too low-intensity for reliable isotope info
+dset %>%
+  filter(mz%between%pmppm(170.11757, 5)) %>%
+  filter(rt%between%c(7, 10)) %>%
+  collect() %>%
+  left_join(filled_file_metadata) %>%
+  ggplot() +
+  geom_line(aes(x=rt, y=int, group=filename, color=station))
+rtr <- c(8.7, 9.2)
+gly_data <- dset %>%
+  filter(rt%between%rtr) %>%
+  collect()
+mzr <- pmppm(170.11756, 10)
+reduce(list(
+  gly_data[mz%between%mzr] %>%
+    group_by(filename) %>%
+    summarise(M=sum(int)),
+  gly_data[mz%between%(mzr+1.003355)] %>%
+    group_by(filename) %>%
+    summarise(C13=sum(int))
+), merge, all.x=TRUE) %>%
+  pivot_longer(C13) %>%
+  mutate(prop=value/M) %>%
+  print(n=Inf)
+
+# But there are useful MS2 fragments
+library(RaMS)
+mz_group <- function(mz_vals, ppm){
+  group_vec <- numeric(length(mz_vals))
+  group_num <- 1L
+  init_vec <- mz_vals
+  names(init_vec) <- seq_along(init_vec)
+  while(length(init_vec)>0){
+    mz_i <- init_vec[1]
+    err <- mz_i*ppm/1000000
+    mz_idxs <- init_vec>mz_i-err & init_vec<mz_i+err
+    group_vec[as.numeric(names(mz_idxs)[mz_idxs])] <- group_num
+    init_vec <- init_vec[!mz_idxs]
+    group_num <- group_num+1L
+  }
+  group_vec
+}
+msmsdata_MC <- "../mzMLs/pos/MSMS" %>%
+  list.files(, full.names = TRUE, pattern = "180205") %>%
+  grabMSdata()
+msmsdata_MC$MS2[premz%between%pmppm(170.11757, 10)] %>%
+  mutate(mz_group=mz_group(fragmz, 5)) %>%
+  group_by(mz_group, filename) %>%
+  summarize(fragmz=mean(fragmz), int=mean(int)) %>%
+  filter(n()>1) %>%
+  group_by(mz_group) %>%
+  summarise(fragmz=mean(fragmz), int=mean(int)) %>%
+  filter(fragmz<170) %>%
+  arrange(desc(int))
+# Top two frags have formulae of 125.1075 = C7H13N2
+# and 89.07148 = C3H9N2O, both requiring 2 N atoms
+```
+
+## Hawaiian Eddy Experiment data reveals a different community and response to mesoscale eddies
+
+Given the strong signals detected in the samples from the MESO-SCOPE cruise both across the eddy transect as well as in the high-resolution DCM sampling, we expected to find similar results in an analogous dataset.
+Samples were collected during a 2018 cruise on the R/V *Falkor* that again targeted both a cyclonic and an anticyclonic eddy in the North Pacific Subtropical Gyre near Station ALOHA as part of the Hawaiian Eddy Experiment (HEE).
+
+```{r falkor nmds and permanova}
+set.seed(123)
+scaled_fk_data <- untarg_peaks %>%
+  # filter(filename!="190715_Smp_FK180310S62C1-DCM_B.mzML") %>%
+  left_join(filled_file_metadata %>% distinct(ms_run, station, filename, depth), 
+          by="filename") %>%
+  filter(ms_run=="FK") %>%
+  mutate(depth=factor(depth, levels=c("Surface", "DCM"))) %>%
+  select(feature, filename, norm_area) %>%
+  group_by(feature) %>%
+  filter(sum(norm_area==0)<n()*0.9)
+metab_mat_fk <- scaled_fk_data %>%
+  mutate(norm_area=scale(norm_area)[,1]) %>%
+  # mutate(norm_area=rank(norm_area)) %>%
+  pivot_wider(names_from = "feature", values_from = norm_area) %>%
+  column_to_rownames("filename") %>%
+  data.matrix()
+env_vec <- pull(filled_file_metadata, sla_corr, filename)[rownames(metab_mat_fk)]
+nmds_fk <- metab_mat_fk %>%
+  vegan::metaMDS(k = 2, distance = "manhattan", trace = 0, autotransform = FALSE,
+                 noshare = FALSE, wascores = FALSE) %>%
+  vegan::MDSrotate(env_vec)
+adon_expl <- data.frame(filename=rownames(metab_mat_fk)) %>%
+  left_join(filled_file_metadata %>% select(filename, sla_class, depth), by="filename")
+adon_fk <- vegan::adonis2(
+  metab_mat_fk~sla_class+depth, data = adon_expl, by = "margin", 
+  permutations = 999, method = "manhattan")
+```
+
+```{r figure falkor nmds}
+rot_mat <- matrix(c(sqrt(2)/2, sqrt(2)/2, -sqrt(2)/2, sqrt(2)/2), ncol=2)
+
+fk_nmdsplot <- nmds_fk$points%*%rot_mat %>%
+  as.data.frame() %>%
+  setNames(c("MDS1", "MDS2")) %>%
+  rownames_to_column("filename") %>%
+  left_join(filled_file_metadata %>% distinct(filename, sla_class, depth)) %>%
+  mutate(MDS1=MDS1*-1) %>%
+  mutate(MDS2=MDS2*-1) %>%
+  mutate(fill_col=case_when(
+    depth=="Surface" & sla_class=="Cyclone"~"25 m Cyclone",
+    depth=="DCM" & sla_class=="Cyclone"~"25 m Anticyclone",
+    depth=="Surface" & sla_class=="Anticyclone"~"DCM Cyclone",
+    depth=="DCM" & sla_class=="Anticyclone"~"DCM Anticyclone"
+  )) %>%
+  ggplot(aes(label=filename)) +
+  geom_vline(xintercept = 0) +
+  geom_hline(yintercept = 0) +
+  geom_point(aes(x=MDS1, y=MDS2, fill=fill_col, shape=sla_class), size=6, color="white") +
+  scale_fill_manual(breaks=c("25 m Cyclone", "25 m Anticyclone", "DCM Cyclone", "DCM Anticyclone"),
+                    values=c("#8989c1", "#343489", "#b57c7c", "#762020")) +
+  scale_shape_manual(breaks=c("Cyclone", "Anticyclone"), values=c(25, 24)) +
+  labs(color="Corr.\nSLA\n(cm)", shape="Depth", x="NMDS axis 1", y="NMDS axis 2") +
+  theme_bw() +
+  annotate("label", x=-Inf, y=-Inf, hjust=0, vjust=0, label.r = unit(0, "in"),
+           label=paste("Stress =", round(nmds_fk$stress, digits = 3))) +
+  annotate("label", x=-Inf, y=Inf, hjust=0, vjust=1, label.r = unit(0, "in"),
+           label=expression(atop(paste("R"[sla]^2, "=.090, p=0.017"),
+                                 paste("R"[depth]^2, "=.251, p<0.001")))) +
+  guides(fill=guide_legend(title = NULL, override.aes = list(
+    fill=c("#8989c1", "#b57c7c", "#343489", "#762020"), color="white",
+    shape=c(25, 24, 25, 24)
+  )), shape="none") +
+  theme(legend.key.spacing.y = unit(6, "pt")) +
+  coord_fixed()
+
+ggsave("fk_nmdsplot.png", plot = fk_nmdsplot, device = "png",
+       path = "figures", width = 6.875, height = 4, units = "in", dpi = 300,
+       bg="white")
+```
+
+The HEE data was characterized by high inter-replicate variability relative to the samples from the MESO-SCOPE cruise, with 25 meter samples in particular highly variable in multivariate space (Figure 6).
+Despite this, we still saw the importance of sea level anomaly as a significant explanatory factor in the dataset (R$^2_{SLA} = 0.090$, p-value = 0.017) as well as the larger depth differences between the surface (25 meters) and DCM (110 - 120 meters) (R$^2_{depth} = 0.251$, p-value << 0.001).
+
+![](figures/fk_nmdsplot.png)
+
+*Figure 6: Non-metric multidimensional scaling (NMDS) plot from the Hawaiian Eddy Experiment cruise data, in which points correspond to individual samples and have been colored and shaped by their source eddy status and shaded by the depth from which they were collected.
+The NMDS stress value has been reported in the bottom left corner, while PERMANOVA R$^2$ and p-values are reported in the top left.
+Samples from 25 meters deep are visibly distinct from the deep chlorophyll maximum (DCM) samples and an SLA signal is visible in the 25 meter samples only.*
+
+```{r top20compounds stacked barplot, eval=FALSE}
+scaled_fk_data %>%
+  ungroup() %>%
+  arrange(desc(norm_area)) %>%
+  filter(feature%in%head(unique(feature), 20)) %>%
+  left_join(filled_file_metadata) %>%
+  left_join(untarg_feats) %>%
+  mutate(feature=ifelse(is.na(compound_name), feature, compound_name)) %>%
+  mutate(feature=fct_inorder(feature)) %>%
+  arrange(depth, filename) %>%
+  mutate(filename=str_remove(filename, "190715_Smp_FK180310")) %>%
+  mutate(filename=str_remove(filename, "-25m_|-DCM_")) %>%
+  mutate(filename=fct_inorder(filename)) %>%
+  ggplot() +
+  geom_col(aes(x=filename, y=norm_area, fill=feature)) +
+  theme(axis.text.x = element_text(angle=90, hjust = 1, vjust=0.5)) +
+  facet_wrap(~depth, scales="free_y", ncol=1)
+```
+
+When analyzed separately as distinct depths, the 25 meter samples had higher variances explained by eddy (R$^2$=.26) and a lower p-value (0.013), while the DCM samples were no longer likely to be distinct between the two eddies (p-value = 0.171) (Figure 6, Supplemental Figure 4).
+
+```{r falkor concentrations comparison to MESO-SCOPE, eval=FALSE}
+fk_targ <- targ_peaks %>% 
+  left_join(filled_file_metadata) %>%
+  filter(ms_run%in%c("FK", "MT")) %>%
+  filter(samp_type=="Smp") %>%
+  select(compound_name, filename, nM, depth, ms_run)
+fk_targ %>%
+  group_by(depth, compound_name, ms_run) %>%
+  summarise(nM=mean(nM)) %>%
+  arrange(desc(nM)) %>%
+  pivot_wider(names_from = c(ms_run, depth), values_from = nM) %>%
+  print(n=Inf)
+cmpd_comp_gp <- fk_targ %>%
+  filter(depth!="Deep") %>%
+  ggplot() +
+  geom_boxplot(aes(x=ms_run, y=nM)) +
+  facet_grid(compound_name~depth, scales="free_y")
+ggsave("cmpd_comp_gp.pdf", cmpd_comp_gp, device = "pdf", width = 6, height = 50, limitsize = FALSE)
+# Visual assessment of boxplots from the above output:
+# Compounds more abundant in Falkor: citrulline, aspartic acid, ornithine, threonine, tyrosine, malic acid
+# Compounds more abundant in MESO: 4-aminobutyric, adenine, gonyol, 4-hydroxyisoleucine, alanine, leucine, nicotinic acid
+```
+
+We were also able to use the HEE dataset to test whether the compounds detected as significantly different in the MESO-SCOPE dataset were also different in this eddy pair.
+We expected to find abundant TMAO and more hydroxyisoleucine at the surface than the DCM in addition to most compounds enriched in the cyclone DCM with biomass, especially the six compounds mentioned above in the results of Figure 5 (trigonelline, homarine, DMS-Ac, DMSP, taurine, and isethionic acid).
+
+```{r abundant TMAO and hydroxyisoleucine depth trends, eval=FALSE}
+fk_targ <- targ_peaks %>% 
+  left_join(filled_file_metadata) %>%
+  filter(ms_run%in%c("FK")) %>%
+  filter(samp_type=="Smp") %>%
+  select(compound_name, filename, nM, depth, ms_run)
+
+fk_targ %>%
+  filter(compound_name=="Hydroxyisoleucine") %>%
+  left_join(filled_file_metadata) %>%
+  aov(formula=nM~depth) %>% summary()
+  # ggplot(aes(x=depth, y=nM, color=station)) +
+  # geom_jitter(width=0.2)
+  # geom_boxplot()
+
+fk_targ %>%
+  filter(compound_name%in%c("Homarine", "Trigonelline", "Dimethylsulfonioacetate", "Dimethylsulfoniopropionate", "Taurine", "Isethionic acid", "Arsenobetaine")) %>%
+  left_join(filled_file_metadata) %>%
+  group_by(filename) %>%
+  mutate(nM=nM/sum(nM)) %>%
+  ungroup() %>%
+  filter(depth=="DCM") %>%
+  nest(data=-compound_name) %>%
+  mutate(aovout=map(data, ~broom::tidy(aov(.x$nM~.x$sla_class)))) %>%
+  unnest(aovout) %>% filter(term==".x$sla_class") %>%
+  mutate(p.value=p.adjust(p.value, method="fdr"))
+  ggplot(aes(x=sla_class, y=nM)) +
+  # geom_boxplot() +
+  geom_point() +
+  facet_wrap(~compound_name, scales="free_y")
+
+untarg_feats %>%
+  filter(mzmed%between%pmppm(170.11756, 5) | mzmed%between%pmppm(173.09211, 5)) %>%
+  filter(startsWith(feature, "FK")) %>% 
+  left_join(untarg_peaks) %>%
+  left_join(filled_file_metadata) %>%
+  nest(data=-feature) %>%
+  mutate(aovout=map(data, ~broom::tidy(aov(.x$norm_area~.x$sla_class)))) %>%
+  unnest(aovout) %>% filter(term==".x$sla_class") %>%
+  mutate(p.value=p.adjust(p.value, method="fdr"))
+  select(feature, sla_class, norm_area) %>% 
+  pivot_wider(names_from = sla_class, values_from = norm_area, values_fn = mean)
+  ggplot(aes(x=sla_class, y=norm_area)) +
+  # geom_boxplot() +
+  geom_point() +
+  facet_wrap(~feature, scales="free_y")
+```
+
+As expected, TMAO was again one of the most abundant metabolites detected in the particulate matter with concentrations around 1.2 nM at 25 meters and 0.4 nM at the DCM, second only to the high levels of the amino acid glycine (2 nM at 25 meters, 0.9 nM DCM), though the differences between 25 meters and DCM were not significant for either compound.
+We also detected a significant difference between depths for hydroxyisoleucine (25 meter mean = 0.32 nM, DCM mean = 0.21 nM, t-test p-value=0.022 with n=12 samples at each depth).
+These results imply that the overall depth structure of the metabolome of the gyre is relatively fixed for the most abundant compounds.
+
+The eddy effects, on the other hand, were much less strong during this cruise.
+Of the six known metabolites with the strong enrichment in the cyclone at the DCM in the MESO-SCOPE cruise, only isethionic acid was significantly different in the HEE dataset (isethionic acid t-test p-value$_{FDR}$ = 0.043, other five p-values > 0.25).
+This difference in isethionic acid concentration was only found after normalizing each sample to the sum of all metabolites in the sample to control for biomass.
+Arsenobetaine was again found to be strongly enriched in the anticyclone at the DCM both when normalizing to biomass and when not, just as in MESO-SCOPE.
+Surprisingly, the 173.0921 *m/z* mass feature was also slightly enriched but this time in the *cyclone* DCM with peak areas approximately 1.4 times the values in the anticyclone (t-test p-value = 0.067, Mann-Whitney p-value = 0.045).
+The 170.1176 *m/z* mass feature was not detected in the HEE samples.
+
+# Discussion
+
+## Multivariate approaches reveal metabolome-wide shifts across pairs of eddies of opposite polarity
+
+We measured the metabolome of samples collected across two sets of adjacent eddies of opposite polarity to explore the effect of sea level anomaly (SLA) on metabolite composition and found that the altered biogeochemistry and microbial community composition between cyclonic and anticyclonic eddies explain a significant portion of the observed variations in their particulate metabolomes.
+
+In all of the datasets analyzed here, we detected a significant difference in the composition of the metabolome between the adjacent eddies.
+This effect was strongest in the samples taken during the Lagrangian stations at the center of each eddy in the 2017 MESO-SCOPE cruise, with nearly a quarter of the total variance explained by the eddy from which the samples were taken.
+In the samples taken along the eddy transect, the largest effect was detected in the deepest samples taken from 175 meters and from the DCM, with much less of a response at 25 meters.
+In the HEE samples, however, the opposite was true with a larger SLA response at 25 meters than that at the DCM.
+In each case, the differences introduced by SLA were smaller than those of depth when compared directly, with even the high-resolution sampling around the DCM finding slightly more variance explained by the 40 meter difference in sampling depth than the 40 centimeter difference in SLA.
+This result agrees well with prior research by @Heal2021 and @Kumler2023 showing the large effect of sampling depth on the metabolome.
+
+Although possible that the observed differences in metabolomes across adjacent eddies differing in polarity are due to latitudinal shifts or simply background variation in the gyre environment, the transect data in particular implies that this is unlikely to be the case.
+Samples taken when the absolute value of the SLA was less than 5 cm were more similar to each other than to those samples taken within the eddies despite larger differences in latitude.
+The transect sampled both outside of each eddy and between the two across a large spatial gradient (~4° latitude and 2° longitude), yet the non-eddy (absolute value of SLA < 5 cm) samples from all three locations grouped together, showed similar median metabolite concentrations, and were visually similar in the most abundant metabolite composition.
+These results imply that strong cyclonic and anticyclonic eddies represent endpoints in NPSG composition, in agreement with results from @Barone2019 which found the most extreme values in the Hawaii Ocean Time series data typically detected when eddies passed over Station ALOHA.
+We also detected the largest differences in the DCM metabolome at the exact center of the eddy, with Station 12 highly distinct from even the nearest stations, while no such stark difference was found in the anticyclone (Figures 1 and 2).
+This perhaps indicates that the exact center of the cyclone has a unique metabolome at the DCM relative to the rest of the transect, rather than existing as a smooth continuation.
+
+Shifts in total biomass were a large driver of metabolomic differences, with earlier work in the same location by @Barone2022 showing that particulate carbon, chlorophyll, and beam attenuation were all 20-80% higher in the DCM of the cyclone relative to the anticyclone.
+We see this reflected particularly well in the targeted metabolites measured here, with clear depth trends and a strong correlation between total metabolite concentration and particulate carbon values.
+However, even after controlling for biomass effects by normalizing each sample to the sum of signal measured within it we still found a significant SLA effect, likely due to shifts in the community composition and in particular the increased eukaryote presence in the cyclone DCM.
+
+```{r multimodel testing maybe, eval=FALSE}
+dist_mat <- untarg_peaks %>%
+  left_join(filled_file_metadata %>% distinct(ms_run, station, filename, depth), 
+            by="filename") %>% filter(ms_run=="MC") %>% 
+  mutate(depth=factor(depth, levels=c("DCMless20m", "DCMless10m", "DCM", 
+                                      "DCMplus10m", "DCMplus20m"))) %>%
+  arrange(feature, station, depth) %>%
+  mutate(filename=fct_inorder(filename)) %>%
+  group_by(feature) %>%
+  mutate(norm_area=rank(norm_area)) %>%
+  select(feature, filename, norm_area) %>%
+  pivot_wider(names_from=feature, values_from="norm_area") %>%
+  mutate(filename=str_remove_all(filename, "180205_Smp_|.mzML")) %>%
+  column_to_rownames("filename") %>%
+  dist(method = "manhattan") %>%
+  as.matrix()
+dist_mat %>%
+  as.data.frame() %>%
+  rownames_to_column("filename_a") %>%
+  pivot_longer(-filename_a, names_to="filename_b", values_to = "dist") %>%
+  mutate(filename_a=fct_inorder(filename_a)) %>%
+  mutate(filename_b=fct_inorder(filename_b)) %>%
+  filter(as.numeric(filename_b)>=as.numeric(filename_a)) %>%
+  ggplot() +
+  geom_raster(aes(x=filename_a, y=filename_b, fill=dist)) +
+  coord_fixed() +
+  scale_fill_gradient2(midpoint = 2200) +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle=90, hjust=1, vjust=0.5))
+
+expl_df <- filled_file_metadata %>% 
+  distinct(ms_run, station, filename, depth, samp_type) %>%
+  left_join(untarg_peaks, by="filename") %>%
+  filter(ms_run=="MC") %>%
+  filter(samp_type=="Smp") %>%
+  mutate(model1_groups=as.numeric(factor(depth))) %>%
+  mutate(model2_groups=case_when(
+    station=="L1" & depth %in% c("DCMless20", "DCMless10") ~ 1,
+    station=="L1" & depth=="DCM" | station=="L2" & depth=="DCMless20" ~ 2,
+    station=="L1" & depth=="DCMplus10" | station=="L2" & depth=="DCMless10" ~ 3,
+    station=="L1" & depth=="DCMplus20" | station=="L2" & depth=="DCM" ~ 4,
+    TRUE ~ 5
+  )) %>%
+  distinct(filename, model1_groups, model2_groups)
+
+vegan::adonis2(
+  dist_mat~model1_groups+model2_groups, data = expl_df, by = "margin", 
+  permutations = 999, method = "manhattan")
+```
+
+The NMDS plots of the high-resolution depth samples were also illustrative of sample dissimilarity with clear SLA and depth trends.
+The anticyclonic samples grouped tightly and showed little variance with depth or triplicate when compared to the cyclonic samples with the exception of DCM triplicate B, which was highly distinct for unknown reasons.
+In contrast, the cyclonic samples were much more variable as is perhaps expected when the biomass is in the form of larger phytoplankton that are less homogenous in the environment.
+One notable aspect of their clustering was the way in which the deepest samples (DCM plus 20m) grouped most closely to the anticyclonic samples, perhaps indicating that below the DCM the metabolome rapidly approaches a uniform deep-water signal.
+This result was noted previously in the 175 meter samples of the eddy transect in @Kumler2023, where samples from the deep euphotic zone showed greater intra-depth similarity than samples from the DCM or 25 meters.
+
+Given the confluence of both depth and SLA signals and the way that the boxplots of Figure 2 would confound the SLA signal with the depth variance, we used k-means clustering to group similar compounds and provide a reduced dimensionality space for visualization.
+This revealed four major patterns of metabolite response, with a majority of compounds responding to eddy state (Clusters 1, 2, and 4 in Figure 5).
+It is interesting to note that the sole cluster in which abundances in the anticyclone are greater than those of the cyclone is also the only cluster to generally increase in abundance with depth (Cluster 4), perhaps indicating that compounds more concentrated in the anticyclone are the same kind of degradation products and recalcitrant carbon typically found at depth.
+Cluster 3, which shows no clear SLA difference and only a slight change with depth, includes compounds which are either unaffected by the environment or simply lack an environmental signal due to large inter-sample variability.
+
+The Hawaiian Eddy Experiment (HEE) data upset many of the expectations [Wei: consider calculating the effect size, as mentioned above.
+Will: Not sure what Wei means by this.
+Should I do effect size calculations here?
+Or just above when we didn't see an effect?] we developed during the MESO-SCOPE cruise in the previous year.
+Most surprising was the large difference between the eddies detected at the *surface*, while the DCM samples were functionally indistinguishable.
+This contrasts directly with the results from MESO-SCOPE and the results in @Gleich2024, who found eddy-driven shifts in protistan community composition to be larger at depth than at the surface.
+However, @Dugenne2023 found differences in nitrogen fixation rate and nitrogen fixer composition varying widely throughout the upper water column.
+The large surface differences were especially surprising given the large inter-replicate differences between the HEE surface samples in this study, while the DCM samples tended to be much more consistent.
+
+## Univariate approaches highlight individual metabolites responding to lifted and depressed isopycnals
+
+```{r kheal comparison, include=FALSE}
+raw_kheal <- read_csv("https://raw.githubusercontent.com/kheal/Gradients1_SemiTargeted3/master/Tables/Manuscript_tables/SuppTables/Full_EnviroCulture_PeakAreas.csv")
+kheal_envirosamp <- raw_kheal %>%
+  filter(Column=="HILIC") %>%
+  mutate(polarity=ifelse(z>0, "pos", "neg")) %>%
+  distinct(MassFeature_Column, compound_name_kath=Identification, Confidence, mz, rt, polarity) %>%
+  mutate(rt=rt/60)
+
+untarg_feats %>%
+  filter(str_detect(feature, "FK")) %>%
+  select(feature, polarity, mzmed, rtmed, compound_name_will=compound_name) %>%
+  mutate(mz_lower=mzmed*(1-5/1e6)) %>%
+  mutate(mz_upper=mzmed*(1+5/1e6)) %>%
+  full_join(kheal_envirosamp, by=join_by(polarity, mz_lower<mz, mz_upper>mz)) %>%
+  filter(compound_name_kath=="Histidine") %>%
+  distinct(feature, MassFeature_Column, mzmed, rtmed, mz, rt, compound_name_will, compound_name_kath)
+
+```
+
+Several molecule- or pathway-specific narratives emerged from the metabolomic data.
+The untargeted approach used here allowed us to describe and characterize signals from small molecules whose identity was unknown, a particularly promising approach in open ocean gyres where the largest fraction of unknowns is found [@Heal2021].
+We found that our list of authentic standards covered fewer metabolites in anticyclonic eddies at all depths, with the molecular diversity of the cyclone much better characterized.
+Many of the trends detected persisted even after normalizing within each sample, indicating that the shifts discussed have implications beyond simple scaling with biomass.
+
+Of those molecules whose identity was known, the clearest response to SLA was that of the enigmatic osmolyte homarine at the DCM.
+This abundant compound increased approximately threefold in concentration from < 50 pM in the anticyclone to around 150 pM in the center of the cyclone in both the eddy transect and eddy center datasets.
+The pattern was weaker in the noisier HEE data but the largest concentrations were still detected in the cyclone and the median cyclone measurement was greater than the maximum value for the anticyclone.
+This molecule has a well-established role as an osmolyte in eukaryotic phytoplankton and cyanobacteria [@Gebser2013; @Dawson2020; @Heal2021; @Durham2022] as well as a documented decrease in abundance with depth [@Heal2021], indicating that its response to SLA is potentially due to differences in physical and chemical attributes that in turn control biomass, community composition, and recycling rates.
+Curiously, the distribution of homarine was also tightly correlated with that of trigonelline.
+Although they have structural similarity they are not known to have a relationship beyond their shared function as osmolytes.
+
+```{r homarine by SLA plot, eval=FALSE}
+targ_hom <- targ_peaks %>%
+  filter(compound_name=="Homarine") %>%
+  left_join(filled_file_metadata) %>%
+  filter(samp_type=="Smp")
+mt_hom_gp <- targ_hom %>%
+  filter(ms_run=="MT") %>%
+  filter(depth=="DCM") %>%
+  ggplot(aes(x=sla_corr, y=nM)) +
+  geom_hline(yintercept = 0) +
+  geom_smooth(method = "lm", formula="y~x", color="black") +
+  geom_point(aes(fill=sla_corr), pch=21, size=4) +
+  scale_fill_gradient2(
+    low = scales::muted("blue"), high = scales::muted("red"),  mid = "grey90",
+    aesthetics = "fill", guide = NULL
+  ) +
+  coord_cartesian(ylim = c(0, 0.24), xlim=c(-15.77, 24.99), expand = FALSE) +
+  labs(x="Sea level anomaly (cm)", y="Concentration of homarine (nM)") +
+  theme_bw()
+mc_hom_gp <- targ_hom %>%
+  filter(ms_run=="MC") %>%
+  ggplot(aes(x=station, y=nM, color=station)) +
+  geom_hline(yintercept = 0) +
+  geom_boxplot(linewidth=0.75) +
+  coord_cartesian(ylim = c(0, 0.24), xlim=c(0.5, 2.5), expand = FALSE) + 
+  scale_x_discrete(breaks = c("L1", "L2"), labels = c("Cyclone\ncenter", "Anticyclone\ncenter")) +
+  scale_color_manual(values = c(scales::muted("blue"), scales::muted("red")),
+                     breaks = c("L1", "L2")) +
+  theme_bw() +
+  theme(axis.text.y = element_blank(), axis.title.y = element_blank(),
+        legend.position = "none", axis.title.x = element_blank())
+egg::ggarrange(mt_hom_gp, mc_hom_gp, nrow=1, widths = c(2, 1))
+
+targ_hom %>%
+  filter(ms_run=="FK") %>%
+  filter(depth=="Surface") %>%
+  # with(t.test(formula=nM~sla_class))
+  # with(wilcox.test(formula=nM~sla_class))
+  ggplot(aes(x=sla_class, y=nM, color=sla_class)) +
+  geom_hline(yintercept = 0) +
+  geom_boxplot(linewidth=0.75) +
+  coord_cartesian(ylim = c(0, 0.5), xlim=c(0.5, 2.5), expand = FALSE) +
+  scale_color_manual(values = c(scales::muted("blue"), scales::muted("red")),
+                     breaks = c("Cyclone", "Anticyclone")) +
+  theme_bw()
+
+```
+
+Isethionate is a known osmolyte thus far found exclusively in eukaryotes [@Durham2022] and strongly associated with a few diatoms in particular [@Heal2021].
+This compound was enriched in the cyclone of the DCM in both the MESO-SCOPE and HEE cruises along with its precursor taurine, validating earlier results from @Barone2022 that showed strong enrichment of eukaryotic phytoplankton at the DCM and *Pseudo-nitzschia* in particular.
+
+```{r isethionate and taurine, eval=FALSE}
+targ_peaks %>%
+  filter(compound_name%in%c("Isethionic acid", "Taurine")) %>%
+  filter(filename!="180821_Smp_MS11C215m_A.mzML") %>%
+  left_join(filled_file_metadata) %>%
+  filter(samp_type=="Smp") %>%
+  mutate(depth=fct_inorder(depth)) %>%
+  mutate(ms_run=factor(ms_run, levels=c("MT", "MC", "FK"))) %>%
+  mutate(sla_class=factor(sla_class, levels=c("Anticyclone", "Neither", "Cyclone"))) %>%
+  # group_by(compound_name, ms_run, sla_class, depth) %>%
+  # summarise(nM=mean(nM)) %>%
+  # pivot_wider(names_from = sla_class, values_from = nM)
+  ggplot() +
+  geom_boxplot(aes(x=depth, y=nM, color=sla_class)) +
+  facet_grid(compound_name~ms_run, scales="free")
+```
+
+The strong SLA signal detected among the 175 meter samples was partially surprising to us given our expectation about the strongest effect at the DCM where eddy effects lift large concentrations of nutrients above the 1% light level.
+Instead, we found more compounds overall to be significantly different across the eddy transect in the 175 meter samples than we did at the DCM (α=0.05, 43 compounds at 175 meters vs 22 at the DCM).
+This was likely driven by the enhancement of organic matter and heterotrophic picoplankton below the DCM as seen in @Barone2019, with the enrichment of nucleobases in the 175 meter samples additionally hint at increases in bacterial biomass given that nucleobases tend to be highly abundant in bacterial cultures [@Heal2021].
+The abundance of a mass feature putatively identified as acetylcholine in the 175 meter cyclone samples followed the same general trend as arsenobetaine, both enriched at depth and in the anticyclone, and may be another compound that results from heterotrophic degradation of phytoplankton metabolites [@Durham2022].
+These results raise important questions about the depth at which the community is isolated from SLA effects, with additional data from @Barone2022 and @Gleich2024 indicating that even at 250 meters eddy effects are still discernable.
+
+Differences between the two cruises could be due to differences in eddy age and/or the different period of the year when sampling took place during the two cruises (March/April in 2018 instead of the June/July sampling in 2017) resulting in shifts in community composition and response.
+@Barone2022 documents the 2017 cruise community composition, reporting abundant Prymnesiophyceae at both 15 meters and the DCM, with other significant contributions from dinoflagellates, diatoms, and green algae.
+In the 2018 cruise both haptophytes and dinoflagellates again dominated the eukaryotic fraction, though surprisingly little was contributed by diatoms [@Gleich2024].
+@Dugenne2023 and @Harke2021 also highlight uncontrolled differences between the two sets of eddies based on their origin and age.
+
+Missing links between the environment with its associated community and the composition of particulate matter make it difficult to estimate this important control on the marine carbon cycle.
+This is especially true for dynamic and short-lived environments such as eddies.
+Metabolites offer a way to bridge this gap in knowledge but the relative impact of eddies on their contribution and fate was not previously estimated.
+In this work, we have shown the sensitivity of the metabolome to various environmental and community factors and contrasted it to the well-established effect of depth to constrain the importance of including eddy effects in models of elemental cycling.
+
+# Conclusions
+
+This study reports how changes in the biogeochemistry and community composition due to mesoscale eddies affect the metabolome and thereby the composition of particulate organic matter.
+A transect across adjacent mesoscale eddies of opposite polarity showed that many metabolites track closely with metrics of overall biomass, with eddy effects stronger at 175 meters than at 15 meters or the deep chlorophyll maximum (DCM).
+High-resolution depth sampling of the DCM at the center of each eddy elucidated several known and unknown biomarkers likely corresponding to the previously documented increase in eukaryotic phytoplankton in the cyclone.
+Metabolite clusters aligned well with expected trends in abundance with depth and between the two eddies, with convergence below the DCM towards a general deep-sea metabolite signal for many metabolites.
+By contrasting these effects with a follow-up analysis in the centers of a separate pair of mesoscale eddies of opposite polarity the following year, we learned that the impact of eddies may be influenced by their origin, age, and period of the year.
+Finally, the metabolites with the strongest responses to eddy effects were unknown, and a relatively smaller proportion of known metabolites in the anticyclone demonstrates the utility of untargeted metabolomics in exploring environmental variation and the large amount of information that is missed by only investigating known compounds.
+
+# Methods
+
+## Cruise information
+Samples were collected from two cruises in the North Pacific Subtropical Gyre near Station ALOHA that targeted strong mesoscale eddy features as described in @Dugenne2023 and @Gleich2024.
+Briefly, the 2017 MESO-SCOPE cruise (Microbial Ecology of the Surface Ocean-Simons Collaboration on Ocean Processes and Ecology, KM1709 on the *R/V Kilo Moana*) consisted of a transect across adjacent cyclonic and anticyclonic eddies as well as two long-term Lagrangian stations at the center of each eddy.
+The cyclonic eddy had a maximum negative sea level anomaly (SLA) of -20 centimeters and the anticyclonic eddy reached +24 centimeters.
+The transect samples were taken at various times of day as the ship transected the adjacent eddies while the eddy center samples were all collected between 5 and 8 pm.
+In 2018, the Hawaiian Eddy Experiment (HEE, FK180310 on the *R/V Falkor*) targeted new cyclonic and anticyclonic eddies in approximately the same location with samples taken at the center of each eddy (maximum negative anomaly in the cyclone = -15 cm, maximum positive anomaly in the anticyclone = +26 cm) (Figure 1).
+
+## Biogeochemical data
+
+Biogeochemical measurements were collected as described in @Barone2022 and @Dugenne2023 mostly following protocols used by the HOT program (http://hahana.soest.hawaii.edu/hot/methods/results.html).
+Briefly, all environmental data was collected via CTD rosette except SLA which was measured via satellite.
+Particulate carbon and particulate nitrogen were measured using an elemental analyzer.
+Nitrate + nitrite (N+N) was measured on an autoanalyzer except where concentrations were below 100 nM in which case they were measured using chemiluminescence.
+Soluble reactive phosphorus (phosphate, PO$_4^{3-}$) was measured on an autoanalyzer or following the magnesium induced coprecipitation method.
+Values were interpolated linearly to the depths at which metabolite samples were taken using all data collected at the given station.
+Two particulate carbon values were determined to be spurious, with concentrations 2-3 times higher than typical Station ALOHA values, and were instead estimated using a best-fit linear regression against beam attenuation.
+
+## Sample collection for particulate metabolites
+Samples were obtained using the onboard CTD rosette to collect water from 15 meters, the deep chlorophyll maximum (DCM), and 175 meters during the MESO-SCOPE eddy transect and from the DCM ±10 and ±20 meters during the eddy center sampling in the MESO-SCOPE cruise and from 25 meters and the DCM during the HEE cruise.
+The DCM was determined visually from fluorometer data during the CTD downcast and Niskin bottles were tripped during the return trip to the surface.
+Seawater from each depth was sampled in triplicate by firing one Niskin bottle for each sample.
+Samples were brought to the surface and decanted into prewashed (3x with DI, 3x with sampled seawater) polycarbonate bottles for filtration.
+10L samples were filtered by peristaltic pump onto 142mm 0.2 µm Durapore filters held by polycarbonate filter holders on a Masterflex tubing line.
+Pressures were kept as low as possible while still producing a reasonable rate of flow through the filter, approximately 250-500 mL per minute.
+Samples were then removed from the filter holder using solvent-washed tweezers and placed into pre-combusted aluminum foil packets that were then flash-frozen in liquid nitrogen before being stored at -80 °C until extraction.
+A methodological blank was also collected by running filtrate through a new filter and then treated identically to the samples.
+
+## Metabolite sample extraction
+Extraction followed a modified Bligh & Dyer approach as detailed in @Boysen2018.
+Briefly, filters were randomized and added to PTFE centrifuge tubes with a 1:1 mix of 100 µm and 400 µm silica beads, approximately 2 mL -20 °C Optima-grade dichloromethane, and approximately 3 mL -20 °C 1:1 methanol/water solution (both also Optima-grade).
+Isotope labeled extraction standards were also added during this step (Supplemental Table 1).
+The samples were then bead-beaten three times, followed by triplicate extraction of the aqueous layer and addition with fresh methanol/water mixture with additional bead-beatings in between.
+Samples were then dried down under ultrapure nitrogen gas and warmed using a Fisher-Scientific Reacti-Therm module.
+Dried aqueous fractions were reconstituted in 380 µL Optima-grade water and amended with 20 µL isotope-labeled injection standards (Supplemental Table 1) to measure the variability introduced by chromatography and ionization.
+The reconstituted fraction was syringe-filtered into precombusted glass inserts in HPLC vials to remove any potential clogging material.
+Samples were then additionally diluted 1:1 with Optima-grade water to prevent overloading on the column and reduce salt effects.
+
+A pooled sample was created by combining 20 µL of each sample into the same HPLC vial, and a 1:1 dilution with water created a half-strength pooled sample from that to assess matrix effects and obscuring variation [@Boysen2018].
+Also run alongside the environmental samples were samples of authentic standards split into two mixes (Supplemental Table 2).
+Standards were dissolved in Optima-grade water to help with identification and into an aliquot of the pooled sample to quantify matrix response factors.
+HPLC vials containing the samples were frozen at -80 °C until thawing shortly before injection.
+
+## HPLC-MS methods
+Separate HPLC-MS runs were used for the MESO-SCOPE eddy transect, the MESO-SCOPE eddy center, and the HEE cruise data.
+Eddy center samples were run in February 2018, eddy transect samples were run later during August of that year, and the HEE samples were run in July 2019.
+Each run was treated as a single batch and injected in sequence while maintaining the randomization that had occurred during extraction to minimize chromatographic shifts from solvent or column switching.
+
+For each run, a Waters Acquity I-Class UPLC with a SeQuant ZIC-pHILIC column (5 µm particle size, 2.1 mm x 150 mm, from Millipore) was used with 10 mM ammonium carbonate in 85:15 acetonitrile to water (Solvent A) and 10 mM ammonium carbonate in 85:15 water to acetonitrile (Solvent B) at a flow rate of 0.15 mL/min.
+The column was held at 100% A for 2 minutes, ramped to 64% B over 18 minutes, ramped to 100% B over 1 minute, held at 100% B for 5 minutes, and equilibrated at 100% A for 25 minutes (50 minutes total).
+The column was maintained at 30 °C.
+The injection volume was 2 µL for samples and standard mixes.
+When starting a batch, the column was equilibrated at the starting conditions for at least 30 minutes.
+To improve the performance of the HILIC column, we maintained the same injection volume, kept the instrument running water blanks between samples as necessary, and injected standards in a representative matrix (the pooled sample) in addition to standards in water.
+After each batch, the column was flushed with 10 mM ammonium carbonate in 85:15 water to acetonitrile for 20 to 30 minutes.
+
+The Waters Acquity UPLC was coupled to a Thermo Q Exactive HF hybrid Orbitrap high resolution mass spectrometer equipped with a heated electrospray ionization source (H-ESI).
+The H-ESI voltage was set to 3.3 kV and sheath gas, auxiliary gas, and sweep gas flow rates were set at 16, 3, and 1, respectively.
+The capillary and auxiliary gas heater temperatures were maintained at 320°C and 100°C, respectively.
+Full scan analyses were performed with polarity switching and a scan range of 60 to 900 *m/z* at a resolution of 60,000.
+The instrument was mass calibrated at 200 *m/z* before each run began and once again during the eddy transect sample set to ensure calibrations were always performed within 3-4 days.
+All data files were then converted to an open-source mzML format and vendor centroided via Proteowizard’s msConvert tool.
+All mzML files have been uploaded to Metabolomics Workbench under Project ID PR001738.
+
+## Targeted metabolomic methods
+
+Compounds for which we had an authentic standard (Supplemental Table 2) were manually integrated using Skyline and MSDIAL with the standard mixes used to ensure the correct peak was integrated by matching retention time.
+Compounds that were visually determined to have poor peak quality in the samples were removed from the downstream analysis.
+Skyline was used for both the eddy center samples and the HEE cruise data while MSDIAL was used for the eddy transect data due to the larger number of files in that dataset.
+Raw peak areas were then normalized to their best-matched internal standard (Supplemental Table 1) per @Boysen2018 except that all compounds were normalized to their best match no matter how minimal the improvement.
+Normalized peaks were then calibrated via a single-point standard curve as determined uniquely for each compound from the authentic standard mixes and used to convert normalized peak area into moles per µL injected and scaled to the amount of seawater filtered to provide a final estimate of environmental concentration for each compound in moles per liter of seawater filtered.
+
+```{r supp table 1 standards list}
+clean_stans <- read_csv("../metadata/clean_stans.csv")
+raafay_notes <- read_csv("../targeted/raw_data/Raafay_HILIC_notes.csv") %>%
+  select(compound_name, in_samps)
+
+IS_table <- clean_stans %>%
+  filter(compound_type=="Internal Standard") %>%
+  select(compound_name, formula, stan_mz, stan_rt, conc_um, date_added, polarity) %>%
+  mutate(ms_run=case_when(
+    date_added<180205~"All",
+    date_added<=180821~"MESO-SCOPE Eddy Transect and HEE",
+    date_added<=190715~"HEE only",
+    TRUE~"None"
+  )) %>%
+  filter(ms_run!="None") %>%
+  mutate(polarity=ifelse(polarity=="pos", "Positive", "Negative")) %>% 
+  mutate(stan_type=case_when(
+    compound_name%in%c("L-Cysteic acid, 2H3", "3-Sulfolactate, 13C3", "Sulfoacetic acid, 13C2", "Isethionic acid, 13C2", "Taurine, 2H4")~"Extraction",
+    TRUE~"Internal"
+  )) %>%
+  arrange(stan_mz) %>%
+  select(`Compound name`=compound_name, `Molecular formula`=formula,
+         `m/z ratio`=stan_mz, `Retention time (min)`=stan_rt,
+         `Concentration (uM)`=conc_um, `Polarity integrated`=polarity,
+         `Which runs`=ms_run, `Extraction or internal`=stan_type)
+
+Auth_table <- clean_stans %>%
+  mutate(ms_run=case_when(
+    date_added<180205~"All",
+    date_added<=180821~"MESO-SCOPE Eddy Transect and HEE",
+    date_added<=190715~"HEE only",
+    TRUE~"None"
+  )) %>%
+  filter(ms_run!="None") %>%
+  left_join(raafay_notes) %>%
+  mutate(polarity=ifelse(polarity=="pos", "Positive", "Negative")) %>%
+  arrange(stan_mz) %>%
+  select(`Compound name`=compound_name, `Molecular formula`=formula,
+         `m/z ratio`=stan_mz, `Retention time (min)`=stan_rt,
+         `Concentration (uM)`=conc_um, `Polarity integrated`=polarity,
+         `Standard mix`=mix, `Which runs`=ms_run, `Good quality`=in_samps)
+
+wb <- createWorkbook(creator = "William Kumler, wkumler@uw.edu")
+addWorksheet(wb, sheetName = "Table 1 Internal Standards")
+writeData(x=IS_table, wb = wb, sheet = "Table 1 Internal Standards")
+
+addWorksheet(wb, sheetName = "Table 2 Authentic Standards")
+writeData(x=Auth_table, wb = wb, sheet = "Table 2 Authentic Standards")
+
+saveWorkbook(wb, file = "supplement/Supplemental Tables.xlsx", overwrite = TRUE)
+```
+
+## Untargeted metabolomic methods
+
+Samples were also processed via the untargeted workflow detailed in @Kumler2023.
+In short, XCMS [@R-xcms] was used for feature detection with CentWave peakpicking, Obiwarp alignment, and peak density correspondence/grouping.
+This was followed by manual inspection of the extracted ion chromatogram for each mass feature by an expert for quality with only those peaks classified as "Good" used in the analysis here.
+Features were then matched with and normalized to an internal standard (Supplemental Table 1) as described above.
+
+## Statistics
+
+All statistics were run in R version 4.3.3.
+We used multivariate statistics provided by the vegan package [@R-vegan] to assess the overall impact of various metrics across the metabolome followed by univariate statistics to investigate individually significant responses.
+Non-parametric tests were used when data were unlikely to obey parametric assumptions and permutational statistics were preferred wherever possible.
+Multiple comparison problems were controlled where necessary using a false-discovery rate correction [@Benjamini1995].
+For the analyses involving the eddy transect samples, sea level anomaly was treated as a continuous variable with the exception of the PERMANOVA, for which SLA was categorized as cyclonic if less than -5 cm and anticyclonic if greater than 5 cm, with a third category for stations in between [@Anderson2001].
+Both the MESO-SCOPE eddy center and HEE analyses treated SLA as a categorical variable.
+
+# References
+
+```{r packages, eval=FALSE}
+packages_used <- list.files("..", pattern = "*\\.R$|*\\.Rmd$", recursive = TRUE) %>%
+  str_subset("old|scripts", negate=TRUE) %>%
+  map(function(file_name){
+    v <- readLines(paste0("../", file_name))
+    libs <- unique(str_extract(v, "(?<=library\\().*?(?=\\)|,)"))
+    refs <- unique(str_extract(v, "[[:alnum:]]*(?=::)"))
+    pkgs <- unique(c(na.omit(libs), na.omit(refs)))
+  }) %>% unlist() %>% unique()
+knitr::write_bib(packages_used[nchar(packages_used)>0], file = "r-packages.bib")
+```
+
+<div id="refs"></div>
+
+# Supplement
+## Supplemental methods: Trichodesmium bloom in MS11C215mA
 
@@
 
-## Particulate organic matter compositional shifts across eddy transect
-
-In addition to absolute shifts in metabolite concentration due to biomass differences, we also explored compositional shifts in the metabolomes by assessing the fraction of the total metabolite pool contributed by each compound across the eddy transect.
-This allowed us to isolate the signal due to shifts in community composition and organismal response from those introduced by changes in biomass of the community as a whole.
-
-```{r regression relative abundance coefficients, eval=FALSE}
-reg_coefs <- untarg_peaks %>%
-  # filter(feature!="MS_NEG_0155") %>%
-  left_join(filled_file_metadata %>% distinct(filename, ms_run, station, depth, sla_corr)) %>%
-  filter(ms_run=="MT") %>%
-  filter(depth%in%c("Surface", "DCM", "Deep")) %>%
-  filter(feature%in%unique(feature)) %>%
-  group_by(filename) %>%
-  mutate(norm_area=norm_area/sum(norm_area)) %>%
-  ungroup() %>%
-  nest(data=-c(depth, feature)) %>%
-  mutate(test=map(data, ~broom::tidy(lm(.x$norm_area~.x$sla_corr)), .progress = TRUE)) %>%
-  unnest(test)
-reg_coefs %>%
-  filter(str_detect(term, "sla_corr")) %>%
-  mutate(p.value=p.adjust(p.value, method="fdr")) %>%
-  arrange(p.value) %>%
-  left_join(untarg_feats) %>%
-  select(feature, depth, estimate, p.value, mzmed, rtmed, compound_name) %>%
-  filter(p.value<0.05) %>%
-  # as.data.frame() %>%
-  filter(depth=="Deep") %>% 
-  print(n=Inf)
-reg_coefs %>%
-  filter(str_detect(term, "sla_corr")) %>%
-  left_join(untarg_feats) %>%
-  filter(compound_name=="Homarine") %>%
-  # filter(feature=="MS_POS_1395") %>%
-  filter(depth=="DCM") %>%
-  select(compound_name, depth, data, estimate) %>%
-  unnest(data) %>%
-  group_by(sla_corr, station) %>%
-  ggplot(aes(x=sla_corr, y=norm_area*100)) +
-  geom_smooth(method="lm", formula="y~x") +
-  geom_boxplot(aes(group=sla_corr)) +
-  scale_y_continuous(limits = c(0, NA))
-```
-
-In the 15 meter surface samples, only seven compounds had significant changes in relative peak area across the eddy transect, two of which were known: 5-oxoproline, which decreased from 0.2-0.3% of the total peak area at the center of the anticyclone to 0.05-0.1% in the cyclone; and a combined peak of sarcosine and beta-alanine which increased from 0.01% to 0.02-0.03%.
-The largest shift significant at an α = 0.05 level was an unknown mass feature with *m/z* = 189.12338 and a retention time around 8.6 minutes, the relative contribution of which increased from approximately 0.2-0.3% in the cyclone to ~1.5% in the anticyclone.
-
-At the DCM, twenty-two compounds changed significantly in relative peak area across the transect.
-Seven of these were known metabolites, of which trigonelline and homarine were most enriched in the cyclone while arsenobetaine was the only metabolite with a larger fraction of the total peak area in the anticyclone.
-Homarine showed a very large and highly significant shift, representing about 2.5% of the total peak area in the anticyclone but 7-8% in the cyclone center.
-Trigonelline (N-methyl niacin), an isomer structurally very similar to homarine but biologically distinct, had a surprisingly strong correlation with homarine (Pearson's $r=.855$).
-Its peak areas were consistently around one third of homarine's but still large enough to have the second-largest shift in relative peak area of the significantly different compounds.
-Similarly, arsenobetaine represented about 0.1% of the peak area in the cyclone and 0.5-0.6% of the total peak area in the anticyclone.
-The most significantly different compounds at the DCM in each direction, however, were both unknowns.
-The lowest p-value ($6.4*10^{-6}$ after FDR correction) in the DCM data was a mass feature with an *m/z* of 173.09211, a retention time also around 8.6 minutes, and enrichment in the anticyclone with a putative chemical formula of [M+H] = C$_7$H$_{13}$N$_2$O$_3$, possibly glycylproline or prolylglycine.
-The next-lowest (p = $2.5*10^{-4}$ after FDR correction) was enriched in the cyclone and had an *m/z* of 275.0712 and a retention time of 11.4 minutes with a putative formula of C$_{18}$H$_{11}$O$_3$.
-
-Finally, in the 175 meter samples we detected 43 mass features with a significant association with SLA.
-Notably, all detected nucleobases (guanine, adenine, and cytosine) were positively associated with SLA (higher values in the anticyclone than the cyclone), though the strongest associations were found in O-acetylcarnitine, betonicine, and tyrosine.
-Both O-acetylcarnitine and betonicine effectively quintupled their contribution to total peak area, shifting from approximately 0.1% to ~0.5% over the transect.
-Acetylcholine was the only known compound more relatively abundant in the cyclone than in the anticyclone along with one unknown of *m/z* 131.0340.
-The unknown at *m/z* 173.0921 with the strongest trend at the DCM again had the strongest trend in the 175 meter samples, here with an FDR-corrected p-value of $2.0 × 10^{-7}$.
-
-```{r homarine/trigonelline correlation, eval=FALSE}
-reg_coefs %>%
-  filter(str_detect(term, "sla_corr")) %>%
-  left_join(untarg_feats) %>%
-  filter(compound_name%in%c("Homarine", "Trigonelline")) %>%
-  filter(depth=="DCM") %>%
-  select(compound_name, depth, data, estimate) %>%
-  unnest(data) %>%
-  select(compound_name, filename, norm_area) %>%
-  pivot_wider(names_from=compound_name, values_from=norm_area) %>%
-  column_to_rownames("filename") %>%
-  cor()
-```
-
-## High vertical resolution sampling near eddy center DCM reveals distinct metabolome between cyclone and anticyclone
-
-```{r MC NMDS and PERMANOVA calculations}
-set.seed(123)
-multivar_analysis_MC <- untarg_peaks %>%
-  left_join(filled_file_metadata %>% distinct(ms_run, station, filename, depth), 
-          by="filename") %>% filter(ms_run=="MC") %>% 
-  mutate(depth=factor(depth, levels=c("DCMless20m", "DCMless10m", "DCM", 
-                                      "DCMplus10m", "DCMplus20m"))) %>%
-  nest(data=everything()) %>%
-  mutate(wide_peaks=map(data, function(metab_df){
-    metab_df %>%
-      select(feature, filename, norm_area) %>%
-      group_by(feature) %>%
-      filter(sum(norm_area==0)<n()*0.9) %>%
-      mutate(norm_area=scale(norm_area)[,1]) %>%
-      pivot_wider(names_from = "feature", values_from = norm_area) %>%
-      column_to_rownames("filename") %>%
-      data.matrix()
-  })) %>%
-  mutate(nmds_output=map(wide_peaks, function(metab_mat){
-    env_vec <- pull(filled_file_metadata, sla_corr, filename)[rownames(metab_mat)]
-    vegan::metaMDS(metab_mat, k = 2, distance = "manhattan", trace = 0, 
-                   autotransform = FALSE, noshare = FALSE, wascores = FALSE) %>%
-    vegan::MDSrotate(env_vec)
-  })) %>%
-  mutate(nmds_stress=map_dbl(nmds_output, pluck, "stress")) %>%
-  mutate(perm_output=map(wide_peaks, function(metab_mat){
-    adon_expl <- data.frame(filename=rownames(metab_mat)) %>%
-      left_join(filled_file_metadata %>% select(filename, sla_class, depth), by="filename")
-    vegan::adonis2(
-      metab_mat~sla_class+depth, data = adon_expl, by = "margin", 
-      permutations = 999, method = "manhattan")
-  })) %>%
-  mutate(perm_r2_sla=map_dbl(perm_output, pluck, "R2", 1)) %>%
-  mutate(perm_r2_depth=map_dbl(perm_output, pluck, "R2", 2)) %>%
-  mutate(perm_pval=map_dbl(perm_output, pluck, "Pr(>F)", 1)) %>%
-  mutate(pca_output=map(wide_peaks, prcomp)) %>%
-  mutate(nmds_stress=paste("NMDS stress =", round(nmds_stress, digits = 3)))
-  # mutate(perm_r2=round(perm_r2, digits = 2)) %>%
-  # mutate(perm_pval=round(perm_pval, digits = 4)) %>%
-  # mutate(perm_label=paste0("R2 = ", perm_r2, "\np = ", perm_pval))
-```
-
-We further investigated the metabolomic response to SLA at the DCM by collecting samples taken at high-resolution depth intervals around the DCM at the center of both eddy poles.
-Here, we found that this ~40 meter depth range and the transition between the eddies explained similar amounts of variation in the data (R$^2_{depth}=0.253$, R$^2_{SLA}=0.238$) and were both highly significant factors (permutational p-values << 0.001, Figure 4).
-Additionally, both sample depth and SLA were correlated with the first principal component of the metabolite matrix ($r_{depth}=.766$, $r_{SLA}=.684$, fraction of variance explained by PC1 = 30.8%).
-The cyclonic samples in particular show a much larger spread than the anticyclonic ones, indicating larger sample variability in the cyclone DCM relative to the anticyclone.
-A depth gradient is visible along NMDS 2, with the deepest samples generally at the top right of the plot and the shallowest ones closer to the bottom (Figure 4).
-
-```{r figure MC NMDS plot}
-MC_nmds_gp <- multivar_analysis_MC$nmds_output %>%
-  pluck(1, "points") %>% 
-  as.data.frame() %>%
-  rownames_to_column("filename") %>%
-  left_join(filled_file_metadata %>% distinct(filename, sla_class, depth, tripl)) %>%
-  mutate(depth=factor(depth, levels=c("DCMless20m", "DCMless10m", "DCM", 
-                                      "DCMplus10m", "DCMplus20m"),
-                      labels=c("-20m", "-10m", "DCM", "+10m", "+20m"))) %>%
-  mutate(fill_col=case_when(
-    depth=="-20m" & sla_class=="Cyclone"~"#c4c4e0",
-    depth=="-10m" & sla_class=="Cyclone"~"#8989c1",
-    depth=="DCM" & sla_class=="Cyclone"~"#6161ad",
-    depth=="+10m" & sla_class=="Cyclone"~"#343489",
-    depth=="+20m" & sla_class=="Cyclone"~"#23235b",
-    depth=="-20m" & sla_class=="Anticyclone"~"#dabdbd",
-    depth=="-10m" & sla_class=="Anticyclone"~"#b57c7c",
-    depth=="DCM" & sla_class=="Anticyclone"~"#9c5050",
-    depth=="+10m" & sla_class=="Anticyclone"~"#762020",
-    depth=="+20m" & sla_class=="Anticyclone"~"#4f1616"
-  )) %>%
-  ggplot() +
-  geom_vline(xintercept = 0) +
-  geom_hline(yintercept = 0) +
-  geom_point(aes(x=MDS1, y=MDS2, fill=fill_col, shape=sla_class, color=depth), size=6) +
-  scale_fill_identity() +
-  scale_color_manual(values = rep("white", 5)) +
-  scale_shape_manual(breaks=c("Cyclone", "Anticyclone"), values=c(25, 24)) +
-  annotate("label", x=-200, y=100, label="Cyclone", color=scales::muted("blue"),
-           label.r = unit(0, "in"), size=6) +
-  annotate("label", x=100, y=120, label="Anticyclone", color=scales::muted("red"),
-           label.r = unit(0, "in"), size=6) +
-  guides(alpha="none", fill="none", shape="none", color=guide_legend(
-    title = "Depth:    ", title.position = "left", title.hjust = 0.5, direction = "horizontal",
-    keywidth=0.2, keyheight=0.1,
-    override.aes = list(fill=c("grey90", "grey80", "grey60", "grey40", "grey20"), 
-                        pch=21, color="black")
-  )) +
-  labs(x="NMDS axis 1", y="NMDS axis 2") +
-  theme_bw() +
-  theme(legend.position = "top", legend.text = element_text(margin = margin(r = 15, l=2, unit = "pt"))) +
-  annotate("label", x=-Inf, y=Inf, hjust=0, vjust=1, label.r = unit(0, "in"),
-           label=expression(atop(paste("R"[SLA]^2, "=.238"), paste("R"[depth]^2, "=.253")))) +
-  annotate("label", x=-Inf, y=-Inf, label=multivar_analysis_MC$nmds_stress, 
-           hjust=0, vjust=-2, label.r = unit(0, "in"))
-
-ggsave("MC_nmds_gp.png", plot = MC_nmds_gp, device = "png", 
-       path = "figures", width = 6.875, height = 4.5, units = "in", dpi = 300)
-```
-
-![](figures/MC_nmds_gp.png)
-
-*Figure 4: NMDS plot of high-resolution depth sampling around the deep chlorophyll maximum (DCM, ~115 meters) at the two eddy centers during MESO-SCOPE.
-Red upward-pointed triangles are from the anticyclone and blue downward-pointed ones are from the cyclone.
-Shading intensity reflects the depth above or below the DCM.
-PERMANOVA estimates of the variance explained by depth and sea level anomaly (SLA) are noted in the upper left corner and the NMDS stress value is reported in the bottom left.*
-
-```{r PCA checks for MC axis regressions, eval=FALSE}
-multivar_analysis_MC$pca_output[[1]]$x %>%
-  as.data.frame() %>%
-  rownames_to_column("filename") %>%
-  left_join(filled_file_metadata) %>%
-  select(abs_depth, sla_corr, PC1:PC3) %>%
-  # cor() %>% round(3)
-  ggplot(aes(x=abs_depth, y=PC1, color=sla_corr)) +
-  geom_point()
-
-head(multivar_analysis_MC$pca_output[[1]]$sdev^2/sum(multivar_analysis_MC$pca_output[[1]]$sdev^2)*100)
-```
-
-To characterize the observed differences in multivariate space we used k-means clustering as an unsupervised way to identify dominant trends within the data (Figure 5).
-We found that a majority of the metabolites (Clusters 1 and 2, 63% of the total) fell into clusters with larger peak areas in the cyclone, as expected, while also detecting 32 metabolites that clustered such that the mean metabolite was enriched in the anticyclone (Cluster 4, Figure 5).
-Cluster 2 had a distinct decrease in relative peak area with depth, while the other clusters had much less clear depth trends.
-Cluster 1 also showed a DCM maximum for the samples from the anticyclone, correlating well with flow cytometry counts of picoeukaryotes (Pearson's $r=.851$) that was not present in the cyclone ($r=.013$).
-
-```{r k-means clusterplot}
-set.seed(123)
-kclust <- untarg_peaks %>%
-  left_join(filled_file_metadata, by="filename") %>%
-  filter(ms_run=="MC") %>%
-  select(feature, filename, norm_area) %>%
-  group_by(feature) %>%
-  filter(sum(norm_area==0)<n()*0.9) %>%
-  mutate(norm_area=scale(norm_area)[,1]) %>%
-  pivot_wider(names_from = "filename", values_from = norm_area) %>%
-  column_to_rownames("feature") %>%
-  data.matrix() %>%
-  kmeans(centers = 4)
-
-kclust_nms <- names(kclust$cluster)
-kclust$cluster <- order(kclust$size, decreasing = TRUE)[kclust$cluster]
-names(kclust$cluster) <- kclust_nms
-kclust$size <- sort(kclust$size, decreasing = TRUE)
-
-kclust_df <- kclust %>%
-  pluck("cluster") %>%
-  data.frame(cluster=.) %>%
-  rownames_to_column("feature") %>%
-  left_join(untarg_peaks, multiple = "all", by="feature") %>%
-  left_join(filled_file_metadata, by="filename") %>%
-  filter(ms_run=="MC") %>%
-  arrange(desc(abs_depth)) %>%
-  mutate(depth=fct_inorder(depth)) %>%
-  group_by(feature) %>%
-  mutate(norm_area=scale(norm_area)[,1]) %>%
-  group_by(cluster, depth, sla_class) %>%
-  summarise(mean_rank=mean(norm_area), iqr_value=sd(norm_area), .groups = "drop")
-
-lab_df <- tribble(
-  ~xlab, ~ylab, ~sla_class, ~clusterlab,
-  0, 5.5, "Anticyclone", "Cluster #1: n=73",
-  0.8, 1, "Cyclone", "Cluster #1: n=73"
-)
-
-library(ggh4x)
-strip_theme <- strip_themed(background_x = elem_list_rect(fill= c(
-  "#414487FF", "#7AD151FF", "#22A884FF", "#2A788EFF")))
-kclust_gp <- kclust_df %>%
-  left_join(data.frame(size=kclust$size) %>% mutate(cluster=row_number()), by="cluster") %>%
-  mutate(clusterlab=paste0("Cluster #", cluster, ": n=", size)) %>%
-  mutate(depth=factor(depth, labels=rev(c("-20m", "-10m", "DCM", "+10m", "+20m")),
-                      levels=rev(c("DCMless20m", "DCMless10m", "DCM", "DCMplus10m", "DCMplus20m")))) %>%
-  ggplot(aes(y=depth, color=sla_class)) +
-  geom_vline(xintercept = 0) +
-  geom_path(aes(x=mean_rank, group=sla_class), linewidth=1) +
-  geom_point(aes(x=mean_rank), size=3) +
-  geom_label(aes(x=xlab, y=ylab, label=sla_class, color=sla_class), 
-             data=lab_df, label.r = unit(0, "in")) +
-  ggh4x::facet_wrap2(~clusterlab, nrow = 1, strip = strip_theme) +
-  scale_color_manual(values = c(scales::muted("blue"), scales::muted("red"), "grey50"), 
-                     breaks = c("Cyclone", "Anticyclone", "Neither"),
-                     guide="none") +
-  labs(x="Mean z-scaled metabolite area", y="Sampling depth", color=NULL) +
-  coord_cartesian(ylim = c(1, 5.5)) +
-  theme_bw() +
-  theme(plot.background = element_rect(fill = "white"),
-        strip.background = element_rect(fill = "white"),
-        strip.text = element_text(face = "bold", color="white"))
-```
-
-```{r pairwise_tests and fold_change_df}
-MC_simpledata <- untarg_peaks %>%
-  left_join(filled_file_metadata %>% distinct(filename, ms_run, station), by="filename") %>%
-  filter(ms_run=="MC") %>%
-  select(feature, filename, norm_area, station) %>%
-  group_by(feature) %>%
-  filter(sum(norm_area==0)<n()*0.9) %>%
-  group_by(filename) %>%
-  # mutate(norm_area=norm_area/sum(norm_area)) %>%
-  ungroup()
-pairwise_tests <- MC_simpledata %>%
-  nest(data=c(filename, norm_area, station)) %>%
-  mutate(stattest=map(data, ~broom::tidy(wilcox.test(.x$norm_area~.x$station)))) %>%
-  unnest(stattest) %>%
-  select(feature, statistic, p.value) %>%
-  mutate(p.value=p.adjust(p.value, method="fdr"))
-
-fold_change_df <- MC_simpledata %>%
-  group_by(feature, station) %>%
-  summarise(mean_area=mean(norm_area, na.rm=TRUE)) %>%
-  pivot_wider(names_from=station, values_from=mean_area) %>%
-  mutate(fold_change=L1/L2) %>%
-  select(feature, fold_change)
-```
-
-```{r eddy center volcano plot}
-label_df <- pairwise_tests %>%
-  left_join(untarg_feats) %>%
-  filter(compound_name%in%c("Isethionic acid", "Trigonelline", "Arsenobetaine",
-                            "Homarine", "Dimethylsulfonioacetate", "Dimethylsulfoniopropionate")) %>%
-  left_join(fold_change_df, by = join_by(feature)) %>%
-  mutate(xspot = log2(fold_change), yspot=-log10(p.value))
-
-volcanoplot_gp <- pairwise_tests %>%
-  left_join(fold_change_df, by = join_by(feature)) %>%
-  left_join(rownames_to_column(data.frame(cluster=kclust$cluster), "feature"), 
-            by = join_by(feature)) %>%
-  left_join(untarg_feats %>% select(feature, compound_name), by = join_by(feature)) %>%
-  mutate(known=ifelse(is.na(compound_name), "Unknown", "Known")) %>%
-  mutate(compound_name=ifelse(is.na(compound_name), feature, compound_name)) %>%
-  ggplot(aes(x=log2(fold_change), y=-log10(p.value), color=factor(cluster), 
-             shape=known, label=compound_name)) +
-  geom_vline(xintercept = 0) +
-  geom_hline(yintercept = 0) +
-  geom_hline(yintercept = 1.3, lty=2) +
-  annotate("segment", x=label_df$xspot[5], y=6.8, xend=label_df$xspot[5], yend=label_df$yspot[5], 
-           # arrow=arrow(type="closed", angle = 25, length = unit(0.05, "npc")), 
-           lwd=0.7, color="grey50") +
-  annotate("label", x=-2, y=6.8, label="Arsenobetaine", hjust=0, label.r=unit(0, "in")) +
-  annotate("segment", x=1.5, y=6.8, xend=label_df$xspot[3], yend=label_df$yspot[3], 
-           lwd=0.7, color="grey50") +
-  annotate("label", x=1.5, y=6.8, label="Homarine", hjust=1, label.r=unit(0, "in")) +
-  annotate("segment", x=2, y=6.8, xend=label_df$xspot[2], yend=label_df$yspot[2], 
-           lwd=0.7, color="grey50") +
-  annotate("label", x=2, y=6.8, label="DMSP", hjust=0, label.r=unit(0, "in")) +
-  annotate("segment", x=Inf, y=label_df$yspot[1], xend=label_df$xspot[1], yend=label_df$yspot[1], 
-           lwd=0.7, color="grey50") +
-  annotate("label", x=Inf, y=label_df$yspot[1], label="DMS-Ac", hjust=1, vjust=0.3, label.r=unit(0, "in")) +
-  annotate("segment", x=3.5, y=5, xend=1.9, yend=5, 
-           lwd=0.7, color="grey50") +
-  annotate("segment", x=1.9, y=5, xend=label_df$xspot[4], yend=label_df$yspot[4], 
-           lwd=0.7, color="grey50") +
-  annotate("label", x=Inf, y=5, label="Trigonelline", hjust=1, label.r=unit(0, "in")) +
-  annotate("segment", x=3.6, y=3, xend=2.2, yend=3, 
-           lwd=0.7, color="grey50") +
-  annotate("segment", x=2.2, y=3, xend=label_df$xspot[6], yend=label_df$yspot[6], 
-           lwd=0.7, color="grey50") +
-  annotate("label", x=Inf, y=3, label="Isethionate", hjust=1, label.r=unit(0, "in")) +
-  geom_point(size=2, stroke=1) +
-  scale_x_continuous(breaks=c(-4:4)) +
-  scale_y_continuous(breaks = 0:6, labels = paste0("10<sup>-", 0:6, "</sup>"), expand = expansion(c(0.01, 0.05))) +
-  scale_shape_manual(values=c(4, 16), breaks = c("Unknown", "Known")) +
-  scale_color_manual(
-    values = c("#414487FF", "#7AD151FF", "#22A884FF", "#2A788EFF"),
-    breaks = 1:4
-  ) +
-  theme_bw() +
-  labs(x="Log-2 fold-change cyclone enrichment", y="Mann-Whitney p-value\n(FDR-corrected)",
-       shape="Metabolite status", color="k-means\ncluster") +
-  guides(color="none", shape=guide_legend(order=1)) +
-  theme(axis.text.y = ggtext::element_markdown(),
-        plot.background = element_rect(fill = "white"),
-        legend.position = c(0, 1), legend.justification = c(0, 1), 
-        legend.background = element_rect(fill="#FFFFFF88", color="black"),
-        legend.spacing.y=unit(0, "in"),
-        legend.key.size = unit(0, "cm"))
-```
-
-```{r kclust volcano plot combine}
-kclust_volcano_gp <- egg::ggarrange(
-  kclust_gp, volcanoplot_gp, 
-  ncol=1, heights = c(1.5, 2), draw = FALSE)
-ggsave("kclust_volcano_gp.png", plot = kclust_volcano_gp, device = "png", 
-       path = "figures", width = 6.875, height = 5, units = "in", dpi = 300,
-       bg="white")
-```
-
-![](figures/kclust_volcano_gp.png)
-
-*Figure 5: Distribution of metabolites in the high-resolution depth samples from the centers of each MESO-SCOPE eddy.
-The upper row of plots shows k-means clusters where points denote the average z-scored peak area for both known and unknown metabolites across the samples and are colored by the eddy from which they were taken.
-Clusters have been ordered by number of metabolites in each group and the total is denoted in the panel titles.
-Both depth trends (mostly a net decrease in metabolites with depth) and eddy effects (cyclonic enrichment in clusters 1 and 2, anticyclone enrichment in cluster 4) are observable.
-The lower plot shows the individual known and unknown metabolites where points correspond to the FDR-corrected p-value estimated by the nonparametric Mann-Whitney U test and the log$_2$ fold-change calculated with the average peak area in the cyclone divided by the average peak area in the anticyclone.
-Colors have been assigned using the k-means clusters and shapes have been assigned based on the status of the mass feature as either a known metabolite that was matched to an authentic standard or an unidentified metabolite.
-The dashed line across the figure represents the 0.05 level of significance as a visual cue for metabolites above which the differences between the eddies are unlikely to be due to chance.
-DMSP = dimethylsulfoniopropionate, DMS-Ac = dimethylsulfonioacetate*
-
-```{r cluster cor w metadata, eval=FALSE}
-filled_file_metadata %>% 
-  filter(ms_run=="MC") %>% 
-  distinct(depth, euk_abund, sla_class) %>% 
-  left_join(kclust_df) %>%
-  filter(cluster==1) %>%
-  filter(sla_class=="Cyclone") %>%
-  # with(cor(mean_rank, euk_abund))
-  ggplot() +
-  geom_point(aes(x=euk_abund, y=mean_rank, color=sla_class))
-```
-
-The majority of the individual metabolites in these high-resolution depth profiles differed significantly between the two eddies (121/228, α = 0.05).
-Of those, 15 were enriched in the anticyclone and 106 were more abundant in the cyclone.
-As expected, all compounds enriched in the anticyclone were part of Cluster 4 and all those enriched in the cyclone belonged to either Cluster 1 or 2 (Figure 5).
-
-Many of the known metabolites enriched in the cyclone function as osmolytes in the cell and might be enriched due to the increased eukaryotic phytoplankton biomass.
-However, some metabolites such as isethionate, the reduced sulfur osmolytes dimethylsulfoniopropionate (DMSP) and dimethylsulfonioacetate (DMS-Ac), and the isomers homarine and trigonelline, were enriched in excess of biomass (Figure 5).
-A few metabolites more abundant in the anticyclone were also given putative identifications based on RT and *m/z* matching with internal standards run at a later time on the mass spectrometer.
-Of these, the putative arsenobetaine was the most significantly different among these with a peak area at the anticyclone DCM nearly quadruple that of the cyclone (Figure 5).
-Of note, the 173.0921 *m/z* mass feature noted above as strongly enriched in the anticyclone was among the most significantly different between the two eddies, along with two isomers at an *m/z* of 170.1176 and retention times around 8-9 minutes that increased by a factor of 1.5 in the anticyclone (putative formula C$_7$H$_{14}$N$_4$O).
-
-```{r putative 170.11756 formula, eval=FALSE}
-# Two possible formula: C9H16NO2 and C7H14N4O
-library(arrow)
-dset <- open_dataset("G:/My Drive/AllMeso/tmzMLs/MC_all_file_arr_pqds")
-
-# Too low-intensity for reliable isotope info
-dset %>%
-  filter(mz%between%pmppm(170.11757, 5)) %>%
-  filter(rt%between%c(7, 10)) %>%
-  collect() %>%
-  left_join(filled_file_metadata) %>%
-  ggplot() +
-  geom_line(aes(x=rt, y=int, group=filename, color=station))
-rtr <- c(8.7, 9.2)
-gly_data <- dset %>%
-  filter(rt%between%rtr) %>%
-  collect()
-mzr <- pmppm(170.11756, 10)
-reduce(list(
-  gly_data[mz%between%mzr] %>%
-    group_by(filename) %>%
-    summarise(M=sum(int)),
-  gly_data[mz%between%(mzr+1.003355)] %>%
-    group_by(filename) %>%
-    summarise(C13=sum(int))
-), merge, all.x=TRUE) %>%
-  pivot_longer(C13) %>%
-  mutate(prop=value/M) %>%
-  print(n=Inf)
-
-# But there are useful MS2 fragments
-library(RaMS)
-mz_group <- function(mz_vals, ppm){
-  group_vec <- numeric(length(mz_vals))
-  group_num <- 1L
-  init_vec <- mz_vals
-  names(init_vec) <- seq_along(init_vec)
-  while(length(init_vec)>0){
-    mz_i <- init_vec[1]
-    err <- mz_i*ppm/1000000
-    mz_idxs <- init_vec>mz_i-err & init_vec<mz_i+err
-    group_vec[as.numeric(names(mz_idxs)[mz_idxs])] <- group_num
-    init_vec <- init_vec[!mz_idxs]
-    group_num <- group_num+1L
-  }
-  group_vec
-}
-msmsdata_MC <- "../mzMLs/pos/MSMS" %>%
-  list.files(, full.names = TRUE, pattern = "180205") %>%
-  grabMSdata()
-msmsdata_MC$MS2[premz%between%pmppm(170.11757, 10)] %>%
-  mutate(mz_group=mz_group(fragmz, 5)) %>%
-  group_by(mz_group, filename) %>%
-  summarize(fragmz=mean(fragmz), int=mean(int)) %>%
-  filter(n()>1) %>%
-  group_by(mz_group) %>%
-  summarise(fragmz=mean(fragmz), int=mean(int)) %>%
-  filter(fragmz<170) %>%
-  arrange(desc(int))
-# Top two frags have formulae of 125.1075 = C7H13N2
-# and 89.07148 = C3H9N2O, both requiring 2 N atoms
-```
-
-## Hawaiian Eddy Experiment data reveals a different community and response to mesoscale eddies
-
-Given the strong signals detected in the samples from the MESO-SCOPE cruise both across the eddy transect as well as in the high-resolution DCM sampling, we expected to find similar results in an analogous dataset.
-Samples were collected during a 2018 cruise on the R/V *Falkor* that again targeted both a cyclonic and an anticyclonic eddy in the North Pacific Subtropical Gyre near Station ALOHA as part of the Hawaiian Eddy Experiment (HEE).
-
-```{r falkor nmds and permanova}
-set.seed(123)
-scaled_fk_data <- untarg_peaks %>%
-  # filter(filename!="190715_Smp_FK180310S62C1-DCM_B.mzML") %>%
-  left_join(filled_file_metadata %>% distinct(ms_run, station, filename, depth), 
-          by="filename") %>%
-  filter(ms_run=="FK") %>%
-  mutate(depth=factor(depth, levels=c("Surface", "DCM"))) %>%
-  select(feature, filename, norm_area) %>%
-  group_by(feature) %>%
-  filter(sum(norm_area==0)<n()*0.9)
-metab_mat_fk <- scaled_fk_data %>%
-  mutate(norm_area=scale(norm_area)[,1]) %>%
-  # mutate(norm_area=rank(norm_area)) %>%
-  pivot_wider(names_from = "feature", values_from = norm_area) %>%
-  column_to_rownames("filename") %>%
-  data.matrix()
-env_vec <- pull(filled_file_metadata, sla_corr, filename)[rownames(metab_mat_fk)]
-nmds_fk <- metab_mat_fk %>%
-  vegan::metaMDS(k = 2, distance = "manhattan", trace = 0, autotransform = FALSE,
-                 noshare = FALSE, wascores = FALSE) %>%
-  vegan::MDSrotate(env_vec)
-adon_expl <- data.frame(filename=rownames(metab_mat_fk)) %>%
-  left_join(filled_file_metadata %>% select(filename, sla_class, depth), by="filename")
-adon_fk <- vegan::adonis2(
-  metab_mat_fk~sla_class+depth, data = adon_expl, by = "margin", 
-  permutations = 999, method = "manhattan")
-```
-
-```{r figure falkor nmds}
-rot_mat <- matrix(c(sqrt(2)/2, sqrt(2)/2, -sqrt(2)/2, sqrt(2)/2), ncol=2)
-
-fk_nmdsplot <- nmds_fk$points%*%rot_mat %>%
-  as.data.frame() %>%
-  setNames(c("MDS1", "MDS2")) %>%
-  rownames_to_column("filename") %>%
-  left_join(filled_file_metadata %>% distinct(filename, sla_class, depth)) %>%
-  mutate(MDS1=MDS1*-1) %>%
-  mutate(MDS2=MDS2*-1) %>%
-  mutate(fill_col=case_when(
-    depth=="Surface" & sla_class=="Cyclone"~"25 m Cyclone",
-    depth=="DCM" & sla_class=="Cyclone"~"25 m Anticyclone",
-    depth=="Surface" & sla_class=="Anticyclone"~"DCM Cyclone",
-    depth=="DCM" & sla_class=="Anticyclone"~"DCM Anticyclone"
-  )) %>%
-  ggplot(aes(label=filename)) +
-  geom_vline(xintercept = 0) +
-  geom_hline(yintercept = 0) +
-  geom_point(aes(x=MDS1, y=MDS2, fill=fill_col, shape=sla_class), size=6, color="white") +
-  scale_fill_manual(breaks=c("25 m Cyclone", "25 m Anticyclone", "DCM Cyclone", "DCM Anticyclone"),
-                    values=c("#8989c1", "#343489", "#b57c7c", "#762020")) +
-  scale_shape_manual(breaks=c("Cyclone", "Anticyclone"), values=c(25, 24)) +
-  labs(color="Corr.\nSLA\n(cm)", shape="Depth", x="NMDS axis 1", y="NMDS axis 2") +
-  theme_bw() +
-  annotate("label", x=-Inf, y=-Inf, hjust=0, vjust=0, label.r = unit(0, "in"),
-           label=paste("Stress =", round(nmds_fk$stress, digits = 3))) +
-  annotate("label", x=-Inf, y=Inf, hjust=0, vjust=1, label.r = unit(0, "in"),
-           label=expression(atop(paste("R"[sla]^2, "=.090, p=0.017"),
-                                 paste("R"[depth]^2, "=.251, p<0.001")))) +
-  guides(fill=guide_legend(title = NULL, override.aes = list(
-    fill=c("#8989c1", "#b57c7c", "#343489", "#762020"), color="white",
-    shape=c(25, 24, 25, 24)
-  )), shape="none") +
-  theme(legend.key.spacing.y = unit(6, "pt")) +
-  coord_fixed()
-
-ggsave("fk_nmdsplot.png", plot = fk_nmdsplot, device = "png",
-       path = "figures", width = 6.875, height = 4, units = "in", dpi = 300,
-       bg="white")
-```
-
-The HEE data was characterized by high inter-replicate variability relative to the samples from the MESO-SCOPE cruise, with 25 meter samples in particular highly variable in multivariate space (Figure 6).
-Despite this, we still saw the importance of sea level anomaly as a significant explanatory factor in the dataset (R$^2_{SLA} = 0.090$, p-value = 0.017) as well as the larger depth differences between the surface (25 meters) and DCM (110 - 120 meters) (R$^2_{depth} = 0.251$, p-value << 0.001).
-
-![](figures/fk_nmdsplot.png)
-
-*Figure 6: Non-metric multidimensional scaling (NMDS) plot from the Hawaiian Eddy Experiment cruise data, in which points correspond to individual samples and have been colored and shaped by their source eddy status and shaded by the depth from which they were collected.
-The NMDS stress value has been reported in the bottom left corner, while PERMANOVA R$^2$ and p-values are reported in the top left.
-Samples from 25 meters deep are visibly distinct from the deep chlorophyll maximum (DCM) samples and an SLA signal is visible in the 25 meter samples only.*
-
-```{r top20compounds stacked barplot, eval=FALSE}
-scaled_fk_data %>%
-  ungroup() %>%
-  arrange(desc(norm_area)) %>%
-  filter(feature%in%head(unique(feature), 20)) %>%
-  left_join(filled_file_metadata) %>%
-  left_join(untarg_feats) %>%
-  mutate(feature=ifelse(is.na(compound_name), feature, compound_name)) %>%
-  mutate(feature=fct_inorder(feature)) %>%
-  arrange(depth, filename) %>%
-  mutate(filename=str_remove(filename, "190715_Smp_FK180310")) %>%
-  mutate(filename=str_remove(filename, "-25m_|-DCM_")) %>%
-  mutate(filename=fct_inorder(filename)) %>%
-  ggplot() +
-  geom_col(aes(x=filename, y=norm_area, fill=feature)) +
-  theme(axis.text.x = element_text(angle=90, hjust = 1, vjust=0.5)) +
-  facet_wrap(~depth, scales="free_y", ncol=1)
-```
-
-When analyzed separately as distinct depths, the 25 meter samples had higher variances explained by eddy (R$^2$=.26) and a lower p-value (0.013), while the DCM samples were no longer likely to be distinct between the two eddies (p-value = 0.171) (Figure 6, Supplemental Figure 4).
-
-```{r falkor concentrations comparison to MESO-SCOPE, eval=FALSE}
-fk_targ <- targ_peaks %>% 
-  left_join(filled_file_metadata) %>%
-  filter(ms_run%in%c("FK", "MT")) %>%
-  filter(samp_type=="Smp") %>%
-  select(compound_name, filename, nM, depth, ms_run)
-fk_targ %>%
-  group_by(depth, compound_name, ms_run) %>%
-  summarise(nM=mean(nM)) %>%
-  arrange(desc(nM)) %>%
-  pivot_wider(names_from = c(ms_run, depth), values_from = nM) %>%
-  print(n=Inf)
-cmpd_comp_gp <- fk_targ %>%
-  filter(depth!="Deep") %>%
-  ggplot() +
-  geom_boxplot(aes(x=ms_run, y=nM)) +
-  facet_grid(compound_name~depth, scales="free_y")
-ggsave("cmpd_comp_gp.pdf", cmpd_comp_gp, device = "pdf", width = 6, height = 50, limitsize = FALSE)
-# Visual assessment of boxplots from the above output:
-# Compounds more abundant in Falkor: citrulline, aspartic acid, ornithine, threonine, tyrosine, malic acid
-# Compounds more abundant in MESO: 4-aminobutyric, adenine, gonyol, 4-hydroxyisoleucine, alanine, leucine, nicotinic acid
-```
-
-We were also able to use the HEE dataset to test whether the compounds detected as significantly different in the MESO-SCOPE dataset were also different in this eddy pair.
-We expected to find abundant TMAO and more hydroxyisoleucine at the surface than the DCM in addition to most compounds enriched in the cyclone DCM with biomass, especially the six compounds mentioned above in the results of Figure 5 (trigonelline, homarine, DMS-Ac, DMSP, taurine, and isethionic acid).
-
-```{r abundant TMAO and hydroxyisoleucine depth trends, eval=FALSE}
-fk_targ <- targ_peaks %>% 
-  left_join(filled_file_metadata) %>%
-  filter(ms_run%in%c("FK")) %>%
-  filter(samp_type=="Smp") %>%
-  select(compound_name, filename, nM, depth, ms_run)
-
-fk_targ %>%
-  filter(compound_name=="Hydroxyisoleucine") %>%
-  left_join(filled_file_metadata) %>%
-  aov(formula=nM~depth) %>% summary()
-  # ggplot(aes(x=depth, y=nM, color=station)) +
-  # geom_jitter(width=0.2)
-  # geom_boxplot()
-
-fk_targ %>%
-  filter(compound_name%in%c("Homarine", "Trigonelline", "Dimethylsulfonioacetate", "Dimethylsulfoniopropionate", "Taurine", "Isethionic acid", "Arsenobetaine")) %>%
-  left_join(filled_file_metadata) %>%
-  group_by(filename) %>%
-  mutate(nM=nM/sum(nM)) %>%
-  ungroup() %>%
-  filter(depth=="DCM") %>%
-  nest(data=-compound_name) %>%
-  mutate(aovout=map(data, ~broom::tidy(aov(.x$nM~.x$sla_class)))) %>%
-  unnest(aovout) %>% filter(term==".x$sla_class") %>%
-  mutate(p.value=p.adjust(p.value, method="fdr"))
-  ggplot(aes(x=sla_class, y=nM)) +
-  # geom_boxplot() +
-  geom_point() +
-  facet_wrap(~compound_name, scales="free_y")
-
-untarg_feats %>%
-  filter(mzmed%between%pmppm(170.11756, 5) | mzmed%between%pmppm(173.09211, 5)) %>%
-  filter(startsWith(feature, "FK")) %>% 
-  left_join(untarg_peaks) %>%
-  left_join(filled_file_metadata) %>%
-  nest(data=-feature) %>%
-  mutate(aovout=map(data, ~broom::tidy(aov(.x$norm_area~.x$sla_class)))) %>%
-  unnest(aovout) %>% filter(term==".x$sla_class") %>%
-  mutate(p.value=p.adjust(p.value, method="fdr"))
-  select(feature, sla_class, norm_area) %>% 
-  pivot_wider(names_from = sla_class, values_from = norm_area, values_fn = mean)
-  ggplot(aes(x=sla_class, y=norm_area)) +
-  # geom_boxplot() +
-  geom_point() +
-  facet_wrap(~feature, scales="free_y")
-```
-
-As expected, TMAO was again one of the most abundant metabolites detected in the particulate matter with concentrations around 1.2 nM at 25 meters and 0.4 nM at the DCM, second only to the high levels of the amino acid glycine (2 nM at 25 meters, 0.9 nM DCM), though the differences between 25 meters and DCM were not significant for either compound.
-We also detected a significant difference between depths for hydroxyisoleucine (25 meter mean = 0.32 nM, DCM mean = 0.21 nM, t-test p-value=0.022 with n=12 samples at each depth).
-These results imply that the overall depth structure of the metabolome of the gyre is relatively fixed for the most abundant compounds.
-
-The eddy effects, on the other hand, were much less strong during this cruise.
-Of the six known metabolites with the strong enrichment in the cyclone at the DCM in the MESO-SCOPE cruise, only isethionic acid was significantly different in the HEE dataset (isethionic acid t-test p-value$_{FDR}$ = 0.043, other five p-values > 0.25).
-This difference in isethionic acid concentration was only found after normalizing each sample to the sum of all metabolites in the sample to control for biomass.
-Arsenobetaine was again found to be strongly enriched in the anticyclone at the DCM both when normalizing to biomass and when not, just as in MESO-SCOPE.
-Surprisingly, the 173.0921 *m/z* mass feature was also slightly enriched but this time in the *cyclone* DCM with peak areas approximately 1.4 times the values in the anticyclone (t-test p-value = 0.067, Mann-Whitney p-value = 0.045).
-The 170.1176 *m/z* mass feature was not detected in the HEE samples.
-
-# Discussion
-
-## Multivariate approaches reveal metabolome-wide shifts across pairs of eddies of opposite polarity
-
-We measured the metabolome of samples collected across two sets of adjacent eddies of opposite polarity to explore the effect of sea level anomaly (SLA) on metabolite composition and found that the altered biogeochemistry and microbial community composition between cyclonic and anticyclonic eddies explain a significant portion of the observed variations in their particulate metabolomes.
-
-In all of the datasets analyzed here, we detected a significant difference in the composition of the metabolome between the adjacent eddies.
-This effect was strongest in the samples taken during the Lagrangian stations at the center of each eddy in the 2017 MESO-SCOPE cruise, with nearly a quarter of the total variance explained by the eddy from which the samples were taken.
-In the samples taken along the eddy transect, the largest effect was detected in the deepest samples taken from 175 meters and from the DCM, with much less of a response at 25 meters.
-In the HEE samples, however, the opposite was true with a larger SLA response at 25 meters than that at the DCM.
-In each case, the differences introduced by SLA were smaller than those of depth when compared directly, with even the high-resolution sampling around the DCM finding slightly more variance explained by the 40 meter difference in sampling depth than the 40 centimeter difference in SLA.
-This result agrees well with prior research by @Heal2021 and @Kumler2023 showing the large effect of sampling depth on the metabolome.
-
-Although possible that the observed differences in metabolomes across adjacent eddies differing in polarity are due to latitudinal shifts or simply background variation in the gyre environment, the transect data in particular implies that this is unlikely to be the case.
-Samples taken when the absolute value of the SLA was less than 5 cm were more similar to each other than to those samples taken within the eddies despite larger differences in latitude.
-The transect sampled both outside of each eddy and between the two across a large spatial gradient (~4° latitude and 2° longitude), yet the non-eddy (absolute value of SLA < 5 cm) samples from all three locations grouped together, showed similar median metabolite concentrations, and were visually similar in the most abundant metabolite composition.
-These results imply that strong cyclonic and anticyclonic eddies represent endpoints in NPSG composition, in agreement with results from @Barone2019 which found the most extreme values in the Hawaii Ocean Time series data typically detected when eddies passed over Station ALOHA.
-We also detected the largest differences in the DCM metabolome at the exact center of the eddy, with Station 12 highly distinct from even the nearest stations, while no such stark difference was found in the anticyclone (Figures 1 and 2).
-This perhaps indicates that the exact center of the cyclone has a unique metabolome at the DCM relative to the rest of the transect, rather than existing as a smooth continuation.
-
-Shifts in total biomass were a large driver of metabolomic differences, with earlier work in the same location by @Barone2022 showing that particulate carbon, chlorophyll, and beam attenuation were all 20-80% higher in the DCM of the cyclone relative to the anticyclone.
-We see this reflected particularly well in the targeted metabolites measured here, with clear depth trends and a strong correlation between total metabolite concentration and particulate carbon values.
-However, even after controlling for biomass effects by normalizing each sample to the sum of signal measured within it we still found a significant SLA effect, likely due to shifts in the community composition and in particular the increased eukaryote presence in the cyclone DCM.
-
-```{r multimodel testing maybe, eval=FALSE}
-dist_mat <- untarg_peaks %>%
-  left_join(filled_file_metadata %>% distinct(ms_run, station, filename, depth), 
-            by="filename") %>% filter(ms_run=="MC") %>% 
-  mutate(depth=factor(depth, levels=c("DCMless20m", "DCMless10m", "DCM", 
-                                      "DCMplus10m", "DCMplus20m"))) %>%
-  arrange(feature, station, depth) %>%
-  mutate(filename=fct_inorder(filename)) %>%
-  group_by(feature) %>%
-  mutate(norm_area=rank(norm_area)) %>%
-  select(feature, filename, norm_area) %>%
-  pivot_wider(names_from=feature, values_from="norm_area") %>%
-  mutate(filename=str_remove_all(filename, "180205_Smp_|.mzML")) %>%
-  column_to_rownames("filename") %>%
-  dist(method = "manhattan") %>%
-  as.matrix()
-dist_mat %>%
-  as.data.frame() %>%
-  rownames_to_column("filename_a") %>%
-  pivot_longer(-filename_a, names_to="filename_b", values_to = "dist") %>%
-  mutate(filename_a=fct_inorder(filename_a)) %>%
-  mutate(filename_b=fct_inorder(filename_b)) %>%
-  filter(as.numeric(filename_b)>=as.numeric(filename_a)) %>%
-  ggplot() +
-  geom_raster(aes(x=filename_a, y=filename_b, fill=dist)) +
-  coord_fixed() +
-  scale_fill_gradient2(midpoint = 2200) +
-  theme_minimal() +
-  theme(axis.text.x = element_text(angle=90, hjust=1, vjust=0.5))
-
-expl_df <- filled_file_metadata %>% 
-  distinct(ms_run, station, filename, depth, samp_type) %>%
-  left_join(untarg_peaks, by="filename") %>%
-  filter(ms_run=="MC") %>%
-  filter(samp_type=="Smp") %>%
-  mutate(model1_groups=as.numeric(factor(depth))) %>%
-  mutate(model2_groups=case_when(
-    station=="L1" & depth %in% c("DCMless20", "DCMless10") ~ 1,
-    station=="L1" & depth=="DCM" | station=="L2" & depth=="DCMless20" ~ 2,
-    station=="L1" & depth=="DCMplus10" | station=="L2" & depth=="DCMless10" ~ 3,
-    station=="L1" & depth=="DCMplus20" | station=="L2" & depth=="DCM" ~ 4,
-    TRUE ~ 5
-  )) %>%
-  distinct(filename, model1_groups, model2_groups)
-
-vegan::adonis2(
-  dist_mat~model1_groups+model2_groups, data = expl_df, by = "margin", 
-  permutations = 999, method = "manhattan")
-```
-
-The NMDS plots of the high-resolution depth samples were also illustrative of sample dissimilarity with clear SLA and depth trends.
-The anticyclonic samples grouped tightly and showed little variance with depth or triplicate when compared to the cyclonic samples with the exception of DCM triplicate B, which was highly distinct for unknown reasons.
-In contrast, the cyclonic samples were much more variable as is perhaps expected when the biomass is in the form of larger phytoplankton that are less homogenous in the environment.
-One notable aspect of their clustering was the way in which the deepest samples (DCM plus 20m) grouped most closely to the anticyclonic samples, perhaps indicating that below the DCM the metabolome rapidly approaches a uniform deep-water signal.
-This result was noted previously in the 175 meter samples of the eddy transect in @Kumler2023, where samples from the deep euphotic zone showed greater intra-depth similarity than samples from the DCM or 25 meters.
-
-Given the confluence of both depth and SLA signals and the way that the boxplots of Figure 2 would confound the SLA signal with the depth variance, we used k-means clustering to group similar compounds and provide a reduced dimensionality space for visualization.
-This revealed four major patterns of metabolite response, with a majority of compounds responding to eddy state (Clusters 1, 2, and 4 in Figure 5).
-It is interesting to note that the sole cluster in which abundances in the anticyclone are greater than those of the cyclone is also the only cluster to generally increase in abundance with depth (Cluster 4), perhaps indicating that compounds more concentrated in the anticyclone are the same kind of degradation products and recalcitrant carbon typically found at depth.
-Cluster 3, which shows no clear SLA difference and only a slight change with depth, includes compounds which are either unaffected by the environment or simply lack an environmental signal due to large inter-sample variability.
-
-The Hawaiian Eddy Experiment (HEE) data upset many of the expectations [Wei: consider calculating the effect size, as mentioned above.
-Will: Not sure what Wei means by this.
-Should I do effect size calculations here?
-Or just above when we didn't see an effect?] we developed during the MESO-SCOPE cruise in the previous year.
-Most surprising was the large difference between the eddies detected at the *surface*, while the DCM samples were functionally indistinguishable.
-This contrasts directly with the results from MESO-SCOPE and the results in @Gleich2024, who found eddy-driven shifts in protistan community composition to be larger at depth than at the surface.
-However, @Dugenne2023 found differences in nitrogen fixation rate and nitrogen fixer composition varying widely throughout the upper water column.
-The large surface differences were especially surprising given the large inter-replicate differences between the HEE surface samples in this study, while the DCM samples tended to be much more consistent.
-
-## Univariate approaches highlight individual metabolites responding to lifted and depressed isopycnals
-
-```{r kheal comparison, include=FALSE}
-raw_kheal <- read_csv("https://raw.githubusercontent.com/kheal/Gradients1_SemiTargeted3/master/Tables/Manuscript_tables/SuppTables/Full_EnviroCulture_PeakAreas.csv")
-kheal_envirosamp <- raw_kheal %>%
-  filter(Column=="HILIC") %>%
-  mutate(polarity=ifelse(z>0, "pos", "neg")) %>%
-  distinct(MassFeature_Column, compound_name_kath=Identification, Confidence, mz, rt, polarity) %>%
-  mutate(rt=rt/60)
-
-untarg_feats %>%
-  filter(str_detect(feature, "FK")) %>%
-  select(feature, polarity, mzmed, rtmed, compound_name_will=compound_name) %>%
-  mutate(mz_lower=mzmed*(1-5/1e6)) %>%
-  mutate(mz_upper=mzmed*(1+5/1e6)) %>%
-  full_join(kheal_envirosamp, by=join_by(polarity, mz_lower<mz, mz_upper>mz)) %>%
-  filter(compound_name_kath=="Histidine") %>%
-  distinct(feature, MassFeature_Column, mzmed, rtmed, mz, rt, compound_name_will, compound_name_kath)
-
-```
-
-Several molecule- or pathway-specific narratives emerged from the metabolomic data.
-The untargeted approach used here allowed us to describe and characterize signals from small molecules whose identity was unknown, a particularly promising approach in open ocean gyres where the largest fraction of unknowns is found [@Heal2021].
-We found that our list of authentic standards covered fewer metabolites in anticyclonic eddies at all depths, with the molecular diversity of the cyclone much better characterized.
-Many of the trends detected persisted even after normalizing within each sample, indicating that the shifts discussed have implications beyond simple scaling with biomass.
-
-Of those molecules whose identity was known, the clearest response to SLA was that of the enigmatic osmolyte homarine at the DCM.
-This abundant compound increased approximately threefold in concentration from < 50 pM in the anticyclone to around 150 pM in the center of the cyclone in both the eddy transect and eddy center datasets.
-The pattern was weaker in the noisier HEE data but the largest concentrations were still detected in the cyclone and the median cyclone measurement was greater than the maximum value for the anticyclone.
-This molecule has a well-established role as an osmolyte in eukaryotic phytoplankton and cyanobacteria [@Gebser2013; @Dawson2020; @Heal2021; @Durham2022] as well as a documented decrease in abundance with depth [@Heal2021], indicating that its response to SLA is potentially due to differences in physical and chemical attributes that in turn control biomass, community composition, and recycling rates.
-Curiously, the distribution of homarine was also tightly correlated with that of trigonelline.
-Although they have structural similarity they are not known to have a relationship beyond their shared function as osmolytes.
-
-```{r homarine by SLA plot, eval=FALSE}
-targ_hom <- targ_peaks %>%
-  filter(compound_name=="Homarine") %>%
-  left_join(filled_file_metadata) %>%
-  filter(samp_type=="Smp")
-mt_hom_gp <- targ_hom %>%
-  filter(ms_run=="MT") %>%
-  filter(depth=="DCM") %>%
-  ggplot(aes(x=sla_corr, y=nM)) +
-  geom_hline(yintercept = 0) +
-  geom_smooth(method = "lm", formula="y~x", color="black") +
-  geom_point(aes(fill=sla_corr), pch=21, size=4) +
-  scale_fill_gradient2(
-    low = scales::muted("blue"), high = scales::muted("red"),  mid = "grey90",
-    aesthetics = "fill", guide = NULL
-  ) +
-  coord_cartesian(ylim = c(0, 0.24), xlim=c(-15.77, 24.99), expand = FALSE) +
-  labs(x="Sea level anomaly (cm)", y="Concentration of homarine (nM)") +
-  theme_bw()
-mc_hom_gp <- targ_hom %>%
-  filter(ms_run=="MC") %>%
-  ggplot(aes(x=station, y=nM, color=station)) +
-  geom_hline(yintercept = 0) +
-  geom_boxplot(linewidth=0.75) +
-  coord_cartesian(ylim = c(0, 0.24), xlim=c(0.5, 2.5), expand = FALSE) + 
-  scale_x_discrete(breaks = c("L1", "L2"), labels = c("Cyclone\ncenter", "Anticyclone\ncenter")) +
-  scale_color_manual(values = c(scales::muted("blue"), scales::muted("red")),
-                     breaks = c("L1", "L2")) +
-  theme_bw() +
-  theme(axis.text.y = element_blank(), axis.title.y = element_blank(),
-        legend.position = "none", axis.title.x = element_blank())
-egg::ggarrange(mt_hom_gp, mc_hom_gp, nrow=1, widths = c(2, 1))
-
-targ_hom %>%
-  filter(ms_run=="FK") %>%
-  filter(depth=="Surface") %>%
-  # with(t.test(formula=nM~sla_class))
-  # with(wilcox.test(formula=nM~sla_class))
-  ggplot(aes(x=sla_class, y=nM, color=sla_class)) +
-  geom_hline(yintercept = 0) +
-  geom_boxplot(linewidth=0.75) +
-  coord_cartesian(ylim = c(0, 0.5), xlim=c(0.5, 2.5), expand = FALSE) +
-  scale_color_manual(values = c(scales::muted("blue"), scales::muted("red")),
-                     breaks = c("Cyclone", "Anticyclone")) +
-  theme_bw()
-
-```
-
-Isethionate is a known osmolyte thus far found exclusively in eukaryotes [@Durham2022] and strongly associated with a few diatoms in particular [@Heal2021].
-This compound was enriched in the cyclone of the DCM in both the MESO-SCOPE and HEE cruises along with its precursor taurine, validating earlier results from @Barone2022 that showed strong enrichment of eukaryotic phytoplankton at the DCM and *Pseudo-nitzschia* in particular.
-Taurine was also among the compounds highly enriched in the anomalous 15 meter sample from Station 11, likely along with its N-trimethylated form taurine betaine (N-trimethyltaurine).
-While putative in our analysis, taurine betaine has been previously reported in corals and red algae, where it probably has a destabilizing influence on proteins [@Impellizzeri1975; @Hill2017; @Bruzdziak2018].
-While *Trichodesmium* are notoriously patchy [@Rouco2014; @Rouco2016], we did not see a strong match to the cultured cyanobacteria in any of the other metabolites with this pattern and the osmolyte homoserine betaine shown by @Pade2016 to be characteristic of *T.
-erythraeum* strain IMS101 was missing.
-
-```{r isethionate and taurine, eval=FALSE}
-targ_peaks %>%
-  filter(compound_name%in%c("Isethionic acid", "Taurine")) %>%
-  filter(filename!="180821_Smp_MS11C215m_A.mzML") %>%
-  left_join(filled_file_metadata) %>%
-  filter(samp_type=="Smp") %>%
-  mutate(depth=fct_inorder(depth)) %>%
-  mutate(ms_run=factor(ms_run, levels=c("MT", "MC", "FK"))) %>%
-  mutate(sla_class=factor(sla_class, levels=c("Anticyclone", "Neither", "Cyclone"))) %>%
-  # group_by(compound_name, ms_run, sla_class, depth) %>%
-  # summarise(nM=mean(nM)) %>%
-  # pivot_wider(names_from = sla_class, values_from = nM)
-  ggplot() +
-  geom_boxplot(aes(x=depth, y=nM, color=sla_class)) +
-  facet_grid(compound_name~ms_run, scales="free")
-```
-
-The strong SLA signal detected among the 175 meter samples was partially surprising to us given our expectation about the strongest effect at the DCM where eddy effects lift large concentrations of nutrients above the 1% light level.
-Instead, we found more compounds overall to be significantly different across the eddy transect in the 175 meter samples than we did at the DCM (α=0.05, 43 compounds at 175 meters vs 22 at the DCM).
-This was likely driven by the enhancement of organic matter and heterotrophic picoplankton below the DCM as seen in @Barone2019, with the enrichment of nucleobases in the 175 meter samples additionally hint at increases in bacterial biomass given that nucleobases tend to be highly abundant in bacterial cultures [@Heal2021].
-The abundance of a mass feature putatively identified as acetylcholine in the 175 meter cyclone samples followed the same general trend as arsenobetaine, both enriched at depth and in the anticyclone, and may be another compound that results from heterotrophic degradation of phytoplankton metabolites [@Durham2022].
-These results raise important questions about the depth at which the community is isolated from SLA effects, with additional data from @Barone2022 and @Gleich2024 indicating that even at 250 meters eddy effects are still discernable.
-
-A final pattern worth noting is the enrichment of metabolites in the urea-ornithine cycle (OUC) among the samples taken during the HEE cruise.
-Aspartate, citrulline, ornithine, and malic acid were all found to be more abundant in the HEE samples relative to MESO-SCOPE.
-The OUC is a staple of metazoan metabolism and is present in stramenopiles as well where it enables rapid response to increases in nitrogen supply [@Allen2011; @Bender2014].
-The enrichment of these metabolites in the HEE cruise may indicate that there was still rapid uptake and turnover of nitrogen in the 2018 eddy pair while the 2017 eddies were less active, despite both eddy pairs being in a similar state of stable anticyclone and weakening cyclone [@Dugenne2023].
-
-Differences between the two cruises could also be due to differences in eddy age and/or the different period of the year when sampling took place during the two cruises (March/April in 2018 instead of the June/July sampling in 2017) resulting in shifts in community composition and response.
-@Barone2022 documents the 2017 cruise community composition, reporting abundant Prymnesiophyceae at both 15 meters and the DCM, with other significant contributions from dinoflagellates, diatoms, and green algae.
-In the 2018 cruise both haptophytes and dinoflagellates again dominated the eukaryotic fraction, though surprisingly little was contributed by diatoms [@Gleich2024].
-@Dugenne2023 and @Harke2021 also highlight uncontrolled differences between the two sets of eddies based on their origin and age.
-
-Missing links between the environment with its associated community and the composition of particulate matter make it difficult to estimate this important control on the marine carbon cycle.
-This is especially true for dynamic and short-lived environments such as eddies.
-Metabolites offer a way to bridge this gap in knowledge but the relative impact of eddies on their contribution and fate was not previously estimated.
-In this work, we have shown the sensitivity of the metabolome to various environmental and community factors and contrasted it to the well-established effect of depth to constrain the importance of including eddy effects in models of elemental cycling.
-
-# Conclusions
-
-This study reports how changes in the biogeochemistry and community composition due to mesoscale eddies affect the metabolome and thereby the composition of particulate organic matter.
-A transect across adjacent mesoscale eddies of opposite polarity showed that many metabolites track closely with metrics of overall biomass, with eddy effects stronger at 175 meters than at 15 meters or the deep chlorophyll maximum (DCM).
-High-resolution depth sampling of the DCM at the center of each eddy elucidated several known and unknown biomarkers likely corresponding to the previously documented increase in eukaryotic phytoplankton in the cyclone.
-Metabolite clusters aligned well with expected trends in abundance with depth and between the two eddies, with convergence below the DCM towards a general deep-sea metabolite signal for many metabolites.
-By contrasting these effects with a follow-up analysis in the centers of a separate pair of mesoscale eddies of opposite polarity the following year, we learned that the impact of eddies may be influenced by their origin, age, and period of the year.
-Finally, the metabolites with the strongest responses to eddy effects were unknown, and a relatively smaller proportion of known metabolites in the anticyclone demonstrates the utility of untargeted metabolomics in exploring environmental variation and the large amount of information that is missed by only investigating known compounds.
-
-# Methods
-
-## Cruise information
-Samples were collected from two cruises in the North Pacific Subtropical Gyre near Station ALOHA that targeted strong mesoscale eddy features as described in @Dugenne2023 and @Gleich2024.
-Briefly, the 2017 MESO-SCOPE cruise (Microbial Ecology of the Surface Ocean-Simons Collaboration on Ocean Processes and Ecology, KM1709 on the *R/V Kilo Moana*) consisted of a transect across adjacent cyclonic and anticyclonic eddies as well as two long-term Lagrangian stations at the center of each eddy.
-The cyclonic eddy had a maximum negative sea level anomaly (SLA) of -20 centimeters and the anticyclonic eddy reached +24 centimeters.
-The transect samples were taken at various times of day as the ship transected the adjacent eddies while the eddy center samples were all collected between 5 and 8 pm.
-In 2018, the Hawaiian Eddy Experiment (HEE, FK180310 on the *R/V Falkor*) targeted new cyclonic and anticyclonic eddies in approximately the same location with samples taken at the center of each eddy (maximum negative anomaly in the cyclone = -15 cm, maximum positive anomaly in the anticyclone = +26 cm) (Figure 1).
-
-## Biogeochemical data
-
-Biogeochemical measurements were collected as described in @Barone2022 and @Dugenne2023 mostly following protocols used by the HOT program (http://hahana.soest.hawaii.edu/hot/methods/results.html).
-Briefly, all environmental data was collected via CTD rosette except SLA which was measured via satellite.
-Particulate carbon and particulate nitrogen were measured using an elemental analyzer.
-Nitrate + nitrite (N+N) was measured on an autoanalyzer except where concentrations were below 100 nM in which case they were measured using chemiluminescence.
-Soluble reactive phosphorus (phosphate, PO$_4^{3-}$) was measured on an autoanalyzer or following the magnesium induced coprecipitation method.
-Values were interpolated linearly to the depths at which metabolite samples were taken using all data collected at the given station.
-Two particulate carbon values were determined to be spurious, with concentrations 2-3 times higher than typical Station ALOHA values, and were instead estimated using a best-fit linear regression against beam attenuation.
-
-## Sample collection for particulate metabolites
-Samples were obtained using the onboard CTD rosette to collect water from 15 meters, the deep chlorophyll maximum (DCM), and 175 meters during the MESO-SCOPE eddy transect and from the DCM ±10 and ±20 meters during the eddy center sampling in the MESO-SCOPE cruise and from 25 meters and the DCM during the HEE cruise.
-The DCM was determined visually from fluorometer data during the CTD downcast and Niskin bottles were tripped during the return trip to the surface.
-Seawater from each depth was sampled in triplicate by firing one Niskin bottle for each sample.
-Samples were brought to the surface and decanted into prewashed (3x with DI, 3x with sampled seawater) polycarbonate bottles for filtration.
-10L samples were filtered by peristaltic pump onto 142mm 0.2 µm Durapore filters held by polycarbonate filter holders on a Masterflex tubing line.
-Pressures were kept as low as possible while still producing a reasonable rate of flow through the filter, approximately 250-500 mL per minute.
-Samples were then removed from the filter holder using solvent-washed tweezers and placed into pre-combusted aluminum foil packets that were then flash-frozen in liquid nitrogen before being stored at -80 °C until extraction.
-A methodological blank was also collected by running filtrate through a new filter and then treated identically to the samples.
-
-## Metabolite sample extraction
-Extraction followed a modified Bligh & Dyer approach as detailed in @Boysen2018.
-Briefly, filters were randomized and added to PTFE centrifuge tubes with a 1:1 mix of 100 µm and 400 µm silica beads, approximately 2 mL -20 °C Optima-grade dichloromethane, and approximately 3 mL -20 °C 1:1 methanol/water solution (both also Optima-grade).
-Isotope labeled extraction standards were also added during this step (Supplemental Table 1).
-The samples were then bead-beaten three times, followed by triplicate extraction of the aqueous layer and addition with fresh methanol/water mixture with additional bead-beatings in between.
-Samples were then dried down under ultrapure nitrogen gas and warmed using a Fisher-Scientific Reacti-Therm module.
-Dried aqueous fractions were reconstituted in 380 µL Optima-grade water and amended with 20 µL isotope-labeled injection standards (Supplemental Table 1) to measure the variability introduced by chromatography and ionization.
-The reconstituted fraction was syringe-filtered into precombusted glass inserts in HPLC vials to remove any potential clogging material.
-Samples were then additionally diluted 1:1 with Optima-grade water to prevent overloading on the column and reduce salt effects.
-
-A pooled sample was created by combining 20 µL of each sample into the same HPLC vial, and a 1:1 dilution with water created a half-strength pooled sample from that to assess matrix effects and obscuring variation [@Boysen2018].
-Also run alongside the environmental samples were samples of authentic standards split into two mixes (Supplemental Table 2).
-Standards were dissolved in Optima-grade water to help with identification and into an aliquot of the pooled sample to quantify matrix response factors.
-HPLC vials containing the samples were frozen at -80 °C until thawing shortly before injection.
-
-## HPLC-MS methods
-Separate HPLC-MS runs were used for the MESO-SCOPE eddy transect, the MESO-SCOPE eddy center, and the HEE cruise data.
-Eddy center samples were run in February 2018, eddy transect samples were run later during August of that year, and the HEE samples were run in July 2019.
-Each run was treated as a single batch and injected in sequence while maintaining the randomization that had occurred during extraction to minimize chromatographic shifts from solvent or column switching.
-
-For each run, a Waters Acquity I-Class UPLC with a SeQuant ZIC-pHILIC column (5 µm particle size, 2.1 mm x 150 mm, from Millipore) was used with 10 mM ammonium carbonate in 85:15 acetonitrile to water (Solvent A) and 10 mM ammonium carbonate in 85:15 water to acetonitrile (Solvent B) at a flow rate of 0.15 mL/min.
-The column was held at 100% A for 2 minutes, ramped to 64% B over 18 minutes, ramped to 100% B over 1 minute, held at 100% B for 5 minutes, and equilibrated at 100% A for 25 minutes (50 minutes total).
-The column was maintained at 30 °C.
-The injection volume was 2 µL for samples and standard mixes.
-When starting a batch, the column was equilibrated at the starting conditions for at least 30 minutes.
-To improve the performance of the HILIC column, we maintained the same injection volume, kept the instrument running water blanks between samples as necessary, and injected standards in a representative matrix (the pooled sample) in addition to standards in water.
-After each batch, the column was flushed with 10 mM ammonium carbonate in 85:15 water to acetonitrile for 20 to 30 minutes.
-
-The Waters Acquity UPLC was coupled to a Thermo Q Exactive HF hybrid Orbitrap high resolution mass spectrometer equipped with a heated electrospray ionization source (H-ESI).
-The H-ESI voltage was set to 3.3 kV and sheath gas, auxiliary gas, and sweep gas flow rates were set at 16, 3, and 1, respectively.
-The capillary and auxiliary gas heater temperatures were maintained at 320°C and 100°C, respectively.
-Full scan analyses were performed with polarity switching and a scan range of 60 to 900 *m/z* at a resolution of 60,000.
-The instrument was mass calibrated at 200 *m/z* before each run began and once again during the eddy transect sample set to ensure calibrations were always performed within 3-4 days.
-All data files were then converted to an open-source mzML format and vendor centroided via Proteowizard’s msConvert tool.
-All mzML files have been uploaded to Metabolomics Workbench under Project ID PR001738.
-
-## Targeted metabolomic methods
-
-Compounds for which we had an authentic standard (Supplemental Table 2) were manually integrated using Skyline and MSDIAL with the standard mixes used to ensure the correct peak was integrated by matching retention time.
-Compounds that were visually determined to have poor peak quality in the samples were removed from the downstream analysis.
-Skyline was used for both the eddy center samples and the HEE cruise data while MSDIAL was used for the eddy transect data due to the larger number of files in that dataset.
-Raw peak areas were then normalized to their best-matched internal standard (Supplemental Table 1) per @Boysen2018 except that all compounds were normalized to their best match no matter how minimal the improvement.
-Normalized peaks were then calibrated via a single-point standard curve as determined uniquely for each compound from the authentic standard mixes and used to convert normalized peak area into moles per µL injected and scaled to the amount of seawater filtered to provide a final estimate of environmental concentration for each compound in moles per liter of seawater filtered.
-
-```{r supp table 1 standards list}
-clean_stans <- read_csv("../metadata/clean_stans.csv")
-raafay_notes <- read_csv("../targeted/raw_data/Raafay_HILIC_notes.csv") %>%
-  select(compound_name, in_samps)
-
-IS_table <- clean_stans %>%
-  filter(compound_type=="Internal Standard") %>%
-  select(compound_name, formula, stan_mz, stan_rt, conc_um, date_added, polarity) %>%
-  mutate(ms_run=case_when(
-    date_added<180205~"All",
-    date_added<=180821~"MESO-SCOPE Eddy Transect and HEE",
-    date_added<=190715~"HEE only",
-    TRUE~"None"
-  )) %>%
-  filter(ms_run!="None") %>%
-  mutate(polarity=ifelse(polarity=="pos", "Positive", "Negative")) %>% 
-  mutate(stan_type=case_when(
-    compound_name%in%c("L-Cysteic acid, 2H3", "3-Sulfolactate, 13C3", "Sulfoacetic acid, 13C2", "Isethionic acid, 13C2", "Taurine, 2H4")~"Extraction",
-    TRUE~"Internal"
-  )) %>%
-  arrange(stan_mz) %>%
-  select(`Compound name`=compound_name, `Molecular formula`=formula,
-         `m/z ratio`=stan_mz, `Retention time (min)`=stan_rt,
-         `Concentration (uM)`=conc_um, `Polarity integrated`=polarity,
-         `Which runs`=ms_run, `Extraction or internal`=stan_type)
-
-Auth_table <- clean_stans %>%
-  mutate(ms_run=case_when(
-    date_added<180205~"All",
-    date_added<=180821~"MESO-SCOPE Eddy Transect and HEE",
-    date_added<=190715~"HEE only",
-    TRUE~"None"
-  )) %>%
-  filter(ms_run!="None") %>%
-  left_join(raafay_notes) %>%
-  mutate(polarity=ifelse(polarity=="pos", "Positive", "Negative")) %>%
-  arrange(stan_mz) %>%
-  select(`Compound name`=compound_name, `Molecular formula`=formula,
-         `m/z ratio`=stan_mz, `Retention time (min)`=stan_rt,
-         `Concentration (uM)`=conc_um, `Polarity integrated`=polarity,
-         `Standard mix`=mix, `Which runs`=ms_run, `Good quality`=in_samps)
-
-wb <- createWorkbook(creator = "William Kumler, wkumler@uw.edu")
-addWorksheet(wb, sheetName = "Table 1 Internal Standards")
-writeData(x=IS_table, wb = wb, sheet = "Table 1 Internal Standards")
-
-addWorksheet(wb, sheetName = "Table 2 Authentic Standards")
-writeData(x=Auth_table, wb = wb, sheet = "Table 2 Authentic Standards")
-
-saveWorkbook(wb, file = "supplement/Supplemental Tables.xlsx", overwrite = TRUE)
-```
-
-## Untargeted metabolomic methods
-
-Samples were also processed via the untargeted workflow detailed in @Kumler2023.
-In short, XCMS [@R-xcms] was used for feature detection with CentWave peakpicking, Obiwarp alignment, and peak density correspondence/grouping.
-This was followed by manual inspection of the extracted ion chromatogram for each mass feature by an expert for quality with only those peaks classified as "Good" used in the analysis here.
-Features were then matched with and normalized to an internal standard (Supplemental Table 1) as described above.
-
-## Statistics
-
-All statistics were run in R version 4.3.3.
-We used multivariate statistics provided by the vegan package [@R-vegan] to assess the overall impact of various metrics across the metabolome followed by univariate statistics to investigate individually significant responses.
-Non-parametric tests were used when data were unlikely to obey parametric assumptions and permutational statistics were preferred wherever possible.
-Multiple comparison problems were controlled where necessary using a false-discovery rate correction [@Benjamini1995].
-For the analyses involving the eddy transect samples, sea level anomaly was treated as a continuous variable with the exception of the PERMANOVA, for which SLA was categorized as cyclonic if less than -5 cm and anticyclonic if greater than 5 cm, with a third category for stations in between [@Anderson2001].
-Both the MESO-SCOPE eddy center and HEE data treated SLA as a categorical variable.
-
-# References
-
-```{r packages, eval=FALSE}
-packages_used <- list.files("..", pattern = "*\\.R$|*\\.Rmd$", recursive = TRUE) %>%
-  str_subset("old|scripts", negate=TRUE) %>%
-  map(function(file_name){
-    v <- readLines(paste0("../", file_name))
-    libs <- unique(str_extract(v, "(?<=library\\().*?(?=\\)|,)"))
-    refs <- unique(str_extract(v, "[[:alnum:]]*(?=::)"))
-    pkgs <- unique(c(na.omit(libs), na.omit(refs)))
-  }) %>% unlist() %>% unique()
-knitr::write_bib(packages_used[nchar(packages_used)>0], file = "r-packages.bib")
-```
-
-<div id="refs"></div>
-
-# Supplement
 ## Supp fig 1: PCA axes 1-3 vs SLA by depth

### 2024-07-01 manuscript/manuscript.Rmd
Manuscript ready to be submitted!

@@
   arrange(desc(s)) %>%
-  left_join(untarg_feats %>% select(feature, mzmed, rtmed, compound_name, polarity))
+  left_join(untarg_feats %>% select(feature, mzmed, rtmed, compound_name, polarity)) %>%
   mutate(compound_name=ifelse(feature=="MS_NEG_0155", "Bromine79", compound_name)) %>%
@@
   filter(feature!="MS_NEG_0068") %>%
-  head(20) %>%
+  head(30) %>%
   print(n=Inf)
@@
 It is interesting to note that the sole cluster in which abundances in the anticyclone are greater than those of the cyclone is also the only cluster to generally increase in abundance with depth (Cluster 4), perhaps indicating that compounds more concentrated in the anticyclone are the same kind of degradation products and recalcitrant carbon typically found at depth.
-Cluster 3, which shows no clear SLA difference and only a slight change with depth, includes compounds which are either unaffected by the environment or simply lack an environmental signal due to large inter-sample variability.
-
-The Hawaiian Eddy Experiment (HEE) data upset many of the expectations [Wei: consider calculating the effect size, as mentioned above.
-Will: Not sure what Wei means by this.
-Should I do effect size calculations here?
-Or just above when we didn't see an effect?] we developed during the MESO-SCOPE cruise in the previous year.
+
+The Hawaiian Eddy Experiment (HEE) data upset many of the expectations we developed during the MESO-SCOPE cruise in the previous year.
 Most surprising was the large difference between the eddies detected at the *surface*, while the DCM samples were functionally indistinguishable.
@@
 # Supplement
-## Supplemental methods: Trichodesmium bloom in MS11C215mA
+## Supplemental methods: Trichodesmium bloom in MS11C215mA?
 

### 2024-07-01 manuscript/manuscript.Rmd
Adding acknowledgements and data availability statements

@@
 Additionally, our targeted list accounted for 13 out of the top 20 largest molecular features by total peak area with an additional 4 features annotated as the inorganic ions for 17/20 of the largest features known.
-However, these calculations should be taken as estimates because peak area does not correspond directly to environmental concentration and it is entirely possible for abundant environmental compounds to have small peak areas if they ionize poorly on the mass spectrometer or degrade during the sampling and extraction process.
+However, these calculations should be taken as estimates because peak area does not correspond directly to environmental concentration and it is entirely possible for abundant environmental compounds to have small peak areas if they ionize poorly on the mass spectrometer or are removed during the sampling and extraction process.
 
@@
 
+# Acknowledgements
+
+The authors would like to thank the other members of the Ingalls Lab who provided assistance in sample processing, integration, and analysis.
+We are also grateful to the SCOPE science team (Tara Clemente and Tim Burrell) for sample collection during the Hawaiian Eddy Experiment.
+We would like to acknowledge the captain and crew of the R/V *Kilo Moana* and R/V *Falkor* during the KM1709 and FK180310 cruises for making this science possible.
+We would also like to acknowledge the analysts involved in the biogeochemical measurements for their willingness to provide data and answer questions (Eric Grabowski, Karin Bjorkman, and Rhea Foreman).
+This work was supported by grants from the Simons Foundation (SCOPE Award ID 329108 to AI, SF Award ID 385428 to AI).
+
+The authors have no conflicts of interest to declare.
+
+# Data availability
+
+Data and code used for this manuscript are all available online.
+Biogeochemical data can be sourced from the Simons SCOPE website at https://scope.soest.hawaii.edu/data/ and metabolomics data has been uploaded to Metabolomics Workbench under Project ID PR001738, where the data can be accessed directly via its DOI: 10.21228/M82719.
+Scripts used to process metabolomics data are available on GitHub at https://github.com/wkumler/MesoscopeMetabolomicsManuscript.
+
 # Methods

### 2024-07-05 manuscript/manuscript.Rmd
Adding Wei's acknowledgements

@@
 We would also like to acknowledge the analysts involved in the biogeochemical measurements for their willingness to provide data and answer questions (Eric Grabowski, Karin Bjorkman, and Rhea Foreman).
-This work was supported by grants from the Simons Foundation (SCOPE Award ID 329108 to AI, SF Award ID 385428 to AI).
+This work was supported by grants from the Simons Foundation (SCOPE Award ID 329108 to AI, SF Award ID 385428 to AI, Postdoctoral Fellowship in Marine Microbial Ecology ID 548565 to WQ)
 

### 2024-07-05 manuscript/manuscript.Rmd
Adding Importance statement

@@
 By analyzing depth variability (accounting for 20-40% of metabolomic variability across ~150 meters) and the vertical displacement of isopycnal surfaces (explaining 10-20% of variability across a sea level anomaly range of 40 centimeters and a spatial distance of 300 kilometers), this analysis constrains the importance of mesoscale eddies in shaping the chemical composition of particulate matter in the largest biomes on the planet.
+
+# Importance
+
+Mesoscale eddies are common ocean surface currents that circulate seawater vertically and horizontally.
+This stirring effect causes shifts in the ocean's biogeochemistry and community composition that presumably alter the functions and fluxes of the organisms in this surface environment, but this has not been measured directly.
+In this article, we used metabolomics to determine the effect eddies have on the quantity and composition of the organic carbon produced in the surface ocean.
+We found that many small, polar molecules track with overall biomass but that there were significant differences between the eddy states.
+We also found that a few metabolites stand out as biomarkers of eukaryotic phytoplankton enrichment under the higher nutrient supply from depth.
+This work answers outstanding questions about the importance of these common ocean features in marine flux models.
 

### 2024-07-10 manuscript/manuscript.Rmd
Updating significance statement w Anitra edits

--- 2024-07-05 manuscript/manuscript.Rmd
@@
 By analyzing depth variability (accounting for 20-40% of metabolomic variability across ~150 meters) and the vertical displacement of isopycnal surfaces (explaining 10-20% of variability across a sea level anomaly range of 40 centimeters and a spatial distance of 300 kilometers), this analysis constrains the importance of mesoscale eddies in shaping the chemical composition of particulate matter in the largest biomes on the planet.
+
+# Importance
+
+Mesoscale eddies are common ocean surface currents that circulate seawater vertically and horizontally.
+This stirring effect alters biogeochemistry and planktonic community composition.
+Here, we use metabolomics to determine how these eddy-induced changes influence the nature of organic carbon across an eddy dipole.
+We found that many small, polar molecules track with the overall particulate carbon in the system and that there were significant differences in metabolite composition between eddy states.
+A few metabolites reflected the increased importance of eukaryotic phytoplankton that were enriched by the higher nutrient supply from depth in the cyclonic eddies.
+Anticyclones contained more compounds that reflected a higher degree of degradation.
+This work answers outstanding questions about the importance of these common ocean features in shaping microbial community function.
 

### 2024-07-10 manuscript/manuscript.Rmd
Updating CSL file

@@
 - r-packages.bib
-csl: limnology-and-oceanography.csl
+csl: msystems.csl
 ---

### 2024-07-10 manuscript/manuscript.Rmd
Minor consistency things noted during proofing

--- 2024-07-10 manuscript/manuscript.Rmd
@@
 The most significantly different compounds at the DCM in each direction, however, were both unknowns.
-The lowest p-value ($6.4*10^{-6}$ after FDR correction) in the DCM data was a mass feature with an *m/z* of 173.09211, a retention time also around 8.6 minutes, and enrichment in the anticyclone with a putative chemical formula of [M+H] = C$_7$H$_{13}$N$_2$O$_3$, possibly glycylproline or prolylglycine.
-The next-lowest (p = $2.5*10^{-4}$ after FDR correction) was enriched in the cyclone and had an *m/z* of 275.0712 and a retention time of 11.4 minutes with a putative formula of C$_{18}$H$_{11}$O$_3$.
+The lowest p-value ($6.4 × 10^{-6}$ after FDR correction) in the DCM data was a mass feature with an *m/z* of 173.09211, a retention time also around 8.6 minutes, and enrichment in the anticyclone with a putative chemical formula of [M+H] = C$_7$H$_{13}$N$_2$O$_3$, possibly glycylproline or prolylglycine.
+The next-lowest (p = $2.5 × 10^{-4}$ after FDR correction) was enriched in the cyclone and had an *m/z* of 275.0712 and a retention time of 11.4 minutes with a putative formula of C$_{18}$H$_{11}$O$_3$.
 
@@
 Samples were collected from two cruises in the North Pacific Subtropical Gyre near Station ALOHA that targeted strong mesoscale eddy features as described in @Dugenne2023 and @Gleich2024.
-Briefly, the 2017 MESO-SCOPE cruise (Microbial Ecology of the Surface Ocean-Simons Collaboration on Ocean Processes and Ecology, KM1709 on the *R/V Kilo Moana*) consisted of a transect across adjacent cyclonic and anticyclonic eddies as well as two long-term Lagrangian stations at the center of each eddy.
+Briefly, the 2017 MESO-SCOPE cruise (Microbial Ecology of the Surface Ocean-Simons Collaboration on Ocean Processes and Ecology, KM1709 on the R/V *Kilo Moana*) consisted of a transect across adjacent cyclonic and anticyclonic eddies as well as two long-term Lagrangian stations at the center of each eddy.
 The cyclonic eddy had a maximum negative sea level anomaly (SLA) of -20 centimeters and the anticyclonic eddy reached +24 centimeters.
 The transect samples were taken at various times of day as the ship transected the adjacent eddies while the eddy center samples were all collected between 5 and 8 pm.
-In 2018, the Hawaiian Eddy Experiment (HEE, FK180310 on the *R/V Falkor*) targeted new cyclonic and anticyclonic eddies in approximately the same location with samples taken at the center of each eddy (maximum negative anomaly in the cyclone = -15 cm, maximum positive anomaly in the anticyclone = +26 cm) (Figure 1).
+In 2018, the Hawaiian Eddy Experiment (HEE, FK180310 on the R/V *Falkor*) targeted new cyclonic and anticyclonic eddies in approximately the same location with samples taken at the center of each eddy (maximum negative anomaly in the cyclone = -15 cm, maximum positive anomaly in the anticyclone = +26 cm) (Figure 1).
 

### 2024-07-11 manuscript/manuscript.Rmd
Adding author attributions to manuscript directly
Using hacky footnote version but I think it looks ok?

@@
 title: "Metabolites reflect variability introduced by mesoscale eddies in the North Pacific Subtropical Gyre"
-author: "William Kumler, Wei Qin, Rachel A. Lundeen, Benedetto Barone, Laura T. Carlson, and Anitra E. Ingalls"
-date: "`r Sys.Date()`"
+author:
+  - William Kumler^[School of Oceanography, University of Washington, Seattle, WA, USA], Wei Qin^[School of Biological Sciences, Institute for Environmental Genomics, University of Oklahoma, Norman, OK, USA], Rachel A.
+Lundeen$^1$, Benedetto Barone^[Daniel K.
+Inouye Center for Microbial Oceanography, Research and Education and Department of Oceanography, University of Hawai'i at Mānoa, Honolulu, HI, USA], Laura T.
+Carlson$^1$, Anitra E.
+Ingalls$^{1*}$ (aingalls@uw.edu)
 output: 

### 2024-07-11 manuscript/manuscript.Rmd
Fiddling with the attributions again for mSystems formatting

--- 2024-07-10 manuscript/manuscript.Rmd
@@
 title: "Metabolites reflect variability introduced by mesoscale eddies in the North Pacific Subtropical Gyre"
-author: "William Kumler, Wei Qin, Rachel A. Lundeen, Benedetto Barone, Laura T. Carlson, and Anitra E. Ingalls"
-date: "`r Sys.Date()`"
+author:
+  - William Kumler,$^a$ Wei Qin,$^{a*}$ Rachel A. Lundeen,$^a$ Benedetto Barone,$^b$ Laura T. Carlson,$^a$ Anitra E. Ingalls$^a$#
 output: 
@@
 ---
+
+$^a$School of Oceanography, University of Washington, Seattle, WA, USA
+
+$^b$Daniel K. Inouye Center for Microbial Oceanography: Research and Education and Department of Oceanography, University of Hawai'i at Mānoa, Honolulu, HI, USA
+
+Running head: Metabolites reflect variability introduced by eddies
+
+\#Address correspondence to Anitra E. Ingalls, aingalls\@uw.edu
+
+\*Present Address: School of Biological Sciences, Institute for Environmental Genomics, University of Oklahoma, Norman, OK, USA
+
+Word counts:
+
+  - Abstract: 216
+  - Importance: 123
+  - Main text (excluding Methods, figure captions, references, etc.): 5453
 

### 2024-08-06 manuscript/manuscript.Rmd
Beginning reformatting for Frontiers
Moving Methods ahead of Results, moving Fig 1 into Methods

@@
 
+# Materials and Methods
+
+## Cruise information
+Samples were collected from two cruises in the North Pacific Subtropical Gyre near Station ALOHA that targeted strong mesoscale eddy features as described in @Dugenne2023 and @Gleich2024 (Figure 1).
+Briefly, the 2017 MESO-SCOPE cruise (Microbial Ecology of the Surface Ocean-Simons Collaboration on Ocean Processes and Ecology, KM1709 on the R/V *Kilo Moana*) consisted of a transect across adjacent cyclonic and anticyclonic eddies as well as two long-term Lagrangian stations at the center of each eddy.
+The cyclonic eddy had a maximum negative sea level anomaly (SLA) of -20 centimeters and the anticyclonic eddy reached +24 centimeters.
+The transect samples were taken at various times of day as the ship transected the adjacent eddies while the eddy center samples were all collected between 5 and 8 pm.
+In 2018, the Hawaiian Eddy Experiment (HEE, FK180310 on the R/V *Falkor*) targeted new cyclonic and anticyclonic eddies in approximately the same location with samples taken at the center of each eddy (maximum negative anomaly in the cyclone = -15 cm, maximum positive anomaly in the anticyclone = +26 cm) (Figure 1).
+
+![](figures/MapForWill_v3.jpg)
+
+*Figure 1: Sampling during MESO-SCOPE and the Hawaiian Eddy Experiment (HEE).
+Cruise bounds are shown in the large central map with Station ALOHA colored as a point in red near the Hawaiian Islands.
+Yellow stars denote sampling locations and station numbers relative to the sea level anomaly contours in the background during 28 June 2017 for the MESO-SCOPE cruise (left) and the 6 April 2018 for the HEE cruise (right).
+Lagrangian sampling near the eddy centers of the MESO-SCOPE cruise was performed by following drifters deployed in proximity of Stations 6 and 12.*
+
+## Biogeochemical data
+
+Biogeochemical measurements were collected as described in @Barone2022 and @Dugenne2023 mostly following protocols used by the HOT program (http://hahana.soest.hawaii.edu/hot/methods/results.html).
+Briefly, all environmental data was collected via CTD rosette except SLA which was measured via satellite.
+Particulate carbon and particulate nitrogen were measured using an elemental analyzer.
+Nitrate + nitrite (N+N) was measured on an autoanalyzer except where concentrations were below 100 nM in which case they were measured using chemiluminescence.
+Soluble reactive phosphorus (phosphate, PO$_4^{3-}$) was measured on an autoanalyzer or following the magnesium induced coprecipitation method.
+Values were interpolated linearly to the depths at which metabolite samples were taken using all data collected at the given station.
+Two particulate carbon values were determined to be spurious, with concentrations 2-3 times higher than typical Station ALOHA values, and were instead estimated using a best-fit linear regression against beam attenuation.
+
+## Sample collection for particulate metabolites
+Samples were obtained using the onboard CTD rosette to collect water from 15 meters, the deep chlorophyll maximum (DCM), and 175 meters during the MESO-SCOPE eddy transect and from the DCM ±10 and ±20 meters during the eddy center sampling in the MESO-SCOPE cruise and from 25 meters and the DCM during the HEE cruise.
+The DCM was determined visually from fluorometer data during the CTD downcast and Niskin bottles were tripped during the return trip to the surface.
+Seawater from each depth was sampled in triplicate by firing one Niskin bottle for each sample.
+Samples were brought to the surface and decanted into prewashed (3x with DI, 3x with sampled seawater) polycarbonate bottles for filtration.
+10L samples were filtered by peristaltic pump onto 142mm 0.2 µm Durapore filters held by polycarbonate filter holders on a Masterflex tubing line.
+Pressures were kept as low as possible while still producing a reasonable rate of flow through the filter, approximately 250-500 mL per minute.
+Samples were then removed from the filter holder using solvent-washed tweezers and placed into pre-combusted aluminum foil packets that were then flash-frozen in liquid nitrogen before being stored at -80 °C until extraction.
+A methodological blank was also collected by running filtrate through a new filter and then treated identically to the samples.
+
+## Metabolite sample extraction
+Extraction followed a modified Bligh & Dyer approach as detailed in @Boysen2018.
+Briefly, filters were randomized and added to PTFE centrifuge tubes with a 1:1 mix of 100 µm and 400 µm silica beads, approximately 2 mL -20 °C Optima-grade dichloromethane, and approximately 3 mL -20 °C 1:1 methanol/water solution (both also Optima-grade).
+Isotope labeled extraction standards were also added during this step (Supplemental Table 1).
+The samples were then bead-beaten three times, followed by triplicate extraction of the aqueous layer and addition with fresh methanol/water mixture with additional bead-beatings in between.
+Samples were then dried down under ultrapure nitrogen gas and warmed using a Fisher-Scientific Reacti-Therm module.
+Dried aqueous fractions were reconstituted in 380 µL Optima-grade water and amended with 20 µL isotope-labeled injection standards (Supplemental Table 1) to measure the variability introduced by chromatography and ionization.
+The reconstituted fraction was syringe-filtered into precombusted glass inserts in HPLC vials to remove any potential clogging material.
+Samples were then additionally diluted 1:1 with Optima-grade water to prevent overloading on the column and reduce salt effects.
+
+A pooled sample was created by combining 20 µL of each sample into the same HPLC vial, and a 1:1 dilution with water created a half-strength pooled sample from that to assess matrix effects and obscuring variation [@Boysen2018].
+Also run alongside the environmental samples were samples of authentic standards split into two mixes (Supplemental Table 2).
+Standards were dissolved in Optima-grade water to help with identification and into an aliquot of the pooled sample to quantify matrix response factors.
+HPLC vials containing the samples were frozen at -80 °C until thawing shortly before injection.
+
+## HPLC-MS methods
+Separate HPLC-MS runs were used for the MESO-SCOPE eddy transect, the MESO-SCOPE eddy center, and the HEE cruise data.
+Eddy center samples were run in February 2018, eddy transect samples were run later during August of that year, and the HEE samples were run in July 2019.
+Each run was treated as a single batch and injected in sequence while maintaining the randomization that had occurred during extraction to minimize chromatographic shifts from solvent or column switching.
+
+For each run, a Waters Acquity I-Class UPLC with a SeQuant ZIC-pHILIC column (5 µm particle size, 2.1 mm x 150 mm, from Millipore) was used with 10 mM ammonium carbonate in 85:15 acetonitrile to water (Solvent A) and 10 mM ammonium carbonate in 85:15 water to acetonitrile (Solvent B) at a flow rate of 0.15 mL/min.
+The column was held at 100% A for 2 minutes, ramped to 64% B over 18 minutes, ramped to 100% B over 1 minute, held at 100% B for 5 minutes, and equilibrated at 100% A for 25 minutes (50 minutes total).
+The column was maintained at 30 °C.
+The injection volume was 2 µL for samples and standard mixes.
+When starting a batch, the column was equilibrated at the starting conditions for at least 30 minutes.
+To improve the performance of the HILIC column, we maintained the same injection volume, kept the instrument running water blanks between samples as necessary, and injected standards in a representative matrix (the pooled sample) in addition to standards in water.
+After each batch, the column was flushed with 10 mM ammonium carbonate in 85:15 water to acetonitrile for 20 to 30 minutes.
+
+The Waters Acquity UPLC was coupled to a Thermo Q Exactive HF hybrid Orbitrap high resolution mass spectrometer equipped with a heated electrospray ionization source (H-ESI).
+The H-ESI voltage was set to 3.3 kV and sheath gas, auxiliary gas, and sweep gas flow rates were set at 16, 3, and 1, respectively.
+The capillary and auxiliary gas heater temperatures were maintained at 320°C and 100°C, respectively.
+Full scan analyses were performed with polarity switching and a scan range of 60 to 900 *m/z* at a resolution of 60,000.
+The instrument was mass calibrated at 200 *m/z* before each run began and once again during the eddy transect sample set to ensure calibrations were always performed within 3-4 days.
+All data files were then converted to an open-source mzML format and vendor centroided via Proteowizard’s msConvert tool.
+All mzML files have been uploaded to Metabolomics Workbench under Project ID PR001738.
+
+## Targeted metabolomic methods
+
+Compounds for which we had an authentic standard (Supplemental Table 2) were manually integrated using Skyline and MSDIAL with the standard mixes used to ensure the correct peak was integrated by matching retention time.
+Compounds that were visually determined to have poor peak quality in the samples were removed from the downstream analysis.
+Skyline was used for both the eddy center samples and the HEE cruise data while MSDIAL was used for the eddy transect data due to the larger number of files in that dataset.
+Raw peak areas were then normalized to their best-matched internal standard (Supplemental Table 1) per @Boysen2018 except that all compounds were normalized to their best match no matter how minimal the improvement.
+Normalized peaks were then calibrated via a single-point standard curve as determined uniquely for each compound from the authentic standard mixes and used to convert normalized peak area into moles per µL injected and scaled to the amount of seawater filtered to provide a final estimate of environmental concentration for each compound in moles per liter of seawater filtered.
+
+```{r supp table 1 standards list}
+clean_stans <- read_csv("../metadata/clean_stans.csv")
+raafay_notes <- read_csv("../targeted/raw_data/Raafay_HILIC_notes.csv") %>%
+  select(compound_name, in_samps)
+
+IS_table <- clean_stans %>%
+  filter(compound_type=="Internal Standard") %>%
+  select(compound_name, formula, stan_mz, stan_rt, conc_um, date_added, polarity) %>%
+  mutate(ms_run=case_when(
+    date_added<180205~"All",
+    date_added<=180821~"MESO-SCOPE Eddy Transect and HEE",
+    date_added<=190715~"HEE only",
+    TRUE~"None"
+  )) %>%
+  filter(ms_run!="None") %>%
+  mutate(polarity=ifelse(polarity=="pos", "Positive", "Negative")) %>% 
+  mutate(stan_type=case_when(
+    compound_name%in%c("L-Cysteic acid, 2H3", "3-Sulfolactate, 13C3", "Sulfoacetic acid, 13C2", "Isethionic acid, 13C2", "Taurine, 2H4")~"Extraction",
+    TRUE~"Internal"
+  )) %>%
+  arrange(stan_mz) %>%
+  select(`Compound name`=compound_name, `Molecular formula`=formula,
+         `m/z ratio`=stan_mz, `Retention time (min)`=stan_rt,
+         `Concentration (uM)`=conc_um, `Polarity integrated`=polarity,
+         `Which runs`=ms_run, `Extraction or internal`=stan_type)
+
+Auth_table <- clean_stans %>%
+  mutate(ms_run=case_when(
+    date_added<180205~"All",
+    date_added<=180821~"MESO-SCOPE Eddy Transect and HEE",
+    date_added<=190715~"HEE only",
+    TRUE~"None"
+  )) %>%
+  filter(ms_run!="None") %>%
+  left_join(raafay_notes) %>%
+  mutate(polarity=ifelse(polarity=="pos", "Positive", "Negative")) %>%
+  arrange(stan_mz) %>%
+  select(`Compound name`=compound_name, `Molecular formula`=formula,
+         `m/z ratio`=stan_mz, `Retention time (min)`=stan_rt,
+         `Concentration (uM)`=conc_um, `Polarity integrated`=polarity,
+         `Standard mix`=mix, `Which runs`=ms_run, `Good quality`=in_samps)
+
+wb <- createWorkbook(creator = "William Kumler, wkumler@uw.edu")
+addWorksheet(wb, sheetName = "Table 1 Internal Standards")
+writeData(x=IS_table, wb = wb, sheet = "Table 1 Internal Standards")
+
+addWorksheet(wb, sheetName = "Table 2 Authentic Standards")
+writeData(x=Auth_table, wb = wb, sheet = "Table 2 Authentic Standards")
+
+saveWorkbook(wb, file = "supplement/Supplemental Tables.xlsx", overwrite = TRUE)
+```
+
+## Untargeted metabolomic methods
+
+Samples were also processed via the untargeted workflow detailed in @Kumler2023.
+In short, XCMS [@R-xcms] was used for feature detection with CentWave peakpicking, Obiwarp alignment, and peak density correspondence/grouping.
+This was followed by manual inspection of the extracted ion chromatogram for each mass feature by an expert for quality with only those peaks classified as "Good" used in the analysis here.
+Features were then matched with and normalized to an internal standard (Supplemental Table 1) as described above.
+
+## Statistics
+
+All statistics were run in R version 4.3.3.
+We used multivariate statistics provided by the vegan package [@R-vegan] to assess the overall impact of various metrics across the metabolome followed by univariate statistics to investigate individually significant responses.
+Non-parametric tests were used when data were unlikely to obey parametric assumptions and permutational statistics were preferred wherever possible.
+Multiple comparison problems were controlled where necessary using a false-discovery rate correction [@Benjamini1995].
+For the analyses involving the eddy transect samples, sea level anomaly was treated as a continuous variable with the exception of the PERMANOVA, for which SLA was categorized as cyclonic if less than -5 cm and anticyclonic if greater than 5 cm, with a third category for stations in between [@Anderson2001].
+Both the MESO-SCOPE eddy center and HEE analyses treated SLA as a categorical variable.
+
 # Results
@@
 We analyzed these three different datasets separately and discuss the findings from each in sequence below.
-
-![](figures/MapForWill_v3.jpg)
-
-*Figure 1: Sampling during MESO-SCOPE and the Hawaiian Eddy Experiment (HEE).
-Cruise bounds are shown in the large central map with Station ALOHA colored as a point in red near the Hawaiian Islands.
-Yellow stars denote sampling locations and station numbers relative to the sea level anomaly contours in the background during 28 June 2017 for the MESO-SCOPE cruise (left) and the 6 April 2018 for the HEE cruise (right).
-Lagrangian sampling near the eddy centers of the MESO-SCOPE cruise was performed by following drifters deployed in proximity of Stations 6 and 12.*
 
@@
 
-# Methods
-
-## Cruise information
-Samples were collected from two cruises in the North Pacific Subtropical Gyre near Station ALOHA that targeted strong mesoscale eddy features as described in @Dugenne2023 and @Gleich2024.
-Briefly, the 2017 MESO-SCOPE cruise (Microbial Ecology of the Surface Ocean-Simons Collaboration on Ocean Processes and Ecology, KM1709 on the R/V *Kilo Moana*) consisted of a transect across adjacent cyclonic and anticyclonic eddies as well as two long-term Lagrangian stations at the center of each eddy.
-The cyclonic eddy had a maximum negative sea level anomaly (SLA) of -20 centimeters and the anticyclonic eddy reached +24 centimeters.
-The transect samples were taken at various times of day as the ship transected the adjacent eddies while the eddy center samples were all collected between 5 and 8 pm.
-In 2018, the Hawaiian Eddy Experiment (HEE, FK180310 on the R/V *Falkor*) targeted new cyclonic and anticyclonic eddies in approximately the same location with samples taken at the center of each eddy (maximum negative anomaly in the cyclone = -15 cm, maximum positive anomaly in the anticyclone = +26 cm) (Figure 1).
-
-## Biogeochemical data
-
-Biogeochemical measurements were collected as described in @Barone2022 and @Dugenne2023 mostly following protocols used by the HOT program (http://hahana.soest.hawaii.edu/hot/methods/results.html).
-Briefly, all environmental data was collected via CTD rosette except SLA which was measured via satellite.
-Particulate carbon and particulate nitrogen were measured using an elemental analyzer.
-Nitrate + nitrite (N+N) was measured on an autoanalyzer except where concentrations were below 100 nM in which case they were measured using chemiluminescence.
-Soluble reactive phosphorus (phosphate, PO$_4^{3-}$) was measured on an autoanalyzer or following the magnesium induced coprecipitation method.
-Values were interpolated linearly to the depths at which metabolite samples were taken using all data collected at the given station.
-Two particulate carbon values were determined to be spurious, with concentrations 2-3 times higher than typical Station ALOHA values, and were instead estimated using a best-fit linear regression against beam attenuation.
-
-## Sample collection for particulate metabolites
-Samples were obtained using the onboard CTD rosette to collect water from 15 meters, the deep chlorophyll maximum (DCM), and 175 meters during the MESO-SCOPE eddy transect and from the DCM ±10 and ±20 meters during the eddy center sampling in the MESO-SCOPE cruise and from 25 meters and the DCM during the HEE cruise.
-The DCM was determined visually from fluorometer data during the CTD downcast and Niskin bottles were tripped during the return trip to the surface.
-Seawater from each depth was sampled in triplicate by firing one Niskin bottle for each sample.
-Samples were brought to the surface and decanted into prewashed (3x with DI, 3x with sampled seawater) polycarbonate bottles for filtration.
-10L samples were filtered by peristaltic pump onto 142mm 0.2 µm Durapore filters held by polycarbonate filter holders on a Masterflex tubing line.
-Pressures were kept as low as possible while still producing a reasonable rate of flow through the filter, approximately 250-500 mL per minute.
-Samples were then removed from the filter holder using solvent-washed tweezers and placed into pre-combusted aluminum foil packets that were then flash-frozen in liquid nitrogen before being stored at -80 °C until extraction.
-A methodological blank was also collected by running filtrate through a new filter and then treated identically to the samples.
-
-## Metabolite sample extraction
-Extraction followed a modified Bligh & Dyer approach as detailed in @Boysen2018.
-Briefly, filters were randomized and added to PTFE centrifuge tubes with a 1:1 mix of 100 µm and 400 µm silica beads, approximately 2 mL -20 °C Optima-grade dichloromethane, and approximately 3 mL -20 °C 1:1 methanol/water solution (both also Optima-grade).
-Isotope labeled extraction standards were also added during this step (Supplemental Table 1).
-The samples were then bead-beaten three times, followed by triplicate extraction of the aqueous layer and addition with fresh methanol/water mixture with additional bead-beatings in between.
-Samples were then dried down under ultrapure nitrogen gas and warmed using a Fisher-Scientific Reacti-Therm module.
-Dried aqueous fractions were reconstituted in 380 µL Optima-grade water and amended with 20 µL isotope-labeled injection standards (Supplemental Table 1) to measure the variability introduced by chromatography and ionization.
-The reconstituted fraction was syringe-filtered into precombusted glass inserts in HPLC vials to remove any potential clogging material.
-Samples were then additionally diluted 1:1 with Optima-grade water to prevent overloading on the column and reduce salt effects.
-
-A pooled sample was created by combining 20 µL of each sample into the same HPLC vial, and a 1:1 dilution with water created a half-strength pooled sample from that to assess matrix effects and obscuring variation [@Boysen2018].
-Also run alongside the environmental samples were samples of authentic standards split into two mixes (Supplemental Table 2).
-Standards were dissolved in Optima-grade water to help with identification and into an aliquot of the pooled sample to quantify matrix response factors.
-HPLC vials containing the samples were frozen at -80 °C until thawing shortly before injection.
-
-## HPLC-MS methods
-Separate HPLC-MS runs were used for the MESO-SCOPE eddy transect, the MESO-SCOPE eddy center, and the HEE cruise data.
-Eddy center samples were run in February 2018, eddy transect samples were run later during August of that year, and the HEE samples were run in July 2019.
-Each run was treated as a single batch and injected in sequence while maintaining the randomization that had occurred during extraction to minimize chromatographic shifts from solvent or column switching.
-
-For each run, a Waters Acquity I-Class UPLC with a SeQuant ZIC-pHILIC column (5 µm particle size, 2.1 mm x 150 mm, from Millipore) was used with 10 mM ammonium carbonate in 85:15 acetonitrile to water (Solvent A) and 10 mM ammonium carbonate in 85:15 water to acetonitrile (Solvent B) at a flow rate of 0.15 mL/min.
-The column was held at 100% A for 2 minutes, ramped to 64% B over 18 minutes, ramped to 100% B over 1 minute, held at 100% B for 5 minutes, and equilibrated at 100% A for 25 minutes (50 minutes total).
-The column was maintained at 30 °C.
-The injection volume was 2 µL for samples and standard mixes.
-When starting a batch, the column was equilibrated at the starting conditions for at least 30 minutes.
-To improve the performance of the HILIC column, we maintained the same injection volume, kept the instrument running water blanks between samples as necessary, and injected standards in a representative matrix (the pooled sample) in addition to standards in water.
-After each batch, the column was flushed with 10 mM ammonium carbonate in 85:15 water to acetonitrile for 20 to 30 minutes.
-
-The Waters Acquity UPLC was coupled to a Thermo Q Exactive HF hybrid Orbitrap high resolution mass spectrometer equipped with a heated electrospray ionization source (H-ESI).
-The H-ESI voltage was set to 3.3 kV and sheath gas, auxiliary gas, and sweep gas flow rates were set at 16, 3, and 1, respectively.
-The capillary and auxiliary gas heater temperatures were maintained at 320°C and 100°C, respectively.
-Full scan analyses were performed with polarity switching and a scan range of 60 to 900 *m/z* at a resolution of 60,000.
-The instrument was mass calibrated at 200 *m/z* before each run began and once again during the eddy transect sample set to ensure calibrations were always performed within 3-4 days.
-All data files were then converted to an open-source mzML format and vendor centroided via Proteowizard’s msConvert tool.
-All mzML files have been uploaded to Metabolomics Workbench under Project ID PR001738.
-
-## Targeted metabolomic methods
-
-Compounds for which we had an authentic standard (Supplemental Table 2) were manually integrated using Skyline and MSDIAL with the standard mixes used to ensure the correct peak was integrated by matching retention time.
-Compounds that were visually determined to have poor peak quality in the samples were removed from the downstream analysis.
-Skyline was used for both the eddy center samples and the HEE cruise data while MSDIAL was used for the eddy transect data due to the larger number of files in that dataset.
-Raw peak areas were then normalized to their best-matched internal standard (Supplemental Table 1) per @Boysen2018 except that all compounds were normalized to their best match no matter how minimal the improvement.
-Normalized peaks were then calibrated via a single-point standard curve as determined uniquely for each compound from the authentic standard mixes and used to convert normalized peak area into moles per µL injected and scaled to the amount of seawater filtered to provide a final estimate of environmental concentration for each compound in moles per liter of seawater filtered.
-
-```{r supp table 1 standards list}
-clean_stans <- read_csv("../metadata/clean_stans.csv")
-raafay_notes <- read_csv("../targeted/raw_data/Raafay_HILIC_notes.csv") %>%
-  select(compound_name, in_samps)
-
-IS_table <- clean_stans %>%
-  filter(compound_type=="Internal Standard") %>%
-  select(compound_name, formula, stan_mz, stan_rt, conc_um, date_added, polarity) %>%
-  mutate(ms_run=case_when(
-    date_added<180205~"All",
-    date_added<=180821~"MESO-SCOPE Eddy Transect and HEE",
-    date_added<=190715~"HEE only",
-    TRUE~"None"
-  )) %>%
-  filter(ms_run!="None") %>%
-  mutate(polarity=ifelse(polarity=="pos", "Positive", "Negative")) %>% 
-  mutate(stan_type=case_when(
-    compound_name%in%c("L-Cysteic acid, 2H3", "3-Sulfolactate, 13C3", "Sulfoacetic acid, 13C2", "Isethionic acid, 13C2", "Taurine, 2H4")~"Extraction",
-    TRUE~"Internal"
-  )) %>%
-  arrange(stan_mz) %>%
-  select(`Compound name`=compound_name, `Molecular formula`=formula,
-         `m/z ratio`=stan_mz, `Retention time (min)`=stan_rt,
-         `Concentration (uM)`=conc_um, `Polarity integrated`=polarity,
-         `Which runs`=ms_run, `Extraction or internal`=stan_type)
-
-Auth_table <- clean_stans %>%
-  mutate(ms_run=case_when(
-    date_added<180205~"All",
-    date_added<=180821~"MESO-SCOPE Eddy Transect and HEE",
-    date_added<=190715~"HEE only",
-    TRUE~"None"
-  )) %>%
-  filter(ms_run!="None") %>%
-  left_join(raafay_notes) %>%
-  mutate(polarity=ifelse(polarity=="pos", "Positive", "Negative")) %>%
-  arrange(stan_mz) %>%
-  select(`Compound name`=compound_name, `Molecular formula`=formula,
-         `m/z ratio`=stan_mz, `Retention time (min)`=stan_rt,
-         `Concentration (uM)`=conc_um, `Polarity integrated`=polarity,
-         `Standard mix`=mix, `Which runs`=ms_run, `Good quality`=in_samps)
-
-wb <- createWorkbook(creator = "William Kumler, wkumler@uw.edu")
-addWorksheet(wb, sheetName = "Table 1 Internal Standards")
-writeData(x=IS_table, wb = wb, sheet = "Table 1 Internal Standards")
-
-addWorksheet(wb, sheetName = "Table 2 Authentic Standards")
-writeData(x=Auth_table, wb = wb, sheet = "Table 2 Authentic Standards")
-
-saveWorkbook(wb, file = "supplement/Supplemental Tables.xlsx", overwrite = TRUE)
-```
-
-## Untargeted metabolomic methods
-
-Samples were also processed via the untargeted workflow detailed in @Kumler2023.
-In short, XCMS [@R-xcms] was used for feature detection with CentWave peakpicking, Obiwarp alignment, and peak density correspondence/grouping.
-This was followed by manual inspection of the extracted ion chromatogram for each mass feature by an expert for quality with only those peaks classified as "Good" used in the analysis here.
-Features were then matched with and normalized to an internal standard (Supplemental Table 1) as described above.
-
-## Statistics
-
-All statistics were run in R version 4.3.3.
-We used multivariate statistics provided by the vegan package [@R-vegan] to assess the overall impact of various metrics across the metabolome followed by univariate statistics to investigate individually significant responses.
-Non-parametric tests were used when data were unlikely to obey parametric assumptions and permutational statistics were preferred wherever possible.
-Multiple comparison problems were controlled where necessary using a false-discovery rate correction [@Benjamini1995].
-For the analyses involving the eddy transect samples, sea level anomaly was treated as a continuous variable with the exception of the PERMANOVA, for which SLA was categorized as cyclonic if less than -5 cm and anticyclonic if greater than 5 cm, with a third category for stations in between [@Anderson2001].
-Both the MESO-SCOPE eddy center and HEE analyses treated SLA as a categorical variable.
-
 # References

### 2024-08-06 manuscript/manuscript.Rmd
Removing mSystems "Importance" statement

@@
 By analyzing depth variability (accounting for 20-40% of metabolomic variability across ~150 meters) and the vertical displacement of isopycnal surfaces (explaining 10-20% of variability across a sea level anomaly range of 40 centimeters and a spatial distance of 300 kilometers), this analysis constrains the importance of mesoscale eddies in shaping the chemical composition of particulate matter in the largest biomes on the planet.
-
-# Importance
-
-Mesoscale eddies are common ocean surface currents that circulate seawater vertically and horizontally.
-This stirring effect alters biogeochemistry and planktonic community composition.
-Here, we use metabolomics to determine how these eddy-induced changes influence the nature of organic carbon across an eddy dipole.
-We found that many small, polar molecules track with the overall particulate carbon in the system and that there were significant differences in metabolite composition between eddy states.
-A few metabolites reflected the increased importance of eukaryotic phytoplankton that were enriched by the higher nutrient supply from depth in the cyclonic eddies.
-Anticyclones contained more compounds that reflected a higher degree of degradation.
-This work answers outstanding questions about the importance of these common ocean features in shaping microbial community function.
 
@@
 
-# Conclusions
+# Conclusion
 

### 2024-08-06 manuscript/manuscript.Rmd
Switching to Frontiers formatting

@@
   word_document:
-    reference_docx: template.docx
+    reference_docx: Frontiers_Template.docx
 bibliography:
@@
 - r-packages.bib
-csl: limnology-and-oceanography.csl
+csl: frontiers-in-marine-science.csl
 ---

### 2024-08-06 manuscript/manuscript.Rmd
Adding new word counts
Fixing attributions and moving running head down

--- 2024-08-06 manuscript/manuscript.Rmd
@@
 author:
-  - William Kumler,$^a$ Wei Qin,$^{a*}$ Rachel A. Lundeen,$^a$ Benedetto Barone,$^b$ Laura T. Carlson,$^a$ Anitra E. Ingalls$^a$#
+  - William Kumler,$^1$ Wei Qin,$^{1+}$ Rachel A. Lundeen,$^1$ Benedetto Barone,$^2$ Laura T. Carlson,$^1$ Anitra E. Ingalls$^{1*}$
 output: 
   word_document:
-    reference_docx: template.docx
+    reference_docx: Frontiers_Template.docx
 bibliography:
@@
 - r-packages.bib
-csl: limnology-and-oceanography.csl
+csl: frontiers-in-marine-science.csl
 ---
 
-$^a$School of Oceanography, University of Washington, Seattle, WA, USA
-
-$^b$Daniel K. Inouye Center for Microbial Oceanography: Research and Education and Department of Oceanography, University of Hawai'i at Mānoa, Honolulu, HI, USA
+$^1$School of Oceanography, University of Washington, Seattle, WA, USA
+
+$^2$Daniel K. Inouye Center for Microbial Oceanography: Research and Education and Department of Oceanography, University of Hawai'i at Mānoa, Honolulu, HI, USA
+
+$^*$Correspondence: Anitra E. Ingalls, aingalls\@uw.edu
+
+$^+$Present Address: School of Biological Sciences, Institute for Environmental Genomics, University of Oklahoma, Norman, OK, USA
+
+Word counts:
+
+  - Main text (excludes the abstract, section titles, figure and table captions, funding statement, acknowledgments, and references): 7,236
+  - Number of figures: 6
+  - Number of tables: 0
 
 Running head: Metabolites reflect variability introduced by eddies
-
-\#Address correspondence to Anitra E. Ingalls, aingalls\@uw.edu
-
-\*Present Address: School of Biological Sciences, Institute for Environmental Genomics, University of Oklahoma, Norman, OK, USA
-
-Word counts:
-
-  - Abstract: 216
-  - Importance: 123
-  - Main text (excluding Methods, figure captions, references, etc.): 5453
 

### 2024-08-14 manuscript/manuscript.Rmd
Adding new supplemental figure and moving MS11 analysis into main text again

@@
 
-  - Main text (excludes the abstract, section titles, figure and table captions, funding statement, acknowledgments, and references): 7,236
+  - Main text (excludes the abstract, section titles, figure and table captions, funding statement, acknowledgments, and references): 7,XXX
   - Number of figures: 6
-  - Number of tables: 0
+  - Number of tables: 1
 
@@
 
+### A single sample highlights a network of related metabolites
+
+One sample, taken from 15 meters at Station 11, was highly distinct from all others (including its paired triplicates) with several metabolites measuring two or more orders of magnitude larger in peak area and concentration both before and after normalization to internal standards (Supplemental figure 3).
+Heuristic exploration of potentially related masses identified a total of 14 compounds highly abundant in this one sample alone.
+The only identified compounds with this trend were taurine and acetyltaurine, though three others with *m/z* ratios of 146.1176, 148.0968, and 206.1388 were discovered to be likely isomers of known compounds (acetylcholine, hydroxyisoleucine, and dexpanthenol, respectively).
+One feature (*m/z* = 184.0637, RT = 7.0 minutes) was identified as choline sulfate using a second run of standards, while an additional mass feature at 168.0689 and eluting at 8.5 minutes was putatively flagged as taurine betaine.
+This mass was also one of the most highly enriched in the sample after taurine itself, with a peak area nearly 30 times that of the next largest sample (Supplemental Figure 4).
+No clear explanation emerged for why some compounds were hyperabundant in this surface sample, though one possibility was the *Trichodesmium* bloom that was also sampled at this time [@Dugenne2023].
+
+```{r MS11A offenders table}
+# All positive mode except N-Acetyltaurine
+ms11_offenders <- untarg_peaks %>%
+  filter(str_detect(filename, "180821")) %>%
+  group_by(feature) %>%
+  filter(sum(norm_area==0)<3) %>%
+  mutate(cursed_area=norm_area[filename=="180821_Smp_MS11C215m_A.mzML"]) %>%
+  filter(cursed_area>0) %>%
+  mutate(rel_area=norm_area/cursed_area) %>%
+  filter(filename!="180821_Smp_MS11C215m_A.mzML") %>%
+  group_by(feature) %>%
+  summarise(mean_rel=mean(rel_area), max_rel=max(rel_area)) %>%
+  # ggplot(aes(x=mean_rel, y=max_rel, label=feature)) + geom_point() + 
+  # scale_y_log10() + scale_x_log10() +
+  # geom_vline(xintercept = 0.25) + geom_hline(yintercept = 0.5)
+  filter(mean_rel<0.25 & max_rel < 0.5) %>%
+  arrange(mean_rel) %>% 
+  add_row(feature="MS_NEG_1123") %>%
+  left_join(untarg_feats %>% select(feature:compound_name)) %>%
+  mutate(mzmed=ifelse(str_detect(feature, "NEG"), mzmed+1.007276*2, mzmed)) %>%
+  filter(feature!="MS_NEG_0768") %>% # Remove taurine duplicate
+  add_row(mzmed=c(198.0794, 184.0637), rtmed=c(5, 7))
+# ms11_offenders$mzmed %>% sort() %>% `names<-`(., round(., digits=3)) %>% dist(diag = TRUE) %>% round(digits=4)
+stans <- read_csv("https://raw.githubusercontent.com/IngallsLabUW/Ingalls_Standards/master/Ingalls_Lab_Standards.csv") %>%
+  filter(Column=="HILIC" & z>0) %>%
+  select(cmpd_name=Compound_Name, formula=Empirical_Formula) %>%
+  group_by(formula) %>%
+  summarise(cmpd_name=paste(cmpd_name, collapse = ", "))
+base_mz <- 85.089149355+1.007276
+iter_mzs <- expand_grid(CH2=0:6, O=2:6) %>%
+  mutate(mz=base_mz+CH2*14.01565+O*15.994915) %>%
+  mutate(mz=round(mz, digits=5)) %>%
+  mutate(cmd=paste0("mz%between%pmppm(", mz, ", 10)")) %>%
+  mutate(formula=paste0("C", 5+CH2, "H", 11+CH2*2, "NO", O)) %>%
+  left_join(stans) %>%
+  mutate(label=ifelse(is.na(cmpd_name), formula, paste0(formula, " (", cmpd_name, ")")))
+ms11_df <- c("C11H23NO6", "C11H23NO6", "C9H19NO5", "C8H17NO5", "C6H13NO6", "C8H17NO4") %>%
+  data.frame(formula=., rtmed=c(8.5, 10, 8.5, 10.2, 5.2, 9.2)) %>%
+  left_join(iter_mzs) %>%
+  select(mzmed=mz, rtmed, formula, cmpd_name) %>%
+  bind_rows(ms11_offenders) %>%
+  arrange(mzmed) %>%
+  select(mzmed, rtmed, formula, compound_name, feature) %>%
+  mutate(formula=case_when(
+    compound_name=="Taurine"~"C2H8NO3S",
+    feature=="MS_POS_1050"~"C7H16NO2",
+    compound_name=="Hydroxyisoleucine"~"C6H14NO3",
+    compound_name=="N-Acetyltaurine"~"C4H10NO4S",
+    feature=="MS_POS_1340"~"C5H14NO3S",
+    feature=="MS_POS_1419"~"C8H18NO3",
+    mzmed%between%pmppm(184.0637, 10)~"C5H14NO4S",
+    mzmed%between%pmppm(198.0794, 10)~"C6H16NO4S",
+    compound_name=="Dexpanthenol"~"C9H20NO4",
+    TRUE~formula
+  )) %>%
+  mutate(compound_name=case_when(
+    compound_name=="(3-Carboxypropyl)trimethylammonium?"~NA,
+    compound_name=="Hydroxyisoleucine"~NA,
+    compound_name=="Dexpanthenol"~NA,
+    formula=="C5H14NO4S"~"Choline sulfate",
+    formula=="C5H14NO3S"~"Taurine betaine?",
+    TRUE~compound_name
+  )) %>%
+  mutate(rtmed=round(rtmed, 1)) %>%
+  mutate(exact_mass=str_replace(formula, "N", "N1")) %>%
+  mutate(exact_mass=str_replace(exact_mass, "S", "S1")) %>%
+  separate(exact_mass, into = c("drop", "C", "H", "N", "O", "S"), sep = "[A-Z]") %>%
+  mutate(across(C:S, as.numeric)) %>%
+  mutate(S=ifelse(is.na(S), 0, 1)) %>%
+  mutate(calc_mass=C*12+H*1.007825+N*14.003074+O*15.994915+S*31.972072) %>%
+  mutate(calc_mass=ifelse(abs(calc_mass-mzmed)>1, calc_mass+1.007276, calc_mass)) %>%
+  mutate(compound_name=ifelse(is.na(compound_name), "", compound_name)) %>%
+  select(-feature)
+
+library(arrow)
+dset <- open_dataset("../tmzMLs/MT_all_file_arr_pqds/")
+arrow_data <- ms11_df %>%
+  pmap(function(...){
+    row_data <- list(...)
+    dset %>%
+      filter(mz%between%pmppm(row_data$mzmed, 20)) %>%
+      collect() %>%
+      mutate(formula=row_data$formula)
+  }, .progress = TRUE)
+Nacetyl_row <- ms11_df %>% filter(formula=="C4H10NO4S")
+neg_msdata <- grabMSdata(list.files("../tmzMLs/neg", pattern = "180821_Smp", full.names = TRUE))
+Nacetyldata <- neg_msdata$MS1[mz%between%pmppm(Nacetyl_row$mzmed-1.007276*2, 10)] %>%
+  mutate(formula=Nacetyl_row$formula) %>%
+  mutate(filename=str_replace(filename, "tmzML", "mzML"))
+peak_lims <- arrow_data %>%
+  bind_rows() %>%
+  filter(str_detect(filename, "Smp")) %>%
+  bind_rows(Nacetyldata) %>%
+  distinct() %>%
+  left_join(ms11_df) %>%
+  filter(rt>rtmed-1 & rt<rtmed+1)
+
+ms11_offenders_gp <- peak_lims %>%
+  mutate(formula=fct_inorder(formula)) %>%
+  ggplot() +
+  geom_line(aes(x=rt, y=int, group=filename, color=filename=="180821_Smp_MS11C215m_A.mzML")) +
+  facet_wrap(~formula, ncol=3, scales="free") +
+  scale_color_manual(name="Anomalous file", breaks = c(TRUE, FALSE), values=c("red", "black")) +
+  labs(x="Retention time (min)", y="Intensity") +
+  theme(legend.position="inside", legend.position.inside = c(0.85, 0.06)) +
+  guides(color=guide_legend(override.aes=list(lwd=1.5)))
+ggsave("ms11_quirky_cmpds.png", ms11_offenders_gp, device = "png", path = "supplement",
+       width = 6.875, height = 5, units = "in", dpi = 300)
+
+mass_diffs <- peak_lims %>%
+  group_by(formula, rtmed, mzmed, calc_mass) %>%
+  summarise(mean_mz=weighted.mean(mz, int)) %>%
+  mutate(calc_mass=ifelse(formula=="C4H10NO4S", calc_mass-2*1.007276, calc_mass)) %>%
+  mutate(delta_ppm=round((mean_mz-calc_mass)*1e6/mzmed, 2)) %>%
+  select(-calc_mass)
+ms11_df %>%
+  left_join(mass_diffs) %>%
+  select(`Median m/z`=mzmed, `Median RT (min)`=rtmed, `Putative formula`=formula,
+         `Delta m/z (ppm)`=delta_ppm, `Compound name`=compound_name) %>%
+  knitr::kable()
+```
+
+| Median m/z| Median RT (min)|Putative formula | Delta m/z (ppm)|Compound name    |
+|----------:|---------------:|:----------------|---------------:|:----------------|
+|   126.0221|            11.1|C2H8NO3S         |           -2.99|Taurine          |
+|   146.1176|             8.5|C7H16NO2         |           -3.58|                 |
+|   148.0968|             8.5|C6H14NO3         |           -3.79|                 |
+|   168.0317|             4.9|C4H10NO4S        |           32.46|N-Acetyltaurine  |
+|   168.0689|             8.5|C5H14NO3S        |           -3.65|Taurine betaine? |
+|   176.1282|             8.1|C8H18NO3         |           -2.86|                 |
+|   184.0637|             7.0|C5H14NO4S        |           -2.95|Choline sulfate  |
+|   192.1230|             9.2|C8H17NO4         |            0.22|                 |
+|   196.0816|             5.2|C6H13NO6         |            7.40|                 |
+|   198.0794|             5.0|C6H16NO4S        |           -2.56|                 |
+|   206.1388|             7.1|C9H20NO4         |           -2.04|                 |
+|   208.1180|            10.2|C8H17NO5         |            0.58|                 |
+|   222.1336|             8.5|C9H19NO5         |            0.31|                 |
+|   266.1598|             8.5|C11H23NO6        |            5.71|                 |
+|   266.1598|            10.0|C11H23NO6        |            4.45|                 |
+
+*Table 1: Detected mass features that were enriched in the 15 meter surface sample from MESO-SCOPE Station 11.
+RT = retention time in minutes.
+Taurine betaine is followed by a ?
+to denote that this identification is putative.
+Mass difference between the measured m/z and putative formula is reported in parts-per-million.*
+
+```{r trichobloomsamples comparison, eval=FALSE}
+library(RaMS)
+trichodata <- list.files("../mzMLs/tricho_pos", pattern=".mzML", full.names = TRUE) %>%
+  grabMSdata()
+tricho_eics <- ms11_df %>%
+  pmap(function(...){
+  row_data <- list(...)
+  peak_data <- trichodata$MS1[mz%between%pmppm(row_data$mzmed, 10)]
+  cbind(peak_data, formula=row_data$formula)
+}) %>%
+  bind_rows()
+tricho_eics %>%
+  filter(rt%between%c(2, 15)) %>%
+  arrange(mz) %>%
+  mutate(formula=fct_inorder(formula)) %>%
+  ggplot() +
+  geom_line(aes(x=rt, y=int, group=filename)) +
+  facet_wrap(~formula, scales="free", ncol=2) +
+  geom_vline(aes(xintercept = rtmed), color="red", data=ms11_df)
+trichodata$MS1[mz%between%pmppm(162.1125, 10)] %>%
+  mutate(samp_type=str_extract(filename, "Smp|Poo|Mix1|Mix2|Mat")) %>%
+  filter(samp_type!="Mat") %>%
+  ggplot() +
+  geom_line(aes(x=rt, y=int, group=filename, color=samp_type)) +
+  ggtitle("Homoserine betaine in Tricho bloom samples")
+```
+
+We explored the possibility that a nearby *Trichodesmium* bloom was responsible for the unusual composition of this one sample by analyzing additional samples taken directly from the bloom to see whether compounds abundant in hand-picked *Trichodesmium* colonies were also abundant in the Station 11 sample.
+After searching for the 15 mass features above in the environmental sample, 7 had good-quality peaks when viewed in the bloom samples as extracted ion chromatograms, though only 4 of these matched closely in retention time.
+We also performed a search of the untargeted data for homoserine betaine, a known *T.
+erythraeum* (strain IMS101) osmolyte [@Pade2016].
+Although this compound was not run in water alongside the sample to confirm a retention time match due to a lack of authentic standards, we identified three mass features in the 162.1125 *m/z* window with retention times of 7.1, 8.4, and 9.9 minutes.
+The feature at RT 9.9 matched our carnitine standard's retention time, leaving the two earlier peaks with putative identifications of threonine betaine at 7.1 and homoserine betaine at 8.4 given the elution order of threonine before homoserine in our authentic standards.
+This order is the reverse of that in @Pade2016, which showed homoserine betaine eluting prior to carnitine.
+This elution order is expected given our use of a normal phase HILIC column and their use of a reversed phase Hypercarb column.
+They did not report an additional peak detected at this mass, leaving our putative identification of the homoserine and threonine betaines as low-confidence.
+None of the three 162.1125 features showed the same degree of enrichment in the anomalous sample as taurine or the other 13 compounds listed in Table 1, though the anomalous sample contained the largest, second-largest, and third-largest peak for carnitine, threonine betaine, and homoserine betaine, respectively.
+These results are fairly inconclusive but we are inclined to believe that a *Trichodesmium* colony is not likely to be the cause of the inflated values in this particular sample.
+
+```{r homoserinebetaine search, eval=FALSE}
+library(arrow)
+dset <- open_dataset("../tmzMLs/MT_all_file_arr_pqds/")
+hmb_eic <- dset %>%
+  filter(mz%between%pmppm(162.1125, 10)) %>%
+  # filter(mz%between%pmppm(120.0655, 10)) %>%
+  collect()
+hmb_eic %>%
+  filter(rt%between%c(6, 11)) %>%
+  # filter(rt%between%c(9, 13)) %>%
+  mutate(samp_type=str_extract(filename, "Poo|Smp|InH2O|Mat")) %>%
+  # distinct(filename, samp_type) %>%
+  # filter(samp_type!="Smp")
+  mutate(plotcol=ifelse(str_detect(filename, "180821_Smp_MS11C215m_A.mzML"), 
+                          "MS11C215m_A", samp_type)) %>%
+  ggplot() +
+  geom_line(aes(x=rt, y=int, group=filename, color=plotcol)) +
+  facet_wrap(~samp_type, ncol=1, scales="free_y")
+untarg_feats %>% 
+  filter(cruise=="MS") %>%
+  filter(mzmed%between%pmppm(162.1125))
+untarg_peaks %>%
+  filter(feature%in%c("MS_POS_1246", "MS_POS_1247", "MS_POS_1248")) %>%
+  mutate(feature=factor(feature, levels=c("MS_POS_1248", "MS_POS_1246", "MS_POS_1247"),
+                        labels=c("Threonine betaine (RT=7.1)", 
+                                 "Homoserine betaine (RT=8.4)",
+                                 "Carnitine (RT=9.9)"))) %>%
+  left_join(filled_file_metadata) %>%
+  filter(ms_run=="MT") %>%
+  mutate(station=factor(station, levels=as.character(4:14))) %>%
+  mutate(depth=factor(depth, levels=c("Surface", "DCM", "Deep"))) %>%
+  mutate(plotcol=ifelse(filename=="180821_Smp_MS11C215m_A.mzML", "red", "black")) %>%
+  ggplot(aes(x=station, y=norm_area, label=filename, color=plotcol)) +
+  geom_point() +
+  facet_grid(feature~depth, scales="free_y") +
+  geom_hline(yintercept = 0) +
+  scale_color_identity()
+untarg_peaks %>%
+  filter(feature%in%c("MS_POS_1246", "MS_POS_1247", "MS_POS_1248")) %>%
+  left_join(filled_file_metadata) %>%
+  filter(ms_run=="MT") %>%
+  group_by(feature) %>%
+  mutate(norm_area=norm_area/norm_area[filename=="180821_Smp_MS11C215m_A.mzML"]) %>%
+  group_by(feature, depth) %>%
+  summarise(mean_area=max(norm_area)) %>%
+  pivot_wider(names_from = "depth", values_from = mean_area)
+```
+
 ## Particulate organic matter compositional shifts across eddy transect
@@
 
-When analyzed separately as distinct depths, the 25 meter samples had higher variances explained by eddy (R$^2$=.26) and a lower p-value (0.013), while the DCM samples were no longer likely to be distinct between the two eddies (p-value = 0.171) (Figure 6, Supplemental Figure 4).
+When analyzed separately as distinct depths, the 25 meter samples had higher variances explained by eddy (R$^2$=.26) and a lower p-value (0.013), while the DCM samples were no longer likely to be distinct between the two eddies (p-value = 0.171) (Figure 6, Supplemental Figure 5).
 
@@
 # Supplement
-## Supplemental methods: Trichodesmium bloom in MS11C215mA?
-
-One sample, taken from 15 meters at Station 11, was highly distinct from all others (including its paired triplicates) with several metabolites measuring two or more orders of magnitude larger in peak area and concentration both before and after normalization to internal standards (Supplemental Figure 3).
-Heuristic exploration of potentially related masses identified a total of 14 compounds highly abundant in this one sample alone.
-The only identified compounds with this trend were taurine and acetyltaurine, though three others with *m/z* ratios of 146.1176, 148.0968, and 206.1388 were discovered to be likely isomers of known compounds (acetylcholine, hydroxyisoleucine, and dexpanthenol, respectively).
-One feature (*m/z* = 184.0637, RT = 7.0 minutes) was identified as choline sulfate using a second run of standards, while an additional mass feature at 168.0689 and eluting at 8.5 minutes was putatively flagged as taurine betaine.
-This mass was also one of the most highly enriched in the sample after taurine itself, with a peak area nearly 30 times that of the next largest sample.
-No clear explanation emerged for why some compounds were hyperabundant in this surface sample, though one possibility was the *Trichodesmium* bloom that was also sampled at this time [@Dugenne2023].
-
-```{r MS11A offenders table}
-ms11_offenders <- untarg_peaks %>%
-  filter(str_detect(filename, "180821")) %>%
-  group_by(feature) %>%
-  filter(sum(norm_area==0)<3) %>%
-  mutate(cursed_area=norm_area[filename=="180821_Smp_MS11C215m_A.mzML"]) %>%
-  filter(cursed_area>0) %>%
-  mutate(rel_area=norm_area/cursed_area) %>%
-  filter(filename!="180821_Smp_MS11C215m_A.mzML") %>%
-  group_by(feature) %>%
-  summarise(mean_rel=mean(rel_area), max_rel=max(rel_area)) %>%
-  # ggplot(aes(x=mean_rel, y=max_rel, label=feature)) + geom_point() + 
-  # scale_y_log10() + scale_x_log10() +
-  # geom_vline(xintercept = 0.25) + geom_hline(yintercept = 0.5)
-  filter(mean_rel<0.25 & max_rel < 0.5) %>%
-  arrange(mean_rel) %>% 
-  add_row(feature="MS_NEG_1123") %>%
-  left_join(untarg_feats %>% select(feature:compound_name)) %>%
-  mutate(mzmed=ifelse(str_detect(feature, "NEG"), mzmed+1.007276*2, mzmed)) %>%
-  filter(feature!="MS_NEG_0768") %>% # Remove taurine duplicate
-  add_row(mzmed=c(198.0794, 184.0637), rtmed=c(5, 7))
-
-# ms11_offenders$mzmed %>% sort() %>% `names<-`(., round(., digits=3)) %>% dist(diag = TRUE) %>% round(digits=4)
-
-stans <- read_csv("https://raw.githubusercontent.com/IngallsLabUW/Ingalls_Standards/master/Ingalls_Lab_Standards.csv") %>%
-  filter(Column=="HILIC" & z>0) %>%
-  select(cmpd_name=Compound_Name, formula=Empirical_Formula) %>%
-  group_by(formula) %>%
-  summarise(cmpd_name=paste(cmpd_name, collapse = ", "))
-base_mz <- 85.089149355+1.007276
-iter_mzs <- expand_grid(CH2=0:6, O=2:6) %>%
-  mutate(mz=base_mz+CH2*14.01565+O*15.994915) %>%
-  mutate(mz=round(mz, digits=5)) %>%
-  mutate(cmd=paste0("mz%between%pmppm(", mz, ", 10)")) %>%
-  mutate(formula=paste0("C", 5+CH2, "H", 11+CH2*2, "NO", O)) %>%
-  left_join(stans) %>%
-  mutate(label=ifelse(is.na(cmpd_name), formula, paste0(formula, " (", cmpd_name, ")")))
-ms11_df <- c("C11H23NO6", "C11H23NO6", "C9H19NO5", "C8H17NO5", "C6H13NO6", "C8H17NO4") %>%
-  data.frame(formula=., rtmed=c(8.5, 10, 8.5, 10.2, 5.2, 9.2)) %>%
-  left_join(iter_mzs) %>%
-  select(mzmed=mz, rtmed, formula, cmpd_name) %>%
-  bind_rows(ms11_offenders) %>%
-  arrange(mzmed) %>%
-  select(mzmed, rtmed, formula, compound_name, feature) %>%
-  mutate(formula=case_when(
-    compound_name=="Taurine"~"C2H8NO3S",
-    feature=="MS_POS_1050"~"C7H16NO2",
-    compound_name=="Hydroxyisoleucine"~"C6H14NO3",
-    compound_name=="N-Acetyltaurine"~"C4H10NO4S",
-    feature=="MS_POS_1340"~"C5H14NO3S",
-    feature=="MS_POS_1419"~"C8H18NO3",
-    mzmed%between%pmppm(184.0637, 10)~"C5H14NO4S",
-    mzmed%between%pmppm(198.0794, 10)~"C6H16NO4S",
-    compound_name=="Dexpanthenol"~"C9H20NO4",
-    TRUE~formula
-  )) %>%
-  mutate(compound_name=case_when(
-    compound_name=="(3-Carboxypropyl)trimethylammonium?"~NA,
-    compound_name=="Hydroxyisoleucine"~NA,
-    compound_name=="Dexpanthenol"~NA,
-    formula=="C5H14NO4S"~"Choline sulfate",
-    formula=="C5H14NO3S"~"Taurine betaine?",
-    TRUE~compound_name
-  )) %>%
-  mutate(rtmed=round(rtmed, 1)) %>%
-  mutate(exact_mass=str_replace(formula, "N", "N1")) %>%
-  mutate(exact_mass=str_replace(exact_mass, "S", "S1")) %>%
-  separate(exact_mass, into = c("drop", "C", "H", "N", "O", "S"), sep = "[A-Z]") %>%
-  mutate(across(C:S, as.numeric)) %>%
-  mutate(S=ifelse(is.na(S), 0, 1)) %>%
-  mutate(calc_mass=C*12+H*1.007825+N*14.003074+O*15.994915+S*31.972072) %>%
-  mutate(calc_mass=ifelse(abs(calc_mass-mzmed)>1, calc_mass+1.007276, calc_mass)) %>%
-  mutate(delta_ppm=ifelse(is.na(feature), 0, round((mzmed-calc_mass)*1e6/mzmed, 2))) %>%
-  mutate(compound_name=ifelse(is.na(compound_name), "", compound_name)) %>%
-  select(-feature) %>%
-  select(`Median m/z`=mzmed, `Median RT (min)`=rtmed, `Putative formula`=formula,
-         `Delta mz (ppm)`=delta_ppm, `Compound name`=compound_name)
-
-# knitr::kable(ms11_df)
-```
-
-| Median m/z| Median RT (min)|Putative formula | Delta mz (ppm)|Compound name    |
-|----------:|---------------:|:----------------|--------------:|:----------------|
-|   126.0221|            11.1|C2H8NO3S         |          -3.35|Taurine          |
-|   146.1176|             8.5|C7H16NO2         |          -3.63|                 |
-|   148.0968|             8.5|C6H14NO3         |          -3.88|                 |
-|   168.0317|             4.9|C4H10NO4S        |          -7.87|N-Acetyltaurine  |
-|   168.0689|             8.5|C5H14NO3S        |          -3.33|Taurine betaine? |
-|   176.1282|             8.1|C8H18NO3         |          -2.88|                 |
-|   184.0637|             7.0|C5H14NO4S        |           0.00|Choline sulfate  |
-|   192.1230|             9.2|C8H17NO4         |           0.00|                 |
-|   196.0816|             5.2|C6H13NO6         |           0.00|                 |
-|   198.0794|             5.0|C6H16NO4S        |           0.00|                 |
-|   206.1388|             7.1|C9H20NO4         |          -1.93|                 |
-|   208.1180|            10.2|C8H17NO5         |           0.00|                 |
-|   222.1336|             8.5|C9H19NO5         |           0.00|                 |
-|   266.1598|             8.5|C11H23NO6        |           0.00|                 |
-|   266.1598|            10.0|C11H23NO6        |           0.00|                 |
-
-*Table 1: Detected mass features that were enriched in the 15 meter surface sample from MESO-SCOPE Station 11.
-RT = retention time in minutes.
-Taurine betaine is followed by a ?
-to denote that this identification is putative.
-Mass difference between the median m/z and putative formula is reported in parts-per-million.*
-
-```{r trichobloomsamples comparison, eval=FALSE}
-library(RaMS)
-trichodata <- list.files("../mzMLs/tricho_pos", pattern=".mzML", full.names = TRUE) %>%
-  grabMSdata()
-tricho_eics <- ms11_df %>%
-  pmap(function(...){
-  row_data <- list(...)
-  peak_data <- trichodata$MS1[mz%between%pmppm(row_data$mzmed, 10)]
-  cbind(peak_data, formula=row_data$formula)
-}) %>%
-  bind_rows()
-
-tricho_eics %>%
-  filter(rt%between%c(2, 15)) %>%
-  arrange(mz) %>%
-  mutate(formula=fct_inorder(formula)) %>%
-  ggplot() +
-  geom_line(aes(x=rt, y=int, group=filename)) +
-  facet_wrap(~formula, scales="free", ncol=2) +
-  geom_vline(aes(xintercept = rtmed), color="red", data=ms11_df)
-
-trichodata$MS1[mz%between%pmppm(162.1125, 10)] %>%
-  mutate(samp_type=str_extract(filename, "Smp|Poo|Mix1|Mix2|Mat")) %>%
-  filter(samp_type!="Mat") %>%
-  ggplot() +
-  geom_line(aes(x=rt, y=int, group=filename, color=samp_type)) +
-  ggtitle("Homoserine betaine in Tricho bloom samples")
-```
-
-We explored the possibility that a nearby *Trichodesmium* bloom was responsible for the unusual composition of this one sample by analyzing additional samples taken directly from the bloom to see whether compounds abundant in hand-picked *Trichodesmium* colonies were also abundant in the Station 11 sample.
-After searching for the 14 mass features above in the environmental sample, 7 had good-quality peaks when viewed in the bloom samples as extracted ion chromatograms, though only 4 of these matched closely in retention time.
-We also performed a search of the untargeted data for homoserine betaine, a known *T.
-erythraeum* (strain IMS101) osmolyte [@Pade2016].
-Although this compound was not run in water alongside the sample to confirm a retention time match due to a lack of authentic standards, we identified three mass features in the 162.1125 *m/z* window with retention times of 7.1, 8.4, and 9.9 minutes.
-The feature at RT 9.9 matched our carnitine standard's retention time, leaving the two earlier peaks with putative identifications of threonine betaine at 7.1 and homoserine betaine at 8.4 given the elution order of threonine before homoserine in our authentic standards.
-This order is the reverse of that in @Pade2016, which showed homoserine betaine eluting prior to carnitine.
-This elution order is expected given our use of a normal phase HILIC column and their use of a reversed phase Hypercarb column.
-They did not report an additional peak detected at this mass, leaving our putative identification of the homoserine and threonine betaines as low-confidence.
-None of the three 162.1125 features showed the same degree of enrichment in the anomalous sample as taurine or the other 13 compounds listed in Table 1, though the anomalous sample contained the largest, second-largest, and third-largest peak for carnitine, threonine betaine, and homoserine betaine, respectively.
-These results are fairly inconclusive but we are inclined to believe that a *Trichodesmium* colony is not likely to be the cause of the inflated values in this particular sample.
-
-```{r homoserinebetaine search, eval=FALSE}
-library(arrow)
-dset <- open_dataset("../tmzMLs/MT_all_file_arr_pqds/")
-hmb_eic <- dset %>%
-  filter(mz%between%pmppm(162.1125, 10)) %>%
-  # filter(mz%between%pmppm(120.0655, 10)) %>%
-  collect()
-hmb_eic %>%
-  filter(rt%between%c(6, 11)) %>%
-  # filter(rt%between%c(9, 13)) %>%
-  mutate(samp_type=str_extract(filename, "Poo|Smp|InH2O|Mat")) %>%
-  # distinct(filename, samp_type) %>%
-  # filter(samp_type!="Smp")
-  mutate(plotcol=ifelse(str_detect(filename, "180821_Smp_MS11C215m_A.mzML"), 
-                          "MS11C215m_A", samp_type)) %>%
-  ggplot() +
-  geom_line(aes(x=rt, y=int, group=filename, color=plotcol)) +
-  facet_wrap(~samp_type, ncol=1, scales="free_y")
-
-untarg_feats %>% 
-  filter(cruise=="MS") %>%
-  filter(mzmed%between%pmppm(162.1125))
-untarg_peaks %>%
-  filter(feature%in%c("MS_POS_1246", "MS_POS_1247", "MS_POS_1248")) %>%
-  mutate(feature=factor(feature, levels=c("MS_POS_1248", "MS_POS_1246", "MS_POS_1247"),
-                        labels=c("Threonine betaine (RT=7.1)", 
-                                 "Homoserine betaine (RT=8.4)",
-                                 "Carnitine (RT=9.9)"))) %>%
-  left_join(filled_file_metadata) %>%
-  filter(ms_run=="MT") %>%
-  mutate(station=factor(station, levels=as.character(4:14))) %>%
-  mutate(depth=factor(depth, levels=c("Surface", "DCM", "Deep"))) %>%
-  mutate(plotcol=ifelse(filename=="180821_Smp_MS11C215m_A.mzML", "red", "black")) %>%
-  ggplot(aes(x=station, y=norm_area, label=filename, color=plotcol)) +
-  geom_point() +
-  facet_grid(feature~depth, scales="free_y") +
-  geom_hline(yintercept = 0) +
-  scale_color_identity()
-
-untarg_peaks %>%
-  filter(feature%in%c("MS_POS_1246", "MS_POS_1247", "MS_POS_1248")) %>%
-  left_join(filled_file_metadata) %>%
-  filter(ms_run=="MT") %>%
-  group_by(feature) %>%
-  mutate(norm_area=norm_area/norm_area[filename=="180821_Smp_MS11C215m_A.mzML"]) %>%
-  group_by(feature, depth) %>%
-  summarise(mean_area=max(norm_area)) %>%
-  pivot_wider(names_from = "depth", values_from = mean_area)
-```
-
 ## Supp fig 1: PCA axes 1-3 vs SLA by depth
@@
 
-## Supp fig 4: HEE NMDS and PERMANOVA broken down by depth
+## Supp fig 4: Chromatograms of MS11 compounds
+
+![](supplement/ms11_quirky_cmpds.png)
+
+*Supplemental Figure 4: Extracted ion chromatograms of features identified in Table 1, separated by their predicted molecular formula.
+An m/z window of 10 ppm and a retention time window of 2 minutes was used to extract the data from the mzML files.
+The anomalous sample taken from 15 meters at Station 11 is shown in red while all other samples are shown in black.*
+
+## Supp fig 5: HEE NMDS and PERMANOVA broken down by depth
 
@@
 
-*Supplemental Figure 4: Plots of the metabolome during the Hawaiian Eddy Experiment, broken down by depth.
+*Supplemental Figure 5: Plots of the metabolome during the Hawaiian Eddy Experiment, broken down by depth.
 The top row of plots are non-metric multidimensional scaling (NMDS) plots, in which points correspond to individual samples and have been colored by their corrected sea level height anomaly.

### 2024-08-15 manuscript/manuscript.Rmd
ce
Adding final word count

@@
 
-  - Main text (excludes the abstract, section titles, figure and table captions, funding statement, acknowledgments, and references): 7,XXX
+  - Main text (excludes the abstract, section titles, figure and table captions, funding statement, acknowledgments, and references): 7,750
   - Number of figures: 6
@@
 The dashed line across the figure represents the 0.05 level of significance as a visual cue for metabolites above which the differences between the eddies are unlikely to be due to chance.
-DMSP = dimethylsulfoniopropionate, DMS-Ac = dimethylsulfonioacetate*
+DMSP = dimethylsulfoniopropionate, DMS-Ac = dimethylsulfonioacetate.*
 

### 2024-08-15 manuscript/manuscript.Rmd
Switching to tif in manuscript

@@
 
-![](figures/MapForWill_v3.jpg)
+![](figures/fig_1_MapForWill_v3.jpg)
 
@@
                                 nrow=1, rel_widths = c(0.9, 0.1))
-ggsave("nmds_and_med_metab.png", plot = nmds_and_med_metab, device = "png", 
+ggsave("fig_2_nmds_and_med_metab.tif", plot = nmds_and_med_metab, device = "tiff", 
        path = "figures", width = 6.875, height = 4.5, units = "in", dpi = 300,
@@
 
-![](figures/nmds_and_med_metab.png)
+![](figures/fig_2_nmds_and_med_metab.tif)
 
@@
                                      draw = FALSE, heights = c(0.4, 0.2, 0.4))
-ggsave("targ_gp_w_sla_frac.png", plot = targ_gp_w_sla_frac, device = "png",
+ggsave("fig_3_targ_gp_w_sla_frac.tif", plot = targ_gp_w_sla_frac, device = "tif",
        path = "figures", width = 6.875, height = 5.5, units = "in", dpi = 300,
@@
 
-![](figures/targ_gp_w_sla_frac.png)
+![](figures/fig_3_targ_gp_w_sla_frac.tif)
 
@@
 
-ggsave("MC_nmds_gp.png", plot = MC_nmds_gp, device = "png", 
+ggsave("fig_4_MC_nmds_gp.tif", plot = MC_nmds_gp, device = "tif", 
        path = "figures", width = 6.875, height = 4.5, units = "in", dpi = 300)
@@
 
-![](figures/MC_nmds_gp.png)
+![](figures/fig_4_MC_nmds_gp.tif)
 
@@
   ncol=1, heights = c(1.5, 2), draw = FALSE)
-ggsave("kclust_volcano_gp.png", plot = kclust_volcano_gp, device = "png", 
+ggsave("fig_5_kclust_volcano_gp.tif", plot = kclust_volcano_gp, device = "tif", 
        path = "figures", width = 6.875, height = 5, units = "in", dpi = 300,
@@
 
-![](figures/kclust_volcano_gp.png)
+![](figures/fig_5_kclust_volcano_gp.tif)
 
@@
 
-ggsave("fk_nmdsplot.png", plot = fk_nmdsplot, device = "png",
+ggsave("fig_6_fk_nmdsplot.tif", plot = fk_nmdsplot, device = "tif",
        path = "figures", width = 6.875, height = 4, units = "in", dpi = 300,
@@
 
-![](figures/fk_nmdsplot.png)
+![](figures/fig_6_fk_nmdsplot.tif)
 

### 2024-08-15 manuscript/manuscript.Rmd
Final recompilation and ce for Frontiers

@@
 ```{r supp table 1 standards list}
-clean_stans <- read_csv("../metadata/clean_stans.csv")
+clean_stans_supp <- read_csv("../metadata/clean_stans.csv")
 raafay_notes <- read_csv("../targeted/raw_data/Raafay_HILIC_notes.csv") %>%
@@
 
-IS_table <- clean_stans %>%
+IS_table <- clean_stans_supp %>%
   filter(compound_type=="Internal Standard") %>%
@@
 
-Auth_table <- clean_stans %>%
+Auth_table <- clean_stans_supp %>%
   mutate(ms_run=case_when(
@@
 
-```{r PC_vs_totnM regression values, include=FALSE}
+```{r frac_pc_in_metab_df creation}
 frac_pc_in_metab_df <- targ_peaks %>%
@@
   mutate(station=factor(station, levels=c(as.character(4:14))))
-
+```
+
+```{r PC_vs_totnM regression values, include=FALSE}
 no_outlier_frac_pc_in_metab_df <- frac_pc_in_metab_df %>%
@@
 The DCM and 175 meter samples had the stronger trends ($\beta_{DCM} = 22.6\pm5.53$, $r= .593$, p-value < 0.001; $\beta_{175m} = 23.9\pm7.36$, $r= .503$, p-value = 0.003) while 15 meter samples showed no significant relationship ($\beta_{15m} = 3.4\pm12.9$, $r= .049$, p-value = 0.792).
-This meant that metabolites contributed a relatively fixed fraction of carbon to the particulate pool with our 53 quantified metabolites representing between 1.5% and 5% of the carbon in the system (Figure 2).
+This meant that metabolites contributed a relatively fixed fraction of carbon to the particulate pool with our 53 quantified metabolites representing between 1.5% and 5% of the carbon in the system (Figure 3).
 Particulate nitrogen was tightly correlated with particulate carbon ($r=.933$) but the fraction of particulate nitrogen represented by the quantified metabolites was slightly higher across the transect with contributions typically between 2% and 6%.
@@
 
-The most abundant intracellular metabolite quantified in the eddy transect samples was trimethylamine N-oxide, with an average concentration of 1.38 nM and a distinct DCM maximum (Figure 2), though a few samples had concentrations an order of magnitude higher for unknown reasons (Supplemental Figure 3).
+The most abundant intracellular metabolite quantified in the eddy transect samples was trimethylamine N-oxide, with an average concentration of 1.38 nM and a distinct DCM maximum (Figure 3), though a few samples had concentrations an order of magnitude higher for unknown reasons (Supplemental Figure 3).
 A majority of the known molecules (37/53) had a similar pattern with a subsurface maximum at the DCM, including glycine betaine, glutamate, and guanine.
@@
       mutate(formula=row_data$formula)
-  }, .progress = TRUE)
+  })
 Nacetyl_row <- ms11_df %>% filter(formula=="C4H10NO4S")
-neg_msdata <- grabMSdata(list.files("../tmzMLs/neg", pattern = "180821_Smp", full.names = TRUE))
+neg_msdata <- grabMSdata(list.files("../tmzMLs/neg", pattern = "180821_Smp", full.names = TRUE), verbosity = 0)
 Nacetyldata <- neg_msdata$MS1[mz%between%pmppm(Nacetyl_row$mzmed-1.007276*2, 10)] %>%
@@
   select(-calc_mass)
-ms11_df %>%
+final_ms11_df <- ms11_df %>%
   left_join(mass_diffs) %>%
   select(`Median m/z`=mzmed, `Median RT (min)`=rtmed, `Putative formula`=formula,
-         `Delta m/z (ppm)`=delta_ppm, `Compound name`=compound_name) %>%
-  knitr::kable()
+         `Delta m/z (ppm)`=delta_ppm, `Compound name`=compound_name)
+# knitr::kable(final_ms11_df)
 ```
@@
 trichodata <- list.files("../mzMLs/tricho_pos", pattern=".mzML", full.names = TRUE) %>%
-  grabMSdata()
+  grabMSdata(verbosity = 0)
 tricho_eics <- ms11_df %>%
@@
 
-Given the confluence of both depth and SLA signals and the way that the boxplots of Figure 2 would confound the SLA signal with the depth variance, we used k-means clustering to group similar compounds and provide a reduced dimensionality space for visualization.
+Given the confluence of both depth and SLA signals and the way that the median metabolite plots of Figure 2 would confound the SLA signal with the depth variance, we used k-means clustering to group similar compounds and provide a reduced dimensionality space for visualization.
 This revealed four major patterns of metabolite response, with a majority of compounds responding to eddy state (Clusters 1, 2, and 4 in Figure 5).

### 2024-10-18 manuscript/manuscript.Rmd
Updating README with low-res figures and relative file paths

@@
        path = "figures", width = 6.875, height = 4.5, units = "in", dpi = 300,
+       bg="white")
+ggsave("fig_2_nmds_and_med_metab.tif", plot = nmds_and_med_metab, device = "tiff", 
+       path = "figures/lowres", width = 6.875, height = 4.5, units = "in", dpi = 72,
        bg="white")
@@
        path = "figures", width = 6.875, height = 5.5, units = "in", dpi = 300,
+       bg = "white")
+ggsave("fig_3_targ_gp_w_sla_frac.tif", plot = targ_gp_w_sla_frac, device = "tif",
+       path = "figures/lowres", width = 6.875, height = 5.5, units = "in", dpi = 72,
        bg = "white")
@@
        path = "figures", width = 6.875, height = 4.5, units = "in", dpi = 300)
+ggsave("fig_4_MC_nmds_gp.tif", plot = MC_nmds_gp, device = "tif", 
+       path = "figures/lowres", width = 6.875, height = 4.5, units = "in", dpi = 72)
 ```
@@
        bg="white")
+ggsave("fig_5_kclust_volcano_gp.tif", plot = kclust_volcano_gp, device = "tif", 
+       path = "figures/lowres", width = 6.875, height = 5, units = "in", dpi = 72,
+       bg="white")
 ```
@@
        path = "figures", width = 6.875, height = 4, units = "in", dpi = 300,
+       bg="white")
+ggsave("fig_6_fk_nmdsplot.tif", plot = fk_nmdsplot, device = "tif",
+       path = "figures/lowres", width = 6.875, height = 4, units = "in", dpi = 72,
        bg="white")

### 2024-10-18 manuscript/manuscript.Rmd
Apparently Github doesn't like tifs?

--- 2024-08-15 manuscript/manuscript.Rmd
@@
        path = "figures", width = 6.875, height = 4.5, units = "in", dpi = 300,
+       bg="white")
+ggsave("fig_2_nmds_and_med_metab.png", plot = nmds_and_med_metab, device = "png", 
+       path = "figures/lowres", width = 6.875, height = 4.5, units = "in", dpi = 72,
        bg="white")
@@
        path = "figures", width = 6.875, height = 5.5, units = "in", dpi = 300,
+       bg = "white")
+ggsave("fig_3_targ_gp_w_sla_frac.png", plot = targ_gp_w_sla_frac, device = "png",
+       path = "figures/lowres", width = 6.875, height = 5.5, units = "in", dpi = 72,
        bg = "white")
@@
        path = "figures", width = 6.875, height = 4.5, units = "in", dpi = 300)
+ggsave("fig_4_MC_nmds_gp.png", plot = MC_nmds_gp, device = "png", 
+       path = "figures/lowres", width = 6.875, height = 4.5, units = "in", dpi = 72)
 ```
@@
        bg="white")
+ggsave("fig_5_kclust_volcano_gp.png", plot = kclust_volcano_gp, device = "png", 
+       path = "figures/lowres", width = 6.875, height = 5, units = "in", dpi = 72,
+       bg="white")
 ```
@@
        path = "figures", width = 6.875, height = 4, units = "in", dpi = 300,
+       bg="white")
+ggsave("fig_6_fk_nmdsplot.png", plot = fk_nmdsplot, device = "png",
+       path = "figures/lowres", width = 6.875, height = 4, units = "in", dpi = 72,
        bg="white")

### 2024-10-18 manuscript/manuscript.Rmd
Increasing lowres figures DPI

--- 2024-08-15 manuscript/manuscript.Rmd
@@
        path = "figures", width = 6.875, height = 4.5, units = "in", dpi = 300,
+       bg="white")
+ggsave("fig_2_nmds_and_med_metab.png", plot = nmds_and_med_metab, device = "png", 
+       path = "figures/lowres", width = 6.875, height = 4.5, units = "in", dpi = 144,
        bg="white")
@@
        path = "figures", width = 6.875, height = 5.5, units = "in", dpi = 300,
+       bg = "white")
+ggsave("fig_3_targ_gp_w_sla_frac.png", plot = targ_gp_w_sla_frac, device = "png",
+       path = "figures/lowres", width = 6.875, height = 5.5, units = "in", dpi = 144,
        bg = "white")
@@
        path = "figures", width = 6.875, height = 4.5, units = "in", dpi = 300)
+ggsave("fig_4_MC_nmds_gp.png", plot = MC_nmds_gp, device = "png", 
+       path = "figures/lowres", width = 6.875, height = 4.5, units = "in", dpi = 144)
 ```
@@
        bg="white")
+ggsave("fig_5_kclust_volcano_gp.png", plot = kclust_volcano_gp, device = "png", 
+       path = "figures/lowres", width = 6.875, height = 5, units = "in", dpi = 144,
+       bg="white")
 ```
@@
        path = "figures", width = 6.875, height = 4, units = "in", dpi = 300,
+       bg="white")
+ggsave("fig_6_fk_nmdsplot.png", plot = fk_nmdsplot, device = "png",
+       path = "figures/lowres", width = 6.875, height = 4, units = "in", dpi = 144,
        bg="white")

### 2024-10-21 manuscript/manuscript.Rmd
Removing MS11/tricho hypothesis results

@@
 
-  - Main text (excludes the abstract, section titles, figure and table captions, funding statement, acknowledgments, and references): 7,750
-  - Number of figures: 6
-  - Number of tables: 1
+  - Main text (excludes the abstract, section titles, figure and table captions, funding statement, acknowledgments, and references): [X,XXX]
+  - Number of figures: [X]
+  - Number of tables: [X]
 
@@
 
-### A single sample highlights a network of related metabolites
-
-One sample, taken from 15 meters at Station 11, was highly distinct from all others (including its paired triplicates) with several metabolites measuring two or more orders of magnitude larger in peak area and concentration both before and after normalization to internal standards (Supplemental figure 3).
-Heuristic exploration of potentially related masses identified a total of 14 compounds highly abundant in this one sample alone.
-The only identified compounds with this trend were taurine and acetyltaurine, though three others with *m/z* ratios of 146.1176, 148.0968, and 206.1388 were discovered to be likely isomers of known compounds (acetylcholine, hydroxyisoleucine, and dexpanthenol, respectively).
-One feature (*m/z* = 184.0637, RT = 7.0 minutes) was identified as choline sulfate using a second run of standards, while an additional mass feature at 168.0689 and eluting at 8.5 minutes was putatively flagged as taurine betaine.
-This mass was also one of the most highly enriched in the sample after taurine itself, with a peak area nearly 30 times that of the next largest sample (Supplemental Figure 4).
-No clear explanation emerged for why some compounds were hyperabundant in this surface sample, though one possibility was the *Trichodesmium* bloom that was also sampled at this time [@Dugenne2023].
-
-```{r MS11A offenders table}
-# All positive mode except N-Acetyltaurine
-ms11_offenders <- untarg_peaks %>%
-  filter(str_detect(filename, "180821")) %>%
-  group_by(feature) %>%
-  filter(sum(norm_area==0)<3) %>%
-  mutate(cursed_area=norm_area[filename=="180821_Smp_MS11C215m_A.mzML"]) %>%
-  filter(cursed_area>0) %>%
-  mutate(rel_area=norm_area/cursed_area) %>%
-  filter(filename!="180821_Smp_MS11C215m_A.mzML") %>%
-  group_by(feature) %>%
-  summarise(mean_rel=mean(rel_area), max_rel=max(rel_area)) %>%
-  # ggplot(aes(x=mean_rel, y=max_rel, label=feature)) + geom_point() + 
-  # scale_y_log10() + scale_x_log10() +
-  # geom_vline(xintercept = 0.25) + geom_hline(yintercept = 0.5)
-  filter(mean_rel<0.25 & max_rel < 0.5) %>%
-  arrange(mean_rel) %>% 
-  add_row(feature="MS_NEG_1123") %>%
-  left_join(untarg_feats %>% select(feature:compound_name)) %>%
-  mutate(mzmed=ifelse(str_detect(feature, "NEG"), mzmed+1.007276*2, mzmed)) %>%
-  filter(feature!="MS_NEG_0768") %>% # Remove taurine duplicate
-  add_row(mzmed=c(198.0794, 184.0637), rtmed=c(5, 7))
-# ms11_offenders$mzmed %>% sort() %>% `names<-`(., round(., digits=3)) %>% dist(diag = TRUE) %>% round(digits=4)
-stans <- read_csv("https://raw.githubusercontent.com/IngallsLabUW/Ingalls_Standards/master/Ingalls_Lab_Standards.csv") %>%
-  filter(Column=="HILIC" & z>0) %>%
-  select(cmpd_name=Compound_Name, formula=Empirical_Formula) %>%
-  group_by(formula) %>%
-  summarise(cmpd_name=paste(cmpd_name, collapse = ", "))
-base_mz <- 85.089149355+1.007276
-iter_mzs <- expand_grid(CH2=0:6, O=2:6) %>%
-  mutate(mz=base_mz+CH2*14.01565+O*15.994915) %>%
-  mutate(mz=round(mz, digits=5)) %>%
-  mutate(cmd=paste0("mz%between%pmppm(", mz, ", 10)")) %>%
-  mutate(formula=paste0("C", 5+CH2, "H", 11+CH2*2, "NO", O)) %>%
-  left_join(stans) %>%
-  mutate(label=ifelse(is.na(cmpd_name), formula, paste0(formula, " (", cmpd_name, ")")))
-ms11_df <- c("C11H23NO6", "C11H23NO6", "C9H19NO5", "C8H17NO5", "C6H13NO6", "C8H17NO4") %>%
-  data.frame(formula=., rtmed=c(8.5, 10, 8.5, 10.2, 5.2, 9.2)) %>%
-  left_join(iter_mzs) %>%
-  select(mzmed=mz, rtmed, formula, cmpd_name) %>%
-  bind_rows(ms11_offenders) %>%
-  arrange(mzmed) %>%
-  select(mzmed, rtmed, formula, compound_name, feature) %>%
-  mutate(formula=case_when(
-    compound_name=="Taurine"~"C2H8NO3S",
-    feature=="MS_POS_1050"~"C7H16NO2",
-    compound_name=="Hydroxyisoleucine"~"C6H14NO3",
-    compound_name=="N-Acetyltaurine"~"C4H10NO4S",
-    feature=="MS_POS_1340"~"C5H14NO3S",
-    feature=="MS_POS_1419"~"C8H18NO3",
-    mzmed%between%pmppm(184.0637, 10)~"C5H14NO4S",
-    mzmed%between%pmppm(198.0794, 10)~"C6H16NO4S",
-    compound_name=="Dexpanthenol"~"C9H20NO4",
-    TRUE~formula
-  )) %>%
-  mutate(compound_name=case_when(
-    compound_name=="(3-Carboxypropyl)trimethylammonium?"~NA,
-    compound_name=="Hydroxyisoleucine"~NA,
-    compound_name=="Dexpanthenol"~NA,
-    formula=="C5H14NO4S"~"Choline sulfate",
-    formula=="C5H14NO3S"~"Taurine betaine?",
-    TRUE~compound_name
-  )) %>%
-  mutate(rtmed=round(rtmed, 1)) %>%
-  mutate(exact_mass=str_replace(formula, "N", "N1")) %>%
-  mutate(exact_mass=str_replace(exact_mass, "S", "S1")) %>%
-  separate(exact_mass, into = c("drop", "C", "H", "N", "O", "S"), sep = "[A-Z]") %>%
-  mutate(across(C:S, as.numeric)) %>%
-  mutate(S=ifelse(is.na(S), 0, 1)) %>%
-  mutate(calc_mass=C*12+H*1.007825+N*14.003074+O*15.994915+S*31.972072) %>%
-  mutate(calc_mass=ifelse(abs(calc_mass-mzmed)>1, calc_mass+1.007276, calc_mass)) %>%
-  mutate(compound_name=ifelse(is.na(compound_name), "", compound_name)) %>%
-  select(-feature)
-
-library(arrow)
-dset <- open_dataset("../tmzMLs/MT_all_file_arr_pqds/")
-arrow_data <- ms11_df %>%
-  pmap(function(...){
-    row_data <- list(...)
-    dset %>%
-      filter(mz%between%pmppm(row_data$mzmed, 20)) %>%
-      collect() %>%
-      mutate(formula=row_data$formula)
-  })
-Nacetyl_row <- ms11_df %>% filter(formula=="C4H10NO4S")
-neg_msdata <- grabMSdata(list.files("../tmzMLs/neg", pattern = "180821_Smp", full.names = TRUE), verbosity = 0)
-Nacetyldata <- neg_msdata$MS1[mz%between%pmppm(Nacetyl_row$mzmed-1.007276*2, 10)] %>%
-  mutate(formula=Nacetyl_row$formula) %>%
-  mutate(filename=str_replace(filename, "tmzML", "mzML"))
-peak_lims <- arrow_data %>%
-  bind_rows() %>%
-  filter(str_detect(filename, "Smp")) %>%
-  bind_rows(Nacetyldata) %>%
-  distinct() %>%
-  left_join(ms11_df) %>%
-  filter(rt>rtmed-1 & rt<rtmed+1)
-
-ms11_offenders_gp <- peak_lims %>%
-  mutate(formula=fct_inorder(formula)) %>%
-  ggplot() +
-  geom_line(aes(x=rt, y=int, group=filename, color=filename=="180821_Smp_MS11C215m_A.mzML")) +
-  facet_wrap(~formula, ncol=3, scales="free") +
-  scale_color_manual(name="Anomalous file", breaks = c(TRUE, FALSE), values=c("red", "black")) +
-  labs(x="Retention time (min)", y="Intensity") +
-  theme(legend.position="inside", legend.position.inside = c(0.85, 0.06)) +
-  guides(color=guide_legend(override.aes=list(lwd=1.5)))
-ggsave("ms11_quirky_cmpds.png", ms11_offenders_gp, device = "png", path = "supplement",
-       width = 6.875, height = 5, units = "in", dpi = 300)
-
-mass_diffs <- peak_lims %>%
-  group_by(formula, rtmed, mzmed, calc_mass) %>%
-  summarise(mean_mz=weighted.mean(mz, int)) %>%
-  mutate(calc_mass=ifelse(formula=="C4H10NO4S", calc_mass-2*1.007276, calc_mass)) %>%
-  mutate(delta_ppm=round((mean_mz-calc_mass)*1e6/mzmed, 2)) %>%
-  select(-calc_mass)
-final_ms11_df <- ms11_df %>%
-  left_join(mass_diffs) %>%
-  select(`Median m/z`=mzmed, `Median RT (min)`=rtmed, `Putative formula`=formula,
-         `Delta m/z (ppm)`=delta_ppm, `Compound name`=compound_name)
-# knitr::kable(final_ms11_df)
-```
-
-| Median m/z| Median RT (min)|Putative formula | Delta m/z (ppm)|Compound name    |
-|----------:|---------------:|:----------------|---------------:|:----------------|
-|   126.0221|            11.1|C2H8NO3S         |           -2.99|Taurine          |
-|   146.1176|             8.5|C7H16NO2         |           -3.58|                 |
-|   148.0968|             8.5|C6H14NO3         |           -3.79|                 |
-|   168.0317|             4.9|C4H10NO4S        |           32.46|N-Acetyltaurine  |
-|   168.0689|             8.5|C5H14NO3S        |           -3.65|Taurine betaine? |
-|   176.1282|             8.1|C8H18NO3         |           -2.86|                 |
-|   184.0637|             7.0|C5H14NO4S        |           -2.95|Choline sulfate  |
-|   192.1230|             9.2|C8H17NO4         |            0.22|                 |
-|   196.0816|             5.2|C6H13NO6         |            7.40|                 |
-|   198.0794|             5.0|C6H16NO4S        |           -2.56|                 |
-|   206.1388|             7.1|C9H20NO4         |           -2.04|                 |
-|   208.1180|            10.2|C8H17NO5         |            0.58|                 |
-|   222.1336|             8.5|C9H19NO5         |            0.31|                 |
-|   266.1598|             8.5|C11H23NO6        |            5.71|                 |
-|   266.1598|            10.0|C11H23NO6        |            4.45|                 |
-
-*Table 1: Detected mass features that were enriched in the 15 meter surface sample from MESO-SCOPE Station 11.
-RT = retention time in minutes.
-Taurine betaine is followed by a ?
-to denote that this identification is putative.
-Mass difference between the measured m/z and putative formula is reported in parts-per-million.*
-
-```{r trichobloomsamples comparison, eval=FALSE}
-library(RaMS)
-trichodata <- list.files("../mzMLs/tricho_pos", pattern=".mzML", full.names = TRUE) %>%
-  grabMSdata(verbosity = 0)
-tricho_eics <- ms11_df %>%
-  pmap(function(...){
-  row_data <- list(...)
-  peak_data <- trichodata$MS1[mz%between%pmppm(row_data$mzmed, 10)]
-  cbind(peak_data, formula=row_data$formula)
-}) %>%
-  bind_rows()
-tricho_eics %>%
-  filter(rt%between%c(2, 15)) %>%
-  arrange(mz) %>%
-  mutate(formula=fct_inorder(formula)) %>%
-  ggplot() +
-  geom_line(aes(x=rt, y=int, group=filename)) +
-  facet_wrap(~formula, scales="free", ncol=2) +
-  geom_vline(aes(xintercept = rtmed), color="red", data=ms11_df)
-trichodata$MS1[mz%between%pmppm(162.1125, 10)] %>%
-  mutate(samp_type=str_extract(filename, "Smp|Poo|Mix1|Mix2|Mat")) %>%
-  filter(samp_type!="Mat") %>%
-  ggplot() +
-  geom_line(aes(x=rt, y=int, group=filename, color=samp_type)) +
-  ggtitle("Homoserine betaine in Tricho bloom samples")
-```
-
-We explored the possibility that a nearby *Trichodesmium* bloom was responsible for the unusual composition of this one sample by analyzing additional samples taken directly from the bloom to see whether compounds abundant in hand-picked *Trichodesmium* colonies were also abundant in the Station 11 sample.
-After searching for the 15 mass features above in the environmental sample, 7 had good-quality peaks when viewed in the bloom samples as extracted ion chromatograms, though only 4 of these matched closely in retention time.
-We also performed a search of the untargeted data for homoserine betaine, a known *T.
-erythraeum* (strain IMS101) osmolyte [@Pade2016].
-Although this compound was not run in water alongside the sample to confirm a retention time match due to a lack of authentic standards, we identified three mass features in the 162.1125 *m/z* window with retention times of 7.1, 8.4, and 9.9 minutes.
-The feature at RT 9.9 matched our carnitine standard's retention time, leaving the two earlier peaks with putative identifications of threonine betaine at 7.1 and homoserine betaine at 8.4 given the elution order of threonine before homoserine in our authentic standards.
-This order is the reverse of that in @Pade2016, which showed homoserine betaine eluting prior to carnitine.
-This elution order is expected given our use of a normal phase HILIC column and their use of a reversed phase Hypercarb column.
-They did not report an additional peak detected at this mass, leaving our putative identification of the homoserine and threonine betaines as low-confidence.
-None of the three 162.1125 features showed the same degree of enrichment in the anomalous sample as taurine or the other 13 compounds listed in Table 1, though the anomalous sample contained the largest, second-largest, and third-largest peak for carnitine, threonine betaine, and homoserine betaine, respectively.
-These results are fairly inconclusive but we are inclined to believe that a *Trichodesmium* colony is not likely to be the cause of the inflated values in this particular sample.
-
 ```{r homoserinebetaine search, eval=FALSE}
@@
 
-When analyzed separately as distinct depths, the 25 meter samples had higher variances explained by eddy (R$^2$=.26) and a lower p-value (0.013), while the DCM samples were no longer likely to be distinct between the two eddies (p-value = 0.171) (Figure 6, Supplemental Figure 5).
+When analyzed separately as distinct depths, the 25 meter samples had higher variances explained by eddy (R$^2$=.26) and a lower p-value (0.013), while the DCM samples were no longer likely to be distinct between the two eddies (p-value = 0.171) (Figure 6, Supplemental Figure 4).
 
@@
 
-## Supp fig 4: Chromatograms of MS11 compounds
-
-![](supplement/ms11_quirky_cmpds.png)
-
-*Supplemental Figure 4: Extracted ion chromatograms of features identified in Table 1, separated by their predicted molecular formula.
-An m/z window of 10 ppm and a retention time window of 2 minutes was used to extract the data from the mzML files.
-The anomalous sample taken from 15 meters at Station 11 is shown in red while all other samples are shown in black.*
-
-## Supp fig 5: HEE NMDS and PERMANOVA broken down by depth
+## Supp fig 4: HEE NMDS and PERMANOVA broken down by depth
 
@@
 
-*Supplemental Figure 5: Plots of the metabolome during the Hawaiian Eddy Experiment, broken down by depth.
+*Supplemental Figure 4: Plots of the metabolome during the Hawaiian Eddy Experiment, broken down by depth.
 The top row of plots are non-metric multidimensional scaling (NMDS) plots, in which points correspond to individual samples and have been colored by their corrected sea level height anomaly.

### 2024-10-21 manuscript/manuscript.Rmd
Adding note about blanks and their use

@@
 Compounds for which we had an authentic standard (Supplemental Table 2) were manually integrated using Skyline and MSDIAL with the standard mixes used to ensure the correct peak was integrated by matching retention time.
-Compounds that were visually determined to have poor peak quality in the samples were removed from the downstream analysis.
+Compounds were removed from the analysis if they were visually assessed to have poor peak quality in the samples or peak areas in the samples similar to those in the methodological blanks.
 Skyline was used for both the eddy center samples and the HEE cruise data while MSDIAL was used for the eddy transect data due to the larger number of files in that dataset.

### 2024-10-21 manuscript/manuscript.Rmd
Rephrasing sample extraction to be more clear

@@
 Isotope labeled extraction standards were also added during this step (Supplemental Table 1).
-The samples were then bead-beaten three times, followed by triplicate extraction of the aqueous layer and addition with fresh methanol/water mixture with additional bead-beatings in between.
-Samples were then dried down under ultrapure nitrogen gas and warmed using a Fisher-Scientific Reacti-Therm module.
-Dried aqueous fractions were reconstituted in 380 µL Optima-grade water and amended with 20 µL isotope-labeled injection standards (Supplemental Table 1) to measure the variability introduced by chromatography and ionization.
+The samples were then bead-beaten three times, followed by triplicate extraction of the aqueous layer into a separate vial after spontaneous separation and replacement with fresh methanol/water mixture using a glass Pasteur pipette with additional bead-beatings in between.
+The aqueous fraction was then dried down under ultrapure nitrogen gas and warmed using a Fisher-Scientific Reacti-Therm module.
+Dry samples were reconstituted in 380 µL Optima-grade water and amended with 20 µL isotope-labeled injection standards (Supplemental Table 1) to measure the variability introduced by chromatography and ionization.
 The reconstituted fraction was syringe-filtered into precombusted glass inserts in HPLC vials to remove any potential clogging material.
-Samples were then additionally diluted 1:1 with Optima-grade water to prevent overloading on the column and reduce salt effects.
+Samples were then additionally diluted 1:1 with Optima-grade water to prevent overloading on the column and to reduce salt effects.
 

### 2024-10-21 manuscript/manuscript.Rmd
Adding 0 in front of correlation coefficients

@@
 At the DCM, approximately 16% of the variation in the metabolome could be explained by sea level anomaly (SLA) alone (PERMANOVA, p = 0.003).
-Additionally, SLA was strongly correlated with the first principal component (PC) of the metabolome  with a Pearson's *r* of .615 and 27% of the variance explained by PC1 (Supplemental Figure 1A), indicating that this was one of the largest sources of variation in the dataset.
+Additionally, SLA was strongly correlated with the first principal component (PC) of the metabolome  with a Pearson's *r* of 0.615 and 27% of the variance explained by PC1 (Supplemental Figure 1A), indicating that this was one of the largest sources of variation in the dataset.
 Notably, the samples taken from the exact center of the cyclonic eddy at Station 12 were highly distinct and likely drove much of the explained variance.
-In the 175 meter samples, 12% of the variance was explained by SLA (PERMANOVA, p = 0.008) and SLA was highly correlated with the second PC of the dataset (r = .697, % variance explained by PC2 = 15.4, Supplemental Figure 1A), though the first PC did not seem to have any visible pattern with metadata and appeared to largely capture variation between biological triplicates (Supplemental Figure 1B).
+In the 175 meter samples, 12% of the variance was explained by SLA (PERMANOVA, p = 0.008) and SLA was highly correlated with the second PC of the dataset (*r* = 0.697, % variance explained by PC2 = 15.4, Supplemental Figure 1A), though the first PC did not seem to have any visible pattern with metadata and appeared to largely capture variation between biological triplicates (Supplemental Figure 1B).
 At 15 meters, SLA trends are less evident with a larger p-value (PERMANOVA, p=0.087) and a lower R$^2$ (0.09) (Figure 2).
@@
 The 175 meter samples, on the other hand, showed the largest metabolite abundances in the anticyclone while the cyclone had consistently lower median z-scores.
-This method highlighted the negative correlation between the average metabolite and SLA at the DCM ($r_{DCM}= -.716$) while the 15 meter and 175 meter samples showed the opposite trend ($r_{15m} = .211$, $r_{175m} = .830$).
+This method highlighted the negative correlation between the average metabolite and SLA at the DCM ($r_{DCM}= -0.716$) while the 15 meter and 175 meter samples showed the opposite trend ($r_{15m} = 0.211$, $r_{175m} = 0.830$).
 
@@
 This shift in metabolite abundance was largely driven by a corresponding shift in biomass.
-Particulate carbon and summed metabolite concentration were tightly correlated across all samples and depths ($\beta = 24.9\pm1.82$ (SE), Pearson's $r$ = .813, p-value < 0.001, Supplemental Figure 2).
+Particulate carbon and summed metabolite concentration were tightly correlated across all samples and depths ($\beta = 24.9\pm1.82$ (SE), Pearson's $r$ = 0.813, p-value < 0.001, Supplemental Figure 2).
 This overall trend was largely driven by higher values at 15 meters and the DCM than at depth, but when analyzed at each depth individually the general correspondence held (Supplemental Figure 2).
-The DCM and 175 meter samples had the stronger trends ($\beta_{DCM} = 22.6\pm5.53$, $r= .593$, p-value < 0.001; $\beta_{175m} = 23.9\pm7.36$, $r= .503$, p-value = 0.003) while 15 meter samples showed no significant relationship ($\beta_{15m} = 3.4\pm12.9$, $r= .049$, p-value = 0.792).
+The DCM and 175 meter samples had the stronger trends ($\beta_{DCM} = 22.6\pm5.53$, $r= 0.593$, p-value < 0.001; $\beta_{175m} = 23.9\pm7.36$, $r= 0.503$, p-value = 0.003) while 15 meter samples showed no significant relationship ($\beta_{15m} = 3.4\pm12.9$, $r= 0.049$, p-value = 0.792).
 This meant that metabolites contributed a relatively fixed fraction of carbon to the particulate pool with our 53 quantified metabolites representing between 1.5% and 5% of the carbon in the system (Figure 3).
-Particulate nitrogen was tightly correlated with particulate carbon ($r=.933$) but the fraction of particulate nitrogen represented by the quantified metabolites was slightly higher across the transect with contributions typically between 2% and 6%.
+Particulate nitrogen was tightly correlated with particulate carbon ($r=0.933$) but the fraction of particulate nitrogen represented by the quantified metabolites was slightly higher across the transect with contributions typically between 2% and 6%.
 
@@
 Homarine showed a very large and highly significant shift, representing about 2.5% of the total peak area in the anticyclone but 7-8% in the cyclone center.
-Trigonelline (N-methyl niacin), an isomer structurally very similar to homarine but biologically distinct, had a surprisingly strong correlation with homarine (Pearson's $r=.855$).
+Trigonelline (N-methyl niacin), an isomer structurally very similar to homarine but biologically distinct, had a surprisingly strong correlation with homarine (Pearson's $r=0.855$).
 Its peak areas were consistently around one third of homarine's but still large enough to have the second-largest shift in relative peak area of the significantly different compounds.
@@
 Here, we found that this ~40 meter depth range and the transition between the eddies explained similar amounts of variation in the data (R$^2_{depth}=0.253$, R$^2_{SLA}=0.238$) and were both highly significant factors (permutational p-values << 0.001, Figure 4).
-Additionally, both sample depth and SLA were correlated with the first principal component of the metabolite matrix ($r_{depth}=.766$, $r_{SLA}=.684$, fraction of variance explained by PC1 = 30.8%).
+Additionally, both sample depth and SLA were correlated with the first principal component of the metabolite matrix ($r_{depth}=0.766$, $r_{SLA}=0.684$, fraction of variance explained by PC1 = 30.8%).
 The cyclonic samples in particular show a much larger spread than the anticyclonic ones, indicating larger sample variability in the cyclone DCM relative to the anticyclone.
@@
 Cluster 2 had a distinct decrease in relative peak area with depth, while the other clusters had much less clear depth trends.
-Cluster 1 also showed a DCM maximum for the samples from the anticyclone, correlating well with flow cytometry counts of picoeukaryotes (Pearson's $r=.851$) that was not present in the cyclone ($r=.013$).
+Cluster 1 also showed a DCM maximum for the samples from the anticyclone, correlating well with flow cytometry counts of picoeukaryotes (Pearson's $r=0.851$) that was not present in the cyclone ($r=0.013$).
 
@@
 
-When analyzed separately as distinct depths, the 25 meter samples had higher variances explained by eddy (R$^2$=.26) and a lower p-value (0.013), while the DCM samples were no longer likely to be distinct between the two eddies (p-value = 0.171) (Figure 6, Supplemental Figure 4).
+When analyzed separately as distinct depths, the 25 meter samples had higher variances explained by eddy (R$^2$=0.26) and a lower p-value (0.013), while the DCM samples were no longer likely to be distinct between the two eddies (p-value = 0.171) (Figure 6, Supplemental Figure 4).
 

### 2024-10-21 manuscript/manuscript.Rmd
Homogenizing w previous graph by capitalizing SLA

@@
   annotate("label", x=-Inf, y=Inf, hjust=0, vjust=1, label.r = unit(0, "in"),
-           label=expression(atop(paste("R"[sla]^2, "=.090, p=0.017"),
+           label=expression(atop(paste("R"[SLA]^2, "=.090, p=0.017"),
                                  paste("R"[depth]^2, "=.251, p<0.001")))) +

### 2024-10-21 manuscript/manuscript.Rmd
Adding CTD data to supplemental tables

@@
 Two particulate carbon values were determined to be spurious, with concentrations 2-3 times higher than typical Station ALOHA values, and were instead estimated using a best-fit linear regression against beam attenuation.
+CTD data is available here in Supplemental Tables 1 and 2.
+
+```{r supplemental CTD data}
+wb <- createWorkbook(creator = "William Kumler, wkumler@uw.edu")
+addWorksheet(wb, sheetName = "Table 1 CTD Cast list")
+ctd_cast_list <- read_csv("../metadata/complete_ctddata.csv") %>%
+  mutate(cruise=ifelse(time<"2018-01-01", "MESOSCOPE", "HEE")) %>%
+  select(cruise, everything())
+writeData(x=ctd_cast_list, wb = wb, sheet = "Table 1 CTD Cast list")
+
+addWorksheet(wb, sheetName = "Table 2 CTD Cast data by bottle")
+ctd_cast_data <- read_csv("../metadata/all_core_gof.csv") %>%
+  mutate(cruise=ifelse(cruise=="FK", "HEE", "MESOSCOPE")) %>%
+  select(cruise, everything())
+writeData(x=ctd_cast_data, wb = wb, sheet = "Table 2 CTD Cast data by bottle")
+
+addWorksheet(wb, sheetName = "Table 3 Interpolated CTD data")
+writeData(x=filled_file_metadata, wb = wb, sheet = "Table 3 Interpolated CTD data")
+saveWorkbook(wb, file = "supplement/Supplemental Tables.xlsx", overwrite = TRUE)
+```
 
@@
 Briefly, filters were randomized and added to PTFE centrifuge tubes with a 1:1 mix of 100 µm and 400 µm silica beads, approximately 2 mL -20 °C Optima-grade dichloromethane, and approximately 3 mL -20 °C 1:1 methanol/water solution (both also Optima-grade).
-Isotope labeled extraction standards were also added during this step (Supplemental Table 1).
+Isotope labeled extraction standards were also added during this step (Supplemental Table 4).
 The samples were then bead-beaten three times, followed by triplicate extraction of the aqueous layer into a separate vial after spontaneous separation and replacement with fresh methanol/water mixture using a glass Pasteur pipette with additional bead-beatings in between.
 The aqueous fraction was then dried down under ultrapure nitrogen gas and warmed using a Fisher-Scientific Reacti-Therm module.
-Dry samples were reconstituted in 380 µL Optima-grade water and amended with 20 µL isotope-labeled injection standards (Supplemental Table 1) to measure the variability introduced by chromatography and ionization.
+Dry samples were reconstituted in 380 µL Optima-grade water and amended with 20 µL isotope-labeled injection standards (Supplemental Table 4) to measure the variability introduced by chromatography and ionization.
 The reconstituted fraction was syringe-filtered into precombusted glass inserts in HPLC vials to remove any potential clogging material.
@@
 A pooled sample was created by combining 20 µL of each sample into the same HPLC vial, and a 1:1 dilution with water created a half-strength pooled sample from that to assess matrix effects and obscuring variation [@Boysen2018].
-Also run alongside the environmental samples were samples of authentic standards split into two mixes (Supplemental Table 2).
+Also run alongside the environmental samples were samples of authentic standards split into two mixes (Supplemental Table 5).
 Standards were dissolved in Optima-grade water to help with identification and into an aliquot of the pooled sample to quantify matrix response factors.
@@
 
-Compounds for which we had an authentic standard (Supplemental Table 2) were manually integrated using Skyline and MSDIAL with the standard mixes used to ensure the correct peak was integrated by matching retention time.
+Compounds for which we had an authentic standard (Supplemental Table 5) were manually integrated using Skyline and MSDIAL with the standard mixes used to ensure the correct peak was integrated by matching retention time.
 Compounds were removed from the analysis if they were visually assessed to have poor peak quality in the samples or peak areas in the samples similar to those in the methodological blanks.
 Skyline was used for both the eddy center samples and the HEE cruise data while MSDIAL was used for the eddy transect data due to the larger number of files in that dataset.
-Raw peak areas were then normalized to their best-matched internal standard (Supplemental Table 1) per @Boysen2018 except that all compounds were normalized to their best match no matter how minimal the improvement.
+Raw peak areas were then normalized to their best-matched internal standard (Supplemental Table 4) per @Boysen2018 except that all compounds were normalized to their best match no matter how minimal the improvement.
 Normalized peaks were then calibrated via a single-point standard curve as determined uniquely for each compound from the authentic standard mixes and used to convert normalized peak area into moles per µL injected and scaled to the amount of seawater filtered to provide a final estimate of environmental concentration for each compound in moles per liter of seawater filtered.
@@
 
-wb <- createWorkbook(creator = "William Kumler, wkumler@uw.edu")
-addWorksheet(wb, sheetName = "Table 1 Internal Standards")
-writeData(x=IS_table, wb = wb, sheet = "Table 1 Internal Standards")
-
-addWorksheet(wb, sheetName = "Table 2 Authentic Standards")
-writeData(x=Auth_table, wb = wb, sheet = "Table 2 Authentic Standards")
+wb <- loadWorkbook("supplement/Supplemental Tables.xlsx")
+addWorksheet(wb, sheetName = "Table 4 Internal Standards")
+writeData(x=IS_table, wb = wb, sheet = "Table 4 Internal Standards")
+
+addWorksheet(wb, sheetName = "Table 5 Authentic Standards")
+writeData(x=Auth_table, wb = wb, sheet = "Table 5 Authentic Standards")
 

### 2024-10-22 manuscript/manuscript.Rmd
Adding version information and citations

@@
 The instrument was mass calibrated at 200 *m/z* before each run began and once again during the eddy transect sample set to ensure calibrations were always performed within 3-4 days.
-All data files were then converted to an open-source mzML format and vendor centroided via Proteowizard’s msConvert tool.
+All data files were then converted to an open-source mzML format and vendor centroided via Proteowizard’s msConvert tool (version 3.0.19297, @Chambers2012).
 All mzML files have been uploaded to Metabolomics Workbench under Project ID PR001738.
@@
 
-Compounds for which we had an authentic standard (Supplemental Table 5) were manually integrated using Skyline and MSDIAL with the standard mixes used to ensure the correct peak was integrated by matching retention time.
+Compounds for which we had an authentic standard (Supplemental Table 5) were manually integrated using Skyline (versions 4.1 and 21.2.0.470, @Adams2020) and MSDIAL (version 4.36, @Tsugawa2015) with the standard mixes used to ensure the correct peak was integrated by matching retention time.
 Compounds were removed from the analysis if they were visually assessed to have poor peak quality in the samples or peak areas in the samples similar to those in the methodological blanks.
@@
 Samples were also processed via the untargeted workflow detailed in @Kumler2023.
-In short, XCMS [@R-xcms] was used for feature detection with CentWave peakpicking, Obiwarp alignment, and peak density correspondence/grouping.
+In short, XCMS (version 3.10.0, @R-xcms) was used for feature detection with CentWave peakpicking, Obiwarp alignment, and peak density correspondence/grouping.
 This was followed by manual inspection of the extracted ion chromatogram for each mass feature by an expert for quality with only those peaks classified as "Good" used in the analysis here.
@@
 
-All statistics were run in R version 4.3.3.
-We used multivariate statistics provided by the vegan package [@R-vegan] to assess the overall impact of various metrics across the metabolome followed by univariate statistics to investigate individually significant responses.
+All statistics were run in R version 4.4.0.
+We used multivariate statistics provided by the vegan package (version 2.6-6.1, @R-vegan) to assess the overall impact of various metrics across the metabolome followed by univariate statistics to investigate individually significant responses.
 Non-parametric tests were used when data were unlikely to obey parametric assumptions and permutational statistics were preferred wherever possible.

### 2024-10-22 manuscript/response_to_reviewers.txt
First pass at responding to reviewer comments is done

> Reviewer 1 | 04 Sep 2024 | 01:18
> #1
> This study examines the impact of mesoscale eddies on the metabolome and composition 
> of particulate organic matter in the North Pacific Subtropical Gyre. It fills a gap 
> in previous research by exploring the molecular composition of particulate organic 
> matter. The use both of targeted and untargeted metabolomics mass spectrometry to 
> detect molecules sensitive to environmental perturbation. The study finds that many 
> metabolites track biomass trends. Metabolites linked to eukaryotic phytoplankton are 
> enriched at the deep chlorophyll maximum of cyclonic features, while degradation 
> products are enriched in anticyclones. Additionally, depth variability and vertical 
> displacement of isopycnal layer constrain the importance of mesoscale eddies in 
> shaping particulate matter composition. Based on these results, we can infer the 
> biological communities and biogeochemical processes driven by mesoscale eddies at 
> both spatial and temporal scales. Viewing this complex dynamic process from the 
> chemical molecular scale is a perspective that sets this article apart from other 
> traditional methods in examining the changes in this system.
This is an accurate assessment of the manuscript.

> However, I still have some questions.
> Currently, the process of the entire system seems to be predictable. Although the 
> author has introduced and discussed some gaps, especially from the perspective of 
> chemical composition, and this article provides detailed data and reliable analysis, 
> it is not prominent enough from the perspective of the entire system. Perhaps the 
> transformation from space to time and the use of chemical composition to observe the 
> aging process of vortices is a great highlight. It is very difficult to focus on 
> scale eddies and the ecosystem evolution driven by them using traditional methods. 
> If that's the case, the structure and writing style of the paper may need to be 
> adjusted.

> This article focuses on the differences in chemical composition to understand this 
> process, but I always want to connect it with specific biological processes, such 
> as structural changes in biological communities, primary production rates, 
> zooplankton grazing rates, or bacterial production and respiration rates etc. 
> The author also mentioned these. I don't know if the author can add more relevant 
> evidence or discussion to allow readers to connect with these specific biological 
> groups and/or processes. I know it's difficult to integrate these data together, 
> but relevant discussions are possible.

> We already know that mesoscale eddies are complex and dynamic multidimensional 
> processes. The three-dimensional space combined with the temporal evolution of 
> ecosystem structure and function. The changes in the density surface driven by 
> physics during this process lead to the coupling reaction of physics chemistry 
> biology, which is the main thread. In fact, many other physical biological coupling 
> processes are similar, such as upwelling, internal waves/tides. Using the author's 
> method, I think similar situations can also be seen in such processes.

> Specific suggestions,
> Table 1, the format is not appropriate
Table 1 has been removed entirely.

> Figures 4, and 6 can be placed together.
While we agree that the data presented in Figures 4 and 6 can be placed side-by-side
for ease of comparison, we believe that keeping these figures in their separate
locations within the manuscript allows for better narrative flow (since the HEE
results are not discussed until several sections later) and graphical design (as
proper visualization of NMDS requires that the x- and y- axes share units). We are
happy to combine the two plots if the editor deems it necessary or helpful.

> Reviewer 2 | 04 Oct 2024 | 12:20
> #1
> Kumler et al. report in their manuscript “Metabolites reflect variability introduced 
> by mesoscale eddies in the North Pacific Subtropical Gyre” on a very impressive 
> dataset, investigating the polar part of the meta-metabolome of a > 0.2 µm fraction 
> of the plankton. They investigate samples taken at the Meso-scope and HEE experiment 
> and discuss their data in light of the eddies found during sampling.
> It is clearly a high-quality study that opens new perspectives. Publication after 
> revision is highly recommended.
This is an accurate assessment of the manuscript.

> Sampling:
> The metabolome at different depths can be strongly affected by the sampling time. 
> Day / night variations are substantial, mainly due to diel vertical migration but 
> also to photosynthetic state of the phytoplankton. Discussing the results sampling 
> times during the transect and during sampling of the HEE should be specified and 
> statistical analysis on the impact (or lack of impact) on the data should be provided.

> Analytics:
> The high-quality analytics is appreciated. It should be verified if the variations 
> between biological replicates could be caused by the order of measurements in the 
> sequence. If >100 samples of 15 min are run in a row there could be a drift that would 
> be hidden because of the randomization. It would be easy to have a look at the outliers 
> (e.g. Fig. 2 lower panels) are measured early / late during the sequence.

> Blanks are mentioned but it is not shown how these were used for the data analysis.
We have added a note about how we used the methodological blanks to visually assess 
whether sample peak areas were above the background signal.

> In line 142 the method for the evaluation of the matrix effect is mentioned, but no 
> results are given.
The Boysen et al. (2018) paper details the method for identifying the magnitude of the 
matrix effect by using a full-strength and water-diluted pooled sample and we believe 
that the reader is best suited to the detailed analysis presented there than a
brief review of it in this manuscript. We have confirmed that references to
this method are cited immediately after its mention here.

> The experimental could be clearer in line 132 ff “The samples were then bead-beaten 
> three times, followed by triplicate extraction of the aqueous layer...” Was the 
> organic layer collected or was the aqueous layer collected (as suggested by the rest 
> of the sentence?). Was phase separation occurring spontaneously or was a 
> centrifugation required? Was separation done by pipetting?
We have added additional text to this line to clarify the requested points of confusion.

> Line 177: I would rather state “targeted analytical methods”. Metabolomics is by 
> definition rather untargeted.
While the exact definition of "targeted" vs "untargeted" metabolomics remains an 
active topic of debate, we believe that the methods used here are sufficiently
targeted (use of authentic standards and manual integration via Skyline/MSDIAL) to 
warrant the use of the term, especially in direct contrast to the automated/untargeted 
pipeline described immediately afterward and thus we have not made this change.

> Line 185: The data are discussed quantitatively but for some compounds no standard 
> was available. The claim that results were normalized “to the best match standard”. 
> How was this decided? Are the different response-factors in the MS considered?
As above, the Boysen et al. (2018) paper describes how internal standards are matched
to compounds for which no authentic standard was available based on the difference
in their response factor between a full-strength and half-strength pooled sample.
The only data presented quantitatively (i.e. as concentrations instead of relative
or absolute peak area) are those for which an authentic standard was available,
ensuring that each quantified compound's response factor was known.

> Chapter 4.1.1 It is strange to read an entire chapter on an outlier. It is really 
> one measurement that could be easily contaminated (e.g. one single zooplankton 
> organism on the filter would change everything) and I would not recommend to write 
> an entire chapter about the metabolome, risking to discuss artifacts and not data 
> important for the large picture. Is table 1 required.
We appreciate the reviewer's identification of this as an unhelpful distraction
from the main narrative and have removed chapter 4.1.1 entirely.

> Minor concerns:
> Could the CTD data be given in the supporting material, this would provide a view of 
> the eddies in a depth profiles.
The CTD data has been added to the Supplemental Tables in 3 forms - first, containing
a list of all the stations and their locations with the number of bottles fired. Second,
with all the associated metadata compiled from Barone 2022. Finally, the values used
in the manuscript's analysis after interpolation to the locations in space and time
from which metabolite samples were collected.

> Software: Give the version number of the software used (e.g. Proteowizard, msConvert, 
> XCMS...)
Done. Software versions for Proteowizard, Skyline, MSDIAL, R, XCMS, and vegan have been
added to the manuscript.

> Statistics: It would facilitate better interpretation if the statistical method used 
> for comparisons would be mentioned in the respective sections of results.

> Line 239 Typo r = 0.697
We have switched to using a 0 in front of Pearson's correlation coefficient throughout
the manuscript to follow modern conventions.

> Line 266: From the figures in the supporting material, it does not become evident that 
> Particulate caron and metabolite concentrations are connected. There could be nested 
> effects due to the selection of data points.

### 2024-10-22 manuscript/manuscript.Rmd
Confirming that analytical order did not cause outliers

@@
 Bottom panels depict the direction and magnitude of this effect by plotting the mean value of all z-scored metabolites across three biological triplicates in color with the raw values in black behind.*
+
+```{r confirm that analytical order does not affect conclusions, eval=FALSE}
+read_csv("../untargeted/IS_integrations/all_is.csv") %>%
+  select(filename=Replicate.Name, compound_name=Precursor.Ion.Name, area=Area) %>%
+  filter(str_detect(filename, "_Smp_")) %>%
+  mutate(filename=paste0(filename, ".mzML")) %>%
+  pivot_wider(names_from=compound_name, values_from = "area") %>%
+  mutate("Inj_vol"=ifelse(str_detect(filename, "Poo.*Half"), 0.5, 1)) %>%
+  pivot_longer(-filename, names_to = "IS_name", values_to = "IS_area") %>%
+  filter(IS_name!="Guanosine monophosphate, 15N5") %>%
+  left_join(filled_file_metadata %>% select(filename, timestamp, sla_corr, ms_run)) %>%
+  filter(ms_run=="MT") %>%
+  arrange(timestamp) %>%
+  drop_na() %>%
+  mutate(filename=fct_inorder(filename)) %>%
+  ggplot(aes(x=timestamp, y=IS_area, color=sla_corr)) +
+  geom_hline(yintercept = 0) +
+  geom_point() +
+  facet_wrap(~IS_name, scales="free_y") +
+  scale_fill_gradient2(
+    low = scales::muted("blue"), high = scales::muted("red"),  mid = "grey90",
+    aesthetics = c("color", "fill"),
+  ) +
+  theme_bw()
+
+read_csv("../untargeted/MS_pos_output/deaddiso_peaks.csv") %>%
+  filter(str_detect(filename, "_Smp_")) %>%
+  left_join(filled_file_metadata %>% distinct(filename, ms_run, station, depth, sla_corr),
+            by="filename") %>%
+  filter(ms_run=="MT") %>%
+  group_by(depth, feature) %>%
+  filter(sum(into==0)<n()*0.9) %>%
+  mutate(into=scale(into)[,1]) %>%
+  ungroup() %>%
+  group_by(filename, depth, station, sla_corr) %>%
+  summarize(avg_amt=mean(into)) %>%
+  ungroup() %>%
+  left_join(filled_file_metadata %>% select(filename, timestamp)) %>%
+  arrange(timestamp) %>%
+  mutate(filename=fct_inorder(filename)) %>%
+  ggplot(aes(x=timestamp, y=avg_amt, color=sla_corr)) +
+  geom_hline(yintercept = 0) +
+  geom_point() +
+  geom_smooth(method="loess", formula=y~x, color="black") +
+  scale_fill_gradient2(
+    low = scales::muted("blue"), high = scales::muted("red"),  mid = "grey90",
+    aesthetics = c("color", "fill"),
+  ) +
+  theme_bw()
+
+med_metab_by_timestamp <- read_csv("../untargeted/all_peaks.csv") %>%
+  filter(str_detect(filename, "_Smp_")) %>%
+  left_join(filled_file_metadata %>% distinct(filename, ms_run, station, depth, sla_corr),
+            by="filename") %>%
+  filter(ms_run=="MT") %>%
+  group_by(depth, feature) %>%
+  filter(sum(norm_area==0)<n()*0.9) %>%
+  mutate(norm_area=scale(norm_area)[,1]) %>%
+  ungroup() %>%
+  group_by(filename, depth, station, sla_corr) %>%
+  summarize(avg_amt=mean(norm_area)) %>%
+  ungroup() %>%
+  left_join(filled_file_metadata %>% select(filename, timestamp)) %>%
+  arrange(timestamp) %>%
+  mutate(filename=fct_inorder(filename))
+
+med_metab_by_timestamp %>%
+  ggplot(aes(x=timestamp, y=avg_amt, color=sla_corr)) +
+  geom_hline(yintercept = 0) +
+  geom_point() +
+  geom_smooth(method="loess", formula=y~x, color="black") +
+  scale_fill_gradient2(
+    low = scales::muted("blue"), high = scales::muted("red"),  mid = "grey90",
+    aesthetics = c("color", "fill"),
+  ) +
+  theme_bw()
+
+med_metab_by_timestamp %>%
+  mutate(station=factor(station, levels=as.character(4:14))) %>%
+  mutate(depth=factor(depth, levels=c("Surface", "DCM", "Deep"))) %>%
+  ggplot() +
+  geom_hline(yintercept = 0) +
+  geom_point(aes(x=station, y=avg_amt, color=timestamp), size=2) +
+  facet_wrap(~depth, nrow=1) +
+  theme_bw() +
+  theme(strip.background = element_blank(), strip.text = element_blank()) +
+  labs(x="Station", y="Median metabolite z-score")
+```
 

### 2024-10-23 manuscript/manuscript.Rmd
Adding zeroes in front of R2 values

@@
   annotate("label", x=-Inf, y=Inf, hjust=0, vjust=1, label.r = unit(0, "in"),
-           label=expression(atop(paste("R"[SLA]^2, "=.238"), paste("R"[depth]^2, "=.253")))) +
+           label=expression(atop(paste("R"[SLA]^2, "=0.238"), paste("R"[depth]^2, "=0.253")))) +
   annotate("label", x=-Inf, y=-Inf, label=multivar_analysis_MC$nmds_stress, 
@@
   annotate("label", x=-Inf, y=Inf, hjust=0, vjust=1, label.r = unit(0, "in"),
-           label=expression(atop(paste("R"[SLA]^2, "=.090, p=0.017"),
-                                 paste("R"[depth]^2, "=.251, p<0.001")))) +
+           label=expression(atop(paste("R"[SLA]^2, "=0.090, p=0.017"),
+                                 paste("R"[depth]^2, "=0.251, p<0.001")))) +
   guides(fill=guide_legend(title = NULL, override.aes = list(

### 2024-10-24 manuscript/response_to_reviewers.txt
First pass at responding to reviewer comments is done
Confirming that analytical order did not cause outliers
Adding discussion of diel effects
New Supplemental Figure
New stats analysis (adding +time_class) to PERMANOVA
New p-values and R^2
New fig. 2 (only difference should be R2 and p-vals)

@@
 results are not discussed until several sections later) and graphical design (as
-proper visualization of NMDS requires that the x- and y- axes share units). We are
-happy to combine the two plots if the editor deems it necessary or helpful.
+proper visualization of NMDS requires that the x- and y- axes share units).
 
@@
 > statistical analysis on the impact (or lack of impact) on the data should be provided.
+We have added the time of day as an explanatory factor to our analysis and reported the
+new values from the PERMANOVA in the text and figures, as well as adding a supplemental
+figure to visually represent some of the variance due to time of day. We do see 
+diel differences at the surface and DCM during the eddy transect and these have been
+reported, with the 175 meter differences more likely to be due to chance and very little
+diel difference in the HEE samples at all. We have additionally added a supplemental
+figure identifying a few select compounds for which diel effects are clearly visible
+(trehalose and sucrose) alongside other highly abundant compounds which show no diel
+effects.
 
@@
 > (e.g. Fig. 2 lower panels) are measured early / late during the sequence.
+We have double-checked that our normalization to internal standards removed any
+analytical drift and can confirm that while significant drift occurred over
+the course of the eddy transect sample run, the normalization removed this
+effect and is not a major factor in the outliers visible in Fig. 2.
 

### 2024-10-24 manuscript/manuscript.Rmd
Adding discussion of diel effects
New Supplemental Figure
New stats analysis (adding +time_class) to PERMANOVA
New p-values and R^2
New fig. 2 (only difference should be R2 and p-vals)

@@
 Both the MESO-SCOPE eddy center and HEE analyses treated SLA as a categorical variable.
+Where time of day was included as an explanatory factor in the PERMANOVA, it was binned into 6-hour increments starting from 3AM (e.g.
+Midday contains casts collected between 9AM and 3PM).
 
@@
 The largest differences in particulate matter composition were detected in the DCM and 175 meter samples, as shown by the NMDS in Figure 2.
-At the DCM, approximately 16% of the variation in the metabolome could be explained by sea level anomaly (SLA) alone (PERMANOVA, p = 0.003).
+
+At the DCM, approximately 14% of the variation in the metabolome could be explained by sea level anomaly (SLA) alone (PERMANOVA, p = 0.002).
 Additionally, SLA was strongly correlated with the first principal component (PC) of the metabolome  with a Pearson's *r* of 0.615 and 27% of the variance explained by PC1 (Supplemental Figure 1A), indicating that this was one of the largest sources of variation in the dataset.
 Notably, the samples taken from the exact center of the cyclonic eddy at Station 12 were highly distinct and likely drove much of the explained variance.
-In the 175 meter samples, 12% of the variance was explained by SLA (PERMANOVA, p = 0.008) and SLA was highly correlated with the second PC of the dataset (*r* = 0.697, % variance explained by PC2 = 15.4, Supplemental Figure 1A), though the first PC did not seem to have any visible pattern with metadata and appeared to largely capture variation between biological triplicates (Supplemental Figure 1B).
-At 15 meters, SLA trends are less evident with a larger p-value (PERMANOVA, p=0.087) and a lower R$^2$ (0.09) (Figure 2).
+
+In the 175 meter samples, 13% of the variance was explained by SLA (PERMANOVA, p = 0.004) and SLA was highly correlated with the second PC of the dataset (*r* = 0.697, % variance explained by PC2 = 15.4, Supplemental Figure 1A), though the first PC did not seem to have any visible pattern with metadata and appeared to largely capture variation between biological triplicates (Supplemental Figure 1B).
+
+At 15 meters, SLA trends are slightly less evident with a larger p-value (PERMANOVA, p=0.010) and a lower R$^2$ (0.12) (Figure 2).
+Time of day was also a significant factor in the PERMANOVA at the surface and DCM (p-values of 0.004 and 0.032, respectively) while the diel effects in the 175 meter samples were unsurprisingly much lower (p = 0.118).
+
+The significance of SLA as an explanatory factor was similar even when excluding time of day, with the surface seeing the largest change to an R$^2_{SLA}$ of 0.09 and a p-value of 0.087, though the fraction of the variance explained by SLA increased to 16% without time of day as a factor.
 
@@
   filter(ms_run=="MT") %>%
+  arrange(depth) %>%
   nest(data=-depth) %>%
@@
     adon_expl <- data.frame(filename=rownames(metab_mat)) %>%
-      left_join(filled_file_metadata %>% select(filename, sla_class), by="filename")
+      left_join(filled_file_metadata %>% select(filename, sla_class, time_class),
+                by="filename")
     vegan::adonis2(
-      metab_mat~sla_class, data = adon_expl, by = "margin", 
+      metab_mat~sla_class, 
+      # metab_mat~sla_class+time_class, 
+      data = adon_expl, by = "margin", 
       permutations = 999, method = "manhattan")
@@
   permutations = 999, method = "manhattan")
+# vegan::adonis2(
+#   metab_mat_fk~sla_class+depth+time_class, data = adon_expl, by = "margin",
+#   permutations = 999, method = "manhattan")
 ```
@@
 Despite this, we still saw the importance of sea level anomaly as a significant explanatory factor in the dataset (R$^2_{SLA} = 0.090$, p-value = 0.017) as well as the larger depth differences between the surface (25 meters) and DCM (110 - 120 meters) (R$^2_{depth} = 0.251$, p-value << 0.001).
+We saw no difference between the 6AM and 6PM samples (R$^2_{time} = 0.031$, p-value = 0.373) despite a few known compounds (trehalose and sucrose) showing large differences, likely due to a majority of the mass features demonstrating no diel effect (Supplemental Figure 4).
 
@@
 
-When analyzed separately as distinct depths, the 25 meter samples had higher variances explained by eddy (R$^2$=0.26) and a lower p-value (0.013), while the DCM samples were no longer likely to be distinct between the two eddies (p-value = 0.171) (Figure 6, Supplemental Figure 4).
+When analyzed separately as distinct depths, the 25 meter samples had higher variances explained by eddy (R$^2$=0.26) and a lower p-value (0.013), while the DCM samples were no longer likely to be distinct between the two eddies (p-value = 0.171) (Figure 6, Supplemental Figure 5).
 
@@
 In the samples taken along the eddy transect, the largest effect was detected in the deepest samples taken from 175 meters and from the DCM, with much less of a response at 25 meters.
-In the HEE samples, however, the opposite was true with a larger SLA response at 25 meters than that at the DCM.
+In the HEE samples, however, the opposite was true with a surprisingly larger SLA response at 25 meters than that at the DCM.
 In each case, the differences introduced by SLA were smaller than those of depth when compared directly, with even the high-resolution sampling around the DCM finding slightly more variance explained by the 40 meter difference in sampling depth than the 40 centimeter difference in SLA.
 This result agrees well with prior research by @Heal2021 and @Kumler2023 showing the large effect of sampling depth on the metabolome.
+Time of day was included as a significant factor in the MESO-SCOPE transect, with trehalose and sucrose in particular showing large differences between 6AM and 6PM as has been noted previously [@Muratore2022], though this effect was diluted at 175 meters.
 
@@
 
-## Supp fig 4: HEE NMDS and PERMANOVA broken down by depth
+## Supp fig 4: Boxplots of selected compounds by depth, time, and SLA
+
+```{r diel, include=FALSE}
+filled_file_metadata %>% 
+  filter(samp_type=="Smp") %>%
+  filter(ms_run=="MT") %>%
+  distinct(station, time_class) %>%
+  count(time_class)
+filled_file_metadata %>% 
+  filter(samp_type=="Smp") %>%
+  filter(ms_run=="MT") %>%
+  distinct(station, sla_class) %>%
+  count(sla_class)
+
+select_sla_time_boxplot_gp <- targ_peaks %>%
+  filter(compound_name%in%c("Sucrose", "Trehalose", "Guanine", "Glycine betaine", "L-Glutamine", "Ectoine")) %>%
+  mutate(compound_name=case_when(
+    compound_name=="Glycine betaine"~"Glycine\nbetaine",
+    compound_name=="L-Glutamine"~"Glutamine",
+    TRUE~compound_name
+  )) %>%
+  select(filename, compound_name, nM) %>%
+  left_join(filled_file_metadata) %>%
+  filter(samp_type=="Smp") %>%
+  filter(ms_run!="MC") %>%
+  select(filename, compound_name, cruise, depth, sla_class, time_class, nM) %>%
+  mutate(time_class=factor(time_class, levels=c("Morning", "Midday", "Evening", "Night"))) %>%
+  mutate(cruise=factor(cruise, levels=c("MS", "FK"), labels=c("MESO-SCOPE transect", "HEE"))) %>%
+  mutate(depth=factor(depth, levels=c("Surface", "DCM", "Deep"))) %>%
+  ggplot() +
+  geom_hline(yintercept = 0) +
+  geom_boxplot(aes(x=time_class, y=nM, color=sla_class)) +
+  facet_nested(compound_name~cruise+depth, scales="free", space = "free_x") +
+  labs(x=NULL, y="Metabolite concentration (nM)", color="SLA condition") +
+  scale_color_manual(breaks = c("Cyclone", "Neither", "Anticyclone"),
+                     values = c(scales::muted("blue"), "grey50", 
+                                scales::muted("red"))) +
+  theme_bw() +
+  theme(axis.text.x = element_text(angle=90, hjust=1, vjust=0.5))
+
+ggsave("diel_effect_boxplots.png", plot = select_sla_time_boxplot_gp, device = "png",
+       path = "supplement", width = 6.875, height = 6, units = "in", dpi = 300,
+       bg="white")
+```
+
+![](supplement/diel_effect_boxplots.png)
+
+*Supplemental Figure 4: Boxplots showing the concentrations of six selected compounds between cruises and depths as well as time and SLA classes.
+The abscissa is separated by the time category into which sampling was binned, referring to 6-hour increments starting from 3AM (e.g.
+Midday contains casts collected between 9AM and 3PM).
+Colors denote the sea level anomaly bins with a ±5cm cutoff.
+For the eddy transect, six stations are included in the anticyclone, three in the cyclone, and two were neither.
+There were two morning stations, 4 midday stations, 2 evening stations, and 3 night stations.
+Diel effects are clearly visible for sucrose and trehalose while other compounds show very little diel differences.
+Trehalose was not quantified during the MESO-SCOPE transect for lack of an authentic standard to quantify the response factor.*
+
+## Supp fig 5: HEE NMDS and PERMANOVA broken down by depth
 

### 2024-10-24 metadata/metadata_control.Rmd
Fixing time zones :upside_down_face:

--- 2023-12-01 metadata/metadata_control.Rmd
@@
     read_table(skip = 3, col_names = col_names, col_types=col_types) %>%
-    mutate(time=strptime(paste(month, day, year, time), format = "%b %d %Y %X")) %>%
+    mutate(time=strptime(paste(month, day, year, time), format = "%b %d %Y %X", tz="UTC")) %>%
     select(-month, -day, -year) %>%
@@
   )) %>%
-  mutate(time=as.POSIXct(time, origin="1970-01-01")) %>%
+  mutate(time=as.POSIXct(time, origin="1970-01-01", tz="UTC")) %>%
+  mutate(local_time=as.POSIXct(time, tz="HST")) %>%
   mutate(time_class=c("Night", "Morning", "Midday", "Evening")[
-    floor((as.numeric(format(time, "%H"))+3)/6)%%4+1
+    floor((as.numeric(format(local_time, "%H"))+3)/6)%%4+1
   ]) %>%

### 2024-10-24 manuscript/manuscript.Rmd
Finalizing p-values for diel discussion

@@
   filter(polarity=="pos") %>%
+  mutate(local_time=as.POSIXct(time, tz="HST")) %>%
   select(-polarity)
@@
 Midday contains casts collected between 9AM and 3PM).
+Marginal effects are reported for the PERMANOVA analyses to avoid influencing the results with the order of the terms in the model formula.
 
@@
 Notably, the samples taken from the exact center of the cyclonic eddy at Station 12 were highly distinct and likely drove much of the explained variance.
-
 In the 175 meter samples, 13% of the variance was explained by SLA (PERMANOVA, p = 0.004) and SLA was highly correlated with the second PC of the dataset (*r* = 0.697, % variance explained by PC2 = 15.4, Supplemental Figure 1A), though the first PC did not seem to have any visible pattern with metadata and appeared to largely capture variation between biological triplicates (Supplemental Figure 1B).
-
-At 15 meters, SLA trends are slightly less evident with a larger p-value (PERMANOVA, p=0.010) and a lower R$^2$ (0.12) (Figure 2).
-Time of day was also a significant factor in the PERMANOVA at the surface and DCM (p-values of 0.004 and 0.032, respectively) while the diel effects in the 175 meter samples were unsurprisingly much lower (p = 0.118).
-
-The significance of SLA as an explanatory factor was similar even when excluding time of day, with the surface seeing the largest change to an R$^2_{SLA}$ of 0.09 and a p-value of 0.087, though the fraction of the variance explained by SLA increased to 16% without time of day as a factor.
+At 15 meters, SLA trends are slightly less evident with a larger p-value (PERMANOVA, p=0.029) and a lower R$^2$ (0.10) (Figure 2).
+Time of day was also a significant factor in the PERMANOVA at the surface and DCM (p-values of 0.018 and 0.033, respectively) while the diel effects in the 175 meter samples were unsurprisingly much lower (p = 0.140).
+The significance of SLA as an explanatory factor was similar even when excluding time of day from the model expression (R$^2_{SLA}$ and p-values within permutation error).
 
@@
     vegan::adonis2(
-      metab_mat~sla_class, 
-      # metab_mat~sla_class+time_class, 
+      # metab_mat~sla_class,
+      metab_mat~sla_class+time_class,
       data = adon_expl, by = "margin", 
@@
 adon_expl <- data.frame(filename=rownames(metab_mat_fk)) %>%
-  left_join(filled_file_metadata %>% select(filename, sla_class, depth), by="filename")
+  left_join(filled_file_metadata %>% select(filename, sla_class, time_class, depth), by="filename")
 adon_fk <- vegan::adonis2(
@@
 Despite this, we still saw the importance of sea level anomaly as a significant explanatory factor in the dataset (R$^2_{SLA} = 0.090$, p-value = 0.017) as well as the larger depth differences between the surface (25 meters) and DCM (110 - 120 meters) (R$^2_{depth} = 0.251$, p-value << 0.001).
-We saw no difference between the 6AM and 6PM samples (R$^2_{time} = 0.031$, p-value = 0.373) despite a few known compounds (trehalose and sucrose) showing large differences, likely due to a majority of the mass features demonstrating no diel effect (Supplemental Figure 4).
+We saw no difference between the 6AM and 6PM samples (R$^2_{time} = 0.031$, p-value = 0.367) despite a few known compounds (trehalose and sucrose) showing large differences, likely due to a majority of the mass features demonstrating no diel effect (Supplemental Figure 4).
 

### 2024-10-25 manuscript/manuscript.Rmd
Adding statistical methods used

@@
 Marginal effects are reported for the PERMANOVA analyses to avoid influencing the results with the order of the terms in the model formula.
+Total sample size (*n*) for the eddy transect was 33 per depth (3 biological triplicates at each of 11 stations), while the eddy centers had an *n* of 15 per eddy (3 biological replicates at 5 depths) and the HEE data had an *n* of 24 (3 biological replicates at 4 stations (2 in the cyclone and 2 in the anticyclone) at 2 depths each).
 
@@
 The HEE data was characterized by high inter-replicate variability relative to the samples from the MESO-SCOPE cruise, with 25 meter samples in particular highly variable in multivariate space (Figure 6).
-Despite this, we still saw the importance of sea level anomaly as a significant explanatory factor in the dataset (R$^2_{SLA} = 0.090$, p-value = 0.017) as well as the larger depth differences between the surface (25 meters) and DCM (110 - 120 meters) (R$^2_{depth} = 0.251$, p-value << 0.001).
-We saw no difference between the 6AM and 6PM samples (R$^2_{time} = 0.031$, p-value = 0.367) despite a few known compounds (trehalose and sucrose) showing large differences, likely due to a majority of the mass features demonstrating no diel effect (Supplemental Figure 4).
+Despite this, we still saw the importance of sea level anomaly as a significant explanatory factor in the dataset (PERMANOVA R$^2_{SLA} = 0.090$, p-value = 0.017) as well as the larger depth differences between the surface (25 meters) and DCM (110 - 120 meters) (PERMANOVA R$^2_{depth} = 0.251$, p-value << 0.001).
+We saw no difference between the 6AM and 6PM samples (PERMANOVA R$^2_{time} = 0.031$, p-value = 0.367) despite a few known compounds (trehalose and sucrose) showing large differences, likely due to a majority of the mass features demonstrating no diel effect (Supplemental Figure 4).
 
@@
 
-When analyzed separately as distinct depths, the 25 meter samples had higher variances explained by eddy (R$^2$=0.26) and a lower p-value (0.013), while the DCM samples were no longer likely to be distinct between the two eddies (p-value = 0.171) (Figure 6, Supplemental Figure 5).
+When analyzed separately as distinct depths, the 25 meter samples had higher variances explained by eddy and a lower p-value (PERMANOVA, R$^2$=0.26, p-value=0.013), while the DCM samples were no longer likely to be distinct between the two eddies (PERMANOVA p-value = 0.171) (Figure 6, Supplemental Figure 5).
 

### 2024-10-25 manuscript/response_to_reviewers.txt
Adding statistical methods used
More stats details

@@
 or absolute peak area) are those for which an authentic standard was available,
-ensuring that each quantified compound's response factor was known.
+ensuring that each quantified compound's response factor was known. For untargeted
+molecular features for which no authentic standard was available, data is only
+discussed explicitly in peak area space or as a relative deviation between stations/
+depths. We have also added links to the Github where the R code used to perform
+these matching and normalization steps is available.
 
@@
 > for comparisons would be mentioned in the respective sections of results.
+We have tried to add statistical methods used for comparisons in each section of the
+results where applicable in this revision, largely annotating the places in which 
+R-squared and p-values were obtained from PERMANOVAs. We have also added the sample
+size to the Statistics section of the Results.
 

### 2024-10-25 manuscript/manuscript.Rmd
More stats details

@@
 Compounds for which we had an authentic standard (Supplemental Table 5) were manually integrated using Skyline (versions 4.1 and 21.2.0.470, @Adams2020) and MSDIAL (version 4.36, @Tsugawa2015) with the standard mixes used to ensure the correct peak was integrated by matching retention time.
-Compounds were removed from the analysis if they were visually assessed to have poor peak quality in the samples or peak areas in the samples similar to those in the methodological blanks.
+Compounds were removed from the analysis if they were visually assessed to have poor peak quality in the samples or peak areas in the methodological blanks similar to those in the samples.
 Skyline was used for both the eddy center samples and the HEE cruise data while MSDIAL was used for the eddy transect data due to the larger number of files in that dataset.
 Raw peak areas were then normalized to their best-matched internal standard (Supplemental Table 4) per @Boysen2018 except that all compounds were normalized to their best match no matter how minimal the improvement.
+While significant analytical drift was detected during the longer eddy transect sample run, the use of internal standards was able to negate the instrument's general increase in sensitivity.
 Normalized peaks were then calibrated via a single-point standard curve as determined uniquely for each compound from the authentic standard mixes and used to convert normalized peak area into moles per µL injected and scaled to the amount of seawater filtered to provide a final estimate of environmental concentration for each compound in moles per liter of seawater filtered.
+Scripts used for this analysis are available at https://github.com/wkumler/MesoscopeMetabolomicsManuscript/tree/master/untargeted.
 
@@
 Features were then matched with and normalized to an internal standard (Supplemental Table 1) as described above.
+Scripts used for this analysis are available at https://github.com/wkumler/MesoscopeMetabolomicsManuscript/tree/master/untargeted.
 
@@
 This shift in metabolite abundance was largely driven by a corresponding shift in biomass.
-Particulate carbon and summed metabolite concentration were tightly correlated across all samples and depths ($\beta = 24.9\pm1.82$ (SE), Pearson's $r$ = 0.813, p-value < 0.001, Supplemental Figure 2).
+Particulate carbon and summed metabolite concentration were tightly correlated across all samples and depths (Type I linear regression $\beta = 24.9\pm1.82$ (SE), Pearson's $r$ = 0.813, p-value < 0.001, Supplemental Figure 2).
 This overall trend was largely driven by higher values at 15 meters and the DCM than at depth, but when analyzed at each depth individually the general correspondence held (Supplemental Figure 2).

### 2024-10-25 manuscript/manuscript.Rmd
Expanding upon the use of IS and including links to github

@@
 This shift in metabolite abundance was largely driven by a corresponding shift in biomass.
-Particulate carbon and summed metabolite concentration were tightly correlated across all samples and depths (Type I linear regression $\beta = 24.9\pm1.82$ (SE), Pearson's $r$ = 0.813, p-value < 0.001, Supplemental Figure 2).
+Particulate carbon and summed metabolite concentration were tightly correlated across all samples and depths ($\beta = 24.9\pm1.82$ (SE), Pearson's $r$ = 0.813, p-value < 0.001, Supplemental Figure 2).
 This overall trend was largely driven by higher values at 15 meters and the DCM than at depth, but when analyzed at each depth individually the general correspondence held (Supplemental Figure 2).

### 2024-10-29 manuscript/manuscript.Rmd
Adding DMS story
Adding phosphorus supplemental figure and discussion

--- 2024-10-25 manuscript/manuscript.Rmd
@@
 library(openxlsx)
+library(arrow)
 
@@
 
+We also measured several phosphorus-containing compounds and their non-phosphorus equivalents (Supplemental Figure 4A).
+Phosphocholine is the polar headgroup for phosphatidylcholine, a class of intact polar diacylglycerols that has been previously described at this study site as showing phosphorus stress relief among eukaryotic organisms in response to the upwelling of phosphate-rich deep sea water [@Bent2024].
+We measured this metabolite as well as choline (the non-phosphate containing equivalent) and glycerophosphocholine (GPC, which contains an additional glycerol group) with the expectation that phosphorus stress would result in higher choline:phosphocholine and choline:GPC ratios, as shown in Seelen et al.
+(in review).
+However, we saw no differences in this ratio across the eddy transect after accounting for depth effects (Supplemental Figure 4C).
+Additionally, choline:GPC ratios were highest at 175m where PO$_4^{3-}$ concentrations are highest, indicating that this metric may not accurately measure P-stress.
+Other phosphorus-containing lipid headgroups glycerophosphoglycerol and glycerophosphoethanolamine also showed no trend across the eddy at a given depth and instead showed trends functionally identical to the average metabolite in Figure 2 (Supplemental Figure 4B).
+
 ## Particulate organic matter compositional shifts across eddy transect
@@
 Despite this, we still saw the importance of sea level anomaly as a significant explanatory factor in the dataset (PERMANOVA R$^2_{SLA} = 0.090$, p-value = 0.017) as well as the larger depth differences between the surface (25 meters) and DCM (110 - 120 meters) (PERMANOVA R$^2_{depth} = 0.251$, p-value << 0.001).
-We saw no difference between the 6AM and 6PM samples (PERMANOVA R$^2_{time} = 0.031$, p-value = 0.367) despite a few known compounds (trehalose and sucrose) showing large differences, likely due to a majority of the mass features demonstrating no diel effect (Supplemental Figure 4).
+We saw no difference between the 6AM and 6PM samples (PERMANOVA R$^2_{time} = 0.031$, p-value = 0.367) despite a few known compounds (trehalose and sucrose) showing large differences, likely due to a majority of the mass features demonstrating no diel effect (Supplemental Figure 5).
 
@@
 
-When analyzed separately as distinct depths, the 25 meter samples had higher variances explained by eddy and a lower p-value (PERMANOVA, R$^2$=0.26, p-value=0.013), while the DCM samples were no longer likely to be distinct between the two eddies (PERMANOVA p-value = 0.171) (Figure 6, Supplemental Figure 5).
+When analyzed separately as distinct depths, the 25 meter samples had higher variances explained by eddy and a lower p-value (PERMANOVA, R$^2$=0.26, p-value=0.013), while the DCM samples were no longer likely to be distinct between the two eddies (PERMANOVA p-value = 0.171) (Figure 6, Supplemental Figure 6).
 
@@
 This compound was enriched in the cyclone of the DCM in both the MESO-SCOPE and HEE cruises along with its precursor taurine, validating earlier results from @Barone2022 that showed strong enrichment of eukaryotic phytoplankton at the DCM and *Pseudo-nitzschia* in particular.
+Additionally, the abundance other sulfur-containing compounds (e.g.
+dimethylsulfonioacetate and dimethylsulfoniopropionate) in the cyclone may indicate links between the formation of eukaryote-rich cyclonic features and the production of dimethylsulfide, a climate-active volatile gas [@Moran2019].
 
@@
 
-*Supplemental Figure 3: Stacked barplots of known metabolite concentration as shown in main text Figure 3 but with individual triplicates shown.
+*Supplemental Figure 4: Stacked barplots of known metabolite concentration as shown in main text Figure 3 but with individual triplicates shown.
 The same data is shown in A) as B) but B) is rendered in relative contribution space instead of absolute concentration.
@@
 
-## Supp fig 4: Boxplots of selected compounds by depth, time, and SLA
+## Supp fig 4: Choline, phosphocholine, and glycerophosphocholine ratios
+
+```{r}
+gp_df <- data.frame(
+  compound_name=c("Ethanolamine", "Glycerol", "Choline"),
+  mz=c(62.060589, 93.05462, 104.107539)
+) %>%
+  expand_grid(data.frame(
+    addition=c("", "Phospho", "Glycerophospho"),
+    mz_diff=c(0, 79.96633, 74.03678+79.96633)
+  )) %>%
+  mutate(base_compound=compound_name) %>%
+  mutate(compound_name=ifelse(addition=="", compound_name, 
+                              paste0(addition, tolower(compound_name))),
+         mz=mz+mz_diff) %>%
+  arrange(mz) %>%
+  mutate(compound_name=fct_inorder(compound_name)) %>%
+  mutate(addition=fct_inorder(addition))
+gp_w_eics <- gp_df %>%
+  mutate(eic=map(mz, function(mz_i){
+    open_dataset("~/My Drive/AllMeso/tmzMLs/MT_all_file_arr_pqds") %>%
+      filter(mz%between%pmppm(mz_i, 10)) %>%
+      collect()
+  }))
+smp_eics <- gp_w_eics %>% 
+  select(-mz) %>% 
+  unnest(eic) %>% 
+  filter(str_detect(filename, "Smp")) %>%
+  left_join(filled_file_metadata)
+all_cmpds <- tribble(
+  ~compound_name, ~rt_start, ~rt_end,
+  "Choline", 10.6, 11.2,
+  "Phosphocholine", 11.8, 13.2,
+  "Glycerophosphocholine", 10.8, 11.5,
+  "Glycerophosphoethanolamine", 11, 12.5,
+  "Glycerophosphoglycerol", 10.2, 11.2
+) %>%
+  mutate(compound_name=fct_inorder(compound_name)) %>%
+  left_join(smp_eics, by=join_by(compound_name, between(y$rt, x$rt_start, x$rt_end))) %>%
+  left_join(filled_file_metadata) %>%
+  mutate(depth=factor(depth, levels=c("Surface", "DCM", "Deep"), labels=c("15m", "DCM", "175m")))
+```
+
+```{r}
+gpc_chrom_gp <- all_cmpds %>%
+  mutate(compound_name=str_replace(compound_name, "Glycerophospho", "Glycerophospho-\n")) %>%
+  mutate(compound_name=fct_inorder(compound_name)) %>%
+  mutate(int=int/1e6) %>%
+  ggplot() +
+  geom_line(aes(x=rt, y=int, group=filename, color=depth)) +
+  facet_wrap(~compound_name, scales="free", nrow=1) +
+  scale_color_manual(values=viridis::viridis(9, direction = -1)[c(2,5,8)], 
+                     breaks = c("15m", "DCM", "175m"),
+                     aesthetics = c("color", "fill"),
+                     name="Sample\ndepth") +
+  theme_bw() +
+  labs(x="Retention time (minutes)", y="Intensity (millions)") +
+  guides(color=guide_legend(override.aes = list(lwd=2)))
+gpc_area_gp <- all_cmpds %>%
+  summarise(area=trapz(rt, int)/1e6, .by = c(compound_name, filename, depth, station, sla_corr)) %>%
+  mutate(compound_name=str_replace(compound_name, "Glycerophospho", "Glycerophospho-\n")) %>%
+  # mutate(station=factor(station, levels=as.character(4:14))) %>%
+  mutate(station=as.numeric(station)) %>%
+  mutate(compound_name=fct_inorder(compound_name)) %>%
+  ggplot() +
+  geom_hline(yintercept = 0) +
+  geom_point(aes(x=station, y=area, fill=sla_corr), color="black", pch=21) +
+  facet_grid2(depth~compound_name, scales="free_y", independent = "y") +
+  labs(x="Station number", y="Peak area (millions)", fill="Corr. SLA\n(cm)") +
+  scale_color_gradient2(low = scales::muted("blue"), 
+                        high = scales::muted("red"),
+                        mid="grey90", aesthetics = c("color", "fill")) +
+  theme_bw()
+
+gpc_ratio_reg_gp <- all_cmpds %>%
+  summarise(area=trapz(rt, int)/1e6, .by = c(compound_name, filename, depth, station, sla_corr)) %>%
+  pivot_wider(names_from = compound_name, values_from=area) %>%
+  pivot_longer(Phosphocholine:last_col()) %>%
+  filter(name%in%c("Phosphocholine", "Glycerophosphocholine")) %>%
+  mutate(name=paste0("Choline:", name)) %>%
+  mutate(name=fct_inorder(name)) %>%
+  ggplot(aes(x=sla_corr, y=Choline/value)) +
+  geom_smooth(method="lm", formula=y~x, color="black") +
+  geom_point(aes(fill=sla_corr), color="black", pch=21) +
+  facet_nested(.~name+depth, scales="free_y", independent = "y") +
+  labs(x="Corrected SLA (cm)", y="Metabolite ratio", fill="Corr. SLA\n(cm)") +
+  scale_fill_gradient2(low = scales::muted("blue"), 
+                        high = scales::muted("red"),
+                        mid="grey90", guide = "none") +
+  theme_bw()
+
+comb_gpc_patch <- gpc_chrom_gp / gpc_area_gp / gpc_ratio_reg_gp + plot_annotation(tag_levels = 'A') + plot_layout(heights = c(0.3, 0.4, 0.3))
+ggsave("supplement/comb_gpc_patch.png", comb_gpc_patch, device = "png", width = 9, height = 6.5, units = "in", dpi = 300, bg="white")
+```
+
+## Supp fig 5: Boxplots of selected compounds by depth, time, and SLA
 
@@
 
-*Supplemental Figure 4: Boxplots showing the concentrations of six selected compounds between cruises and depths as well as time and SLA classes.
+*Supplemental Figure 5: Boxplots showing the concentrations of six selected compounds between cruises and depths as well as time and SLA classes.
 The abscissa is separated by the time category into which sampling was binned, referring to 6-hour increments starting from 3AM (e.g.
@@
 
-## Supp fig 5: HEE NMDS and PERMANOVA broken down by depth
+## Supp fig 6: HEE NMDS and PERMANOVA broken down by depth
 
@@
 
-*Supplemental Figure 5: Plots of the metabolome during the Hawaiian Eddy Experiment, broken down by depth.
+*Supplemental Figure 6: Plots of the metabolome during the Hawaiian Eddy Experiment, broken down by depth.
 The top row of plots are non-metric multidimensional scaling (NMDS) plots, in which points correspond to individual samples and have been colored by their corrected sea level height anomaly.

### 2024-10-29 manuscript/response_to_reviewers.txt
Adding response to nested effects concern

@@
 > effects due to the selection of data points.
+It is unclear what the reviewer means by the selection of data points. We are aware that
+the strong trend between particulate carbon and metabolite concentration is largely due
+to the effect of depth and that was our inspiration for performing both the overall
+regression and independent regressions at each depth.

### 2024-10-29 manuscript/manuscript.Rmd
Copyedits for readability

@@
 This was followed by manual inspection of the extracted ion chromatogram for each mass feature by an expert for quality with only those peaks classified as "Good" used in the analysis here.
-Features were then matched with and normalized to an internal standard (Supplemental Table 1) as described above.
+Features were then matched with and normalized to an internal standard (Supplemental Table 4) as described above.
 Scripts used for this analysis are available at https://github.com/wkumler/MesoscopeMetabolomicsManuscript/tree/master/untargeted.
@@
 In the samples taken along the eddy transect, the largest effect was detected in the deepest samples taken from 175 meters and from the DCM, with much less of a response at 25 meters.
+This lack of SLA effect at the surface is unsurprising given the similarity in the biogeochemistry of the well-lit upper ocean previously reported [@Barone2022].
 In the HEE samples, however, the opposite was true with a surprisingly larger SLA response at 25 meters than that at the DCM.
 In each case, the differences introduced by SLA were smaller than those of depth when compared directly, with even the high-resolution sampling around the DCM finding slightly more variance explained by the 40 meter difference in sampling depth than the 40 centimeter difference in SLA.
-This result agrees well with prior research by @Heal2021 and @Kumler2023 showing the large effect of sampling depth on the metabolome.
-Time of day was included as a significant factor in the MESO-SCOPE transect, with trehalose and sucrose in particular showing large differences between 6AM and 6PM as has been noted previously [@Muratore2022], though this effect was diluted at 175 meters.
+This result agrees well with prior research showing the large effect of sampling depth on the metabolome [@Heal2021; @Kumler2023; @Bent2024].
+Time of day was included as a significant factor in the MESO-SCOPE transect, with trehalose and sucrose in particular showing large differences between 6AM and 6PM as has been previously noted [@Muratore2022], though this effect was diminished at 175 meters.
 
@@
 We also detected the largest differences in the DCM metabolome at the exact center of the eddy, with Station 12 highly distinct from even the nearest stations, while no such stark difference was found in the anticyclone (Figures 1 and 2).
-This perhaps indicates that the exact center of the cyclone has a unique metabolome at the DCM relative to the rest of the transect, rather than existing as a smooth continuation.
+This indicates that the exact center of the cyclone has a unique metabolome at the DCM relative to the rest of the transect, rather than existing as a smooth gradient with SLA.
 
@@
 This revealed four major patterns of metabolite response, with a majority of compounds responding to eddy state (Clusters 1, 2, and 4 in Figure 5).
-It is interesting to note that the sole cluster in which abundances in the anticyclone are greater than those of the cyclone is also the only cluster to generally increase in abundance with depth (Cluster 4), perhaps indicating that compounds more concentrated in the anticyclone are the same kind of degradation products and recalcitrant carbon typically found at depth.
+It is interesting to note that the sole cluster in which abundances in the anticyclone are greater than those of the cyclone is also the only cluster to generally increase in abundance with depth (Cluster 4), likely because compounds more concentrated in the anticyclone are the same kind of degradation products and recalcitrant carbon typically found at depth.
 

### 2024-10-30 manuscript/manuscript.Rmd
Responding to reviewer #1 concerns
Needs additional feedback from coauthors

@@
 Instead, we found more compounds overall to be significantly different across the eddy transect in the 175 meter samples than we did at the DCM (α=0.05, 43 compounds at 175 meters vs 22 at the DCM).
-This was likely driven by the enhancement of organic matter and heterotrophic picoplankton below the DCM as seen in @Barone2019, with the enrichment of nucleobases in the 175 meter samples additionally hint at increases in bacterial biomass given that nucleobases tend to be highly abundant in bacterial cultures [@Heal2021].
+This was likely driven by the enhancement of organic matter and heterotrophic picoplankton below the DCM as seen in @Barone2019, with the enrichment of nucleobases in the 175 meter samples additionally hinting at increases in bacterial biomass given that nucleobases tend to be highly abundant in bacterial cultures [@Heal2021].
 The abundance of a mass feature putatively identified as acetylcholine in the 175 meter cyclone samples followed the same general trend as arsenobetaine, both enriched at depth and in the anticyclone, and may be another compound that results from heterotrophic degradation of phytoplankton metabolites [@Durham2022].
@@
 
-Differences between the two cruises could be due to differences in eddy age and/or the different period of the year when sampling took place during the two cruises (March/April in 2018 instead of the June/July sampling in 2017) resulting in shifts in community composition and response.
-@Barone2022 documents the 2017 cruise community composition, reporting abundant Prymnesiophyceae at both 15 meters and the DCM, with other significant contributions from dinoflagellates, diatoms, and green algae.
-In the 2018 cruise both haptophytes and dinoflagellates again dominated the eukaryotic fraction, though surprisingly little was contributed by diatoms [@Gleich2024].
-@Dugenne2023 and @Harke2021 also highlight uncontrolled differences between the two sets of eddies based on their origin and age.
-
-Missing links between the environment with its associated community and the composition of particulate matter make it difficult to estimate this important control on the marine carbon cycle.
+## Metabolites link changes in biogeochemistry to community biology
+
+Missing links between the environment with its associated community and the composition of particulate matter make it difficult to estimate important controls on the marine carbon cycle.
 This is especially true for dynamic and short-lived environments such as eddies.
-Metabolites offer a way to bridge this gap in knowledge but the relative impact of eddies on their contribution and fate was not previously estimated.
-In this work, we have shown the sensitivity of the metabolome to various environmental and community factors and contrasted it to the well-established effect of depth to constrain the importance of including eddy effects in models of elemental cycling.
+Metabolites offer a way to bridge this gap in knowledge by directly measuring the units of exchange between organisms and their environment.
+
+Here, we saw how changes in the density surface driven by physics during this process led to the coupling of physics, chemistry, and biology.
+The enrichment of eukaryotic organisms at the DCM of the cyclone mirrors other ocean processes where upwelling of deep, nutrient-rich water leads to increases in nitrogen-rich metabolites representative of a shift towards organisms such as diatoms and dinoflagellates that undergo bloom dynamics.
+@Barone2022 established a "chemical wake" hypothesis in which cyclone formation introduces a transient phytoplankton community with net photosynthetic production that induces export, a conceptual model that can be extended to include many of the metabolites named here, as some metabolites are more bioavailable than others and the "flavor" of this chemical wake will change over time.
+
+Almost all large phytoplankton classes in the 2017 eddy pair were enriched in the cyclone DCM relative to the anticyclone (particularly prymnesiophytes, dinoflagellates, diatoms, and green algae), while *Prochlorococcus* was significantly less abundant and *Synechococcus* was approximately equal [@Barone2022].
+Thus, the metabolite data presented here does not help distinguish between large phytoplankton classes but instead provides additional cues about the shift from a metabolome dominated by cyanobacteria producers into one dominated by eukaryotic organisms.
+This shift occurs anywhere
+
+A shift from targeted to untargeted metabolites may also in part explain why the increased biomass in the cyclone did not result in large increases in export [@Barone2022; @Bent2024].
+Our list of known compounds covers a large fraction of the small molecules common to most organisms, indicating that unknown molecules are more likely to be rare and therefore more difficult to degrade.
+If the increased biomass in the cyclone is dominantly in the form of highly bioavailable compounds, it is not surprising that much of it can be easily degraded.
+This hypothesis is supported by @Gleich2024, who found that heterotrophy-associated transcripts were more abundant in the 2018 cruise cyclone than the anticyclone.
+
+Both cruises occurred in weakening cyclones and stable anticyclones, limiting the degree to which the results presented here can be extrapolated to the full life cycle of an eddy [@Dugenne2023].
+Differences between the 2017 and 2018 cruises in both biogeochemistry and community biology additionally make it difficult to generalize the links between these two factors via metabolites and establish the demand for additional characterization of the metabolome in the North Pacific Subtropical Gyre.
 
@@
 Finally, the metabolites with the strongest responses to eddy effects were unknown, and a relatively smaller proportion of known metabolites in the anticyclone demonstrates the utility of untargeted metabolomics in exploring environmental variation and the large amount of information that is missed by only investigating known compounds.
+In this work, we have shown the sensitivity of the metabolome to various environmental and community factors and constrained the importance of eddy effects in models of elemental cycling.
 

### 2024-10-30 manuscript/response_to_reviewers.txt
Adding possible response to Reviewer #1

@@
 > adjusted.
-
 > This article focuses on the differences in chemical composition to understand this 
@@
 > but relevant discussions are possible.
-
 > We already know that mesoscale eddies are complex and dynamic multidimensional 
@@
 > method, I think similar situations can also be seen in such processes.
+We have expanded the Discussion to touch on the concerns raised by this reviewer.
+Specifically, we have mentioned possible links between the increased sulfonium
+compounds in the eddy and DMS production, a hypothesis about the reason why
+increased cyclonic biomass was not matched by increased export, and extensive
+references to additional manuscripts with relevant data.
 

### 2024-10-30 manuscript/manuscript.Rmd
Forgot to add the GPC fig to the manuscript

@@
 
+![](supplement/comb_gpc_patch.png)
+
+*Supplemental Figure 4: Evidence against phosphorus limitation across the MESO-SCOPE eddy transect.
+Top row of plots (A) correspond to extracted ion chromatograms for 5 metabolites useful for estimating the phosphorus stress of an environment, colored by the depth from which the sample was taken.
+Middle row of plots shows the integrated peak area for each metabolite across the 11 stations of the transect with sub-panels for each depth independently, with each point colored by the corrected sea level anomaly (SLA).
+Bottom row of plots shows the ratio of choline to phosphocholine (left) and choline to glycerophosphocholine (right) as a function of corrected SLA with no statistically significant trend visible at any of the three depths.*
+
 ## Supp fig 5: Boxplots of selected compounds by depth, time, and SLA

### 2024-11-14 manuscript/response_to_reviewers.txt
Adding line numbers
Might regret this if I add a bunch of text above and it moves things down :(

--- 2024-10-29 manuscript/response_to_reviewers.txt
@@
 > method, I think similar situations can also be seen in such processes.
+We have expanded the Discussion to touch on the concerns raised by this reviewer.
+Specifically, we have mentioned possible links between the increased sulfonium
+compounds in the eddy and DMS production, a hypothesis about the reason why
+increased cyclonic biomass was not matched by increased export, and extensive
+references to additional manuscripts with relevant data in lines 651-655 and 
+689-718.
 
@@
 We have added the time of day as an explanatory factor to our analysis and reported the
-new values from the PERMANOVA in the text and figures, as well as adding a supplemental
+new values from the PERMANOVA in the text (Lines 220-226, 262-266 and 505-509) and figures, 
+as well as adding a supplemental
 figure to visually represent some of the variance due to time of day. We do see 
@@
 the course of the eddy transect sample run, the normalization removed this
-effect and is not a major factor in the outliers visible in Fig. 2.
+effect and is not a major factor in the outliers visible in Fig. 2. (Lines 190-192)
 
@@
 We have added a note about how we used the methodological blanks to visually assess 
-whether sample peak areas were above the background signal.
+whether sample peak areas were above the background signal. (Lines 184-186)
 
@@
 > centrifugation required? Was separation done by pipetting?
-We have added additional text to this line to clarify the requested points of confusion.
+We have added additional text to this line to clarify the requested points of 
+confusion. (Lines 133-138)
 
@@
 depths. We have also added links to the Github where the R code used to perform
-these matching and normalization steps is available.
+these matching and normalization steps is available. (Lines 196-198 and 205-208)
 
@@
 Done. Software versions for Proteowizard, Skyline, MSDIAL, R, XCMS, and vegan have been
-added to the manuscript.
+added to the manuscript. (Sections 3.6 and 3.7)
 
@@
 R-squared and p-values were obtained from PERMANOVAs. We have also added the sample
-size to the Statistics section of the Results.
+size to the Statistics section of the Results. (Lines 220-226)
 

### 2024-11-14 manuscript/manuscript.Rmd
Forgot to add the GPC fig to the manuscript
Adding spaces between 6AM and 6PM per Wei's recommendation

@@
 Despite this, we still saw the importance of sea level anomaly as a significant explanatory factor in the dataset (PERMANOVA R$^2_{SLA} = 0.090$, p-value = 0.017) as well as the larger depth differences between the surface (25 meters) and DCM (110 - 120 meters) (PERMANOVA R$^2_{depth} = 0.251$, p-value << 0.001).
-We saw no difference between the 6AM and 6PM samples (PERMANOVA R$^2_{time} = 0.031$, p-value = 0.367) despite a few known compounds (trehalose and sucrose) showing large differences, likely due to a majority of the mass features demonstrating no diel effect (Supplemental Figure 5).
+We saw no difference between the 6 AM and 6 PM samples (PERMANOVA R$^2_{time} = 0.031$, p-value = 0.367) despite a few known compounds (trehalose and sucrose) showing large differences, likely due to a majority of the mass features demonstrating no diel effect (Supplemental Figure 5).
 
@@
 This result agrees well with prior research showing the large effect of sampling depth on the metabolome [@Heal2021; @Kumler2023; @Bent2024].
-Time of day was included as a significant factor in the MESO-SCOPE transect, with trehalose and sucrose in particular showing large differences between 6AM and 6PM as has been previously noted [@Muratore2022], though this effect was diminished at 175 meters.
+Time of day was included as a significant factor in the MESO-SCOPE transect, with trehalose and sucrose in particular showing large differences between 6 AM and 6 PM as has been previously noted [@Muratore2022], though this effect was diminished at 175 meters.
 
@@
 ![](supplement/comb_gpc_patch.png)
-
 *Supplemental Figure 4: Evidence against phosphorus limitation across the MESO-SCOPE eddy transect.

### 2024-11-14 manuscript/manuscript.Rmd
Adding refs for N-metabs in euk blooms

@@
 Here, we saw how changes in the density surface driven by physics during this process led to the coupling of physics, chemistry, and biology.
-The enrichment of eukaryotic organisms at the DCM of the cyclone mirrors other ocean processes where upwelling of deep, nutrient-rich water leads to increases in nitrogen-rich metabolites representative of a shift towards organisms such as diatoms and dinoflagellates that undergo bloom dynamics.
+The enrichment of eukaryotic organisms at the DCM of the cyclone mirrors other ocean processes where upwelling of deep, nutrient-rich water leads to increases in nitrogen-rich metabolites representative of a shift towards organisms such as diatoms and dinoflagellates that undergo bloom dynamics [@Durham2022; @Kulisch2024].
 @Barone2022 established a "chemical wake" hypothesis in which cyclone formation introduces a transient phytoplankton community with net photosynthetic production that induces export, a conceptual model that can be extended to include many of the metabolites named here, as some metabolites are more bioavailable than others and the "flavor" of this chemical wake will change over time.
@@
 
-Both cruises occurred in weakening cyclones and stable anticyclones, limiting the degree to which the results presented here can be extrapolated to the full life cycle of an eddy [@Dugenne2023].
-Differences between the 2017 and 2018 cruises in both biogeochemistry and community biology additionally make it difficult to generalize the links between these two factors via metabolites and establish the demand for additional characterization of the metabolome in the North Pacific Subtropical Gyre.
+Both cruises occurred in weakening cyclones and stable anticyclones
 

### 2024-11-14 manuscript/manuscript.Rmd
Adding DMSP/Ac abbreviations

@@
 A majority of the known molecules (37/53) had a similar pattern with a subsurface maximum at the DCM, including glycine betaine, glutamate, and guanine.
-Thirteen molecules obeyed a different pattern and decreased monotonically with depth, including the molecules hydroxyisoleucine, gonyol, and dimethylsulfoniopropionate.
+Thirteen molecules obeyed a different pattern and decreased monotonically with depth, including the molecules hydroxyisoleucine, gonyol, and dimethylsulfoniopropionate (DMSP).
 Two known molecules (arsenobetaine and O-acetylcarnitine) increased with depth, potentially representing particle degradation during export.
@@
 Additionally, the abundance other sulfur-containing compounds (e.g.
-dimethylsulfonioacetate and dimethylsulfoniopropionate) in the cyclone may indicate links between the formation of eukaryote-rich cyclonic features and the production of dimethylsulfide, a climate-active volatile gas [@Moran2019].
+dimethylsulfonioacetate, DMS-Ac and dimethylsulfoniopropionate, DMSP) in the cyclone may indicate links between the formation of eukaryote-rich cyclonic features and the production of dimethylsulfide, a climate-active volatile gas [@Moran2019].
 

### 2024-11-14 manuscript/manuscript.Rmd
Responding to Rachel's comments

@@
 Previous studies have explored eddy effects on biogeochemistry and microbial community composition but not on the molecular composition of particulate organic matter.
-This study reports the absolute concentration of 67 metabolites and relative abundances for 640 molecular features to understand how mesoscale eddies impact the metabolome of the North Pacific Subtropical Gyre during two cruises in 2017 and 2018.
+This study reports the absolute concentration of 67 metabolites and relative abundances for 640 molecular features, measured using liquid chromatography-mass spectrometry (LC-MS) following both targeted and untargeted approaches.
+This approach allowed us to better understand how mesoscale eddies impact the metabolome of the North Pacific Subtropical Gyre during two cruises in 2017 and 2018.
 We find that many metabolites track biomass trends, but metabolites like isethionic acid, homarine, and trigonelline linked to eukaryotic phytoplankton were enriched at the deep chlorophyll maximum of the cyclonic features, while degradation products such as arsenobetaine were enriched in anticyclones.
-In every analysis, metabolites with the strongest responses were detected via untargeted mass spectrometry, indicating that the molecules most sensitive to environmental perturbation were not among the characterized metabolome.
+In every analysis, the metabolites with the strongest responses were identified using LC-MS through untargeted metabolomics approaches, highlighting that the molecules most sensitive to environmental perturbation were not among the previously characterized metabolome.
 By analyzing depth variability (accounting for 20-40% of metabolomic variability across ~150 meters) and the vertical displacement of isopycnal surfaces (explaining 10-20% of variability across a sea level anomaly range of 40 centimeters and a spatial distance of 300 kilometers), this analysis constrains the importance of mesoscale eddies in shaping the chemical composition of particulate matter in the largest biomes on the planet.
@@
 
-The uplift of deep, nutrient-rich seawater into the euphotic zone alters microbial communities [@Rii2022].
+The uplift of deep, nutrient-rich seawater into the euphotic zone alters microbial communities [@Rii2008; @Rii2022].
 Measurements of chlorophyll in cyclones reveal a shallower and more intense maximum [@Barone2022; @Cornec2021] as a result of eukaryotic phytoplankton thriving in the higher nutrient concentrations while cyanobacterial biomass is reduced [@Hawco2021].
@@
 This shift in community structure and productivity should result in a corresponding shift in the composition of the particulate organic matter produced.
-Eukaryotic organisms have distinct chemical fingerprints from those of cyanobacteria and heterotrophic bacteria [@Heal2021; @Durham2022] and the relative contribution of these different taxa to total biomass should reflect their chemical composition.
+Eukaryotic organisms have distinct chemical fingerprints from those of cyanobacteria and heterotrophic bacteria [@Heal2021; @Durham2022; @Kuhlisch2023] and the relative contribution of these different taxa to total biomass should reflect their chemical composition.
 However, recent work from @Harke2021 showed that overall community function may be robust across eddy types at the surface while @Gleich2024 detected significant differences in protistan community composition and metabolic potential down to 250 meters, with heterotrophy-associated protistan transcripts enriched in the cyclone.
@@
 In this work, we directly tested for differences in the chemical composition of particulate matter due to changes in eddy state in the NPSG.
-We sampled from cyclonic and anticyclonic eddies spatially near each other and used targeted and untargeted mass spectrometry to explore changes in the composition of small, polar molecules.
+We sampled from cyclonic and anticyclonic eddies spatially near each other and used liquid chromatography-mass spectrometry (LC-MS) for both targeted and untargeted metabolomics to explore changes in the composition of small, polar molecules.
 We expected to find that overall metabolite abundance would reflect shifts in biomass over multiple depths.
@@
 Here, we saw how changes in the density surface driven by physics during this process led to the coupling of physics, chemistry, and biology.
-The enrichment of eukaryotic organisms at the DCM of the cyclone mirrors other ocean processes where upwelling of deep, nutrient-rich water leads to increases in nitrogen-rich metabolites representative of a shift towards organisms such as diatoms and dinoflagellates that undergo bloom dynamics [@Durham2022; @Kulisch2024].
+The enrichment of eukaryotic organisms at the DCM of the cyclone mirrors other ocean processes where upwelling of deep, nutrient-rich water leads to increases in nitrogen-rich metabolites representative of a shift towards organisms such as diatoms and dinoflagellates that undergo bloom dynamics [@Durham2022; @Kulisch2023].
 @Barone2022 established a "chemical wake" hypothesis in which cyclone formation introduces a transient phytoplankton community with net photosynthetic production that induces export, a conceptual model that can be extended to include many of the metabolites named here, as some metabolites are more bioavailable than others and the "flavor" of this chemical wake will change over time.

### 2024-11-15 manuscript/manuscript.Rmd
Tidying LCMS terminology

@@
 Dry samples were reconstituted in 380 µL Optima-grade water and amended with 20 µL isotope-labeled injection standards (Supplemental Table 4) to measure the variability introduced by chromatography and ionization.
-The reconstituted fraction was syringe-filtered into precombusted glass inserts in HPLC vials to remove any potential clogging material.
+The reconstituted fraction was syringe-filtered into precombusted glass inserts in liquid chromatography (LC) vials to remove any potential clogging material.
 Samples were then additionally diluted 1:1 with Optima-grade water to prevent overloading on the column and to reduce salt effects.
 
-A pooled sample was created by combining 20 µL of each sample into the same HPLC vial, and a 1:1 dilution with water created a half-strength pooled sample from that to assess matrix effects and obscuring variation [@Boysen2018].
+A pooled sample was created by combining 20 µL of each sample into the same LC vial, and a 1:1 dilution with water created a half-strength pooled sample from that to assess matrix effects and obscuring variation [@Boysen2018].
 Also run alongside the environmental samples were samples of authentic standards split into two mixes (Supplemental Table 5).
 Standards were dissolved in Optima-grade water to help with identification and into an aliquot of the pooled sample to quantify matrix response factors.
-HPLC vials containing the samples were frozen at -80 °C until thawing shortly before injection.
+LC vials containing the samples were frozen at -80 °C until thawing shortly before injection.
 
 ## HPLC-MS methods
-Separate HPLC-MS runs were used for the MESO-SCOPE eddy transect, the MESO-SCOPE eddy center, and the HEE cruise data.
+Separate LC-MS runs were used for the MESO-SCOPE eddy transect, the MESO-SCOPE eddy center, and the HEE cruise data.
 Eddy center samples were run in February 2018, eddy transect samples were run later during August of that year, and the HEE samples were run in July 2019.
@@
 Normalized peaks were then calibrated via a single-point standard curve as determined uniquely for each compound from the authentic standard mixes and used to convert normalized peak area into moles per µL injected and scaled to the amount of seawater filtered to provide a final estimate of environmental concentration for each compound in moles per liter of seawater filtered.
-Scripts used for this analysis are available at https://github.com/wkumler/MesoscopeMetabolomicsManuscript/tree/master/untargeted.
+Scripts used for this analysis are available at https://github.com/wkumler/MesoscopeMetabolomicsManuscript/tree/master/targeted.
 

### 2024-11-15 manuscript/manuscript.Rmd
Cleaning up discussion, adding refs

@@
 Here, we saw how changes in the density surface driven by physics during this process led to the coupling of physics, chemistry, and biology.
-The enrichment of eukaryotic organisms at the DCM of the cyclone mirrors other ocean processes where upwelling of deep, nutrient-rich water leads to increases in nitrogen-rich metabolites representative of a shift towards organisms such as diatoms and dinoflagellates that undergo bloom dynamics [@Durham2022; @Kulisch2023].
+The enrichment of eukaryotic organisms at the DCM of the cyclone mirrors other ocean processes where upwelling of deep, nutrient-rich water leads to increases in nitrogen-rich metabolites representative of a shift towards organisms such as diatoms and dinoflagellates that undergo bloom dynamics [@Durham2022; @Kuhlisch2023].
 @Barone2022 established a "chemical wake" hypothesis in which cyclone formation introduces a transient phytoplankton community with net photosynthetic production that induces export, a conceptual model that can be extended to include many of the metabolites named here, as some metabolites are more bioavailable than others and the "flavor" of this chemical wake will change over time.
-
-Almost all large phytoplankton classes in the 2017 eddy pair were enriched in the cyclone DCM relative to the anticyclone (particularly prymnesiophytes, dinoflagellates, diatoms, and green algae), while *Prochlorococcus* was significantly less abundant and *Synechococcus* was approximately equal [@Barone2022].
-Thus, the metabolite data presented here does not help distinguish between large phytoplankton classes but instead provides additional cues about the shift from a metabolome dominated by cyanobacteria producers into one dominated by eukaryotic organisms.
-This shift occurs anywhere
 
@@
 
-Both cruises occurred in weakening cyclones and stable anticyclones
+Almost all large phytoplankton classes in the 2017 eddy pair were enriched in the cyclone DCM relative to the anticyclone (particularly prymnesiophytes, dinoflagellates, diatoms, and green algae), while *Prochlorococcus* was significantly less abundant and *Synechococcus* was approximately equal [@Barone2022].
+Thus, the metabolite data presented here does not help distinguish between large phytoplankton classes but instead provides additional cues about the shift from a metabolome dominated by cyanobacteria producers into one dominated by eukaryotic organisms.
+This shift occurs naturally over the life cycle of an eddy, with distinctions typically drawn between the eddy intensification, its mature phase, and eventual decay [@Rii2008; @McGillicuddy2016; @Kuhlisch2023; @Barone2022].
+Both cruises occurred in weakening cyclones and mature anticyclones, limiting our ability to make direct conclusions about all eddy life cycle phases, but the general alignment of the results presented here with our initial hypotheses strengthens our confidence in metabolites as links between eddy biogeochemistry and their biological shifts anywhere.
 

### 2024-11-15 manuscript/manuscript.Rmd
Cleaning up discussion, adding refs, final edits

@@
 To characterize the observed differences in multivariate space we used k-means clustering as an unsupervised way to identify dominant trends within the data (Figure 5).
-We found that a majority of the metabolites (Clusters 1 and 2, 63% of the total) fell into clusters with larger peak areas in the cyclone, as expected, while also detecting 32 metabolites that clustered such that the mean metabolite was enriched in the anticyclone (Cluster 4, Figure 5).
+We found that a majority of the metabolites (Clusters 1 and 2, 63% of the total) fell into clusters with larger peak areas in the cyclone while also detecting 32 metabolites that clustered such that the mean metabolite was enriched in the anticyclone (Cluster 4, Figure 5).
 Cluster 2 had a distinct decrease in relative peak area with depth, while the other clusters had much less clear depth trends.
@@
 This shift occurs naturally over the life cycle of an eddy, with distinctions typically drawn between the eddy intensification, its mature phase, and eventual decay [@Rii2008; @McGillicuddy2016; @Kuhlisch2023; @Barone2022].
-Both cruises occurred in weakening cyclones and mature anticyclones, limiting our ability to make direct conclusions about all eddy life cycle phases, but the general alignment of the results presented here with our initial hypotheses strengthens our confidence in metabolites as links between eddy biogeochemistry and their biological shifts anywhere.
+The initial introduction of inorganic nutrients during intensification of cyclonic eddies followed by diapycnal mixing can still stimulate the DCM community without large effects on integrate primary productivity or export even months after formation, as seen here.
+Both cruises occurred in weakening cyclones and mature anticyclones, limiting our ability to make direct conclusions about all eddy life cycle phases.
+However, the general alignment of the results presented here with our initial hypotheses strengthens our confidence in metabolites as links between eddy biogeochemistry and the associated biological shifts anywhere.
 

### 2024-11-15 manuscript/manuscript.Rmd
Additional edits and tidying from Benedetto

@@
 A shift from targeted to untargeted metabolites may also in part explain why the increased biomass in the cyclone did not result in large increases in export [@Barone2022; @Bent2024].
-Our list of known compounds covers a large fraction of the small molecules common to most organisms, indicating that unknown molecules are more likely to be rare and therefore more difficult to degrade.
+Our list of known compounds covers a large fraction of the small molecules common to most organisms, indicating that unknown molecules are more likely to be rare and therefore more difficult to degrade or a result of prokaryote dominance.
 If the increased biomass in the cyclone is dominantly in the form of highly bioavailable compounds, it is not surprising that much of it can be easily degraded.
@@
 We would also like to acknowledge the analysts involved in the biogeochemical measurements for their willingness to provide data and answer questions (Eric Grabowski, Karin Bjorkman, and Rhea Foreman).
-This work was supported by grants from the Simons Foundation (SCOPE Award ID 329108 to AI, SF Award ID 385428 to AI, Postdoctoral Fellowship in Marine Microbial Ecology ID 548565 to WQ)
+This work was supported by grants from the Simons Foundation (SCOPE Award ID 329108 to AI, SF Award ID 385428 to AI, Postdoctoral Fellowship in Marine Microbial Ecology ID 548565 to WQ, and SCOPE Award ID 721264 to David Karl which supported BB).
 

### 2024-11-15 manuscript/manuscript.Rmd
Additional edits and tidying from Benedetto

@@
 Data and code used for this manuscript are all available online.
-Biogeochemical data can be sourced from the Simons SCOPE website at https://scope.soest.hawaii.edu/data/ and metabolomics data has been uploaded to Metabolomics Workbench under Project ID PR001738, where the data can be accessed directly via its DOI: 10.21228/M82719.
+Biogeochemical data can be sourced from the Simons SCOPE website at https://scope.soest.hawaii.edu/data/ and has been provided in Supplemental Tables 1-3.
+Metabolomics data has been uploaded to Metabolomics Workbench under Project ID PR001738, where the data can be accessed directly via its DOI: 10.21228/M82719.
 Scripts used to process metabolomics data are available on GitHub at https://github.com/wkumler/MesoscopeMetabolomicsManuscript.

### 2024-11-15 manuscript/manuscript.Rmd
typo :upside_down_face:

@@
 This shift occurs naturally over the life cycle of an eddy, with distinctions typically drawn between the eddy intensification, its mature phase, and eventual decay [@Rii2008; @McGillicuddy2016; @Kuhlisch2023; @Barone2022].
-The initial introduction of inorganic nutrients during intensification of cyclonic eddies followed by diapycnal mixing can still stimulate the DCM community without large effects on integrate primary productivity or export even months after formation, as seen here.
+The initial introduction of inorganic nutrients during intensification of cyclonic eddies followed by diapycnal mixing can still stimulate the DCM community without large effects on integrated primary productivity or export even months after formation, as seen here.
 Both cruises occurred in weakening cyclones and mature anticyclones, limiting our ability to make direct conclusions about all eddy life cycle phases.

### 2024-11-15 manuscript/response_to_reviewers.txt
Cleaning up discussion, adding refs
Switching section headings per Rachel and reviewer recs

@@
 > method, I think similar situations can also be seen in such processes.
-We have expanded the Discussion to touch on the concerns raised by this reviewer.
+We have expanded the Discussion to touch on the concerns raised by this reviewer,
+adding section 5.3 to respond to this directly.
 Specifically, we have mentioned possible links between the increased sulfonium
@@
 > definition rather untargeted.
-While the exact definition of "targeted" vs "untargeted" metabolomics remains an 
-active topic of debate, we believe that the methods used here are sufficiently
-targeted (use of authentic standards and manual integration via Skyline/MSDIAL) to 
-warrant the use of the term, especially in direct contrast to the automated/untargeted 
-pipeline described immediately afterward and thus we have not made this change.
+We have restructured the headings for this section of the methods and added
+additional detail to respond to this concern.
 

### 2024-11-15 manuscript/manuscript.Rmd
Switching section headings per Rachel and reviewer recs
Adding Anitra and Benedetto recs for section 5.3

--- 2024-11-15 manuscript/manuscript.Rmd
@@
 
-## Targeted metabolomic methods
+## Metabolomic data analysis
+
+### Full scan feature extraction for known molecules
 
@@
 
-## Untargeted metabolomic methods
+### Automated processing for detection of unknown features
 
@@
 
-Missing links between the environment with its associated community and the composition of particulate matter make it difficult to estimate important controls on the marine carbon cycle.
-This is especially true for dynamic and short-lived environments such as eddies.
-Metabolites offer a way to bridge this gap in knowledge by directly measuring the units of exchange between organisms and their environment.
-
-Here, we saw how changes in the density surface driven by physics during this process led to the coupling of physics, chemistry, and biology.
-The enrichment of eukaryotic organisms at the DCM of the cyclone mirrors other ocean processes where upwelling of deep, nutrient-rich water leads to increases in nitrogen-rich metabolites representative of a shift towards organisms such as diatoms and dinoflagellates that undergo bloom dynamics [@Durham2022; @Kuhlisch2023].
-@Barone2022 established a "chemical wake" hypothesis in which cyclone formation introduces a transient phytoplankton community with net photosynthetic production that induces export, a conceptual model that can be extended to include many of the metabolites named here, as some metabolites are more bioavailable than others and the "flavor" of this chemical wake will change over time.
-
-A shift from targeted to untargeted metabolites may also in part explain why the increased biomass in the cyclone did not result in large increases in export [@Barone2022; @Bent2024].
+The composition of particulate matter controls the turnover of organic carbon in the ocean, but it is still unclear how this composition is altered in the environment and the plankton community.
+Measuring metabolites can quantify the variability of the central molecules responsible for the flow of energy and matter through the marine ecosystem at many different scales.
+In these results, we focused on mesoscale eddies, which perturb the ocean ecosystem on relatively small horizontal scales on the order of tens to hundreds of kilometers.
+
+Here, we reported how changes in the distribution of water layers of different density driven by mesoscale physical dynamics cascading effects in the biological and chemical characteristics of the eddies.
+The enrichment of eukaryotic organisms at the DCM of the cyclone mirrors other ocean processes where upwelling of deep, nutrient-rich water leads to increases in sulfur-rich metabolites representative of a shift towards organisms such as diatoms and dinoflagellates that undergo bloom dynamics [@Durham2022; @Kuhlisch2023].
+@Barone2022 proposed that a "chemical wake" is formed after cyclone intensification when a transient phytoplankton community with net photosynthetic production induces export and leaves deficits and excesses of inorganic nutrients and oxygen, respectively.
+This conceptual model that can be extended to include many of the metabolites named here, as some metabolites are more bioavailable than others and the "flavor" of the chemical wake will change throughout the eddy life cycle.
+
+The prevalence of unknown metabolites relative to known metabolites in the anticyclone may also in part explain why the increased biomass in the cyclone did not result in large increases in export [@Barone2022; @Bent2024].
 Our list of known compounds covers a large fraction of the small molecules common to most organisms, indicating that unknown molecules are more likely to be rare and therefore more difficult to degrade or a result of prokaryote dominance.
 If the increased biomass in the cyclone is dominantly in the form of highly bioavailable compounds, it is not surprising that much of it can be easily degraded.
-This hypothesis is supported by @Gleich2024, who found that heterotrophy-associated transcripts were more abundant in the 2018 cruise cyclone than the anticyclone.
+This hypothesis is supported by @Gleich2024, who found that heterotrophy-associated transcripts were more abundant in the 2018 cruise cyclone than the anticyclone, indicating that there was fresh material for the heterotrophic bacteria to readily consume.
 
@@
 This shift occurs naturally over the life cycle of an eddy, with distinctions typically drawn between the eddy intensification, its mature phase, and eventual decay [@Rii2008; @McGillicuddy2016; @Kuhlisch2023; @Barone2022].
-The initial introduction of inorganic nutrients during intensification of cyclonic eddies followed by diapycnal mixing can still stimulate the DCM community without large effects on integrate primary productivity or export even months after formation, as seen here.
-Both cruises occurred in weakening cyclones and mature anticyclones, limiting our ability to make direct conclusions about all eddy life cycle phases.
-However, the general alignment of the results presented here with our initial hypotheses strengthens our confidence in metabolites as links between eddy biogeochemistry and the associated biological shifts anywhere.
+Increased diapycnal mixing of nutrients in mature cyclones can alter the DCM community without large effects on integrated primary productivity or export even months after eddy intensification [@Barone2022].
+We here demonstrated how this dynamic is linked with large changes in the chemical characteristics of the particulate material.
 

### 2024-11-15 manuscript/manuscript.Rmd
Adding BMIS info to supplemental table

@@
 raafay_notes <- read_csv("../targeted/raw_data/Raafay_HILIC_notes.csv") %>%
-  select(compound_name, in_samps)
+  select(compound_name, in_samps, polarity)
+bmis_info <- read_csv("../targeted/made_data/BMIS_areas.csv") %>%
+  distinct(compound_name, cruise, BMIS, polarity)
 
@@
   left_join(raafay_notes) %>%
+  left_join(bmis_info) %>%
   mutate(polarity=ifelse(polarity=="pos", "Positive", "Negative")) %>%
@@
          `Concentration (uM)`=conc_um, `Polarity integrated`=polarity,
-         `Standard mix`=mix, `Which runs`=ms_run, `Good quality`=in_samps)
+         `Standard mix`=mix, `Which runs`=ms_run, `Good quality`=in_samps,
+         `MS run`=cruise, `Best-matched IS`=BMIS)
 

### 2024-11-15 manuscript/manuscript.Rmd
Adding word and figure counts

@@
 
-  - Main text (excludes the abstract, section titles, figure and table captions, funding statement, acknowledgments, and references): [X,XXX]
-  - Number of figures: [X]
-  - Number of tables: [X]
+  - Main text (excludes the abstract, section titles, figure and table captions, funding statement, acknowledgments, and references): 8,200
+  - Number of figures: 6
+  - Number of tables: 0
 

### 2024-11-15 manuscript/response_to_reviewers.txt
Fixing line numbers and tidying response

@@
 We have expanded the Discussion to touch on the concerns raised by this reviewer,
-adding section 5.3 to respond to this directly.
+adding section 5.3 (starting at line 624) to respond to this directly.
 Specifically, we have mentioned possible links between the increased sulfonium
-compounds in the eddy and DMS production, a hypothesis about the reason why
-increased cyclonic biomass was not matched by increased export, and extensive
-references to additional manuscripts with relevant data in lines 651-655 and 
-689-718.
+compounds in the eddy and DMS production (lines 605-609), a hypothesis about the reason why
+increased cyclonic biomass was not matched by increased export (lines 642-651), and extensive
+references to additional manuscripts with relevant data throughout the section.
 
@@
 results are not discussed until several sections later) and graphical design (as
-proper visualization of NMDS requires that the x- and y- axes share units).
+proper visualization of NMDS requires that the x- and y- axes share units, making 
+it awkward to place them side-by-side).
 
@@
 We have added the time of day as an explanatory factor to our analysis and reported the
-new values from the PERMANOVA in the text (Lines 220-226, 262-266 and 505-509) and figures, 
-as well as adding a supplemental
-figure to visually represent some of the variance due to time of day. We do see 
-diel differences at the surface and DCM during the eddy transect and these have been
-reported, with the 175 meter differences more likely to be due to chance and very little
+new values from the PERMANOVA in the text (Lines 221-224, 263-267 and 522-525) and figures. 
+We do see diel differences at the surface and DCM during the eddy transect and these have 
+been reported, with the 175 meter differences more likely to be due to chance and very little
 diel difference in the HEE samples at all. We have additionally added a supplemental
@@
 the course of the eddy transect sample run, the normalization removed this
-effect and is not a major factor in the outliers visible in Fig. 2. (Lines 190-192)
+effect and is not a major factor in the outliers visible in Fig. 2. (Lines 193-195)
 
@@
 We have added a note about how we used the methodological blanks to visually assess 
-whether sample peak areas were above the background signal. (Lines 184-186)
+whether sample peak areas were above the background signal. (Lines 187-189)
 
@@
 We have added additional text to this line to clarify the requested points of 
-confusion. (Lines 133-138)
+confusion. (Lines 135-140)
 
@@
 depths. We have also added links to the Github where the R code used to perform
-these matching and normalization steps is available. (Lines 196-198 and 205-208)
+these matching and normalization steps is available. (Lines 195-200 and 206-209)
 
@@
 R-squared and p-values were obtained from PERMANOVAs. We have also added the sample
-size to the Statistics section of the Results. (Lines 220-226)
+size to the Statistics section of the Results. (Lines 224-227)
 
@@
 We have switched to using a 0 in front of Pearson's correlation coefficient throughout
-the manuscript to follow modern conventions.
+the manuscript.
 

### 2024-11-15 manuscript/response_to_reviewers.txt

@@
 adding section 5.3 (starting at line 624) to respond to this directly.
-Specifically, we have mentioned possible links between the increased sulfonium
-compounds in the eddy and DMS production (lines 605-609), a hypothesis about the reason why
-increased cyclonic biomass was not matched by increased export (lines 642-651), and extensive
-references to additional manuscripts with relevant data throughout the section.
+Specifically, we performed some additional analysis of phosphorus-containing lipid
+headgroups to assess metrics of phosphorus stress based on new research published
+since initial submission (Bent et al., 2024: lines 338-351 and Supplemental figure 4), 
+have mentioned possible links between the increased sulfonium compounds in the eddy 
+and DMS production (lines 605-609), a hypothesis about the reason why increased 
+cyclonic biomass was not matched by increased export (lines 642-651), and 
+extensive references to additional manuscripts with relevant data throughout the section.
 
